### 2020-10-20 Ant Matrix.Rmd
initial commit of Ali's `Shared Code' folder contents.

---
title: "Ant Transition Matrix"
author: "Ali Campbell"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
```

```{r}
#import the data
cactus <- read.csv("/Users/alicampbell/Desktop/cholla.csv", header = TRUE)
#str(cactus)
```

##Data Filtering
```{r}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

#hist(log(cactus$volume_t))
```

```{r}
##Year T
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)
cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t))
cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t))
cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t))

##Year T1
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)
cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t1))
cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t1))
cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t1))

#unique(cactus$Antsp_t)
```

```{r}
#Bin all ant species in year t to a value 1 - 5 without taking away the species denotation
cactus$bin1 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t[i] == "LIOM"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "L"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "LIOM "){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "CREM"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "C"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "FOR"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "F"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "VACANT"){cactus$bin1[i] <- 4}
  else if(cactus$Antsp_t[i] == ""){cactus$bin1[i] <- 4}
  else{cactus$bin1[i] <- 5}
}
#Bin all ant species in yeay t1 to a value 1 - 5 without taking away species denotation
cactus$bin2 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t1[i] == "LIOM"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "L"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "LIOM "){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "CREM"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "C"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "FOR"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "F"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "VACANT"){cactus$bin2[i] <- 4}
  else if(cactus$Antsp_t1[i] == ""){cactus$bin2[i] <- 4}
  else{cactus$bin2[i] <- 5}
}
```

##Ant Transition Matrix
#### Hard-Coded Matrix
```{r}
transMat <- as.data.frame(prop.table(with(cactus, table(cactus$bin1, cactus$bin2)), 2))

Transitions <- matrix(c(0.66064710	, 0.04138450	, 0.01504891	, 0.27012792	, 0.01279157 ,	0.14760148	, 0.29889299	, 0.01660517	, 0.50553506	, 0.03136531	, 0.27131783	, 0.09302326	, 0.06201550	, 0.50387597	, 0.06976744	, 0.21302405	, 0.06086401	, 0.01532241	, 0.69035965	, 0.02042988	, 0.27950311	, 0.21118012	, 0.03105590	, 0.43478261	, 0.04347826),ncol = 5, nrow = 5, byrow = FALSE)
#Transitions
#colSums(Transitions)
```

#### Not Hard-Coded matrix
```{r}
# If you have for example a matrix like this
m <- matrix(c(1:25), ncol=5, nrow=5)

for(i in 1:5){
  for(j in 1:5){
    m[i,j] <- transMat[i,3]
  }
}
```

##Plot Survival by Volume probability Data
```{r}
plot(log(cactus$volume_t),cactus$Survival_t1)

survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
  group_by(size_bin) %>%
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

years <- unique(cactus$Year_t)
n_years <- length(years)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")

```

## Vital Rate Plots
#### Survival Rate

Three models were considered here. First A simple survival vs log(volume) model. This had an AIC value of 2380.825.  The second included a second explanatory variable of the ant species (as an independent variable). This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume. This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement. 
```{r}
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

```{r}
#plot the basic surv_model with plot
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
#now with ggplot
ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

This includes an ant interaction term
```{r}
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

#anova(surv_model_ant_interact, test = "Chisq")
#summary(surv_model_ant_interact)
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#### Flowering Model
```{r}
flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)

cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
for(i in 1:8189){
  if(cactus$TotFlowerbuds_t[i] == NA){
    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
  }
  else if(cactus$TotFlowerbuds_t[i] == 0){
    cactus$Flow[i] <- 0
  }
  else if(cactus$TotFlowerbuds_t[i] >0){
    cactus$Flow[i] <- 1
  }
}

cactus$Flow
ggplot(cactus, aes(y=Flow, x=log(volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE)

plot(log(cactus$volume_t),cactus$Flow)
```

#### Growth Model
```{r}
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
```
## Ant Transition Model
```{r}
ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

```{r}
ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
#AIC(ant_trans_model)
#AIC(ant_trans_size_model)
#AIC(ant_trans_size_model2)
##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
summary(ant_trans_size_model2)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
#model.matrix(ant_trans_size_model2)
```

## Multinomial Models

```{r}
#Will be ant vs size
x1 <- factor(as.factor(cactus$bin1), levels = c("5","1","2","3","4"))
m1 <- multinom(as.factor(cactus$bin1) ~ 0+ 1, data = cactus)
m2 <- multinom(x1 ~ 1, data = cactus)
z1<- summary(m1)$coefficients/summary(m1)$standard.errors
z2<- summary(m2)$coefficients/summary(m2)$standard.errors
summary(m1)
summary(m2)
#coef(m1)
#coef(m2)
p1 <- (1 - pnorm(abs(z1), 0, 1)) * 2
p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
ggplot(cactus, aes(x = cactus$bin2, y = cactus$bin1, colour = cactus$bin1)) + geom_line() 
#model.matrix(m1)s
```

```{r}
#Will be ant vs ant
m2 <- multinom(as.factor(bin1) ~ 0+as.factor(bin2), data = cactus)
z<- summary(m2)$coefficients/summary(m2)$standard.errors
#summary(m2)
p <- (1 - pnorm(abs(z), 0, 1)) * 2
ggplot(cactus, aes(x = bin2, y = bin1, colour = bin1)) + geom_line()
#model.matrix(m2)
```

### 2020-10-20 Cholla_Analysis.Rmd
initial commit of Ali's `Shared Code' folder contents.

--- 2020-10-20 Ant Matrix.Rmd
@@
 
+## Ant Transition Model
+```{r}
+ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
+ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
+  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
+```
+
+```{r}
+ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
+ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
+#AIC(ant_trans_model)
+#AIC(ant_trans_size_model)
+#AIC(ant_trans_size_model2)
+##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
+summary(ant_trans_size_model2)
+ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
+  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
+#model.matrix(ant_trans_size_model2)
+```
+
 ##Plot Survival by Volume probability Data
@@
 ```
+```{r}
+survival_plot <- cactus %>% 
+  mutate(size_bin = cut_interval(log(volume_t),10)) %>% 
+  group_by(size_bin, cactus$bin1) %>% 
+  summarise(mean_size = mean(log(volume_t),na.rm=T),
+    surv = mean(Survival_t1,na.rm=T))
 
-## Vital Rate Plots
+plot(log(cactus$volume_t),cactus$Survival_t1)
+points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=1, col=1:length(cactus$bin1))
+legend(12,1,unique(cactus$bin1),col=1:length(cactus$bin1),pch=1)
+```
+
+## Vital Rate Binomial Models
 #### Survival Rate
-
 Three models were considered here.
@@
 ```{r}
-surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
+surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                   family="binomial",
@@
 
-#anova(surv_model_ant_interact, test = "Chisq")
-#summary(surv_model_ant_interact)
+anova(surv_model_ant_interact, test = "Chisq")
+summary(surv_model_ant_interact)
 ```
@@
 ```{r}
-ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
+ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=factor(bin1, label = c("Liom","Crem","For","Vacant","Other")))) +  geom_point(alpha = 0.1) +
+  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) +
+  labs(title = "Survival Rate",colour = "Ant Species") + 
+  xlab("Volume at t") + 
+  ylab("Survival Rate")
 ```
@@
 #### Flowering Model
+The three models considered here model the flowering rates.
+The first considers only the flowering rates vs the log(volume).
+The AIC value for this model is 246.7699.
+The second includes the ant species as a second explanatory variable.
+The AIC value for this is 250.4667.
+The third model includes the ant species as an interacting variable.
+The AIC value for this is 254.5773.
+Therefore there is no statistical evidence that the models including the ant species, either as an independent or dependent variable, are more relevant.
+```{r}
+flo_model <- glm(TotFlowerbuds_t > 0 ~ log(volume_t),
+                  family="binomial",
+                  data=cactus)
+flo_model_ant <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) + as.factor(bin1),
+                  family="binomial",
+                  data=cactus)
+flo_model_ant_interact <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) * as.factor(bin1),
+                  family="binomial",
+                  data=cactus)
+```
+
 ```{r}
@@
                   data=cactus)
+summary(flo_model)
+```
 
+```{r}
 cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
-for(i in 1:8189){
-  if(cactus$TotFlowerbuds_t[i] == NA){
-    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
+for(i in seq_len(nrow(cactus))){
+  if(na.omit(cactus$TotFlowerbuds_t[i]) > 0){cactus$Flow[i] <- 1}
 }
-  else if(cactus$TotFlowerbuds_t[i] == 0){
-    cactus$Flow[i] <- 0
-  }
-  else if(cactus$TotFlowerbuds_t[i] >0){
-    cactus$Flow[i] <- 1
-  }
-}
+
+na.omit(cactus$TotFlowerbuds_t[100]) > 0
 
@@
 #### Growth Model
+Three models were considered for the growth rates vs the log(volume) of the cacti.
+The first is just the log(volue) of the second year vs the log(volume) of the first.
+The AIC value for this is 16126.65.
+The second includes the ant species as a second explanatory variable.
+The AIC value for this is 16039.89.
+The third model includes the ant species as an interacting variable.
+The AIC value for this is 16007.67.
+This means that the third model is statistically improved from the other two.
 ```{r}
+gro_model <- glm(log(volume_t1) ~ log(volume_t),
+                  data=cactus)
+gro_model_ant <- glm(log(volume_t1) ~ 0 + log(volume_t) + as.factor(bin1),
+                  data=cactus)
 gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                   data=cactus)
-
-ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
-```
-## Ant Transition Model
-```{r}
-ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
-ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
 ```
@@
 ```{r}
-ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
-ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
-#AIC(ant_trans_model)
-#AIC(ant_trans_size_model)
-#AIC(ant_trans_size_model2)
-##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
-summary(ant_trans_size_model2)
-ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
-#model.matrix(ant_trans_size_model2)
+#summary(gro_model_ant_interact)
+ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=factor(bin1, label = c("Liom","Crem","For","Vacant","Other")))) +  geom_point(alpha = 0) +
+  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) +
+  labs(title = "Growth Rate",colour = "Ant Species") + 
+  xlab("Volume at t") + 
+  ylab("Volume at t1")
 ```
 
-## Multinomial Models
+```{r}
+par(mfrow=c(2,2),mar=c(4,4,2,1))
+xx=seq(0,8,by=.01)
+hist(cactus$volume_t1[is.na(cactus$volume_t)],main='Size of Recruits',freq=FALSE)
+   # lines(xx,dnorm(xx,mean(cactus$volume_t1[is.na(cactus$volume_t)]),sd(cactus$volume_t1[is.na(cactus$volume_t)])),col='red',lwd=3)
+```
+
+## Multinomial Models of Vital Rates
 
@@
 ```
+
+##Vital Rate Functions
+List of Parameters:
+```{r}
+params=data.frame(
+    ##Survival Functions (by ant bin)
+    surv1.int=NA,
+    surv1.slope=NA,
+    surv2.int=NA,
+    surv2.slope=NA,
+    surv3.int=NA,
+    surv3.slope=NA,
+    surv4.int=NA,
+    surv4.slope=NA,
+    surv5.int=NA,
+    surv5.slope=NA,
+    ##Growth Function
+    flo.int=NA,
+    flo.slope=NA,
+    ##Flower Function
+    gro1.int=NA,
+    gro1.slope=NA,
+    gro2.int=NA,
+    gro2.slope=NA,
+    gro3.int=NA,
+    gro3.slope=NA,
+    gro4.int=NA,
+    gro4.slope=NA,
+    gro5.int=NA,
+    gro5.slope=NA
+)
+params$flo.int = coef(flo_model)[1]
+params$flo.slope = coef(flo_model)[2]
+```
+
+Probability of Surviving:
+```{r}
+s1.x=function(x,params) {
+    u=exp(params$surv1.int+params$surv1.slope*x)
+    return(u/(1+u))
+}
+s2.x=function(x,params) {
+    u=exp(params$surv2.int+params$surv2.slope*x)
+    return(u/(1+u))
+}
+s3.x=function(x,params) {
+    u=exp(params$surv3.int+params$surv3.slope*x)
+    return(u/(1+u))
+}
+s4.x=function(x,params) {
+    u=exp(params$surv4.int+params$surv4.slope*x)
+    return(u/(1+u))
+}
+s5.x=function(x,params) {
+    u=exp(params$surv5.int+params$surv5.slope*x)
+    return(u/(1+u))
+}
+```
+
+Probability of Flowering:
+```{r}
+f.x=function(x,params) {
+    u=exp(params$flo.int+params$flo.slope*x)
+    return(u/(1+u))
+}
+```
+
+Probability of Growth:
+```{r}
+g1.x=function(x,params){
+    u=exp(params$growth1.int+params$growth1.slope*x)
+    return(u/(1+u))
+}
+```

### 2020-10-20 Survival Graphs Binned.Rmd
initial commit of Ali's `Shared Code' folder contents.

---
title: "Survival Graphs Binned"
author: "Ali Campbell"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
```

```{r}
#import the data
cactus <- read.csv("/Users/alicampbell/Desktop/cholla.csv", header = TRUE)
#str(cactus)
```

##Filter Ant Species to standardized
```{r}
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)

##for loop to change them
#Bin all ant species in year t to a value 1 - 5 without taking away the species denotation
cactus$bin1 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t[i] == "LIOM"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "L"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "LIOM "){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "CREM"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "C"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "FOR"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "F"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "VACANT"){cactus$bin1[i] <- 4}
  else if(cactus$Antsp_t[i] == ""){cactus$bin1[i] <- 4}
  else{cactus$bin1[i] <- 5}
}
#Bin all ant species in yeay t1 to a value 1 - 5 without taking away species denotation
cactus$bin2 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t1[i] == "LIOM"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "L"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "LIOM "){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "CREM"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "C"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "FOR"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "F"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "VACANT"){cactus$bin2[i] <- 4}
  else if(cactus$Antsp_t1[i] == ""){cactus$bin2[i] <- 4}
  else{cactus$bin2[i] <- 5}
}
```

##Create Volume
```{r}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

hist(log(cactus$volume_t))
```

##Plot Survival by Volume probability
```{r}
plot(log(cactus$volume_t),cactus$Survival_t1)

survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
  group_by(size_bin) %>%
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

years <- unique(cactus$Year_t)
n_years <- length(years)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")

```

##Fit statistical Binom Model Tier 1
```{r}
#This is the simplest possible model
##The simplest possible model for size dep model
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
AIC(surv_model_ant)
AIC(surv_model)
AIC(surv_model_ant_interact)
#Lower is better for AIC, providing statistical support for the ant model
#AIC suggests that there is a size ant interaction
#AIC difference of 2 is a statistically supported improvement
```

```{r}
summary(surv_model_ant_interact)
b0 <- coef(surv_model_ant_interact)[1]
b1 <- coef(surv_model_ant_interact)[2]
b2 <- coef(surv_model_ant_interact)[3]
b3 <- coef(surv_model_ant_interact)[4]
b4 <- coef(surv_model_ant_interact)[5]
b5 <- coef(surv_model_ant_interact)[6]
b6 <- coef(surv_model_ant_interact)[7]
b7 <- coef(surv_model_ant_interact)[8]
b8 <- coef(surv_model_ant_interact)[9]
b9 <- coef(surv_model_ant_interact)[10]

```

```{r}
vol_range <- seq(from=min(log(na.omit(cactus$volume_t))), to=max(log(na.omit(cactus$volume_t))), by=.01)
mean
1logits <- 0 + b0*vol_range + 
```

```{r}
#plot(surv_model)
plot(log(cactus$volume_t),cactus$Survival_t1)
curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=T)

ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

```{r}
flo_model <- glm(TotFlowerbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
flo_model_ant <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
flo_model_ant_interact <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

AIC(flo_model)
AIC(flo_model_ant)
AIC(flo_model_ant_interact)

##In this case the ant models are not better by AIC
```

```{r}
summary(flo_model)
b0=coef(flo_model)[1]
b1=coef(flo_model)[2]
plot(log(cactus$volume_t1),cactus$TotFlowerbuds_t)
curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=T)
```

```{r}
gro_model <- glm(log(volume_t1) ~ log(volume_t),
                  data=cactus)
gro_model_ant <- glm(log(volume_t1) ~ 0 + log(volume_t) + as.factor(bin1),
                  data=cactus)
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

AIC(gro_model)
AIC(gro_model_ant)
AIC(gro_model_ant_interact)
##The ant interaction model proves the best again
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#Plot survival probability by volume, grouped by ant species (the bins)
```{r}
survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>% 
  group_by(size_bin, cactus$bin1) %>% 
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

plot(log(cactus$volume_t),cactus$Survival_t1)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=1, col=1:length(cactus$bin1))
legend(12,1,unique(cactus$bin1),col=1:length(cactus$bin1),pch=1)
```

##Fit Statistical Binom Model Tier 2
####Independent
```{r}
surv_model_2 <- glmer(Survival_t1 ~ log(volume_t) + (1|Year_t) + (1|Plot),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy_2 <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy_2,invlogit(fixef(surv_model_2)[1] + fixef(surv_model_2)[2]*size_dummy_2), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="blue")
```

####Interaction
```{r}
surv_model_2 <- glmer(Survival_t1 ~ log(volume_t) + ((log(volume_t)|Year_t) + (log(volume_t)|Plot)),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy_2 <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy_2,invlogit(fixef(surv_model_2)[1] + fixef(surv_model_2)[2]*size_dummy_2), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="blue")
```

```{r}
cactus_split <- split(cactus,cactus$bin1)
surv_model_bin1 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`1`)
size_dummy <- seq(min(log(cactus_split$`1`$volume_t),na.rm=T),max(log(cactus_split$`1`$volume_t),na.rm=T),0.01)
surv_model_bin2 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`2`)
plot(log(cactus_split$`2`$volume_t),cactus_split$`2`$Survival_t1, xlab = "log(Volume)", ylab = "Survival")
surv_model_bin3 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`3`)
surv_model_bin4 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`4`)
surv_model_bin5 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`5`)

size_dummy <- seq(min(log(cactus_split$`2`$volume_t),na.rm=T),max(log(cactus_split$`2`$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model_bin1)[1] + coef(surv_model_bin1)[2]*size_dummy), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model_bin2)[1] + coef(surv_model_bin2)[2]*size_dummy), lwd=3, col="blue")
lines(size_dummy,invlogit(coef(surv_model_bin3)[1] + coef(surv_model_bin3)[2]*size_dummy), lwd=3, col="green")
lines(size_dummy,invlogit(coef(surv_model_bin4)[1] + coef(surv_model_bin4)[2]*size_dummy), lwd=3, col="yellow")
lines(size_dummy,invlogit(coef(surv_model_bin5)[1] + coef(surv_model_bin5)[2]*size_dummy), lwd=3, col="orange")
legend(11.5,.4,legend = c("Liom","Crem","For","Vacant","Other"),col = c("red","blue","green","yellow","orange"),lty = 1,cex = .8)
```

```{r}
AIC(surv_model_bin1)
AIC(surv_model_bin2)
```

### 2020-10-21 Ant Matrix_TOM.Rmd
progress on new file

--- 2020-10-20 Ant Matrix.Rmd
@@
 
-```{r}
+```{r read_data}
 #import the data
-cactus <- read.csv("/Users/alicampbell/Desktop/cholla.csv", header = TRUE)
-#str(cactus)
+cactus <- read.csv("cholla_demography_20042019.csv", header = TRUE)
+#str(cactus) ##<- problem: antcount is a factor
+#levels(cactus$Antcount_t);levels(cactus$Antcount_t1)
+## drop the offending rows -- this is a 2019 data entry problem
+cactus <- cactus[-c(which(cactus$Antcount_t=="2-Jan"),which(cactus$Antcount_t1=="2-Jan")),]
+cactus$antcount_t <- as.numeric(as.character(cactus$Antcount_t))
+cactus$antcount_t1 <- as.numeric(as.character(cactus$Antcount_t1))
 ```
@@
 ##Data Filtering
-```{r}
+```{r volume}
 ## function for the volume of a cone
@@
 
-```{r}
-##Year T
-cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
-                              function(variables) {
-                                if (is.character(variables)) {
-                                  return(toupper(variables))
-                                } else {
-                                  return(variables)
-                                }
-                              }),
-                       stringsAsFactors = FALSE)
-cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t))
-cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t))
-cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t))
+```{r year_t_ants}
+## assign ant counts of zero as vacant
+cactus$Antsp_t[cactus$antcount_t==0] <- "vacant"
+# here are the ordered levels of the current variable
+Antsp_t_levels <- levels(cactus$Antsp_t)
+# here is how I would like to collapse these into fewer bins -- most will be "other"
+ant_t_levels <- rep("other",times=length(Antsp_t_levels))
+## crem levels - elements 5,8,9
+ant_t_levels[c(5,8,9)] <- "crem"
+## liom levels - elements 15,18,19,20,21
+ant_t_levels[c(15,18,19,20,21)] <- "liom"
+## forelius levels - 12,13
+ant_t_levels[c(12,13)] <- "for"
+## vacant - 31
+ant_t_levels[31] <- "vacant"
+## create new variable merging levels as above
+cactus$ant_t <- factor(cactus$Antsp_t,levels=Antsp_t_levels,labels=ant_t_levels)
+summary(cactus$ant_t) ## there are too many "others". check these out.
+filter(cactus,ant_t=="other")
+## first, there are no ant data from 2007:
+cactus$Antcount_t[cactus$Year_t==2007]
+#so give these observations NA for ant status:
+cactus$ant_t[cactus$Year_t==2007] <- NA
+# also, if a plant was new in year t+1 it's year t ant status should be NA
+cactus$ant_t[cactus$Newplant==1] <- NA
+# plots 7 and 8 were added in 2011, so their year_t==2010 ant status should be NA
+cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==7] <- NA
+cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==8] <- NA
+# check to see how many others this leaves -- still a lot
+summary(cactus$ant_t)
+## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
+## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
+cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
+## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
+cactus$ant_t[cactus$Antsp_t==""]<-NA
+## inspect the "others" visually
+cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
+## I am satisfied that everything assigned other is correctly assigned
 
-##Year T1
-cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
-                              function(variables) {
-                                if (is.character(variables)) {
-                                  return(toupper(variables))
-                                } else {
-                                  return(variables)
-                                }
-                              }),
-                       stringsAsFactors = FALSE)
-cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t1))
-cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t1))
-cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t1))
-
-#unique(cactus$Antsp_t)
 ```
 
-```{r}
-#Bin all ant species in year t to a value 1 - 5 without taking away the species denotation
-cactus$bin1 <- 0 #blank column 
-for(i in seq_len(nrow(cactus))){
-  if(cactus$Antsp_t[i] == "LIOM"){cactus$bin1[i] <- 1}
-  else if(cactus$Antsp_t[i] == "L"){cactus$bin1[i] <- 1}
-  else if(cactus$Antsp_t[i] == "LIOM "){cactus$bin1[i] <- 1}
-  else if(cactus$Antsp_t[i] == "CREM"){cactus$bin1[i] <- 2}
-  else if(cactus$Antsp_t[i] == "C"){cactus$bin1[i] <- 2}
-  else if(cactus$Antsp_t[i] == "FOR"){cactus$bin1[i] <- 3}
-  else if(cactus$Antsp_t[i] == "F"){cactus$bin1[i] <- 3}
-  else if(cactus$Antsp_t[i] == "VACANT"){cactus$bin1[i] <- 4}
-  else if(cactus$Antsp_t[i] == ""){cactus$bin1[i] <- 4}
-  else{cactus$bin1[i] <- 5}
-}
-#Bin all ant species in yeay t1 to a value 1 - 5 without taking away species denotation
-cactus$bin2 <- 0 #blank column 
-for(i in seq_len(nrow(cactus))){
-  if(cactus$Antsp_t1[i] == "LIOM"){cactus$bin2[i] <- 1}
-  else if(cactus$Antsp_t1[i] == "L"){cactus$bin2[i] <- 1}
-  else if(cactus$Antsp_t1[i] == "LIOM "){cactus$bin2[i] <- 1}
-  else if(cactus$Antsp_t1[i] == "CREM"){cactus$bin2[i] <- 2}
-  else if(cactus$Antsp_t1[i] == "C"){cactus$bin2[i] <- 2}
-  else if(cactus$Antsp_t1[i] == "FOR"){cactus$bin2[i] <- 3}
-  else if(cactus$Antsp_t1[i] == "F"){cactus$bin2[i] <- 3}
-  else if(cactus$Antsp_t1[i] == "VACANT"){cactus$bin2[i] <- 4}
-  else if(cactus$Antsp_t1[i] == ""){cactus$bin2[i] <- 4}
-  else{cactus$bin2[i] <- 5}
-}
+```{r ant_time_series}
+## plot a time series of ant frequency by year
+ant_freq_ts <- cactus %>% 
+  select(Year_t,ant_t) %>% 
+  drop_na() %>% 
+  group_by(Year_t,ant_t) %>% 
+  summarise(n = n()) %>% 
+  mutate(freq = n / sum(n))
+
+plot(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="vacant"],ant_freq_ts$freq[ant_freq_ts$ant_t=="vacant"],type="b",ylim=c(0,1),pch=16,cex=1.5,
+     xlab="Year",ylab="Frequency")
+legend("topleft",legend=c("Vacant","Liom.","Crem.","For.","Other"),pch=16,col=c("black","blue","red","green","pink"))
+points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="liom"],ant_freq_ts$freq[ant_freq_ts$ant_t=="liom"],type="b",pch=16,cex=1.5,col="blue")
+points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="crem"],ant_freq_ts$freq[ant_freq_ts$ant_t=="crem"],type="b",pch=16,cex=1.5,col="red")
+points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="for"],ant_freq_ts$freq[ant_freq_ts$ant_t=="for"],type="b",pch=16,cex=1.5,col="green")
+points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="other"],ant_freq_ts$freq[ant_freq_ts$ant_t=="other"],type="b",pch=16,cex=1.5,col="pink")
+
+```
+
+```{r year_t1_ants}
+## assign ant counts of zero as vacant
+cactus$Antsp_t1[cactus$antcount_t1==0] <- "vacant"
+## repeat for t1 -- these indices are different
+Antsp_t1_levels <- levels(cactus$Antsp_t1)
+# here is how I would like to collapse these into fewer bins -- most will be "other"
+ant_t1_levels <- rep("other",times=length(Antsp_t1_levels))
+## crem levels 
+ant_t1_levels[c(8,9,10)] <- "crem"
+## liom levels 
+ant_t1_levels[c(19,20,21,22)] <- "liom"
+## forelius levels 
+ant_t1_levels[c(13,14,15)] <- "for"
+## vacant 
+ant_t1_levels[32] <- "vacant"
+## create new variable merging levels as above
+cactus$ant_t1 <- factor(cactus$Antsp_t1,levels=Antsp_t1_levels,labels=ant_t1_levels)
+summary(cactus$ant_t1) ## there are too many "others" -- same deal as above
+## first, there are no ant data from 2007 so give these observations NA for ant status:
+cactus$ant_t1[cactus$Year_t1==2007] <- NA
+# check to see how many others this leaves -- still a lot
+summary(cactus$ant_t1)
+## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
+## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
+cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
+## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
+cactus$ant_t[cactus$Antsp_t==""]<-NA
+## inspect the "others" visually
+cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
+## I am satisfied that everything assigned other is correctly assigned
+
 ```

### 2020-10-21 Ant Matrix_TOM.Rmd
got through multinomial fits, now passing back to Ali

@@
 library(effects)
+library(bbmle)
 ```
@@
 cactus$ant_t1[cactus$Year_t1==2007] <- NA
+## if we did not get size in t1 assume we did not get ant
+cactus$ant_t1[is.na(cactus$Height_t1)] <- NA
 # check to see how many others this leaves -- still a lot
 summary(cactus$ant_t1)
-## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
-## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
-cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
-## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
-cactus$ant_t[cactus$Antsp_t==""]<-NA
+## there are some Antsp_t1=="" that show up as "other"
+cactus %>% filter(Antsp_t1=="" & ant_t1=="other")
+## not sure why this happens but apparently there are some plants for which we have ant counts but no ant sp. Could be a mix in comments. For now make these NA.
+cactus$ant_t1[cactus$Antsp_t1==""] <- NA
 ## inspect the "others" visually
-cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
-## I am satisfied that everything assigned other is correctly assigned
+cactus %>% filter(ant_t1=="other") %>% select(Antsp_t1,ant_t1)
+## I am satisfied that everything assigned other in t1 is correctly assigned
+
+```
+
+## Vital rates
+
+```{r survival}
+surv_model <- glm(Survival_t1 ~ log(volume_t),family="binomial",data=cactus)
+surv_model_ant <- glm(Survival_t1 ~ log(volume_t) + ant_t,family="binomial",data=cactus)
+surv_model_ant_interact <- glm(Survival_t1 ~ log(volume_t) * ant_t,family="binomial",data=cactus)
+AICtab(surv_model,surv_model_ant,surv_model_ant_interact)
+
+ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t)) +  geom_point(alpha = 0.1) +
+  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE)
+
+```
+
+```{r growth}
+grow_model <- lm(log(volume_t1) ~ log(volume_t),data=cactus)
+grow_model_ant <- lm(log(volume_t1) ~ log(volume_t) + ant_t,data=cactus)
+grow_model_ant_interact <- lm(log(volume_t1) ~ log(volume_t) * ant_t,data=cactus)
+AICtab(grow_model,grow_model_ant,grow_model_ant_interact)
+
+plot(log(cactus$volume_t),log(cactus$volume_t1),type="n")
+points(log(cactus$volume_t[cactus$ant_t=="vacant"]),log(cactus$volume_t1[cactus$ant_t=="vacant"]),col=alpha("black",0.1))
+abline(coef(grow_model_ant_interact)[1]+coef(grow_model_ant_interact)[6],coef(grow_model_ant_interact)[2]+coef(grow_model_ant_interact)[10],col="black")
+points(log(cactus$volume_t[cactus$ant_t=="liom"]),log(cactus$volume_t1[cactus$ant_t=="liom"]),col=alpha("blue",0.1))
+abline(coef(grow_model_ant_interact)[1]+coef(grow_model_ant_interact)[5],coef(grow_model_ant_interact)[2]+coef(grow_model_ant_interact)[9],col="blue")
+points(log(cactus$volume_t[cactus$ant_t=="crem"]),log(cactus$volume_t1[cactus$ant_t=="crem"]),col=alpha("red",0.1))
+abline(coef(grow_model_ant_interact)[1]+coef(grow_model_ant_interact)[3],coef(grow_model_ant_interact)[2]+coef(grow_model_ant_interact)[7],col="red")
+points(log(cactus$volume_t[cactus$ant_t=="for"]),log(cactus$volume_t1[cactus$ant_t=="for"]),col=alpha("green",0.1))
+abline(coef(grow_model_ant_interact)[1]+coef(grow_model_ant_interact)[4],coef(grow_model_ant_interact)[2]+coef(grow_model_ant_interact)[8],col="green")
+points(log(cactus$volume_t[cactus$ant_t=="other"]),log(cactus$volume_t1[cactus$ant_t=="other"]),col=alpha("pink",0.1))
+abline(coef(grow_model_ant_interact)[1],coef(grow_model_ant_interact)[2],col="pink")
+
+```
+
+```{r floral_abortion}
+## see what we have in terms of counts of total, good, and aborted flowerbuds
+cactus %>% 
+  group_by(Year_t) %>% 
+  select(Goodbuds_t,TotFlowerbuds_t,ABFlowerbuds_t) %>% 
+  summarise(mean1 = mean(Goodbuds_t,na.rm=T),
+            mean2 = mean(TotFlowerbuds_t,na.rm=T),
+            mean3 = mean(ABFlowerbuds_t,na.rm=T))
+## only goodbuds in 2009-2012 so no abortion data
+## before that we have all three counts
+## 2013 and later, just counts of goodbuds and aborted
+
+abort_model <- glm(cbind(Goodbuds_t,ABFlowerbuds_t) ~ 1,family="binomial",data=cactus)
+abort_model_ant <- glm(cbind(Goodbuds_t,ABFlowerbuds_t) ~ ant_t,family="binomial",data=cactus)
+AICtab(abort_model,abort_model_ant)
+
+## view the raw proportions
+cactus %>% 
+  mutate(viab_rate = Goodbuds_t/(Goodbuds_t+ABFlowerbuds_t)) %>% 
+  group_by(ant_t) %>% 
+  summarise(mean_viab = mean(viab_rate,na.rm=T),
+            n_viab = n())
 
@@
 ##Ant Transition Matrix
-#### Hard-Coded Matrix
-```{r}
-transMat <- as.data.frame(prop.table(with(cactus, table(cactus$bin1, cactus$bin2)), 2))
 
-Transitions <- matrix(c(0.66064710	, 0.04138450	, 0.01504891	, 0.27012792	, 0.01279157 ,	0.14760148	, 0.29889299	, 0.01660517	, 0.50553506	, 0.03136531	, 0.27131783	, 0.09302326	, 0.06201550	, 0.50387597	, 0.06976744	, 0.21302405	, 0.06086401	, 0.01532241	, 0.69035965	, 0.02042988	, 0.27950311	, 0.21118012	, 0.03105590	, 0.43478261	, 0.04347826),ncol = 5, nrow = 5, byrow = FALSE)
-#Transitions
-#colSums(Transitions)
+```{r ant_matrix}
+table(cactus$ant_t, cactus$ant_t1)
+plot(table(cactus$ant_t, cactus$ant_t1))
+
+transition_mat <- prop.table(table(cactus$ant_t, cactus$ant_t1),margin=2)
+## check that columns sum to 1
+colSums(transition_mat)
 ```
 
-#### Not Hard-Coded matrix
-```{r}
-# If you have for example a matrix like this
-m <- matrix(c(1:25), ncol=5, nrow=5)
+```{r multinom}
+## for multinomial model, first relevel the ant factor to have a reference level, which will be vacant
+cactus$ant_t1_relevel <- relevel(cactus$ant_t1,ref = "vacant")
+## create new variable for plant reproductive state transition (we know nectar goes up when plants start flowering)
+cactus$repro_trans <- c()
+cactus$repro_trans[cactus$Goodbuds_t==0 & cactus$Goodbuds_t1==0]<-"VV"
+cactus$repro_trans[cactus$Goodbuds_t==0 & cactus$Goodbuds_t1>0]<-"VF"
+cactus$repro_trans[cactus$Goodbuds_t>0 & cactus$Goodbuds_t1==0]<-"FV"
+cactus$repro_trans[cactus$Goodbuds_t>0 & cactus$Goodbuds_t1>0]<-"FF"
 
-for(i in 1:5){
-  for(j in 1:5){
-    m[i,j] <- transMat[i,3]
-  }
-}
+## here is a base model, no predictor variables
+ant_transition_models<-list()
+## null model
+ant_transition_models[[1]] <- multinom(ant_t1_relevel ~ 1, data = cactus)
+## previous state dependence
+ant_transition_models[[2]] <- multinom(ant_t1_relevel ~ 0+ant_t, data = cactus)
+## previous state and size
+ant_transition_models[[3]] <- multinom(ant_t1_relevel ~ 0+ant_t + log(volume_t), data = cactus)
+ant_transition_models[[4]] <- multinom(ant_t1_relevel ~ 0+ant_t * log(volume_t), data = cactus)
+## previous state and reproductive transition
+ant_transition_models[[5]] <- multinom(ant_t1_relevel ~ 0+ant_t + as.factor(repro_trans), data = cactus)
+ant_transition_models[[6]] <- multinom(ant_t1_relevel ~ 0+ant_t * as.factor(repro_trans), data = cactus)
+## all three
+ant_transition_models[[7]] <- multinom(ant_t1_relevel ~ 0+ant_t + log(volume_t) + as.factor(repro_trans), data = cactus)
+ant_transition_models[[8]] <- multinom(ant_t1_relevel ~ 0+ant_t * as.factor(repro_trans) + log(volume_t), data = cactus) 
+ant_transition_models[[9]] <- multinom(ant_t1_relevel ~ 0+ant_t * log(volume_t) + as.factor(repro_trans), data = cactus) 
+ant_transition_models[[10]] <- multinom(ant_t1_relevel ~ 0+ant_t + log(volume_t) * as.factor(repro_trans), data = cactus) 
+##3-way intx does not converge
+
+AICtab(ant_transition_models)
 ```
-
-##Plot Survival by Volume probability Data
-```{r}
-plot(log(cactus$volume_t),cactus$Survival_t1)
-
-survival_plot <- cactus %>% 
-  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
-  group_by(size_bin) %>%
-  summarise(mean_size = mean(log(volume_t),na.rm=T),
-    surv = mean(Survival_t1,na.rm=T))
-
-years <- unique(cactus$Year_t)
-n_years <- length(years)
-points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")
+```{r visualize multinomial}
+summary(ant_transition_models[[5]])
 
 ```
-
-## Vital Rate Plots
-#### Survival Rate
-
-Three models were considered here.
-First A simple survival vs log(volume) model.
-This had an AIC value of 2380.825.
-The second included a second explanatory variable of the ant species (as an independent variable).
-This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume.
-This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement.
-```{r}
-surv_model <- glm(Survival_t1 ~ log(volume_t),
-                  family="binomial",
-                  data=cactus)
-surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
-                  family="binomial",
-                  data=cactus)
-surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
-                  family="binomial",
-                  data=cactus)
-```
-
-```{r}
-#plot the basic surv_model with plot
-plot(log(cactus$volume_t),cactus$Survival_t1)
-size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
-lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
-#now with ggplot
-ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
-```
-
-This includes an ant interaction term
-```{r}
-surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
-                  family="binomial",
-                  data=cactus)
-
-#anova(surv_model_ant_interact, test = "Chisq")
-#summary(surv_model_ant_interact)
-```
-
-```{r}
-ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
-```
-
-#### Flowering Model
-```{r}
-flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
-                  family="binomial",
-                  data=cactus)
-
-cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
-for(i in 1:8189){
-  if(cactus$TotFlowerbuds_t[i] == NA){
-    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
-  }
-  else if(cactus$TotFlowerbuds_t[i] == 0){
-    cactus$Flow[i] <- 0
-  }
-  else if(cactus$TotFlowerbuds_t[i] >0){
-    cactus$Flow[i] <- 1
-  }
-}
-
-cactus$Flow
-ggplot(cactus, aes(y=Flow, x=log(volume_t))) +  geom_point() +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE)
-
-plot(log(cactus$volume_t),cactus$Flow)
-```
-
-#### Growth Model
-```{r}
-gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
-                  data=cactus)
-
-ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
-```
-## Ant Transition Model
-```{r}
-ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
-ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
-```
-
-```{r}
-ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
-ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
-#AIC(ant_trans_model)
-#AIC(ant_trans_size_model)
-#AIC(ant_trans_size_model2)
-##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
-summary(ant_trans_size_model2)
-ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
-  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
-#model.matrix(ant_trans_size_model2)
-```
-
-## Multinomial Models
-
-```{r}
-#Will be ant vs size
-x1 <- factor(as.factor(cactus$bin1), levels = c("5","1","2","3","4"))
-m1 <- multinom(as.factor(cactus$bin1) ~ 0+ 1, data = cactus)
-m2 <- multinom(x1 ~ 1, data = cactus)
-z1<- summary(m1)$coefficients/summary(m1)$standard.errors
-z2<- summary(m2)$coefficients/summary(m2)$standard.errors
-summary(m1)
-summary(m2)
-#coef(m1)
-#coef(m2)
-p1 <- (1 - pnorm(abs(z1), 0, 1)) * 2
-p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
-ggplot(cactus, aes(x = cactus$bin2, y = cactus$bin1, colour = cactus$bin1)) + geom_line() 
-#model.matrix(m1)s
-```
-
-```{r}
-#Will be ant vs ant
-m2 <- multinom(as.factor(bin1) ~ 0+as.factor(bin2), data = cactus)
-z<- summary(m2)$coefficients/summary(m2)$standard.errors
-#summary(m2)
-p <- (1 - pnorm(abs(z), 0, 1)) * 2
-ggplot(cactus, aes(x = bin2, y = bin1, colour = bin1)) + geom_line()
-#model.matrix(m2)
-```

### 2020-10-23 Ant Matrix_TOM.Rmd
ADded a size distribution figure.

@@
 ```
+
+```{r ant_size_distributions}
+plot(density(log(cactus$volume_t[cactus$ant_t=="vacant"]),na.rm=T),type="l",lwd=3,ylim=c(0,0.3),main=" ",
+     xlab="Plant size (log volume)")
+lines(density(log(cactus$volume_t[cactus$ant_t=="liom"]),na.rm=T),type="l",lwd=3,col="blue")
+lines(density(log(cactus$volume_t[cactus$ant_t=="crem"]),na.rm=T),type="l",lwd=3,col="red")
+lines(density(log(cactus$volume_t[cactus$ant_t=="for"]),na.rm=T),type="l",lwd=3,col="green")
+lines(density(log(cactus$volume_t[cactus$ant_t=="other"]),na.rm=T),type="l",lwd=3,col="pink")
+legend("topleft",legend=c("Vacant","Liom.","Crem.","For.","Other"),pch=16,col=c("black","blue","red","green","pink"),bty="n",lwd=3)
+
+```
+
+```{r}
+
+par(mfrow=c(2,3))
+plot(density(log(cactus$volume_t[cactus$ant_t=="vacant"]),na.rm=T))
+points(log(cactus$volume_t[cactus$ant_t=="vacant"]),
+       rep(0,times=length(log(cactus$volume_t[cactus$ant_t=="vacant"]))),pch="|")
+plot(density(log(cactus$volume_t[cactus$ant_t=="liom"]),na.rm=T))
+points(log(cactus$volume_t[cactus$ant_t=="liom"]),
+       rep(0,times=length(log(cactus$volume_t[cactus$ant_t=="liom"]))),pch="|")
+plot(density(log(cactus$volume_t[cactus$ant_t=="crem"]),na.rm=T))
+points(log(cactus$volume_t[cactus$ant_t=="crem"]),
+       rep(0,times=length(log(cactus$volume_t[cactus$ant_t=="crem"]))),pch="|")
+plot(density(log(cactus$volume_t[cactus$ant_t=="for"]),na.rm=T))
+points(log(cactus$volume_t[cactus$ant_t=="for"]),
+       rep(0,times=length(log(cactus$volume_t[cactus$ant_t=="for"]))),pch="|")
+plot(density(log(cactus$volume_t[cactus$ant_t=="other"]),na.rm=T))
+points(log(cactus$volume_t[cactus$ant_t=="other"]),
+       rep(0,times=length(log(cactus$volume_t[cactus$ant_t=="other"]))),pch="|")
+```

### 2020-10-23 Ant Matrix_TOM.Rmd
ADded a size distribution figure.

@@
 ```{r ant_size_distributions}
-plot(density(log(cactus$volume_t[cactus$ant_t=="vacant"]),na.rm=T),type="l",lwd=3,ylim=c(0,0.3),main=" ",
-     xlab="Plant size (log volume)")
-lines(density(log(cactus$volume_t[cactus$ant_t=="liom"]),na.rm=T),type="l",lwd=3,col="blue")
-lines(density(log(cactus$volume_t[cactus$ant_t=="crem"]),na.rm=T),type="l",lwd=3,col="red")
-lines(density(log(cactus$volume_t[cactus$ant_t=="for"]),na.rm=T),type="l",lwd=3,col="green")
-lines(density(log(cactus$volume_t[cactus$ant_t=="other"]),na.rm=T),type="l",lwd=3,col="pink")
-legend("topleft",legend=c("Vacant","Liom.","Crem.","For.","Other"),pch=16,col=c("black","blue","red","green","pink"),bty="n",lwd=3)
-
-```
-
-```{r}
+plot(density(log(cactus$volume_t[cactus$ant_t=="vacant"]),na.rm=T),type="l",lwd=2,ylim=c(0,0.3))
+lines(density(log(cactus$volume_t[cactus$ant_t=="liom"]),na.rm=T),type="l",lwd=2,col="blue")
+lines(density(log(cactus$volume_t[cactus$ant_t=="crem"]),na.rm=T),type="l",lwd=2,col="red")
+lines(density(log(cactus$volume_t[cactus$ant_t=="for"]),na.rm=T),type="l",lwd=2,col="green")
+lines(density(log(cactus$volume_t[cactus$ant_t=="other"]),na.rm=T),type="l",lwd=2,col="pink")
 

### 2021-02-08 Stan/Stan Base Models.Rmd
Stan work is here

---
title: "Stan Practice"
author: "Ali Campbell"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
```

## Basic Models (No Ant Effects)
```{r Growth Model}
## Now rename variables and index
x <- log(data$volume_t)
z <- as.numeric(data$ant_t)
y <- log(data$volume_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
#summary(grow_model)
#summary(mod1)
## Now we can extract some important statistics
lm_alpha <- summary(grow_model)$coeff[1]  # the intercept
lm_beta <- summary(grow_model)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model)  # the residual error
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y, z = z, K = K)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
 int < lower = 1 > N; // Sample size
 vector[N] x; // Predictor
 vector[N] y; // Outcome
}

parameters {
 real alpha; // Intercept
 real beta; // Slope (regression coefficients)
 real < lower = 0 > sigma; // Error SD
}

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~ normal(alpha + x * beta , sigma); // The Model
}

generated quantities {
} // The posterior predictive distribution",

"grow_lm1.stan")
## Check that you wrote a file
#stanc("grow_lm1.stan")
## Save the file path
stanny <- "grow_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
#class(fit)
posterior <- as.data.frame(fit)
print(colnames(posterior))

par(mfrow = c(1,3))
plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r Total Flowers}
x <- log(data$volume_t)
y <- data$TotFlowerbuds_t
N <- length(data$volume_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
summary(flower_model)
#summary(mod1)
## Now we can extract some important statistics
lm_lambda = mean(data$TotFlowerbuds_t)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple total flowers regression

data {
 int N;
 real x[N];
 int y[N];
}
parameters {
 real<lower = 0> lambda;
}
model {
// Priors
 lambda ~ cauchy(0, 1);
 //lambda ~ gamma(0, 0);
 // Model
 y ~ poisson(lambda);
}
",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$lambda), main = "Lambda")
abline(v = lm_lambda, col = 4, lty = 2)
```

```{r Survival}
x <- log(data$volume_t)
y <- data$Survival_t1
N <- length(data$volume_t)
#import things from the r regression models for comparison
lm_alpha <- summary(surv_model)$coeff[1]  # the intercept
lm_beta <- summary(surv_model)$coeff[2]  # the slope
summary(surv_model)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple survival regression

data {
  int<lower=0> N;
  vector[N] x;
  int<lower=0,upper=1> y[N];
}
parameters {
  real alpha;
  real beta;
}
model {
 // Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(alpha + beta * x);
}
",

"surv_lm1.stan")
## Check that you wrote a file
stanc("surv_lm1.stan")
## Save the file path
stanny <- "surv_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

par(mfrow = c(1,2))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
```

#### Include ants
######Growth
```{r Growth Model}
## Now rename variables and index
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$volume_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
lm_alpha <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_beta <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

write("// Stan model for ant growth regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 real < lower = 0 > y[N]; // outcome vector
}

parameters {
 real alpha; // Intercept
 vector[K] beta; // coefficients for predictors
 real < lower = 0 > sigma; // Error SD
}

transformed parameters{
real mu[N];
for(i in 1:N){
mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
             }
}

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~  normal(mu, sigma);   // The Model
}

generated quantities {
} // The posterior predictive distribution",

"grow_ant_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000 , chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

par(mfrow = c(1,2))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

#plot(density(posterior$beta[i]), main = "Beta")
#abline(v = lm_beta, col = 4, lty = 2)

plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r grow attempt 1}
vol = data$volume_t
y = data$volume_t1
crem = data$crem
liom = data$liom
vacant = data$vacant
other = data$other
N = nrow(data)
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants

write("
      data{
      int <lower = 1> N; // Number of observations
      vector[N] vol; // volume predictor
      int <lower = 0, upper = 1> crem[N]; // crem predictor
      int <lower = 0, upper = 1> liom[N]; // liom predictor
      int <lower = 0, upper = 1> other[N]; // other predictor
      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
      vector[N] y; // outcome
      }
      parameters{
      real alpha; // intercept 
      real b_crem; // crem slope
      real b_liom; // liom slope
      real b_other; // other slope
      real b_vacant; // vacant slope
      real <lower = 0> sigma; // standard deviation
      }
      model{
      // Priors
      alpha ~ normal(0,100);
      beta ~ normal(0,100);
      // Likelihood
      y ~ normal(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol ,sigma);
      }
      ",
      "tester1.stan")

stanc("tester1.stan")

fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))
par(mfrow = c(2,3))
plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
plot(density(posterior$b_other), main = "Beta Other")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_vacant), main = "Beta Vacant")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_crem), main = "Beta Crem")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_liom), main = "Beta Liom")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r grow attempt 2&3}
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
## Write the model in stan
lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
write("// Stan model for simple linear regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 int <lower=0,upper=1> y[N]; // outcome vector
}

parameters {
  real alpha;
  vector[K] beta; // coefficients for predictors
}

transformed parameters{
 real mu[N];
 for(i in 1:N){
 mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
 }
}

model {
//Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
```

######Survival

```{r survival}
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
## Write the model in stan
lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
write("// Stan model for simple ant survival regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 int <lower=0,upper=1> y[N]; // outcome vector
}

parameters {
  real alpha;
  vector[K] beta; // coefficients for predictors
}

transformed parameters{
 real mu[N];
 for(i in 1:N){
 mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
 }
}

model {
//Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
```

```{r surv attempt 1}
vol = data$volume_t
y = as.integer(data$Survival_t1)
crem = data$crem
liom = data$liom
vacant = data$vacant
other = data$other
N = nrow(data)
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants

write("
      data{
      int <lower = 1> N; // Number of observations
      vector[N] vol; // volume predictor
      int <lower = 0, upper = 1> crem[N]; // crem predictor
      int <lower = 0, upper = 1> liom[N]; // liom predictor
      int <lower = 0, upper = 1> other[N]; // other predictor
      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
      int <lower = 0, upper = 1> y[N]; // outcome
      }
      parameters{
      real alpha; // intercept 
      real b_crem; // crem slope
      real b_liom; // liom slope
      real b_other; // other slope
      real b_vacant; // vacant slope
      real <lower = 0> sigma; // standard deviation
      }
      model{
      // Priors
      alpha ~ normal(0,100);
      // Likelihood
      y ~ bernoulli_logit(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol);
      }
      ",
      "tester1.stan")

stanc("tester1.stan")

fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit
```

### 2021-02-08 Simulate Data copy/Data_Simulation_ALI.Rmd
Stan work is here

---
title: "Simulate the Models"
author: "Ali Campbell"
date: "12/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
```

###Growth model

```{r}
#Model:
grow_model_ant_interact
#Simulated Data:
coefs <- mvrnorm(n = 10000, mu = coefficients(grow_model_ant_interact), Sigma = vcov(grow_model_ant_interact))
plot(grow_model_ant_interact)

##Check this against original model
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(grow_model_ant_interact)
#These are pretty similar, so I will now look at standard errors
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(grow_model_ant_interact)))
#These are also pretty good. So now we simulate the data

##simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(grow_model_ant_interact)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)
##Now we compare the simualted and original variance and mean
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
c(mean(cactus$volume_t1,na.rm=TRUE), var(cactus$volume_t1,na.rm = TRUE))
```

The variance is not that far off
The mean is significantly different. This to me indicates that likely the model is not a good one

```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
grow_model_ant_interact <- glm(formula = log(volume_t1) ~ log(volume_t) * ant_t, data = cactus, family = "gaussian")
anova(grow_model_ant_interact, test = "Chisq")
summary(grow_model_ant_interact)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
# note that terms is for a linear model and response is for a binomial model
y_pred <- predict(grow_model_ant_interact, newdata = newdat, type = "terms")
y_CI <- predict(grow_model_ant_interact, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(grow_model_ant_interact)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = volume_t1, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) 
  #geom_smooth(data=cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE) 
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

###Survival Model
```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
surv_model_ant_interact
anova(surv_model_ant_interact, test = "Chisq")
summary(surv_model_ant_interact)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(surv_model_ant_interact, newdata = newdat, type = "response")
y_CI <- predict(surv_model_ant_interact, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(surv_model_ant_interact)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) 
  #geom_smooth(data=cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE) 
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

###Viable Model (OKAY. I DO NOT KNOW HOW TO SIMULATE THIS ONE. AT ALL. PLEASE HELP)
```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
abort_model_ant
anova(abort_model_ant, test = "Chisq")
summary(abort_model_ant)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(abort_model_ant, newdata = newdat, type = "response")
y_CI <- predict(abort_model_ant, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(abort_model_ant)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = ant_t, y = cbind(Goodbuds_t,ABFlowerbuds_t), pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) +
  geom_line(data = newpred, aes(x = ant_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = ant_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) 
  #geom_smooth(data=cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE) 
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

```{r}
#Model:
abort_model_ant
#Simulated Data:
coefs <- mvrnorm(n = 10000, mu = coefficients(abort_model_ant), Sigma = vcov(abort_model_ant))

##Check this against original model
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(abort_model_ant)
#These are pretty similar, so I will now look at standard errors
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(abort_model_ant)))
#These are also pretty good. So now we simulate the data

##simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(abort_model_ant)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)

#Now we compare the simualted and original variance and mean
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
x <- cbind(cactus$Goodbuds_t, cactus$ABFlowerbuds_t)
c(mean(x,na.rm=TRUE), var(x,na.rm = TRUE))
```
Sooooo, This is not quite right. Clearly.

###Reproductive State
```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
repro_model_ant
anova(repro_model_ant, test = "Chisq")
summary(repro_model_ant)
plot(repro_model_ant)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(repro_model_ant, newdata = newdat, type = "response")
y_CI <- predict(repro_model_ant, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(repro_model_ant)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = repro_state_t, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) 
  #geom_smooth(data=cactus, aes(y=repro_state_t, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE)
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

###Total Flowers

```{r}
#Model:
flower_model_ant_interact
#Simulated Data:
coefs <- mvrnorm(n = 10000, mu = coefficients(flower_model_ant_interact), Sigma = vcov(flower_model_ant_interact))

##Check this against original model
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(flower_model_ant_interact)
#These are pretty similar, so I will now look at standard errors
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(flower_model_ant_interact)))
#These are also pretty good. So now we simulate the data

##simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(flower_model_ant_interact)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)

##Now we compare the simualted and original variance and mean
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
c(mean(cactus$TotFlowerbuds_t,na.rm=TRUE), var(cactus$TotFlowerbuds_t,na.rm = TRUE))
```

Ok, So not perfect, but really not that bad!!!
## Visualize it!
```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
flower_model_ant_interact
anova(flower_model_ant_interact, test = "Chisq")
summary(flower_model_ant_interact)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(flower_model_ant_interact, newdata = newdat, type = "response")
y_CI <- predict(flower_model_ant_interact, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(flower_model_ant_interact)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = TotFlowerbuds_t, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) 
  #geom_smooth(data=cactus, aes(y=repro_state_t, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE)
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

###

### 2021-02-08 Simulate Data copy/Tester.Rmd
Stan work is here

---
title: "Tester"
author: "Ali Campbell"
date: "12/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create the data:
```{r}
library(MASS) # For multivariate normal distribution, handy later on
n <- 50
set.seed(18050518)
dat <- data.frame(xc = rnorm(n), xb = rbinom(n, 1, .5))

dat <- within(dat, y <- rpois(n, exp(.5 * xc + xb)))
summary(dat)
```

Model:
```{r}
mod <- glm(y ~ xc + xb, poisson, dat)
```

Simulated Data:
```{r}
coefs <- mvrnorm(n = 10000, mu = coefficients(mod), Sigma = vcov(mod))
```

Check this against original model
```{r}
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(mod)
```
These are pretty similar, so I will now look at standard errors

```{r}
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(mod)))
```
These are also pretty good. So now we simulate the data

```{r}
#simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(mod)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)
```

Now we compare the simualted and original variance and mean
```{r}
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
c(mean(dat$y), var(dat$y))
```
These are pretty close, particularly teh means

###Now try this same process with the growth model that is just a glm
Model:
```{r}
grow_model
```

Simulated Data:
```{r}
coefs <- mvrnorm(n = 10000, mu = coefficients(grow_model), Sigma = vcov(grow_model))
```

Check this against original model
```{r}
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(grow_model)
```
These are pretty similar, so I will now look at standard errors

```{r}
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(grow_model)))
```
These are also pretty good. So now we simulate the data

```{r}
#simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(grow_model)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)
```

Now we compare the simualted and original variance and mean
```{r}
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
c(mean(cactus$volume_t1,na.rm=TRUE), var(cactus$volume_t1,na.rm = TRUE))
```

This mean is not close, but the variances are not THAT bad. We know though that this model is not good.

###Try again with the right growth model
Model:
```{r}
grow_model_ant_interact
```

Simulated Data:
```{r}
coefs <- mvrnorm(n = 10000, mu = coefficients(grow_model_ant_interact), Sigma = vcov(grow_model_ant_interact))
```

Check this against original model
```{r}
#Simulated Coefficients
colMeans(coefs)
#Original Coefficients
coefficients(grow_model_ant_interact)
```
These are pretty similar, so I will now look at standard errors

```{r}
#Simulated Standard Error
apply(coefs, 2, sd)
#Original Standard Error
sqrt(diag(vcov(grow_model_ant_interact)))
```
These are also pretty good. So now we simulate the data

```{r}
#simulate the data
sim.dat <- matrix(nrow = n, ncol = nrow(coefs))
mod.mat <- model.matrix(grow_model_ant_interact)
for (i in 1:nrow(coefs)) {
  sim.dat[, i] <- rpois(n, exp(mod.mat %*% coefs[i, ]))
}
rm(i, mod.mat)
```

Now we compare the simualted and original variance and mean
```{r}
#Simulated mean and var
c(mean(colMeans(sim.dat)), mean(apply(sim.dat, 2, var)))
#Original mean and var
c(mean(cactus$volume_t1,na.rm=TRUE), var(cactus$volume_t1,na.rm = TRUE))
```

The variance is not that far off
The mean is significantly different. This to me indicates that likely the model is not a good one

###Now Josh's Method

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
require(lattice)
elvi_fitness_data <- read.csv("/Users/alicampbell/Desktop/endo_herb_georef.csv", header = TRUE,stringsAsFactors=T)
```

```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
infl_mod <- glm(mean_infl_length ~ year*Endo_status_liberal, data = (subset(elvi_fitness_data)), family = "gaussian")
anova(infl_mod, test = "Chisq")
summary(infl_mod)

# plot(infl_mod)

# This is our simulated predictor data
newdat1<- data.frame(year = seq(1900,2016,1), Endo_status_liberal = 1)
newdat0 <- data.frame(year= seq(1900,2016,1), Endo_status_liberal = 0)
newdat <- rbind(newdat0, newdat1)

# use the predict function for to get the model output
y_pred <- predict(infl_mod, newdata = newdat, type = "response")
y_CI <- predict(infl_mod, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(infl_mod)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ELVI_outcomeyear <- ggplot() +
  geom_point(data = elvi_fitness_data, aes(x = year, y =mean_infl_length, pch = as.factor(Endo_status_liberal)), color = "#4c90dd", lwd = 2)+
  geom_line(data = newpred, aes(x = year, y = pred, group = Endo_status_liberal, linetype = as.factor(Endo_status_liberal)), color = "#4c90dd") +
  geom_ribbon(data = newpred, aes(x = year, ymin = lwr, ymax = upr, group = Endo_status_liberal, fill = as.factor(Endo_status_liberal)), alpha = .2) +
  theme_classic() + labs(y = "Mean Inflorescence Length", x = "Year", linetype = "Endophyte_status", shape = "Endophyte_status")+
  scale_linetype_manual(breaks = c("1", "0"),
                        values = c("solid", "dashed"))+
  scale_shape_manual(breaks = c("1", "0"),
                     values = c(20,1))+
  scale_fill_manual(breaks = c("1", "0"),
                    values = c(  "#4c90dd", "#636363")) +
  xlim(1910,2016) + guides(fill = FALSE)

ELVI_outcomeyear
#ggsave(ELVI_outcome, filename = "~/Documents/ELVIinflyear.tiff", width = 4, height = 3)
```

```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
surv_model_ant_interact
anova(surv_model_ant_interact, test = "Chisq")
summary(surv_model_ant_interact)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(surv_model_ant_interact, newdata = newdat, type = "response")
y_CI <- predict(surv_model_ant_interact, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(surv_model_ant_interact)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- invlogit(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=invlogit(pred0-sc*y_CI$se.fit),
                     upr=invlogit(pred0+sc*y_CI$se.fit),
                     lwr2=invlogit(pred0-sc2*y_CI$se.fit),
                     upr2=invlogit(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) + 
  geom_smooth(data=cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE)
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

```{r}
# This is were I run the model
# This is a gaussian model, but you can change the family to be poisson or whatever
surv_model_ant_interact
anova(surv_model_ant_interact, test = "Chisq")
summary(surv_model_ant_interact)
# plot(infl_mod)

# This is our simulated predictor data
min(cactus$volume_t, na.rm = T)
newdat1<- data.frame(volume_t = seq(log(0.1),log(3129321),0.002076463), ant_t = "crem")
newdat2 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "liom")
newdat3 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "vacant")
newdat4 <- data.frame(volume_t= seq(log(0.1),log(3129321),0.002076463), ant_t = "other")
newdat <- rbind(newdat1, newdat1, newdat2, newdat3)

# use the predict function for to get the model output
y_pred <- predict(surv_model, newdata = newdat, type = "response")
y_CI <- predict(surv_model, newdata = newdat, interval = "confidence", type = "link", se.fit=TRUE)

#This is one way to pull out the link function automatically from the fitted object
linkinv <- family(surv_model_ant_interact)$linkinv ## inverse-link function

# you can also define your own link function, which is useful for stuff like Stan, where you just get posterior draws, and not a R model object
# and so for my other project, where I have categorical data, I define these functions at the top of the script
# invlogit<-function(x){exp(x)/(1+exp(x))}
# and then use that transform the model predictions to the appropriate scale
# I think for poisson, you can use exp() to exponentiate the results

#Now make a dataframe that combines all the predictions and confidence intervals
newpred <- newdat
newpred$pred0 <- y_CI$fit
newpred$pred <- linkinv(y_CI$fit) #This is the link function
alpha <- 0.95
sc <- abs(qnorm((1-alpha)/2))  ## Normal approx. to likelihood
alpha2 <- 0.5
sc2 <- abs(qnorm((1-alpha2)/2))  ## Normal approx. to likelihood
newpred <- transform(newpred,
                     lwr=linkinv(pred0-sc*y_CI$se.fit),
                     upr=linkinv(pred0+sc*y_CI$se.fit),
                     lwr2=linkinv(pred0-sc2*y_CI$se.fit),
                     upr2=linkinv(pred0+sc2*y_CI$se.fit))

# Then you can plot it
ggplot() + 
  geom_point(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2) + 
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t))) +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) + 
  geom_smooth(data=cactus, aes(y=Survival_t1, x=log(volume_t), col=ant_t), method = "glm", formula='y~x', method.args = list(family = "binomial"), se = TRUE)
  #geom_smooth(data = cactus, aes(x = log(volume_t), y = Survival_t1, pch = as.factor(ant_t)))
```

Ok, so it appears that this follows the general trend, sort of.

```{r}
ELVI_outcomeyear <- ggplot() +
  geom_point(data = cactus, aes(x = volume_t, y =Survival_t1, pch = as.factor(ant_t)), color = "#4c90dd", lwd = 2)+
  geom_line(data = newpred, aes(x = volume_t, y = pred, group = ant_t, linetype = as.factor(ant_t)), color = "#4c90dd") +
  geom_ribbon(data = newpred, aes(x = volume_t, ymin = lwr, ymax = upr, group = ant_t, fill = as.factor(ant_t)), alpha = .2) +
  theme_classic() + labs(y = "Mean Inflorescence Length", x = "Year", linetype = "Endophyte_status", shape = "Endophyte_status")+
  scale_linetype_manual(breaks = c("crem", "liom", "vacant", "other"),
                        values = c("solid", "dashed", "solid", "dashed"))+
  scale_shape_manual(breaks = c("crem", "liom", "vacant", "other"),
                     values = c(20,1))+
  scale_fill_manual(breaks = c("crem", "liom", "vacant", "other"),
                    values = c(  "#4c90dd", "#636363", "red", "green")) +
  xlim(1910,2016) + guides(fill = FALSE)

ELVI_outcomeyear
ggsave(ELVI_outcome, filename = "~/Documents/ELVIinflyear.tiff", width = 4, height = 3)
```

### 2021-02-08 Stan/Practice.Rmd
Stan work is here

---
title: "Stan Practice"
author: "Ali Campbell"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
data <- data.frame(cactus$volume_t, cactus$volume_t1, cactus$ant_t, cactus$Goodbuds_t, cactus$Goodbuds_t1, cactus$ABFlowerbuds_t, cactus$ABFlowerbuds_t1, cactus$TotFlowerbuds_t, cactus$TotFlowerbuds_t1, cactus$Survival_t1)
data <- na.omit(data)
colnames(data) <- c("volume_t","volume_t1", "ant_t", "Goodbuds_t", "Goodbuds_t1", "ABFlowerbuds_t", "ABFlowerbuds_t1", "TotFlowerbuds_t", "TotFlowerbuds_t1", "Survival_t1")
```

## Basic Models (No Ant Effects)
```{r Growth Model}
## Now rename variables and index
x <- log(data$volume_t)
y <- log(data$volume_t1)
N <- length(data$volume_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
#summary(grow_model)
#summary(mod1)
## Now we can extract some important statistics
lm_alpha <- summary(grow_model)$coeff[1]  # the intercept
lm_beta <- summary(grow_model)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model)  # the residual error
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple linear regression

data {
 int < lower = 1 > N; // Sample size
 vector[N] x; // Predictor
 vector[N] y; // Outcome
}

parameters {
 real alpha; // Intercept
 real beta; // Slope (regression coefficients)
 real < lower = 0 > sigma; // Error SD
}

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~ normal(alpha + x * beta , sigma); // The Model
}

generated quantities {
} // The posterior predictive distribution",

"grow_lm1.stan")
## Check that you wrote a file
#stanc("grow_lm1.stan")
## Save the file path
stanny <- "grow_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

par(mfrow = c(1,3))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)

plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

These appear pretty good. Roughly centered around the previous estimates.

```{r Total Flowers}
x <- log(data$volume_t)
y <- data$TotFlowerbuds_t
N <- length(data$volume_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
summary(flower_model)
#summary(mod1)
## Now we can extract some important statistics
lm_lambda = mean(data$TotFlowerbuds_t)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple linear regression

data {
 int N;
 real x[N];
 int y[N];
}
parameters {
 real<lower = 0> lambda;
}
model {
// Priors
 lambda ~ cauchy(0, 1);
 //lambda ~ gamma(0, 0);
 // Model
 y ~ poisson(lambda);
}
",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

##plot(density(posterior$lambda), main = "Lambda")
##abline(v = lm_lambda, col = 4, lty = 2)
```

```{r Survival}
x <- log(data$volume_t)
y <- data$Survival_t1
N <- length(data$volume_t)
#import things from the r regression models for comparison
lm_alpha <- summary(surv_model)$coeff[1]  # the intercept
lm_beta <- summary(surv_model)$coeff[2]  # the slope
summary(surv_model)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple linear regression

data {
  int<lower=0> N;
  vector[N] x;
  int<lower=0,upper=1> y[N];
}
parameters {
  real alpha;
  real beta;
}
model {
  y ~ bernoulli_logit(alpha + beta * x);
}
",

"survr_lm1.stan")
## Check that you wrote a file
stanc("surv_lm1.stan")
## Save the file path
stanny <- "surv_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

par(mfrow = c(1,2))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
```

#### Mixed models
```{r}
## Write in Data
data <- data.frame(cactus$volume_t, cactus$volume_t1, cactus$ant_t, cactus$TotFlowerbuds_t)
data <- na.omit(data)
colnames(data) <- c("volume_t","volume_t1", "ant_t", "TotFlowerbuds_t", "Survival_t1")
## Write the models in
totflow_stan_mixex_ant <- stan_glmer(TotFlowerbuds_t ~log(volume_t) + (1|ant_t), family = "poisson", data = cactus, chains = 4, cores = 4)
totflow_stan_mixed_ant_interact <- stan_glmer(TotFlowerbuds_t ~log(volume_t)*(1|ant_t), family = "poisson", data = cactus, chains = 4, cores = 4)
## Now check AIC values
AIC(totflow_stan)
AIC(totflow_stan_ant)
AIC(totflow_stan_ant_interact)
#Note that all of these say 0
## Check convergence
plot(totflow_stan, plotfun = "trace")
plot(totflow_stan_ant, plotfun = "trace")
plot(totflow_stan_ant_interact, plotfun = "trace")

##Now Code it in stan
write("// Stan model for simple linear regression

data {
  int<lower=1> N;            //number of observations
  real RT[N];                //reaction times

  int<lower=0,upper=9> Days[N];   //predictor (days of sleep deprivation)

  // grouping factor
  int<lower=1> J;                   //number of subjects
  int<lower=1,upper=J> Subject[N];  //subject id
}

parameters {
  vector[2] beta;                   // fixed-effects parameters
  real<lower=0> sigma_e;            // residual std
  vector<lower=0>[2] sigma_u;       // random effects standard deviations

  // declare L_u to be the Choleski factor of a 2x2 correlation matrix
  cholesky_factor_corr[2] L_u;

  matrix[2,J] z_u;                  // random effect matrix
}

transformed parameters {
  // this transform random effects so that they have the correlation
  // matrix specified by the correlation matrix above
  matrix[2,J] u;
  u = diag_pre_multiply(sigma_u, L_u) * z_u;

}

model {
  real mu; // conditional mean of the dependent variable

  //priors
  L_u ~ lkj_corr_cholesky(1.5); // LKJ prior for the correlation matrix
  to_vector(z_u) ~ normal(0,2);
  sigma_e ~ normal(0, 5);       // prior for residual standard deviation
  beta[1] ~ normal(0.3, 0.5);   // prior for fixed-effect intercept
  beta[2] ~ normal(0.2, 2);     // prior for fixed-effect slope

  //likelihood
  for (i in 1:N){
    mu = beta[1] + u[1,Subject[i]] + (beta[2] + u[2,Subject[i]])*Days[i];
    RT[i] ~ normal(mu, sigma_e);
  }
} // The posterior predictive distribution

generated quantities {
  matrix[2, 2] Omega;
  Omega = L_u * L_u'; // so that it return the correlation matrix
}",

"totflow_mixed1.stan")

stan_model1 <- "totflow_mixed1.stan"
## Now we fit the model
fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
```

### 2021-02-08 Stan/Stan Mix Models.Rmd
Stan work is here

---
title: "Mixed Models Stan"
author: "Ali Campbell"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
```

#### Mixed models
```{r growth}
## Write in Data
vol <- log(data$volume_t)
vol1 <- log(data$volume_t1)
ant <- as.numeric(data$ant_t)
plot <- as.factor(data$Plot)
year <- data$Year_t
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- unique(data$ant_t)
J <- unique(data$Plot)
L <- unique(data$Year_t)
## Now make it a dataframe
stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
write("// Stan model for simple linear regression

data {
  int <lower = 0> N; // number of observations
  int <lower = 0> K; // number of ant species
  real < lower = 0 > vol[N]; // volume t
  real < lower = 0 > vol1[N]; // volume t1
  int < lower = 0 > ant[K]; // ants
  int <lower=0,upper=1> y[N]; // outcome vector
  real rt[N];
  int<lower=1> J; //number of plots
  int<lower=1> L; //number of years
  int<lower=1, upper=J> plot[N]; // plot
  int<lower=1, upper=L> year[N]; // year
}

parameters {
  real alpha; // intercept
  vector[K] beta; // coefficients for ants
  vector[J] u; //subject intercepts
  vector[L] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters {
  real mu;
}

model {
  // Priors
  alpha ~ normal(0,100);
  beta ~ normal(0,100); 
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  //sigma ~ exponential(0,0);
  // The Model
  for (i in 1:N){
    mu = beta[1] + u[plot[i]] + w[year[i]] //+ beta[2] * vol[i] + beta[3]*ant[i];
    rt[i] ~ lognormal(mu, sigma_e);
  }
}",

"grow_mixed1.stan")

stan_model1 <- "grow_mixed1.stan"
## Now we fit the model
fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
```

```{r}
## Write in Data
vol <- log(data$volume_t)
vol1 <- log(data$volume_t1)
ant <- as.numeric(data$ant_t)
plot <- as.factor(data$Plot)
year <- data$Year_t
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- as.factor(unique(data$ant_t))
J <- unique(as.factor(data$Plot))
L <- unique(data$Year_t)
## Now make it a dataframe
stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
write("// Stan model for simple mixed model growth regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> K; // number of ant species
  real < lower = log(0) > vol[N]; // volume t
  int < lower = 1 > ant[K]; // ants
  real rt[N]; // response variable
  int < lower=1 > J; //number of plots
  int < lower=1 > L; //number of years
  int < lower=1, upper=J > plot[N]; // plot
  int < lower=1, upper=L > year[N]; // year
}

parameters {
  real alpha; // intercept
  vector[N] beta; // coefficients for ants
  vector[J] u; //subject intercepts
  vector[L] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters {
  real mu[N];
  for (i in 1:N)
    mu[i] = beta[1] + u[plot[i]] + w[year[i]]+ beta[2] * vol[i] + beta[3]*ant[i];
}

model {
  // Priors
  alpha ~ normal(0,100);
  beta ~ normal(0,100); 
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  //sigma ~ exponential(0,0);
  // The Model
  for(i in 1:N)
    rt[i] ~ lognormal(mu[i], sigma);
}",

"grow_mixed1.stan")

stan_model1 <- "grow_mixed1.stan"
## Now we fit the model
fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
```

### 2021-02-16 Stan/Stan Base Models.Rmd
Mixed Models (With Ant Interaction)

@@
 ```{r}
-data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1")]
+data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1")]
 data <- na.omit(data)
@@
 }
+
+flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1")]
+flower <- na.omit(flower)
+flower$ant <- as.integer(flower$ant_t)
 ```
@@
 ## Basic Models (No Ant Effects)
+####Growth
 ```{r Growth Model}
 ## Now rename variables and index
-x <- log(data$volume_t)
-z <- as.numeric(data$ant_t)
-y <- log(data$volume_t1)
-N <- length(data$volume_t)
-K <- length(data$ant_t)
+vol <- data$volume_t
+y <- data$volume_t1
+N <- as.integer(nrow(data))
 ## Rerun linear model with that new data
@@
 ## Make it a dataframe
-stan_data <- list(N = N, x = x, y = y, z = z, K = K)
+stan_data <- list(N = N, vol = vol, y = y)
 ## Write the model in stan
@@
  int < lower = 1 > N; // Sample size
- vector[N] x; // Predictor
+ vector[N] vol; // Predictor
  vector[N] y; // Outcome
@@
 //sigma ~ exponential(0,0);
- y ~ normal(alpha + x * beta , sigma); // The Model
+ y ~ normal(alpha + vol * beta , sigma); // The Model
 }
@@
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
-fit
+fit_grow <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_grow
 #class(fit)
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
+posterior_grow <- as.data.frame(fit_grow)
+print(colnames(posterior_grow))
 
 par(mfrow = c(1,3))
-plot(density(posterior$alpha), main = "Alpha")
+plot(density(posterior_grow$alpha), main = "Alpha")
 abline(v = lm_alpha, col = 4, lty = 2)
-plot(density(posterior$beta), main = "Beta")
+plot(density(posterior_grow$beta), main = "Beta", xlim = c(0.85,0.93))
 abline(v = lm_beta, col = 4, lty = 2)
-plot(density(posterior$sigma), main = "Sigma")
+plot(density(posterior_grow$sigma), main = "Sigma")
 abline(v = lm_sigma, col = 4, lty = 2)
@@
 
+####Total Flowers
 ```{r Total Flowers}
-x <- log(data$volume_t)
-y <- data$TotFlowerbuds_t
-N <- length(data$volume_t)
-## Rerun linear model with that new data
-#mod1 <- lm(y~x, data = data)
-summary(flower_model)
-#summary(mod1)
-## Now we can extract some important statistics
-lm_lambda = mean(data$TotFlowerbuds_t)
-## Make it a dataframe
-stan_data <- list(N = N, x = x, y = y)
-## Write the model in stan
-#note that we are using flat priors here which are non restrictive
+## Ok so this one is difficult because the data does not include this variable because of all of the NAs. 
+## Ok so I have created a second data set. Obviously this will be a problem down the line when it all has to go into one model, but here we go for now. 
+## Next problem is that the dimensions do not work now. 
+y <- flower$TotFlowerbuds_t
+N <- nrow(flower)
+stan_data <- list(N = N, ## number of observations
+                  y = y ## response survival next year
+                  )
+
 write("// Stan model for simple total flowers regression
@@
 data {
- int N;
- real x[N];
- int y[N];
+ int < lower = 1 > N; // Sample size
+ int y[N]; // Outcome
 }
 parameters {
- real<lower = 0> lambda;
+  real lambda; //intercept, unique to ant
 }
@@
 // Priors
- lambda ~ cauchy(0, 1);
- //lambda ~ gamma(0, 0);
+ lambda ~ beta(1,1);
  // Model
@@
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
-fit
+fit_flow <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_flow
 
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
+posterior_flow <- as.data.frame(fit_flow)
+print(colnames(posterior_flow))
 
-plot(density(posterior$lambda), main = "Lambda")
+plot(density(posterior_flow$lambda), main = "Lambda")
 abline(v = lm_lambda, col = 4, lty = 2)
 ```
+####Abort
+```{r abort model}
+y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
+vol = flower$volume_t
+N = nrow(flower)
+stan_data <- list(N = N, ## number of observations
+                  y = y, ## response survival next year
+                  vol = vol
+                  )
 
+write("// Stan model for simple total flowers regression
+
+data {
+ int < lower = 1 > N; // Sample size
+ int <lower = 0, upper = 1> y[N]; // Outcome
+ vector[N] vol;	//size_t
+}
+parameters {
+  real beta0; //intercept, unique to ant sp
+  real beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
+}
+model {
+ // Priors
+ beta0 ~ normal(0,100);
+ beta1 ~ normal(0,100); 
+ // Model
+ y ~ bernoulli_logit(beta0 + beta1 * vol);
+}
+",
+
+"abort_lm1.stan")
+## Check that you wrote a file
+stanc("abort_lm1.stan")
+## Save the file path
+stanny <- "abort_lm1.stan"
+## Now we fit the model
+fit_abort <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_abort
+
+posterior_abort <- as.data.frame(fit_abort)
+print(colnames(posterior_abort))
+```
+
+####Survival
 ```{r Survival}
-x <- log(data$volume_t)
+vol = data$volume_t
 y <- data$Survival_t1
-N <- length(data$volume_t)
+N <- nrow(data)
 #import things from the r regression models for comparison
@@
 data {
-  int<lower=0> N;
-  vector[N] x;
-  int<lower=0,upper=1> y[N];
+  int <lower=0> N;
+  vector[N] vol;
+  int <lower=0,upper=1> y[N];
 }
@@
  // Model
- y ~ bernoulli_logit(alpha + beta * x);
+ y ~ bernoulli_logit(alpha + beta * vol);
 }
@@
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
-fit
+fit_surv <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_surv
 
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
+posterior_surv <- as.data.frame(fit_surv)
+print(colnames(posterior_surv))
 
@@
 
-plot(density(posterior$alpha), main = "Alpha")
+plot(density(posterior_surv$alpha), main = "Alpha")
 abline(v = lm_alpha, col = 4, lty = 2)
 
-plot(density(posterior$beta), main = "Beta")
+plot(density(posterior_surv$beta), main = "Beta")
 abline(v = lm_beta, col = 4, lty = 2)
@@
 
-#### Include ants
-######Growth
-```{r Growth Model}
-## Now rename variables and index
-vol <- log(data$volume_t)
-ant <- as.numeric(data$ant_t)
-y <- log(data$volume_t1)
-N <- length(data$volume_t)
-K <- length(data$ant_t)
-## Make it a dataframe
-stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
-lm_alpha <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
-lm_beta <- summary(grow_model_ant_interact)$coeff[2]  # the slope
-lm_sigma <- sigma(grow_model_ant_interact)  # the residual error
+####Repro
 
-write("// Stan model for ant growth regression
+```{r}
+vol1 <- flower$volume_t1
+y = flower$repro_state_t1
+N = nrow(flower)
+stan_data <- list(N = N, ## number of observations
+                  y = y, ## response survival next year
+                  vol1 = vol1
+                  )
+
+write("// Stan model for simple total flowers regression
 
@@
  int < lower = 1 > N; // Sample size
- int < lower = 1 > K; // Number of Predictors (Ant Species)
- real < lower = 0 > vol[N]; // volume
- int < lower = 0 > ant[K]; // ants
- real < lower = 0 > y[N]; // outcome vector
+ int <lower = 0, upper = 1> y[N]; // Outcome
+ vector[N] vol1;	//size_t
 }
+parameters {
+  real beta0; //intercept, unique to ant sp
+  real beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
+}
+model {
+ // Priors
+ beta0 ~ normal(0,100);
+ beta1 ~ normal(0,100); 
+ // Model
+ y ~ bernoulli_logit(beta0 + beta1 * vol1);
+}
+",
 
+"repro_lm1.stan")
+
+## Check that you wrote a file
+stanc("repro_lm1.stan")
+## Save the file path
+stanny <- "repro_lm1.stan"
+## Now we fit the model
+fit_repro <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_repro
+
+posterior_repro <- as.data.frame(fit_repro)
+print(colnames(posterior_repro))
+```
+
+## Include ants
+####Growth
+```{r Growth Model}
+vol = data$volume_t
+y = data$volume_t1
+N = nrow(data)
+N_ant = 4
+ant = data$ant
+stan_data <- list(N = N, ## number of observations
+                  vol = vol, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant
+                  ) 
+## Write the model in stan
+#note that we are using flat priors here which are non restrictive
+write("// Stan model for simple growth regression
+
+data {
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  vector[N] y; // size_t1
+}
 parameters {
- real alpha; // Intercept
- vector[K] beta; // coefficients for predictors
- real < lower = 0 > sigma; // Error SD
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
 }
@@
 transformed parameters{
-real mu[N];
+  vector[N] mu; //linear predictor for the mean
   for(i in 1:N){
-mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
   }
@@
 model {
-// Priors
-alpha ~ normal(0,100);
-beta ~ normal(0,100); 
-//sigma ~ exponential(0,0);
- y ~  normal(mu, sigma);   // The Model
-}
-
-generated quantities {
-} // The posterior predictive distribution",
+  y ~ normal(mu, sigma); // likelihood
+}",
 
@@
 ## Save the file path
-stanny <- "grow_ant_lm1.stan"
+#stanny <- "grow_ant_lm1.stan"
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000 , chains = 4, cores = 2, thin = 1)
-fit
+fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_grow_ant
+#class(fit)
+posterior_grow_ant <- as.data.frame(fit_grow_ant)
+print(colnames(posterior_grow_ant))
+lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
+lm_sigma <- sigma(grow_model_ant_interact)  # the residual error
 
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
-
-par(mfrow = c(1,2))
-
-plot(density(posterior$alpha), main = "Alpha")
-abline(v = lm_alpha, col = 4, lty = 2)
-
-#plot(density(posterior$beta[i]), main = "Beta")
-#abline(v = lm_beta, col = 4, lty = 2)
-
-plot(density(posterior$sigma), main = "Sigma")
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_grow_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_grow_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_grow_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+#Sigma
+plot(density(posterior_grow_ant$sigma), main = "Sigma")
 abline(v = lm_sigma, col = 4, lty = 2)
@@
 
-```{r grow attempt 1}
+####Survival
+
+```{r survival}
 vol = data$volume_t
-y = data$volume_t1
-crem = data$crem
-liom = data$liom
-vacant = data$vacant
-other = data$other
+y = data$Survival_t1
 N = nrow(data)
+N_ant = 4
+ant = data$ant
 stan_data <- list(N = N, ## number of observations
                   vol = vol, ## predictors volume
-                  y = y, ## response volume next year
-                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants
+                  y = y, ## response survival next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant
+                  )
 
-write("
-      data{
-      int <lower = 1> N; // Number of observations
-      vector[N] vol; // volume predictor
-      int <lower = 0, upper = 1> crem[N]; // crem predictor
-      int <lower = 0, upper = 1> liom[N]; // liom predictor
-      int <lower = 0, upper = 1> other[N]; // other predictor
-      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
-      vector[N] y; // outcome
-      }
-      parameters{
-      real alpha; // intercept 
-      real b_crem; // crem slope
-      real b_liom; // liom slope
-      real b_other; // other slope
-      real b_vacant; // vacant slope
-      real <lower = 0> sigma; // standard deviation
-      }
-      model{
-      // Priors
-      alpha ~ normal(0,100);
-      beta ~ normal(0,100);
-      // Likelihood
-      y ~ normal(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol ,sigma);
-      }
-      ",
-      "tester1.stan")
-
-stanc("tester1.stan")
-
-fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
-fit
-
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
-par(mfrow = c(2,3))
-plot(density(posterior$alpha), main = "Alpha")
-abline(v = lm_alpha, col = 4, lty = 2)
-plot(density(posterior$b_other), main = "Beta Other")
-abline(v = lm_beta, col = 4, lty = 2)
-plot(density(posterior$b_vacant), main = "Beta Vacant")
-abline(v = lm_beta, col = 4, lty = 2)
-plot(density(posterior$b_crem), main = "Beta Crem")
-abline(v = lm_beta, col = 4, lty = 2)
-plot(density(posterior$b_liom), main = "Beta Liom")
-abline(v = lm_beta, col = 4, lty = 2)
-plot(density(posterior$sigma), main = "Sigma")
-abline(v = lm_sigma, col = 4, lty = 2)
-```
-
-```{r grow attempt 2&3}
-vol <- log(data$volume_t)
-ant <- as.numeric(data$ant_t)
-y <- log(data$Survival_t1)
-N <- length(data$volume_t)
-K <- length(data$ant_t)
-## Make it a dataframe
-stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
-## Write the model in stan
-lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
-lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
-lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
-write("// Stan model for simple linear regression
+write("// Stan model for simple ant survival regression
 
 data {
- int < lower = 1 > N; // Sample size
- int < lower = 1 > K; // Number of Predictors (Ant Species)
- real < lower = 0 > vol[N]; // volume
- int < lower = 0 > ant[K]; // ants
- int <lower=0,upper=1> y[N]; // outcome vector
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
 }
@@
 parameters {
-  real alpha;
-  vector[K] beta; // coefficients for predictors
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
 }
@@
 transformed parameters{
- real mu[N];
+  vector[N] mu; //linear predictor for the mean
   for(i in 1:N){
- mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
   }
@@
 //Priors
- alpha ~ normal(0,100);
- beta ~ normal(0,100);
+ beta0 ~ normal(0,100);
+ beta1 ~ normal(0,100);
  //Model
@@
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
-fit
+fit_surv_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_surv_ant
 
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
+#class(fit)
+posterior_surv_ant <- as.data.frame(fit_surv_ant)
+print(colnames(posterior_surv_ant))
+lm_alpha_other <- summary(surv_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(surv_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(surv_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(surv_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(surv_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(surv_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(surv_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(surv_model_ant_interact)$coeff[8]
+lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
 
-plot(density(posterior$alpha), main = "Alpha")
-abline(v = lm_alpha, col = 4, lty = 2)
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_surv_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_surv_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_surv_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+#Sigma
+plot(density(posterior_surv_ant$sigma), main = "Sigma")
+abline(v = lm_sigma, col = 4, lty = 2)
 ```
 
-######Survival
+####Total Flowers
 
-```{r survival}
-vol <- log(data$volume_t)
-ant <- as.numeric(data$ant_t)
-y <- log(data$Survival_t1)
-N <- length(data$volume_t)
-K <- length(data$ant_t)
-## Make it a dataframe
-stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
-## Write the model in stan
-lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
-lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
-lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
-write("// Stan model for simple ant survival regression
+```{r total flowers}
+y <- flower$TotFlowerbuds_t
+vol = flower$volume_t
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
+stan_data <- list(N = N, ## number of observations
+                  y = y, ## response survival next year
+                  vol = vol,
+                  N_ant = N_ant,
+                  ant = ant
+)
+
+write("// Stan model for simple total flowers regression
 
 data {
- int < lower = 1 > N; // Sample size
- int < lower = 1 > K; // Number of Predictors (Ant Species)
- real < lower = 0 > vol[N]; // volume
- int < lower = 0 > ant[K]; // ants
- int <lower=0,upper=1> y[N]; // outcome vector
+ int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0> y[N]; // survival in year t1
 }
@@
 transformed parameters{
- real mu[N];
+  vector[N] mu; //linear predictor for the mean
   for(i in 1:N){
- mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
   }
 }
+model {
+ // Model
+ y ~ poisson(mu);
+}
+",
 
+"flower_lm1.stan")
+## Check that you wrote a file
+stanc("flower_lm1.stan")
+## Save the file path
+stanny <- "flower_lm1.stan"
+## Now we fit the model
+fit_flow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_flow_ant
+
+posterior_flow_ant <- as.data.frame(fit_flow_ant)
+print(colnames(posterior_flow_ant))
+
+lm_alpha_other <- summary(flower_model_ant)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(flower_model_ant)$coeff[3]
+lm_alpha_liom <- summary(flower_model_ant)$coeff[4]
+lm_alpha_vacant <- summary(flower_model_ant)$coeff[5]
+lm_beta_other <- summary(flower_model_ant)$coeff[2]  # the slope
+lm_beta_crem <- summary(flower_model_ant)$coeff[6]
+lm_beta_liom <- summary(flower_model_ant)$coeff[7]
+lm_beta_vacant <- summary(flower_model_ant)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_flow_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_flow_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_flow_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+```
+
+####Abortion Model
+
+```{r abortion model}
+y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
+vol = flower$volume_t
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
+stan_data <- list(N = N, ## number of observations
+                  y = y, ## response survival next year
+                  vol = vol,
+                  N_ant = N_ant,
+                  ant = ant
+)
+
+write("// Stan model for simple total flowers regression
+
+data {
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
+}
+parameters {
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
+}
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
+  }
+}
 model {
-//Priors
- alpha ~ normal(0,100);
- beta ~ normal(0,100);
- //Model
+ // Priors
+ beta0 ~ normal(0,100);
+ beta1 ~ normal(0,100); 
+ // Model
  y ~ bernoulli_logit(mu);
@@
 
-"surv_ant_lm1.stan")
+"abort_ant_lm1.stan")
 ## Check that you wrote a file
-stanc("surv_ant_lm1.stan")
+stanc("abort_ant_lm1.stan")
 ## Save the file path
-stanny <- "surv_ant_lm1.stan"
+stanny <- "abort_ant_lm1.stan"
 ## Now we fit the model
-fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
-fit
+fit_abort_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_abort_ant
 
-posterior <- as.data.frame(fit)
-print(colnames(posterior))
+posterior_abort_ant <- as.data.frame(fit_abort_ant)
+print(colnames(posterior_abort_ant))
 
-plot(density(posterior$alpha), main = "Alpha")
-abline(v = lm_alpha, col = 4, lty = 2)
+lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
+lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
+lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
+lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
+lm_beta_crem <- summary(abort_model_ant)$coeff[6]
+lm_beta_liom <- summary(abort_model_ant)$coeff[7]
+lm_beta_vacant <- summary(abort_model_ant)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_abort_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_abort_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_abort_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
 ```
 
-```{r surv attempt 1}
-vol = data$volume_t
-y = as.integer(data$Survival_t1)
-crem = data$crem
-liom = data$liom
-vacant = data$vacant
-other = data$other
-N = nrow(data)
+####Repro Model
+```{r}
+vol1 <- flower$volume_t1
+y = flower$repro_state_t1
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
 stan_data <- list(N = N, ## number of observations
-                  vol = vol, ## predictors volume
-                  y = y, ## response volume next year
-                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants
+                  y = y, ## response survival next year
+                  vol1 = vol1,
+                  N_ant = N_ant,
+                  ant = ant
+)
 
-write("
-      data{
-      int <lower = 1> N; // Number of observations
-      vector[N] vol; // volume predictor
-      int <lower = 0, upper = 1> crem[N]; // crem predictor
-      int <lower = 0, upper = 1> liom[N]; // liom predictor
-      int <lower = 0, upper = 1> other[N]; // other predictor
-      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
-      int <lower = 0, upper = 1> y[N]; // outcome
+write("// Stan model for simple total flowers regression
+
+data {
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol1;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
 }
-      parameters{
-      real alpha; // intercept 
-      real b_crem; // crem slope
-      real b_liom; // liom slope
-      real b_other; // other slope
-      real b_vacant; // vacant slope
-      real <lower = 0> sigma; // standard deviation
+
+parameters {
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  real<lower = 0> sigma;
 }
-      model{
+
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i];
+  }
+}
+model {
  // Priors
-      alpha ~ normal(0,100);
-      // Likelihood
-      y ~ bernoulli_logit(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol);
+ beta0 ~ normal(0,100);
+ beta1 ~ normal(0,100); 
+ // Model
+ y ~ bernoulli_logit(mu);
 }
 ",
-      "tester1.stan")
 
-stanc("tester1.stan")
+"repro_lm1.stan")
 
-fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
-fit
+## Check that you wrote a file
+stanc("repro_lm1.stan")
+## Save the file path
+stanny <- "repro_lm1.stan"
+## Now we fit the model
+fit_repro_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_repro_ant
+
+posterior_repro_ant <- as.data.frame(fit_repro_ant)
+print(colnames(posterior_repro_ant))
+
+lm_alpha_other <- summary(model3)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(model3)$coeff[3]
+lm_alpha_liom <- summary(model3)$coeff[4]
+lm_alpha_vacant <- summary(model3)$coeff[5]
+lm_beta_other <- summary(model3)$coeff[2]  # the slope
+lm_beta_crem <- summary(model3)$coeff[6]
+lm_beta_liom <- summary(model3)$coeff[7]
+lm_beta_vacant <- summary(model3)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_repro_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_repro_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_repro_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
 ```

### 2021-02-16 Stan/Stan Mix Models.Rmd
Mixed Models (With Ant Interaction)

@@
 }
+data$Year_t <- as.factor(data$Year_t)
+data$year <- as.integer(data$Year_t)
+data$Plot <- as.factor(data$Plot)
+data$plot <- as.integer(data$Plot)
+
+flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1", "Year_t", "Plot")]
+flower <- na.omit(flower)
+flower$ant <- as.integer(flower$ant_t)
+flower$Year_t <- as.factor(flower$Year_t)
+flower$year <- as.integer(flower$Year_t)
+flower$Plot <- as.factor(flower$Plot)
+flower$plot <- as.integer(flower$Plot)
 ```
 
-#### Mixed models
-```{r growth}
-## Write in Data
-vol <- log(data$volume_t)
-vol1 <- log(data$volume_t1)
-ant <- as.numeric(data$ant_t)
-plot <- as.factor(data$Plot)
-year <- data$Year_t
-y <- log(data$Survival_t1)
-N <- length(data$volume_t)
-K <- unique(data$ant_t)
-J <- unique(data$Plot)
-L <- unique(data$Year_t)
-## Now make it a dataframe
-stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
-write("// Stan model for simple linear regression
+#### Growth
+```{r Growth Model}
+vol = data$volume_t
+y = data$volume_t1
+N = nrow(data)
+N_ant = 4
+ant = data$ant
+N_Year <- max(data$year)
+N_Plot <- max(data$plot)
+plot <- data$plot
+year <- data$year
+stan_data <- list(N = N, ## number of observations
+                  vol = vol, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant, ## number of ant states
+                  N_Year = N_Year, ## number of years
+                  N_Plot = N_Plot, ## number of plots
+                  plot = plot, ## predictor plots
+                  year = year ## predictor years
+                  ) 
+## Write the model in stan
+#note that we are using flat priors here which are non restrictive
+write("// Stan model for simple growth regression
 
 data {
-  int <lower = 0> N; // number of observations
-  int <lower = 0> K; // number of ant species
-  real < lower = 0 > vol[N]; // volume t
-  real < lower = 0 > vol1[N]; // volume t1
-  int < lower = 0 > ant[K]; // ants
-  int <lower=0,upper=1> y[N]; // outcome vector
-  real rt[N];
-  int<lower=1> J; //number of plots
-  int<lower=1> L; //number of years
-  int<lower=1, upper=J> plot[N]; // plot
-  int<lower=1, upper=L> year[N]; // year
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  vector[N] y; // size_t1
+  int<lower=1> N_Year; //number of plots
+  int<lower=1> N_Plot; //number of years
+  int<lower=1, upper=N_Plot> plot[N]; // plot
+  int<lower=1, upper=N_Year> year[N]; // year
 }
-
 parameters {
-  real alpha; // intercept
-  vector[K] beta; // coefficients for ants
-  vector[J] u; //subject intercepts
-  vector[L] w; //item intercepts
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  vector[N_Plot] u; //subject intercepts
+  vector[N_Year] w; //item intercepts
   real < lower = 0 > sigma; // Error SD
@@
 
-transformed parameters {
-  real mu;
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
+  }
 }
@@
 model {
-  // Priors
-  alpha ~ normal(0,100);
-  beta ~ normal(0,100); 
   u ~ normal(0, sigma_u); // plot random effects
   w ~ normal(0, sigma_w); // year random effects
-  //sigma ~ exponential(0,0);
-  // The Model
-  for (i in 1:N){
-    mu = beta[1] + u[plot[i]] + w[year[i]] //+ beta[2] * vol[i] + beta[3]*ant[i];
-    rt[i] ~ lognormal(mu, sigma_e);
+  beta0 ~ normal(0,100); // intercept distribution
+  beta1 ~ normal(0,100); // slope distribution
+  for(i in 1:N){
+    y[i] ~ lognormal(mu[i], sigma);
   }
@@
 
-"grow_mixed1.stan")
+"grow_ant_mix_lm1.stan")
+## Check that you wrote a file
+stanc("grow_ant_mix_lm1.stan")
+## Save the file path
+#stanny <- "grow_ant_mix_lm1.stan"
+## Now we fit the model
+fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_grow_mix_ant
+#class(fit)
+posterior_grow_mix_ant <- as.data.frame(fit_grow_mix_ant)
+print(colnames(posterior_grow_mix_ant))
+lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
+lm_sigma <- sigma(grow_model_ant_interact)  # the residual error
 
-stan_model1 <- "grow_mixed1.stan"
-## Now we fit the model
-fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
-fit
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_grow_mix_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_grow_mix_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_grow_mix_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+#Sigma
+plot(density(posterior_grow_mix_ant$sigma), main = "Sigma")
+abline(v = lm_sigma, col = 4, lty = 2)
 ```
 
-```{r}
-## Write in Data
-vol <- log(data$volume_t)
-vol1 <- log(data$volume_t1)
-ant <- as.numeric(data$ant_t)
-plot <- as.factor(data$Plot)
-year <- data$Year_t
-y <- log(data$Survival_t1)
-N <- length(data$volume_t)
-K <- as.factor(unique(data$ant_t))
-J <- unique(as.factor(data$Plot))
-L <- unique(data$Year_t)
-## Now make it a dataframe
-stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
-write("// Stan model for simple mixed model growth regression
+####Survival
+```{r survival}
+vol = data$volume_t
+y = data$Survival_t1
+N = nrow(data)
+N_ant = 4
+ant = data$ant
+N_Year <- max(data$year)
+N_Plot <- max(data$plot)
+plot <- data$plot
+year <- data$year
+stan_data <- list(N = N, ## number of observations
+                  vol = vol, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant, ## number of ant states
+                  N_Year = N_Year, ## number of years
+                  N_Plot = N_Plot, ## number of plots
+                  plot = plot, ## predictor plots
+                  year = year ## predictor years
+                  )
+
+write("// Stan model for simple ant survival regression
 
@@
   int <lower = 1> N; // number of observations
-  int <lower = 1> K; // number of ant species
-  real < lower = log(0) > vol[N]; // volume t
-  int < lower = 1 > ant[K]; // ants
-  real rt[N]; // response variable
-  int < lower=1 > J; //number of plots
-  int < lower=1 > L; //number of years
-  int < lower=1, upper=J > plot[N]; // plot
-  int < lower=1, upper=L > year[N]; // year
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
+  int<lower=1> N_Year; //number of plots
+  int<lower=1> N_Plot; //number of years
+  int<lower=1, upper=N_Plot> plot[N]; // plot
+  int<lower=1, upper=N_Year> year[N]; // year
 }
@@
 parameters {
-  real alpha; // intercept
-  vector[N] beta; // coefficients for ants
-  vector[J] u; //subject intercepts
-  vector[L] w; //item intercepts
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  vector[N_Plot] u; //subject intercepts
+  vector[N_Year] w; //item intercepts
   real < lower = 0 > sigma; // Error SD
@@
 
-transformed parameters {
-  real mu[N];
-  for (i in 1:N)
-    mu[i] = beta[1] + u[plot[i]] + w[year[i]]+ beta[2] * vol[i] + beta[3]*ant[i];
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
+  }
 }
@@
 model {
-  // Priors
-  alpha ~ normal(0,100);
-  beta ~ normal(0,100); 
+//Priors
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
-  //sigma ~ exponential(0,0);
-  // The Model
-  for(i in 1:N)
-    rt[i] ~ lognormal(mu[i], sigma);
-}",
+ beta0 ~ normal(0,100); // intercept distribution
+ beta1 ~ normal(0,100); // slope distribution
+ //Model
+ for(i in 1:N){
+ y[i] ~ bernoulli_logit(mu[i]);
+ }
+}
+",
 
-"grow_mixed1.stan")
+"surv_ant_mix_lm1.stan")
+## Check that you wrote a file
+stanc("surv_ant_mix_lm1.stan")
+## Save the file path
+stanny <- "surv_ant_mix_lm1.stan"
+## Now we fit the model
+fit_surv_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_surv_mix_ant
 
-stan_model1 <- "grow_mixed1.stan"
+#class(fit)
+posterior_surv_mix_ant <- as.data.frame(fit_surv_mix_ant)
+print(colnames(posterior_surv_mix_ant))
+lm_alpha_other <- summary(surv_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(surv_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(surv_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(surv_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(surv_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(surv_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(surv_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(surv_model_ant_interact)$coeff[8]
+lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_surv_mix_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_surv_mix_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_surv_mix_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+#Sigma
+plot(density(posterior_surv_mix_ant$sigma), main = "Sigma")
+abline(v = lm_sigma, col = 4, lty = 2)
+```
+
+####Total Flowers
+
+```{r total flowers}
+y <- flower$TotFlowerbuds_t
+vol = flower$volume_t
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
+N_Year <- max(flower$year)
+N_Plot <- max(flower$plot)
+plot <- flower$plot
+year <- flower$year
+stan_data <- list(N = N, ## number of observations
+                  vol = vol, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant, ## number of ant states
+                  N_Year = N_Year, ## number of years
+                  N_Plot = N_Plot, ## number of plots
+                  plot = plot, ## predictor plots
+                  year = year ## predictor years
+                  )
+
+write("// Stan model for simple total flowers regression
+
+data {
+ int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0> y[N]; // survival in year t1
+  int<lower=1> N_Year; //number of plots
+  int<lower=1> N_Plot; //number of years
+  int<lower=1, upper=N_Plot> plot[N]; // plot
+  int<lower=1, upper=N_Year> year[N]; // year
+}
+
+parameters {
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  vector[N_Plot] u; //subject intercepts
+  vector[N_Year] w; //item intercepts
+  real < lower = 0 > sigma; // Error SD
+  real < lower = 0 > sigma_u; // plot SD
+  real < lower = 0 > sigma_w; // year SD
+}
+
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
+  }
+}
+model {
+ // Model
+  u ~ normal(0, sigma_u); // plot random effects
+  w ~ normal(0, sigma_w); // year random effects
+  beta0 ~ normal(0,100); // intercept distribution
+  beta1 ~ normal(0,100); // slope distribution
+  for(i in 1:N){
+    y[i] ~ poisson(mu[i]);
+  }
+}
+",
+
+"flower_ant_mix_lm1.stan")
+## Check that you wrote a file
+stanc("flower_ant_mix_lm1.stan")
+## Save the file path
+stanny <- "flower_ant_mix_lm1.stan"
 ## Now we fit the model
-fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
-fit
+fit_flow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_flow_mix_ant
+
+posterior_flow_ant <- as.data.frame(fit_flow_mix_ant)
+print(colnames(posterior_flow_mix_ant))
+
+lm_alpha_other <- summary(flower_model_ant)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(flower_model_ant)$coeff[3]
+lm_alpha_liom <- summary(flower_model_ant)$coeff[4]
+lm_alpha_vacant <- summary(flower_model_ant)$coeff[5]
+lm_beta_other <- summary(flower_model_ant)$coeff[2]  # the slope
+lm_beta_crem <- summary(flower_model_ant)$coeff[6]
+lm_beta_liom <- summary(flower_model_ant)$coeff[7]
+lm_beta_vacant <- summary(flower_model_ant)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_flow_mix_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_flow_mix_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_flow_mix_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
 ```
+
+####Abortion Model
+
+```{r abortion model}
+y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
+vol = flower$volume_t
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
+N_Year <- max(flower$year)
+N_Plot <- max(flower$plot)
+plot <- flower$plot
+year <- flower$year
+stan_data <- list(N = N, ## number of observations
+                  vol = vol, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant, ## number of ant states
+                  N_Year = N_Year, ## number of years
+                  N_Plot = N_Plot, ## number of plots
+                  plot = plot, ## predictor plots
+                  year = year ## predictor years
+                  )
+
+write("// Stan model for simple total flowers regression
+
+data {
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
+  int<lower=1> N_Year; //number of plots
+  int<lower=1> N_Plot; //number of years
+  int<lower=1, upper=N_Plot> plot[N]; // plot
+  int<lower=1, upper=N_Year> year[N]; // year
+}
+parameters {
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  vector[N_Plot] u; //subject intercepts
+  vector[N_Year] w; //item intercepts
+  real < lower = 0 > sigma; // Error SD
+  real < lower = 0 > sigma_u; // plot SD
+  real < lower = 0 > sigma_w; // year SD
+}
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
+  }
+}
+model {
+ // Priors
+  u ~ normal(0, sigma_u); // plot random effects
+  w ~ normal(0, sigma_w); // year random effects
+  beta0 ~ normal(0,100); // intercept distribution
+  beta1 ~ normal(0,100); // slope distribution
+ // Model
+ for(i in 1:N){
+  y[i] ~ bernoulli_logit(mu[i]);
+ }
+}
+",
+
+"abort_ant_mix_lm1.stan")
+## Check that you wrote a file
+stanc("abort_ant_mix_lm1.stan")
+## Save the file path
+stanny <- "abort_ant_mix_lm1.stan"
+## Now we fit the model
+fit_abort_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_abort_mix_ant
+
+posterior_abort_mix_ant <- as.data.frame(fit_abort_mix_ant)
+print(colnames(posterior_abort_mix_ant))
+
+lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
+lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
+lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
+lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
+lm_beta_crem <- summary(abort_model_ant)$coeff[6]
+lm_beta_liom <- summary(abort_model_ant)$coeff[7]
+lm_beta_vacant <- summary(abort_model_ant)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_abort_mix_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_abort_mix_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_abort_mix_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+```
+
+####Repro Model
+```{r}
+vol1 <- flower$volume_t1
+y = flower$repro_state_t1
+N = nrow(flower)
+N_ant = 4
+ant = flower$ant
+N_Year <- max(flower$year)
+N_Plot <- max(flower$plot)
+plot <- flower$plot
+year <- flower$year
+stan_data <- list(N = N, ## number of observations
+                  vol1 = vol1, ## predictors volume
+                  y = y, ## response volume next year
+                  ant = ant,## predictors ants
+                  N_ant = N_ant, ## number of ant states
+                  N_Year = N_Year, ## number of years
+                  N_Plot = N_Plot, ## number of plots
+                  plot = plot, ## predictor plots
+                  year = year ## predictor years
+                  )
+
+write("// Stan model for simple total flowers regression
+
+data {
+  int <lower = 1> N; // number of observations
+  int <lower = 1> N_ant; // number of ant states
+  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+  vector[N] vol1;	//size_t
+  int <lower = 0, upper = 1> y[N]; // survival in year t1
+  int<lower=1> N_Year; //number of plots
+  int<lower=1> N_Plot; //number of years
+  int<lower=1, upper=N_Plot> plot[N]; // plot
+  int<lower=1, upper=N_Year> year[N]; // year
+}
+
+parameters {
+  vector[N_ant] beta0; //intercept, unique to ant sp
+  vector[N_ant] beta1; //slope, unique to ant sp
+  vector[N_Plot] u; //subject intercepts
+  vector[N_Year] w; //item intercepts
+  real < lower = 0 > sigma; // Error SD
+  real < lower = 0 > sigma_u; // plot SD
+  real < lower = 0 > sigma_w; // year SD
+}
+
+transformed parameters{
+  vector[N] mu; //linear predictor for the mean
+  for(i in 1:N){
+    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i] + u[plot[i]] + w[year[i]];
+  }
+}
+model {
+ // Priors
+ `u ~ normal(0, sigma_u); // plot random effects
+  w ~ normal(0, sigma_w); // year random effects
+  beta0 ~ normal(0,100); // intercept distribution
+  beta1 ~ normal(0,100); // slope distribution
+ // Model
+ for(i in 1:N){
+  y[i] ~ bernoulli_logit(mu[i]);
+ }
+}
+",
+
+"repro_lm1.stan")
+
+## Check that you wrote a file
+stanc("repro_ant_mix_lm1.stan")
+## Save the file path
+stanny <- "repro_ant_mix_lm1.stan"
+## Now we fit the model
+fit_repro_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_repro_mix_ant
+
+posterior_repro_mix_ant <- as.data.frame(fit_repro_mix_ant)
+print(colnames(posterior_repro_mix_ant))
+
+lm_alpha_other <- summary(model3)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(model3)$coeff[3]
+lm_alpha_liom <- summary(model3)$coeff[4]
+lm_alpha_vacant <- summary(model3)$coeff[5]
+lm_beta_other <- summary(model3)$coeff[2]  # the slope
+lm_beta_crem <- summary(model3)$coeff[6]
+lm_beta_liom <- summary(model3)$coeff[7]
+lm_beta_vacant <- summary(model3)$coeff[8]
+
+par(mfrow = c(2,4))
+#Plot Intercepts
+plot(density(posterior_repro_mix_ant$`beta0[1]`), main = "Intercept Other")
+abline(v = lm_alpha_other, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta0[2]`), main = "Intercept Crem")
+abline(v = lm_alpha_crem, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta0[3]`), main = "Intercept Liom")
+abline(v = lm_alpha_liom, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta0[4]`), main = "Intercept Vacant")
+abline(v = lm_alpha_vacant, col = 4, lty = 2)
+#Plot Slopes
+plot(density(posterior_repro_mix_ant$`beta1[1]`), main = "Intercept Other")
+abline(v = lm_beta_other, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta1[2]`), main = "Intercept Crem")
+abline(v = lm_beta_crem, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta1[3]`), main = "Intercept Liom")
+abline(v = lm_beta_liom, col = 4, lty = 2)
+plot(density(posterior_repro_mix_ant$`beta1[4]`), main = "Intercept Vacant")
+abline(v = lm_beta_vacant, col = 4, lty = 2)
+```

### 2021-02-16 Stan/Stan Base Models.Rmd
Stan Models (With and Without Ant Interaction)

@@
 }
+data$Year_t <- as.factor(data$Year_t)
+data$year <- as.integer(data$Year_t)
+data$Plot <- as.factor(data$Plot)
+data$plot <- as.integer(data$Plot)
 

### 2021-03-02 Stan/Stan Mix Models.Rmd
Ant Vital Rate Models (Stan)

@@
 #### Growth
-```{r Growth Model}
+```{r Growth Mix Ant Model}
 vol = data$volume_t
@@
 ## Save the file path
-#stanny <- "grow_ant_mix_lm1.stan"
+stanny <- "grow_ant_mix_lm1.stan"
 ## Now we fit the model
-fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
 fit_grow_mix_ant
@@
 posterior_grow_mix_ant <- as.data.frame(fit_grow_mix_ant)
-print(colnames(posterior_grow_mix_ant))
+#print(colnames(posterior_grow_mix_ant))
 lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
@@
 abline(v = lm_sigma, col = 4, lty = 2)
+
 ```
@@
 
-####Abortion Model
+####Viability Model
 
-```{r abortion model}
-y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
-vol = flower$volume_t
+```{r Viability model}
+good <- flower$Goodbuds_t
+abort <- flower$ABFlowerbuds_t
+tot <- flower$TotFlowerbuds_t
+vol = log(flower$volume_t)
 N = nrow(flower)
@@
 year <- flower$year
-stan_data <- list(N = N, ## number of observations
-                  vol = vol, ## predictors volume
-                  y = y, ## response volume next year
-                  ant = ant,## predictors ants
-                  N_ant = N_ant, ## number of ant states
+stan_data <- list(N = N, ## number of observations = trials
+                  N_ant = N_ant,
+                  ant = ant,
+                  good = good, ## good buds = successes
+                  abort = abort, ## aborted buds data
+                  tot = tot, ## number of trials
+                  vol = vol,
                   N_Year = N_Year, ## number of years
@@
 write("// Stan model for simple total flowers regression
-
 data {
-  int <lower = 1> N; // number of observations
+ int < lower = 1 > N; // Sample size
+ int <lower = 0> good[N]; // Number of viable seeds
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+ int<lower=0> tot[N]; // number of trials
+ int<lower=0> abort[N];
  vector[N] vol;	//size_t
-  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
@@
 parameters {
-  vector[N_ant] beta0; //intercept, unique to ant sp
-  vector[N_ant] beta1; //slope, unique to ant sp
+  real <lower = 0, upper = 1> theta[N]; // probability of viability for each bud
+  //
+  real<lower=0,upper=1> v0;
   vector[N_Plot] u; //subject intercepts
@@
 transformed parameters{
-  vector[N] mu; //linear predictor for the mean
+  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
+  // Prediction for seed viability
   for(i in 1:N){
-    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
+    predV[i] = v0 * pow(theta[ant[i]],good[i]) * (1-theta[ant[i]])^abort[i] + u[plot[i]] + w[year[i]];
   }
@@
  // Priors
+ v0  ~ beta(10,1);  // intercept viability model
   u ~ normal(0, sigma_u); // plot random effects
   w ~ normal(0, sigma_w); // year random effects
-  beta0 ~ normal(0,100); // intercept distribution
-  beta1 ~ normal(0,100); // slope distribution
- // Model
- for(i in 1:N){
-  y[i] ~ bernoulli_logit(mu[i]);
+  theta ~ normal(0,100); // intercept distribution
+
+  good ~ binomial(tot, predV);
 }
-}
-",
+", "viab_mix_ant.stan")
+## Check that you wrote a file
+stanc("viab_mix_ant.stan")
+## Save the file path
+stanny <- "viab_mix_ant.stan"
+## Now we fit the model
+fit_viab_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
 
-"abort_ant_mix_lm1.stan")
-## Check that you wrote a file
-stanc("abort_ant_mix_lm1.stan")
-## Save the file path
-stanny <- "abort_ant_mix_lm1.stan"
-## Now we fit the model
-fit_abort_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
-fit_abort_mix_ant
+posterior_viab_mix_ant <- as.data.frame(fit_viab_mix_ant)
 
-posterior_abort_mix_ant <- as.data.frame(fit_abort_mix_ant)
-print(colnames(posterior_abort_mix_ant))
+lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]
 
-lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
-lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
-lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
-lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
-lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
-lm_beta_crem <- summary(abort_model_ant)$coeff[6]
-lm_beta_liom <- summary(abort_model_ant)$coeff[7]
-lm_beta_vacant <- summary(abort_model_ant)$coeff[8]
-
-par(mfrow = c(2,4))
-#Plot Intercepts
-plot(density(posterior_abort_mix_ant$`beta0[1]`), main = "Intercept Other")
-abline(v = lm_alpha_other, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta0[2]`), main = "Intercept Crem")
-abline(v = lm_alpha_crem, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta0[3]`), main = "Intercept Liom")
-abline(v = lm_alpha_liom, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta0[4]`), main = "Intercept Vacant")
-abline(v = lm_alpha_vacant, col = 4, lty = 2)
+par(mfrow = c(1,2))
 #Plot Slopes
-plot(density(posterior_abort_mix_ant$`beta1[1]`), main = "Intercept Other")
-abline(v = lm_beta_other, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta1[2]`), main = "Intercept Crem")
-abline(v = lm_beta_crem, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta1[3]`), main = "Intercept Liom")
-abline(v = lm_beta_liom, col = 4, lty = 2)
-plot(density(posterior_abort_mix_ant$`beta1[4]`), main = "Intercept Vacant")
-abline(v = lm_beta_vacant, col = 4, lty = 2)
+plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim =  c(0,2))
+abline(v = lm_alpha_other, col = "black", lty = 2)
+lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
+abline(v = lm_alpha_crem, col = "red", lty = 2)
+lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
+abline(v = lm_alpha_liom, col = "blue", lty = 2)
+lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
+abline(v = lm_alpha_vacant, col = "pink", lty = 2)
+legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
 ```

### 2021-03-02 Stan/Stan Multinomial Models.Rmd
Ant Vital Rate Models (Stan)

---
title: "Multinomial Models"
author: "Ali Campbell"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
#library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

#### Data

```{r}
## Data Set
cactus$ant_t1_relevel <- relevel(cactus$ant_t1,ref = "vacant")
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1", "ant_t1_relevel")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
data$ant1 <- as.integer(data$ant_t1_relevel)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
data$Year_t <- as.factor(data$Year_t)
data$year <- as.integer(data$Year_t)
data$Plot <- as.factor(data$Plot)
data$plot <- as.integer(data$Plot)

## Flower Data Set
flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1","Year_t","Plot")]
flower <- na.omit(flower)
flower$ant <- as.integer(flower$ant_t)
flower$Year_t <- as.factor(flower$Year_t)
flower$year <- as.integer(flower$Year_t)
flower$Plot <- as.factor(flower$Plot)
flower$plot <- as.integer(flower$Plot)
```

#### 1
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]]  + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial1.stan")
#stanc("multinomial1.stan")
## Save the file path
stanny <- "multinomial1.stan"
## Now we fit the model
fit_multi1 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 4, cores = 2, thin = 1)
#class(fit)
posterior_multi1 <- as.data.frame(fit_multi1)
```

#### 2 Ants no interaction
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1 * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial2.stan")
stanc("multinomial2.stan")
## Save the file path
stanny <- "multinomial2.stan"
## Now we fit the model
fit_multi2 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi2 <- as.data.frame(fit_multi2)
```

#### 3

```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial3.stan")
#stanc("multinomial3.stan")
## Save the file path
stanny <- "multinomial3.stan"
## Now we fit the model
fit_multi3 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi3 <- as.data.frame(fit_multi3)
```

### 2021-03-02 Stan/Visualize Stan Mixed.Rmd
Ant Vital Rate Models (Stan)

---
title: "Visualize Stan Mixed Models"
author: "Ali Campbell"
date: "2/25/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
#library(rstansim)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
#library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
cactus$ant_t1_relevel <- relevel(cactus$ant_t1,ref = "vacant")
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1", "ant_t1_relevel")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
data$ant1 <- as.integer(data$ant_t1_relevel)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
data$Year_t <- as.factor(data$Year_t)
data$year <- as.integer(data$Year_t)
data$Plot <- as.factor(data$Plot)
data$plot <- as.integer(data$Plot)

## Flower Data Set
flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1","Year_t","Plot")]
flower <- na.omit(flower)
flower$ant <- as.integer(flower$ant_t)
flower$Year_t <- as.factor(flower$Year_t)
flower$year <- as.integer(flower$Year_t)
flower$Plot <- as.factor(flower$Plot)
flower$plot <- as.integer(flower$Plot)
```

#### Growth
```{r Growth Mixed Ant Model}
vol = log(data$volume_t)
y = log(data$volume_t1)
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
) 
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression
data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  vector[N] y; // size_t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ normal(mu[i], sigma);
  }
}
generated quantities{
  real  y_rep[N] = normal_rng(mu, sigma);
  real  mean_y_rep = mean(to_vector(y_rep));
  real  sd_y_rep = sd(to_vector(y_rep));
}
",
"grow_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("grow_ant_mix_lm1.stan")
## Save the file path
stanny <- "grow_ant_mix_lm1.stan"
## Now we fit the model
fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#fit_grow_mix_ant
#class(fit)
posterior_grow_mix_ant <- as.data.frame(fit_grow_mix_ant)
```

```{r Growth Visualizations}
lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

par(mfrow = c(1,2))
#Plot Intercepts
plot(density(posterior_grow_mix_ant$`beta0[1]`), main = "Growth Intercept Distributions", col = "black", lwd = 2, xlim = c(-4,2), ylim =  c(0,10))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_grow_mix_ant$`beta0[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_grow_mix_ant$`beta0[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_grow_mix_ant$`beta0[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
#Plot Slopes
plot(density(posterior_grow_mix_ant$`beta1[1]`), main = "Growth Slope Distributions", col = "black", lwd = 2, xlim = c(-4,5), ylim = c(0,1.4))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_grow_mix_ant$`beta1[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_grow_mix_ant$`beta1[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_grow_mix_ant$`beta1[4]`), col = "pink", lwd = 2 )
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

```{r}
bayesplot::mcmc_trace(As.mcmc.list(fit_grow_mix_ant, pars=c("beta0", "beta1")))
```

#### Survival
```{r Survival Ant Mixed Model}
vol = log(data$volume_t)
y = data$Survival_t1
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple ant survival regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
//Priors
 u ~ normal(0, sigma_u); // plot random effects
 w ~ normal(0, sigma_w); // year random effects
 beta0 ~ normal(0,100); // intercept distribution
 beta1 ~ normal(0,100); // slope distribution
 //Model
 for(i in 1:N){
 y[i] ~ bernoulli_logit(mu[i]);
 }
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",

"surv_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("surv_ant_mix_lm1.stan")
## Save the file path
stanny <- "surv_ant_mix_lm1.stan"
## Now we fit the model
fit_surv_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

#class(fit)
posterior_surv_mix_ant <- as.data.frame(fit_surv_mix_ant)
```

```{r Survival Visualize}
lm_alpha_other <- summary(surv_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(surv_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(surv_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(surv_model_ant_interact)$coeff[5]
lm_beta_other <- summary(surv_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(surv_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(surv_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(surv_model_ant_interact)$coeff[8]
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error

par(mfrow = c(1,2))
#Plot Intercepts
plot(density(posterior_surv_mix_ant$`beta0[1]`), main = "Survival Intercept Distributions", col = "black", lwd = 2, xlim = c(-2,3), ylim =  c(0,7))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_surv_mix_ant$`beta0[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_surv_mix_ant$`beta0[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_surv_mix_ant$`beta0[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
#Plot Slopes
plot(density(posterior_surv_mix_ant$`beta1[1]`), main = "Survival Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim = c(0,1.4))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_surv_mix_ant$`beta1[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_surv_mix_ant$`beta1[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_surv_mix_ant$`beta1[4]`), col = "pink", lwd = 2 )
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```
The mean here is too large and the standard deviation is also too large

```{r}
bayesplot::mcmc_trace(As.mcmc.list(fit_surv_mix_ant, pars=c("beta0", "beta1")))
```

#### Total Flowers
```{r}
y <- flower$TotFlowerbuds_t
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression

data {
 int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ poisson(mu[i]);
  }
}
generated quantities {
  int<lower = 0> y_rep[N] = poisson_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"flower_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("flower_ant_mix_lm1.stan")
## Save the file path
stanny <- "flower_ant_mix_lm1.stan"
## Now we fit the model
fit_flow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_flow_mix_ant <- as.data.frame(fit_flow_mix_ant)
```

```{r Ants Post Visualize}

par(mfrow = c(1,2))
#Plot Intercepts
plot(density(posterior_flow_mix_ant$`beta0[1]`), main = "Total Flowers Intercept Distributions", col = "black", lwd = 2, xlim = c(-2,4), ylim =  c(0,4.5))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_flow_mix_ant$`beta0[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_flow_mix_ant$`beta0[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_flow_mix_ant$`beta0[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topright", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
#Plot Slopes
plot(density(posterior_flow_mix_ant$`beta1[1]`), main = "Total Flowers Slope Distributions", col = "black", lwd = 2, xlim = c(-1,4.5), ylim = c(0,1.7))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_flow_mix_ant$`beta1[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_flow_mix_ant$`beta1[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_flow_mix_ant$`beta1[4]`), col = "pink", lwd = 2 )
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topright", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

```{r Total Visualize}
par(mfrow = c(1,2))
plot(density(posterior_flow_mix_ant$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(flower$TotFlowerbuds_t), col = "red", lty = 2)
plot(density(posterior_flow_mix_ant$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(flower$TotFlowerbuds_t), col = "red", lty = 2)

flow_mix_ant_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot),family = "poisson", warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Total Flowerbuds Model Simulated Data")
bayesplot::ppc_dens_overlay(y = flow_mix_ant_mod$y,
                 yrep = posterior_predict(flow_mix_ant_mod, draws = 50)) + plot_title

bayesplot::mcmc_trace(As.mcmc.list(fit_flow_mix_ant, pars=c("beta0", "beta1")))
```

#### Repro
```{r Repro Ant Mixed Model}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol1 = vol1, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol1;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Priors
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
 // Model
 for(i in 1:N){
  y[i] ~ bernoulli_logit(mu[i]);
 }
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",

"repro_ant_mix_lm1.stan")

## Check that you wrote a file
#stanc("repro_ant_mix_lm1.stan")
## Save the file path
stanny <- "repro_ant_mix_lm1.stan"
## Now we fit the model
fit_repro_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_repro_mix_ant <- as.data.frame(fit_repro_mix_ant)
```

```{r Ants Post Visualize}

par(mfrow = c(1,2))
#Plot Intercepts
plot(density(posterior_repro_mix_ant$`beta0[1]`), main = "Reproductive State Intercept Distributions", col = "black", lwd = 2, xlim = c(-6,4), ylim =  c(0,25))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_repro_mix_ant$`beta0[2]`),  col = "red", lwd = 2)
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_repro_mix_ant$`beta0[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_repro_mix_ant$`beta0[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
#Plot Slopes
plot(density(posterior_repro_mix_ant$`beta1[1]`), main = "Reproductive State Slope Distributions", col = "black", lwd = 2, xlim = c(-6,4.5), ylim = c(0,25 ))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_repro_mix_ant$`beta1[2]`),  col = "red", lwd = 2)
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_repro_mix_ant$`beta1[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_repro_mix_ant$`beta1[4]`), col = "pink", lwd = 2 )
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

```{r Reproduction Visualize}
par(mfrow = c(1,2))
plot(density(posterior_repro_mix_ant$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(flower$repro_state_t1), col = "red", lty = 2)
plot(density(posterior_repro_mix_ant$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(flower$repro_state_t1), col = "red", lty = 2)

repro_mix_ant_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot),family = "binomial", warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Reproductive State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = repro_mix_ant_mod$y,
                 yrep = posterior_predict(repro_mix_ant_mod, draws = 50)) + plot_title

bayesplot::mcmc_trace(As.mcmc.list(fit_repro_mix_ant, pars=c("beta0", "beta1")))
```

####Viability Model

```{r Viability model}
good <- flower$Goodbuds_t
abort <- flower$ABFlowerbuds_t
tot <- flower$TotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations = trials
                  N_ant = N_ant,
                  ant = ant,
                  good = good, ## good buds = successes
                  abort = abort, ## aborted buds data
                  tot = tot, ## number of trials
                  vol = vol,
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression
data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> good[N]; // Number of viable seeds
 int <lower = 1> N_ant; // number of ant states
 int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
 int<lower=0> tot[N]; // number of trials
 int<lower=0> abort[N];
 vector[N] vol;	//size_t
 int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(good));
  real<lower = 0> sd_y = sd(to_vector(good));
}
parameters {
  real theta[N]; // probability of viability for each bud
  //
  real<lower=0,upper=1> v0;
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
  // Prediction for seed viability
  for(i in 1:N){
    logit(predV[i]) = theta[ant[i]] + u[ant[i]] + w[ant[i]];
  }
}
model {
 // Priors
  v0  ~ beta(10,1);  // intercept viability model
  theta ~ normal(0,100); // intercept distribution
  //Model Statements
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects

  good ~ binomial(tot, predV);
}
generated quantities {
  int<lower = 0> y_rep[N] = binomial_rng(tot,predV);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
", "viab_mix_ant.stan")
## Check that you wrote a file
stanc("viab_mix_ant.stan")
## Save the file path
stanny <- "viab_mix_ant.stan"
## Now we fit the model
fit_viab_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_viab_mix_ant <- as.data.frame(fit_viab_mix_ant)
```

```{r Visualize}
par(mfrow = c(1,2))
plot(density(posterior_viab_mix_ant$mean_y_rep), xlim = c(2.19,10))
abline(v = mean(flower$Goodbuds_t), col = "red", lty = 2)
plot(density(posterior_viab_mix_ant$sd_y_rep), xlim = c(3.3,8.7))
abline(v = sd(flower$Goodbuds_t), col = "red", lty = 2)

viab_mix_ant_mod <- stan_glmer(cbind(flower$Goodbuds_t,flower$ABFlowerbuds_t) ~ flower$ant + (1|flower$year) + (1|flower$plot),family = "binomial", warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Viability Model Simulated Data")
bayesplot::ppc_dens_overlay(y = viab_mix_ant_mod$y,
                 yrep = posterior_predict(viab_mix_ant_mod, draws = 5)) + plot_title
fit_viab_mix_ant$v
bayesplot::mcmc_trace(As.mcmc.list(fit_viab_mix_ant, pars=c("v0")))
```

```{r Visualize}
lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]

#Plot Slopes
plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,2), ylim =  c(0,4))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

## Multinomial
#### 1
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]]  + u[plot[i]] + w[year[i]];
  }
}
model {
  // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = multinomail_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial1.stan")
#stanc("multinomial1.stan")
## Save the file path
stanny <- "multinomial1.stan"
## Now we fit the model
fit_multi1 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 4, cores = 2, thin = 1)
#class(fit)
posterior_multi1 <- as.data.frame(fit_multi1)
```

```{r 1 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi1$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi1$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi1_mod <- stan_glmer(y ~ 0 + ant + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi1_mod$y,
                 yrep = posterior_predict(multi1_mod, draws = 5)) + plot_title

bayesplot::mcmc_trace(As.mcmc.list(fit_multi1, pars=c("beta0")))
```

```{r 1}
traceplot(fit_multi1, pars = c("beta0", "sigma"))

plot(posterior_multi1$`beta0[1]`, type = "l", col = "black", lwd = 2, ylim = c(-0.8,1.5))
lines(posterior_multi1$`beta0[2]`, type = "l", col = "red", lwd = 2)
lines(posterior_multi1$`beta0[3]`, type = "l", col = "blue", lwd = 2)
lines(posterior_multi1$`beta0[4]`, type = "l", col = "pink", lwd = 2)
```

#### 2 Ants no interaction
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1 * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
  // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = multinomial_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial2.stan")
stanc("multinomial2.stan")
## Save the file path
stanny <- "multinomial2.stan"
## Now we fit the model
fit_multi2 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi2 <- as.data.frame(fit_multi2)
```

```{r 2 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi2$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi2$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi2_mod <- stan_glmer(y ~ 0 + ant + vol + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi2_mod$y,
                 yrep = posterior_predict(multi2_mod, draws = 5)) + plot_title

bayesplot::mcmc_trace(As.mcmc.list(fit_multi2, pars=c("beta0", "beta1")))
```

```{r}
traceplot(fit_multi2, pars = c("beta0", "beta1", "sigma"))

par(mfrow = c(1,2))
plot(posterior_multi2$`beta0[1]`, type = "l", col = "black", lwd = 2, ylim = c(-2.2, 0))
lines(posterior_multi2$`beta0[2]`, type = "l", col = "red", lwd = 2)
lines(posterior_multi2$`beta0[3]`, type = "l", col = "blue", lwd = 2)
lines(posterior_multi2$`beta0[4]`, type = "l", col = "pink", lwd = 2)
plot(posterior_multi2$`beta1`, type = "l", col = "black", lwd = 2, lty = 2)
```

#### 3

```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
  // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = multinomial_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial3.stan")
#stanc("multinomial3.stan")
## Save the file path
stanny <- "multinomial3.stan"
## Now we fit the model
fit_multi3 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi3 <- as.data.frame(fit_multi3)
```

```{r 3 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi3$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi3$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi3_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Interaction Effects Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi3_mod$y,
                 yrep = posterior_predict(multi3_mod, draws = 5)) + plot_title

bayesplot::mcmc_trace(As.mcmc.list(fit_multi3, pars=c("beta0", "beta1")))
```

```{r}
par(mfrow = c(1,2))
plot(posterior_multi3$`beta0[1]`, type = "l", col = "black", lwd = 2, ylim = c(-2,1))
lines(posterior_multi3$`beta0[2]`, type = "l", col = "red", lwd = 2)
lines(posterior_multi3$`beta0[3]`, type = "l", col = "blue", lwd = 2)
lines(posterior_multi3$`beta0[4]`, type = "l", col = "pink", lwd = 2)
plot(posterior_multi3$`beta1[1]`, type = "l", col = "black", lwd = 2, lty = 2, ylim = c(-1.5,1))
lines(posterior_multi3$`beta1[2]`, type = "l", col = "red", lwd = 2, lty = 2)
lines(posterior_multi3$`beta1[3]`, type = "l", col = "blue", lwd = 2, lty = 2)
lines(posterior_multi3$`beta1[4]`, type = "l", col = "pink", lwd = 2, lty = 2)

```

### 2021-03-09 Data Analysis/Data_Setup.Rmd
Data Setup
This takes the original data file and produces the derived data

---
title: "Setup Data"
author: "Ali Campbell"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
require(lattice)
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Cholla-Analysis_files")
```

```{r read_data}
#import the data
cactus <- read.csv("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/cholla_demography_20042019.csv", header = TRUE,stringsAsFactors=T)
#str(cactus) ##<- problem: antcount is a factor
#levels(cactus$Antcount_t);levels(cactus$Antcount_t1)
## drop the offending rows -- this is a 2019 data entry problem
cactus <- cactus[-c(which(cactus$Antcount_t=="2-Jan"),which(cactus$Antcount_t1=="2-Jan")),]
cactus$antcount_t <- as.numeric(as.character(cactus$Antcount_t))
cactus$antcount_t1 <- as.numeric(as.character(cactus$Antcount_t1))
```

##Data Filtering
```{r volume}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

#hist(log(cactus$volume_t))
```

```{r year_t_ants}
## assign ant counts of zero as vacant
cactus$Antsp_t[cactus$antcount_t==0] <- "vacant"
# here are the ordered levels of the current variable
Antsp_t_levels <- levels(cactus$Antsp_t)
# here is how I would like to collapse these into fewer bins -- most will be "other"
ant_t_levels <- rep("other",times=length(Antsp_t_levels))
## crem levels - elements 5,8,9
ant_t_levels[c(5,8,9)] <- "crem"
## liom levels - elements 15,18,19,20,21
ant_t_levels[c(15,18,19,20,21)] <- "liom"
## forelius levels - 12,13
#ant_t_levels[c(12,13)] <- "for"
## vacant - 31
ant_t_levels[31] <- "vacant"
## create new variable merging levels as above
cactus$ant_t <- factor(cactus$Antsp_t,levels=Antsp_t_levels,labels=ant_t_levels)
summary(cactus$ant_t) ## there are too many "others". check these out.
filter(cactus,ant_t=="other")
## first, there are no ant data from 2007:
cactus$Antcount_t[cactus$Year_t==2007]
#so give these observations NA for ant status:
cactus$ant_t[cactus$Year_t==2007] <- NA
# also, if a plant was new in year t+1 it's year t ant status should be NA
cactus$ant_t[cactus$Newplant==1] <- NA
# plots 7 and 8 were added in 2011, so their year_t==2010 ant status should be NA
cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==7] <- NA
cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==8] <- NA
# check to see how many others this leaves -- still a lot
summary(cactus$ant_t)
## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
cactus$ant_t[cactus$Antsp_t==""]<-NA
## inspect the "others" visually
cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
## I am satisfied that everything assigned other is correctly assigned

```

```{r year_t1_ants}
## assign ant counts of zero as vacant
cactus$Antsp_t1[cactus$antcount_t1==0] <- "vacant"
## repeat for t1 -- these indices are different
Antsp_t1_levels <- levels(cactus$Antsp_t1)
# here is how I would like to collapse these into fewer bins -- most will be "other"
ant_t1_levels <- rep("other",times=length(Antsp_t1_levels))
## crem levels 
ant_t1_levels[c(8,9,10)] <- "crem"
## liom levels 
ant_t1_levels[c(19,20,21,22)] <- "liom"
## forelius levels 
#ant_t1_levels[c(13,14,15)] <- "for"
## vacant 
ant_t1_levels[32] <- "vacant"
## create new variable merging levels as above
cactus$ant_t1 <- factor(cactus$Antsp_t1,levels=Antsp_t1_levels,labels=ant_t1_levels)
summary(cactus$ant_t1) ## there are too many "others" -- same deal as above
## first, there are no ant data from 2007 so give these observations NA for ant status:
cactus$ant_t1[cactus$Year_t1==2007] <- NA
## if we did not get size in t1 assume we did not get ant
cactus$ant_t1[is.na(cactus$Height_t1)] <- NA
# check to see how many others this leaves -- still a lot
summary(cactus$ant_t1)
## there are some Antsp_t1=="" that show up as "other"
cactus %>% filter(Antsp_t1=="" & ant_t1=="other")
## not sure why this happens but apparently there are some plants for which we have ant counts but no ant sp. Could be a mix in comments. For now make these NA.
cactus$ant_t1[cactus$Antsp_t1==""] <- NA
## inspect the "others" visually
cactus %>% filter(ant_t1=="other") %>% select(Antsp_t1,ant_t1)
## I am satisfied that everything assigned other in t1 is correctly assigned

```

```{r}
cactus$repro_state_t <- NA
for(i in 1:8187){
  if(is.na(cactus$TotFlowerbuds_t[i]) == FALSE & cactus$TotFlowerbuds_t[i] > 0) {
    cactus$repro_state_t[i] <- 1
  }
  if(is.na(cactus$Goodbuds_t[i]) == FALSE & cactus$Goodbuds_t[i] > 0){
    cactus$repro_state_t[i] <- 1
  }
  if(is.na(cactus$Goodbuds_t[i]) == FALSE & is.na(cactus$TotFlowerbuds_t[i]) == FALSE & cactus$Goodbuds_t[i] < 1 & cactus$TotFlowerbuds_t[i] < 1){
  cactus$repro_state_t[i] <- 0
  }
  if(is.na(cactus$TotFlowerbuds_t1[i]) == FALSE & cactus$TotFlowerbuds_t1[i] > 0) {
    cactus$repro_state_t1[i] <- 1
  }
  if(is.na(cactus$Goodbuds_t1[i]) == FALSE & cactus$Goodbuds_t1[i] > 0){
    cactus$repro_state_t1[i] <- 1
  }
  if(is.na(cactus$Goodbuds_t1[i]) == FALSE & is.na(cactus$TotFlowerbuds_t1[i]) == FALSE & cactus$Goodbuds_t1[i] < 1 & cactus$TotFlowerbuds_t1[i] < 1){
  cactus$repro_state_t1[i] <- 0
}
}
```

### 2021-03-22 Stan/Simulate_Data.Rmd

---
title: "Simulate Data"
author: "Ali Campbell"
date: "2/17/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
#library(rstansim)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
#library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
cactus$ant_t1_relevel <- relevel(cactus$ant_t1,ref = "vacant")
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1", "ant_t1_relevel")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
data$ant1 <- as.integer(data$ant_t1_relevel)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
data$Year_t <- as.factor(data$Year_t)
data$year <- as.integer(data$Year_t)
data$Plot <- as.factor(data$Plot)
data$plot <- as.integer(data$Plot)

## Flower Data Set
flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1","Year_t","Plot")]
flower <- na.omit(flower)
flower$ant <- as.integer(flower$ant_t)
flower$Year_t <- as.factor(flower$Year_t)
flower$year <- as.integer(flower$Year_t)
flower$Plot <- as.factor(flower$Plot)
flower$plot <- as.integer(flower$Plot)
```

## Simulate for Base Models (no ants)
#### Growth
```{r Growth Model}
vol <- data$volume_t
y <- data$volume_t1
N <- as.integer(nrow(data))
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
#summary(grow_model)
#summary(mod1)
## Now we can extract some important statistics
lm_alpha <- summary(grow_model)$coeff[1]  # the intercept
lm_beta <- summary(grow_model)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model)  # the residual error
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression
data {
 int < lower = 1 > N; // Sample size
 vector[N] vol; // Predictor
 vector[N] y; // Outcome
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
 real alpha; // Intercept
 real beta; // Slope (regression coefficients)
 real < lower = 0 > sigma; // Error SD
}
model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~ normal(alpha + vol * beta , sigma); // The Model
}
generated quantities {
  real  y_rep[N] = normal_rng(alpha + beta * vol, sigma);
  real  mean_y_rep = mean(to_vector(y_rep));
  real  sd_y_rep = sd(to_vector(y_rep));
} // The posterior predictive distribution",
"grow_lm1.stan")
stanc("grow_lm1.stan")
## Save the file path
stanny <- "grow_lm1.stan"
## Now we fit the model
fit_grow <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_grow
#class(fit)
posterior_grow <- as.data.frame(fit_grow)
print(colnames(posterior_grow))
mat_grow <- as.matrix(fit_grow)

par(mfrow = c(1,1))
plot(density(posterior_grow$mean_y_rep), xlim = c(55000, 67000))
abline(v = mean(na.omit(cactus$volume_t1)), col = "red", lty = 2)

```

#### Survival
```{r Survival Model}
vol = data$volume_t
y <- data$Survival_t1
N <- nrow(data)
#import things from the r regression models for comparison
lm_alpha <- summary(surv_model)$coeff[1]  # the intercept
lm_beta <- summary(surv_model)$coeff[2]  # the slope
summary(surv_model)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple survival regression

data {
  int <lower=0> N;
  vector[N] vol;
  int <lower=0,upper=1> y[N];
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  real alpha;
  real beta;
}
model {
 // Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(alpha + beta * vol);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(alpha + beta * vol);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"surv_lm1.stan")
## Check that you wrote a file
stanc("surv_lm1.stan")
## Save the file path
stanny <- "surv_lm1.stan"
## Now we fit the model
fit_surv <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_surv

posterior_surv <- as.data.frame(fit_surv)
print(colnames(posterior_surv))

par(mfrow = c(1,1))
plot(density(posterior_surv$mean_y_rep), xlim = c(0.934,1.0002))
abline(v = mean(na.omit(cactus$Survival_t1)), col = "red", lty = 2)
```

#### Total Flowers
```{r Total Flowers Model}
y <- flower$TotFlowerbuds_t
N <- nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y ## response survival next year
)

write("// Stan model for simple total flowers regression

data {
 int < lower = 1 > N; // Sample size
 int y[N]; // Outcome
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  real<lower = 0> lambda; //intercept, unique to ant
}
model {
// Priors
 lambda ~ exponential(0.2);
 // Model
 y ~ poisson(lambda);
}
generated quantities {
  int<lower = 0> y_rep[N] = poisson_rng(rep_array(lambda, N));
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
// The posterior predictive distribution",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit_flow <- stan(file = stanny, data = stan_data, warmup = 50, iter = 100, chains = 1, cores = 2, thin = 1)
fit_flow

posterior_flow <- as.data.frame(fit_flow)
print(colnames(posterior_flow))
lm_lambda <- summary(flower_model)$coeff[2]

plot(density(posterior_flow$mean_y_rep), xlim = c(4,12))
abline(v = mean(na.omit(cactus$TotFlowerbuds_t1)), col = "red", lty = 2)

```

#### Reproduction
```{r Reproductive Model}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol1 = vol1
)

write("// Stan model for simple total flowers regression

data {
 int < lower = 1 > N; // Sample size
 int <lower = 0, upper = 1> y[N]; // Outcome
 vector[N] vol1;	//size_t
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  real beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(beta0 + beta1 * vol1);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(beta0 + beta1 * vol1);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"repro_lm1.stan")

## Check that you wrote a file
stanc("repro_lm1.stan")
## Save the file path
stanny <- "repro_lm1.stan"
## Now we fit the model
fit_repro <- stan(file = stanny, data = stan_data, warmup = 50, iter = 100, chains = 1, cores = 2, thin = 1)
fit_repro

posterior_repro <- as.data.frame(fit_repro)
print(colnames(posterior_repro))

par(mfrow = c(1,1))
plot(density(posterior_repro$mean_y_rep))
abline(v = mean(na.omit(cactus$repro_state_t1)), col = "red", lty = 2)
```

## Simulate for Base Models (ants)
#### Growth
```{r Growth Ant Model}
vol = data$volume_t
y = data$volume_t1
N = nrow(data)
N_ant = 4
ant = data$ant
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant
) 
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  vector[N] y; // size_t1
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}

model {
  y ~ normal(mu, sigma); // likelihood
}
generated quantities{
  real  y_rep[N] = normal_rng(mu, sigma);
  real  mean_y_rep = mean(to_vector(y_rep));
  real  sd_y_rep = sd(to_vector(y_rep));
}
",

"grow_ant_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
fit_grow_ant
#class(fit)
posterior_grow_ant <- as.data.frame(fit_grow_ant)
print(colnames(posterior_grow_ant))

par(mfrow = c(1,1))
plot(density(posterior_grow_ant$mean_y_rep), xlim = c(53375, 66738))
abline(v = mean(na.omit(cactus$volume_t1)), col = "red", lty = 2)
```

#### Survival
```{r Survival Model Ant}
vol = data$volume_t
y = data$Survival_t1
N = nrow(data)
N_ant = 4
ant = data$ant
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response survival next year
                  ant = ant,## predictors ants
                  N_ant = N_ant
)

write("// Stan model for simple ant survival regression
data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}
model {
//Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit_surv_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
fit_surv_ant

#class(fit)
posterior_surv_ant <- as.data.frame(fit_surv_ant)
print(colnames(posterior_surv_ant))

par(mfrow = c(1,1))
plot(density(posterior_surv_ant$mean_y_rep), xlim = c(0.879,0.94))
abline(v = mean(na.omit(cactus$Survival_t1)), col = "red", lty = 2)
```

#### Total Flowers
```{r}
y <- flower$TotFlowerbuds_t
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol,
                  N_ant = N_ant,
                  ant = ant
)

write("// Stan model for simple total flowers regression
data {
 int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0> y[N]; // survival in year t1
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}
model {
 // Model
 y ~ poisson(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = poisson_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"flower_ant_lm1.stan")
## Check that you wrote a file
stanc("flower_ant_lm1.stan")
## Save the file path
stanny <- "flower_ant_lm1.stan"
## Now we fit the model
fit_flow_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
fit_flow_ant

posterior_flow_ant <- as.data.frame(fit_flow_ant)
print(colnames(posterior_flow_ant))

par(mfrow = c(1,1))
plot(density(posterior_flow_ant$mean_y_rep), xlim = c(0, 133450))
abline(v = mean(y), col = "red", lty = 2)
```

#### Reproduction
```{r}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol1 = vol1,
                  N_ant = N_ant,
                  ant = ant
)

write("// Stan model for simple total flowers regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol1;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i];
  }
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"repro_ant_lm1.stan")

## Check that you wrote a file
stanc("repro_ant_lm1.stan")
## Save the file path
stanny <- "repro_ant_lm1.stan"
## Now we fit the model
fit_repro_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
fit_repro_ant

posterior_repro_ant <- as.data.frame(fit_repro_ant)
print(colnames(posterior_repro_ant))

par(mfrow = c(1,1))
plot(density(posterior_repro_ant$mean_y_rep))
abline(v = mean(y), col = "red", lty = 2)
```

## Simulate for Mixed Models 
#### Growth
```{r Growth Mixed Ant Model}
vol = log(data$volume_t)
y = log(data$volume_t1)
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
) 
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression
data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  vector[N] y; // size_t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ normal(mu[i], sigma);
  }
}
generated quantities{
  real  y_rep[N] = normal_rng(mu, sigma);
  real  mean_y_rep = mean(to_vector(y_rep));
  real  sd_y_rep = sd(to_vector(y_rep));
}
",
"grow_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("grow_ant_mix_lm1.stan")
## Save the file path
stanny <- "grow_ant_mix_lm1.stan"
## Now we fit the model
fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
#fit_grow_mix_ant
#class(fit)
posterior_grow_mix_ant <- as.data.frame(fit_grow_mix_ant)
mat_grow_mix_ant <- as.matrix(fit_grow_mix_ant)
#print(colnames(posterior_grow_mix_ant))
```

```{r Growth Visualizations}
par(mfrow = c(1,2))
plot(density(posterior_grow_mix_ant$mean_y_rep),
     main = "Simulated Mean and Data Mean")
abline(v = mean(y), col = "red", lty = 2)
plot(density(posterior_grow_mix_ant$sd_y_rep),
     main = "Simulated SD and Data SD")
abline(v = sd(y), col = "red", lty = 2)

grow_mix_ant_mod <- stan_glmer(y ~ vol*ant + (1|year) + (1|plot),warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Growth Model Simulated Data")
bayesplot::ppc_dens_overlay(y = grow_mix_ant_mod$y,
                 yrep = posterior_predict(grow_mix_ant_mod, draws = 50)) + plot_title
```

#### Survival
```{r Survival Ant Mixed Model}
vol = log(data$volume_t)
y = data$Survival_t1
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple ant survival regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
//Priors
 u ~ normal(0, sigma_u); // plot random effects
 w ~ normal(0, sigma_w); // year random effects
 beta0 ~ normal(0,100); // intercept distribution
 beta1 ~ normal(0,100); // slope distribution
 //Model
 for(i in 1:N){
 y[i] ~ bernoulli_logit(mu[i]);
 }
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",

"surv_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("surv_ant_mix_lm1.stan")
## Save the file path
stanny <- "surv_ant_mix_lm1.stan"
## Now we fit the model
fit_surv_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
fit_surv_mix_ant

#class(fit)
posterior_surv_mix_ant <- as.data.frame(fit_surv_mix_ant)
#print(colnames(posterior_surv_mix_ant))
```

```{r Survival Visualize}
par(mfrow = c(1,2))
plot(density(posterior_surv_mix_ant$mean_y_rep), xlim = c(0.9639, 0.9997))
abline(v = mean(data$Survival_t1), col = "red", lty = 2)
plot(density(posterior_surv_mix_ant$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$Survival_t1), col = "red", lty = 2)

surv_mix_ant_mod <- stan_glmer(y ~ vol * ant + (1|plot) + (1|year),warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Survival Model Simulated Data")
bayesplot::ppc_dens_overlay(y = surv_mix_ant_mod$y,
                 yrep = posterior_predict(surv_mix_ant_mod, draws = 50)) + plot_title
```
The mean here is too large and the standard deviation is also too large

#### Total Flowers
```{r}
y <- flower$TotFlowerbuds_t
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression

data {
 int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ poisson(mu[i]);
  }
}
generated quantities {
  int<lower = 0> y_rep[N] = poisson_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
} 
",

"flower_ant_mix_lm1.stan")
## Check that you wrote a file
#stanc("flower_ant_mix_lm1.stan")
## Save the file path
stanny <- "flower_ant_mix_lm1.stan"
## Now we fit the model
fit_flow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)

posterior_flow_mix_ant <- as.data.frame(fit_flow_mix_ant)
```

```{r Total Visualize}
par(mfrow = c(1,2))
plot(density(posterior_flow_mix_ant$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(flower$TotFlowerbuds_t), col = "red", lty = 2)
plot(density(posterior_flow_mix_ant$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(flower$TotFlowerbuds_t), col = "red", lty = 2)

flow_mix_ant_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot),family = "poisson", warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Total Flowerbuds Model Simulated Data")
bayesplot::ppc_dens_overlay(y = flow_mix_ant_mod$y,
                 yrep = posterior_predict(flow_mix_ant_mod, draws = 50)) + plot_title
```

#### Repro
```{r Repro Ant Mixed Model}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol1 = vol1, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol1;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Priors
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
 // Model
 for(i in 1:N){
  y[i] ~ bernoulli_logit(mu[i]);
 }
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",

"repro_ant_mix_lm1.stan")

## Check that you wrote a file
stanc("repro_ant_mix_lm1.stan")
## Save the file path
stanny <- "repro_ant_mix_lm1.stan"
## Now we fit the model
fit_repro_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_repro_mix_ant <- as.data.frame(fit_repro_mix_ant)
```

```{r Reproduction Visualize}
par(mfrow = c(1,2))
plot(density(posterior_repro_mix_ant$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(flower$repro_state_t1), col = "red", lty = 2)
plot(density(posterior_repro_mix_ant$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(flower$repro_state_t1), col = "red", lty = 2)

repro_mix_ant_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot),family = "binomial", warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Reproductive State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = repro_mix_ant_mod$y,
                 yrep = posterior_predict(repro_mix_ant_mod, draws = 50)) + plot_title
```

####Viability Model

```{r Viability model}
good <- flower$Goodbuds_t
abort <- flower$ABFlowerbuds_t
tot <- flower$TotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations = trials
                  N_ant = N_ant,
                  ant = ant,
                  good = good, ## good buds = successes
                  abort = abort, ## aborted buds data
                  tot = tot, ## number of trials
                  vol = vol,
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression
data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> good[N]; // Number of viable seeds
 int <lower = 1> N_ant; // number of ant states
 int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
 int<lower=0> tot[N]; // number of trials
 int<lower=0> abort[N];
 vector[N] vol;	//size_t
 int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(good));
  real<lower = 0> sd_y = sd(to_vector(good));
}
parameters {
  real <lower = 0, upper = 1> theta[N]; // probability of viability for each bud
  //
  real<lower=0,upper=1> v0;
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
  // Prediction for seed viability
  for(i in 1:N){
    predV[i] = v0 * pow(theta[ant[i]],good[i]) * (1-theta[ant[i]])^abort[i];
  }
}
model {
 // Priors
 v0  ~ beta(10,1);  // intercept viability model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  theta ~ normal(0,100); // intercept distribution

  good ~ binomial(tot, predV);
}
generated quantities {
  int<lower = 0> y_rep[N] = binomial_rng(tot,predV);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
", "viab_mix_ant.stan")
## Check that you wrote a file
stanc("viab_mix_ant.stan")
## Save the file path
stanny <- "viab_mix_ant.stan"
## Now we fit the model
fit_viab_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_viab_mix_ant <- as.data.frame(fit_viab_mix_ant)
```

```{r Viability Visualize}
par(mfrow = c(1,2))
plot(density(posterior_viab_mix_ant$mean_y_rep), xlim = c(2.19,10))
abline(v = mean(flower$Goodbuds_t), col = "red", lty = 2)
plot(density(posterior_viab_mix_ant$sd_y_rep), xlim = c(3.3,8.7))
abline(v = sd(flower$Goodbuds_t), col = "red", lty = 2)

viab_mix_ant_mod <- stan_glmer(cbind(flower$Goodbuds_t,flower$ABFlowerbuds_t) ~ flower$ant + (1|flower$year) + (1|flower$plot),family = "binomial", warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Viability Model Simulated Data")
bayesplot::ppc_dens_overlay(y = viab_mix_ant_mod$y,
                 yrep = posterior_predict(viab_mix_ant_mod, draws = 5)) + plot_title
```

```{r}
lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]

par(mfrow = c(1,2))
#Plot Slopes
plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim =  c(0,2))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

## Multinomial
#### 1
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]]  + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial1.stan")
#stanc("multinomial1.stan")
## Save the file path
stanny <- "multinomial1.stan"
## Now we fit the model
fit_multi1 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi1 <- as.data.frame(fit_multi1)
```

```{r 1 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi1$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi1$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi1_mod <- stan_glmer(y ~ 0 + ant + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi1_mod$y,
                 yrep = posterior_predict(multi1_mod, draws = 5)) + plot_title
```

#### 2
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1 * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial2.stan")
stanc("multinomial2.stan")
## Save the file path
stanny <- "multinomial2.stan"
## Now we fit the model
fit_multi2 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi2 <- as.data.frame(fit_multi2)
```

```{r 2 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi2$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi2$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi2_mod <- stan_glmer(y ~ 0 + ant + vol + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Effects Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi2_mod$y,
                 yrep = posterior_predict(multi2_mod, draws = 5)) + plot_title
```

```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = bernoulli_logit_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial3.stan")
#stanc("multinomial3.stan")
## Save the file path
stanny <- "multinomial3.stan"
## Now we fit the model
fit_multi3 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
#class(fit)
posterior_multi3 <- as.data.frame(fit_multi3)
```

```{r 3 Visualize}
par(mfrow = c(1,2))
plot(density(posterior_multi3$mean_y_rep), xlim = c(10.7, 73310))
abline(v = mean(data$ant1), col = "red", lty = 2)
plot(density(posterior_multi3$sd_y_rep), xlim = c(0,0.190))
abline(v = sd(data$ant1), col = "red", lty = 2)

multi3_mod <- stan_glmer(y ~ 0 + ant*vol + (1|year) + (1|plot), warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1 )
plot_title <- ggtitle("Mixed Ant Interaction Effects Ant T1 State Model Simulated Data")
bayesplot::ppc_dens_overlay(y = multi3_mod$y,
                 yrep = posterior_predict(multi3_mod, draws = 5)) + plot_title
```

```{r Viability model}
y_good <- flower$Goodbuds_t
y_abort <- flower$ABFlowerbuds_t
y_tot <- flower$TotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations = trials
                  N_ant = N_ant,
                  ant = ant,
                  y_good = y_good, ## good buds = successes
                  y_abort = y_abort, ## aborted buds data
                  y_tot = y_tot, ## number of trials
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  vol = vol
)

write("// Stan model for simple total flowers regression
data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> y_good[N]; // Number of viable seeds
 int <lower = 1> N_ant; // number of ant states
 int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
 int<lower=0> y_tot[N]; // number of trials
 vector[N] vol;	//size_t
 int<lower=1> N_Year; //number of plots
 int<lower=1> N_Plot; //number of years
 int<lower=1, upper=N_Plot> plot[N]; // plot
 int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y_good));
  real<lower = 0> sd_y = sd(to_vector(y_good));
}
parameters {
  vector[N] beta0; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  real<lower=0,upper=1> predV[N];
  // Prediction for seed viability
  for(i in 1:N){
    predV[i] = beta0[ant[i]] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Priors
 u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
 beta0 ~ normal (0,100);
  y_good ~ binomial(y_tot, predV);
}
generated quantities {
  int<lower = 0> y_rep[N] = binomial_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
", "viab_mix_ant.stan")
## Check that you wrote a file
#stanc("viab_ant.stan")
## Save the file path
stanny <- "viab_mix_ant.stan"
## Now we fit the model
fit_viab_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_viab_mix_ant <- as.data.frame(fit_viab_mix_ant)
```

### 2021-03-22 Stan/Stan Base Models.Rmd

@@
 library(shinystan)
-library(bayesplot)
+#library(bayesplot)
 library(devtools)
@@
 ```
-####Abort
-```{r abort model}
-y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
-vol = flower$volume_t
+####Viability
+```{r Viability model}
+y_good <- flower$Goodbuds_t
+y_abort <- flower$AbABFlowerbuds_t
+y_tot <- flower$TTotFlowerbuds_t
+vol = log(flower$volume_t)
 N = nrow(flower)
-stan_data <- list(N = N, ## number of observations
-                  y = y, ## response survival next year
+stan_data <- list(N = N, ## number of observations = trials
+                  y_good = y_good, ## good buds = successes
+                  y_abort = y_abort, ## aborted buds data
+                  y_tot = y_tot, ## number of trials
                   vol = vol
@@
  int < lower = 1 > N; // Sample size
- int <lower = 0, upper = 1> y[N]; // Outcome
+ int <lower = 0> y_good[N]; // Number of viable seeds
+ int<lower=0> y_tot[N]; // number of trials
  vector[N] vol;	//size_t
@@
   real<lower = 0> sigma;
+  real<lower=0,upper=1> v0;
+  real<lower=0> a_v;
+}
+transformed parameters{
+  real<lower=0,upper=1> predV[n_v];
+  for(i in 1:N){
+    predV[i] = v0 * (1 - pow(SR_v[i],a_v) ) + 0.00001;
+  }
 }
@@
  // Model
- y ~ bernoulli_logit(beta0 + beta1 * vol);
+ y ~ binomial_logit(beta0 + beta1 * vol);
 }
@@
 
-"abort_lm1.stan")
+"viab_lm1.stan")
 ## Check that you wrote a file
-stanc("abort_lm1.stan")
+stanc("vian_lm1.stan")
 ## Save the file path
-stanny <- "abort_lm1.stan"
+stanny <- "viab_lm1.stan"
 ## Now we fit the model
-fit_abort <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
-fit_abort
+fit_viab <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
 
-posterior_abort <- as.data.frame(fit_abort)
-print(colnames(posterior_abort))
+posterior_viab <- as.data.frame(fit_viab)
 ```
@@
 "grow_ant_lm1.stan")
+
 ## Check that you wrote a file
@@
 ## Save the file path
-#stanny <- "grow_ant_lm1.stan"
+stanny <- "grow_ant_lm1.stan"
 ## Now we fit the model
-fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
+fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 50, iter = 100, chains = 1, cores = 2, thin = 1)
 fit_grow_ant
@@
 posterior_grow_ant <- as.data.frame(fit_grow_ant)
-print(colnames(posterior_grow_ant))
 lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
@@
 
-####Abortion Model
-
-```{r abortion model}
-y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
-vol = flower$volume_t
+####Viability Model
+```{r Viability model}
+good <- flower$Goodbuds_t
+abort <- flower$ABFlowerbuds_t
+tot <- flower$TotFlowerbuds_t
+vol = log(flower$volume_t)
 N = nrow(flower)
@@
 ant = flower$ant
-stan_data <- list(N = N, ## number of observations
-                  y = y, ## response survival next year
-                  vol = vol,
+stan_data <- list(N = N, ## number of observations = trials
                   N_ant = N_ant,
-                  ant = ant
+                  ant = ant,
+                  good = good, ## good buds = successes
+                  abort = abort, ## aborted buds data
+                  tot = tot, ## number of trials
+                  vol = vol
 )
@@
 write("// Stan model for simple total flowers regression
-
 data {
-  int <lower = 1> N; // number of observations
+ int < lower = 1 > N; // Sample size
+ int <lower = 0> good[N]; // Number of viable seeds
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
+ int<lower=0> tot[N]; // number of trials
+ int<lower=0> abort[N];
  vector[N] vol;	//size_t
-  int <lower = 0, upper = 1> y[N]; // survival in year t1
 }
 parameters {
-  vector[N_ant] beta0; //intercept, unique to ant sp
-  vector[N_ant] beta1; //slope, unique to ant sp
-  real<lower = 0> sigma;
+  real <lower = 0, upper = 1> theta[N]; // probability of viability for each bud
+  //
+  real<lower=0,upper=1> v0;
+  real<lower=0> a_v;
 }
 transformed parameters{
-  vector[N] mu; //linear predictor for the mean
+  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
+  // Prediction for seed viability
   for(i in 1:N){
-    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
+    predV[i] = v0 * pow(theta[ant[i]],good[i]) * (1-theta[ant[i]])^abort[i] ;
   }
@@
  // Priors
- beta0 ~ normal(0,100);
- beta1 ~ normal(0,100); 
- // Model
- y ~ bernoulli_logit(mu);
+ v0  ~ beta(10,1);  // intercept viability model
+  a_v ~ gamma(1,1);  // decay in viability with SR
+
+  good ~ binomial(tot, predV);
 }
-",
+", "viab_ant.stan")
+## Check that you wrote a file
+stanc("viab_ant.stan")
+## Save the file path
+stanny <- "viab_ant.stan"
+## Now we fit the model
+fit_viab <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)
 
-"abort_ant_lm1.stan")
-## Check that you wrote a file
-stanc("abort_ant_lm1.stan")
-## Save the file path
-stanny <- "abort_ant_lm1.stan"
-## Now we fit the model
-fit_abort_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
-fit_abort_ant
+posterior_viab_ant <- as.data.frame(fit_viab)
 
-posterior_abort_ant <- as.data.frame(fit_abort_ant)
-print(colnames(posterior_abort_ant))
+lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
+lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
+lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
+lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
+lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
+lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
+lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
+lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]
 
-lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
-lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
-lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
-lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
-lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
-lm_beta_crem <- summary(abort_model_ant)$coeff[6]
-lm_beta_liom <- summary(abort_model_ant)$coeff[7]
-lm_beta_vacant <- summary(abort_model_ant)$coeff[8]
-
-par(mfrow = c(2,4))
-#Plot Intercepts
-plot(density(posterior_abort_ant$`beta0[1]`), main = "Intercept Other")
-abline(v = lm_alpha_other, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta0[2]`), main = "Intercept Crem")
-abline(v = lm_alpha_crem, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta0[3]`), main = "Intercept Liom")
-abline(v = lm_alpha_liom, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta0[4]`), main = "Intercept Vacant")
-abline(v = lm_alpha_vacant, col = 4, lty = 2)
+par(mfrow = c(1,2))
 #Plot Slopes
-plot(density(posterior_abort_ant$`beta1[1]`), main = "Intercept Other")
-abline(v = lm_beta_other, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta1[2]`), main = "Intercept Crem")
-abline(v = lm_beta_crem, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta1[3]`), main = "Intercept Liom")
-abline(v = lm_beta_liom, col = 4, lty = 2)
-plot(density(posterior_abort_ant$`beta1[4]`), main = "Intercept Vacant")
-abline(v = lm_beta_vacant, col = 4, lty = 2)
+plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim =  c(0,2))
+abline(v = lm_alpha_other, col = "black", lty = 2)
+lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
+abline(v = lm_alpha_crem, col = "red", lty = 2)
+lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
+abline(v = lm_alpha_liom, col = "blue", lty = 2)
+lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
+abline(v = lm_alpha_vacant, col = "pink", lty = 2)
+legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
 ```

### 2021-03-22 Stan/Visualize Stan Mixed.Rmd

@@
 }
-transformed data {
-  real<lower = 0> mean_y = mean(to_vector(y));
-  real<lower = 0> sd_y = sd(to_vector(y));
-}
+
 parameters {
   vector[N_ant] beta0; //intercept, unique to ant sp
-  vector[N_ant] beta1; //slope, unique to ant sp
   vector[N_Plot] u; //subject intercepts
   vector[N_Year] w; //item intercepts
-  real < lower = 0 > sigma; // Error SD
   real < lower = 0 > sigma_u; // plot SD
@@
 generated quantities {
-  int<lower = 0> y_rep[N] = multinomail_rng(mu);
-  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
-  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
+  int<lower = 0> y_rep[N] = multinomial_rng(mu);
 }

### 2022-04-14 Paper/Cholla Paper Draft.Rmd
Updates

---
title: "Cholla Ant Paper"
author: "Ali Campbell"
date: "4/7/2022"
output: pdf_document
bibliography: references.bib
document class: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
tinytex::install_tinytex()
```

#	Abstract
	Understanding interspecific mutualisms is challenging, because they are among the most widespread species interactions with dynamic consequences. One of the most important steps towards understanding lifelong demography impacts of mutualisms is quantifying the costs and benefits partners provide. In this study system Cilindropuntia imbracata (cholla cacti) attract ant partners, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), as well as small populations of other ant species, by producing extrafloral nectar (EFN) and in return are defended from herbivores. The ant partners may be attracted to different qualities of EFN and differ in defensive effectiveness. If the cacti use significant energy to produce high quality nectar they may attract the most effective defender, but will have less energy for reproduction and growth. We ask “Which ant partner, or set of ant partners, is most effective and efficient for the cacti in this system?”. To determine this, we developed an Integral Projection Model (IPM) with Bayesian estimates, and used observational data from a 13 year study to parameterize the model. This lets us examine how interactions with different ant partners changes the fitness of the cacti population. This same framework is applicable to analyze other multispecies mutualisms to investigate partner contributions.

	Our results demonstrate the importance of evaluating mutualisms as dynamic systems. Our modeling found the vital rates of cholla vary depending on the partner present and that no partner is the “best” across the entire lifespan. The “best” partner appears to change across ontogeny. When these cholla cacti are young, they are unable to reproduce, instead all of their energy goes towards growth and EFN. We found that small cacti with the highest growth rates and survival rates are most commonly tended by Crem. ants. This indicates that Crem. is a very effective partner for small cacti. We found that larger cacti have the highest reproductive success when tended by Liom. ants, the most aggressive species observed interacting with these cacti. This aligns with the life state of reproduction, at which the cactus is most vulnerable to herbivory. Together these results show that the optimal partner will likely shift across the life stages of cholla cacti. There are a number of biodiversity hypotheses that may explain these phenomena, including complementarity and sampling effect. As we continue this project, we will investigate the grounds for these hypotheses.

# Introduction

# Bibliography

### 2022-04-14 Cholla_Analysis_Vital_Rates.R

@@ -1,785 +1,1635 @@
 #setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
 #######################################################################################################
 ##
-##                  The purpose of this file is to run each vital rate sub model separately
-##                          and run the vital rate models as one large model 
+##                  The purpose of this file is to run each vital rate sub model separately,
+##                          save the outputs, and check the posterior distributions 
 ##
 #######################################################################################################
 options(mc.cores = parallel::detectCores())
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
 source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Setup_Script.R")
+cactus <- read.csv("cholla_demography_20042019_cleaned.csv", header = TRUE,stringsAsFactors=T)
 
-#######################################################################################################
-#### Growth Model ############################################################################
 ##############################################################################################
+#### Growth Model -- What size will the cacti be next time step? #############################
+##############################################################################################
+growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
+growth_data <- na.omit(growth_data_orig)
+## Lose 2032 rows (due to plant death & recruit status)
+nrow(growth_data_orig)
+nrow(growth_data)
+# check that you are happy with the subsetting
+plot(growth_data$logsize_t, growth_data$logsize_t1)
+points((cactus$logsize_t), (cactus$logsize_t1), col = "red")
+
 ## Create Stan Data for all ant states
 stan_data_grow <- list(N = nrow(growth_data), ## number of observations
-                       y_grow = (growth_data$logsize_t1), ## response volume next year
+                       vol = (growth_data$logsize_t), ## predictors volume
+                       y_grow = (growth_data$logsize_t1), ## response survival next year
+                       ant = as.integer(as.factor(growth_data$ant_t)),## predictors ants
                        K = 4, ## number of ant states
-                       D = 6, ## number of predictors
[…]

### 2022-04-14 Create_Clean_Data_Script.R

@@ -5,8 +5,44 @@
 ##                                                                                                    ##
 ########################################################################################################
 ########################################################################################################
-## load packages
+########################################################################################################
+##############          Load All Necessary Packages Here        ########################################
+########################################################################################################
+library(brms)  # for models
+library(tidyverse)
+library(boot)
+library(rstan)
+library(StanHeaders)
+library(shinystan)
+library(devtools)
+library(tidyr)
+library(lattice)
+library(lme4)
+library(dplyr)
+library(tidyverse)
+library(nnet)
+library(fixest)
+library(mlogit)
+library(dfidx)
+library(effects)
+library(bbmle)
 library(magrittr)
+library(Gmisc)
+library(RColorBrewer)
+library(ggeffects)
+library(heplots)
+library("posterior")
+library(Rlab)
+library(extraDistr)
+library(gridExtra)
+library(grid)
+library(gbm)
+library(dismo)
+library(popbio)
+#library(sjPlot)
+knitr::opts_chunk$set(echo = TRUE)
+options(mc.cores = parallel::detectCores())
+
 knitr::opts_chunk$set(echo = TRUE)
 options(mc.cores = parallel::detectCores())
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Data Analysis")
@@ -14,21 +50,17 @@ setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Data Analysis")
 ################        import the data -- Cacti (main)        #######################################################
 #######################################################################################################
[…]

### 2022-04-14 Data Analysis/Cholla_Analysis_Vital_Rates_One_Ant.R

@@ -1,583 +1,0 @@
-options(mc.cores = parallel::detectCores())
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
-#setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
-#######################################################################################################
-##
-##                  The purpose of this file is to run each vital rate sub model separately
-##                          and run the vital rate models as one large model 
-##
-#######################################################################################################
-source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Setup_Data.R")
-
-#### Growth Model ############################################################################
-##############################################################################################
-## Create Stan Data Crem
-stan_data_grow_crem <- list(N_grow = nrow(growth_data_crem), ## number of observations
-                       vol_grow = log(growth_data_crem$volume_t), ## predictors volume
-                       y_grow = log(growth_data_crem$volume_t1), ## response volume next year
-                       ant_grow = growth_data_crem$ant,## predictors ants
-                       N_ant = 1, ## number of ant states
-                       N_Year_grow = max(growth_data_crem$year), ## number of years
-                       N_Plot_grow = max(growth_data_crem$Plot), ## number of plots
-                       plot_grow = growth_data_crem$Plot, ## predictor plots
-                       year_grow = growth_data_crem$year ## predictor years
-) 
-stan_data_grow_liom <- list(N_grow = nrow(growth_data_liom), ## number of observations
-                            vol_grow = log(growth_data_liom$volume_t), ## predictors volume
-                            y_grow = log(growth_data_liom$volume_t1), ## response volume next year
-                            ant_grow = growth_data_liom$ant,## predictors ants
-                            N_ant = 1, ## number of ant states
-                            N_Year_grow = max(growth_data_liom$year), ## number of years
-                            N_Plot_grow = max(growth_data_liom$Plot), ## number of plots
-                            plot_grow = growth_data_liom$Plot, ## predictor plots
-                            year_grow = growth_data_liom$year ## predictor years
-) 
-stan_data_grow_other <- list(N_grow = nrow(growth_data_other), ## number of observations
-                            vol_grow = log(growth_data_other$volume_t), ## predictors volume
[…]

### 2022-04-14 Data Analysis/Cholla_Analysis_Vital_Rates_Two_Ant.R

@@ -1,573 +1,0 @@
-options(mc.cores = parallel::detectCores())
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
-#setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
-#######################################################################################################
-##
-##                  The purpose of this file is to run each vital rate sub model separately
-##                          and run the vital rate models as one large model 
-##
-#######################################################################################################
-source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Setup_Data.R")
-
-#### Growth Model ############################################################################
-##############################################################################################
-## Create Stan Data Crem Liom
-stan_data_grow_crem_liom <- list(N_grow = nrow(growth_data_crem_liom), ## number of observations
-                       vol_grow = log(growth_data_crem_liom$volume_t), ## predictors volume
-                       y_grow = log(growth_data_crem_liom$volume_t1), ## response volume next year
-                       ant_grow = growth_data_crem_liom$ant,## predictors ants
-                       N_ant = 2, ## number of ant states
-                       N_Year_grow = max(growth_data_crem_liom$year), ## number of years
-                       N_Plot_grow = max(growth_data_crem_liom$Plot), ## number of plots
-                       plot_grow = growth_data_crem_liom$Plot, ## predictor plots
-                       year_grow = growth_data_crem_liom$year ## predictor years
-) 
-#Check that you are happy with the subsetting
-plot(stan_data_grow_crem_liom$vol_grow, stan_data_grow_crem_liom$y_grow)
-plot(log(cactus_crem_liom$volume_t), log(cactus_crem_liom$volume_t1))
-## Run the Model
-#Check if the model is written to the right place
-#stanc("STAN Models/grow_mix_ant.stan")
-fit_grow_mix_ant_crem_liom <- stan(file = "STAN Models/grow_mix_ant_crem_liom.stan", data = stan_data_grow_crem_liom, warmup = 500, iter = 1000, chains = 3, cores = 3, thin = 1)
-grow_outputs_crem_liom <- rstan::extract(fit_grow_mix_ant_crem_liom, pars = c("beta0","beta1","sigma"))
-grow_yrep_crem_liom <- rstan::extract(fit_grow_mix_ant_crem_liom, pars = c("y_rep"))$y_rep
-write.csv(grow_outputs_crem_liom, "grow_outputs_crem_liom.csv")
-
[…]

### 2022-04-14 Data Analysis/STAN Models/multinomialR.R

@@ -1,0 +1,589 @@
+## simulation code from: https://stats.stackexchange.com/questions/103728/simulating-multinomial-logit-data-with-r
+#Adding library for multinomial logit regression
+setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
+library("nnet")
+library(rstan)
+library(bayesplot)
+
+###### Simulate Fake Data 
+#Genarating 500 random numbers with zero mean
+x = rnorm(3885,0)
+#Assigning the values of beta1 and beta2
+Beta1 = 20
+Beta2 = .05
+#Calculation of denominator for probability calculation
+Denominator= exp(Beta1*x)+exp(Beta2*x)
+#Calculating the matrix of probabilities for three choices
+vProb = cbind(1/Denominator, exp(x*Beta1)/Denominator, exp(x*Beta2)/Denominator )
+## check that all rows sum to 1
+sum(vProb[100,])
+# Assigning the value one to maximum probability and zero for rest to get the appropriate choices for value of x
+mChoices = t(apply(vProb, 1, rmultinom, n = 1, size = 1))
+# Value of Y and X together
+dfM = cbind.data.frame(y = apply(mChoices, 1, function(x) which(x==1)), x)
+#We want zero intercept hence x+0 hence the foumula of regression as below
+fit<-(multinom(y ~ x + 0, dfM))
+#This function uses first y as base class 
+#hence upper probability calculation is changed
+summary(fit)
+#In case we do not keep intercept as zero
+fit2<-multinom(y ~ x, dfM)
+summary(fit2)
+
+
+###### Import real data and format it
+cactus_real <- cactus[,c("ant_t","ant_t1","logsize_t","Year_t","Plot")]
+cactus_real <- na.omit(cactus_real)
+cactus_real$ant_t1_relevel <- relevel(cactus_real$ant_t1,ref = "vacant")
+cactus_real$ant_t_relevel <- relevel(cactus_real$ant_t, ref = "vacant")
+cactus_real <- cactus_real[,c("ant_t_relevel","ant_t1_relevel","logsize_t", "ant_t", "ant_t1","Year_t","Plot")]
+
+#### fit model of the mean using the nnet function
+table(cactus_real$ant_t1)/nrow(cactus_real)
[…]

### 2022-04-14 Extras/Max Lik Models/Max Like Models.R

@@ -162,7 +162,7 @@ ant_transition_models[[3]] <- multinom(ant_t1_relevel ~ 0+ant_t + logsize_t, dat
 ant_transition_models[[4]] <- multinom(ant_t1_relevel ~ 0+ant_t * logsize_t, data = cactus)
 AICtab(ant_transition_models)
 ## The selected model is 4
-predict(ant_transition_models[[4]],newdata=cactus[1,],type="probs")
+predict(ant_transition_models[[3]],newdata=cactus[1,],type="probs")
 
 summary(ant_transition_models[[4]])
 other_other <- coef(ant_transition_models[[4]])[1]

### 2022-04-14 Extras/multinomial.R

@@ -33,11 +33,11 @@ summary(fit2)
 
 
 ###### Import real data and format it
-cactus_real <- cactus[,c("ant_t","ant_t1","logsize_t")]
+cactus_real <- cactus[,c("ant_t","ant_t1","logsize_t","Year_t","Plot")]
 cactus_real <- na.omit(cactus_real)
 cactus_real$ant_t1_relevel <- relevel(cactus_real$ant_t1,ref = "vacant")
 cactus_real$ant_t_relevel <- relevel(cactus_real$ant_t, ref = "vacant")
-cactus_real <- cactus_real[,c("ant_t_relevel","ant_t1_relevel","logsize_t", "ant_t", "ant_t1")]
+cactus_real <- cactus_real[,c("ant_t_relevel","ant_t1_relevel","logsize_t", "ant_t", "ant_t1","Year_t","Plot")]
 
 #### fit model of the mean using the nnet function
 table(cactus_real$ant_t1)/nrow(cactus_real)
@@ -399,7 +399,7 @@ multi_dat_real <- list(K = length(unique(cactus_real$ant_t1_relevel)), #number o
                        N = dim(cactus_real)[1], #number of observations
                        D = 5, #number of predictors
                        y = as.integer(as.factor(cactus_real$ant_t1_relevel)), #observations
-                       x = model.matrix(~ 0 + (as.factor(ant_t_relevel)) + logsize_t, cactus_real)) #design matrix
+                       x = model.matrix(~ 0 + (as.factor(ant_t_relevel)) + logsize_t + as.factor(Year_t) + Plot, cactus_real)) #design matrix
 ## Run the model & save the results
 real_ant_size <- stan(file = "Data Analysis/STAN Models/multi_prac_tom_Km1.stan", 
                  data = multi_dat_real, warmup = 100, iter = 1000, chains = 1)
@@ -474,13 +474,13 @@ pred_liom<-cbind(
   exp(mean(real_ant_size_out$beta[,4,4]) + size_dummy_real*mean(real_ant_size_out$beta[,5,4]))/Denominator_liom)
 sum(pred_liom[1,])
                       ## vac -> vac       vac -> other    vac -> crem       vac -> liom
-pred_probs_vac <- cbind((pred_vac[,1]) , (pred_vac[,2]) , (pred_vac[,3]) , (pred_vac[,4]))
[…]

### 2022-04-14 Figures/Cholla_Analysis_Visualization_Transition_Rates.R

@@ -1,0 +1,641 @@
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
+source( "/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Cholla_Analysis_Vital_Rates.R")
+
+
+################################################################################################################################################
+##################### Binomial Visuals ########################################################################
+###############################################################################################################################################
+size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm = T), by = 0.1)
+binom_out <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/binom_outputs.csv", header = TRUE,stringsAsFactors=T)
+occ_occ = (quantile(binom_out$beta1.1, 0.5)) + size_dummy * quantile(binom_out$beta1.2, 0.5)
+occ_occ_low = (quantile(binom_out$beta1.1, 0.05)) + size_dummy * quantile(binom_out$beta1.2, 0.05)
+occ_occ_high = (quantile(binom_out$beta1.1, 0.95)) + size_dummy * quantile(binom_out$beta1.2, 0.95)
+occ_vac = 1 - occ_occ
+occ_vac_low = 1 - occ_occ_low
+occ_vac_high = 1 - occ_occ_high
+vac_occ_low = quantile(binom_out$beta0.1,0.05) + size_dummy * quantile(binom_out$beta0.2, 0.05)
+vac_occ = quantile(binom_out$beta0.1,0.5) + size_dummy * quantile(binom_out$beta0.2, 0.5)
+vac_occ_high = quantile(binom_out$beta0.1,0.95) + size_dummy * quantile(binom_out$beta0.2, 0.95)
+vac_vac = 1 - vac_occ
+vac_vac_low = 1 - vac_occ_low
+vac_vac_high = 1 - vac_occ_high
+
+
+png("binom_panels.png")
+plot(size_dummy, invlogit(occ_occ), type = "l", ylim = c(0,1), lty = 2, ylab = "Probability of being Occupied", xlab = "Log(Volume)", main = "Transition Probabilities \n Occupied vs. Vacant")
[…]

### 2022-04-14 Figures/Cholla_Analysis_Visualization_Vital_Rates.R

@@ -32,13 +32,15 @@ table(cactus_herb$herb, cactus_herb$ant_t)
 #### Growth Visuals #####################################################################################################
 ##########################################################################################################################################################################################################
 ## Extract & Format Data
+growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
+growth_data <- na.omit(growth_data_orig)
 #extract from original data
 
-y_subset <- growth_data[,c("logsize_t1","ant_t_relevel", "logsize_t")]
-y_crem_subset_grow <- subset(y_subset, ant_t_relevel == "crem")
-y_liom_subset_grow <- subset(y_subset, ant_t_relevel == "liom")
-y_vac_subset_grow <- subset(y_subset, ant_t_relevel == "vac")
-y_other_subset_grow <- subset(y_subset, ant_t_relevel == "other")
+y_subset <- growth_data[,c("logsize_t1","ant_t", "logsize_t")]
+y_crem_subset_grow <- subset(y_subset, ant_t == "crem")
+y_liom_subset_grow <- subset(y_subset, ant_t == "liom")
+y_vac_subset_grow <- subset(y_subset, ant_t == "vacant")
+y_other_subset_grow <- subset(y_subset, ant_t == "other")
 grow_out <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/grow_outputs.csv", header = TRUE,stringsAsFactors=T)
 ## Formulas
 y_other_mean_grow <- quantile(grow_out$beta0.3,0.5) + size_dummy * quantile(grow_out$beta1.3,0.5)
@@ -67,8 +69,8 @@ plot(x = size_dummy  ,y = y_other_mean_grow, type = "l", col = "black", lwd = 4,
 for(i in 1:samp){
   lines(x = size_dummy, y = (grow_data$beta0.3[i] + size_dummy * grow_data$beta1.3[i]),col = "lightgrey", alpha = 0.1)
 }
[…]

### 2022-04-14 Figures/Post_Dist_Check.R

@@ -1,196 +1,0 @@
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
-source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Cholla_Analysis_Vital_Rates.R")
-#### Y_rep checks
-
-## Growth ################################################################################################
-#For overlay plots
-y <- growth_data$logsize_t1
-grow_out_all <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/grow_outputs_all.csv", header = TRUE,stringsAsFactors=T)
-yrep_grow <- grow_yrep_all#read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/grow_yrep_all.csv", header = TRUE,stringsAsFactors=T)-all
-samp100 <- sample(nrow(yrep_grow), 500)
-## Overlay Plots
-png(file = "grow_post1.png")
-bayesplot::ppc_dens_overlay(y, yrep_grow[samp100,])
-dev.off()
-png(file = "grow_ant_post1.png")
-bayesplot::ppc_dens_overlay_grouped(y, yrep_grow[samp100,], group = ant_grow)
-dev.off()
-## Convergence Plots
-png("grow_conv2.png")
-bayesplot::mcmc_trace(As.mcmc.list(fit_grow_all, pars=c("beta0", "beta1")))
-dev.off()
-
-## Survival ################################################################################################
-#overlay plot data
-y <- survival_data$Survival_t1
-surv_data <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/surv_outputs.csv", header = TRUE,stringsAsFactors=T)
-yrep_surv <- surv_yrep_all
-samp100 <- sample(nrow(yrep_surv), 500)
-## Overlay Plots
-png(file = "surv_post1.png")
-bayesplot::ppc_dens_overlay(y, yrep_surv[samp100,])
-dev.off()
-png(file = "surv_ant_post1.png")
-bayesplot::ppc_dens_overlay_grouped(y, yrep_surv[samp100,],group = as.integer(as.factor(survival_data$ant_t)))
-dev.off()
-## Convergence Plots
-png(file = "surv_conv1.png")
[…]

### 2022-04-14 Figures/fruit_analysis_visualizations.R

@@ -1,13 +1,0 @@
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
-
-################# Seed Abortion Rates ########################################################################################################
-
-
-
-
-
-
-
-
-
-################# Seed Production  ########################################################################################################

### 2022-04-14 IPM Code/Analysis.R

@@ -8,65 +8,83 @@ setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
 source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code/IPM.R")
 
 #### One ant results
-##Crem
-for(i in 1:100){
-  params = cholla[,i]
-  big_list_c[[i]]<-bigmatrix.1(x_dum,y_dum,cholla, lower, upper, matsize)$IPMmat_c
-  mat_c <- big_list_c[[i]]
-  eig_c <- eigen(mat_c)
-  lambda_c[i]<-Re(eig_c$values[1])
-  
-  stable_c[[i]] <- stable.stage(mat_c)
+bigmatrix.1(4,4.5,lower,upper,matsize,1,"crem",(params))$IPMmat
+## Check that it works
+i = c("liom","vac","crem","other")
+j = c("vac","crem","crem","liom")
+x = c(-1,-5,4,3)
+y = c(-1,-4,4.5,3.01)
+big1 <- list()
+lambda1 <- list()
+for(n in 1:length(i)){
+  big1[[n]] <- bigmatrix.1(x[n],y[n],lower,upper,matsize,1,i[n],params)
+  lambda1[[n]] <- Re(eigen(big1[[n]]$IPMmat)$values[1])
 }
-Re(eigen(bigmatrix.1(x_dum,y_dum,cholla, lower, upper, matsize)$IPMmat_c)$values[1])
-hist(lambda_c)
-lambda_c
+lambda1
 
-##Liom
-Re(eigen(bigmatrix.1(x_dum,y_dum,cholla,lower,upper,matsize,1)$IPMmat_l)$values[1])
+## Crem
+crem_data <- subset(cactus, cactus$ant_t == "crem" & cactus$ant_t1 == "crem")
+i <- crem_data$ant_t
+j <- crem_data$ant_t1
+x <- crem_data$logsize_t
+y <- crem_data$logsize_t1
+big1 <- list()
+lambda1 <- list()
+lambda <- vector()
+for(n in 1:length(i)){
+  big1[[n]] <- bigmatrix.1(x[n],y[n],lower,upper,matsize,1,i[n],params)
+  lambda1[[n]] <- Re(eigen(big1[[n]]$IPMmat)$values[1])
+}
+lambda1
+## 0.9929766
 
-##Other
[…]

### 2022-04-14 IPM Code/IPM.R

@@ -9,36 +9,54 @@ invlogit<-function(x){exp(x)/(1+exp(x))}
 ## ----------- Vital rate functions. Parameter indices are hard-coded and must correspond to rows of MCMC matrix ------------- ##
 ##############################################
 #GROWTH FROM SIZE X TO Y
-gxy<-function(x,y,params,i){
+gxy<-function(x,y,i,params){
   xb=pmin(pmax(x,cholla_min),cholla_max) #Transforms all values below/above limits in min/max size (So the params are the minimums and maximums of size?)
-  g_crem = dnorm(y,mean=params[1] + params[2]*xb,sd=params[3])
-  g_vac = dnorm(y,mean=params[301] + params[302]*xb,sd=params[3])
-  g_other = dnorm(y,mean=params[201] + params[202]*xb,sd=params[3])
-  g_liom = dnorm(y,mean=params[101] + params[102]*xb,sd=params[3])
+  g_crem = dnorm(y,mean=(params$grow_beta01) + (params$grow_beta11)*xb,sd=(params$grow_sig))
+  g_vac = dnorm(y,mean=(params$grow_beta04) + (params$grow_beta14)*xb,sd=(params$grow_sig))
+  g_other = dnorm(y,mean=(params$grow_beta03) + (params$grow_beta13)*xb,sd=(params$grow_sig))
+  g_liom = dnorm(y,mean=(params$grow_beta02) + (params$grow_beta12)*xb,sd=(params$grow_sig))
   if(i == "crem"){ return(g_crem)}
   if(i == "liom"){ return(g_liom)}
   if(i == "other"){ return(g_other)}
   if(i == "vac"){ return(g_vac)}
 }
 
+gxy<-function(x,y,i,params){
+  xb=pmin(pmax(x,cholla_min),cholla_max) #Transforms all values below/above limits in min/max size (So the params are the minimums and maximums of size?)
+  g_crem = dnorm(y,mean=(params[,4]) + (params[,5])*xb,sd=(params[,1]))
+  g_vac = dnorm(y,mean=(params[,10]) + (params[,11])*xb,sd=(params[,1]))
+  g_other = dnorm(y,mean=(params[,8]) + (params[,9])*xb,sd=(params[,1]))
[…]

### 2022-04-14 IPM Code/Params.R

@@ -2,24 +2,10 @@
 
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
 
-## -------- read in MCMC output ---------------------- ##
 
-##This file contains random draws from the joint posterior distribution of all parameters
-post.params <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/params_outputs.csv", header = TRUE,stringsAsFactors=T)          
-Ndraws<-min(100,nrow(post.params))
-##This file contains random draws from the joint posterior distribution of the occupied vs vacant transition model 
-trans.params_occ <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/ant_outputs_occ.csv", header = TRUE,stringsAsFactors=T)
-post.params$beta0_ant.1 <- trans.params_occ$beta0.1[1:Ndraws]
-post.params$beta1_ant.1 <- trans.params_occ$beta1.1[1:Ndraws]
-post.params$beta0_ant.2 <- trans.params_occ$beta1.1[1:Ndraws]
-post.params$beta1_ant.2 <- trans.params_occ$beta1.2[1:Ndraws]
 
-## Number of draws to take from the joint posterior distribution of the parameters. 
-## Cannot be greater than the number of draws provided in the .csv file, which is 1500.
-post.params<-post.params[1:Ndraws,]
-
-cholla_min<- min(log(cactus$volume_t), na.rm = TRUE)  ## minsize 
-cholla_max<- max(log(cactus$volume_t), na.rm = TRUE)  ## maxsize 
+cholla_min<- min((cactus$logsize_t), na.rm = TRUE)  ## minsize 
+cholla_max<- max((cactus$logsize_t), na.rm = TRUE)  ## maxsize 
 
 Nplots <- length(unique(cactus$Plot))
 Nyears <- length(unique(cactus$Year_t))
@@ -36,120 +22,300 @@ h<-(U-L)/n                   #Bin size
 b<-L+c(0:n)*h;               #Lower boundaries of bins 
 y<-0.5*(b[1:n]+b[2:(n+1)]);  #Bin midpoints
 
-## -------- load IPM source functions ---------------------- ##
 
-#source("IPM.R")
+## -------- read in MCMC output ---------------------- ##
 
-## -------- Set up IPM parameter vector ---------------------- ##
[…]

### 2022-04-14 IPM Shtuff.R

@@ -1,0 +1,958 @@
+#########################################################################################################
+##            This will be an IPM which allows you to choose how many ants are present
+#########################################################################################################
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
+source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code/Params.R")
+## ----------- Miscellany...we'll need an inverse logit functions ------------- ##
+invlogit<-function(x){exp(x)/(1+exp(x))}
+
+## ----------- Vital rate functions. Parameter indices are hard-coded and must correspond to rows of MCMC matrix ------------- ##
+##############################################
+#GROWTH FROM SIZE X TO Y
+gxy<-function(x,y,params,i){
+  xb=pmin(pmax(x,cholla_min),cholla_max) #Transforms all values below/above limits in min/max size (So the params are the minimums and maximums of size?)
+  g_crem = dnorm(y,mean=params[4,] + params[5,]*xb,sd=params[1,])
+  g_vac = dnorm(y,mean=params[10,] + params[11,]*xb,sd=params[1,])
+  g_other = dnorm(y,mean=params[8,] + params[9,]*xb,sd=params[1,])
+  g_liom = dnorm(y,mean=params[6,] + params[7,]*xb,sd=params[1,])
+  if(i == "crem"){ return(g_crem)}
+  if(i == "liom"){ return(g_liom)}
+  if(i == "other"){ return(g_other)}
+  if(i == "vac"){ return(g_vac)}
+}
+
+##Check that it works properly
+i = c("liom","vac","crem","other")
+x = c(-1,-5,4,3)
+y = c(-1,-4,4.5,3.01)
+crem<-vector()
+for(n in seq(1:length(i))){
+  crem[n]<-gxy(x[n],y[n],cholla.grow,i[n])
+  g[,n] <- gxy(x[n],y[n],cholla.grow,i[n])
+}
+crem
+
+#################################################
+#SURVIVAL AT SIZE X.
+sx<-function(x,params,i){
+  xb=pmin(pmax(x,cholla_min),cholla_max)
+  s_crem = invlogit(params[4] + params[5]*xb)
+  s_vac = invlogit(params[10] + params[11]*xb)
+  s_other = invlogit(params[8] + params[9]*xb)
+  s_liom = invlogit(params[6] + params[7]*xb)
[…]

### 2022-04-14 Params Shtuff.R

@@ -1,0 +1,140 @@
+## -------- set working directory ---------------------- ##
+
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
+
+## -------- read in MCMC output ---------------------- ##
+
+##This file contains random draws from the joint posterior distribution of all parameters
+post.params <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/params_outputs.csv", header = TRUE,stringsAsFactors=T)          
+Ndraws<-min(100,nrow(post.params))
+##This file contains random draws from the joint posterior distribution of the occupied vs vacant transition model 
+trans.params_occ <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/ant_outputs_occ.csv", header = TRUE,stringsAsFactors=T)
+post.params$beta0_ant.1 <- trans.params_occ$beta0.1[1:Ndraws]
+post.params$beta1_ant.1 <- trans.params_occ$beta1.1[1:Ndraws]
+post.params$beta0_ant.2 <- trans.params_occ$beta1.1[1:Ndraws]
+post.params$beta1_ant.2 <- trans.params_occ$beta1.2[1:Ndraws]
+
+## Number of draws to take from the joint posterior distribution of the parameters. 
+## Cannot be greater than the number of draws provided in the .csv file, which is 1500.
+post.params<-post.params[1:Ndraws,]
+
+cholla_min<- min(log(cactus$volume_t), na.rm = TRUE)  ## minsize 
+cholla_max<- max(log(cactus$volume_t), na.rm = TRUE)  ## maxsize 
+
+Nplots <- length(unique(cactus$Plot))
+Nyears <- length(unique(cactus$Year_t))
+iter <- 1000
+matsize<-200
+floor.extend=0.9*cholla_min
+ceiling.extend=1.1*cholla_max
+lower<- cholla_min - floor.extend
+upper<- cholla_max + ceiling.extend
+
+n<-matsize
+L<-lower; U<-upper
+h<-(U-L)/n                   #Bin size
+b<-L+c(0:n)*h;               #Lower boundaries of bins 
+y<-0.5*(b[1:n]+b[2:(n+1)]);  #Bin midpoints
+
+## -------- load IPM source functions ---------------------- ##
+
+#source("IPM.R")
+
[…]

### 2022-04-14 Setup_Script.R

@@ -7,7 +7,7 @@
 ########################################################################################################
 
 ## call the data cleaned in the Create_Clean_Data_Script
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Data Analysis")
 source("Create_Clean_Data_Script.R")
 
 ########################################################################################################
@@ -49,7 +49,7 @@ knitr::opts_chunk$set(echo = TRUE)
 options(mc.cores = parallel::detectCores())
 
 ## Import the data
-cactus <- read.csv("cholla_demography_20042021_cleaned.csv", header = TRUE,stringsAsFactors=T)
+cactus <- read.csv("cholla_demography_20042019_cleaned.csv", header = TRUE,stringsAsFactors=T)
 
 #######################################################################################################
 ################        import the data -- Germination  (No ant state)      ###########################
@@ -66,13 +66,18 @@ germ.dat <- germ.dat[-c(42,39,40),]
 #Check that you are happy with the subsetting
 plot(germ.dat$rate)
 points(germ.dat_orig$rate, col = "red")
+## Most of the rows are included. We only lose 3, but it is still a small dataset. 
+nrow(germ.dat)
+nrow(germ.dat_orig)
 #######################################################################################################
 ################        import the data -- Fruit Survival  (No ant state)      ########333#############
 #######################################################################################################
-seedling.dat <- cactus[,c("logsize_t","logsize_t1","Recruit")]
-filter(cactus, Recruit == 1)
-seedling.dat <- filter(seedling.dat, Recruit == 1)
-seedling.dat <- na.omit(seedling.dat)
+seedling.dat_orig <- subset(cactus, cactus$Recruit == 1)
+seedling.dat_orig <- seedling.dat_orig[,c("logsize_t1","Recruit")]
+nrow(seedling.dat_orig)
+seedling.dat <- na.omit(seedling.dat_orig)
+nrow(seedling.dat)
+## losing 19 rows
 # check that you are happy with the subsetting
 plot(seedling.dat$logsize_t1, seedling.dat$Recruit, xlim = c(-5,15), ylim = c(0,1))
 points(cactus$logsize_t1, cactus$Recruit, col = "red")
@@ -82,7 +87,9 @@ points(cactus$logsize_t1, cactus$Recruit, col = "red")
 precensus.dat.orig<-read.csv("PrecensusSurvival.csv") 
[…]

### 2022-05-05 Paper/main.tex
moving ms to github

\documentclass[12pt,a4paper]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{multirow}
\usepackage{multicol}
\setlength{\columnsep}{1.5cm}
\usepackage{setspace}
\usepackage{url}
\usepackage{array}
\usepackage{xfrac}
\usepackage{mwe}
\usepackage{tikz}
\usepackage{lmodern}
\usepackage{flowchart}\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{color,soul}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{hyperref}
\usepackage{biblatex}
\addbibresource{References.bib}
\title{Thank You, Next}
\author{Ali M. Campbell\\
Tom E.X. Miller}
\begin{document}
	\maketitle
%	\begin{multicols}{2}

	\section*{Abstract}

	Understanding interspecific mutualisms is a challenge, because they are among the most widespread species interactions with diverse and dynamic consequences. Depending on the environmental contexts the outcome of an interspecific cooperation can range from parasitic to mutualistic. Most studies focus on a pair of species interacting in a mutualism, however often this is a simplification of the reality. These mutualistic interactions become even more complex when put into the context of a multi-species mutualism, where there are multiple partners interacting with the same focal species. The partners can directly impact the focal species through the interactions, and indirectly impact them through interactions with each other. The cactus \textit{Cylindriopuntia imbricata} forms symbioses with multiple ant species, but each individual plant is able to interact with only one species at a time. Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny.Our results demonstrate the importance of evaluating a mutualism within a community context and suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti.

	Each of the ant partners in this system are within the same guild, all offering defense from herbivores, with only one ant species interacting with each individual cactus in a given period of time. Though these ants may offer very similar rewards, they likely differ in some way. Two possible ways they could differ are the fitness boost they offer the focal species or the period of ontogeny in which they offer the greatest fitness boost. Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny. Our results demonstrate the importance of evaluating a mutualism within a community context and suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti.

	\section*{Context/Introduction}
	\subsection*{Context}

	Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences, yet they are still poorly understood. They form a complex foundation to many natural systems which involve both costs and benefits, despite their nature of being beneficial.

	There are benefits of mutualistic interactions, but there are costs as well, some direct and some indirect. Directly, mutualistic partners can sometimes become parasitic, causing harm to the focal mutualist\cite{Bronstein2001a}. More indirectly, these interactions require energy. Within a mutualism, each participant offers rewards and in return receives benefits from their partners. These rewards come in many forms, nectar, protection, habitat, etc., all of which require some investment from the individual\cite{Bronstein2001}. These investments can be seen as the costs of mutualisms, they are traded for the rewards provided by a partner. Because of these trade-offs, there are both costs and benefits to a mutualistic interaction\cite{Bronstein2001,Bronstein2001a}, particularly as the mutualism becomes more complex in the context of its environment.

	Mutualistic interactions can occur simply between two species, a pairwise mutualism, or between a group of species, a multi-species mutualism\cite{Stanton2013}. Pairwise mutualisms may provide benefits of efficiency in the context of mutualisms, but they are not a good investment due to temporal and spatial variation\cite{Waser1996}. Despite this, pairwise mutualisms are historically well studied, with most mutualism papers being focused on pairwise interactions. These studies are often used to predict outcomes of mutualistic interaction, which may underestimate the combined impacts of multiple partners on the focal mutualist\cite{Stanton2013,Palmer2010}. Pairwise interactions do not always indicate what the net interaction will be within a multi-species context\cite{Chamberlain2014,Song2020}. When many species are included, variation between partners can occur in functional benefit, lifespan, and competitive rankings, among other factors\cite{Stanton2013}. This difference is because in addition to the expected direct benefits of interactions, they can also indirectly impact the focal mutualist through other partner interactions. One reason for this lack of predictability is that interactions between partners changes the relationship with the focal mutualist\cite{Afkhami2014}.  Often the partners interact with each other, either increasing their performances, with extra rewards, or decreasing them, often by fighting\cite{Boucher1982}. This means predicting the rewards of a multi-species mutualism is very difficult.

	Multi-species mutualisms are not always additive: In theory if multiple partners can all form pairwise mutualisms with a focal species, in a multi-species mutualism one would expect more benefits for the focal partner than in any one pairwise scenario. The reality is much less predictable\cite{Stanton2013,Palmer2010,Song2020}. The rewards can be as simple as the sum of all pairwise rewards, aka. additivity, they can be increased beyond the sum of all pairwise rewards, synergy, or they can be reduced below the sum of pairwise rewards, sub-additivity. Specifically, it can often change the quality of benefits.

	Sometimes a partner can confer benefits on the focal mutualist in a pairwise interaction but be a neutral or even antagonistic interactor in a multi-species interaction as partners can vary in quality\cite{Afkhami2014}. Some mutualists are of so low a quality that they are known as freeloaders or cheaters\cite{Song2020,West2007,Frederickson2013} because they return very little benefit, possibly going as far as to be a cost to the focal mutualist in exchange for the benefits they receive. This change in quality of partners is often due to competition between the partners\cite{Amarasekare2003}. Some partners reduce in quality so drastically that they become antagonistic towards the focal mutualist in these relationships\cite{Afkhami2014,Frederickson2013,Bronstein1994}. Despite the potential to reduce fitness of the focal mutualist they often remain a partner within the mutualism\cite{Jones2015}.

    There are several underlying mechanisms which promote partner diversity that may shed some light on this; specifically, sampling effect,portfolio effect, and complementarity, three branches of biodiversity ecosystem function theory\cite{Batstone2018,Hooper2005}.

    Sampling effect states that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the most beneficial one\cite{Afkhami2014}. Particularly, this means that increased diversity of partners means increased probability that some of the partners will be highly beneficial. The fitness of a multi-species mutualistic population experiencing sampling effect will be equal to the fitness of the same population if they interacted only with their strongest partner.

    Complementarity states that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in a broader benefit\cite{Winfree2020}. This means that in populations where there is sufficient interaction with the diverse partners there is likely to be increased rewards and fitness. The fitness of a multi-species mutualistic population experiencing complementarity is likely to be higher than that of any pairwise mutualistic population.

    Portfolio effect states that each partner is the best under a different set of conditions, meaning across all conditions, meaning there is likely to be an increase in potential benefits\cite{Winfree2020}. Thus, in populations experiencing highly diverse conditions, diverse partners are likely to increase the rewards and fitness of the focal mutualist. The fitness of a multi-species mutualistic population experiencing portfolio effect is likely to have a different strongest partner under different conditions, such as year to year, and will experience similar fitness as they would in a pairwise mutualism with their best partner under any condition.

    These biodiversity ecosystem functions are three which can be used to explain the diversity of mutualistic relationships in nature\cite{Afkhami2014}. If a system benefits from the existence of multiple partners, one of these functions may be at play. Within the section of mutualistic interaction which have multiple partners, there are many different classifications. One type of mutualism that is historically well documented is the interaction between ants and plants\cite{Boege2005,Barton2010}. These interactions can take many forms, including pollination, defense, dispersal, etc\cite{Bronstein2006}. This paper focuses on defensive ant-plant mutualisms. These often involve long-lived plants providing extra-floral nectar and/or habitats to ants which in turn defend them from herbivores\cite{Bronstein1998}.  These are very common interactions and have provided many opportunities for empirical studies of mutualisms. However, many of these studies have been pairwise rather than investigating the impacts that multiple partners may have on the costs and benefits of a system.

    This study focuses on the long lived, faculative \textit{Cylindriopuntia imbricata} (Cholla cacti) and its ant mutualists. The cacti interact with several ant partners, \textit{Crematogaster opuntiae} (\textit{Crem.}) and \textit{Liometopum apiculatum} (\textit{Liom.}) are the most common partners though there are a number of other rarer ant partners, which visit these cacti in smaller population sizes. The ant partners all provide defense to the cacti from predators in return for exclusive interaction with the plant to take the extra-floral nectar, meaning they are considered to be within one guild. Predators of the cholla include insect herbivores and seed predators, which influence the reproductive output of the cacti. Though many multi-species mutualisms result in both costs and benefits, it has been shown that the relationship between these cacti and ants does not result in any major costs for the cacti\cite{Miller2007}. However, there are still some small costs which could influence the optimal set of partners. This is part of what we will be analyzing in this paper.

    Though all the ants associated with the Cholla are within one guild, they may not all be equal, meaning the impacts on the Cholla are complex. In this study, we will answer a number of questions: 
    \begin{enumerate}
        \item Do the Cholla cacti benefit from the multi-species mutualism with multiple ant partners?
        \item What combination of, or individual, ant(s) result in the highest fitness for the Cholla cacti?
        \item Is there evidence of sampling effect, complementarity, or portfolio effect to encourage the current multi-species state of the mutualism?
    \end{enumerate}

    In order to answer these questions, I am constructing an Integral Projection Model (IPM) to calculate the fitness of the Cholla cacti under different conditions. I will be able to quantitatively estimate the fitness of the cholla population if they were to interact with any combination of ant partners, allowing me to determine the best combination of partners. Similar approaches have been used to assess the individual qualities of multiple partners\cite{Palmer2010} but not to determine the optimal combination of partners in a multi-species mutualism.

    We hypothesize that the cacti will experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism is the optimal outcome. We also hypothesize that complementarity and sampling effect will play roles in encouraging this diverse mutualism.

	\section*{Methods}
		\subsection*{Study System}
    To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of Cholla cacti annually since 2004, this is still ongoing. This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico. The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA *** Benson 1982***. These arborescent cacti produce new cylindrical segments every year which develop either into fruits which produce seeds or into vegetative segments with large spine which increase the size of the cactus ***Miller, Tenhumberg & Louda 2008***. In central New Mexico, vegetative growth occurs from May -- August.

	The cholla cacti secrete extra-floral nectar (EFN) from glands on young vegetative segments and reproductive segments. This extra-floral nectar varies in chemical makeup across the lifespan of the cactus *** Miller ??***. 
	At the Sevilleta, the cacti are visited by multiple species of ground-nesting ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and many other species which occur at small frequencies. These ants are attracted to the EFN and in return fight off the seed predators which also visit the cacti. These ants never co-occur on the same plant, instead competing for plant services. Over the lifetime of the cacti each plant interacts with multiple partners as they tend to switch out season to season.

	There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments. There are a number of insects which occur commonly on these cacti at the Sevilleta. These include the weevil, the NP, and the *** (I think there is a moth as well?). The weevil feeds on ***. The NP feeds on ***. Finally, the moth?? feeds on ***. All of these predators can have significant impacts on the fitness of both individual cacti and the population. *** Miller 2009***.

	The data collected on these cacti are from an 18-year data-set. There are 8 30 $\times$ 30 meter plots from which this data is taken. Each plot is surveyed, and each plant within each plot is tagged and logged each year. For every plant we take data on their survival, size, ant state, herbivore presence, flowering data, and seed data. This allows us to track the progression of individual plants across many years. The length of this data-set

	\subsection*{Statistical Modeling}
	To understand the impacts of multiple species interactions on the cacti populations, we constructed bayesian demographic models of many vital rates. With these models, we were able to ask (i) Do the impacts of ant presence on cacti fitness differ across ant species? (ii) What are the probabilities of an individual cactus interacting with each ant species?.

	The vital rates of the cacti considered in this model include survival rate, growth rate, reproductive state, floral abortion rate, seed production, and germination. We fit generalized linear mixed effects models in a heirarchical Bayesian framework to quantify each of the vital rates. Each model takes the data from year \textit{t} and estimates the values in year \textit{t+1}. This allows us to predict a vital rate for any plant with a given set of characteristics, such as size and ant state. 
	\vspace{1cm}

	\subsection*{Demographic Modeling}

	The statistical models described above parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.  This model was used to estimate the growth rate of the cactus population, effectively a quantitative measure of the fitness of the population. The fitness was estimated under the range of all different conditions, no ants present, one ant species present, two ant species present, all ant species present. These allow us to compare the fitness of the cacti under no mutualistic interactions, pairwise mutualisms, and various multi-species mutualisms to show which case results in the highest fitness for the cacti. Following previous studies, we modeled the life cycle of \textit{C. imbracata} using continuously size-structured plants,$n_i(x)$, and two discrete seed banks ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
$$
B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) A_i(x) F(x)n_i(x) dx \\
$$
$$
B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
$$

    The functions $P$, $F$, and $A$ give the probability of flowering, the number of flowerbuds produced, and the proportion of flowerbuds which will create seeds. Each of these functions, estimated by a Bayesian model calculates these values for an $x$-sized plant in year $t$. The proportion of flowerbuds which will produce seeds ($A$) is also dependent on the ant species present on the plant $i$ in year $t$. The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the 1-year old seed bank. Parameters $U$ and $L$ are the upper and lower  bounds, respectively, of the plant size distribution. Plants can recruit out of the 1-year seed bank with the probability of $\gamma_1$ or transition to the 2-year seed bank with a probability of $1 - \gamma_1$. Seeds in the 2-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.

    The size dynamics of the plants are given by:

$$
n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
$$
$$
\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
$$

    The final equation gives the size of the \textit{C. imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$. The first term gives the recruitment from 1 and 2-year seedbanks to size $y$. $\eta(y)$ gives the seedling size distribution, which is assumed to be normally distributed, and $\omega$ gives the proportion of seedlings which survive from germination (late summer) to the census (May). The second term gives the changes in the population of the cacti which are not recruits. The functions $S$ and $G$ give the probabilities of surviving from year $t$ to $t+1$ and growing to size $y$ from year $t$ to year $t+1$, respectively. Each depends on the size $x$ in year $t$ and the ant state $j$ in year $t+1$. Finally, $\tau_{ij}$ is the probability of a cactus which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.

    \subsection*{Model analysis}

	\section*{Results}
	\subsection*{Vital Rates}

	We used field data on cholla cacti to evaluate the differing impacts that the ant mutualism has on the fitness of the cacti. The fitness can be broken down into a number of vital rate components, including survival, growth, probability of flowering, number of fruits and seeds produced, etc. called vital rates. We found that many of these vital rates differ in the presence of different ant species.

	\begin{figure}[!ht]
	\centering
	\includegraphics[width = 0.85\linewidth]{Figures/surv_panels_cropped.png}
	\caption{Probability of cactus survival from year t to year t+1 when tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow). Panels (a-d) show the best-fit line estimating the survival probabilities of cacti, with a grey area representing the uncertainty, and circles representing the actual data (scaled to represent number of observations). (e) shows the best-fit probabilities of survival of different-sized cacti with all ant partners.}
	\label{fig:surv}
	\end{figure}

	The first vital rate we analyzed was the survival of the cacti. As you can see in Figure \ref{fig:surv}, the survival rate of cacti vary significantly across their lifetimes as they grow. Young (small) cacti have relatively low survival rates, near $30\%$ while the older mature cacti have much higher survival rates, near $100\%$. This means cacti are at their most vulnerable to death when they are young. The survival rates differ with more than just the size of the cacti, the ant species present on the cactus is also important.

	There are no very small cacti tended by ant partners because they do not produce EFN. This is reflected in Figure \ref{fig:surv} where only the vacant data (yellow) in panel (d) extends to the smallest sizes. Once cacti reach a size large enough to attract ant partners with EFN, the cacti which are partnered with \textit{Crem.} have the highest survival rates compared to all other cacti. When the cacti reach their larger sizes, the partnered plants all reach a nearly $100\%$ survival rate while plants with no partners have a slightly lower survival rate for all large cacti. This result indicates that \textit{Crem.} tended plants may have an advantage, when they are relatively small, over other cacti.

	The next vital rate we analyzed was the viability rate of the cacti, the proportion of fruits produced which are viable for reproduction. It is clear from Figure \ref{fig:viab} that the cacti tended by \textit{Liom.} ants experience significantly higher proportions of viable fruits. This result aligns with much of the literature which shows that \textit{Liom.} are more aggressive against herbivores and seed predators, therefore are very high quality defenders. This high-quality defense is imperative for high viability rates on reproducing cacti which are likely to attract many antagonists due to the quality of their EFN.

	\begin{figure}[!ht]
	    \centering
	    \includegraphics{Figures/viab_hist.png}
	    \caption{Proportion of viable fruits produced by a cactus in year t+1 when tended in year t by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow). Each panel (a-d) shows a histogram of the actual data overlayed with a density distribution of the predicted viability rate of the cacti. Panel (e) shows the predicted distributions of viability rates separated by ant species.}
	    \label{fig:viab}
	\end{figure}

	A third vital rate considered in this paper is the growth rate. Cacti who are attacked by herbivores are likely to lose sections of their growth, leading to a lower, or even negative, growth rate. In Figure \ref{fig:grow} you can clearly see that untended plants are the only cacti which experience a negative growth rate at any size. This again aligns with expectations that tended cacti are being defended from herbivores and are therefore unlikely to lose enough vegetation to result in a negative growth rate.

	\begin{figure}[!ht]
	    \centering
	    \includegraphics{Figures/grow_panel.png}
	    \caption{The size of a cactus in year t+1 given a previous size in year t and being tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow). Each panel (a-d) shows a best-fit prediction of the future size based on the previous size, the uncertainty in grey surrounding the best-fit, and the actual data in (circles scaled to represent the number of observations). Panel (e) shows all best-fit lines together in the same colors as above and a line (grey, dashed) above which plants are growing (have a positive growth rate) and below which plants have a negative growth rate.}
	    \label{fig:grow}
	\end{figure}

	\begin{figure}[!ht]
	    \centering
	    \includegraphics{Figures/grow_dist_panel2.png}
	    \caption{The size of a cactus in year t+1 given a previous size in year t and being tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow). Each panel (a-d) shows a distribution of the predicted future size based on the previous sizeand a histogram of the actual data. Panel (e) shows all predicted distributions together in the same colors as above.}
	    \label{fig:grow2}
	\end{figure}

	\subsection*{Ant Transition Rates}

	In addition to the vital rates of the cacti, we also analyzed the transition rates from one ant partner to another across years.

	\begin{table}[]
	    \centering
	    \begin{tabular}{|p{0.25 \linewidth}   |p{0.25 \linewidth}|p{0.25 \linewidth}|}
	    \hline
	        \textbf{Ant States Possible} & \textbf{Growth Rate ($\frac{n_{t+1}}{n_t}$)} & \textbf{Population $\lambda$}\\
	        \hline
	        No ants & &\\
	        \hline
	        \textit{Liom.} & & \\
	        \hline
	        \textit{Crem.}  & & \\
	        \hline
	        Other & & \\
	        \hline
	        \textit{Liom.} \& \textit{Crem.} & & \\
	        \hline
	        \textit{Liom.} \& Other & & \\
	        \hline 
	        \textit{Crem.} \& Other & & \\
	        \hline
	        \textit{Liom.}, \textit{Crem.} \& Other & & \\
	        \hline
	    \end{tabular}
	    \caption{Caption}
	    \label{tab:table_sims}
	\end{table}
%	\begin{multicols}{2}

	\section*{Discussion}

     There is a constant tension within life cycles between energy and resource allocation. Individuals often have the ability to either put significant resources towards one process or split their resources between multiple. Full energy allocation to one process generally means better results, but splitting resources is a form of bet hedging ***. The cacti studied here experience a certain tension between their many vital rates, particularly growth and reproduction because each year they produce new floral buds and new vegetative growth. Small plants rarely put out flowers and instead focus on growth, but larger plants must choose each year. The analyses of the individual vital rates indicate that different ant species may protect different processes better than others. In particular \textit{Crem.} may offer advantages to smaller plants by protecting their new vegetation well. This is seen in the higher survival rates and growth rates of small and medium cacti with \textit{Crem.} tenders in Figure \ref{fig:grow} and Figure \ref{fig:surv}. \textit{Liom.} ants alternatively may protect reproductive focus more than other partners. The \textit{Crem.} tended plants however, have lower viability rates than any other ant tended plants, indicating they do not protect flowers well from herbivores(Figure \ref{fig:viab}). This is seen in Figure \ref{fig:viab} where it is shown that plants tended by \textit{Liom.} have the highest viability rates of their flowers, resulting in higher pollination rates. Larger plants are also less likely to lose branches when they are tended by \textit{Liom.} (Figure \ref{fig:grow}), keeping them healthy and more able to generate new plants. Each of these ants individually offer both benefits and drawbacks but considered together may indicate complementarity across ontogeny.

%	\end{multicols}

\end{document}

\typeout{\bibliography{References.bib}}

### 2022-05-10 Manuscript/manuscript_draft.tex
Create manuscript_draft.tex

--- 2022-05-05 Paper/main.tex
@@
 \usepackage{lmodern}
-\usepackage{flowchart}\usetikzlibrary{shapes,arrows,positioning,calc,fit}
+\usepackage{flowchart}
+\usetikzlibrary{shapes,arrows,positioning,calc,fit}
 \usepackage{graphics}
@@
 \usepackage{color,soul}
-\usepackage[numbers,sort&compress]{natbib}
+%\usepackage[numbers,sort&compress]{natbib}
 \usepackage{hyperref}
+%\usepackage{cite}
 \usepackage{biblatex}
@@
 	\maketitle
-%	\begin{multicols}{2}
 
@@
 \subsection*{Study System}
+
 To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of Cholla cacti annually since 2004, this is still ongoing.
@@
 The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA *** Benson 1982***.
-These arborescent cacti produce new cylindrical segments every year which develop either into fruits which produce seeds or into vegetative segments with large spine which increase the size of the cactus ***Miller, Tenhumberg & Louda 2008***.
+These arborescent cacti produce new cylindrical segments every year which develop either into fruits which produce seeds or into vegetative segments with large spine which increase the size of the cactus %***Miller, Tenhumberg \& Louda 2008***.
 In central New Mexico, vegetative growth occurs from May -- August.
@@
 
-%	\end{multicols}
-
 \end{document}

### 2022-06-17 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex
finished commenting on intro

@@
 \addbibresource{References.bib}
+\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
+
 \title{Thank You, Next}
@@
 
-Understanding interspecific mutualisms is a challenge, because they are among the most widespread species interactions with diverse and dynamic consequences.
-Depending on the environmental contexts the outcome of an interspecific cooperation can range from parasitic to mutualistic.
-Most studies focus on a pair of species interacting in a mutualism, however often this is a simplification of the reality.
-These mutualistic interactions become even more complex when put into the context of a multi-species mutualism, where there are multiple partners interacting with the same focal species.
-The partners can directly impact the focal species through the interactions, and indirectly impact them through interactions with each other.
-The cactus \textit{Cylindriopuntia imbricata} forms symbioses with multiple ant species, but each individual plant is able to interact with only one species at a time.
-Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny.Our results demonstrate the importance of evaluating a mutualism within a community context and suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti.
+\tom{Understanding interspecific mutualisms is a challenge, because they are among the most widespread species interactions with diverse and dynamic consequences.}{This sentence is pretty vague and does not have much information content.}
+Depending on environmental context the outcome of the interaction can range from parasitic to mutualistic. 
+Most studies focus on a pair of species interacting in a mutualism, \tom{however often this is a simplification of the reality}{awkward and unclear -- pairs are ``real'' after all}. 
+These mutualistic interactions \tom{become even more complex}{more complex than what?} when put into the context of a multi-species mutualism, where there are multiple partners interacting with the same focal species.
+The partners can directly impact the focal species through the interactions, \tom{and indirectly impact them through interactions with each other}{This is true but not something we are well poised to address with our data so might be risky to bring it up.}.
+The cactus \textit{Cylindriopuntia imbricata} forms \tom{symbioses}{technically it's not a symbiosis because the partners live independently} with multiple ant species, but each individual plant is able to \tom{interact}{This sentence does not sufficiently convey the nature of the interaction - the plants reward ants with extrafloral nectar in exchange for defense} with \tom{only one species at a time}{There is not enough context to know why this information is relevant here.
+Abstracts should be brutally efficient in communicating the essentials.}.
+Using \tom{a long-term data-set}{This should be unpacked a bit more.
+How long, where, what type of data?}, we show that there are differences in the \tom{impacts of different ant partners}{How many partners?} on \tom{various vital rates of the cacti across ontogeny}{This sounds a little nebulous, be more specific.}.
+Our results demonstrate the importance of evaluating a mutualism within a community context and \tom{suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti}{Nothing you have reported in the abstract so far really supports this statement.}.
 
-Each of the ant partners in this system are within the same guild, all offering defense from herbivores, with only one ant species interacting with each individual cactus in a given period of time.
-Though these ants may offer very similar rewards, they likely differ in some way.
-Two possible ways they could differ are the fitness boost they offer the focal species or the period of ontogeny in which they offer the greatest fitness boost.
-Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny.
+\tom{Each of the ant partners}{Again, abstract should briefly state how many and which.} in this system are within the same guild, all offering defense from herbivores, \tom{with only one ant species interacting with each individual cactus in a given period of time}{Redundant with statement above.}.
+Though these ants may offer very similar rewards, \tom{they likely differ in some way}{Vague}. 
+Two possible ways they could differ are the \tom{fitness boost}{Vague and too colloquial.} they offer the focal species or the \tom{period of ontogeny}{We have size structure data but we are not really dealing with ontogeny per se, so I think it is safer to discuss things in terms of size.} in which they offer the greatest fitness boost.
+\tom{Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny.
 Our results demonstrate the importance of evaluating a mutualism within a community context and suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti.
+}{I am a little confused - maybe these two paragraphs are two different versions of the abstract?}
+\tom{}{I know this was just a first stab at the abstract, and written when you don't yet have your final results.
+But eventually this will need to be tightened up and edited to better describe the conceptual motivation, what was done, and what was found.}
 
@@
 \subsection*{Context}
+\tom{}{The Intro section should generally not have subsections.}
+Mutualisms are species interactions in which all involved participants benefit.
+They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences, \tom{yet they are still poorly understood}{What, specifically, is poorly understood?}.
+\tom{They form a complex foundation to many natural systems}{Not sure what this means} \tom{which involve both costs and benefits, despite their nature of being beneficial.}{I get what you are trying to say here but the phrasing is a little awkward, could be sharpened.}
 
-Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences, yet they are still poorly understood.
-They form a complex foundation to many natural systems which involve both costs and benefits, despite their nature of being beneficial.
+\tom{There are benefits of mutualistic interactions, but there are costs as well}{redundant with last sentence of previous paragraph}, some direct and some indirect.
+\tom{}{This topic sentence does not effectively build on the multi-species emphasis of the previous paragraph.}
+Directly, mutualistic partners can sometimes become parasitic, causing harm to the focal mutualist\cite{Bronstein2001a}. 
+\tom{More indirectly, these interactions require energy.}{Awkward sentence, and I'm not sure why investment is reward is an indirect cost.} 
+\tom{Within a mutualism, each participant offers rewards and in return receives benefits from their partners.}{No new information content in this sentence.} 
+These rewards come in many forms (nectar, protection, habitat, etc.), all of which require some investment \cite{Bronstein2001}. 
+These investments can be seen as the costs of mutualisms, they are traded for the rewards provided by a partner. 
+\tom{Because of these trade-offs}{trade-offs between what and what?}, there are both costs and benefits to a mutualistic interaction\cite{Bronstein2001,Bronstein2001a}, \tom{particularly as the mutualism becomes more complex in the context of its environment}{I don't know what this means.}.\tom{}{Having made it to the end of this paragraph I can't really find any new information here.
+You are covering very basic aspects of mutualism that were already described in the first paragraph.
+I would get to the question/problem at the core of this paper more quickly.}
 
-There are benefits of mutualistic interactions, but there are costs as well, some direct and some indirect.
-Directly, mutualistic partners can sometimes become parasitic, causing harm to the focal mutualist\cite{Bronstein2001a}.
-More indirectly, these interactions require energy.
-Within a mutualism, each participant offers rewards and in return receives benefits from their partners.
-These rewards come in many forms, nectar, protection, habitat, etc., all of which require some investment from the individual\cite{Bronstein2001}.
-These investments can be seen as the costs of mutualisms, they are traded for the rewards provided by a partner.
-Because of these trade-offs, there are both costs and benefits to a mutualistic interaction\cite{Bronstein2001,Bronstein2001a}, particularly as the mutualism becomes more complex in the context of its environment.
-
-Mutualistic interactions can occur simply between two species, a pairwise mutualism, or between a group of species, a multi-species mutualism\cite{Stanton2013}.
-Pairwise mutualisms may provide benefits of efficiency in the context of mutualisms, but they are not a good investment due to temporal and spatial variation\cite{Waser1996}.
+Mutualistic interactions can occur simply between two species, a pairwise mutualism, or between a group of species, a multi-species mutualism\cite{Stanton2013}.\tom{}{Technically all mutualisms are pairwise (++) but they can be combined into many pairs.}
+Pairwise mutualisms may provide \tom{benefits of efficiency in the context of mutualisms}{I don't know what this means}, but \tom{they are not a good investment due to temporal and spatial variation}{Variation in what?}\cite{Waser1996}.
 Despite this, pairwise mutualisms are historically well studied, with most mutualism papers being focused on pairwise interactions. 
@@
 When many species are included, variation between partners can occur in functional benefit, lifespan, and competitive rankings, among other factors\cite{Stanton2013}. 
-This difference is because in addition to the expected direct benefits of interactions, they can also indirectly impact the focal mutualist through other partner interactions.
+This difference is because in addition to the expected direct benefits of interactions, \tom{they can also indirectly impact the focal mutualist through other partner interactions.
+}{Again, I would avoid too much discussion of interaction between partners.
+You don't need interactions between partners to have non-additive effects, and you are priming readers to expect something that the manuscript does not deliver.}
 One reason for this lack of predictability is that interactions between partners changes the relationship with the focal mutualist\cite{Afkhami2014}.  
-Often the partners interact with each other, either increasing their performances, with extra rewards, or decreasing them, often by fighting\cite{Boucher1982}.
+Often the partners interact with each other, either \tom{increasing their performances}{awkward}, with extra rewards, or decreasing them, often by fighting\cite{Boucher1982}. 
 This means predicting the rewards of a multi-species mutualism is very difficult.
 
-Multi-species mutualisms are not always additive: In theory if multiple partners can all form pairwise mutualisms with a focal species, in a multi-species mutualism one would expect more benefits for the focal partner than in any one pairwise scenario.
-The reality is much less predictable\cite{Stanton2013,Palmer2010,Song2020}.
-The rewards can be as simple as the sum of all pairwise rewards, aka.
-additivity, they can be increased beyond the sum of all pairwise rewards, synergy, or they can be reduced below the sum of pairwise rewards, sub-additivity.
-Specifically, it can often change the quality of benefits.
+Multi-species mutualisms are not always \tom{additive}{You have not introduced or defined what you mean by additive / non-additive.}: In theory if multiple partners can all form pairwise mutualisms with a focal species, in a multi-species mutualism \tom{one would expect more benefits for the focal partner than in any one pairwise scenario.
+}{The logic for this expectation is not sufficiently clear.}
+\tom{The reality is much less predictable\cite{Stanton2013,Palmer2010,Song2020}.}{While there is nothing grammatically wrong with sentences like this, they don't really provide any new information.
+Try to streamline your writing so that every word is working in service of communicating information.}
+The \tom{rewards}{Unclear what you mean by rewards.
+That word can describe something like nectar or protection, but I think you are referring to fitness benefits} can be as simple as the sum of all pairwise rewards, aka.
+additivity, they can be increased beyond the sum of all pairwise rewards, synergy, or they can be reduced below the sum of pairwise rewards, sub-additivity\tom{}{Needs editing for grammar}.
+\tom{Specifically, it can often change the quality of benefits.
+}{Vague and confusing sentence.
+What is ``it''?}
 
-Sometimes a partner can confer benefits on the focal mutualist in a pairwise interaction but be a neutral or even antagonistic interactor in a multi-species interaction as partners can vary in quality\cite{Afkhami2014}.
-Some mutualists are of so low a quality that they are known as freeloaders or cheaters\cite{Song2020,West2007,Frederickson2013} because they return very little benefit, possibly going as far as to be a cost to the focal mutualist in exchange for the benefits they receive.
-This change in quality of partners is often due to competition between the partners\cite{Amarasekare2003}.
-Some partners reduce in quality so drastically that they become antagonistic towards the focal mutualist in these relationships\cite{Afkhami2014,Frederickson2013,Bronstein1994}.
-Despite the potential to reduce fitness of the focal mutualist they often remain a partner within the mutualism\cite{Jones2015}.
+Sometimes a partner can confer benefits on the focal mutualist in a pairwise interaction but be a neutral or even antagonistic interactor in a multi-species interaction as partners can vary in quality\cite{Afkhami2014}.\tom{}{This topic sentence feels like it is revisiting points from the previous paragraph.
+Try to use topic sentence to demarcate narrative movement.}
+Some mutualists are of so low a quality that they are known as freeloaders or cheaters\cite{Song2020,West2007,Frederickson2013} because they return very little benefit, possibly going as far as to be costly to the focal mutualist in exchange for the benefits they receive.
+This change in quality of partners is often due to competition between the partners\cite{Amarasekare2003}.\tom{}{I'm confused here - what change in quality of partners?
+And how could the existence of freeloaders be driven by competition?}
+Some partners \tom{reduce in quality}{I don't follow.
+How does a partner reduce in quality?
+What is the reference point for the reduction?} so drastically that they become antagonistic towards the focal mutualist in these relationships\cite{Afkhami2014,Frederickson2013,Bronstein1994}.\tom{}{This is restating points made earlier that some partners can be costly.}
+\tom{Despite the potential to reduce fitness of the focal mutualist they often remain a partner within the mutualism\cite{Jones2015}.}{Not sure what point you are trying to make with this sentence.
+Are you trying to suggest something about partner choice?
+It is not surprising that cheaters associate with their ostensible mutualists, because that is what cheaters are selected to do.}
 
-There are several underlying mechanisms which promote partner diversity that may shed some light on this; specifically, sampling effect,portfolio effect, and complementarity, three branches of biodiversity ecosystem function theory\cite{Batstone2018,Hooper2005}.
+There are several underlying mechanisms which \tom{promote partner diversity}{These things do not promote partner diversity.
+They are mechanisms that could account for benefits of partner diversity, if such benefits exist in the first place.
+That's an important distinction.
+Therefore if you want to discussed these things then that discussion should be preceded by evidence that partner diversity is or can be beneficial.} that may \tom{shed some light on this}{Shed light on what?
+What's in the dark?}; specifically, sampling effect,portfolio effect, and complementarity, \tom{three branches of biodiversity ecosystem function theory}{This is good, but I suggest you be more explicit in how this body of theory relates to mutualism, rather than mentioning it in an off-handed way.}\cite{Batstone2018,Hooper2005}.
 
-Sampling effect states that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the most beneficial one\cite{Afkhami2014}.
+Sampling effect \tom{states}{The effect does not state anything.} that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the most beneficial one\cite{Afkhami2014}.
 Particularly, this means that increased diversity of partners means increased probability that some of the partners will be highly beneficial. 
-The fitness of a multi-species mutualistic population experiencing sampling effect will be equal to the fitness of the same population if they interacted only with their strongest partner.
+The fitness of a \tom{multi-species mutualistic population}{awkward} experiencing sampling effect will be equal to the fitness of the same population if they interacted only with their strongest partner.
 
 Complementarity states that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in a broader benefit\cite{Winfree2020}. 
-This means that in populations where there is sufficient interaction with the diverse partners there is likely to be increased rewards and fitness.
-The fitness of a multi-species mutualistic population experiencing complementarity is likely to be higher than that of any pairwise mutualistic population.
+This means that in populations where there is \tom{sufficient interaction}{What do you mean by ``sufficient''?} with the diverse partners there is likely to be increased rewards and \tom{fitness}{Whose fitness?}.
+The fitness of a multi-species mutualistic population \tom{experiencing complementarity}{I don't think sampling effect, complementarity, etc.
+are ``experienced''.} is likely to be higher than that of any pairwise \tom{mutualistic population}{What is a mutualistic population?}.
 
-Portfolio effect states that each partner is the best under a different set of conditions, meaning across all conditions, meaning there is likely to be an increase in potential benefits\cite{Winfree2020}.
+Portfolio effect states that each partner is the best under a \tom{different set of conditions}{I could be wrong but I think the standard use of portfolio effect specifically refers to temporal variation.}, meaning across all conditions, meaning there is likely to be an \tom{increase in potential benefits}{Increase relative to what?
+Benefits to who?
+What are they potential?}\cite{Winfree2020}.
 Thus, in populations experiencing highly diverse conditions, diverse partners are likely to increase the rewards and fitness of the focal mutualist. 
-The fitness of a multi-species mutualistic population experiencing portfolio effect is likely to have a different strongest partner under different conditions, such as year to year, and will experience similar fitness as they would in a pairwise mutualism with their best partner under any condition.
+The fitness of a multi-species mutualistic population experiencing portfolio effect is likely to have a different strongest partner under different conditions, such as year to year, \tom{and will experience similar fitness as they would in a pairwise mutualism with their best partner under any condition}{I don't follow this last part.}.\tom{}{You do a good job describing the mechanisms for positive diversity effects.
+I think you could rely more heavily on literature for these mechanisms, which is substantial.}
 
-These biodiversity ecosystem functions are three which can be used to explain the diversity of mutualistic relationships in nature\cite{Afkhami2014}.
-If a system benefits from the existence of multiple partners, one of these functions may be at play.
-Within the section of mutualistic interaction which have multiple partners, there are many different classifications.
-One type of mutualism that is historically well documented is the interaction between ants and plants\cite{Boege2005,Barton2010}.
-These interactions can take many forms, including pollination, defense, dispersal, etc\cite{Bronstein2006}.
+These \tom{biodiversity ecosystem functions}{This is not what they are.
+They are mechanisms for positive effects of diversity that are most closely associated with the biodiversity-ecosystem function literature, which asks how changes in biodiversity affect ecosystem function} are three which \tom{can be used to explain the diversity of mutualistic relationships in nature}{This is not correct, and it's an important point.
+These mechanisms cannot explain diversity in mutualisms.
+They \emph{could} explain positive effects of partner diversity on a focal species, if any such effects occur}\cite{Afkhami2014}.
+If a \tom{system benefits}{I don't think a ``system'' can benefit from anything.} from the existence of multiple partners, one of these functions may be at play. 
+\tom{Within the section of mutualistic interaction which have multiple partners, there are many different classifications.}{Awkward and vague - what do you mean by ``classification''?} One type of mutualism that is historically well documented is the interaction between ants and plants\cite{Boege2005,Barton2010}.
+These \tom{interactions can take many forms}{I know what you mean, but use more precise language.}, including \tom{pollination}{ant pollination is rare}, defense, dispersal, \tom{etc}{are these any others besides these three?}\cite{Bronstein2006}.
 This paper focuses on defensive ant-plant mutualisms.
-These often involve long-lived plants providing extra-floral nectar and/or habitats to ants which in turn defend them from herbivores\cite{Bronstein1998}.
-These are very common interactions and have provided many opportunities for empirical studies of mutualisms.
-However, many of these studies have been pairwise rather than investigating the impacts that multiple partners may have on the costs and benefits of a system.
+These often involve \tom{long-lived}{not necessarily} plants providing extra-floral nectar and/or \tom{habitats}{better to call this ``housing''} to ants which in turn defend them from herbivores\cite{Bronstein1998}.
+\tom{These are very common interactions and have provided many opportunities for empirical studies of mutualisms.}{No information here.} \tom{However, many of these studies have been pairwise rather than investigating the impacts that multiple partners may have on the costs and benefits of a system.
+}{Hard to follow what gap in the literature you are trying to identify here.
+There are in fact many studies of ant-plant mutualism that include multiple partners.
+I think the new thing that you are contributing is asking how diversity in the ant defender guild affects the overall benefits of mutualism for the plant partner.}
 
-This study focuses on the long lived, faculative \textit{Cylindriopuntia imbricata} (Cholla cacti) and its ant mutualists.
-The cacti interact with several ant partners, \textit{Crematogaster opuntiae} (\textit{Crem.}) and \textit{Liometopum apiculatum} (\textit{Liom.}) are the most common partners though there are a number of other rarer ant partners, which visit these cacti in smaller population sizes.
-The ant partners all provide defense to the cacti from predators in return for exclusive interaction with the plant to take the extra-floral nectar, meaning they are considered to be within one guild.
-Predators of the cholla include insect herbivores and seed predators, which influence the reproductive output of the cacti.
-Though many multi-species mutualisms result in both costs and benefits, it has been shown that the relationship between these cacti and ants does not result in any major costs for the cacti\cite{Miller2007}.
-However, there are still some small costs which could influence the optimal set of partners.
-This is part of what we will be analyzing in this paper.
+This study focuses on the long lived, \tom{faculative}{this is not an adjective that can be applied to a species} \textit{Cylindriopuntia imbricata} (\tom{Cholla cacti}{There are many species of cholla cacti.
+The common name of this on is tree cholla.}) and its ant mutualists.
+\tom{The}{weird grammar in this sentence} cacti interact with several ant partners, \textit{Crematogaster opuntiae} (\textit{Crem.}) and \textit{Liometopum apiculatum} (\textit{Liom.}) are the most common partners though there are a number of other rarer ant partners, which visit these cacti \tom{in smaller population sizes}{We don't know anything about their population sizes.}.\tom{}{These are the ant partners at the SEV.
+But the plants have a very wide distribution and likely interact with other ant species elsewhere.
+So I would not make it sound like this is an exhaustive list of partners for the entire species.}
+The ant partners all provide defense to the cacti from \tom{predators}{insect herbivores and seed predators} in return for \tom{exclusive interaction}{explain} with the plant to take the extra-floral nectar\tom{, meaning they are considered to be within one guild}{awkward}.
+Predators of the cholla include insect herbivores and seed predators, which influence the \tom{reproductive output of the cacti}{also growth, survival, etc.}. 
+\tom{Though many multi-species mutualisms result in both costs and benefits, it has been shown that the relationship between these cacti and ants does not result in any major costs for the cacti\cite{Miller2007}.}{I don't know what you mean.
+EFN production is a pretty clear cost.}
+\tom{However, there are still some small costs which could influence the optimal set of partners.}{This is vague, and tbh I don't know what you mean when you talk about costs.} \tom{This is part of what we will be analyzing in this paper.}{Awkward and unnecessary.}
 
-Though all the ants associated with the Cholla are within one guild, they may not all be equal, meaning the impacts on the Cholla are complex. In this study, we will answer a number of questions: 
+Though all the ants associated with the \tom{Cholla}{common names are not capitalized} are within one guild, they may not all be equal, \tom{meaning the impacts on the Cholla are complex}{no information here}.
+In this study, we will answer a number of \tom{questions}{I think this is a good list of questions.
+Following up on our conversation, we will probably want to re-orient the language to be clear that we are essentially asking about one vs two partners, which is still technically a diversity question but we don't want to oversell the `diversity'.}:
 \begin{enumerate}
@@
 
-In order to answer these questions, I am constructing an Integral Projection Model (IPM) to calculate the fitness of the Cholla cacti under different conditions.
-I will be able to quantitatively estimate the fitness of the cholla population if they were to interact with any combination of ant partners, allowing me to determine the best combination of partners.
-Similar approaches have been used to assess the individual qualities of multiple partners\cite{Palmer2010} but not to determine the optimal combination of partners in a multi-species mutualism.
+\tom{In order to answer these questions, I am constructing an Integral Projection Model (IPM)}{This is only one piece of what you are doing.
+I think the value of the long-term data should be better emphasized.} to calculate the fitness of the Cholla cacti \tom{under different conditions}{What conditions?}.
+\tom{I will be able to quantitatively estimate}{Use plural pronouns and don't use future tense, this is not a proposal.} the fitness of the cholla population if they were to interact with any combination of ant partners, allowing me to determine the best combination of partners.
+Similar approaches have been used to assess the individual qualities of multiple partners\cite{Palmer2010} \tom{but not to determine the optimal combination of partners in a multi-species mutualism}{The Palmer paper and perhaps others have done this.
+I think the novel contribution of this paper needs to be better explained.}.
 
-We hypothesize that the cacti will experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism is the optimal outcome.
-We also hypothesize that complementarity and sampling effect will play roles in encouraging this diverse mutualism.
+We hypothesize that the cacti will experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism is the \tom{optimal outcome}{I would avoid mention of optimality, which has an evolutionary implication that we are not addressing here}.
+We also hypothesize that complementarity and sampling effect will \tom{play roles in encouraging this}{vague - sharpen this} \tom{diverse mutualism}{A mutualism cannot be `diverse'.}.
 

### 2022-06-17 Manuscript/manuscript_draft.tex
edited through end of methods

@@
 
-To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of Cholla cacti annually since 2004, this is still ongoing.
+To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of Cholla cacti annually since 2004. 
 This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico. 
@@
 
-The cholla cacti secrete extra-floral nectar (EFN) from glands on young vegetative segments and reproductive segments.
-This extra-floral nectar varies in chemical makeup across the lifespan of the cactus *** Miller ??***.
-At the Sevilleta, the cacti are visited by multiple species of ground-nesting ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and many other species which occur at small frequencies.
-These ants are attracted to the EFN and in return fight off the seed predators which also visit the cacti.
-These ants never co-occur on the same plant, instead competing for plant services.
-Over the lifetime of the cacti each plant interacts with multiple partners as they tend to switch out season to season.
+Cholla cacti secrete extra-floral nectar (EFN) from glands on young vegetative segments and reproductive segments. 
+\tom{This extra-floral nectar varies in chemical makeup across the lifespan of the cactus *** Miller ??***. }{This is pretty vague, so if you are going to say this you should add more detail.}
+At the Sevilleta, the cacti are visited by \tom{multiple species}{I think it would be better to say primarily two species plus other rare species, and giving the frequencies} of ground-nesting ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and \tom{many other species}{which ones?} which occur at \tom{small}{low} frequencies.
+These ants are attracted to the EFN and in return fight off the \tom{seed predators}{which ones?} \tom{which also visit the cacti}{An example of words that make sentences clunky without adding any information}.
+These ants never co-occur on the same plant, \tom{instead competing for plant services}{This is a hypothesis, not something that has been demonstrated.}. 
+\tom{Over the lifetime of the cacti each plant interacts with multiple partners as they tend to switch out season to season.}{I would guess that most plants do not have much switching.
+I think you also need to talk about vacancy and the start of EFN production at larger reproductive sizes, more detail on transitions in ant state between vs within years.
+All of this should be in a separate subsection about natural history.}
 
-There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments.
+\tom{There}{This paragraph should also be in a natural history subsection and obviously needs more detail.
+See my earlier papers and references therein and let me know if you have specific questions.} are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments.
 There are a number of insects which occur commonly on these cacti at the Sevilleta. 
 These include the weevil, the NP, and the *** (I think there is a moth as well?). 
-The weevil feeds on ***.
+The \tom{weevil}{what species?} feeds on ***. 
 The NP feeds on ***. 
-Finally, the moth??
-feeds on ***.
-All of these predators can have significant impacts on the fitness of both individual cacti and the population.
-*** Miller 2009***.
+Finally, the moth?? feeds on ***. 
+All of these predators can have significant impacts on the fitness of both individual cacti and the population. *** Miller 2009***.
 
-The data collected on these cacti are from an 18-year data-set.
+\tom{The}{This should be a subsection for data collection, with significantly more information than what you have here so that the methods could be reproduced.} data collected on these cacti are from an \tom{18-year}{Which years?} data-set.
 There are 8 30 $\times$ 30 meter plots from which this data is taken. 
-Each plot is surveyed, and each plant within each plot is tagged and logged each year.
-For every plant we take data on their survival, size, ant state, herbivore presence, flowering data, and seed data.
-This allows us to track the progression of individual plants across many years.
+Each plot is \tom{surveyed}{what does this mean?}, and each plant within each plot is tagged and \tom{logged}{what does this mean} each year. 
+For every plant we take data on their survival, \tom{size}{how is size measured?}, \tom{ant state}{what is this?}, \tom{herbivore presence}{this makes it sounds like presence/absence which is not what we record}, \tom{flowering data}{what does this mean?}, and \tom{seed data}{what does this mean?}.
+\tom{This allows us to track the progression of individual plants across many years.}{No new information here.} 
 The length of this data-set
@@
 \subsection*{Statistical Modeling}
-To understand the impacts of multiple species interactions on the cacti populations, we constructed bayesian demographic models of many vital rates.
-With these models, we were able to ask (i) Do the impacts of ant presence on cacti fitness differ across ant species?
-(ii) What are the probabilities of an individual cactus interacting with each ant species?.
+To understand the impacts of multiple species interactions on the cacti populations, we constructed \tom{bayesian demographic models of many vital rates}{What are bayesian demographic models?
+WHat are the many vital rates?}.
+With these models, we were able to ask (i) Do the impacts of ant presence on \tom{cacti fitness}{these models are not measuring fitness} differ across ant species?
+(ii) What are the probabilities of an individual cactus interacting with each ant species?.\tom{}{I think it's good that you have these as separate questions but I think each one needs a lot more context and detail.}
 
-The vital rates of the cacti considered in this model include survival rate, growth rate, reproductive state, floral abortion rate, seed production, and germination.
-We fit generalized linear mixed effects models in a heirarchical Bayesian framework to quantify each of the vital rates.
-Each model takes the data from year \textit{t} and estimates the values in year \textit{t+1}.
-This allows us to predict a vital rate for any plant with a given set of characteristics, such as size and ant state.
+The vital rates of the cacti considered in this model include survival rate, growth rate, reproductive state, floral abortion rate, \tom{seed production, and germination}{You have not described where these data come from.}.
+\tom{We fit generalized linear mixed effects models in a heirarchical Bayesian framework to quantify each of the vital rates.}{Not a reproducible description of the methods.
+e.g., what distributions were used, what were the fixed and random effects, which models included ant effects, what were the priors, how many chain and iterations, and were convergence and fit assessed, etc?}
+\tom{Each model takes the data from year \textit{t} and estimates the values in year \textit{t+1}.}{This is true only for growth and survival.} 
+This allows us to predict a vital rate for any plant with a given set of characteristics, \tom{such as size and ant state}{Not such as, it is really just these two things.}. 
 \vspace{1cm}
@@
 This model was used to estimate the growth rate of the cactus population, effectively a quantitative measure of the fitness of the population. 
-The fitness was estimated under the range of all different conditions, no ants present, one ant species present, two ant species present, all ant species present.
-These allow us to compare the fitness of the cacti under no mutualistic interactions, pairwise mutualisms, and various multi-species mutualisms to show which case results in the highest fitness for the cacti.
+The fitness was estimated under the \tom{range of all different conditions, no ants present, one ant species present, two ant species present, all ant species present.
+These allow us to compare the fitness of the cacti under no mutualistic interactions, pairwise mutualisms, and various multi-species mutualisms to show which case results in the highest fitness for the cacti.}{This needs a more thorough explanation, but I would also do this later after the model has been introduced and explained.}
+
 Following previous studies, we modeled the life cycle of \textit{C.
-imbracata} using continuously size-structured plants,$n_i(x)$, and two discrete seed banks ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
+imbricata} using continuously size-structured plants,$n_i(x)$, and two discrete seed banks ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
 $$
@@
 
-The final equation gives the size of the \textit{C.
+The final equation gives the \tom{size}{not just population size but also composition in terms of size and ant state - this needs to be explained better} of the \textit{C.
 imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$.
@@
  \subsection*{Model analysis}
+
+\textbf{TOM STOPPED EDITING HERE}
 

### 2022-06-28 IPM Code/IPM Post.R
Update IPM Post.R

@@ -27,7 +27,7 @@ gxy(3,3,"vacant",params[m,])
 i = c("vacant","crem","liom","other")
 x = c(-1,-5,3,4)
 y = c(-1,-4,3,4)
-i = "vacant"
+i = "crem"
 x = 1
 y = 3
 g <- matrix(NA,ncol = length(i), nrow = 100)
@@ -525,8 +525,32 @@ bigmatrix.1 <- function(params,lower,upper,matsize,num_ants,i){
   Tmat[3:(n+2),3:(n+2)]<-t(outer(y,y,pxy,i,params))*h 
   # Put it all together
   IPMmat<-Fmat+Tmat  
-  return(IPMmat)
+  #return(IPMmat)
+  lambda <- Re(eigen(IPMmat)$values[1])
+  return(lambda)
 }
+n<-matsize
+L<-lower; U<-upper
+h<-(U-L)/n                   #Bin size
+b<-L+c(0:n)*h;               #Lower boundaries of bins 
+y<-0.5*(b[1:n]+b[2:(n+1)]);  #Bin midpoints
+i = c("liom")
+a <- matrix(NA,ncol = length(i), nrow = (Ndraws))
+b <- matrix(NA,ncol = length(i), nrow = (Ndraws))
+for(m in seq(1:nrow(params))){
+  for(n in seq(1:length(i))){
+    f[m,n] <- fx(x[n],i[n],params[m,])
+  }
+}
+f
+for(m in 1:10){
+  for(n in 1:length(i)){
+    a[m,n] <- fx(x[n],i[n],params[m,]) 
+    b[m,n] <- bigmatrix.1(params[m,],lower,upper,matsize,num_ants,i[n])
+  }
+}
+a
+b
 
 bigmatrix.1(params[1,],lower,upper,matsize,1,"vacant")
 
@@ -546,6 +570,12 @@ for(m in 1:nrow(params)){
     lambda[m,n] <- Re(eigen(big1[[m]])$values[1])
   }
 }
+for(m in seq(1:nrow(params))){
+  for(n in seq(1:length(i))){
+    f[m,n] <- fx(x[n],i[n],params[m,])
+  }
+}
+f
 big1[[2]]
 
 

### 2022-06-28 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

@@
 
-\tom{Understanding interspecific mutualisms is a challenge, because they are among the most widespread species interactions with diverse and dynamic consequences.}{This sentence is pretty vague and does not have much information content.}
+Mutuaisms are among the most widespread species interactions with diverse and dynamic consequences.
 Depending on environmental context the outcome of the interaction can range from parasitic to mutualistic. 
-Most studies focus on a pair of species interacting in a mutualism, \tom{however often this is a simplification of the reality}{awkward and unclear -- pairs are ``real'' after all}. 
-These mutualistic interactions \tom{become even more complex}{more complex than what?} when put into the context of a multi-species mutualism, where there are multiple partners interacting with the same focal species.
-The partners can directly impact the focal species through the interactions, \tom{and indirectly impact them through interactions with each other}{This is true but not something we are well poised to address with our data so might be risky to bring it up.}.
-The cactus \textit{Cylindriopuntia imbricata} forms \tom{symbioses}{technically it's not a symbiosis because the partners live independently} with multiple ant species, but each individual plant is able to \tom{interact}{This sentence does not sufficiently convey the nature of the interaction - the plants reward ants with extrafloral nectar in exchange for defense} with \tom{only one species at a time}{There is not enough context to know why this information is relevant here.
-Abstracts should be brutally efficient in communicating the essentials.}.
-Using \tom{a long-term data-set}{This should be unpacked a bit more.
-How long, where, what type of data?}, we show that there are differences in the \tom{impacts of different ant partners}{How many partners?} on \tom{various vital rates of the cacti across ontogeny}{This sounds a little nebulous, be more specific.}.
+Most studies focus on a pair of species interacting in a mutualism, however most observed mutualistic interactions include more than two species. 
+These are multi-species mutualisms, where there are multiple partners interacting with the same focal species. 
+\hl{The partners can directly impact the focal species through the interactions, \tom{and indirectly impact them through interactions with each other}{***This is true but not something we are well poised to address with our data so might be risky to bring it up.***}.
+}
+The cactus \textit{Cylindriopuntia imbricata} reward partners with extrafloral nectar in exchange for defense by various ant species (primarily \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, though there are many other ant species with smaller populations) from herbivores and seed predators.
+\hl{Using a 20-year collection of demographic data (growth, survival, reproduction, ant partners, herbivory, etc.) from these cacti in New Mexico, we estimate the fitness of the cacti populations with every combination of partners, allowing us to *** understand what accounts for partner diversity in this system *** understand if the current partner diversity of the system is associated with the optimal fitness *** and the impacts of these partners on different vital plant processes across their life history (such as partner impacts on reproduction, survival, or growth).
+}
+We found that different ant partners had different impacts on cacti vital rates suggesting there may be a different 'best' partner during pre-reprodctive life stages than during reproductive life stages, and that the highest cacti fitness is associated with the current partner set (all possible ant partners are available).
+Together, these results suggest that complementarity may account for the benefits that these plants experience with partner diversity. 
 Our results demonstrate the importance of evaluating a mutualism within a community context and \tom{suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti}{Nothing you have reported in the abstract so far really supports this statement.}.
@@
 \section*{Context/Introduction}
-\subsection*{Context}
-\tom{}{The Intro section should generally not have subsections.}
+BASELINE OF MUTUALISMS
 Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences, \tom{yet they are still poorly understood}{What, specifically, is poorly understood?}.
-\tom{They form a complex foundation to many natural systems}{Not sure what this means} \tom{which involve both costs and benefits, despite their nature of being beneficial.}{I get what you are trying to say here but the phrasing is a little awkward, could be sharpened.}
+They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences.
 
-\tom{There are benefits of mutualistic interactions, but there are costs as well}{redundant with last sentence of previous paragraph}, some direct and some indirect.
-\tom{}{This topic sentence does not effectively build on the multi-species emphasis of the previous paragraph.}
+PAIRWISE VS MULTISPECIES
+
+There are benefits of mutualistic interactions, but there are costs as well, some direct and some indirect. 
 Directly, mutualistic partners can sometimes become parasitic, causing harm to the focal mutualist\cite{Bronstein2001a}. 

### 2022-07-26 Manuscript/manuscript_draft.tex

@@
 They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences.
+The benefits from these interactions are called rewards, but they require investment of resources into the interaction {I'm honestly not sure if I need to explain this idea?}.
 
-PAIRWISE VS MULTISPECIES
+PARTNER DIVERSITY IN MUTUALISMS
+
+Historically, scientists focused on pairwise mutualisms *** citations ***, however, multi-species mutualisms are becoming better studied
 
@@
 \tom{More indirectly, these interactions require energy.}{Awkward sentence, and I'm not sure why investment is reward is an indirect cost.} 
-\tom{Within a mutualism, each participant offers rewards and in return receives benefits from their partners.}{No new information content in this sentence.} 
 These rewards come in many forms (nectar, protection, habitat, etc.), all of which require some investment \cite{Bronstein2001}. 

### 2022-08-03 Figures/Cholla_Analysis_Visualization_Vital_Rates.R
Update Cholla_Analysis_Visualization_Vital_Rates.R

@@ -562,6 +562,9 @@ filled.contour(x,y,other, xlab = "",ylab = "", main = "Other Tended",col = gray.
 ## Vacant
 filled.contour(x,y,crem, xlab = "",ylab = "", main = " Not Tended",col = gray.colors(24, start =1, end = 0))
 
+
+x <- seq(min(cactus$logsize_t, na.rm = T),max(cactus$logsize_t,na.rm = T), length = 25); # three columns
+y <- seq(min(cactus$logsize_t1, na.rm = T),max(cactus$logsize_t1,na.rm = T), length = 25); # five rows
 ## Contour plots
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
 png("grow_contour_lines.png")
@@ -575,6 +578,7 @@ contour(x,y,crem, nlevels = 7
         , xlim = c(2,10), ylim = c(0,10)
         , col = gray.colors(10, start =1, end = 0))
 lines(size_crem, y_crem_mean_grow, col = cremcol, lwd = 3)
+points(y_crem_subset_grow$logsize_t,y_crem_subset_grow$logsize_t1,pch=16,col= alpha("black", 0.4))
 ## Liom
 contour(x,y,liom, nlevels = 8
         , xlim = c(2,10), ylim = c(0,10)

### 2022-08-03 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

--- 2022-06-28 Manuscript/manuscript_draft.tex
@@
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
+\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
 
@@
 
-Mutuaisms are among the most widespread species interactions with diverse and dynamic consequences.
+Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
 Depending on environmental context the outcome of the interaction can range from parasitic to mutualistic. 
@@
 They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences.
+\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
+Do I need to include more basic mutualism background?
+I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.
 
-PAIRWISE VS MULTISPECIES
+Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied. 
+It has become apparent that partner diversity in mutualisms often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}. 
+This means the benefits can be greater than, less than, or equal to the sum of the benefits observed in each of the pairwise interactions with the focal mutualist that makes up the multi-species mutualism.
+This is true for many complicated reasons, including interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability\cite{stanton2013, Afkhami2014,Boucher1982}.
+Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and fighting, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
+For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020} which are more common.
+\ali{***** Biodiversity ecosystem function?
+}{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explainations of negative effects?
+Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?}
+Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
 
-There are benefits of mutualistic interactions, but there are costs as well, some direct and some indirect. 
-Directly, mutualistic partners can sometimes become parasitic, causing harm to the focal mutualist\cite{Bronstein2001a}. 
-\tom{More indirectly, these interactions require energy.}{Awkward sentence, and I'm not sure why investment is reward is an indirect cost.} 
-\tom{Within a mutualism, each participant offers rewards and in return receives benefits from their partners.}{No new information content in this sentence.} 
-These rewards come in many forms (nectar, protection, habitat, etc.), all of which require some investment \cite{Bronstein2001}. 
-These investments can be seen as the costs of mutualisms, they are traded for the rewards provided by a partner. 
-\tom{Because of these trade-offs}{trade-offs between what and what?}, there are both costs and benefits to a mutualistic interaction\cite{Bronstein2001,Bronstein2001a}, \tom{particularly as the mutualism becomes more complex in the context of its environment}{I don't know what this means.}.\tom{}{Having made it to the end of this paragraph I can't really find any new information here.
-You are covering very basic aspects of mutualism that were already described in the first paragraph.
-I would get to the question/problem at the core of this paper more quickly.}
+The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \{Stanton2003}.
+Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year. 
+When partner quality remains consistent across time, meaning one partner is always better or worse than another, the effects of partner diversity on the focal mutualist can be explained by a mechanism called sampling effect.
+This is the concept that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
+When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner.
+In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, such as resource shortages and niche shifts.
+The effects of partner diversity in these systems can be explained by the portfolio effect, the idea that each partner is the ``best" under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}.
+This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
+Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
+The effects of partner functional diversity can be explained by the mechanism complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
+This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\{Stanton2003}.
+Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\{Batstone2018}.
+\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.}
 
-Mutualistic interactions can occur simply between two species, a pairwise mutualism, or between a group of species, a multi-species mutualism\cite{Stanton2013}.\tom{}{Technically all mutualisms are pairwise (++) but they can be combined into many pairs.}
-Pairwise mutualisms may provide \tom{benefits of efficiency in the context of mutualisms}{I don't know what this means}, but \tom{they are not a good investment due to temporal and spatial variation}{Variation in what?}\cite{Waser1996}.
-Despite this, pairwise mutualisms are historically well studied, with most mutualism papers being focused on pairwise interactions. 
-These studies are often used to predict outcomes of mutualistic interaction, which may underestimate the combined impacts of multiple partners on the focal mutualist\cite{Stanton2013,Palmer2010}. 
-Pairwise interactions do not always indicate what the net interaction will be within a multi-species context\cite{Chamberlain2014,Song2020}. 
-When many species are included, variation between partners can occur in functional benefit, lifespan, and competitive rankings, among other factors\cite{Stanton2013}. 
-This difference is because in addition to the expected direct benefits of interactions, \tom{they can also indirectly impact the focal mutualist through other partner interactions.
-}{Again, I would avoid too much discussion of interaction between partners.
-You don't need interactions between partners to have non-additive effects, and you are priming readers to expect something that the manuscript does not deliver.}
-One reason for this lack of predictability is that interactions between partners changes the relationship with the focal mutualist\cite{Afkhami2014}.  
-Often the partners interact with each other, either \tom{increasing their performances}{awkward}, with extra rewards, or decreasing them, often by fighting\cite{Boucher1982}. 
-This means predicting the rewards of a multi-species mutualism is very difficult.
+Multi-species mutualisms are a huge group of interactions which vary considerably. 
+While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.
+These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
+While these interactions have been well studied\cite{} in the literature, few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.
 
-Multi-species mutualisms are not always \tom{additive}{You have not introduced or defined what you mean by additive / non-additive.}: In theory if multiple partners can all form pairwise mutualisms with a focal species, in a multi-species mutualism \tom{one would expect more benefits for the focal partner than in any one pairwise scenario.
-}{The logic for this expectation is not sufficiently clear.}
-\tom{The reality is much less predictable\cite{Stanton2013,Palmer2010,Song2020}.}{While there is nothing grammatically wrong with sentences like this, they don't really provide any new information.
-Try to streamline your writing so that every word is working in service of communicating information.}
-The \tom{rewards}{Unclear what you mean by rewards.
-That word can describe something like nectar or protection, but I think you are referring to fitness benefits} can be as simple as the sum of all pairwise rewards, aka.
-additivity, they can be increased beyond the sum of all pairwise rewards, synergy, or they can be reduced below the sum of pairwise rewards, sub-additivity\tom{}{Needs editing for grammar}.
-\tom{Specifically, it can often change the quality of benefits.
-}{Vague and confusing sentence.
-What is ``it''?}
+This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
+Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.
+These cacti flower in the spring throughout the summer, their reproductive season, outside of which they are not tended by ants. 
+During the reproductive season these plants experience increased visitation by seed predators and herbivores, which influence their reproductive outputs, survival, growth, etc.
+There are a number of observed ant partners which interact with the cacti in this part of their range, including \textit{Crematogaster opuntiae} (\textit{Crem.}), \textit{Liometopum apiculatum} (\textit{Liom.}), \textit{Forelius pruinosus}, an unidentified \textit{Camponotus} species, an unidentified \textit{Phenogaster} species, among others.
+The ants which most commonly interact with the cacti are \textit{Liom.} and \textit{Crem.}, while others are seen with low enough frequency that for the purpose of this paper we refer to them as ``other" ants.
+These ant partners are all in the same guild, meaning they all provide denfense to the cacti from insect herbivores and seed predators in return for extrafloral nectar (EFN).
+The cacti in this mutualism are each tended by one species at a time, for their entire reproductive season with little turnover until the winter between reproductive seasons. 
+Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
+This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.
 
-Sometimes a partner can confer benefits on the focal mutualist in a pairwise interaction but be a neutral or even antagonistic interactor in a multi-species interaction as partners can vary in quality\cite{Afkhami2014}.\tom{}{This topic sentence feels like it is revisiting points from the previous paragraph.
-Try to use topic sentence to demarcate narrative movement.}
-Some mutualists are of so low a quality that they are known as freeloaders or cheaters\cite{Song2020,West2007,Frederickson2013} because they return very little benefit, possibly going as far as to be costly to the focal mutualist in exchange for the benefits they receive.
-This change in quality of partners is often due to competition between the partners\cite{Amarasekare2003}.\tom{}{I'm confused here - what change in quality of partners?
-And how could the existence of freeloaders be driven by competition?}
-Some partners \tom{reduce in quality}{I don't follow.
-How does a partner reduce in quality?
-What is the reference point for the reduction?} so drastically that they become antagonistic towards the focal mutualist in these relationships\cite{Afkhami2014,Frederickson2013,Bronstein1994}.\tom{}{This is restating points made earlier that some partners can be costly.}
-\tom{Despite the potential to reduce fitness of the focal mutualist they often remain a partner within the mutualism\cite{Jones2015}.}{Not sure what point you are trying to make with this sentence.
-Are you trying to suggest something about partner choice?
-It is not surprising that cheaters associate with their ostensible mutualists, because that is what cheaters are selected to do.}
-
-There are several underlying mechanisms which \tom{promote partner diversity}{These things do not promote partner diversity.
-They are mechanisms that could account for benefits of partner diversity, if such benefits exist in the first place.
-That's an important distinction.
-Therefore if you want to discussed these things then that discussion should be preceded by evidence that partner diversity is or can be beneficial.} that may \tom{shed some light on this}{Shed light on what?
-What's in the dark?}; specifically, sampling effect,portfolio effect, and complementarity, \tom{three branches of biodiversity ecosystem function theory}{This is good, but I suggest you be more explicit in how this body of theory relates to mutualism, rather than mentioning it in an off-handed way.}\cite{Batstone2018,Hooper2005}.
-
-Sampling effect \tom{states}{The effect does not state anything.} that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the most beneficial one\cite{Afkhami2014}.
-Particularly, this means that increased diversity of partners means increased probability that some of the partners will be highly beneficial. 
-The fitness of a \tom{multi-species mutualistic population}{awkward} experiencing sampling effect will be equal to the fitness of the same population if they interacted only with their strongest partner.
-
-Complementarity states that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in a broader benefit\cite{Winfree2020}. 
-This means that in populations where there is \tom{sufficient interaction}{What do you mean by ``sufficient''?} with the diverse partners there is likely to be increased rewards and \tom{fitness}{Whose fitness?}.
-The fitness of a multi-species mutualistic population \tom{experiencing complementarity}{I don't think sampling effect, complementarity, etc.
-are ``experienced''.} is likely to be higher than that of any pairwise \tom{mutualistic population}{What is a mutualistic population?}.
-
-Portfolio effect states that each partner is the best under a \tom{different set of conditions}{I could be wrong but I think the standard use of portfolio effect specifically refers to temporal variation.}, meaning across all conditions, meaning there is likely to be an \tom{increase in potential benefits}{Increase relative to what?
-Benefits to who?
-What are they potential?}\cite{Winfree2020}.
-Thus, in populations experiencing highly diverse conditions, diverse partners are likely to increase the rewards and fitness of the focal mutualist. 
-The fitness of a multi-species mutualistic population experiencing portfolio effect is likely to have a different strongest partner under different conditions, such as year to year, \tom{and will experience similar fitness as they would in a pairwise mutualism with their best partner under any condition}{I don't follow this last part.}.\tom{}{You do a good job describing the mechanisms for positive diversity effects.
-I think you could rely more heavily on literature for these mechanisms, which is substantial.}
-
-These \tom{biodiversity ecosystem functions}{This is not what they are.
-They are mechanisms for positive effects of diversity that are most closely associated with the biodiversity-ecosystem function literature, which asks how changes in biodiversity affect ecosystem function} are three which \tom{can be used to explain the diversity of mutualistic relationships in nature}{This is not correct, and it's an important point.
-These mechanisms cannot explain diversity in mutualisms.
-They \emph{could} explain positive effects of partner diversity on a focal species, if any such effects occur}\cite{Afkhami2014}.
-If a \tom{system benefits}{I don't think a ``system'' can benefit from anything.} from the existence of multiple partners, one of these functions may be at play. 
-\tom{Within the section of mutualistic interaction which have multiple partners, there are many different classifications.}{Awkward and vague - what do you mean by ``classification''?} One type of mutualism that is historically well documented is the interaction between ants and plants\cite{Boege2005,Barton2010}.
-These \tom{interactions can take many forms}{I know what you mean, but use more precise language.}, including \tom{pollination}{ant pollination is rare}, defense, dispersal, \tom{etc}{are these any others besides these three?}\cite{Bronstein2006}.
-This paper focuses on defensive ant-plant mutualisms.
-These often involve \tom{long-lived}{not necessarily} plants providing extra-floral nectar and/or \tom{habitats}{better to call this ``housing''} to ants which in turn defend them from herbivores\cite{Bronstein1998}.
-\tom{These are very common interactions and have provided many opportunities for empirical studies of mutualisms.}{No information here.} \tom{However, many of these studies have been pairwise rather than investigating the impacts that multiple partners may have on the costs and benefits of a system.
-}{Hard to follow what gap in the literature you are trying to identify here.
-There are in fact many studies of ant-plant mutualism that include multiple partners.
-I think the new thing that you are contributing is asking how diversity in the ant defender guild affects the overall benefits of mutualism for the plant partner.}
-
-This study focuses on the long lived, \tom{faculative}{this is not an adjective that can be applied to a species} \textit{Cylindriopuntia imbricata} (\tom{Cholla cacti}{There are many species of cholla cacti.
-The common name of this on is tree cholla.}) and its ant mutualists.
-\tom{The}{weird grammar in this sentence} cacti interact with several ant partners, \textit{Crematogaster opuntiae} (\textit{Crem.}) and \textit{Liometopum apiculatum} (\textit{Liom.}) are the most common partners though there are a number of other rarer ant partners, which visit these cacti \tom{in smaller population sizes}{We don't know anything about their population sizes.}.\tom{}{These are the ant partners at the SEV.
-But the plants have a very wide distribution and likely interact with other ant species elsewhere.
-So I would not make it sound like this is an exhaustive list of partners for the entire species.}
-The ant partners all provide defense to the cacti from \tom{predators}{insect herbivores and seed predators} in return for \tom{exclusive interaction}{explain} with the plant to take the extra-floral nectar\tom{, meaning they are considered to be within one guild}{awkward}.
-Predators of the cholla include insect herbivores and seed predators, which influence the \tom{reproductive output of the cacti}{also growth, survival, etc.}. 
-\tom{Though many multi-species mutualisms result in both costs and benefits, it has been shown that the relationship between these cacti and ants does not result in any major costs for the cacti\cite{Miller2007}.}{I don't know what you mean.
-EFN production is a pretty clear cost.}
-\tom{However, there are still some small costs which could influence the optimal set of partners.}{This is vague, and tbh I don't know what you mean when you talk about costs.} \tom{This is part of what we will be analyzing in this paper.}{Awkward and unnecessary.}
-
-Though all the ants associated with the \tom{Cholla}{common names are not capitalized} are within one guild, they may not all be equal, \tom{meaning the impacts on the Cholla are complex}{no information here}.
-In this study, we will answer a number of \tom{questions}{I think this is a good list of questions.
-Following up on our conversation, we will probably want to re-orient the language to be clear that we are essentially asking about one vs two partners, which is still technically a diversity question but we don't want to oversell the `diversity'.}:
+In this paper we will answer a number of questions about the demographic effects of partner diversity in multi-species mutualisms:
 \begin{enumerate}
-	\item Do the Cholla cacti benefit from the multi-species mutualism with multiple ant partners?
-	\item What combination of, or individual, ant(s) result in the highest fitness for the Cholla cacti?
-	\item Is there evidence of sampling effect, complementarity, or portfolio effect to encourage the current multi-species state of the mutualism?
+  \item{What are the contrasting demographic effects of multiple partners?}
+  \item{What combination of ant partners is associated with the highest fitness for the tree cholla?}
+  item{What mechanisms explain the effects of multiple ant partners on the tree cholla?}
 \end{enumerate}
 
-\tom{In order to answer these questions, I am constructing an Integral Projection Model (IPM)}{This is only one piece of what you are doing.
-I think the value of the long-term data should be better emphasized.} to calculate the fitness of the Cholla cacti \tom{under different conditions}{What conditions?}.
-\tom{I will be able to quantitatively estimate}{Use plural pronouns and don't use future tense, this is not a proposal.} the fitness of the cholla population if they were to interact with any combination of ant partners, allowing me to determine the best combination of partners.
-Similar approaches have been used to assess the individual qualities of multiple partners\cite{Palmer2010} \tom{but not to determine the optimal combination of partners in a multi-species mutualism}{The Palmer paper and perhaps others have done this.
-I think the novel contribution of this paper needs to be better explained.}.
+To answer these questions we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
+We analyzed the impacts of partner diversity on each individual vital rate (survivial, growth, floral-viability, etc.) with bayesian models, which we used to parameterize an Integral Projection Model (IPM).
+With the IPM we moved beyond what has been done in previous studies\cite{Palmer2010} to identify which mechanisms explain the effects of partner diversity on the tree cholla.
 
-We hypothesize that the cacti will experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism is the \tom{optimal outcome}{I would avoid mention of optimality, which has an evolutionary implication that we are not addressing here}.
-We also hypothesize that complementarity and sampling effect will \tom{play roles in encouraging this}{vague - sharpen this} \tom{diverse mutualism}{A mutualism cannot be `diverse'.}.
-
+We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
+We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.
 \section*{Methods}

### 2022-10-18 Manuscript/manuscript_draft_ALI.tex
Manuscript work

\documentclass[12pt,a4paper]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{multirow}
\usepackage{multicol}
\setlength{\columnsep}{1.5cm}
\usepackage{setspace}
\usepackage{url}
\usepackage{array}
\usepackage{xfrac}
\usepackage{mwe}
\usepackage{tikz}
\usepackage{lmodern}
\usepackage{flowchart}
\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{color,soul}
%\usepackage[numbers,sort&compress]{natbib}
\usepackage{hyperref}
%\usepackage{cite}
\usepackage{biblatex}
\addbibresource{References.bib}
\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}

\title{Thank You, Next}
\author{Ali M. Campbell\\
	Tom E.X. Miller}
\begin{document}
	\maketitle

\section*{Abstract}

Mutualisms are among the most widespread species interactions with diverse and dynamic consequences . Mutualisms are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity. Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies. It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms, portfolio effect, complementarity, and sampling effect. In this paper, I will focus on defensive ant-plant mutualisms. These involve plants which provide extra-floral nectar and/or “housing” to ants which in turn defend them from herbivores. While these interactions have been well studied in the literature, few have considered how diversity within partner guilds affect the overall benefits of mutualism for the plant partner. 
I use the plant, Cylindropuntia imbracata (tree cholla), and ant, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), and more, multispecies mutualism in which the cacti provide extrafloral nectar in exchange for defense from various herbivores and seed predators. I used 18 years of data collected from demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, ant partner, and herbivory for 8 30x30 m plots, in New Mexico. With this data I parameterize a series of Bayesian generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. I found that different ant partners did have different impacts on the vital rates of the tree cholla. Specifically, Crem. tended plants had advantages in both growth and survival when small, and Liom. tended plants had advantages in colonizing partners as well as floral viability. With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest fitness, and which of the above mechanisms may explain the effects of partner diversity in this system. I found that the real-life scenario (all possible ants are present) lead to the highest fitness for the tree cholla, indicating that partner diversity is beneficial in this system. It also shows that complementarity is at play in this system, meaning different partners offer different benefits leading to synergistic benefits for the tree cholla.

\section*{Context/Introduction}
BASELINE OF MUTUALISMS
Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences. 
\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea? Do I need to include more basic mutualism background? I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.

Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied. 
It has become apparent that partner diversity in mutualisms often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}. 
This means the benefits can be greater than, less than, or equal to the sum of the benefits observed in each of the pairwise interactions with the focal mutualist that makes up the multi-species mutualism. 
This is true for many complicated reasons, including interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability\cite{stanton2013, Afkhami2014,Boucher1982}.
Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and fighting, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020} which are more common.
\ali{***** Biodiversity ecosystem function? }{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explanations of negative effects? Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?} 
Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.

The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \{Stanton2003}.
Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year. 
When partner quality remains consistent across time, meaning one partner is always better or worse than another, the effects of partner diversity on the focal mutualist can be explained by a mechanism called sampling effect.
This is the concept that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner. 
In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, such as resource shortages and niche shifts.
The effects of partner diversity in these systems can be explained by the portfolio effect, the idea that each partner is the ``best" under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}. 
This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
The effects of partner functional diversity can be explained by the mechanism complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\{Stanton2003}. 
Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\{Batstone2018}. 
\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.}

Multi-species mutualisms are a huge group of interactions which vary considerably. 
While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.
These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
While these interactions have been well studied\cite{} in the literature, few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.

This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.

These cacti flower in the spring throughout the summer (May to August), their reproductive season, outside of which they are not tended by ants. 
During the reproductive season these plants experience increased visitation by seed predators and herbivores, which influence their reproductive outputs, survival, growth, etc.
There are a number of observed ant partners which interact with the cacti in this part of their range, including \textit{Crematogaster opuntiae} (\textit{Crem.}), \textit{Liometopum apiculatum} (\textit{Liom.}), \textit{Forelius pruinosus}, an unidentified \textit{Camponotus} species, an unidentified \textit{Phenogaster} species, among others. 
The ants which most commonly interact with the cacti are \textit{Liom.} and \textit{Crem.}, while others are seen with low enough frequency that for the purpose of this paper we refer to them as ``other" ants. 
These ant partners are all in the same guild, meaning they all provide denfense to the cacti from insect herbivores and seed predators in return for extrafloral nectar (EFN).
The cacti in this mutualism are each tended by one species at a time, for their entire reproductive season with little turnover until the winter between reproductive seasons. 
Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.

\subsection{Questions}
In this paper we will answer several questions about the demographic effects of partner diversity in multi-species mutualisms:
\begin{enumerate}
  \item{What are the contrasting demographic effects of multiple partners?}
  \item{What combination of ant partners is associated with the highest fitness for the tree cholla?}
  item{What mechanisms explain the effects of multiple ant partners on the tree cholla?}
\end{enumerate}

To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti. 
We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
With the IPM we moved beyond what has been done in previous studies\cite{Palmer2010} to identify which mechanisms explain the effects of partner diversity on the tree cholla.

We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.

\section*{Methods}
\subsection*{Study System}
\subsection*{Natural History}
To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
These arborescent cacti produce cylindrical segments with large spines. 
In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
While most plants produce new segments every season, only those which are mature enough produce flower buds. 
Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.

This EFN is harvested by various ant species in return for defense. 
At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) } and \textit{Camponotus spp. (Camp.)}.

\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year \ref{Fig:Ant_freq}. 
It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.

These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next. 
Across the entire lifespan cacti may interact with multiple species of partners. 
At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\{Miller2014}. 
Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed. 
\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.

There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}. 
More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
These include an unidentified weevil, a cactus bug, and a seed predator. 
The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
The cactus bug, \textit{Narnia pallidicornis, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary. 
These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller 2009}.

\subsection*{Data}

The data collected on these cacti are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
Six of the 30 $\times$ 30 meter plots were created with naturally occurring plants, censused from 2009 to 2014. 
The final two plots were added to this census from 2011 onwards. 
The distribution of these eight plots are shown in figure \ref{fig:map.
Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
For each plant, we recorded the height (cm), maximum crown width, and perpendicular crown width, which are used to calculate plant volume (cm3) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter. 
We recorded plant survival from the last survey to the current survey as a binary data point, 0 (died) or 1 (survived). 
We recorded the total number of flower buds, including how many were aborted and how many were not, which we used to calculate the viability rates of the cacti. 
We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
We also recorded the species of herbivore or seed predator, the number present on the plant, and if there is any clear, new evidence of herbivory on the tree cholla segments.

In addition to this primary dataset, I used several other datasets throughout the analysis in this paper.  Look at how Tom has described these. Cite these other papers which this data comes from. 
The second dataset was *** germination rates.
The third dataset was *** pre-census survival.
The fourth dataset was *** JO_fruit_data_final_dropplant0.csv.

\subsection*{Demographic Modeling}

The statistical models described above parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.  
IPMs are used to estimate fitness of a population across a continuous variable, rather than using a stage- or size-specific variable to categorize the population. 
This IPM was used to estimate the growth rate of the tree cholla population, effectively a quantitative measure of the fitness of the population.

Following previous studies, we modeled the life cycle of \textit{C. imbricata} using continuously size-structured plants,$n_i(x)$, and two discrete seed banks ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
$$
B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) A_i(x) F(x)n_i(x) dx \\
$$
$$
B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
$$

The functions $P$, $F$, and $A$ give the probability of flowering, the number of flowerbuds produced, and the proportion of flowerbuds which will create seeds. 
Each of these functions, estimated by a Bayesian model calculates these values for an $x$-sized plant in year $t$. 
The proportion of flowerbuds which will produce seeds ($A$) is also dependent on the ant species present on the plant $i$ in year $t$. 
The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the 1-year old seed bank. 
Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
Plants can recruit out of the 1-year seed bank with the probability of $\gamma_1$ or transition to the 2-year seed bank with a probability of $1 - \gamma_1$. 
Seeds in the 2-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.

The size dynamics of the plants are given by:

$$
n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
$$
$$
\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
$$

The final equation gives the stochastic population composition in terms of size and ant state of the \textit{C. imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$.
The first term gives the recruitment from 1 and 2-year seedbanks to size $y$.
$\eta(y)$ gives the seedling size distribution, which is assumed to be normally distributed, and $\omega$ gives the proportion of seedlings which survive from germination (late summer) to the census (May). 
The second term gives the changes in the population of the cacti which are not recruits. 
The functions $S$ and $G$ give the probabilities of surviving from year $t$ to $t+1$ and growing to size $y$ from year $t$ to year $t+1$, respectively. 
Each depends on the size $x$ in year $t$ and the ant state $j$ in year $t+1$. 
Finally, $\tau_{ij}$ is the probability of a cactus which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.

Because many of the vital rates were ant-specific, we were able to consider the composition of the population with the presence of any single ant species, or any combination of ant species (with vacancy always included).
These options include: complete vacancy; \textit{Liom.} and vacancy; \textit{Crem.} and vacancy; other and vacancy; \textit{Liom.}, \textit{Crem.}, and vacancy, \textit{Liom.}, other, and vacancy; \textit{Crem.}, other, and vacancy; and all ant partners and vacancy.
We used this option to calculate the distribution of stochastic population fitness for each combination of ant partners to determine if there is benefit from partner diversity and if there is an optimal combination of partners for this system. 
We then used these fitnesses and the yearly break-downs of the model to determine if there was evidence of any biodiversity ecosystem function mechanisms to explain any partner diversity benefits.

\subsection*{Statistical Modeling}
At the start of this section say for all models unless otherwise noted we used vague priors. 
Also need to introduce the names of these functions ex: survival = s(x)
With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of the interaction between various factors on the highlighted variable rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.
Unless otherwise mentioned, all models used vague priors. 
The growth model ($G_j(y,x)$) estimated the growth rate of cacti, with fixed effects for the size and ant partner and random effects for plot and year, using a Normal distribution, with standard deviation dependent on the size of the cactus also with a Normal distribution. 
The survival model ($S_j(x)$) estimated the probability of survival, with fixed effects for the size of the cactus and ant partner and random effects for plot and year, using a Bernoulli distribution. 
The reproduction model ($P(x)$) estimated the probability of reproducing each year, with fixed effects for the size of the cactus and random effects for plot and year, using a Bernoulli distribution. 
The total flowers model ($F(x)$) estimated the total flowers produced by a plant, with fixed effects for its size and random effects for plot and year, using a Truncated Negative Binomial distribution. 
The viability model ($A_i(x)$) estimated the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects for the size of the cactus and the presence of ant partners and random effects for plot and year, using a Binomial distribution.
The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects for the size of the cactus and the previous ant partner and random effects for plot and year, using a Multinomial distribution.  
The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a year, with no fixed or random effects, using a Normal distribution. 
With the second dataset above, we fit a two Bayesian generalized linear models for the 
probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) using a Binomial distribution.
With the third dataset above, we fit a Bayesian generalized linear model for the probability of a seedling, which has sprouted each year, surviving to May ($\delta$) when we visit (accounting for missed mortality events), with fixed effects for the previous size and transect random effects, using a Bernoulli distribution. 
With the fourth dataset above, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cactus ($\kappa$) based on the ant partner, using a Negative Binomial distribution.

These Bayesian GLMMs were run with STAN through version 4.0.2  of R using 1000 iterations, 150 warmups, 3 chains, and 1 thins. 
To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ___).
To assess the fit we simulated data from the models and compared this to the real data (the fit plots are shown in the supplemental documents in Figures __).

With these models, we investigated the impacts of different ant species on the processes of vital rates and the probabilities of individual cacti interacting with each ant species.
To address our first goal, we compared the effects of ant partners on the growth rate, survival rate, and the viability of flowers across sizes. 
We also broke these effects down by every year of data to see if the impacts of ant partners varied across years. 
To address the second goal, we compared the probability of being visited by a given ant based on each possible previous partner to note any patterns of ant tending. 
We also broke these interactions down by every year of data to see if the patterns changed across years.

\vspace{1cm}

\section*{Results}
\subsection*{Vital Rates}
\includegraphics{"Figures/grow_countour_lines_color.png"}
The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
Plants tended by \textit{Crem.} ants had higher growth rates than tree cholla tended by any other ant or vacant. 
The next highest growth rate was seen in plants tended by other ants, followed by \textit{Liom.} ants, then vacant plants. 
The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
\ali{ }{I’m really unsure about what else I should say about this. }

\includegraphics{"Figures/grow_year_ant_timeseries.png"}
We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}. 
In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
In 2006, \textit{Crem.} has the most positive effect on the growth rate of cacti, followed by Vacant, Other, and \textit{Liom.}. 
In 2007, 2008, and 2016, vacant had the most negative effect on the growth rates of cacti, while in 2012, vacant plants had experienced the most positive effects on their growth rates. 
In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative).

\includegraphics{"Figures/surv_panels_cropped.png"}
Our analyses of the bayesian hierarchical survival model showed that large cacti have higher survival rates than small, regardless of the ant partner. 
It also showed that small cacti had significantly different survival rates depending on the ant partner. 
Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:surv}.
As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last. 
\textit{Crem.} tended plants have survival rates ranging from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.  
\textit{Liom.} tended plants have survival rates ranging from $35.997\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
Other tended plants have survival rates ranging from $15.078\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
Vacant plants have survival rates ranging from $22.031\%$ to $99.647\%$, with the rates increasing with the size of the cacti.

\includegraphics{"Figures/surv_year_ant_timeseries.png"}
We broke down the survival rates by year to determine the differences in ant effects across time. 
In 2004, 2011, 2014, and 2014, vacant tree cholla experienced more positive effects on their survival rates than any other cacti, while vacant cacti experienced the most negative effects on survival rates in 2005 and 2010.
In 2007 and 2013, tree cholla tended by \texit{Liom.} experienced significantly more positive effects on the survival rate than any other cacti, whereas in 2018 and 2019 \texit{Liom.} tended cacti experienced the most negative effects on the survival rates. 
In 2004 and 2009, plants tended by ants in the category of other experienced the most negative effects on survival rates, while in 2017 these plants experienced the most positive effects on the survival rates. 
\textit{Crem.} tended tree cholla experienced more negative effects on the survival rates than any other cacti. 
\textit{Crem.}, \texit{Liom.}, and Other-tended plants all experience similar patterns of positive and negative effects on survival rates through most years, with exceptions between the years of 2009-2010, 2011-2012, and 2017-2019.

\includegraphics{"Figures/viab_bars.png"}
Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viab}. 
The lowest observed viability rate of tree cholla flower buds ranged from $62.83\%$ to $83.78\%$ with a mean of $74.604\%$ when there were no ant partners. 
Using a chi squared test I determined that there is a $13.6\%$ chance that this difference between the mean viability rates of \textit{Liom.} tended plants and vacant plants.

\includegraphics{"Figures/viab_year_ant_timeseries.png"}
We broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly. 
IN 2004 vacant cacti experienced the most positive effects on their viability rates, whereas in 2005, vacant cacti experienced the most negative effects on their viability rates compared to other cacti. 
In 2006 and 2017, \textit{Crem.} tended cacti experienced the most positive effects on viability rates, while in 2012, 2014, 2016, and 2018, \textit{Crem.} tended cacti experienced the most negative effects on viability rates. 
In 2005,2013, 2015, and 2019, cacti tended by ants in the other category experienced the most positive effects on the viability rates, while in 2006 they experienced the most negative effects. 
In 2012, \textit{Liom} tended cacti experienced the most positive effects on their viability rates, while in 2019 they experienced the most negative. 
From the years 2013 to 2019, the effects of all ant partners on the viability rates of cacti are tightly coupled in patterns from positive to negative.

\subsection*{Ant Transition Rates}
\includegraphics{"Figures/Ant_Size_Multi.png"}
All very small cacti are vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
Most large tree cholla are tended by \textit{Liom.} ants even if they had a different previous partner.
The largest vacant cacti have a $42.875\%$ probability of being tended by \textit{Liom.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, $28.571\%$ probability of being tended by \textit{Crem.}, and $21.429\%$ probability of being vacant in the next season.
Previously vacant cacti are most likely to stay vacant until the cacti reach about $10 log(m)^3$, at which point they are more likely to be tended by \textit{Liom.} ants in the next season. 
Large cacti tended by \textit{Liom.} ants are likely to be tended by \textit{Liom.} ants again ($90.476\%$) in the next season.
They have a $2.372\%$ probability of being tended by \textit{Crem.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, and $$ probability of being vacant.  
Previously \textit{Liom.} tended cacti are most likely to be vacant until they reach the size of about $7 log(m^3)$, at which point they are most likely to be tended by \textit{Liom.} in the next season.
Only large tree cholla previously tended by \textit{Crem.} are more likely to be tended by \textit{Crem.} again than be tended by \textit{Liom.} ants in the next year. 
Large tree cholla tended by \textit{Crem.} have a $47.059\%$ probability of being tended by \textit{Crem.} in the next season, $33.333\%$ probability of being tended by \texit{Liom.} in the next season, $30\%$ probability of being tended by other ants in the next season, and $33.333\%$ probability of being vacant. 
Previously \texit{Crem.} tended plants are most likely to become vacant in the next season until they reach the size of about $15 log(m^3)$, after which they are more likely to remain tended by \texit{Crem.} ants. 
Cacti previously tended by other ants have a $24.138\%$ probability of being tended by \textit{Crem.} in the next season, $69.231\%$ probability of being tended by \texit{Liom.} in the next season, $7.692\%$ probability of being tended by other ants in the next season, and $23.076\%$ probability of being vacant. 
Medium cacti follow the same patterns as large cacti in partner transitions.

\subsection*{Demographic Modeling}
\includegraphics{"Figures/lambda_post2.png"}
We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences. 
In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda}.
The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.
Similarly, only a few other scenarios had means within the interquartile range of this high partner diversity scenario. 
Honestly have no idea how much detail I should go into in this?

\section*{Discussion}
\subsection*{Vital Rates}
\textbf{Ant partners have contrasting effects on vital rate processes of tree cholla.}
Regression analyses showed that the different ant partners had significantly different impacts on the various vital rate processes of the tree cholla cacti. 
Specifically, \textit{Crem.} tended cacti have advantages, at small sizes, over cacti tended by any other ant or vacant. 
\textit{Crem.} tended tree cholla had the highest survival rates (Figure \ref{fig:surv})  at small sizes and the highest growth rates (Figure \ref{fig:grow}), two of the most important vital rates for small cacti which are not yet reproducing. 
On the other hand, reproducing cacti which have \textit{Liom.} partners  have advantages with the highest viability rates of flower buds (Figure \ref{fig:viab}). 
Together this indicates that the best partner may change as the cacti grow and begin to reproduce, when the most vulnerable part of the plant is the flower with the seeds. 
This reflects the changes in the resource use of tree cholla as they begin to use their resources for reproduction rather than growth. 
The fact that different ant partners have significantly different effects on the various vital rates of tree cholla indicates that none of them are the “perfect” partner, and that the “best” partner may in fact change over the lifespan of the cacti. 
As the tree cholla grew, the best partner changed from \textit{Crem.} ants, partners known for ***, to \textit{Liom.} ants, partners best known for defensive benefits to the cacti, particularly against the seed predators which most impact viability. 
The difference in ant partners made a significant difference in the observed vital rates of the cacti, indicating that considering the interaction between the tree cholla and any individual partner would fail to capture the extent of the benefits to the cacti.
Like many other systems, pairwise perspectives do not fully encompass the complex impacts that multispecies mutualisms have on the focal mutualists vital rates.

\subsection*{Ant Transition Rates}
\textbf{Small cacti remain vacant, while large cacti are most likely to be tended by \textit{Liom.} ants.}
Small cacti are unlikely to be tended because most do not produce EFN, of if they do it is a very small amount. 
Medium-sized to large cacti are much more likely to be tended than vacant as they begin producing EFN, with many factors that determine the most likely partner. 
Some of the ant partners appear to have high turnover rates, meaning they are unlikely to tend the same plant multiple years in a row. 
Ants in the other category have high turnover rates, shown by the fact that medium and large cacti tended by other are unlikely to be tended by other ants again in the next season. 
The reason for these turnovers could be due to the inability to defend their territories, an ability to find and colonize new resources, or an ability to colonize resources that have been left unclaimed. 
This discovery-dominancy tradeoff is a well-studied hypothesis in ant literature \cite{lach2010} that could explain the remaining presence of ants in the other category despite high turnover rates.

On the other hand, \texit{Crem.} ants appear to have lower turnover rates, since large cacti have up to $47.059\%$ probability of being tended by \texit{Crem.} in the next year.
In addition to turnover rates, there are also colonization rates, the probability of a species taking over a cactus that was previously tended by different ants. 
\textit{Liom.} are the only ant partners we observed with high colonization rates. 
Most cacti tended by non-\textit{Liom.} ants have a high probability of being taken over by \texit{Liom} ants in the next season, as shown in Figure \ref{fig:ant}.
This pattern could be due to the well-known high levels of aggression displayed by \textit{Liom.} ants in this system, however there are other possible explanations, such as nectar composition.
The exception to this rule are plants tended by \textit{Crem.} ants which are more likely to remain tended by \textit{Crem.} than be taken over by \textit{Liom.} in the next season. 
The trend that most large ants are likely to be tended by \textit{Liom.} ants reflects the findings that large cacti benefit most from \textit{Liom.} as ant partners. 
As explained in the introduction, there are many factors that determine the colonization rates of different cacti by their ant partners, including EFN quantity and quality, ability to seek out cacti, aggression, and more.
These patterns could be explained by the \textit{Liom.} ants ability to dominate at a resource site and therefore takeover from different ants which originally found the site.

An alternative, or parallel, explanation for these ant transitions could be the changes in EFN composition across ontogeny of the cacti.
As cacti grow the chemistry of EFN produced changes\cite{Miller2013}, changing from more *** composition to *** composition. 
Different ants have preferences for different nectar compositions\cite{Lach2010}, specifically, ***.
This could provide a potential route for the cacti to select their own ideal partners
This indicates a potential future avenue of research into the correlations between ant partners and the chemistry of the EFN produced by the tree cholla.

\subsection*{Demographic Modeling}

\end{document}

\typeout{\bibliography{References.bib}}

### 2022-10-18 Setup_Script.R
Update Setup_Script.R

@@ -302,6 +302,14 @@ cactus <- cactus[ , c("Plot","TagID","Year_t","Goodbuds_t","TotFlowerbuds_t","AB
 ## Export cactus to a csv
 write.csv(cactus, "cholla_demography_20042021_cleaned.csv")
 
+## How many plants do we survey annually (on average)
+yr <- vector()
+for(i in 1:length(unique(cactus$Year_t))){
+  yr[i] <- nrow(subset(cactus, cactus$Year_t == 2003+i))
+}
+yr
+mean(yr[1:17])
+
 ##########################################################################################
 ########          Set up functions and subsets of data          ##########################
 ##########################################################################################

### 2022-10-18 Manuscript/manuscript_draft.tex

@@
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
-\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+\newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}
 
@@
 
-\section*{Context/Introduction}
-BASELINE OF MUTUALISMS
-Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences.
-\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
+\section*{Introduction}
+%BASELINE OF MUTUALISMS
+Mutualisms are pairwise species interactions in which both participants benefit. 
+They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} \tom{with diverse and dynamic consequences}{This sounds nice but does not carry much information.
+Consequences for what?
+And in what way are they diverse and dynamic?}.
+%\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
 Do I need to include more basic mutualism background?
 I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.
+%no, I don't think you need this
+\tom{Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied.
+}{It is not clear from this sentence whether you are implying that some mutualisms are pairwise and some are multi-species, or whether this distinction is just a matter of perspective.
+Either way, this is a sentence about how mutualisms are studied and not about mutualisms per se.}
+While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species may engage with multiple partner species.
+\tom{It has become apparent that partner diversity in mutualisms}{General comment in the spirit of constructive criticism: the ratio of words to information in your writing is high.
+Imagine this sentence started ``Partner diversity in mutualisms often leads...''.
+This has the same meaning, is more efficient, and sounds more declarative and authoritative.} often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}.
+\tom{This means benefits to a focal mutualist can be greater than, less than, or equal to the sum of the benefits observed in each of the  pairwise interactions that makes up the multi-species mutualism.}{This is correct, and interesting, but just noticing that you are priming the read for an analysis that you do not currently do (though you could and we could thinking about how).}
+\tom{This is true for many complicated reasons}{Not a very scientific thing to say}, including \tom{interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability}{I think this list requires more explanation to be informative, but this is probably not the place for it.}\cite{stanton2013, Afkhami2014,Boucher1982}.
+Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and \tom{fighting}{You are probably thinking about ants here.
+At this point in the paper you should be writing more generally.}, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
+For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013.
+Chamberlain2014, Song2020} \tom{which are more common}{This relates to my comment above about how it's not clear whether you are describing mutualisms as intrinsically pairwise or multi-species, or whether this is about research approached.
+I don't think you have a solid foundation for saying that multi-species is more common.}.
+\ali{***** Biodiversity ecosystem function?
+-- TOM: I would save this for next paragraph.}{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explainations of negative effects?
+Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?}
+\tom{Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
+}{Good closing sentence but I would probably not emphasize observational.}
 
-Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied. 
-It has become apparent that partner diversity in mutualisms often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}. 
-This means the benefits can be greater than, less than, or equal to the sum of the benefits observed in each of the pairwise interactions with the focal mutualist that makes up the multi-species mutualism.
-This is true for many complicated reasons, including interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability\cite{stanton2013, Afkhami2014,Boucher1982}.
-Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and fighting, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
-For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020} which are more common.
-\ali{***** Biodiversity ecosystem function?
-}{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explainations of negative effects?
-Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?}
-Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
-
-The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \{Stanton2003}.
-Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year. 
-When partner quality remains consistent across time, meaning one partner is always better or worse than another, the effects of partner diversity on the focal mutualist can be explained by a mechanism called sampling effect.
-This is the concept that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
-When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner.
-In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, such as resource shortages and niche shifts.
-The effects of partner diversity in these systems can be explained by the portfolio effect, the idea that each partner is the ``best" under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}.
+The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of \tom{the focal mutualist}{I think you should introduce the idea of a `focal mutualist' more clearly.} which can be explained by a number of mechanisms.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the functions of beneficial partners can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
+\tom{Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year or in response to different types of stressors. }{SHould cite.}
+When partner quality remains consistent across time, meaning one partner is always better or worse than another, benefits of partner diversity may reflect a sampling effect, where a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
+When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word - not the right word.
+A focal mutualist does not experience a sampling effect.
+The sampling effect is a mechanism for positive effects of diversity.} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner.
+In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, \tom{such as resource shortages and niche shifts}{Unclear what this means.}.
+The effects of partner diversity in these systems can be explained by the \tom{portfolio effect}{I would discuss this last, after sampling effect and complementarity.}, the idea that each partner is the \tom{``best"}{I would not keep using this in quotations and instead find the right scientific language for what you are trying to say.} under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}.
 This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
 Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
-The effects of partner functional diversity can be explained by the mechanism complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
-This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\{Stanton2003}.
-Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\{Batstone2018}.
-\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.}
+The effects of partner functional diversity can be explained by complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
+\tom{This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\cite{Stanton2003}.
+}{I got stuck on this sentence and I am unsure what you are trying to communicate.}
+Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\cite{Batstone2018}.
+\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.
+Tom: there is no new information in this sentence, so I don't think you need it.}
 
-Multi-species mutualisms are a huge group of interactions which vary considerably. 
-While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.
+\tom{For reasons I mentioned, you need to provide conceptual context and background information on partner turnover, probably as a new paragraph here (or elsewhere in the Intro).}{}
+
+\tom{Multi-species mutualisms are a huge group of interactions which vary considerably.}{There is no information in this sentence.}
+\tom{While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.}{Not an effective sentence.
+Don't list the things the paper is not about.
+Instead, explain why ant-plant systems are a relevant and valuable context for studying effects of partner diversity.}
 These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
-While these interactions have been well studied\cite{} in the literature, few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.
+While these interactions have been well studied\cite{} in the literature, \tom{few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.
+}{Cite.
+Also worth pointing out that I don't think any previous study has incorporated temporal fluctuations, i.e.
+portfolio effect.}
 
 This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
-Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.
+\tom{Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.}{Like much of this paragraph, this information is more appropriate for the Study System section of the methods.
+I am not flagging all of that here, but note that much of this paragraph should change.}
 These cacti flower in the spring throughout the summer, their reproductive season, outside of which they are not tended by ants. 
@@
 Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
-This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.
+This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.\tom{}{This paragraph is a little chaotic.
+There is natural history info, methods, and results!
+Needs to be tightened up.
+What do you notice in other papers that you read about how study systems are first described?
+This should help guide your writing.}
 
-In this paper we will answer a number of questions about the demographic effects of partner diversity in multi-species mutualisms:
+In this paper we will answer \tom{a number}{WHat is the number?} of questions about the demographic effects of partner diversity in multi-species mutualisms:
 \begin{enumerate}
@@
 
-To answer these questions we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We analyzed the impacts of partner diversity on each individual vital rate (survivial, growth, floral-viability, etc.) with bayesian models, which we used to parameterize an Integral Projection Model (IPM).
-With the IPM we moved beyond what has been done in previous studies\cite{Palmer2010} to identify which mechanisms explain the effects of partner diversity on the tree cholla.
+To answer these \tom{questions}{I think the questions can be refined further, plus we need the turnover question in there.} we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
+We analyzed the impacts of partner diversity on each individual vital rate (survivial, growth, floral-viability, etc.) with Bayesian statistical models, which we used to parameterize an Integral Projection Model (IPM).
+With the IPM we moved \tom{beyond what has been done in previous studies\cite{Palmer2010}}{If you are going to say this it's important to describe what has been done and how this study goes beyond it.} to identify which mechanisms explain the effects of partner diversity on the tree cholla.
+\tom{}{Generally this paragraph is a good, quick overview of what we did - good way to wrap up the Intro.}
 
-We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
-We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.
+tom{We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
+We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.}{This is a stylistic thing but I don't think that explicit hypotheses are necessary as long as there are clear and compelling questions.}
+
 \section*{Methods}

### 2022-11-17 Figures/Cholla_Analysis_Visualization_Vital_Rates.R
Update Cholla_Analysis_Visualization_Vital_Rates.R

@@ -46,6 +46,31 @@ legend("bottomright", legend = c("A","B"), fill = c("Chartreuse4","Pink"))
 dev.off()
 
 ########################################################################################################
+#### Timeseries Visuals ################################################################################
+########################################################################################################
+##Number of Cacti by ant partners
+ant_freq_ts <- cactus %>% 
+  select(Year_t,ant_t) %>% 
+  drop_na() %>% 
+  group_by(Year_t,ant_t) %>% 
+  summarise(n = n()) %>% 
+  mutate(freq = n / sum(n))
+png("Timeseries.png")
+plot(ant_freq_ts$Year_t[ant_freq_ts$ant_t == "crem"], ant_freq_ts$freq[ant_freq_ts$ant_t == "crem"], 
+     type = "b",pch = 20, cex = 3,col = cremcol, ylim =c(0,1),
+     xlab = "", ylab = "")
+lines(ant_freq_ts$Year_t[ant_freq_ts$ant_t == "liom"], ant_freq_ts$freq[ant_freq_ts$ant_t == "liom"], 
+      type = "b",pch = 20, cex = 3,col = liomcol)
+lines(ant_freq_ts$Year_t[ant_freq_ts$ant_t == "vacant"], ant_freq_ts$freq[ant_freq_ts$ant_t == "vacant"], 
+      type = "b",pch = 20, cex = 3,col = vaccol)
+lines(ant_freq_ts$Year_t[ant_freq_ts$ant_t == "other"], ant_freq_ts$freq[ant_freq_ts$ant_t == "other"], 
+      type = "b",pch = 20, cex = 3,col = othercol)
+legend("topright", legend = c("Liom.","Crem.","Other","Vacant"), fill = c(liomcol,cremcol,othercol,vaccol),
+       cex = 1.5)
+mtext("Year",side=1,line=-1.5,outer=TRUE,cex=1.5)
+mtext("Frequency of Ant Species",side=2,line=-1.5,outer=TRUE,cex=1.5,las=0)
+dev.off()
+########################################################################################################
[…]

### 2022-11-17 Manuscript/References_Cacti.tex
10/17/22 commit

@article{Jones2015,
abstract = {Cheating is a focal concept in the study of mutualism, with the majority of researchers considering cheating to be both prevalent and highly damaging. However, current definitions of cheating do not reliably capture the evolutionary threat that has been a central motivation for the study of cheating. We describe the development of the cheating concept and distill a relative-fitness-based definition of cheating that encapsulates the evolutionary threat posed by cheating, i.e. that cheaters will spread and erode the benefits of mutualism. We then describe experiments required to conclude that cheating is occurring and to quantify fitness conflict more generally. Next, we discuss how our definition and methods can generate comparability and integration of theory and experiments, which are currently divided by their respective prioritisations of fitness consequences and traits. To evaluate the current empirical evidence for cheating, we review the literature on several of the best-studied mutualisms. We find that although there are numerous observations of low-quality partners, there is currently very little support from fitness data that any of these meet our criteria to be considered cheaters. Finally, we highlight future directions for research on conflict in mutualisms, including novel research avenues opened by a relative-fitness-based definition of cheating.},
author = {Jones, Emily I. and Afkhami, Michelle E. and Ak{\c{c}}ay, Erol and Bronstein, Judith L. and Bshary, Redouan and Frederickson, Megan E. and Heath, Katy D. and Hoeksema, Jason D. and Ness, Joshua H. and Pankey, M. Sabrina and Porter, Stephanie S. and Sachs, Joel L. and Scharnagl, Klara and Friesen, Maren L.},
doi = {10.1111/ele.12507},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Jones_2015.pdf:pdf},
issn = {14610248},
journal = {Ecology Letters},
keywords = {Ant-plant,Cleaner fish-client,Cooperation,Fig-fig wasp,Fitness conflict,Legume-rhizobia,Nectar larceny,Partner quality,Plant-mycorrhizae,Yucca-yucca moth},
number = {11},
pages = {1270--1284},
title = {{Cheaters must prosper: Reconciling theoretical and empirical perspectives on cheating in mutualism}},
volume = {18},
year = {2015}
}
@article{Bronstein2001,
abstract = {Mutualisms arc of central importance in biological systems. Despite growing attention in recent years, however, few conceptual themes have yet to be identified that span mutualisms differing in natural history. Here I examine the idea that the ecology and evolution of mutualisms are shaped by diverse costs, not only by the benefits they confer. This concept helps link mutualism to antagonisms such as herbivory, pr{\'{e}}dation, and parasitism, interactions defined largely by the existence of costs. I first briefly review the range of costs associated with mutualisms, then describe how one cost, the consumption of seeds by pollinator offspring, was quantified for one fig/pollinator mutualism. I compare this cost to published values for other fig/pollinator mutualisms and for other kinds of pollinating seed parasite mutualisms, notably the yucca/yucca moth interaction. I then discuss four issues that fundamentally complicate comparative studies of the cost of mutualism:. problems of knowing how to measure the magnitude of any one cost accurately; problems associated with using average estimates in the absence of data on sources of variation; complications arising from the complex correlates of costs, such as functional linkages between costs and benefits; and problems that arise from considering the cost of mutualism as a unilateral issue in what is fundamentally a reciprocal interaction. The rich diversity of as-yet unaddressed questions surrounding the costs of mutualism may best be investigated via detailed studies of individual interactions.},
author = {Bronstein, Judith L.},
doi = {10.1093/icb/41.4.825},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Bronstein2001.pdf:pdf},
issn = {00031569},
journal = {American Zoologist},
number = {4},
pages = {825--839},
title = {{The costs of mutualism}},
volume = {41},
year = {2001}
}
@article{Bronstein2001a,
author = {Bronstein, Judith L.},
journal = {Ecology Letters},
number = {3},
pages = {277--287},
title = {{The Exploitation of Mutualisms}},
volume = {4},
year = {2001}
}
@article{Bronstein2006,
author = {Bronstein, Judith L and Bronstein, Judith L and Alarc{\'{o}}n, Ruben and Geber, Monica},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Bronstein 2006.pdf:pdf},
pages = {412--428},
title = {{The evolution of plant – insect mutualisms}},
year = {2006}
}
@article{Bronstein1998,
author = {Bronstein, Judith L.},
journal = {bioTropica},
number = {2},
pages = {150--161},
title = {{The Contribution of Ant-Plant Protection Studies to Our Understanding of Mutualism}},
volume = {30},
year = {1998}
}
@article{Bascompte2019,
abstract = {Mutualism is a type of interaction in which both partners benefit from each other. For example, a butterfly receives nectar, a rich source of food, from the flower of a plant and in turn moves pollen from that plant to another far away (Figure 1). In order to reflect about the widespread nature of mutualism, John N. Thompson proposed the following thought experiment: try to imagine a plant species that is viable in its natural habitat without using on top of its own nuclear genome the genomes of a mitochondrion, a chloroplast, of several mycorrhizal fungi, of several insects to pollinate it and of several species of bird to disperse its seeds. Mutualism is everywhere and it is assumed that mutualistic interactions have played a major role in the diversification of life on Earth. An often-cited example is the rich adaptive radiation of the flowering plants (angiosperms), with about 300,000 described species. Flowering plants originated around 160 million years ago and diversified fast during the Early Cretaceous, so that by around 120 million years ago they had become widespread. It is generally assumed that such a fast diversification is largely the result of the mutualistic interaction with pollinators. Thus, mutualism has been most likely shaping the diversity of species on Earth from an early stage. But the relationships between mutualism and diversity are not yet clear, mainly because mutualism has traditionally been studied within pairs or small groups of species. Also, mutualism has historically been studied in isolation from competition, so it is unclear how these two forces balance each other in ecological communities. Jordi Bascompte introduces mutualistic networks and how they affect biodiversity patterns.},
author = {Bascompte, Jordi},
doi = {10.1016/j.cub.2019.03.062},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Bascompote_.pdf:pdf},
issn = {09609822},
journal = {Current Biology},
number = {11},
pages = {R467--R470},
pmid = {31163160},
publisher = {Elsevier},
title = {{Mutualism and biodiversity}},
url = {http://dx.doi.org/10.1016/j.cub.2019.03.062},
volume = {29},
year = {2019}
}
@article{Stachowicz2001,
abstract = {POSITIVE INTERACTIONS PLAY A CRITICAL, BUT UNDERAPPRECIATED, ROLE IN ECOLOGICAL COMMUNITIES BY REDUCING PHYSICAL OR BIOTIC STRESSES IN EXISTING HABITATS AND BY CREATING NEW HABITATS ON WHICH MANY SPECIES DEPEND},
author = {Stachowicz, John J.},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Stachowicz_2001.pdf:pdf},
issn = {0006-3568},
journal = {BioScience},
number = {3},
pages = {235--246},
title = {{the Structure of Ecological Communities}},
volume = {51},
year = {2001}
}
@article{Chamberlain2014,
abstract = {The net effects of interspecific species interactions on individuals and populations vary in both sign (-, 0, +) and magnitude (strong to weak). Interaction outcomes are context-dependent when the sign and/or magnitude change as a function of the biotic or abiotic context. While context dependency appears to be common, its distribution in nature is poorly described. Here, we used meta-analysis to quantify variation in species interaction outcomes (competition, mutualism, or predation) for 247 published articles. Contrary to our expectations, variation in the magnitude of effect sizes did not differ among species interactions, and while mutualism was most likely to change sign across contexts (and predation least likely), mutualism did not strongly differ from competition. Both the magnitude and sign of species interactions varied the most along spatial and abiotic gradients, and least as a function of the presence/absence of a third species. However, the degree of context dependency across these context types was not consistent among mutualism, competition and predation studies. Surprisingly, study location and ecosystem type varied in the degree of context dependency, with laboratory studies showing the highest variation in outcomes. We urge that studying context dependency per se, rather than focusing only on mean outcomes, can provide a general method for describing patterns of variation in nature. {\textcopyright} 2014.},
author = {Chamberlain, Scott A. and Bronstein, Judith L. and Rudgers, Jennifer A.},
doi = {10.1111/ele.12279},
file = {:Users/alicampbell/Box Sync/Grad/Endophytes/Papers/Chamberlain_2014.pdf:pdf},
issn = {14610248},
journal = {Ecology Letters},
keywords = {Coefficient of variation,Community context,Conditionality,Distributed outcomes,Interaction strength,Meta-analysis},
number = {7},
pages = {881--890},
pmid = {24735225},
title = {{How context dependent are species interactions?}},
volume = {17},
year = {2014}
}
@incollection{Winfree2020,
author = {Winfree, Rachel},
booktitle = {Unsolved Problems in Ecology},
pages = {337--354},
publisher = {Princeton University Press},
title = {{How Does Biodiversity Relate to Ecosystem Functioning in Natural Ecosystems?}},
year = {2020}
}
@article{West2007,
author = {West, Stuart A. and Griffin, Ashleigh S. and Gardner, Andy},
journal = {Current Biology},
number = {16},
pages = {R661--R672},
title = {{Evolutionary Explanations for Cooperation}},
volume = {17},
year = {2007}
}
@article{Stanton2013,
author = {Stanton, Maureen L.},
journal = {The American Naturalist},
pages = {510--523},
title = {{Interacting Guilds: Moving beyond the Pairwise Perspective on Mutualisms}},
volume = {162},
year = {2013}
}
@article{Stachowicz2005,
author = {Stachowicz, John J. and Whitlatch, Robert B.},
journal = {Ecology},
number = {9},
pages = {2418--2427},
title = {{Multiple mutualists provide complementary benefits to their seaweed host}},
volume = {86},
year = {2005}
}
@article{Song2020,
author = {Song, Chuliang and Ahn, Sarah Von and Rohr, Rudolf P. and Saavedra, Serguei},
journal = {Trends in Ecology and Evolution},
number = {5},
pages = {384--396},
title = {{Towards Probabilistic Understanding About the Context-Dependency of Species Interactions}},
volume = {35},
year = {2020}
}
@article{Rodriguez-Rodriguez2017,
author = {Rodriguez-Rodriguez, Maria C. and Pedro, Jordano and Valido, Alfredo},
journal = {Ecology},
number = {5},
pages = {1266--1276},
title = {{Functional consequences of plant-animal interactions along the mutualism-antagonism gradient}},
volume = {98},
year = {2017}
}
@article{Palmer2010,
author = {Palmer, Todd M. and Doak, Daniel F. and Stanton, Maureen L. and Bronstein, Judith L. and Kiers, Toby E. and Young, Truman P. and Goheen, Jacob R. and Pringle, Robert M.},
journal = {PNAS},
number = {40},
pages = {17234--17239},
title = {{Synergy of multiple partners, including freeloaders, increases host fitness in a multispecies mutualism}},
volume = {107},
year = {2010}
}
@article{Ohm2014,
annote = {These plants have tradeoffs over time between the Crem and Liom -equal at young -Liom only at reproduction The ants have different defense abilities and pollinator tradeoffs -liom does not have as big of negative effects on pollinators and does a better job defending from enemies -crem has big pollination effects (negative) as well as does a worse job defending -part of the reason why this is possible is that the ants are active at different times of day than the pollinators The Liom are the better ant defender because they keep away more enemies as well has have less negative impacts on the reproduction of the plants (which is the most important mechanism for fitness). The Crem are good enough for young plants because pollination is not important and it has been shown in different papers that having any defender is better than having no defender It appears that the presence of aggressive defenders has little impact on pollination.},
author = {Ohm, Johanna R. and Miller, Tom E.X.},
journal = {Ecology},
mendeley-groups = {Cactus Info},
number = {10},
pages = {2924--2935},
title = {{Balancing Anti-Herbivore Benefits and Anti-Pollinator Costs of Defensive Mutualists}},
volume = {95},
year = {2014}
}
@article{Miller2007,
author = {Miller, Tom E.X.},
journal = {Oikos},
number = {3},
pages = {500--512},
title = {{Does Having Multiple Partners Weaken the Benefits of Faculative Mutualism? A Test with Cacti and Cactus-Tending Ants}},
volume = {116},
year = {2007}
}
@article{Miller2014,
abstract = {Many plants engage ants in defensive mutualisms by offering extrafloral nectar (EFN). Identifying sources of variation in EFN quantity (amount) and quality (composition) is important because they can affect ant visitation and identity and hence effectiveness of plant defence. I investigated plant size and reproductive state (vegetative or flowering) as sources of variation in EFN quantity and quality. I focused on Opuntia imbricata and two ant partners, Crematogaster opuntiae and Liometopum apiculatum. I tested the influence of plant size and nectary type (vegetative vs. reproductive structure) on the probability and rate of EFN secretion, concentrations of total carbohydrates (CH) and amino acids (AAs), and relative abundances of constituent CH and AAs. I also examined how traits of individual nectaries scaled up to influence total plant-level rewards. Parallel observations documented associations between plant demographic state and ant visitation and species identity. EFN quantity and quality were generally greater for larger, reproductive plants. At the scale of individual nectaries, probability of EFN secretion was positively size-dependent and greater for nectaries on reproductive vs. vegetative structures. Rate of EFN secretion, carbohydrate and amino acid concentrations, and the relative abundance of disaccharide vs. monosaccharide sugars were greater for reproductive nectaries but were unaffected by plant size. Nectary-level traits scaled up to influence rewards at the whole-plant level in ways that corresponded to ant visitation: the probability of ant occupancy increased with plant size and reproduction, as did the likelihood of being tended by the superior guard, L. apiculatum. Variability in EFN traits may contribute to changes in ant occupancy and identity across plant sizes and reproductive states. Synthesis. This study provides a thorough examination of how plant investment in biotic defence varies over the life cycle. Explicit consideration of plant demography may enhance understanding of ant-plant mutualisms. Populations of long-lived plants are demographically heterogeneous, spanning sizes and reproductive states. The rewards offered to animal mutualists can track demographic heterogeneity with consequences for plant defence and the dynamics of multispecies mutualisms. This study provides a thorough examination of how plant investment in biotic defence varies over the life cycle. Explicit consideration of plant demography may enhance understanding of ant-plant mutualisms. Populations of long-lived plants are demographically heterogeneous, spanning sizes and reproductive states. The rewards offered to animal mutualists can track demographic heterogeneity with consequences for plant defence and the dynamics of multispecies mutualisms. {\textcopyright} 2013 British Ecological Society.},
author = {Miller, Tom E.X.},
doi = {10.1111/1365-2745.12210},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Miller 2014.pdf:pdf},
issn = {00220477},
journal = {Journal of Ecology},
keywords = {Ant-plant mutualism,Biotic defence,Demography,Extrafloral nectar,Invertase,Ontogeny,Plant development and life-history traits,Stage structure},
mendeley-groups = {Cactus Info},
number = {2},
pages = {496--507},
title = {{Plant size and reproductive state affect the quantity and quality of rewards to animal mutualists}},
volume = {102},
year = {2014}
}
@article{Benson1982,
author = {Benson, L.},
journal = {Stanford University Press, Stanford, CA.},
mendeley-groups = {Cactus Info},
title = {{Cacti of the United States and Canada}},
year = {1982}
}
@article{Ness2006,
author = {Ness, Joshua H. and Morris, W.F. and Bronstein, Judith L.},
journal = {Ecology},
mendeley-groups = {Cactus Info},
number = {4},
pages = {912--921},
title = {{INTEGRATING QUALITY AND QUANTITY OF MUTUALISTIC SERVICE TO CONTRAST ANT SPECIES PROTECTING FEROCACTUS WISLIZENI}},
volume = {87},
year = {2006}
}
@article{Oliveira1999,
abstract = {1. This study examines the anti-herbivore effect of ants visiting the extrafloral nectaries (EFNs) of Opuntia stricta (Cactaceae) and its possible influence on the plant's reproductive output in Mexican coastal sand dunes. Opuntia's EFNs are located in the areoles of the developing tissue of emerging cladodes and flower buds. 2. Ants visited the EFNs of O. stricta on a round-the-clock basis. The associated ant assemblage was formed by nine species distributed in four subfamilies, and the species composition of the principal ant visitors changed markedly from clay to night period. 3. Cladodes of control (ants present) and treatment (ants excluded) plants of Opuntia were equally infested by sucking bugs and mining dipterans. Damage to buds by a pyralid moth, however, was significantly higher on treatment than on control plants. Ant visitation to Opuntia's EFNs translated into a 50% increase in the plant's reproductive output, as expressed by the number of fruits produced by experimental control and treatment branches. Moreover, fruit production by ant-visited branches was positively and significantly associated with the mean monthly rate of ant visitation to EFNs. 4. This is the first demonstration of ant protection leading to increased fruit set in the Cactaceae under natural conditions. Although the consequences of damage by sucking and mining insects remain unclear for Opuntia, the results show how the association of EFNs with vulnerable reproductive plant organs can result in a direct ant-derived benefit to plant fitness.},
author = {Oliveira, P. S. and Rico-Gray, V. and D{\'{i}}az-Castelazo, C. and Castillo-Guevara, C.},
doi = {10.1046/j.1365-2435.1999.00360.x},
file = {:Users/alicampbell/Box Sync/Grad/Endophytes/Papers/Oliveira and Rico-Gray etal. 1999.pdf:pdf},
issn = {02698463},
journal = {Functional Ecology},
keywords = {Ant foraging schedule,Ant protection,Ant-plant mutualism,Cactus,Herbivory,Reproductive output},
mendeley-groups = {Cactus Info},
number = {5},
pages = {623--631},
title = {{Interaction between ants, extrafloral nectaries and insect herbivores in Neotropical coastal sand dunes: Herbivore deterrence by visiting ants increases fruit set in Opuntia stricta (Cactaceae)}},
volume = {13},
year = {1999}
}
@book{Lach2010,
author = {Lach, Lori and Parr, Catherine L. and Abbott, Kirsti L.},
mendeley-groups = {Cactus Info},
pages = {1--23,75--137,157--230},
title = {{Ant Ecology}},
year = {2010}
}
@article{Mann1969,
author = {Mann, J},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Mann 1969.pdf:pdf},
journal = {Smithsonian Inst.},
mendeley-groups = {Cactus Info},
title = {{Cactus-feeding insects and mites.}},
year = {1969}
}
@article{Miller2013,
annote = {This paper shows that the benefits to the ants in our system come primarily in nectar production. The nectar production, however can change in quantity and quality. Reproductive state influences both quantity and quality of the EFN. The probability of EFN secretion in increased by size. This indicates that potentially there is a feedback with ant species and these factors for the EFN.},
author = {Miller, Tom E.X.},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Miller_2014_JEcol (2).pdf:pdf},
journal = {Journal of Ecology},
mendeley-groups = {Cactus Info},
pages = {1--11},
title = {{Plant size and reproductive state affect the quantity and quality of rewards to animal mutualists}},
year = {2013}
}
@article{Miller2006,
abstract = {Life-history theory suggests that iteroparous plants should be flexible in their allocation of resources toward growth and reproduction. Such plasticity could have consequences for herbivores that prefer or specialize on vegetative versus reproductive structures. To test this prediction, we studied the response of the cactus bug (Narnia pallidicornis) to meristem allocation by tree cholla cactus (Opuntia imbricata). We evaluated the explanatory power of demographic models that incorporated variation in cactus relative reproductive effort (RRE; the proportion of meristems allocated toward reproduction). Field data provided strong support for a single model that denned herbivore fecundity as a time-varying, increasing function of host RRE. High-RRE plants were predicted to support larger insect populations, and this effect was strongest late in the season. Independent field data provided strong support for these qualitative predictions and suggested that plant allocation effects extend across temporal and spatial scales. Specifically, late-season insect abundance was positively associated with interannual changes in cactus RRE over 3 years. Spatial variation in insect abundance was correlated with variation in RRE among five cactus populations across New Mexico. We conclude that plant allocation can be a critical component of resource quality for insect herbivores and, thus, an important mechanism underlying variation in herbivore abundance across time and space. {\textcopyright} 2006 by The University of Chicago.},
author = {Miller, Tom E.X. and Tyre, Andrew J. and Louda, Svata M.},
doi = {10.1086/509610},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Miller etal 2006.pdf:pdf},
issn = {00030147},
journal = {American Naturalist},
keywords = {Demographic model,Herbivory,Insect-plant interactions,Opuntia,Population dynamics,Resource allocation},
mendeley-groups = {Cactus Info},
number = {5},
pages = {608--616},
pmid = {17080360},
title = {{Plant reproductive allocation predicts herbivore dynamics across spatial and temporal scales}},
volume = {168},
year = {2006}
}
@article{Miller2009,
abstract = {Understanding the role of consumers in plant population dynamics is important, both conceptually and practically. Yet, while the negative effects of herbivory on plant performance have been well documented, we know much less about how individuallevel damage translates to impacts on population growth or whether spatial variation in herbivory affects patterns of plant distribution. We studied the role of insect herbivory in the dynamics and distribution of the tree cholla cactus (Opuntia imbricata), a long-lived perennial plant, across an elevational gradient in central New Mexico, USA, from low-elevation grassland (1670 m) to a grassland-mountain transition zone (1720 m) to the rocky slopes of the Los Pinos Mountains (1790 m). Tree cholla density increased significantly with elevation, while abundance of and damage by a suite of native, cactus-feeding insects decreased. We combined field experiments and demographic models to test the hypothesis that systematic spatial variation in chronic insect herbivory limits the tree cholla distribution to a subset of suitable habitat across the gradient. Our results support this hypothesis. We found that key demographic functions (survival, growth, fecundity) and the responses of these functions to experimental reductions in insect herbivory varied across the gradient. The effects of insect exclusion on plant growth and seed production were strongest in the lowelevation grassland and decreased in magnitude with increasing elevation. We used the experimental data to parameterize integral projection models (IPM), which predict the asymptotic rate of population increase ($\lambda$). The modeling results showed that insect herbivory depressed $\lambda$ and that the magnitude of this effect was context-dependent. The effect of insect herbivory on population growth was strongest at low elevation ($\delta$$\lambda$klow = 0.095), intermediate at mid elevation ($\delta$$\lambda$mid = 0.046), and weakest at high elevation ($\delta$$\lambda$high = -0.0089). The total effect of insects on $\lambda$ was due to a combination of reductions in growth and in fecundity and their combination; the relative contribution of each of these effects varied spatially. Our results, generated by experimental demography across a heterogeneous landscape, provide new insights into the role of native consumers in the population dynamics and distribution of abundance of long-lived native plants.},
author = {Miller, Tom E.X. and Louda, Svata M. and Rose, Karen A. and Eckberg, James O.},
doi = {10.1890/07-1550.1},
file = {:Users/alicampbell/Box Sync/Grad/Cactus/Papers/Miller etal 2009.pdf:pdf},
issn = {00129615},
journal = {Ecological Monographs},
keywords = {Cahela ponderosella,Chihuahuan Desert,Elevation gradient,Herbivory,Integral projection model,New Mexico,Opuntia imbricata,Plant-Insect interactions,Population dynamics,USA,cactus-feeding insects},
mendeley-groups = {Cactus Info},
number = {1},
pages = {155--172},
title = {{Impacts of insect herbivory on cactus population dynamics: Experimental demography across an environmental gradient}},
volume = {79},
year = {2009}
}
@article{Isbell2011,
author = {Isbell, Forest},
journal = {Nature},
pages = {199--202},
title = {{High plant diversity is needed to maintain ecosystem services}},
volume = {477},
year = {2011}
}
@article{Hooper2005,
author = {Hooper, D. U.},
journal = {Ecological Monographs},
number = {1},
pages = {3--35},
title = {{Effects of Biodiversity on Ecosystem Functioning: A Consensus of Current Knowledge}},
volume = {75},
year = {2005}
}
@article{Frederickson2013,
author = {Frederickson, Megan E.},
journal = {Quarterly Review of Biology},
number = {4},
pages = {269--295},
title = {{Rethinking Mutualism Stability: Cheaters and the Evolution of Sanctions}},
volume = {88},
year = {2013}
}
@article{Cardinale2012,
author = {Cardinale, Bradley J.},
journal = {American Journal of Botany},
number = {3},
pages = {572--592},
title = {{The Functional Role of Producer Diversity in Ecosystems}},
volume = {98},
year = {2012}
}
@article{Bronstein1994,
author = {Bronstein, Judith L.},
journal = {TREE},
number = {6},
pages = {214--217},
title = {{Conditional Outcomes in Mutualistic Interactions}},
volume = {9},
year = {1994}
}
@article{Boucher1982,
author = {Boucher, Douglas H.},
journal = {Annual Review of Ecological Systems},
pages = {317--347},
title = {{The Ecology of Mutualism}},
volume = {13},
year = {1982}
}
@article{Boege2005,
author = {Boege, Karina and Marquis, Robert J.},
journal = {Trends in Ecology and Evolution},
pages = {441--448},
title = {{Facing Herbivory as You Grow up: The Ontogeny of Resistance in Plants}},
volume = {20},
year = {2005}
}
@article{Batstone2018,
author = {Batstone, Rebecca T.},
journal = {Ecology},
number = {5},
pages = {1039--1050},
title = {{Using niche breadth theory to explain generalization in mutualisms}},
volume = {99},
year = {2018}
}
@article{Barton2010,
author = {Barton, Kasey E. and Koricheva, Julia},
journal = {The American Naturalist},
number = {4},
pages = {481--493},
title = {{The Ontogeny of Plant Defense and Herbivory: Characterizing General Patterns Using Meta-Analysis}},
volume = {175},
year = {2010}
}
@article{Amarasekare2003,
author = {Amarasekare},
journal = {Ecology Letters},
pages = {1109--1122},
title = {{Competitive coexistence in spatially structured environments: a synthesis}},
volume = {6},
year = {2003}
}
@article{Afkhami2014,
author = {Afkhami},
journal = {Ecology},
number = {4},
pages = {833--844},
title = {{Multiple mutualist effects: conflict and synergy in multispecies mutualisms}},
volume = {95},
year = {2014}
}

### 2022-11-17 Manuscript/manuscript_draft.tex
10/17/22 commit

@@
 \subsection*{Study System}
+\subsection*{Natural History}
+To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
+This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
+This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
+The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
+These arborescent cacti produce cylindrical segments with large spines. 
+In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+While most plants produce new segments every season, only those which are mature enough produce flower buds. 
+Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
+Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.
 
-To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of Cholla cacti annually since 2004. 
-This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico. 
-The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA *** Benson 1982***. 
-These arborescent cacti produce new cylindrical segments every year which develop either into fruits which produce seeds or into vegetative segments with large spine which increase the size of the cactus %***Miller, Tenhumberg \& Louda 2008***.
-In central New Mexico, vegetative growth occurs from May -- August.
+This EFN is harvested by various ant species in return for defense. 
+At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) } and \textit{Camponotus spp.
+(Camp.)}.
 
-Cholla cacti secrete extra-floral nectar (EFN) from glands on young vegetative segments and reproductive segments. 
-\tom{This extra-floral nectar varies in chemical makeup across the lifespan of the cactus *** Miller ??***. }{This is pretty vague, so if you are going to say this you should add more detail.}
-At the Sevilleta, the cacti are visited by \tom{multiple species}{I think it would be better to say primarily two species plus other rare species, and giving the frequencies} of ground-nesting ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and \tom{many other species}{which ones?} which occur at \tom{small}{low} frequencies.
-These ants are attracted to the EFN and in return fight off the \tom{seed predators}{which ones?} \tom{which also visit the cacti}{An example of words that make sentences clunky without adding any information}.
-These ants never co-occur on the same plant, \tom{instead competing for plant services}{This is a hypothesis, not something that has been demonstrated.}. 
-\tom{Over the lifetime of the cacti each plant interacts with multiple partners as they tend to switch out season to season.}{I would guess that most plants do not have much switching.
-I think you also need to talk about vacancy and the start of EFN production at larger reproductive sizes, more detail on transitions in ant state between vs within years.
-All of this should be in a separate subsection about natural history.}
+\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year \ref{Fig:Ant_freq}.
+It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.
 
-\tom{There}{This paragraph should also be in a natural history subsection and obviously needs more detail.
-See my earlier papers and references therein and let me know if you have specific questions.} are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments.
-There are a number of insects which occur commonly on these cacti at the Sevilleta. 
-These include the weevil, the NP, and the *** (I think there is a moth as well?). 
-The \tom{weevil}{what species?} feeds on ***. 
-The NP feeds on ***. 
-Finally, the moth?? feeds on ***. 
-All of these predators can have significant impacts on the fitness of both individual cacti and the population. *** Miller 2009***.
+These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next.
+Across the entire lifespan cacti may interact with multiple species of partners. 
+At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
+They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
+In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
+The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
+Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
+Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
+Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
+Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
+These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed.
+\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
+These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.
 
-\tom{The}{This should be a subsection for data collection, with significantly more information than what you have here so that the methods could be reproduced.} data collected on these cacti are from an \tom{18-year}{Which years?} data-set.
-There are 8 30 $\times$ 30 meter plots from which this data is taken. 
-Each plot is \tom{surveyed}{what does this mean?}, and each plant within each plot is tagged and \tom{logged}{what does this mean} each year. 
-For every plant we take data on their survival, \tom{size}{how is size measured?}, \tom{ant state}{what is this?}, \tom{herbivore presence}{this makes it sounds like presence/absence which is not what we record}, \tom{flowering data}{what does this mean?}, and \tom{seed data}{what does this mean?}.
-\tom{This allows us to track the progression of individual plants across many years.}{No new information here.} 
-The length of this data-set
+There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}.
+More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
+These include an unidentified weevil, a cactus bug, and a seed predator. 
+The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
+The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
+The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
+These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller 2009}.
 
-\subsection*{Statistical Modeling}
-To understand the impacts of multiple species interactions on the cacti populations, we constructed \tom{bayesian demographic models of many vital rates}{What are bayesian demographic models?
-WHat are the many vital rates?}.
-With these models, we were able to ask (i) Do the impacts of ant presence on \tom{cacti fitness}{these models are not measuring fitness} differ across ant species?
-(ii) What are the probabilities of an individual cactus interacting with each ant species?.\tom{}{I think it's good that you have these as separate questions but I think each one needs a lot more context and detail.}
+	\subsection*{Data}
 
-The vital rates of the cacti considered in this model include survival rate, growth rate, reproductive state, floral abortion rate, \tom{seed production, and germination}{You have not described where these data come from.}.
-\tom{We fit generalized linear mixed effects models in a heirarchical Bayesian framework to quantify each of the vital rates.}{Not a reproducible description of the methods.
-e.g., what distributions were used, what were the fixed and random effects, which models included ant effects, what were the priors, how many chain and iterations, and were convergence and fit assessed, etc?}
-\tom{Each model takes the data from year \textit{t} and estimates the values in year \textit{t+1}.}{This is true only for growth and survival.} 
-This allows us to predict a vital rate for any plant with a given set of characteristics, \tom{such as size and ant state}{Not such as, it is really just these two things.}. 
-\vspace{1cm}
+	The data collected on these cacti are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+	The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
+	Six of the 30 $\times$ 30 meter plots were created with naturally occurring plants, censused from 2009 to 2014. 
+	The final two plots were added to this census from 2011 onwards. 
+	The distribution of these eight plots are shown in figure \ref{fig:map}.
+		Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
+		For each plant, we recorded the height (cm), maximum crown width, and perpendicular crown width, which are used to calculate plant volume (cm3) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
+		We recorded plant survival from the last survey to the current survey as a binary data point, 0 (died) or 1 (survived). 
+		We recorded the total number of flower buds, including how many were aborted and how many were not, which we used to calculate the viability rates of the cacti. 
+		We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
+		We also recorded the species of herbivore or seed predator, the number present on the plant, and if there is any clear, new evidence of herbivory on the tree cholla segments.
+
+		In addition to this primary dataset, I used several other datasets throughout the analysis in this paper.  Look at how Tom has described these. Cite these other papers which this data comes from. 
+		The second dataset was *** germination rates.
+		The third dataset was *** pre-census survival.
+		The fourth dataset was *** $JO_fruit_data_final_dropplant0.csv$.
 
@@
 		The statistical models described above parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.  
-This model was used to estimate the growth rate of the cactus population, effectively a quantitative measure of the fitness of the population. 
-The fitness was estimated under the \tom{range of all different conditions, no ants present, one ant species present, two ant species present, all ant species present.
-These allow us to compare the fitness of the cacti under no mutualistic interactions, pairwise mutualisms, and various multi-species mutualisms to show which case results in the highest fitness for the cacti.}{This needs a more thorough explanation, but I would also do this later after the model has been introduced and explained.}
+		IPMs are used to estimate fitness of a population across a continuous variable, rather than using a stage- or size-specific variable to categorize the population. 
+		This IPM was used to estimate the growth rate of the tree cholla population, effectively a quantitative measure of the fitness of the population.
 
@@
 
-The final equation gives the \tom{size}{not just population size but also composition in terms of size and ant state - this needs to be explained better} of the \textit{C.
+		The final equation gives the stochastic population composition in terms of size and ant state of the \textit{C.
 imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$.
@@
 
- \subsection*{Model analysis}
+		Because many of the vital rates were ant-specific, we were able to consider the composition of the population with the presence of any single ant species, or any combination of ant species (with vacancy always included).
+		These options include: complete vacancy; \textit{Liom.} and vacancy; \textit{Crem.} and vacancy; other and vacancy; \textit{Liom.}, \textit{Crem.}, and vacancy, \textit{Liom.}, other, and vacancy; \textit{Crem.}, other, and vacancy; and all ant partners and vacancy.
+		We used this option to calculate the distribution of stochastic population fitness for each combination of ant partners to determine if there is benefit from partner diversity and if there is an optimal combination of partners for this system.
+		We then used these fitnesses and the yearly break-downs of the model to determine if there was evidence of any biodiversity ecosystem function mechanisms to explain any partner diversity benefits.
 
-\textbf{TOM STOPPED EDITING HERE}
+		\subsection*{Statistical Modeling}
+		At the start of this section say for all models unless otherwise noted we used vague priors. 
+		Also need to introduce the names of these functions ex: survival = s(x)
+		With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of the interaction between various factors on the highlighted variable rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.
+		Unless otherwise mentioned, all models used vague priors. 
+		The growth model ($G_j(y,x)$) estimated the growth rate of cacti, with fixed effects for the size and ant partner and random effects for plot and year, using a Normal distribution, with standard deviation dependent on the size of the cactus also with a Normal distribution.
+		The survival model ($S_j(x)$) estimated the probability of survival, with fixed effects for the size of the cactus and ant partner and random effects for plot and year, using a Bernoulli distribution.
+		The reproduction model ($P(x)$) estimated the probability of reproducing each year, with fixed effects for the size of the cactus and random effects for plot and year, using a Bernoulli distribution.
+		The total flowers model ($F(x)$) estimated the total flowers produced by a plant, with fixed effects for its size and random effects for plot and year, using a Truncated Negative Binomial distribution.
+		The viability model ($A_i(x)$) estimated the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects for the size of the cactus and the presence of ant partners and random effects for plot and year, using a Binomial distribution.
+		The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects for the size of the cactus and the previous ant partner and random effects for plot and year, using a Multinomial distribution.
+		The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a year, with no fixed or random effects, using a Normal distribution. 
+		With the second dataset above, we fit a two Bayesian generalized linear models for the 
+		probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) using a Binomial distribution.
+		With the third dataset above, we fit a Bayesian generalized linear model for the probability of a seedling, which has sprouted each year, surviving to May ($\delta$) when we visit (accounting for missed mortality events), with fixed effects for the previous size and transect random effects, using a Bernoulli distribution.
+		With the fourth dataset above, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cactus ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+
+		These Bayesian GLMMs were run with STAN through version 4.0.2  of R using 1000 iterations, 150 warmups, 3 chains, and 1 thins. 
+		To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ***).
+		To assess the fit we simulated data from the models and compared this to the real data (the fit plots are shown in the supplemental documents in Figures ***).
+
+		With these models, we investigated the impacts of different ant species on the processes of vital rates and the probabilities of individual cacti interacting with each ant species.
+		To address our first goal, we compared the effects of ant partners on the growth rate, survival rate, and the viability of flowers across sizes. 
+		We also broke these effects down by every year of data to see if the impacts of ant partners varied across years. 
+		To address the second goal, we compared the probability of being visited by a given ant based on each possible previous partner to note any patterns of ant tending. 
+		We also broke these interactions down by every year of data to see if the patterns changed across years.
+
+		\vspace{1cm}
 
 		\section*{Results}
+		I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???
+
 		\subsection*{Vital Rates}
+		The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
+		Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
+		Plants tended by \textit{Crem.} ants had higher growth rates than tree cholla tended by any other ant or vacant. 
+		The next highest growth rate was seen in plants tended by other ants, followed by \textit{Liom.} ants, then vacant plants. 
+		The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
+		\ali{ }{I’m really unsure about what else I should say about this. }
 
-We used field data on cholla cacti to evaluate the differing impacts that the ant mutualism has on the fitness of the cacti.
-The fitness can be broken down into a number of vital rate components, including survival, growth, probability of flowering, number of fruits and seeds produced, etc.
-called vital rates.
-We found that many of these vital rates differ in the presence of different ant species.
+		We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}.
+		In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
+		In 2006, \textit{Crem.} has the most positive effect on the growth rate of cacti, followed by Vacant, Other, and \textit{Liom.}. 
+		In 2007, 2008, and 2016, vacant had the most negative effect on the growth rates of cacti, while in 2012, vacant plants had experienced the most positive effects on their growth rates. 
+		In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
+		From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative).
 
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[width = 0.85\linewidth]{Figures/surv_panels_cropped.png}
-	\caption{Probability of cactus survival from year t to year t+1 when tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow).
-Panels (a-d) show the best-fit line estimating the survival probabilities of cacti, with a grey area representing the uncertainty, and circles representing the actual data (scaled to represent number of observations).
-(e) shows the best-fit probabilities of survival of different-sized cacti with all ant partners.}
-	\label{fig:surv}
-\end{figure}
+		Our analyses of the bayesian hierarchical survival model showed that large cacti have higher survival rates than small, regardless of the ant partner. 
+		It also showed that small cacti had significantly different survival rates depending on the ant partner. 
+		Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:surv}.
+		As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last.
+		\textit{Crem.} tended plants have survival rates ranging from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.  
+		\textit{Liom.} tended plants have survival rates ranging from $35.997\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
+		Other tended plants have survival rates ranging from $15.078\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
+		Vacant plants have survival rates ranging from $22.031\%$ to $99.647\%$, with the rates increasing with the size of the cacti.
 
-	The first vital rate we analyzed was the survival of the cacti.
-As you can see in Figure \ref{fig:surv}, the survival rate of cacti vary significantly across their lifetimes as they grow.
-Young (small) cacti have relatively low survival rates, near $30\%$ while the older mature cacti have much higher survival rates, near $100\%$.
-This means cacti are at their most vulnerable to death when they are young.
-The survival rates differ with more than just the size of the cacti, the ant species present on the cactus is also important.
+		We broke down the survival rates by year to determine the differences in ant effects across time. 
+		In 2004, 2011, 2014, and 2014, vacant tree cholla experienced more positive effects on their survival rates than any other cacti, while vacant cacti experienced the most negative effects on survival rates in 2005 and 2010.
+		In 2007 and 2013, tree cholla tended by \textit{Liom.} experienced significantly more positive effects on the survival rate than any other cacti, whereas in 2018 and 2019 \textit{Liom.} tended cacti experienced the most negative effects on the survival rates.
+		In 2004 and 2009, plants tended by ants in the category of other experienced the most negative effects on survival rates, while in 2017 these plants experienced the most positive effects on the survival rates.
+		\textit{Crem.} tended tree cholla experienced more negative effects on the survival rates than any other cacti. 
+		\textit{Crem.}, \textit{Liom.}, and Other-tended plants all experience similar patterns of positive and negative effects on survival rates through most years, with exceptions between the years of 2009-2010, 2011-2012, and 2017-2019.
 
-There are no very small cacti tended by ant partners because they do not produce EFN.
-This is reflected in Figure \ref{fig:surv} where only the vacant data (yellow) in panel (d) extends to the smallest sizes.
-Once cacti reach a size large enough to attract ant partners with EFN, the cacti which are partnered with \textit{Crem.} have the highest survival rates compared to all other cacti.
-When the cacti reach their larger sizes, the partnered plants all reach a nearly $100\%$ survival rate while plants with no partners have a slightly lower survival rate for all large cacti.
-This result indicates that \textit{Crem.} tended plants may have an advantage, when they are relatively small, over other cacti.
+		Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viab}.
+		The lowest observed viability rate of tree cholla flower buds ranged from $62.83\%$ to $83.78\%$ with a mean of $74.604\%$ when there were no ant partners. 
+		Using a chi squared test I determined that there is a $13.6\%$ chance that this difference between the mean viability rates of \textit{Liom.} tended plants and vacant plants.
 
-The next vital rate we analyzed was the viability rate of the cacti, the proportion of fruits produced which are viable for reproduction.
-It is clear from Figure \ref{fig:viab} that the cacti tended by \textit{Liom.} ants experience significantly higher proportions of viable fruits.
-This result aligns with much of the literature which shows that \textit{Liom.} are more aggressive against herbivores and seed predators, therefore are very high quality defenders.
-This high-quality defense is imperative for high viability rates on reproducing cacti which are likely to attract many antagonists due to the quality of their EFN.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics{Figures/viab_hist.png}
-	\caption{Proportion of viable fruits produced by a cactus in year t+1 when tended in year t by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow).
-Each panel (a-d) shows a histogram of the actual data overlayed with a density distribution of the predicted viability rate of the cacti.
-Panel (e) shows the predicted distributions of viability rates separated by ant species.}
-	\label{fig:viab}
-\end{figure}
-
-A third vital rate considered in this paper is the growth rate.
-Cacti who are attacked by herbivores are likely to lose sections of their growth, leading to a lower, or even negative, growth rate.
-In Figure \ref{fig:grow} you can clearly see that untended plants are the only cacti which experience a negative growth rate at any size.
-This again aligns with expectations that tended cacti are being defended from herbivores and are therefore unlikely to lose enough vegetation to result in a negative growth rate.
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics{Figures/grow_panel.png}
-	\caption{The size of a cactus in year t+1 given a previous size in year t and being tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow).
-Each panel (a-d) shows a best-fit prediction of the future size based on the previous size, the uncertainty in grey surrounding the best-fit, and the actual data in (circles scaled to represent the number of observations).
-Panel (e) shows all best-fit lines together in the same colors as above and a line (grey, dashed) above which plants are growing (have a positive growth rate) and below which plants have a negative growth rate.}
-	\label{fig:grow}
-\end{figure}
-
-\begin{figure}[!ht]
-	\centering
-	\includegraphics{Figures/grow_dist_panel2.png}
-	\caption{The size of a cactus in year t+1 given a previous size in year t and being tended by (a) \textit{Crem.} (Purple), (b) \textit{Liom.} (Teal), (c) ants in the other category (Pink), or (d) no ants (Yellow).
-Each panel (a-d) shows a distribution of the predicted future size based on the previous sizeand a histogram of the actual data.
-Panel (e) shows all predicted distributions together in the same colors as above.}
-	\label{fig:grow2}
-\end{figure}
+		We broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly.
+		IN 2004 vacant cacti experienced the most positive effects on their viability rates, whereas in 2005, vacant cacti experienced the most negative effects on their viability rates compared to other cacti.
+		In 2006 and 2017, \textit{Crem.} tended cacti experienced the most positive effects on viability rates, while in 2012, 2014, 2016, and 2018, \textit{Crem.} tended cacti experienced the most negative effects on viability rates.
+		In 2005,2013, 2015, and 2019, cacti tended by ants in the other category experienced the most positive effects on the viability rates, while in 2006 they experienced the most negative effects. 
+		In 2012, \textit{Liom} tended cacti experienced the most positive effects on their viability rates, while in 2019 they experienced the most negative. 
+		From the years 2013 to 2019, the effects of all ant partners on the viability rates of cacti are tightly coupled in patterns from positive to negative.
 
 		\subsection*{Ant Transition Rates}
+		All very small cacti are vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
+		Most large tree cholla are tended by \textit{Liom.} ants even if they had a different previous partner.
+		The largest vacant cacti have a $42.875\%$ probability of being tended by \textit{Liom.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, $28.571\%$ probability of being tended by \textit{Crem.}, and $21.429\%$ probability of being vacant in the next season.
+		Previously vacant cacti are most likely to stay vacant until the cacti reach about $10 log(m)^3$, at which point they are more likely to be tended by \textit{Liom.} ants in the next season. 
+		Large cacti tended by \textit{Liom.} ants are likely to be tended by \textit{Liom.} ants again ($90.476\%$) in the next season.
+		They have a $2.372\%$ probability of being tended by \textit{Crem.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, and $asdfasdf$ probability of being vacant.
+		Previously \textit{Liom.} tended cacti are most likely to be vacant until they reach the size of about $7 log(m^3)$, at which point they are most likely to be tended by \textit{Liom.} in the next season.
+		Only large tree cholla previously tended by \textit{Crem.} are more likely to be tended by \textit{Crem.} again than be tended by \textit{Liom.} ants in the next year. 
+		Large tree cholla tended by \textit{Crem.} have a $47.059\%$ probability of being tended by \textit{Crem.} in the next season, $33.333\%$ probability of being tended by \textit{Liom.} in the next season, $30\%$ probability of being tended by other ants in the next season, and $33.333\%$ probability of being vacant.
+		Previously \textit{Crem.} tended plants are most likely to become vacant in the next season until they reach the size of about $15 log(m^3)$, after which they are more likely to remain tended by \textit{Crem.} ants.
+		Cacti previously tended by other ants have a $24.138\%$ probability of being tended by \textit{Crem.} in the next season, $69.231\%$ probability of being tended by \textit{Liom.} in the next season, $7.692\%$ probability of being tended by other ants in the next season, and $23.076\%$ probability of being vacant.
+		Medium cacti follow the same patterns as large cacti in partner transitions.
 
-In addition to the vital rates of the cacti, we also analyzed the transition rates from one ant partner to another across years.
-
-\begin{table}[]
-	\centering
-	\begin{tabular}{|p{0.25 \linewidth}   |p{0.25 \linewidth}|p{0.25 \linewidth}|}
-		\hline
-		\textbf{Ant States Possible} & \textbf{Growth Rate ($\frac{n_{t+1}}{n_t}$)} & \textbf{Population $\lambda$}\\
-		\hline
-		No ants & &\\
-		\hline
-		\textit{Liom.} & & \\
-		\hline
-		\textit{Crem.}  & & \\
-		\hline
-		Other & & \\
-		\hline
-		\textit{Liom.} \& \textit{Crem.} & & \\
-		\hline
-		\textit{Liom.} \& Other & & \\
-		\hline 
-		\textit{Crem.} \& Other & & \\
-		\hline
-		\textit{Liom.}, \textit{Crem.} \& Other & & \\
-		\hline
-	\end{tabular}
-	\caption{Caption}
-	\label{tab:table_sims}
-\end{table}
-%	\begin{multicols}{2}
+		\subsection*{Demographic Modeling}
+		We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences. 
+		In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda}.
+		The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.
+		Similarly, only a few other scenarios had means within the interquartile range of this high partner diversity scenario. 
+		Honestly have no idea how much detail I should go into in this?
 
 		\section*{Discussion}
+		\subsection*{Vital Rates}
+		\textbf{Ant partners have contrasting effects on vital rate processes of tree cholla.}
+		Regression analyses showed that the different ant partners had significantly different impacts on the various vital rate processes of the tree cholla cacti. 
+		Specifically, \textit{Crem.} tended cacti have advantages, at small sizes, over cacti tended by any other ant or vacant. 
+		\textit{Crem.} tended tree cholla had the highest survival rates (Figure \ref{fig:surv})  at small sizes and the highest growth rates (Figure \ref{fig:grow}), two of the most important vital rates for small cacti which are not yet reproducing.
+		On the other hand, reproducing cacti which have \textit{Liom.} partners  have advantages with the highest viability rates of flower buds (Figure \ref{fig:viab}). 
+		Together this indicates that the best partner may change as the cacti grow and begin to reproduce, when the most vulnerable part of the plant is the flower with the seeds. 
+		This reflects the changes in the resource use of tree cholla as they begin to use their resources for reproduction rather than growth. 
+		The fact that different ant partners have significantly different effects on the various vital rates of tree cholla indicates that none of them are the “perfect” partner, and that the “best” partner may in fact change over the lifespan of the cacti.
+		As the tree cholla grew, the best partner changed from \textit{Crem.} ants, partners known for ***, to \textit{Liom.} ants, partners best known for defensive benefits to the cacti, particularly against the seed predators which most impact viability.
+		The difference in ant partners made a significant difference in the observed vital rates of the cacti, indicating that considering the interaction between the tree cholla and any individual partner would fail to capture the extent of the benefits to the cacti.
+		Like many other systems, pairwise perspectives do not fully encompass the complex impacts that multispecies mutualisms have on the focal mutualists vital rates.
 
-There is a constant tension within life cycles between energy and resource allocation.
-Individuals often have the ability to either put significant resources towards one process or split their resources between multiple.
-Full energy allocation to one process generally means better results, but splitting resources is a form of bet hedging ***.
-The cacti studied here experience a certain tension between their many vital rates, particularly growth and reproduction because each year they produce new floral buds and new vegetative growth.
-Small plants rarely put out flowers and instead focus on growth, but larger plants must choose each year.
-The analyses of the individual vital rates indicate that different ant species may protect different processes better than others.
-In particular \textit{Crem.} may offer advantages to smaller plants by protecting their new vegetation well.
-This is seen in the higher survival rates and growth rates of small and medium cacti with \textit{Crem.} tenders in Figure \ref{fig:grow} and Figure \ref{fig:surv}.
-\textit{Liom.} ants alternatively may protect reproductive focus more than other partners.
-The \textit{Crem.} tended plants however, have lower viability rates than any other ant tended plants, indicating they do not protect flowers well from herbivores(Figure \ref{fig:viab}).
-This is seen in Figure \ref{fig:viab} where it is shown that plants tended by \textit{Liom.} have the highest viability rates of their flowers, resulting in higher pollination rates.
-Larger plants are also less likely to lose branches when they are tended by \textit{Liom.} (Figure \ref{fig:grow}), keeping them healthy and more able to generate new plants.
-Each of these ants individually offer both benefits and drawbacks but considered together may indicate complementarity across ontogeny.
+		\subsection*{Ant Transition Rates}
+		\textbf{Small cacti remain vacant, while large cacti are most likely to be tended by \textit{Liom.} ants.}
+		Small cacti are unlikely to be tended because most do not produce EFN, of if they do it is a very small amount. 
+		Medium-sized to large cacti are much more likely to be tended than vacant as they begin producing EFN, with many factors that determine the most likely partner. 
+		Some of the ant partners appear to have high turnover rates, meaning they are unlikely to tend the same plant multiple years in a row. 
+		Ants in the other category have high turnover rates, shown by the fact that medium and large cacti tended by other are unlikely to be tended by other ants again in the next season. 
+		The reason for these turnovers could be due to the inability to defend their territories, an ability to find and colonize new resources, or an ability to colonize resources that have been left unclaimed.
+		This discovery-dominancy tradeoff is a well-studied hypothesis in ant literature \cite{lach2010} that could explain the remaining presence of ants in the other category despite high turnover rates.
+
+		On the other hand, \textit{Crem.} ants appear to have lower turnover rates, since large cacti have up to $47.059\%$ probability of being tended by \textit{Crem.} in the next year.
+		In addition to turnover rates, there are also colonization rates, the probability of a species taking over a cactus that was previously tended by different ants. 
+		\textit{Liom.} are the only ant partners we observed with high colonization rates. 
+		Most cacti tended by non-\textit{Liom.} ants have a high probability of being taken over by \textit{Liom} ants in the next season, as shown in Figure \ref{fig:ant}.
+		This pattern could be due to the well-known high levels of aggression displayed by \textit{Liom.} ants in this system, however there are other possible explanations, such as nectar composition.
+		The exception to this rule are plants tended by \textit{Crem.} ants which are more likely to remain tended by \textit{Crem.} than be taken over by \textit{Liom.} in the next season. 
+		The trend that most large ants are likely to be tended by \textit{Liom.} ants reflects the findings that large cacti benefit most from \textit{Liom.} as ant partners. 
+		As explained in the introduction, there are many factors that determine the colonization rates of different cacti by their ant partners, including EFN quantity and quality, ability to seek out cacti, aggression, and more.
+		These patterns could be explained by the \textit{Liom.} ants ability to dominate at a resource site and therefore takeover from different ants which originally found the site.
+
+		An alternative, or parallel, explanation for these ant transitions could be the changes in EFN composition across ontogeny of the cacti.
+		As cacti grow the chemistry of EFN produced changes ***, changing from more *** composition to *** composition. 
+		Different ants have preferences for different nectar compositions\cite{Lach2010}, specifically, ***.
+		This could provide a potential route for the cacti to select their own ideal partners
+		This indicates a potential future avenue of research into the correlations between ant partners and the chemistry of the EFN produced by the tree cholla.
+
+		\subsection*{Demographic Modeling}
 

### 2022-11-17 Manuscript/manuscript_draft_ALI.tex
10/17/22 commit

@@
 \usepackage{color,soul}
-%\usepackage[numbers,sort&compress]{natbib}
 \usepackage{hyperref}
-%\usepackage{cite}
 \usepackage{biblatex}
@@
 
-\title{Thank You, Next}
+\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms}
 \author{Ali M. Campbell\\
@@
 
-Mutualisms are among the most widespread species interactions with diverse and dynamic consequences .
-Mutualisms are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity.
+Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
+They are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity.
 Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies.
@@
 \section*{Context/Introduction}
-BASELINE OF MUTUALISMS
+
 Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences.
+They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007}.
 \ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
@@
 I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.
+While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species.
+%Understanding why these relationships deteriorate or persist is important as mutualism is a cornerstone in maintaining the biodiversity of many systems\cite{Isbell2011}.
 
-Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied. 
-It has become apparent that partner diversity in mutualisms often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}. 
-This means the benefits can be greater than, less than, or equal to the sum of the benefits observed in each of the pairwise interactions with the focal mutualist that makes up the multi-species mutualism.
-This is true for many complicated reasons, including interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability\cite{stanton2013, Afkhami2014,Boucher1982}.
-Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and fighting, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
-For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020} which are more common.
-\ali{***** Biodiversity ecosystem function?
-}{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explanations of negative effects?
-Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?}
-Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
+Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
+%Interactions between partners, partner niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability all impact the fitness benefits to the focal mutualist\cite{stanton2013, Afkhami2014,Boucher1982}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
+Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
 
-The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \{Stanton2003}.
-Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year. 
-When partner quality remains consistent across time, meaning one partner is always better or worse than another, the effects of partner diversity on the focal mutualist can be explained by a mechanism called sampling effect.
-This is the concept that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
-When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner.
-In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, such as resource shortages and niche shifts.
-The effects of partner diversity in these systems can be explained by the portfolio effect, the idea that each partner is the ``best" under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}.
-This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
-Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
-The effects of partner functional diversity can be explained by the mechanism complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
-This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\{Stanton2003}.
-Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\{Batstone2018}.
-\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.}
+Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
+When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure ***, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
+Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
+When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
+Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
+Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure ***, through the portfolio effect\cite{Batstone2018,Lazaro2022}.
 
-Multi-species mutualisms are a huge group of interactions which vary considerably. 
-While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.
-These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
-While these interactions have been well studied\cite{} in the literature, few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.
+When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005}. 
+Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
+The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
+High degrees of partner turnover, also known as oportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
+The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
+For example, susceptibility to enemies can change across life stages\cite{Boege2005}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.
 
-This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
-Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.
+Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
+Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms.
+While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{} affect the overall benefits of mutualism for the plant partner.
 
-These cacti flower in the spring throughout the summer (May to August), their reproductive season, outside of which they are not tended by ants. 
-During the reproductive season these plants experience increased visitation by seed predators and herbivores, which influence their reproductive outputs, survival, growth, etc.
-There are a number of observed ant partners which interact with the cacti in this part of their range, including \textit{Crematogaster opuntiae} (\textit{Crem.}), \textit{Liometopum apiculatum} (\textit{Liom.}), \textit{Forelius pruinosus}, an unidentified \textit{Camponotus} species, an unidentified \textit{Phenogaster} species, among others.
-The ants which most commonly interact with the cacti are \textit{Liom.} and \textit{Crem.}, while others are seen with low enough frequency that for the purpose of this paper we refer to them as ``other" ants.
-These ant partners are all in the same guild, meaning they all provide denfense to the cacti from insect herbivores and seed predators in return for extrafloral nectar (EFN).
-The cacti in this mutualism are each tended by one species at a time, for their entire reproductive season with little turnover until the winter between reproductive seasons. 
+This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla), long-lived EFN-bearing plants that associate with multiple species of ant partners.
+Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
+The ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
+The multiple ant partners do not co-occur on plants, but a single cactus may interact with multiple species over its lifetime.
 Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
@@
 
-\subsection{Questions}
-In this paper we will answer several questions about the demographic effects of partner diversity in multi-species mutualisms:
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{herb_ant_only_flow.png}
+	\caption{asdasdf}
+	\label{fig:herb}
+\end{figure}
+
+In this paper we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
 \begin{enumerate}
 	\item{What are the contrasting demographic effects of multiple partners?}
-  \item{What combination of ant partners is associated with the highest fitness for the tree cholla?}
-  item{What mechanisms explain the effects of multiple ant partners on the tree cholla?}
+	\item{What is the directionality of turnover between multiple partners?}
+	\item{What mechanisms explain the benefits of partner diversity in a multispecies mutualism?}
 \end{enumerate}
@@
 We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
-With the IPM we moved beyond what has been done in previous studies\cite{Palmer2010} to identify which mechanisms explain the effects of partner diversity on the tree cholla.
-
-We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
-We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.
+In previous studies the effects of multiple partners have been identified and quantified\cite{Palmer2010}, but none have identified the mechanisms that explain the benefits of partner diversity.
 
@@
 \subsection*{Study System}
-\subsection*{Natural History}
+
 To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
@@
 This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) } and \textit{Camponotus spp.
+At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \textit{Phenogaster} ant, and \textit{Camponotus spp.
 (Camp.)}.
 
-\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year \ref{Fig:Ant_freq}.
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{Timeseries.png}
+	\caption{asdfasdfsadfasdf}
+	\label{fig:timeseries}
+\end{figure}
+
+\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in Figure \ref{fig:timeseries}.
 It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.
@@
 The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
-Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\{Miller2014}. 
+Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
 Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
@@
 The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
-The cactus bug, \textit{Narnia pallidicornis, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
+The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
 The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller 2009}.
+These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.
 
@@
 The final two plots were added to this census from 2011 onwards. 
-The distribution of these eight plots are shown in figure \ref{fig:map.
+The distribution of these eight plots are shown in figure \ref{fig:map}.
 Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
@@
 The third dataset was *** pre-census survival.
-The fourth dataset was *** JO_fruit_data_final_dropplant0.csv.
+The fourth dataset was *** $JO_fruit_data_final_dropplant0.csv$.
+
+\begin{figure}[h]
+%	\includegraphics[width=\linewidth]{map.png}
+	\caption{asdfasdf}
+	\label{fig:map}
+\end{figure}
 
@@
 \subsection*{Statistical Modeling}
-At the start of this section say for all models unless otherwise noted we used vague priors. 
-Also need to introduce the names of these functions ex: survival = s(x)
 With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of the interaction between various factors on the highlighted variable rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.
@@
 These Bayesian GLMMs were run with STAN through version 4.0.2  of R using 1000 iterations, 150 warmups, 3 chains, and 1 thins. 
-To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ___).
-To assess the fit we simulated data from the models and compared this to the real data (the fit plots are shown in the supplemental documents in Figures __).
+To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ***).
+To assess the fit we simulated data from the models and compared this to the real data (the fit plots are shown in the supplemental documents in Figures ***).
 
@@
 
-\vspace{1cm}
+\section*{Results}
+I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???
 
-\section*{Results}
 \subsection*{Vital Rates}
-\includegraphics{"Figures/grow_countour_lines_color.png"}
-The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
+The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:growth}.
 Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
@@
 
-\includegraphics{"Figures/grow_year_ant_timeseries.png"}
-We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}. 
-In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
-In 2006, \textit{Crem.} has the most positive effect on the growth rate of cacti, followed by Vacant, Other, and \textit{Liom.}. 
-In 2007, 2008, and 2016, vacant had the most negative effect on the growth rates of cacti, while in 2012, vacant plants had experienced the most positive effects on their growth rates. 
-In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
-From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative).
+\begin{figure}[h]
+	\includegraphics[width=.5\linewidth]{grow_contour_lines_color.png}
+	\caption{asdfasdf}
+	\label{fig:growth}
+\end{figure}
 
-\includegraphics{"Figures/surv_panels_cropped.png"}
 Our analyses of the bayesian hierarchical survival model showed that large cacti have higher survival rates than small, regardless of the ant partner. 
 It also showed that small cacti had significantly different survival rates depending on the ant partner. 
-Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:surv}.
+Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:survival}.
 As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last.
@@
 
-\includegraphics{"Figures/surv_year_ant_timeseries.png"}
-We broke down the survival rates by year to determine the differences in ant effects across time. 
-In 2004, 2011, 2014, and 2014, vacant tree cholla experienced more positive effects on their survival rates than any other cacti, while vacant cacti experienced the most negative effects on survival rates in 2005 and 2010.
-In 2007 and 2013, tree cholla tended by \texit{Liom.} experienced significantly more positive effects on the survival rate than any other cacti, whereas in 2018 and 2019 \texit{Liom.} tended cacti experienced the most negative effects on the survival rates.
-In 2004 and 2009, plants tended by ants in the category of other experienced the most negative effects on survival rates, while in 2017 these plants experienced the most positive effects on the survival rates.
-\textit{Crem.} tended tree cholla experienced more negative effects on the survival rates than any other cacti. 
-\textit{Crem.}, \texit{Liom.}, and Other-tended plants all experience similar patterns of positive and negative effects on survival rates through most years, with exceptions between the years of 2009-2010, 2011-2012, and 2017-2019.
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{surv_panels.png}
+	\caption{asdfasdf}
+	\label{fig:survival}
+\end{figure}
 
-\includegraphics{"Figures/viab_bars.png"}
-Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viab}.
+Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viability}.
 The lowest observed viability rate of tree cholla flower buds ranged from $62.83\%$ to $83.78\%$ with a mean of $74.604\%$ when there were no ant partners. 
@@
 
-\includegraphics{"Figures/viab_year_ant_timeseries.png"}
-We broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly.
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{Viability.png}
+	\caption{asdfasdf}
+	\label{fig:viability}
+\end{figure}
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{year_ant_timeseries.png}
+	\caption{asdfasdf}
+	\label{fig:year-ant}
+\end{figure}
+
+We broke down the impacts of ant partner on the growth rate, survival rate, and viability rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}.
+In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
+In 2006, \textit{Crem.} has the most positive effect on the growth rate of cacti, followed by Vacant, Other, and \textit{Liom.}. 
+In 2007, 2008, and 2016, vacant had the most negative effect on the growth rates of cacti, while in 2012, vacant plants had experienced the most positive effects on their growth rates. 
+In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
+From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative). 
+In 2004, 2011, 2014, and 2014, vacant tree cholla experienced more positive effects on their survival rates than any other cacti, while vacant cacti experienced the most negative effects on survival rates in 2005 and 2010.
+In 2007 and 2013, tree cholla tended by \textit{Liom.} experienced significantly more positive effects on the survival rate than any other cacti, whereas in 2018 and 2019 \textit{Liom.} tended cacti experienced the most negative effects on the survival rates.
+In 2004 and 2009, plants tended by ants in the category of other experienced the most negative effects on survival rates, while in 2017 these plants experienced the most positive effects on the survival rates.
+\textit{Crem.} tended tree cholla experienced more negative effects on the survival rates than any other cacti. 
+\textit{Crem.}, \textit{Liom.}, and Other-tended plants all experience similar patterns of positive and negative effects on survival rates through most years, with exceptions between the years of 2009-2010, 2011-2012, and 2017-2019.
+We found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly. 
 IN 2004 vacant cacti experienced the most positive effects on their viability rates, whereas in 2005, vacant cacti experienced the most negative effects on their viability rates compared to other cacti.
@@
 \subsection*{Ant Transition Rates}
-\includegraphics{"Figures/Ant_Size_Multi.png"}
 All very small cacti are vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
@@
 Large cacti tended by \textit{Liom.} ants are likely to be tended by \textit{Liom.} ants again ($90.476\%$) in the next season.
-They have a $2.372\%$ probability of being tended by \textit{Crem.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, and $$ probability of being vacant.
+They have a $2.372\%$ probability of being tended by \textit{Crem.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, and $asdf$ probability of being vacant.
 Previously \textit{Liom.} tended cacti are most likely to be vacant until they reach the size of about $7 log(m^3)$, at which point they are most likely to be tended by \textit{Liom.} in the next season.
 Only large tree cholla previously tended by \textit{Crem.} are more likely to be tended by \textit{Crem.} again than be tended by \textit{Liom.} ants in the next year. 
-Large tree cholla tended by \textit{Crem.} have a $47.059\%$ probability of being tended by \textit{Crem.} in the next season, $33.333\%$ probability of being tended by \texit{Liom.} in the next season, $30\%$ probability of being tended by other ants in the next season, and $33.333\%$ probability of being vacant.
-Previously \texit{Crem.} tended plants are most likely to become vacant in the next season until they reach the size of about $15 log(m^3)$, after which they are more likely to remain tended by \texit{Crem.} ants.
-Cacti previously tended by other ants have a $24.138\%$ probability of being tended by \textit{Crem.} in the next season, $69.231\%$ probability of being tended by \texit{Liom.} in the next season, $7.692\%$ probability of being tended by other ants in the next season, and $23.076\%$ probability of being vacant.
+Large tree cholla tended by \textit{Crem.} have a $47.059\%$ probability of being tended by \textit{Crem.} in the next season, $33.333\%$ probability of being tended by \textit{Liom.} in the next season, $30\%$ probability of being tended by other ants in the next season, and $33.333\%$ probability of being vacant.
+Previously \textit{Crem.} tended plants are most likely to become vacant in the next season until they reach the size of about $15 log(m^3)$, after which they are more likely to remain tended by \textit{Crem.} ants.
+Cacti previously tended by other ants have a $24.138\%$ probability of being tended by \textit{Crem.} in the next season, $69.231\%$ probability of being tended by \textit{Liom.} in the next season, $7.692\%$ probability of being tended by other ants in the next season, and $23.076\%$ probability of being vacant.
 Medium cacti follow the same patterns as large cacti in partner transitions.
 
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{Ant_Size_Multi.png}
+	\caption{asdfasdf}
+	\label{fig:ant}
+\end{figure}
+
 \subsection*{Demographic Modeling}
-\includegraphics{"Figures/lambda_post2.png"}
 We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences. 
-In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda}.
+In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda-det}.
 The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.
@@
 Honestly have no idea how much detail I should go into in this?
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{lambda_det_full.png}
+	\caption{asdfasdf}
+	\label{fig:lambda-det}
+\end{figure}
 
@@
 
-On the other hand, \texit{Crem.} ants appear to have lower turnover rates, since large cacti have up to $47.059\%$ probability of being tended by \texit{Crem.} in the next year.
+On the other hand, \textit{Crem.} ants appear to have lower turnover rates, since large cacti have up to $47.059\%$ probability of being tended by \textit{Crem.} in the next year.
 In addition to turnover rates, there are also colonization rates, the probability of a species taking over a cactus that was previously tended by different ants. 
 \textit{Liom.} are the only ant partners we observed with high colonization rates. 
-Most cacti tended by non-\textit{Liom.} ants have a high probability of being taken over by \texit{Liom} ants in the next season, as shown in Figure \ref{fig:ant}.
+Most cacti tended by non-\textit{Liom.} ants have a high probability of being taken over by \textit{Liom} ants in the next season, as shown in Figure \ref{fig:ant}.
 This pattern could be due to the well-known high levels of aggression displayed by \textit{Liom.} ants in this system, however there are other possible explanations, such as nectar composition.
@@
 An alternative, or parallel, explanation for these ant transitions could be the changes in EFN composition across ontogeny of the cacti.
-As cacti grow the chemistry of EFN produced changes\cite{Miller2013}, changing from more *** composition to *** composition. 
+As cacti grow the chemistry of EFN produced changes ***, changing from more *** composition to *** composition. 
 Different ants have preferences for different nectar compositions\cite{Lach2010}, specifically, ***.
@@
 
+\printbibliography
+
 \end{document}

### 2022-11-17 IPM Code/Analysis.R

@@ -7,7 +7,14 @@
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
 source("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code/IPM.R")
 
-
+vcol <- "#ad90ec"
+lcol <- "#084f98"
+ccol <- "#e9a67a"
+ocol <- "#93022f"
+lccol <- "#5dc9cf"
+locol <- "#cf3545"
+cocol <- "#ab59c8"
+acol <- "#5d906b"
 ##########################################################################################################
 ####################### DETERMINISTIC MEAN IPM    ########################################################
 ## Check that every function written in the IPM source code runs properly and that all of the outputs   ##
@@ -324,13 +331,13 @@ layout(matrix(c(1,2,3,4),
 plot(density(lams[,1]), col = "Red",lwd = 2, 
      cex.main = 2.3, xlim = c(0.91,1.05), ylim = c(0,100), cex.lab = 2,
      xlab = "Lambda Values", ylab = "Density Probability", main = "Fitness of Ants")
-lines(density(lams[,2]), col = "Blue", lwd = 2)
+plot(density(lams[,2]), col = "Blue", lwd = 2)
 lines(density(lams[,3]), col = "Green", lwd = 2)
 lines(density(lams[,4]), col = "Yellow", lwd = 2)
-lines(density(lams[,5]), col = "Orange", lwd = 2)
+plot(density(lams[,5]), col = "Orange", lwd = 2)
 lines(density(lams[,6]), col = "Brown", lwd = 2)
 lines(density(lams[,7]), col = "Black", lwd = 2)
-lines(density(lams[,8]), col = "Grey", lwd = 2)
+plot(density(lams[,8]), col = "Grey", lwd = 2)
 legend("topleft",legend = scenario,fill = c("Red","Blue","Green","Yellow","Orange","Brown","Black","Grey"),
        cex = 1.5)
 dev.off()
@@ -341,6 +348,40 @@ boxplot(lams, xlim = c(-2,12),ylim = c(0.9, 1.1),
         at = c(1,4,5,6,8,8,10,12),
         col = c("Red","Blue","Green","Yellow","Orange","Brown","Black","Grey"))
 legend("topleft",legend = scenario,fill = c("Red","Blue","Green","Yellow","Orange","Brown","Black","Grey"),
+       cex = 1.5)
[…]

### 2022-11-17 IPM Code/IPM Post.R

@@ -41,6 +41,7 @@ for(m in 1:nrow(params)){
 }
 
 g
+gxy(1,1,"crem",params[1,])
 
 
 #########################################################################################################

### 2022-11-17 IPM Code/IPM_Stochastic.R

@@ -28,11 +28,11 @@ gxy<-function(x,y,i,params,grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4){
   if(i == "other"){ return(g_other)}
   if(i == "vacant"){ return(g_vac)}
 }
-vac <- vector()
-for(a in 1:10){
-  vac[a] <- gxy(1,1.2,"crem",params,grow_rfx1[,a],grow_rfx2[,a],grow_rfx3[,a],grow_rfx4[,a])
-}
-vac
+# vac <- vector()
+# for(a in 1:10){
+#   vac[a] <- gxy(1,1.2,"crem",params,grow_rfx1[,a],grow_rfx2[,a],grow_rfx3[,a],grow_rfx4[,a])
+# }
+# vac
 
 #########################################################################################################
 ## SURVIVAL AT SIZE X. Returns the probability of survival of a cactus based on size and ant state   ####
@@ -55,11 +55,11 @@ sx<-function(x,i,params,surv_rfx1,surv_rfx2,surv_rfx3,surv_rfx4){
    if(i == "other"){ return(s_other)}
    if(i == "vacant"){ return(s_vac)}
 }
-vac <- vector()
-for(a in 1:10){
-  vac[a] <- sx(1,"liom",params,surv_rfx1[,a],surv_rfx2[,a],surv_rfx3[,a],surv_rfx4[,a])
-}
-vac
+# vac <- vector()
+# for(a in 1:10){
+#   vac[a] <- sx(1,"liom",params,surv_rfx1[,a],surv_rfx2[,a],surv_rfx3[,a],surv_rfx4[,a])
+# }
+# vac
 
 #################################################
 #SURVIVAL*GROWTH. Combine the survival and growth probabilities
@@ -70,11 +70,11 @@ pxy<-function(x,y,i,params,grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4,surv_rfx1,sur
   pxy = sx(xb,i,params,surv_rfx1,surv_rfx2,surv_rfx3,surv_rfx4)*gxy(xb,y,i,params,grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4)
   return(pxy)
 }
-vac <- vector()
-for(a in 1:10){
[…]

### 2022-11-17 IPM Code/IPM_Stochastic_Post.R

@@ -29,24 +29,24 @@ gxy<-function(x,y,i,params,grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4){
   if(i == "vacant"){ return(g_vac)}
 }
 
-##Check that it works properly
-i = c("vacant","crem","liom","other")
-x = c(-1,-5,3,4)
-y = c(-1,-4,3,4)
-g <- matrix(NA,ncol = length(i), nrow = 100)
-l <- list()
-
-for(a in 1:17){ ## year
-for(m in 1:nrow(params)){ ## iteration
-  for(n in seq(1:length(i))){ ## input info
-    g[m,n] <- gxy(x[n],y[n],i[n],params[m,],grow_rfx1[m,a],grow_rfx2[m,a],grow_rfx3[m,a],grow_rfx4[m,a])
-    }
-}
-  l[[a]] <- g
-}
-
-g
-l
+# ##Check that it works properly
+# i = c("vacant","crem","liom","other")
+# x = c(-1,-5,3,4)
+# y = c(-1,-4,3,4)
+# g <- matrix(NA,ncol = length(i), nrow = 100)
+# l <- list()
+# 
+# for(a in 1:17){ ## year
+# for(m in 1:nrow(params)){ ## iteration
+#   for(n in seq(1:length(i))){ ## input info
+#     g[m,n] <- gxy(x[n],y[n],i[n],params[m,],grow_rfx1[m,a],grow_rfx2[m,a],grow_rfx3[m,a],grow_rfx4[m,a])
+#     }
+# }
+#   l[[a]] <- g
+# }
+# 
+# g
+# l
 
 #########################################################################################################
 ## SURVIVAL AT SIZE X. Returns the probability of survival of a cactus based on size and ant state   ####
@@ -70,23 +70,23 @@ sx<-function(x,i,params,surv_rfx1,surv_rfx2,surv_rfx3,surv_rfx4){
   if(i == "vacant"){ return(s_vac)}
 }
 
-##Check that it works properly
-i = c("liom","vacant","crem","other")
-x = c(-1,-5,4,3)
-s <- matrix(NA,ncol = length(i), nrow = 100)
-l <- list()
-
-for(a in 1:17){ ## year
-for(m in 1:nrow(params)){ ## iteration
-  for(n in seq(1:length(i))){ ## input info
[…]

### 2022-11-17 IPM Code/Params.R

@@ -52,7 +52,7 @@ germ2.params<-germ2.params[draws,]
 rec.params<-rec.params[draws,]
 
 ##This file contains random draws from the posterior distributions of the transition models 
-multi.params <- read.csv(paste0(mcmc_dir,"multi_outputs.csv"), header = TRUE,stringsAsFactors=T)
+multi.params <- read.csv(paste0(mcmc_dir,"multi_outputs1.csv"), header = TRUE,stringsAsFactors=T)
 ## Pull the random draws from all posterior distributions
 multi.params<-multi.params[draws,]
 
@@ -286,6 +286,10 @@ params$multi_betalc <- multi.params$beta.4.3
 params$multi_betall <- multi.params$beta.4.4
 params$multi_betal <- multi.params$beta.5.4
 
-
+multi_rfx1 <- cbind(rep(0,100),rep(0,100),rep(0,100),rep(0,100),
+                    rep(0,100),rep(0,100),rep(0,100),rep(0,100),
+                    rep(0,100),rep(0,100),rep(0,100),rep(0,100),
+                    rep(0,100),rep(0,100),rep(0,100),rep(0,100),
+                    rep(0,100))
 
   

### 2022-11-17 Setup_Script.R

@@ -36,6 +36,7 @@ library(magrittr)
 # graphics
 library(RColorBrewer)
 library(corrplot)
+library(ggpubr)
 
 
 #library(shinystan)
@@ -395,7 +396,7 @@ nrow(germ.dat_orig)
 ################        import the data -- Fruit Survival  (No ant state)      ########333#############
 #######################################################################################################
 seedling.dat_orig <- subset(cactus, cactus$Recruit == 1)
-seedling.dat_orig <- seedling.dat_orig[,c("logsize_t1","Recruit")]
+seedling.dat_orig <- seedling.dat_orig[,c("logsize_t1","Recruit","Year_t")]
 nrow(seedling.dat_orig)
 seedling.dat <- na.omit(seedling.dat_orig)
 nrow(seedling.dat)

### 2022-11-17 Manuscript/manuscript_draft_ALI.tex
Manuscript updates -- Ali

@@
 Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007}.
-\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
-Do I need to include more basic mutualism background?
-I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.
-While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species.
-%Understanding why these relationships deteriorate or persist is important as mutualism is a cornerstone in maintaining the biodiversity of many systems\cite{Isbell2011}.
+They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+The benefits from these interactions are called rewards, but they require investment of resources into the interaction\cite{Bronstein2001}.
+The presence of mutualists can shift the available habitat, fitness, or resource uptake of a partner\cite{Stachowicz2001,West2007}.
+While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.
 
-Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-%Interactions between partners, partner niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability all impact the fitness benefits to the focal mutualist\cite{stanton2013, Afkhami2014,Boucher1982}.
+Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
 For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
-Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
+Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
 
 Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
 When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
@@
 The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
-For example, susceptibility to enemies can change across life stages\cite{Boege2005}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.
+For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.
 
-Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
+Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
 Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms.

### 2022-11-17 Manuscript/manuscript_draft_ALI.tex
Manuscript updates -- Ali

@@
 Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-The benefits from these interactions are called rewards, but they require investment of resources into the interaction\cite{Bronstein2001}.
+They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+The benefits from these interactions are called rewards, but they require investment of resources or services into the interaction\cite{Bronstein2001}.
 The presence of mutualists can shift the available habitat, fitness, or resource uptake of a partner\cite{Stachowicz2001,West2007}.
@@
 Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
 Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
@@
 Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
-When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
-When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure ***, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
+When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
 Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
-When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
+When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}.
 Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
-Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure ***, through the portfolio effect\cite{Batstone2018,Lazaro2022}.
+Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.
 
-When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005}. 
+\begin{figure}[h]
+	\includegraphics[width=0.58\linewidth]{complementarity.png}
+	\includegraphics[width=0.40\linewidth]{sampling_effect.png}
+	\caption{asdasdf}
+	\label{fig:comp-samp}
+\end{figure}
+\begin{figure}[h]
+	\includegraphics[width=0.60\linewidth]{portfolio_effect.png}
+	\caption{asdasdf}
+	\label{fig:portfolio}
+\end{figure}
+
+When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
 Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
 The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
-High degrees of partner turnover, also known as oportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
+High degrees of partner turnover, also known as opportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
 The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
@@
 Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms.
-While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{} affect the overall benefits of mutualism for the plant partner.
+While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
 
@@
 \begin{figure}[h]
-	\includegraphics[width=\linewidth]{herb_ant_only_flow.png}
+	\includegraphics[width=.8\linewidth]{herb_ant_only_flow.png}
 	\caption{asdasdf}
@@
 	\item{What are the contrasting demographic effects of multiple partners?}
-	\item{What is the directionality of turnover between multiple partners?}
+	\item{How do size and partner identity impact the directionality of turnover between multiple partners?}
 	\item{What mechanisms explain the benefits of partner diversity in a multispecies mutualism?}
@@
 We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
-In previous studies the effects of multiple partners have been identified and quantified\cite{Palmer2010}, but none have identified the mechanisms that explain the benefits of partner diversity.
+In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but none have identified the mechanisms that explain the benefits of partner diversity.
 

### 2022-11-21 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

@@
 
-\section*{Introduction}
-%BASELINE OF MUTUALISMS
-Mutualisms are pairwise species interactions in which both participants benefit. 
-They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} \tom{with diverse and dynamic consequences}{This sounds nice but does not carry much information.
-Consequences for what?
-And in what way are they diverse and dynamic?}.
-%\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea?
-Do I need to include more basic mutualism background?
-I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.
-%no, I don't think you need this
-\tom{Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied.
-}{It is not clear from this sentence whether you are implying that some mutualisms are pairwise and some are multi-species, or whether this distinction is just a matter of perspective.
-Either way, this is a sentence about how mutualisms are studied and not about mutualisms per se.}
-While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species may engage with multiple partner species.
-\tom{It has become apparent that partner diversity in mutualisms}{General comment in the spirit of constructive criticism: the ratio of words to information in your writing is high.
-Imagine this sentence started ``Partner diversity in mutualisms often leads...''.
-This has the same meaning, is more efficient, and sounds more declarative and authoritative.} often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}.
-\tom{This means benefits to a focal mutualist can be greater than, less than, or equal to the sum of the benefits observed in each of the  pairwise interactions that makes up the multi-species mutualism.}{This is correct, and interesting, but just noticing that you are priming the read for an analysis that you do not currently do (though you could and we could thinking about how).}
-\tom{This is true for many complicated reasons}{Not a very scientific thing to say}, including \tom{interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability}{I think this list requires more explanation to be informative, but this is probably not the place for it.}\cite{stanton2013, Afkhami2014,Boucher1982}.
-Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and \tom{fighting}{You are probably thinking about ants here.
-At this point in the paper you should be writing more generally.}, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
-For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013.
-Chamberlain2014, Song2020} \tom{which are more common}{This relates to my comment above about how it's not clear whether you are describing mutualisms as intrinsically pairwise or multi-species, or whether this is about research approached.
-I don't think you have a solid foundation for saying that multi-species is more common.}.
-\ali{***** Biodiversity ecosystem function?
--- TOM: I would save this for next paragraph.}{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explainations of negative effects?
-Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?}
-\tom{Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.
-}{Good closing sentence but I would probably not emphasize observational.}
+\section*{Context/Introduction}
 
-The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of \tom{the focal mutualist}{I think you should introduce the idea of a `focal mutualist' more clearly.} which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the functions of beneficial partners can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
-\tom{Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year or in response to different types of stressors. }{SHould cite.}
-When partner quality remains consistent across time, meaning one partner is always better or worse than another, benefits of partner diversity may reflect a sampling effect, where a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
-When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word - not the right word.
-A focal mutualist does not experience a sampling effect.
-The sampling effect is a mechanism for positive effects of diversity.} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner.
-In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, \tom{such as resource shortages and niche shifts}{Unclear what this means.}.
-The effects of partner diversity in these systems can be explained by the \tom{portfolio effect}{I would discuss this last, after sampling effect and complementarity.}, the idea that each partner is the \tom{``best"}{I would not keep using this in quotations and instead find the right scientific language for what you are trying to say.} under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}.
-This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
-Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
-The effects of partner functional diversity can be explained by complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
-\tom{This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\cite{Stanton2003}.
-}{I got stuck on this sentence and I am unsure what you are trying to communicate.}
-Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\cite{Batstone2018}.
-\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.
-Tom: there is no new information in this sentence, so I don't think you need it.}
+Mutualisms are species interactions in which all involved participants benefit.
+They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+The benefits from these interactions are called rewards, but they require investment of resources or services into the interaction\cite{Bronstein2001}.
+The presence of mutualists can shift the available habitat, fitness, or resource uptake of a partner\cite{Stachowicz2001,West2007}.
+While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.
 
-\tom{For reasons I mentioned, you need to provide conceptual context and background information on partner turnover, probably as a new paragraph here (or elsewhere in the Intro).}{}
+Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
+Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
 
-\tom{Multi-species mutualisms are a huge group of interactions which vary considerably.}{There is no information in this sentence.}
-\tom{While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.}{Not an effective sentence.
-Don't list the things the paper is not about.
-Instead, explain why ant-plant systems are a relevant and valuable context for studying effects of partner diversity.}
-These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
-While these interactions have been well studied\cite{} in the literature, \tom{few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.
-}{Cite.
-Also worth pointing out that I don't think any previous study has incorporated temporal fluctuations, i.e.
-portfolio effect.}
+Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
+When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
+Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
+When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}.
+Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
+Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.
 
-This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
-\tom{Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.}{Like much of this paragraph, this information is more appropriate for the Study System section of the methods.
-I am not flagging all of that here, but note that much of this paragraph should change.}
-These cacti flower in the spring throughout the summer, their reproductive season, outside of which they are not tended by ants. 
-During the reproductive season these plants experience increased visitation by seed predators and herbivores, which influence their reproductive outputs, survival, growth, etc.
-There are a number of observed ant partners which interact with the cacti in this part of their range, including \textit{Crematogaster opuntiae} (\textit{Crem.}), \textit{Liometopum apiculatum} (\textit{Liom.}), \textit{Forelius pruinosus}, an unidentified \textit{Camponotus} species, an unidentified \textit{Phenogaster} species, among others.
-The ants which most commonly interact with the cacti are \textit{Liom.} and \textit{Crem.}, while others are seen with low enough frequency that for the purpose of this paper we refer to them as ``other" ants.
-These ant partners are all in the same guild, meaning they all provide denfense to the cacti from insect herbivores and seed predators in return for extrafloral nectar (EFN).
-The cacti in this mutualism are each tended by one species at a time, for their entire reproductive season with little turnover until the winter between reproductive seasons. 
+\begin{figure}[h]
+	\includegraphics[width=0.58\linewidth]{complementarity.png}
+	\includegraphics[width=0.40\linewidth]{sampling_effect.png}
+	\caption{asdasdf}
+	\label{fig:comp-samp}
+\end{figure}
+\begin{figure}[h]
+	\includegraphics[width=0.60\linewidth]{portfolio_effect.png}
+	\caption{asdasdf}
+	\label{fig:portfolio}
+\end{figure}
+
+When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
+Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
+The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
+High degrees of partner turnover, also known as opportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
+The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
+For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.
+
+Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
+Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms.
+While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
+
+This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla), long-lived EFN-bearing plants that associate with multiple species of ant partners.
+Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
+The ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
+The multiple ant partners do not co-occur on plants, but a single cactus may interact with multiple species over its lifetime.
 Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
-This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.\tom{}{This paragraph is a little chaotic.
-There is natural history info, methods, and results!
-Needs to be tightened up.
-What do you notice in other papers that you read about how study systems are first described?
-This should help guide your writing.}
+This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.
 
-In this paper we will answer \tom{a number}{WHat is the number?} of questions about the demographic effects of partner diversity in multi-species mutualisms:
+\begin{figure}[h]
+	\includegraphics[width=.8\linewidth]{herb_ant_only_flow.png}
+	\caption{asdasdf}
+	\label{fig:herb}
+\end{figure}
+
+In this paper we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
 \begin{enumerate}
 	\item{What are the contrasting demographic effects of multiple partners?}
-  \item{What combination of ant partners is associated with the highest fitness for the tree cholla?}
-  item{What mechanisms explain the effects of multiple ant partners on the tree cholla?}
+	\item{How do size and partner identity impact the directionality of turnover between multiple partners?}
+	\item{What mechanisms explain the benefits of partner diversity in a multispecies mutualism?}
 \end{enumerate}
 
-To answer these \tom{questions}{I think the questions can be refined further, plus we need the turnover question in there.} we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We analyzed the impacts of partner diversity on each individual vital rate (survivial, growth, floral-viability, etc.) with Bayesian statistical models, which we used to parameterize an Integral Projection Model (IPM).
-With the IPM we moved \tom{beyond what has been done in previous studies\cite{Palmer2010}}{If you are going to say this it's important to describe what has been done and how this study goes beyond it.} to identify which mechanisms explain the effects of partner diversity on the tree cholla.
-\tom{}{Generally this paragraph is a good, quick overview of what we did - good way to wrap up the Intro.}
+To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
+We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
+In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but none have identified the mechanisms that explain the benefits of partner diversity.
 
-tom{We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
-We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.}{This is a stylistic thing but I don't think that explicit hypotheses are necessary as long as there are clear and compelling questions.}
+\section*{Methods}
+\subsection*{Study System}
+
+To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
+This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
+This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
+The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
+These arborescent cacti produce cylindrical segments with large spines. 
+In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+While most plants produce new segments every season, only those which are mature enough produce flower buds. 
+Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
+Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.
+
+This EFN is harvested by various ant species in return for defense. 
+At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \textit{Phenogaster} ant, and \textit{Camponotus spp.
+(Camp.)}.
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{Timeseries.png}
+	\caption{asdfasdfsadfasdf}
+	\label{fig:timeseries}
+\end{figure}
+
+\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in Figure \ref{fig:timeseries}.
+It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.
+
+These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next.
+Across the entire lifespan cacti may interact with multiple species of partners. 
+At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
+They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
+In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
+The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
+Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
+Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
+Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
+Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
+These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed.
+\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
+These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.
+
+There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}.
+More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
+These include an unidentified weevil, a cactus bug, and a seed predator. 
+The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
+The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
+The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
+These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.
 

### 2022-11-21 Manuscript/manuscript_draft.tex
Compile the whole manuscript

@@
 \usepackage{graphics}
+\usepackage{graphicx}
 \usepackage{subfig}
@@
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
-\newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}
+\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
 
-\title{Thank You, Next}
+\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms}
 \author{Ali M. Campbell\\
@@
 	Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
-Depending on environmental context the outcome of the interaction can range from parasitic to mutualistic. 
-Most studies focus on a pair of species interacting in a mutualism, however most observed mutualistic interactions include more than two species. 
-These are multi-species mutualisms, where there are multiple partners interacting with the same focal species. 
-\hl{The partners can directly impact the focal species through the interactions, \tom{and indirectly impact them through interactions with each other}{***This is true but not something we are well poised to address with our data so might be risky to bring it up.***}.
-}
-The cactus \textit{Cylindriopuntia imbricata} reward partners with extrafloral nectar in exchange for defense by various ant species (primarily \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, though there are many other ant species with smaller populations) from herbivores and seed predators.
-\hl{Using a 20-year collection of demographic data (growth, survival, reproduction, ant partners, herbivory, etc.) from these cacti in New Mexico, we estimate the fitness of the cacti populations with every combination of partners, allowing us to *** understand what accounts for partner diversity in this system *** understand if the current partner diversity of the system is associated with the optimal fitness *** and the impacts of these partners on different vital plant processes across their life history (such as partner impacts on reproduction, survival, or growth).
-}
-We found that different ant partners had different impacts on cacti vital rates suggesting there may be a different 'best' partner during pre-reprodctive life stages than during reproductive life stages, and that the highest cacti fitness is associated with the current partner set (all possible ant partners are available).
-Together, these results suggest that complementarity may account for the benefits that these plants experience with partner diversity. 
-Our results demonstrate the importance of evaluating a mutualism within a community context and \tom{suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti}{Nothing you have reported in the abstract so far really supports this statement.}.
-
-\tom{Each of the ant partners}{Again, abstract should briefly state how many and which.} in this system are within the same guild, all offering defense from herbivores, \tom{with only one ant species interacting with each individual cactus in a given period of time}{Redundant with statement above.}.
-Though these ants may offer very similar rewards, \tom{they likely differ in some way}{Vague}. 
-Two possible ways they could differ are the \tom{fitness boost}{Vague and too colloquial.} they offer the focal species or the \tom{period of ontogeny}{We have size structure data but we are not really dealing with ontogeny per se, so I think it is safer to discuss things in terms of size.} in which they offer the greatest fitness boost.
-\tom{Using a long-term data-set, we show that there are differences in the impacts of different ant partners on various vital rates of the cacti across ontogeny.
-Our results demonstrate the importance of evaluating a mutualism within a community context and suggest that even slight differences of rewards between mutualist partners can help promote fitness of the cacti.
-}{I am a little confused - maybe these two paragraphs are two different versions of the abstract?}
-\tom{}{I know this was just a first stab at the abstract, and written when you don't yet have your final results.
-But eventually this will need to be tightened up and edited to better describe the conceptual motivation, what was done, and what was found.}
+They are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity.
+Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies.
+It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions.
+The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms, portfolio effect, complementarity, and sampling effect.
+In this paper, I will focus on defensive ant-plant mutualisms.
+These involve plants which provide extra-floral nectar and/or “housing” to ants which in turn defend them from herbivores.
+While these interactions have been well studied in the literature, few have considered how diversity within partner guilds affect the overall benefits of mutualism for the plant partner.
+	I use the plant, Cylindropuntia imbracata (tree cholla), and ant, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), and more, multispecies mutualism in which the cacti provide extrafloral nectar in exchange for defense from various herbivores and seed predators.
+I used 18 years of data collected from demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, ant partner, and herbivory for 8 30x30 m plots, in New Mexico.
+With this data I parameterize a series of Bayesian generalized linear vital rate models to determine the impacts of different partners on the focal mutualists.
+I found that different ant partners did have different impacts on the vital rates of the tree cholla.
+Specifically, Crem.
+tended plants had advantages in both growth and survival when small, and Liom.
+tended plants had advantages in colonizing partners as well as floral viability.
+With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest fitness, and which of the above mechanisms may explain the effects of partner diversity in this system.
+I found that the real-life scenario (all possible ants are present) lead to the highest fitness for the tree cholla, indicating that partner diversity is beneficial in this system.
+It also shows that complementarity is at play in this system, meaning different partners offer different benefits leading to synergistic benefits for the tree cholla.
 

### 2022-12-02 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

@@
 		\subsection*{Vital Rates}
+		\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{grow_contour_lines_color.png}
+			\caption{asdasdf}
+			\label{fig:grow}
+		\end{figure}
 		The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
@@
 		The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
+
 		\ali{ }{I’m really unsure about what else I should say about this. }
+			\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{year_ant_timeseries.png}
+			\caption{asdasdf}
+			\label{fig:year_ant}
+		\end{figure}
 
@@
 		From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative).
+
+		\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{surv_panels.png}
+			\caption{asdasdf}
+			\label{fig:surv}
+		\end{figure}
 
@@
 
+		\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{Viability.png}
+			\caption{asdasdf}
+			\label{fig:viab}
+		\end{figure}
+
 		Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viab}.
@@
 		\subsection*{Ant Transition Rates}
+			\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{Ant_Size_Multi.png}
+			\caption{asdasdf}
+			\label{fig:ant}
+		\end{figure}
 		All very small cacti are vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
@@
 		\subsection*{Demographic Modeling}
+		\begin{figure}[h]
+			\includegraphics[width=0.58\linewidth]{lambda_det_full.png}
+			\caption{asdasdf}
+			\label{fig:lambda_det}
+		\end{figure}
+		\begin{figure}[h]
+		\includegraphics[width=0.58\linewidth]{lambda_st_full3.png}
+		\caption{asdasdf}
+		\label{fig:lambda_stoch}
+	\end{figure}
 		We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences. 
-		In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda}.
+		In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda_det}.
 		The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.

### 2023-02-02 Manuscript/manuscript_draft.tex
jan 6 comments from tom

@@
 
-Mutualisms are species interactions in which all involved participants benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-The benefits from these interactions are called rewards, but they require investment of resources or services into the interaction\cite{Bronstein2001}.
-The presence of mutualists can shift the available habitat, fitness, or resource uptake of a partner\cite{Stachowicz2001,West2007}.
-While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.
+Mutualisms are species interactions in which all \tom{involved participants}{Notice the redundancy here.
+You don't need ``involved''.} benefit.
+They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985}, and are notoriously \tom{context dependent}{You need to define what you mean by this.} \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+\tom{The benefits from these interactions are called rewards}{This is a pretty rudimentary definition, not necessary.}, but they require investment of resources or services into the interaction\cite{Bronstein2001}.
+The presence of mutualists can shift the available \tom{habitat, fitness, or resource uptake}{These are very different things.
+I don't see why this sentence is needed in the context of your study.} of a partner\cite{Stachowicz2001,West2007}.
+While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.\tom{}{I suggest you re-think whether most of this first paragraph is necessary.}
 
-Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
+Partner diversity in mutualisms \tom{often leads to variable fitness benefits}{Not sure what you are saying here: different partners provide different fitness benefits?
+Or diversity per se has fitness benefits?} for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of \tom{multi-species mutualisms}{I think you need to operationally define this.}\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
 Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
@@
 Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
-When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
-When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
+Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, \tom{some freeloaders, and even some parasites}{What is the difference between a freeloader and a parasite?} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, \tom{the function of the partner can vary}{I think I know what you mean but you should clarify for other readers what is the difference between quality and function.}, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
+\tom{When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.}{I think the content of this paragraph needs to be better organized.
+Describing ways in which partner species differ is conceptually very different than describing mechanisms by which partner diversity can be beneficial -- yet this paragraph seems to glom them together.
+The ``partners are different'' content seems to fit better in the previous paragraph.}
+When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in \tom{Figure}{This figure is great for talks but I don't recommend it for a publication.} \ref{fig:comp-samp}b, sampling effect can explain \tom{the positive effects}{When you say \emph{the} positive effects of partner diversity, does that imply that diversity effects are always positive?
+I think this is the first you've said that partner diversity is beneficial.} of partner diversity\cite{Batstone2018}.
 Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
@@
 
-When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
+When multiple partners are present, \tom{partner turnover}{I suggest introducing this idea a little differently.
+Partner diversity can have different effects depending on whether it manifests all at once or sequentially.
+So this leads to the question of turnover.} is also important to consider, \tom{as partner identity has been shown to impact benefits in mutualisms}{You have made this point several times.}\cite{Djieto-Lordon2005, Ness2006, Bruna2014}.
 Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
 The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
-High degrees of partner turnover, also known as opportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
-The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
-For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.
+High degrees of partner turnover, \tom{also known as opportunism}{I've never seen partner turnover referred to as ``opportunism''.}, may offer \tom{buffering in the face of environmental variation}{This sounds a lot like portfolio effect, so I think you need to clarify how it relates.}\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, \tom{can increase the benefits}{This makes it sounds like diversity is harmful - since low turnover means low diversity.
+Is that an example of negative effects of partner diversity?} received in a mutualism\cite{Sachs2004}.
+\tom{The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
+For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.}{This is an excellent point.
+The second sentence here is a little clunky - see if you can fine-tune it.}
 
-Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
-Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms.
-While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
+\tom{Defensive}{Better transition from concepts to ants?} ant-plant mutualisms are common interactions in which plants provide \tom{extra-floral nectar}{I would just say ``food'' -- see Beltian bodies as another food resource} and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
+\tom{Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}}{This seems like too general/vague a statement to be useful here.}, types of rewards offered\cite{Beattie1985}, and \tom{quality of interaction}{What do you mean by this and how does it differ from types of rewards?} with their plant partners\cite{Rico-Gray1989,Palmer2010}, \tom{making these interactions great systems to study the effects of partner diversity}{A simpler and more to-the-point observation is that these interactions are typically generalized: most EFN plants have multiple ant partners} within mutualisms.
+\tom{While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} \st{in the literature}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.}{Good sentence.}
 
-This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla), long-lived EFN-bearing plants that associate with multiple species of ant partners.
+This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, \st{long-lived} EFN-bearing plant that associates with multiple species of ant partners.
 Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
-The ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
-The multiple ant partners do not co-occur on plants, but a single cactus may interact with multiple species over its lifetime.
-Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
-This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.
+\st{The} ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
+The multiple ant partners do not co-occur on \tom{individual}{} plants, but a single cactus may interact with multiple species over its lifetime.
+Some of the ant partners are more effective defenders than others, with \tom{\textit{Liom.}}{Notice that you have not yet introduced the species, let alone the abbreviation of its latin name, so a reader will be very confused here.
+Try to anticipate what your readers need to stay with you.} tended reproducing plants experiencing the lowest level of herbivory as shown in \tom{Figure \ref{fig:herb}}{You have put a result in the Introduction.
+Sometimes this is appropriate but I don't think it is here.
+You have not described any methods for where this result comes from, so it is not reproducible.
+Make this a result.}.
+\tom{This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.}{I think you can make this point without needing the figure, based on what we have already published from this system.
+It is also worth pointig out that one of the defenders was suggested to be net-negative because it deterred pollinators (Ohm and Miller 2014) but that study did not integrate demographic effects over the life cycle, which is what you do here and worth emphasizing as a source of novelty.}
 
@@
 
-In this paper we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
+In this \tom{paper}{Refer to this as a study and not a paper.} we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
 \begin{enumerate}
 	\item{What are the contrasting demographic effects of multiple partners?}
-	\item{How do size and partner identity impact the directionality of turnover between multiple partners?}
-	\item{What mechanisms explain the benefits of partner diversity in a multispecies mutualism?}
+	\item{How do \tom{size}{size of what?} and partner identity impact the directionality of turnover between multiple partners?}
+	\item{\tom{What mechanisms explain the benefits}{This question pre-supposes that there are benefits of diversity.} of partner diversity in a multispecies mutualism?}
 \end{enumerate}
 
-To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
-In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but none have identified the mechanisms that explain the benefits of partner diversity.
+\tom{To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.}{I think this sentence can be sharpened.
+I don't think that details like flower abortion and ant number are important to include here, but I also don't think this sentence really communicates the nature or volume of the data,}
+We analyzed the \tom{impacts of partner diversity}{The vital rate models do not reflect the impact of partner diversity, they are simply vital rate estimates fit with ant category as a factor -- be careful with how you present this.} on each individual vital rate (survival, growth, floral viability, etc.) with \tom{Bayesian models}{``Bayesian models'' does not carry much meaning.
+A model cannot be intrinsically Bayesian.
+The model is the model, and Bayesian analysis is one way to estimate it's parameters.}, which we used to parameterize an Integral Projection Model (IPM).
+In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but \tom{none have identified the mechanisms that explain the benefits of partner diversity}{Not sure I agree here.
+How are you ``identifying mechanisms'' in a way that Palmer et al.
+did not, for example?}.
 
@@
 
-To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
-This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
-This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
-The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
+\st{To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004.} 
+This study was conducted in the Los \tom{Pinos}{there is an accent on the n} mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico \tom{(34*20’5.3’’N, 106*37’53.2’’W)}{I would not give lat-long here.
+We should provied lat-long of each plot as part of the full data that we make available.}.
+This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
+\st{The} tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
 These arborescent cacti produce cylindrical segments with large spines. 
 In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
-While most plants produce new segments every season, only those which are mature enough produce flower buds. 
+While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
 Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
-Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.
+Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative and reproductive structures\cite{Ness2006,Oliveira1999}.
 
 This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \textit{Phenogaster} ant, and \textit{Camponotus spp.
+At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum \tom{(Liom.)}{I know I have used these abbreviations in previous papers but I would not recommend doing so again.
+More professional and appropriate to refer to \emph{L.
+apiculatum}, etc.}, } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \tom{\textit{Phenogaster}}{Incorrect spelling.} \tom{ant}{Obviously it's an ant.}, and \textit{Camponotus spp.
 (Camp.)}.
@@
 
-\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in Figure \ref{fig:timeseries}.
-It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.
+\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti \tom{annually}{Unclear what you mean by ``annually''?
+Think about the most technically correct and most efficient way to communicate these results.}, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in \tom{Figure}{Again, you have slipped in a result without providing any methods.
+There is a similar figure in Donald and Miller 2022, so you can cite that paper when you report frequencies of association.} \ref{fig:timeseries}.
+\tom{\st{It is important to note that not all cacti are visited every season, they can remain vacant, in fact} }{Notice that this part of the sentence does not provide any information that is not in the second part.}up to 80\% of cacti are vacant in any given year.
 
-These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next.
-Across the entire lifespan cacti may interact with multiple species of partners. 
+These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}\tom{, rather}{Grammatically awkward.
+I would start a new sentence Instead, each cactus} each cactus is visited by a single ant species for the duration of a season, and the \tom{visitor}{What do you mean by visitor?
+Is each individual ant a visitor?} can change from one season to the next.
+\tom{Across the entire lifespan cacti may interact with multiple species of partners. }{This is implied by the sentences above so there is no new information here.}
 At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
+They will visit the cacti every day during the season from early morning *** to sunset *** \tom{abandoning the cacti at night for their nests}{Incorrect. See Ohm and Miller 2014.}. 
 In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
-The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
+\tom{The process by which ants are sorted out onto different cacti is somewhat of a mystery}{too colloquial}, but we know several facts which likely influence the process. 
 Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
-Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
-Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
+Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more \tom{desirable}{I don't yhink we should speculate about ants' desires.}.
+Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in \tom{going after}{too colloquial} more desirable cacti. 
 Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
 These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed.
-\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
-These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.
+\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}\tom{}{Yes this is what I was thinnking as i was reading.
+Much of this fall out of the results of your analysis.
+I would significantly reduce this paragraph.
+Think about what your readers need to know to process what's to come.}
+\st{These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.}
 
-There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}.
-More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
-These include an unidentified weevil, a cactus bug, and a seed predator. 
+There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments \tom{throughout the entire range}{range info not really pertinent} of the tree cholla\cite{Mann1969}.
+\tom{More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
+These include an unidentified weevil, a cactus bug, and a seed predator. }{Notice how these sentences add no information to the paragraph.}
 The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
@@
 The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.
+These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.\tom{}{I think it is also important to summarize what is known about ant defense against these insects from previous papers.
+The fact that we have experimental evidence that ants provide defense is important to include, because your data are all observational.}
 
-\section*{Methods}
+\tom{\section*{Methods}}{Some duplication here...?}
 \subsection*{Study System}
@@
 
-	The data collected on these cacti are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
-	The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
-	Six of the 30 $\times$ 30 meter plots were created with naturally occurring plants, censused from 2009 to 2014. 
+	The data collected \tom{on these cacti}{notice how these words add nothing} are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+	\tom{The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.}{this seems to contradict what you just said about eight plots.}
+	Six of the 30 $\times$ 30 meter plots \tom{were created with naturally occurring plants}{this sounds funny, meaning unclear}, censused from 2009 to \tom{2014}{Not sure why you have these ending in 2014.
+These are plots 1-6, which we continue to collect data from}.
 	The final two plots were added to this census from 2011 onwards. 
-	The distribution of these eight plots are shown in figure \ref{fig:map}.
+	The distribution of these eight plots are shown in \tom{figure}{you probably don't need a plot map - maybe in the supplement} \ref{fig:map}.
 		Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
-		For each plant, we recorded the height (cm), maximum crown width, and perpendicular crown width, which are used to calculate plant volume (cm3) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-		We recorded plant survival from the last survey to the current survey as a binary data point, 0 (died) or 1 (survived). 
-		We recorded the total number of flower buds, including how many were aborted and how many were not, which we used to calculate the viability rates of the cacti. 
+		For each plant, we recorded the height (cm), maximum crown width, and \st{perpendicular }crown width perpendicular to the maximum, which are used to calculate plant volume (\tom{cm3}{make this a superscript}) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
+		We recorded plant \tom{survival}{I would start with survival because if it's dead nothing else is recorded} from the last survey to the current survey\st{ as a binary data point, 0 (died) or 1 (survived)}.
+		We recorded the total number of flower buds, including how many were aborted and how many were not\st{, which we used to calculate the viability rates of the cacti}. 
 		We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
-		We also recorded the species of herbivore or seed predator, the number present on the plant, and if there is any clear, new evidence of herbivory on the tree cholla segments.
+		We also recorded the species of herbivore or \tom{seed predator}{we do not record seed predator data}, the number present on the plant, and \tom{if there is any clear, new evidence of herbivory}{beware that this has not been very consistent} on the tree cholla segments.
 
-		In addition to this primary dataset, I used several other datasets throughout the analysis in this paper.  Look at how Tom has described these. Cite these other papers which this data comes from. 
+		In addition to this primary dataset, I used \tom{several other datasets}{I don't think you need to say much here, you can cite prevous papers when you describe these parts of the IPM} throughout the analysis in this paper.
+Look at how Tom has described these.
+Cite these other papers which this data comes from.
 		The second dataset was *** germination rates.
@@
 
-		The statistical models described above parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.  
-		IPMs are used to estimate fitness of a population across a continuous variable, rather than using a stage- or size-specific variable to categorize the population. 
-		This IPM was used to estimate the growth rate of the tree cholla population, effectively a quantitative measure of the fitness of the population.
+		The \tom{statistical models described above}{they are actually below - but i would put them above.} parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.
+		IPMs are used to estimate fitness of a population across a \tom{continuous variable}{you actually have both continuous and discrete variables}, \tom{rather than using a stage- or size-specific variable}{not an accurate description of how IPMs differ from MPMs} to categorize the population.
+		This IPM was used to estimate the \tom{growth rate }{not just growth rate, but growthrate as a function of partner identity and diversity -- important to include}of the tree cholla population, \tom{effectively}{why effectively?} a quantitative measure of the fitness of the population.
 
-		Following previous studies, we modeled the life cycle of \textit{C.
-imbricata} using continuously size-structured plants,$n_i(x)$, and two discrete seed banks ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
+		Following previous studies, we modeled the life cycle of \tom{\textit{C.
+imbricata}}{sometimes you say cholla, sometimes cacti, sometimes the latin name -- stay consistent} using continuously size-structured plants,\tom{$n_i(x)$}{what is n?
+what is i?
+what is x?}, and two discrete \tom{seed banks}{there is no beta in the model below.
+what is t?} ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
 		$$
@@
 		The functions $P$, $F$, and $A$ give the probability of flowering, the number of flowerbuds produced, and the proportion of flowerbuds which will create seeds. 
-		Each of these functions, estimated by a Bayesian model calculates these values for an $x$-sized plant in year $t$. 
+		Each of these functions, \st{estimated by a Bayesian model} \tom{calculates}{the function does not do any calculation} these values for an $x$-sized plant in year $t$. 
 		The proportion of flowerbuds which will produce seeds ($A$) is also dependent on the ant species present on the plant $i$ in year $t$. 
@@
 
-		The final equation gives the stochastic population composition in terms of size and ant state of the \textit{C.
+		The final equation gives the \tom{stochastic population composition }{what does this mean?}in terms of size and ant state of the \textit{C.
 imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$.
@@
 		Each depends on the size $x$ in year $t$ and the ant state $j$ in year $t+1$. 
-		Finally, $\tau_{ij}$ is the probability of a cactus which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
+		Finally, $\tau_{ij}$ is the probability of a cactus which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.\tom{}{the ants dynamics of the model need to be presented more thoroughly and clearly.}
 
-		Because many of the vital rates were ant-specific, we were able to consider the composition of the population with the presence of any single ant species, or any combination of ant species (with vacancy always included).
+		\tom{}{This paragraph should get its own subsection.
+This is about model analysis, not model structure.}\tom{Because many of the vital rates were ant-specific}{you need to explain why some are and some are not}, we were able to consider the composition of the population with the presence of any single ant species, or any combination of ant species (with vacancy always included).
 		These options include: complete vacancy; \textit{Liom.} and vacancy; \textit{Crem.} and vacancy; other and vacancy; \textit{Liom.}, \textit{Crem.}, and vacancy, \textit{Liom.}, other, and vacancy; \textit{Crem.}, other, and vacancy; and all ant partners and vacancy.
 		We used this option to calculate the distribution of stochastic population fitness for each combination of ant partners to determine if there is benefit from partner diversity and if there is an optimal combination of partners for this system.
-		We then used these fitnesses and the yearly break-downs of the model to determine if there was evidence of any biodiversity ecosystem function mechanisms to explain any partner diversity benefits.
+		We then used these fitnesses and \tom{the yearly break-downs of the model}{what does this mean?} to determine if there was \tom{evidence of any biodiversity ecosystem function mechanisms to explain any partner diversity benefits}{you never describe any criteria for detecting the different mechanisms.
+What does one need to see in data to conclude that one or another is operating?}.
 
@@
 		Also need to introduce the names of these functions ex: survival = s(x)
-		With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of the interaction between various factors on the highlighted variable rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.
+		With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of \tom{the interaction between various factors on the highlighted variable}{confusing} rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.\tom{}{this is a confusing sentence.
+try to sharpen this.}
 		Unless otherwise mentioned, all models used vague priors. 
-		The growth model ($G_j(y,x)$) estimated the growth rate of cacti, with fixed effects for the size and ant partner and random effects for plot and year, using a Normal distribution, with standard deviation dependent on the size of the cactus also with a Normal distribution.
+		The growth model ($G_j(y,x)$) estimated the \tom{growth rate}{this is not what you modeled} of cacti, with fixed effects for the size and ant partner and random effects for plot and year, using a Normal distribution, with standard deviation dependent on the size of the cactus \tom{also with a Normal distribution}{??}.
 		The survival model ($S_j(x)$) estimated the probability of survival, with fixed effects for the size of the cactus and ant partner and random effects for plot and year, using a Bernoulli distribution.
 		The reproduction model ($P(x)$) estimated the probability of reproducing each year, with fixed effects for the size of the cactus and random effects for plot and year, using a Bernoulli distribution.
-		The total flowers model ($F(x)$) estimated the total flowers produced by a plant, with fixed effects for its size and random effects for plot and year, using a Truncated Negative Binomial distribution.
-		The viability model ($A_i(x)$) estimated the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects for the size of the cactus and the presence of ant partners and random effects for plot and year, using a Binomial distribution.
+		The total flowers model ($F(x)$) estimated the total flowers produced by a plant, with fixed effects for its size and random effects for plot and year, using a \tom{Truncated}{truncated how?} Negative Binomial distribution.
+		The viability model ($A_i(x)$) estimated the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects for the size of the cactus and the \tom{presence of ant partners}{this is not the variable in the model} and random effects for plot and year, using a Binomial distribution.
 		The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects for the size of the cactus and the previous ant partner and random effects for plot and year, using a Multinomial distribution.
 		The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a year, with no fixed or random effects, using a Normal distribution. 
-		With the second dataset above, we fit a two Bayesian generalized linear models for the 
+		With the \tom{second dataset}{what second data set? same question for 3 and 4} above, we fit a two Bayesian generalized linear models for the 
 		probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) using a Binomial distribution.
@@
 
-		These Bayesian GLMMs were run with STAN through version 4.0.2  of R using 1000 iterations, 150 warmups, 3 chains, and 1 thins. 
+		These Bayesian GLMMs were run with STAN through version 4.0.2  of R using \tom{1000 iterations}{do you mean 10000?
+1000 is not very many}, \tom{150 warmups, 3 chains, and 1 thins.}{``warmups'' and ``thins'' are not the right word choices}
 		To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ***).
@@
 
-		With these models, we investigated the impacts of different ant species on the processes of vital rates and the probabilities of individual cacti interacting with each ant species.
+		\tom{With these models, we investigated the impacts of different ant species on the processes of vital rates and the probabilities of individual cacti interacting with each ant species.
 		To address our first goal, we compared the effects of ant partners on the growth rate, survival rate, and the viability of flowers across sizes. 
@@
 		We also broke these interactions down by every year of data to see if the patterns changed across years.
+}{This needs a much more thorough explanation.
+You have not described any methods for a stochastic analysis.
+You need a new section here explaining these methods in a way that is reproducible.
+Keep in mind that reproducibility is the standard we are aiming for, and in the current draft, just about all of the methods fall short of that standard.}
 
@@
 		\section*{Results}
-		I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???
+		\tom{I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???}{Let's talk about this.}
 
@@
 		\end{figure}
-		The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
+		The results of the \tom{Bayesian}{the model is not intrinsically Bayesian} growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in \tom{Figure}{this figure needs data} \ref{fig:grow}.
 		Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
@@
 
-		We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}.
+		\tom{We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}.
 		In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
@@
 		In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
-		From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative).
+		From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative). }{This is confusing and I think incorrect.}
 
@@
 		As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last.
-		\textit{Crem.} tended plants have survival rates ranging from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.  
+		\textit{Crem.} tended plants have survival rates \tom{ranging}{These ranges are not meaningful because survival is so dependent on size.} from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.
 		\textit{Liom.} tended plants have survival rates ranging from $35.997\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
@@
 		The lowest observed viability rate of tree cholla flower buds ranged from $62.83\%$ to $83.78\%$ with a mean of $74.604\%$ when there were no ant partners. 
-		Using a chi squared test I determined that there is a $13.6\%$ chance that this difference between the mean viability rates of \textit{Liom.} tended plants and vacant plants.
+		Using a \tom{chi squared test}{I did not know you did this but i don't think you should} I determined that there is a $13.6\%$ chance that this difference between the mean viability rates of \textit{Liom.} tended plants and vacant plants.
 
-		We broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly.
+		\tom{We}{I think you need to re-think these paragraphs describing the time series, because I don't think you have given readers enough info to interpret these.} broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly.
 		IN 2004 vacant cacti experienced the most positive effects on their viability rates, whereas in 2005, vacant cacti experienced the most negative effects on their viability rates compared to other cacti.
@@
 		\end{figure}
-		All very small cacti are vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
+		All very small cacti \tom{are}{your tense is inconsistent} vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
 		Most large tree cholla are tended by \textit{Liom.} ants even if they had a different previous partner.
-		The largest vacant cacti have a $42.875\%$ probability of being tended by \textit{Liom.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, $28.571\%$ probability of being tended by \textit{Crem.}, and $21.429\%$ probability of being vacant in the next season.
+		The largest vacant cacti have a \tom{$42.875\%$}{we should talk about this but I don't think these point values are very effectivein the text} probability of being tended by \textit{Liom.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, $28.571\%$ probability of being tended by \textit{Crem.}, and $21.429\%$ probability of being vacant in the next season.
 		Previously vacant cacti are most likely to stay vacant until the cacti reach about $10 log(m)^3$, at which point they are more likely to be tended by \textit{Liom.} ants in the next season. 
@@
 		Medium cacti follow the same patterns as large cacti in partner transitions.
+
+		\tom{}{What about the other vital rates?}
 
@@
 	\end{figure}
-		We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences. 
-		In the deterministic model, we found that simulations with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda_det}.
+		\tom{We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences.}{Just emphasizing again that the methods does not convey this.} 
+		In the deterministic model, we found that \tom{simulations}{you did not provide reproducible methods for these simulations} with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda_det}.
 		The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.
 		Similarly, only a few other scenarios had means within the interquartile range of this high partner diversity scenario. 
-		Honestly have no idea how much detail I should go into in this?
+		\tom{Honestly have no idea how much detail I should go into in this?
+}{It's less a matter of detail (though more would be good) and more about having a point.
+Your text seems to describe the results in a quasi-random, surface-level way.
+Bring more intention to your decisions about what to write here.
+What does a reader need to know to understand the story you are trying to tell?}
 
 		\section*{Discussion}
-		\subsection*{Vital Rates}
+		\tom{\subsection*{Vital Rates}}{I don't think these are the right subsection headings because they are simply redescribing results.}
 		\textbf{Ant partners have contrasting effects on vital rate processes of tree cholla.}
@@
 		As the tree cholla grew, the best partner changed from \textit{Crem.} ants, partners known for ***, to \textit{Liom.} ants, partners best known for defensive benefits to the cacti, particularly against the seed predators which most impact viability.
-		The difference in ant partners made a significant difference in the observed vital rates of the cacti, indicating that considering the interaction between the tree cholla and any individual partner would fail to capture the extent of the benefits to the cacti.
+		The difference in ant partners made a significant difference in the observed vital rates of the cacti, indicating that \tom{considering the interaction between the tree cholla and any individual partner would fail to capture the extent of the benefits}{great point and well made -- I am going to stop commenting on the rest of the discussion because I think it needs to be re-worked.
+We can discuss more.} to the cacti.
 		Like many other systems, pairwise perspectives do not fully encompass the complex impacts that multispecies mutualisms have on the focal mutualists vital rates.

### 2023-02-28 IPM Code/IPM.R
Update IPM.R

@@ -1,4 +1,4 @@
-#########################################################################################################
+g#########################################################################################################
 ##            This will be an IPM which allows you to choose how many ants are present
 #########################################################################################################
 #setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")

### 2023-02-28 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

@@
 
-Mutualisms are species interactions in which all \tom{involved participants}{Notice the redundancy here.
-You don't need ``involved''.} benefit.
-They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985}, and are notoriously \tom{context dependent}{You need to define what you mean by this.} \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-\tom{The benefits from these interactions are called rewards}{This is a pretty rudimentary definition, not necessary.}, but they require investment of resources or services into the interaction\cite{Bronstein2001}.
-The presence of mutualists can shift the available \tom{habitat, fitness, or resource uptake}{These are very different things.
-I don't see why this sentence is needed in the context of your study.} of a partner\cite{Stachowicz2001,West2007}.
-While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.\tom{}{I suggest you re-think whether most of this first paragraph is necessary.}
+Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
+They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing\cite{Willmer1997}, pollination and dispersal for food\cite{Sakai2002,Burns2004}, resource uptake for housing\cite{Holland2010}, etc.
+Mutualisms are typically considered more context dependent than other species interactions\cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
 
-Partner diversity in mutualisms \tom{often leads to variable fitness benefits}{Not sure what you are saying here: different partners provide different fitness benefits?
-Or diversity per se has fitness benefits?} for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
+\ali{}{I don't love this paragraph. Help. I particularly don't like how I defined multispecies mutualisms vs pairwise}
+Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
+The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
 For these reasons, pairwise studies cannot be accurately used to predict the outcomes of \tom{multi-species mutualisms}{I think you need to operationally define this.}\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.

### 2023-04-20 growth work.R
Update growth work.R

@@ -11,17 +11,17 @@ stan_data <- list(y = a,              ## the response variable
                   N = length(a),      ## the number of observations
                   vol = vol           ## size data
                   )
-stan_model <- stan(file = "Data Analysis/STAN Models/stan_prac.stan", 
+stan_null_model <- stan(file = "Data Analysis/STAN Models/stan_prac.stan", 
                    data = stan_data, warmup = 150, iter = 1000, chains = 3, 
                    cores = 3, thin = 1)
 ##### export the required variables so they may be used
 ## export
-summary(stan_model)
-stan_mu <- rstan::extract(stan_model, pars = c("beta0"))$beta0
-stan_outputs <- rstan::extract(stan_model, pars = c("alpha","sigma","beta0","beta1"))
+summary(stan_null_model)
+stan_mu <- rstan::extract(stan_null_model, pars = c("beta0"))$beta0
+stan_outputs <- rstan::extract(stan_null_model, pars = c("alpha","sigma","beta0","beta1"))
 write.csv(stan_outputs, "stan_outputs.csv")
 write.csv(stan_mu, "stan_mu.csv")
-y_rep <- rstan::extract(stan_model, pars = "y_rep")[["y_rep"]]
+y_rep <- rstan::extract(stan_null_model, pars = "y_rep")[["y_rep"]]
 samp100 <- sample(nrow(y_rep), 100)
 ## format
 others <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/stan_outputs.csv", header = TRUE,stringsAsFactors=T)
@@ -37,7 +37,7 @@ for(i in 1:1000){
 ## the convergence of the parameters
 png("null_convergence_try4.png")
 bayesplot::color_scheme_set(scheme = "pink")
-bayesplot::mcmc_trace(stan_model, pars = c("beta0","beta1","sigma","alpha")) 
+bayesplot::mcmc_trace(stan_null_model, pars = c("beta0","beta1","sigma","alpha")) 
 dev.off()
 ## converges fine
 ## the ability of the model to match the data
@@ -124,7 +124,7 @@ dev.off()
 ## the ability of the model to match the data
 png("overlay_data7.png")
 bayesplot::color_scheme_set(scheme = "pink")
-bayesplot::ppc_dens_overlay(as.vector(y), y_rep[samp100,])
+bayesplot::ppc_dens_overlay(as.vector(y), y_rep[,])
[…]

### 2023-04-20 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

@@
 The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of \tom{multi-species mutualisms}{I think you need to operationally define this.}\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
 Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}. 
+Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
+\ali{In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host}{Is this the best way to introduce this or should I talk about quality of benefits compared to cost of rewards?
+AKA some partners offer a net benefit when the benefits outweigh the cost of rewards to the partner ...
+etc.} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
+In cases where the partner offers benefits, \ali{the function, or type of benefits, offered by the partner can vary}{Is this enough detail or do I need to go into examples?} \cite{Stanton2003}.
 
-Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, \tom{some freeloaders, and even some parasites}{What is the difference between a freeloader and a parasite?} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, \tom{the function of the partner can vary}{I think I know what you mean but you should clarify for other readers what is the difference between quality and function.}, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
-\tom{When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.}{I think the content of this paragraph needs to be better organized.
-Describing ways in which partner species differ is conceptually very different than describing mechanisms by which partner diversity can be beneficial -- yet this paragraph seems to glom them together.
-The ``partners are different'' content seems to fit better in the previous paragraph.}
-When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in \tom{Figure}{This figure is great for talks but I don't recommend it for a publication.} \ref{fig:comp-samp}b, sampling effect can explain \tom{the positive effects}{When you say \emph{the} positive effects of partner diversity, does that imply that diversity effects are always positive?
-I think this is the first you've said that partner diversity is beneficial.} of partner diversity\cite{Batstone2018}.
+\begin{figure}[h]
+	\includegraphics[width=0.58\linewidth]{Complementarity copy.png}
+	\includegraphics[width=0.40\linewidth]{Sampling_Effect copy.png}
+	\caption{asdasdf}
+	\label{fig:comp-samp}
+\end{figure}
+\begin{figure}[h]
+	\includegraphics[width=0.60\linewidth]{Portfolio_Effect copy.png}
+	\caption{asdasdf}
+	\label{fig:portfolio}
+\end{figure}
+
+This variation in partners can lead to \ali{net benefits, no net benefits, or net costs for the focal mutualist}{Is this the best way to start this paragraph? If so should I include a figure panel?}.
+In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
+When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
 Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
@@
 
-\begin{figure}[h]
-	\includegraphics[width=0.58\linewidth]{complementarity.png}
-	\includegraphics[width=0.40\linewidth]{sampling_effect.png}
-	\caption{asdasdf}
-	\label{fig:comp-samp}
-\end{figure}
-\begin{figure}[h]
-	\includegraphics[width=0.60\linewidth]{portfolio_effect.png}
-	\caption{asdasdf}
-	\label{fig:portfolio}
-\end{figure}
-
-When multiple partners are present, \tom{partner turnover}{I suggest introducing this idea a little differently.
-Partner diversity can have different effects depending on whether it manifests all at once or sequentially.
-So this leads to the question of turnover.} is also important to consider, \tom{as partner identity has been shown to impact benefits in mutualisms}{You have made this point several times.}\cite{Djieto-Lordon2005, Ness2006, Bruna2014}.
+Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
 Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
 The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
-High degrees of partner turnover, \tom{also known as opportunism}{I've never seen partner turnover referred to as ``opportunism''.}, may offer \tom{buffering in the face of environmental variation}{This sounds a lot like portfolio effect, so I think you need to clarify how it relates.}\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, \tom{can increase the benefits}{This makes it sounds like diversity is harmful - since low turnover means low diversity.
-Is that an example of negative effects of partner diversity?} received in a mutualism\cite{Sachs2004}.
-\tom{The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
-For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.}{This is an excellent point.
-The second sentence here is a little clunky - see if you can fine-tune it.}
+\ali{The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.}{Not quite sure what to say here.
+}
+The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
+For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.
 
-\tom{Defensive}{Better transition from concepts to ants?} ant-plant mutualisms are common interactions in which plants provide \tom{extra-floral nectar}{I would just say ``food'' -- see Beltian bodies as another food resource} and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
-\tom{Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}}{This seems like too general/vague a statement to be useful here.}, types of rewards offered\cite{Beattie1985}, and \tom{quality of interaction}{What do you mean by this and how does it differ from types of rewards?} with their plant partners\cite{Rico-Gray1989,Palmer2010}, \tom{making these interactions great systems to study the effects of partner diversity}{A simpler and more to-the-point observation is that these interactions are typically generalized: most EFN plants have multiple ant partners} within mutualisms.
-\tom{While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} \st{in the literature}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.}{Good sentence.}
+\ali{Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns}{...
+Is this a better transition?
+I tried to tie it in a bit more conceptually.}, making these interactions great systems to study the effects of partner diversity.
+A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
+While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
 
-This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, \st{long-lived} EFN-bearing plant that associates with multiple species of ant partners.
+This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.
 Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
-\st{The} ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
-The multiple ant partners do not co-occur on \tom{individual}{} plants, but a single cactus may interact with multiple species over its lifetime.
-Some of the ant partners are more effective defenders than others, with \tom{\textit{Liom.}}{Notice that you have not yet introduced the species, let alone the abbreviation of its latin name, so a reader will be very confused here.
-Try to anticipate what your readers need to stay with you.} tended reproducing plants experiencing the lowest level of herbivory as shown in \tom{Figure \ref{fig:herb}}{You have put a result in the Introduction.
-Sometimes this is appropriate but I don't think it is here.
-You have not described any methods for where this result comes from, so it is not reproducible.
-Make this a result.}.
-\tom{This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.}{I think you can make this point without needing the figure, based on what we have already published from this system.
-It is also worth pointig out that one of the defenders was suggested to be net-negative because it deterred pollinators (Ohm and Miller 2014) but that study did not integrate demographic effects over the life cycle, which is what you do here and worth emphasizing as a source of novelty.}
+Ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
+The multiple ant partners do not co-occur on individual plants, but a single cactus may interact with multiple types of partners over its lifetime.
+While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these cacti.
 
@@
 
-In this \tom{paper}{Refer to this as a study and not a paper.} we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
+In this study we will answer three questions about the demographic effects of partner diversity across host lifetimes in multi-species mutualisms:
 \begin{enumerate}
 	\item{What are the contrasting demographic effects of multiple partners?}
-	\item{How do \tom{size}{size of what?} and partner identity impact the directionality of turnover between multiple partners?}
-	\item{\tom{What mechanisms explain the benefits}{This question pre-supposes that there are benefits of diversity.} of partner diversity in a multispecies mutualism?}
+	\item{How do host size and partner identity impact the directionality of turnover between multiple partners?}
+	\item{What mechanisms explain the effects of partner diversity in a multispecies mutualism?}
 \end{enumerate}
 
-\tom{To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.}{I think this sentence can be sharpened.
-I don't think that details like flower abortion and ant number are important to include here, but I also don't think this sentence really communicates the nature or volume of the data,}
-We analyzed the \tom{impacts of partner diversity}{The vital rate models do not reflect the impact of partner diversity, they are simply vital rate estimates fit with ant category as a factor -- be careful with how you present this.} on each individual vital rate (survival, growth, floral viability, etc.) with \tom{Bayesian models}{``Bayesian models'' does not carry much meaning.
-A model cannot be intrinsically Bayesian.
-The model is the model, and Bayesian analysis is one way to estimate it's parameters.}, which we used to parameterize an Integral Projection Model (IPM).
-In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but \tom{none have identified the mechanisms that explain the benefits of partner diversity}{Not sure I agree here.
-How are you ``identifying mechanisms'' in a way that Palmer et al.
-did not, for example?}.
+To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
+We analyzed the correlations between each individual vital rate (survival, growth, floral viability, etc.) and ant partners with Bayesian analysis of heirarchical models, which we used to parameterize an Integral Projection Model (IPM).
 

### 2023-05-18 Cholla_Analysis_Vital_Rates.R
Update Cholla_Analysis_Vital_Rates.R

@@ -136,7 +136,7 @@ alpha <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualis
 #alpha <- read.csv("/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/grow_alpha_skew.csv", header = TRUE,stringsAsFactors=T)
 alpha <- alpha[,c(-1)]
 alpha <- as.matrix(alpha)
-
+n_post_draws = 100
 y_sim <- matrix(NA, n_post_draws,length(y))
 for(i in 1:n_post_draws){
   y_sim[i,] <- rsn(n=length(y), xi = (xi[i,]), omega = (omega[i,]), alpha = alpha[i,])
@@ -144,23 +144,23 @@ for(i in 1:n_post_draws){
 View(y_sim)
 samp100 <- sample(nrow(y_sim), 100)
 ## Overlay Plots
-png(file = "grow_post_test.png")
+png(file = "grow_post.png")
 bayesplot::color_scheme_set(scheme = "pink")
 bayesplot::ppc_dens_overlay_grouped(y, y_sim[samp100,], group = ant)
 dev.off()
 ## Convergence Plots
-png("grow_conv_test.png")
+png("grow_conv.png")
 bayesplot::color_scheme_set(scheme = "pink")
-bayesplot::mcmc_trace(As.mcmc.list(fit_grow_skew, pars=c("beta0", "beta1","alpha","d_0","d_size")))
+bayesplot::mcmc_trace(As.mcmc.list(fit_grow_skew, pars=c("beta0", "beta1","a_0","a_size","d_0","d_size")))
 dev.off()
 ## They all converge
 ## Histograms
-png("grow_hist_post_test.png")
+png("grow_hist_post.png")
 bayesplot::color_scheme_set(scheme = "pink")
 bayesplot::ppc_stat_grouped(y, y_sim[samp100,], stat = "mean",group = ant)
 dev.off()
 #### Skew, Kurtosis, ETC.
-png("grow_moments_test.png")
+png("grow_moments.png")
 size_moments_ppc(growth_data, 
                  "logsize_t1", 
                  y_sim[samp100,], 

### 2023-05-18 Figures/Cholla_Analysis_Visualization_Vital_Rates.R
Update Cholla_Analysis_Visualization_Vital_Rates.R

@@ -308,23 +308,23 @@ grow_out <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutua
 ## Formulas
 ## Other
 
-y_other_mean_grow <- quantile(grow_out$beta0.2,0.5) + size_other * quantile(grow_out$beta1.2,0.5)
-y_other_low_grow <- quantile(grow_out$beta0.2,0.05) + size_other * quantile(grow_out$beta1.2,0.05)
-y_other_high_grow <- quantile(grow_out$beta0.2,0.95) + size_other * quantile(grow_out$beta1.2,0.95)
+y_other_mean_grow <- quantile(grow_out$beta0.2,0.5) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.5)
+y_other_low_grow <- quantile(grow_out$beta0.2,0.05) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.05)
+y_other_high_grow <- quantile(grow_out$beta0.2,0.95) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.95)
 ## Crem
-y_crem_mean_grow <- quantile(grow_out$beta0.3,0.5) + size_crem * quantile(grow_out$beta1.3,0.5)
-y_crem_low_grow <- quantile(grow_out$beta0.3,0.05) + size_crem * quantile(grow_out$beta1.3,0.05)
-y_crem_high_grow <- quantile(grow_out$beta0.3,0.95) + size_crem * quantile(grow_out$beta1.3,0.95)
+y_crem_mean_grow <- quantile(grow_out$beta0.3,0.5) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.5)
+y_crem_low_grow <- quantile(grow_out$beta0.3,0.05) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.05)
+y_crem_high_grow <- quantile(grow_out$beta0.3,0.95) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.95)
 ## Liom
-y_liom_mean_grow <- quantile(grow_out$beta0.4,0.5) + size_liom * quantile(grow_out$beta1.4,0.5)
-y_liom_low_grow <- quantile(grow_out$beta0.4,0.05) + size_liom * quantile(grow_out$beta1.4,0.05)
[…]

### 2023-05-18 Manuscript/manuscript_draft.tex
May Updates

@@
 
-\ali{}{I don't love this paragraph. Help. I particularly don't like how I defined multispecies mutualisms vs pairwise}
 Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
@@
 This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.
-Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
-Ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
-The multiple ant partners do not co-occur on individual plants, but a single cactus may interact with multiple types of partners over its lifetime.
-While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these cacti.
+Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates\cite{Miller2009}. 
+Ant partners defend the cholla from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
+The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.
+While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
 
@@
 In this study we will answer three questions about the demographic effects of partner diversity across host lifetimes in multi-species mutualisms:
+
 \begin{enumerate}
@@
 To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We analyzed the correlations between each individual vital rate (survival, growth, floral viability, etc.) and ant partners with Bayesian analysis of heirarchical models, which we used to parameterize an Integral Projection Model (IPM).
+We used vital rate functions to evaluate the first two questions. 
+We then used the composite integral projection model (IPM) to evaluate how vital rate differences affect intrinsic population growth rates and identify mechanisms at play.
 
@@
 \subsection*{Study System}
-
-\st{To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004.} 
-This study was conducted in the Los \tom{Pinos}{there is an accent on the n} mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico \tom{(34*20’5.3’’N, 106*37’53.2’’W)}{I would not give lat-long here.
-We should provied lat-long of each plot as part of the full data that we make available.}.
+This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
 This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
-\st{The} tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
-These arborescent cacti produce cylindrical segments with large spines. 
+Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
+These arborescent cholla produce cylindrical segments with large spines. 
 In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
 While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
-Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
-Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative and reproductive structures\cite{Ness2006,Oliveira1999}.
+Tree cholla generally reach at least 9 years of age before beginning to reproduce \cite{Ohm2014}.
+Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures\cite{Ness2006,Oliveira1999}.
 
 This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum \tom{(Liom.)}{I know I have used these abbreviations in previous papers but I would not recommend doing so again.
-More professional and appropriate to refer to \emph{L.
-apiculatum}, etc.}, } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \tom{\textit{Phenogaster}}{Incorrect spelling.} \tom{ant}{Obviously it's an ant.}, and \textit{Camponotus spp.
-(Camp.)}.
+At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, a \tom{\textit{Phenogaster spp.}}{Incorrect spelling.}, and \textit{Camponotus spp.}.
+\textit{L.
+apiculatum} are the most frequent visitors with 25\% - 60\% of tree cholla tended by these ants, followed by \textit{C.
+opuntiae} visiting between 0\% - 20\% of cacti\cite{Donald2022} depending on the year.
+Up to 80\% of cacti remain vacant in any given year.
 
-\begin{figure}[h]
-	\includegraphics[width=\linewidth]{Timeseries.png}
-	\caption{asdfasdfsadfasdf}
-	\label{fig:timeseries}
-\end{figure}
+These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}.
+Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
+At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
+They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset\cite{Ohm2014}. 
+Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended\cite{Miller2014}. 
+In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti \tom{annually}{Unclear what you mean by ``annually''?
-Think about the most technically correct and most efficient way to communicate these results.}, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in \tom{Figure}{Again, you have slipped in a result without providing any methods.
-There is a similar figure in Donald and Miller 2022, so you can cite that paper when you report frequencies of association.} \ref{fig:timeseries}.
-\tom{\st{It is important to note that not all cacti are visited every season, they can remain vacant, in fact} }{Notice that this part of the sentence does not provide any information that is not in the second part.}up to 80\% of cacti are vacant in any given year.
-
-These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}\tom{, rather}{Grammatically awkward.
-I would start a new sentence Instead, each cactus} each cactus is visited by a single ant species for the duration of a season, and the \tom{visitor}{What do you mean by visitor?
-Is each individual ant a visitor?} can change from one season to the next.
-\tom{Across the entire lifespan cacti may interact with multiple species of partners. }{This is implied by the sentences above so there is no new information here.}
-At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cacti every day during the season from early morning *** to sunset *** \tom{abandoning the cacti at night for their nests}{Incorrect. See Ohm and Miller 2014.}. 
-In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
-\tom{The process by which ants are sorted out onto different cacti is somewhat of a mystery}{too colloquial}, but we know several facts which likely influence the process. 
-Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
-Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more \tom{desirable}{I don't yhink we should speculate about ants' desires.}.
-Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in \tom{going after}{too colloquial} more desirable cacti. 
-Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
-These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed.
-\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}\tom{}{Yes this is what I was thinnking as i was reading.
-Much of this fall out of the results of your analysis.
-I would significantly reduce this paragraph.
-Think about what your readers need to know to process what's to come.}
-\st{These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.}
-
-There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments \tom{throughout the entire range}{range info not really pertinent} of the tree cholla\cite{Mann1969}.
-\tom{More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
-These include an unidentified weevil, a cactus bug, and a seed predator. }{Notice how these sentences add no information to the paragraph.}
-The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
-The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
-The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.\tom{}{I think it is also important to summarize what is known about ant defense against these insects from previous papers.
-The fact that we have experimental evidence that ants provide defense is important to include, because your data are all observational.}
-
-\tom{\section*{Methods}}{Some duplication here...?}
-\subsection*{Study System}
-\subsection*{Natural History}
-To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
-This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
-This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
-The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
-These arborescent cacti produce cylindrical segments with large spines. 
-In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
-While most plants produce new segments every season, only those which are mature enough produce flower buds. 
-Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
-Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.
-
-This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) } and \textit{Camponotus spp.
-(Camp.)}.
-
-\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year \ref{Fig:Ant_freq}.
-It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.
-
-These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next.
-Across the entire lifespan cacti may interact with multiple species of partners. 
-At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
-In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
-The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
-Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
-Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
-Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
-Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
-These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed.
-\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
-These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.
-
-There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}.
-More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
-These include an unidentified weevil, a cactus bug, and a seed predator. 
-The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
-The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
-The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller 2009}.
+There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments\cite{Mann1969}. 
+An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
+A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \cite{Miller2006}.
+A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
+These predators can have significant negative impacts on the fitness of individual cholla and depress population growth\cite{Miller2009}.
+There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators\cite{Miller2007}.
 
 	\subsection*{Data}
+The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
+Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
+The final two plots were added to this census from 2011 onwards. 
+Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
+For each cholla, we recorded plant survival from the last survey to the current survey. 
+We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
+We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
+We also recorded the species of herbivore and the number present on the plant.
 
-	The data collected \tom{on these cacti}{notice how these words add nothing} are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
-	\tom{The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.}{this seems to contradict what you just said about eight plots.}
-	Six of the 30 $\times$ 30 meter plots \tom{were created with naturally occurring plants}{this sounds funny, meaning unclear}, censused from 2009 to \tom{2014}{Not sure why you have these ending in 2014.
-These are plots 1-6, which we continue to collect data from}.
-	The final two plots were added to this census from 2011 onwards. 
-	The distribution of these eight plots are shown in \tom{figure}{you probably don't need a plot map - maybe in the supplement} \ref{fig:map}.
-		Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
-		For each plant, we recorded the height (cm), maximum crown width, and \st{perpendicular }crown width perpendicular to the maximum, which are used to calculate plant volume (\tom{cm3}{make this a superscript}) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-		We recorded plant \tom{survival}{I would start with survival because if it's dead nothing else is recorded} from the last survey to the current survey\st{ as a binary data point, 0 (died) or 1 (survived)}.
-		We recorded the total number of flower buds, including how many were aborted and how many were not\st{, which we used to calculate the viability rates of the cacti}. 
-		We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
-		We also recorded the species of herbivore or \tom{seed predator}{we do not record seed predator data}, the number present on the plant, and \tom{if there is any clear, new evidence of herbivory}{beware that this has not been very consistent} on the tree cholla segments.
+	\subsection*{Statistical Modeling -- Construction}
+With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
+Many of the vital rates are estimated as a function of plant size, ant partner, or both.
+Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
+The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
+Unless otherwise mentioned, all models use vague priors. 
+The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
+The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
+Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these. 
+The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
+The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
+The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
+Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion. 
+The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
+Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
+The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
+With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
+With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
+With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds.
 
-		In addition to this primary dataset, I used \tom{several other datasets}{I don't think you need to say much here, you can cite prevous papers when you describe these parts of the IPM} throughout the analysis in this paper.
-Look at how Tom has described these.
-Cite these other papers which this data comes from.
-		The second dataset was *** germination rates.
-		The third dataset was *** pre-census survival.
-		The fourth dataset was *** $JO_fruit_data_final_dropplant0.csv$.
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
+For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
+The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
+We did not thin the chains, thus all samples were retained. 
+To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
+To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
+Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
+All estimated parameters are described in table 1.
+Data and code for all vital rate models is included in the supplemental information.
 
-		\subsection*{Demographic Modeling}
+\subsection*{Integral Projection Model Construction}
+Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
+We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
+Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
+There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
+This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
 
-		The \tom{statistical models described above}{they are actually below - but i would put them above.} parameterize the Integral Projection Model (IPM) that we used to estimate population growth under various partner conditions.
-		IPMs are used to estimate fitness of a population across a \tom{continuous variable}{you actually have both continuous and discrete variables}, \tom{rather than using a stage- or size-specific variable}{not an accurate description of how IPMs differ from MPMs} to categorize the population.
-		This IPM was used to estimate the \tom{growth rate }{not just growth rate, but growthrate as a function of partner identity and diversity -- important to include}of the tree cholla population, \tom{effectively}{why effectively?} a quantitative measure of the fitness of the population.
+Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
+The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
 
-		Following previous studies, we modeled the life cycle of \tom{\textit{C.
-imbricata}}{sometimes you say cholla, sometimes cacti, sometimes the latin name -- stay consistent} using continuously size-structured plants,\tom{$n_i(x)$}{what is n?
-what is i?
-what is x?}, and two discrete \tom{seed banks}{there is no beta in the model below.
-what is t?} ($\beta_{1,t}$ and $\beta_{2,t}$) corresponding to 1 and 2-year old seeds.
 		$$
-		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) A_i(x) F(x)n_i(x) dx \\
+		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
 		$$
@@
 
-		The functions $P$, $F$, and $A$ give the probability of flowering, the number of flowerbuds produced, and the proportion of flowerbuds which will create seeds. 
-		Each of these functions, \st{estimated by a Bayesian model} \tom{calculates}{the function does not do any calculation} these values for an $x$-sized plant in year $t$. 
-		The proportion of flowerbuds which will produce seeds ($A$) is also dependent on the ant species present on the plant $i$ in year $t$. 
-		The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the 1-year old seed bank. 
+The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
+The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
+The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
 Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
-		Plants can recruit out of the 1-year seed bank with the probability of $\gamma_1$ or transition to the 2-year seed bank with a probability of $1 - \gamma_1$. 
-		Seeds in the 2-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
+Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
+Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
 
-		The size dynamics of the plants are given by:
+The size dynamics of the cholla are given by:
 
@@
 
-		The final equation gives the \tom{stochastic population composition }{what does this mean?}in terms of size and ant state of the \textit{C.
-imbracata} population $n$ in year $t+1$ based on the vital rates and size $n$ of the population in year $t$.
-		The first term gives the recruitment from 1 and 2-year seedbanks to size $y$.
-		$\eta(y)$ gives the seedling size distribution, which is assumed to be normally distributed, and $\omega$ gives the proportion of seedlings which survive from germination (late summer) to the census (May).
-		The second term gives the changes in the population of the cacti which are not recruits. 
-		The functions $S$ and $G$ give the probabilities of surviving from year $t$ to $t+1$ and growing to size $y$ from year $t$ to year $t+1$, respectively. 
-		Each depends on the size $x$ in year $t$ and the ant state $j$ in year $t+1$. 
-		Finally, $\tau_{ij}$ is the probability of a cactus which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.\tom{}{the ants dynamics of the model need to be presented more thoroughly and clearly.}
+The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
+The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
+$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
+Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
 
-		\tom{}{This paragraph should get its own subsection.
-This is about model analysis, not model structure.}\tom{Because many of the vital rates were ant-specific}{you need to explain why some are and some are not}, we were able to consider the composition of the population with the presence of any single ant species, or any combination of ant species (with vacancy always included).
-		These options include: complete vacancy; \textit{Liom.} and vacancy; \textit{Crem.} and vacancy; other and vacancy; \textit{Liom.}, \textit{Crem.}, and vacancy, \textit{Liom.}, other, and vacancy; \textit{Crem.}, other, and vacancy; and all ant partners and vacancy.
-		We used this option to calculate the distribution of stochastic population fitness for each combination of ant partners to determine if there is benefit from partner diversity and if there is an optimal combination of partners for this system.
-		We then used these fitnesses and \tom{the yearly break-downs of the model}{what does this mean?} to determine if there was \tom{evidence of any biodiversity ecosystem function mechanisms to explain any partner diversity benefits}{you never describe any criteria for detecting the different mechanisms.
-What does one need to see in data to conclude that one or another is operating?}.
+\subsection*{Deterministic IPM Analysis}
+Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
+Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix \as shown in figure ***. 
+The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
+We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
+We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
+apiculatum} and vacancy; \textit{C.
+opuntiae} and vacancy; other and vacancy; \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and vacancy, \textit{L.
+apiculatum}, other, and vacancy; \textit{C.
+opuntiae}, other, and vacancy; and all ant partners and vacancy.
+The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
+Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
+Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner.
 
-		\subsection*{Statistical Modeling}
-		At the start of this section say for all models unless otherwise noted we used vague priors. 
-		Also need to introduce the names of these functions ex: survival = s(x)
-		With the primary dataset above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework, meaning the output is a distribution of \tom{the interaction between various factors on the highlighted variable}{confusing} rather than a maximum likelihood estimate with both fixed and random effects, which predicts the probability of a given process based on the size of the cacti, the presence of different ant partners, the year, and the plot.\tom{}{this is a confusing sentence.
-try to sharpen this.}
-		Unless otherwise mentioned, all models used vague priors. 
-		The growth model ($G_j(y,x)$) estimated the \tom{growth rate}{this is not what you modeled} of cacti, with fixed effects for the size and ant partner and random effects for plot and year, using a Normal distribution, with standard deviation dependent on the size of the cactus \tom{also with a Normal distribution}{??}.
-		The survival model ($S_j(x)$) estimated the probability of survival, with fixed effects for the size of the cactus and ant partner and random effects for plot and year, using a Bernoulli distribution.
-		The reproduction model ($P(x)$) estimated the probability of reproducing each year, with fixed effects for the size of the cactus and random effects for plot and year, using a Bernoulli distribution.
-		The total flowers model ($F(x)$) estimated the total flowers produced by a plant, with fixed effects for its size and random effects for plot and year, using a \tom{Truncated}{truncated how?} Negative Binomial distribution.
-		The viability model ($A_i(x)$) estimated the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects for the size of the cactus and the \tom{presence of ant partners}{this is not the variable in the model} and random effects for plot and year, using a Binomial distribution.
-		The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects for the size of the cactus and the previous ant partner and random effects for plot and year, using a Multinomial distribution.
-		The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a year, with no fixed or random effects, using a Normal distribution. 
-		With the \tom{second dataset}{what second data set? same question for 3 and 4} above, we fit a two Bayesian generalized linear models for the 
-		probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) using a Binomial distribution.
-		With the third dataset above, we fit a Bayesian generalized linear model for the probability of a seedling, which has sprouted each year, surviving to May ($\delta$) when we visit (accounting for missed mortality events), with fixed effects for the previous size and transect random effects, using a Bernoulli distribution.
-		With the fourth dataset above, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cactus ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+\subsection*{Stochastic IPM Analysis}
+The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
+The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
+In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
+For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
+With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
+This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
+We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
+These two relative differences allow us to determine if portfolio effect is at play.
+Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
 
-		These Bayesian GLMMs were run with STAN through version 4.0.2  of R using \tom{1000 iterations}{do you mean 10000?
-1000 is not very many}, \tom{150 warmups, 3 chains, and 1 thins.}{``warmups'' and ``thins'' are not the right word choices}
-		To assess the convergence of our models we checked the trace of each mcmc chain (the convergence graphs are shown in the supplemental documents in Figures ***).
-		To assess the fit we simulated data from the models and compared this to the real data (the fit plots are shown in the supplemental documents in Figures ***).
+% I am creating a table here to include all parameter estimations and descriptions
+\begin{table}[]
+\begin{tabular}{l|l|l}
+\textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
+\hline
+growth xi intercept vacant $\Beta_01^g$ & & \\
+growth xi intercept other $\Beta_02^g$ & & \\
+growth xi intercept \textit{C. opuntiae} $\Beta_03^g$ & & \\
+growth xi intercept \textit{L. apiculatum} $\Beta_04^g$ & & \\
+growth xi size dependent vacant $\Beta_11^g$ & & \\
+growth xi size dependent other $\Beta_12^g$ & & \\
+growth xi size dependent \textit{C. opuntiae} $\Beta_13^g$ & & \\
+growth xi size dependent \textit{L. apiculatum} $\Beta_14^g$ & & \\
+growth omega intercept $\omega_0^g$ & & \\
+growth omega size dependent $\omega_1^g$ & & \\
+growth alpha intercept $\alpha_0^g$ & & \\
+growth alpha size dependent $\alpha_1^g$ & & \\
+1-year germination intercept $\alpha^{\gamma_1}$ & & \\
+2-year germination intercept $\alpha^{\gamma_2}$ & & \\
 
-		\tom{With these models, we investigated the impacts of different ant species on the processes of vital rates and the probabilities of individual cacti interacting with each ant species.
-		To address our first goal, we compared the effects of ant partners on the growth rate, survival rate, and the viability of flowers across sizes. 
-		We also broke these effects down by every year of data to see if the impacts of ant partners varied across years. 
-		To address the second goal, we compared the probability of being visited by a given ant based on each possible previous partner to note any patterns of ant tending. 
-		We also broke these interactions down by every year of data to see if the patterns changed across years.
-}{This needs a much more thorough explanation.
-You have not described any methods for a stochastic analysis.
-You need a new section here explaining these methods in a way that is reproducible.
-Keep in mind that reproducibility is the standard we are aiming for, and in the current draft, just about all of the methods fall short of that standard.}
-
-		\vspace{1cm}
+\end{tabular}
+\end{table}
 

### 2023-05-18 IPM Code/IPM_Stochastic_Post.R

@@ -1012,58 +1012,98 @@ bigmatrix<-function(params,lower,upper,matsize,scenario,grow_rfx1,grow_rfx2,grow
 # lambdaS Simulations for different Years Rands #########################################################
 # This is the original code as taken from the Climate IPM Online ########################################
 #########################################################################################################
-# lambdaSim=function(params,                                  ## parameters
-#                    grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4, ## growth model year rfx
-#                    surv_rfx1,surv_rfx2,surv_rfx3,surv_rfx4, ## survival model year rfx
-#                    flow_rfx,                                ## flower model year rfx
-#                    repro_rfx,                               ## repro model year rfx
-#                    viab_rfx1,viab_rfx2,viab_rfx3,viab_rfx4, ## viability model year rfx
-#                    max_yrs,                                 ## the # years you want to iterate
-#                    matsize,                                 ## size of transition matrix
-#                    scenario,                                ## partner diversity scenario
-#                    lower,upper                              ## extensions to avoid eviction
-# ){           
-#   
-#   ## Create an actual matrix filled with 0 of the right size based on scenarios
-#   if(scenario == "none"){K_t <- matrix(0,matsize+2,matsize+2)}
-#   if(scenario == "cremvac"|scenario == "liomvac"|scenario == "othervac"){K_t <- matrix(0,2*matsize+2,2*matsize+2)}
-#   if(scenario == "liomcremvac"|scenario == "liomvacother"|scenario == "othercremvac"){K_t <- matrix(0,3*matsize+2,3*matsize+2)}
-#   if(scenario == "all"){K_t <- matrix(0,4*matsize+2,4*matsize+2)}
-#   matdim        <- ncol(K_t)
-#   rtracker      <- matrix(rep(0,max_yrs*10),nrow = 10)  ## Empty vector to store growth rates in
-#   n0            <- rep(1/matdim,matdim)  ## Create dummy initial growth rate vector that sums to 1
-# 
-# 
-#   for(t in 1:max_yrs){ ## In this loop I call the IPMmat and store it in the K_t matrix then
[…]

### 2023-05-29 Cholla_Analysis_Vital_Rates.R
Update Cholla_Analysis_Vital_Rates.R

@@ -1,232 +1,180 @@
-#setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
 #######################################################################################################
 ##
 ##                  The purpose of this file is to run each vital rate sub model separately,
 ##                          save the outputs, and check the posterior distributions 
 ##
 #######################################################################################################
-cactus <- read.csv("cholla_demography_20042021_cleaned.csv", header = TRUE,stringsAsFactors=T)
+
+## First read the data in 
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
+#setwd("C:/Users/tm9/Dropbox/github/ant_cactus_demography")
+cactus <- read.csv("cholla_demography_20042021_cleaned.csv", header = TRUE,stringsAsFactors=T)
 
 ##############################################################################################
-#### Growth Model -- What size will the cacti be next time step? #############################
+##
+##   Skew Growth Model -- What size will the cacti be next time step?
+##
 ##############################################################################################
-# growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
-# growth_data <- na.omit(growth_data_orig)
-# ## Lose 2032 rows (due to plant death & recruit status)
-# nrow(growth_data_orig)
-# nrow(growth_data)
-# # check that you are happy with the subsetting
-# plot(growth_data$logsize_t, growth_data$logsize_t1)
-# points((cactus$logsize_t), (cactus$logsize_t1), col = "red")
-# levels(growth_data$ant_t)
-# ## Create Stan Data for all ant states
-# stan_data_grow <- list(N = nrow(growth_data), ## number of observations
-#                        vol = (growth_data$logsize_t), ## predictors volume
-#                        y_grow = (growth_data$logsize_t1), ## response survival next year
-#                        ant = as.integer(as.factor(growth_data$ant_t)),## predictors ants
-#                        K = 4, ## number of ant states
-#                        N_Year = max(as.integer(as.factor(growth_data$Year_t))), ## number of years
[…]

### 2023-05-29 IPM Code/Analysis.R
Update Analysis.R

@@ -572,49 +572,34 @@ dev.off()
 ##############################################################################
 scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
 max_scenario = length(scenario)
-max_iter = 100
+max_rep = 100
+scenario = c("none","all")
 max_yrs = 10
-lams <- matrix(rep(NA,max_iter*max_scenario), nrow = max_iter)
+lam <- matrix(nrow = max_rep, ncol = max_scenario)
 for(n in 1:max_scenario){
-  ## must create kt matrix based on the scenario
-  if(scenario[n] == "none"){K_t <- matrix(0,matsize+2,matsize+2)}
-  if(scenario[n] == "cremvac"|scenario[n] == "liomvac"|scenario[n] == "othervac"){K_t <- matrix(0,2*matsize+2,2*matsize+2)}
-  if(scenario[n] == "liomcremvac"|scenario[n] == "liomvacother"|scenario[n] == "othercremvac"){K_t <- matrix(0,3*matsize+2,3*matsize+2)}
-  if(scenario[n] == "all"){K_t <- matrix(0,4*matsize+2,4*matsize+2)}
-  ## reinitialize matdim based on the new KT
-  matdim <- ncol(K_t)
-  for(m in 1:max_iter){  
-    ## Must reinitialize rtracker and n0 for every iteration of parameters
-    rtracker      <- c(rep(0,max_yrs))  ## Empty vector to store growth rates in 
-    n0            <- rep(1/matdim,matdim)  ## Create dummy initial growth rate vector that sums to 1
-    for(t in 1:max_yrs){ ## In this loop I call the IPMmat and store it in the K_t matrix then 
-      ## scale this to the stochastic growth rate
-      ## Randomly sample the years we have data for by calling column r in all matricies of 
-      ## the year random effects
-      ## Must sample a new r value for every year of the simulation we run
-      r <- sample(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17),1,replace = TRUE,prob = NULL)
-      ## Call and store matrix for every year of the simulation
-      K_t[,]<-bigmatrix(params[m,],lower,upper,matsize,scenario[n],
-                        grow_rfx1[m,r],grow_rfx2[m,r],grow_rfx3[m,r],grow_rfx4[m,r],
[…]

### 2023-05-29 IPM Code/IPM_Stochastic_Post.R
Update IPM_Stochastic_Post.R

@@ -44,24 +44,24 @@ gxy<-function(x,y,i,params,grow_rfx1,grow_rfx2,grow_rfx3,grow_rfx4){
   if(i == "vacant"){ return(g_vac)}
 }
 
-##Check that it works properly
-i = c("vacant","crem","liom","other")
-x = c(-1,-5,3,4)
-y = c(-1,-4,3,4)
-g <- matrix(NA,ncol = length(i), nrow = 100)
-l <- list()
-
-for(a in 1:17){ ## year
-for(m in 1:nrow(params)){ ## iteration
-  for(n in seq(1:length(i))){ ## input info
-    xb=pmin(pmax(x,cholla_min),cholla_max) #Transforms all values below/above limits in min/max size (So the params are the minimums and maximums of size?)
-    g[m,n] <- gxy(x[n],y[n],i[n],params[m,],grow_rfx1[m,a],grow_rfx2[m,a],grow_rfx3[m,a],grow_rfx4[m,a])
-    }
-}
-  l[[a]] <- g
-}
-g
-l
+# ##Check that it works properly
+# i = c("vacant","crem","liom","other")
+# x = c(-1,-5,3,4)
+# y = c(-1,-4,3,4)
+# g <- matrix(NA,ncol = length(i), nrow = 100)
+# l <- list()
+# 
+# for(a in 1:17){ ## year
+# for(m in 1:nrow(params)){ ## iteration
+#   for(n in seq(1:length(i))){ ## input info
+#     xb=pmin(pmax(x,cholla_min),cholla_max) #Transforms all values below/above limits in min/max size (So the params are the minimums and maximums of size?)
+#     g[m,n] <- gxy(x[n],y[n],i[n],params[m,],grow_rfx1[m,a],grow_rfx2[m,a],grow_rfx3[m,a],grow_rfx4[m,a])
+#     }
+# }
+#   l[[a]] <- g
+# }
+# g
+# l
 
 #########################################################################################################
 ## SURVIVAL AT SIZE X. Returns the probability of survival of a cactus based on size and ant state   ####
@@ -836,22 +836,22 @@ bigmatrix.3 <- function(params,lower,upper,matsize,scenario,grow_rfx1,grow_rfx2,
     return(list(IPMmat = IPMmat, Tmat = Tmat, Fmat = Fmat))
   }
 }
-# Scenario options are "liomvacother", "liomcremother", "liomcremvac", "othercremvac"
[…]

### 2023-05-29 Manuscript/AmNat_MS_template (1).tex
Create AmNat_MS_template (1).tex

\documentclass[11pt]{article}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

%%%%%%%%%%%%%%%%%%%%%
% Line numbering
%%%%%%%%%%%%%%%%%%%%%
%
% Please use line numbering with your initial submission and
% subsequent revisions. After acceptance, please turn line numbering
% off by adding percent signs to the lines %\usepackage{lineno} and
% to %\linenumbers{} and %\modulolinenumbers[3] below.
%
% To avoid line numbering being thrown off around math environments,
% the math environments have to be wrapped using
% \begin{linenomath*} and \end{linenomath*}
%
% (Thanks to Vlastimil Krivan for pointing this out to us!)

\title{Template and Guidelines for Using \LaTeX{} in \textit{The~American~Naturalist} }

% This version of the LaTeX template was last updated on
% May 11, 2023.

%%%%%%%%%%%%%%%%%%%%%
% Authorship
%%%%%%%%%%%%%%%%%%%%%
% Please remove authorship information while your paper is under review,
% unless you wish to waive your anonymity under double-blind review. You
% will need to add this information back in to your final files after
% acceptance.

\author{Owen E. Cook$^{1,\ast}$ \\ 
Generic H. Collaborator$^{2,\dag}$ \\ 
Additional Q. Expert$^{3}$}

\date{}

\begin{document}

\maketitle

\noindent{} 1. University of Chicago, Chicago, Illinois 60637;

\noindent{} 2. University of Toronto, Toronto, Ontario M5S 1A5, Canada;

\noindent{} 3. Middle Eastern Technical University, Çankaya, Ankara 06800, Turkey.

\noindent{} $\ast$ Corresponding author; e-mail: amnat@uchicago.edu.

\noindent{} $\dag$ Deceased.

\bigskip

\textit{Manuscript elements}: Figure~1, figure~2, table~1, appendix~A (for print; including figure~A1, figure~A2, and table~A1), supplemental PDF. Figure~2 is to print in color.

\bigskip

\textit{Keywords}: Examples, model, template, guidelines.

\bigskip

\textit{Manuscript type}: Article. %Or note, natural history miscellany note, comment, reply, invited symposium, featured topic, or historical perspective.

\bigskip

\noindent{\footnotesize Prepared using the suggested \LaTeX{} template for \textit{Am.\ Nat.}}

%\linenumbers{}
%\modulolinenumbers[3]

\newpage{}

\section*{Abstract}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed non risus. Suspendisse lectus tortor, dignissim sit amet, adipiscing nec, ultricies sed, dolor. Cras elementum ultrices diam. Praesent quis dolor in dolor molestie cursus et ac nisi. Vestibulum ante purus, semper eget est vitae, vehicula ornare nisl. Morbi efficitur euismod enim, nec feugiat tellus cursus eget.

\newpage{}

\section*{Introduction}

% The journal does not have numbered sections in the main portion of
% articles. Please refrain from using section references (à la
% section~\ref{section:CountingOwlEggs}), and refer to sections by name
% (e.g. section ``Counting Owl Eggs'').

The quick red fox jumps over the lazy brown dog. Furthermore, the quick brown fox jumps over the lazy red dog. In addition, the quick R\"{u}ppell's fox (\textit{Vulpes rueppellii}) jumps over the lazy golden retriever.

\section*{Methods}

Lorem ipsum ut velit mauris, egestas sed, gravida nec, ornare ut, mi. Aenean ut orci vel massa suscipit pulvinar. Nulla sollicitudin. Fusce varius, ligula non tempus aliquam, nunc turpis ullamcorper nibh, in tempus sapien eros vitae ligula (\citealt{OriginSpec}).

\subsection*{Second-order heading}

Lorem ipsum nulla facilisi, pace \citet{LemKapEx07}. Etiam semper, orci sit amet facilisis interdum, tellus nunc consequat erat, quis viverra nisi diam ut metus. Pellentesque cursus, sapien malesuada euismod iaculis, mauris purus interdum diam, vel vestibulum justo enim vitae tellus. Nunc interdum lorem sit amet diam volutpat tristique. Quisque pulvinar ac metus commodo lacinia (\citealt{Ing11,Xiao2015}).

\subsubsection*{Third-order heading}

Usually two or three levels of heading will be all you need. Journal style even permits a fourth level in case you need it.

\paragraph*{Fourth-order heading}
The quick red fox jumps over the lazy brown dog in this paragraph as well. Donec mauris nibh, volutpat vehicula viverra at, iaculis congue sem. Praesent eget erat rhoncus erat sollicitudin volutpat.

\begin{equation}
{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
\end{equation}

Praesent quis dolor in dolor molestie cursus et ac nisi. Vestibulum ante purus, semper eget est vitae, vehicula ornare nisl. Morbi efficitur euismod enim, nec feugiat tellus cursus eget. Donec mauris nibh, volutpat vehicula viverra at, iaculis congue sem.

\subsection*{Another second-order heading}

As \citet{Xiao2015} argued, phasellus porttitor eros et ante condimentum, eget facilisis orci condimentum. Nulla facilisi. Proin placerat elit blandit, euismod dolor nec, dapibus diam. Mauris posuere malesuada lacus, at elementum lacus auctor eu (fig~\ref{Fig:Jumps}A).

\section*{Results}

Lorem ipsum dolor sit amet. Aenean pulvinar malesuada commodo (see \citealt{DavisEtAl2011}; table~\ref{Table:Founders}). Sed aliquet mauris odio, in tristique dui egestas a. Etiam eu malesuada quam. Suspendisse tincidunt eu erat sit amet vulputate. Duis at arcu et nisl dictum mattis. Maecenas vel cursus ante. Cras eleifend elit nec velit sollicitudin fermentum in ac mauris. Pellentesque rutrum magna vel elit maximus hendrerit.

\subsection*{The height of the jump}

Aenean eu pellentesque quam (fig.~\ref{Fig:OkapiHorn}). Nam pellentesque augue eu finibus lacinia. Nullam nec justo vitae odio imperdiet rhoncus vitae vitae quam. Pellentesque porttitor metus et lectus ornare, ac cursus urna efficitur (fig~\ref{Fig:Jumps}B).

\subsection*{The laziness of the dog}

Example paragraph with embedded references (video~\ref{VideoExample}, fig.~\ref{Fig:AnotherFigure}). 
If you have deposited data to Dryad, it is advisable to cite them somewhere in the main text (usually in the Methods or Results sections). A sentence like the following will do: All data are available in the Dryad Digital Repository (\citealt{CookEtAl2015}).

\section*{Discussion}

Nam pulvinar lorem at lorem ultrices, vel accumsan massa feugiat (\citealt{Ing11}). Proin tristique velit eget lacus iaculis, in pellentesque nulla varius. Phasellus sodales est odio, eu pulvinar magna pellentesque eu. Sed ut lobortis eros. Aliquam eget metus turpis. Sed et convallis lectus, id tincidunt enim. In porta nibh ut lacus feugiat, non consequat orci rhoncus. Morbi blandit at augue nec tempor. Sed fringilla ipsum ut justo viverra, ut euismod nisi gravida.

Curabitur non posuere augue, id suscipit orci. Nunc luctus accumsan aliquam. Cras egestas turpis vitae nisl vulputate interdum. Donec pellentesque libero egestas tortor pharetra laoreet. Phasellus facilisis auctor ligula, eu sollicitudin mi sagittis non.

\section*{Conclusion}

Duis pharetra enim at libero cursus, eu commodo mi vestibulum. Nullam eget velit nec lectus viverra sodales. Suspendisse egestas, eros at dictum tincidunt, mi orci laoreet libero, eget rutrum sapien arcu blandit odio.

%%%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%%%
% You may wish to remove the Acknowledgments section while your paper 
% is under review (unless you wish to waive your anonymity under
% double-blind review) if the Acknowledgments reveal your identity.
% If you remove this section, you will need to add it back in to your
% final files after acceptance.

\section*{Acknowledgments}

OEC would like to thank Madlen Wilmes, Gyuri Barab\'{a}s, Flo D\'{e}barre, Vlastimil K\v{r}ivan, and Greg Dwyer for their comments and suggestions on this template.

%%%%%%%%%%%%%%%%%%%%%
% Statement of Authorship
%%%%%%%%%%%%%%%%%%%%%
% This section should also be commented out while your MS is undergoing
% double-blind review. The specifics should of course be adapted to
% your paper, but the paragraph below gives some hints of possible
% contributions.

\section*{Statement of Authorship}

OEC conceived the experiments, collected the data, and wrote the original draft.
GHC provided specimens and analyzed the model.
AQE oversaw data analysis and developed the code. 
All authors reviewed and edited the writing at all stages of revision.

\section*{Data and Code Availability}

On initial submission, you may use this section to provide a URL for editors and reviewers that is `private for peer review'. After acceptance, this section must be updated with correct, working DOIs for data deposits (such as on the Dryad Digital Repository, \citealt{CookEtAl2015}) and code deposits (such as in Zenodo).

\section*{Appendix A: Additional Methods and Parameters}

% In most cases, authors should typeset supplementary material in a separate,
% author-supplied PDF. For author-supplied PDFs, please consult the
% AmNat_supp_template.tex document, available from
% https://www.journals.uchicago.edu/journals/an/instruct 
%
% By contrast, the Appendix instructions below apply to cases in which
% a brief appendix is to appear in print after the main body of the article.
% That notably includes descriptions of methods, tables defining parameters,
% and other material necessary for reproducing the MS's results.
%
% Please reset counters for the appendix (thus normally figure A1, 
% figure A2, table A1, etc.).
%
% Most AmNat articles have no more than one print appendix. If your article
% has more than one, counters for each appendix should match the letter of
% that appendix. For example, tables in Appendix B should be numbered table B1, % table C2, etc. This applies to tables, equations, and figures.
%
% It's better not to use the \appendix command, because we have some
% formatting peculiarities that \appendix conflicts with.

\renewcommand{\theequation}{A\arabic{equation}}
% redefine the command that creates the equation number.
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{equation}{0}  % reset counter 
\setcounter{figure}{0}
\setcounter{table}{0}

\subsection*{Fox--dog encounters through the ages}

The quick red fox jumps over the lazy brown dog. The quick red fox has always jumped over the lazy brown dog. The quick red fox began jumping over the lazy brown dog in the 19th century and has never ceased from so jumping, as we shall see in figure~\ref{Fig:Jumps}. But there can be surprises (figure~\ref{Fig:JumpsOk}).

If the order and location of figures is not otherwise clear, feel free to include explanatory dummy text like this:

[Figure A1 goes here.]

[Figure A2 goes here.]

\subsection*{Further insights}

Tables in the appendices can appear in the appendix text (see table~\ref{Table:Rivers} for an example), unlike appendix figure legends which should be grouped at the end of the document together with the other figure legends.

\begin{table}[h]
\caption{Various rivers, cities, and animals}
\label{Table:Rivers}
\centering
\begin{tabular}{lll}\hline
River        & City        & Animal            \\ \hline
Chicago      & Chicago     & Raccoon           \\
Des Plaines  & Joliet      & Coyote            \\
Illinois     & Peoria      & Cardinal          \\
Kankakee     & Bourbonnais & White-tailed deer \\
Mississippi  & Galena      & Bald eagle        \\ \hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: See table~\ref{Table:Founders} below for further table formatting hints.}
\end{table}

Lorem ipsum dolor sit amet, as we have seen in figures~\ref{Fig:Jumps} and \ref{Fig:JumpsOk}.

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
% You can either type your references following the examples below, or
% compile your BiBTeX database and paste the contents of your .bbl file
% here. The amnatnat.bst style file should work for this---but please
% let us know if you run into any hitches with it!
%
% If you upload a .bib file with your submission, please upload the .bbl
% file as well; this will be required for typesetting.
%
% The list below includes sample journal articles, book chapters, and
% Dryad references.

\begin{thebibliography}{}

\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
\newblock Data from: Template and guidelines for using \LaTeX{} in \textit{The American Naturalist}.
\newblock American Naturalist, Dryad Digital Repository, https://dx.doi.org/10.5061/dryad.XYZAB123.

\bibitem[{Darwin(1859)Darwin}]{OriginSpec}
Darwin, C. 1859.
\newblock On the origin of species by means of natural selection, or the preservation of favoured races in the struggle for life.
\newblock J.~Murray, London.

\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
\newblock Evolution of ruminant headgear: a review.
\newblock Proceedings of the Royal Society~B 278:2857--2865.

\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
\newblock Spite and the scale of competition in \textit{Pseudomonas
  aeruginosa}.
\newblock American Naturalist 178:276--285.

\bibitem[{Fastovsky(2009)Fastovsky}]{LemKapEx07}
Fastovsky, D.~E. 2009.
\newblock Ideas in dinosaur paleontology: resonating to social and political context.
\newblock Pages 239--253 \emph{in} D. Sepkoski and M. Ruse, eds. The Paleobiological Revolution. University of Chicago Press, Chicago~IL.

\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
\newblock A strong test of the maximum entropy theory of ecology.
\newblock American Naturalist 185:E705--E80.

\section*{References Cited Only in the Online Enhancements}

% This section is for works cited in your supplemental PDF but
% not in the main text.

\bibitem[{Tytler(1759)Tytler}]{tytler-mqos}
Tytler, W. 1759.
\newblock The Inquiry, Historical and Critical, into the Evidence against Mary Queen of Scots, and an Examination of the Histories of Dr. Robertson and David Hume with respect to that Evidence.
\newblock W.~Creech, Edinburgh.

\end{thebibliography}

\newpage{}

\section*{Tables}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}

\begin{table}[h]
\caption{Founders of \textit{The~American Naturalist}}
\label{Table:Founders}
\centering
\begin{tabular}{lll}\hline
Early editor            & Years with the journal \\ \hline
Alpheus S. Packard Jr.  & 1867--1886 \\
Frederick W. Putnam     & 1867--1874 \\ 
Edward S. Morse         & 1867--1871 \\ 
Alpheus Hyatt           & 1867--1871 \\
Edward Drinker Cope$^a$ & 1878--1897 \\
J.~S. Kingsley          & 1887--1896 \\ \hline 
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Table titles should be short. Further details should go in a `notes' area after the tabular environment, like this. $^a$ Published the first description of \textit{Dimetrodon}.}
\end{table}

\newpage{}

\section*{Figure legends}

\begin{figure}[h!]
%\includegraphics{horn-of-okapi}
\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long---in most cases, they should be no more than 100 words each.}
\label{Fig:OkapiHorn}
\end{figure}

\begin{figure}[h!]
%\includegraphics{elegance}
\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance. Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
\label{Fig:AnotherFigure}
\end{figure}

% Figure legends for any print appendices you might have.

\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}[h!]
%\includegraphics{jumps20m}
\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs. \textit{B}, the quick red fox landing gracefully despite the skepticism of naysayers.}
\label{Fig:Jumps}
\end{figure}

\begin{figure}[h!]
%\includegraphics{jumps20m}
\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details, see \citet{LemKapEx07}.}
\label{Fig:JumpsOk}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%
% Videos
%%%%%%%%%%%%%%%%%%%%%
% If you have videos, journal style for them is generally similar to that for
% figures.

%%%%% Include the text below if you have videos

\renewcommand{\figurename}{Video} 
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}[h!]
%\includegraphics{VideoScreengrab.jpg}
\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
\label{VideoExample}
\end{figure}

\renewcommand{\figurename}{Figure}
\setcounter{figure}{1}

%%%%% Include the above if you have videos

\end{document}

### 2023-05-29 Manuscript/Manuscript_Am_Nat.tex
Create Manuscript_Am_Nat.tex

--- 2023-05-18 Manuscript/manuscript_draft.tex
@@
-\documentclass[12pt,a4paper]{article}
-\usepackage[a4paper, margin=1in]{geometry}
+\documentclass[11pt]{article}\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
+\usepackage{fullpage}
+\usepackage[authoryear,sectionbib,sort]{natbib}\linespread{1.7}
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{latexsym}
-\usepackage{amsfonts,amssymb,amsmath,amsthm}
-\usepackage{multirow}
-\usepackage{multicol}
-\setlength{\columnsep}{1.5cm}
-\usepackage{setspace}
-\usepackage{url}
-\usepackage{array}
-\usepackage{xfrac}
-\usepackage{mwe}
-\usepackage{tikz}
-\usepackage{lmodern}
-\usepackage{flowchart}
-\usetikzlibrary{shapes,arrows,positioning,calc,fit}
-\usepackage{graphics}
-\usepackage{graphicx}
-\usepackage{subfig}
-\usepackage{booktabs}
-\usepackage{float}
-\usepackage{color,soul}
-\usepackage{hyperref}
+\usepackage{lineno}\usepackage{titlesec}
+\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
+\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
+\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
+\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}
 \usepackage{biblatex}
@@
 \newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}  
+%% Following is used only for line numbers for pre-acceptance submissions
+\usepackage{lineno}
+%%%%%%%%%%%%%%%%%%%%%
+% Line numbering
+%%%%%%%%%%%%%%%%%%%%%
+%
+% Please use line numbering with your initial submission and
+% subsequent revisions. After acceptance, please turn line numbering
+% off by adding percent signs to the lines %\usepackage{lineno} and
+% to %\linenumbers{} and %\modulolinenumbers[3] below.
+%
+% To avoid line numbering being thrown off around math environments,
+% the math environments have to be wrapped using
+% \begin{linenomath*} and \end{linenomath*}
+%
+% (Thanks to Vlastimil Krivan for pointing this out to us!)
 
-\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms}
-\author{Ali M. Campbell\\
-	Tom E.X. Miller}
+\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms }
+
+% This version of the LaTeX template was last updated on
+% May 11, 2023.
+
+%%%%%%%%%%%%%%%%%%%%%
+% Authorship
+%%%%%%%%%%%%%%%%%%%%%
+% Please remove authorship information while your paper is under review,
+% unless you wish to waive your anonymity under double-blind review. You
+% will need to add this information back in to your final files after
+% acceptance.
+
+\author{Alexandra Campbell$^{1,\ast}$ \\ 
+Tom E.X. Miller$^{1,\dag}$}
+
+\date{}
+
 \begin{document}
+
 \maketitle
 
+\noindent{} 1. Rice University, Houston, Texas 77004;
+
+\noindent{} $\ast$ Corresponding author; e-mail: amc49@rice.edu.
+
+\noindent{} $\dag$ Corresponding author; e-mail: tom.miller@rice.edu.
+
+\textit{Manuscript elements}: Figure~1, figure~2, table~1, appendix~A (for print; including figure~A1, figure~A2, and table~A1), supplemental PDF. Figure~2 is to print in color.
+
+\bigskip
+
+\textit{Keywords}: Examples, model, template, guidelines.
+
+\bigskip
+
+\textit{Manuscript type}: Article.
+
+\bigskip
+
+\noindent{\footnotesize Prepared using the suggested \LaTeX{} template for \textit{Am.\ Nat.}}
+
+%% Comment out the following two lines after acceptance
+\linenumbers{}
+\modulolinenumbers[3]
+
+\newpage{}
 \section*{Abstract}  
-
   Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
@@
 
-\section*{Context/Introduction}
-
+\newpage{}
+\section*{Introduction}
   Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates.
-They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing\cite{Willmer1997}, pollination and dispersal for food\cite{Sakai2002,Burns2004}, resource uptake for housing\cite{Holland2010}, etc.
+They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing\cite{Willmer1997}, pollination and dispersal for food\cite{Sakai2002,Burns2004}, resource uptake for housing\cite{Holland2010}, etc.
 Mutualisms are typically considered more context dependent than other species interactions\cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
-
-Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
-The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
-Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}. 
-Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-\ali{In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host}{Is this the best way to introduce this or should I talk about quality of benefits compared to cost of rewards?
+Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
+Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.\ali{In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host}{Is this the best way to introduce this or should I talk about quality of benefits compared to cost of rewards?
 AKA some partners offer a net benefit when the benefits outweigh the cost of rewards to the partner ...
@@
     \includegraphics[width=0.58\linewidth]{Complementarity copy.png}
-	\includegraphics[width=0.40\linewidth]{Sampling_Effect copy.png}
-	\caption{asdasdf}
-	\label{fig:comp-samp}
+    \includegraphics[width=0.40\linewidth]{Sampling_Effect copy.png}\caption{asdasdf}\label{fig:comp-samp}
   \end{figure}
@@
 
-This variation in partners can lead to \ali{net benefits, no net benefits, or net costs for the focal mutualist}{Is this the best way to start this paragraph? If so should I include a figure panel?}.
-In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
-When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
-When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
-Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
-When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}.
-Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
-Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.
-
-Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
+  This variation in partners can lead to \ali{net benefits, no net benefits, or net costs for the focal mutualist}{Is this the best way to start this paragraph?
+If so should I include a figure panel?}.In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below.
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
+Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}.
+Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}.
 Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years.
-The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
-\ali{The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.}{Not quite sure what to say here.
-}
-The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
-For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.
-
-\ali{Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns}{...
+The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.\ali{The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.}{Not quite sure what to say here.
+} The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualists life\cite{Fonseca2003}.For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.\ali{Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns}{...
 Is this a better transition?
-I tried to tie it in a bit more conceptually.}, making these interactions great systems to study the effects of partner diversity.
-A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
-While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
-
-This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.
-Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates\cite{Miller2009}. 
+I tried to tie it in a bit more conceptually.}, making these interactions great systems to study the effects of partner diversity.A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
+While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates\cite{Miller2009}.
 Ant partners defend the cholla from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}.
-The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.
-While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
+The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
 
@@
   In this study we will answer three questions about the demographic effects of partner diversity across host lifetimes in multi-species mutualisms:
-
   \begin{enumerate}	
@@
     \end{enumerate}
-
-To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We used vital rate functions to evaluate the first two questions. 
+    To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.We used vital rate functions to evaluate the first two questions.
 We then used the composite integral projection model (IPM) to evaluate how vital rate differences affect intrinsic population growth rates and identify mechanisms at play.
@@
 \textit{L.
-apiculatum} are the most frequent visitors with 25\% - 60\% of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between 0\% - 20\% of cacti\cite{Donald2022} depending on the year.
-Up to 80\% of cacti remain vacant in any given year.
+apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
+opuntiae} visiting between $0\% - 20\%$ of cacti\cite{Donald2022} depending on the year.
+Up to $80\%$ of cacti remain vacant in any given year.
 
@@
 
+  \begin{linenomath*}
 		$$
@@
 		$$
+  \end{linenomath*}
 
@@
 
+	\begin{linenomath*}
 		$$
@@
 		$$
+	\end{linenomath*}
 
@@
     \hline
-growth xi intercept vacant $\Beta_01^g$ & & \\
-growth xi intercept other $\Beta_02^g$ & & \\
-growth xi intercept \textit{C. opuntiae} $\Beta_03^g$ & & \\
-growth xi intercept \textit{L. apiculatum} $\Beta_04^g$ & & \\
+    %% Growth Parameters
+    growth xi intercept vacant $\Beta_01^g$ & -5.210899 (-5.686865, -5.491787) & \\
+    growth xi intercept other $\Beta_02^g$ & -5.8288 (-5.956217, 1.766021) & \\
+    growth xi intercept \textit{C. opuntiae} $\Beta_03^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
+    growth xi intercept \textit{L. apiculatum} $\Beta_04^g$ -5.106802 (-5.4499944, 0.5453901) & \\
     growth xi size dependent vacant $\Beta_11^g$ & & \\
@@
     growth alpha size dependent $\alpha_1^g$ & & \\
+    \hline
+    %% Germination Parameters
     1-year germination intercept $\alpha^{\gamma_1}$ & & \\
     2-year germination intercept $\alpha^{\gamma_2}$ & & \\
-
+    \hline
+    %% Survival Parameters
+    survival intercept vacant $\Beta_01^s$ & & \\
+    survival intercept other $\Beta_02^s$ & & \\
+    survival intercept \textit{C.opuntiae} $\Beta_03^s$ & & \\
+    survival intercept \textit{L. apiculatum} $\Beta_04^s$ & & \\
+    survival size dependent vacant $\Beta_11^s$ & & \\
+    survival size dependent other $\Beta_12^s$ & & \\
+    survival size dependent \textit{C. opuntiae} $\Beta_13^s$ & & \\
+    survival size dependent \textit{L. apiculatum} $\Beta_14^s$ & & \\
+    \hline
+    %% Probability of Flowering Parameters
+    flowering intercept $\Beta_0^f$ & & \\
+    flowering size dependent $\Beta_1^f$ & & \\
+    \hline
+    %% Floral Viability Parameters
+    viability intercept vacant $\Beta_01^v$ & & \\
+    viability intercept other $\Beta02^v$ & & \\
+    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
+    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
   \end{tabular}
@@
 
-		\section*{Results}
-		\tom{I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???}{Let's talk about this.}
+%%%%%%%%%%%%%%%%%%%%%
+% Acknowledgments
+%%%%%%%%%%%%%%%%%%%%%
+% You may wish to remove the Acknowledgments section while your paper 
+% is under review (unless you wish to waive your anonymity under
+% double-blind review) if the Acknowledgments reveal your identity.
+% If you remove this section, you will need to add it back in to your
+% final files after acceptance.
 
-		\subsection*{Vital Rates}
-		\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{grow_contour_lines_color.png}
-			\caption{asdasdf}
-			\label{fig:grow}
-		\end{figure}
-		The results of the \tom{Bayesian}{the model is not intrinsically Bayesian} growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in \tom{Figure}{this figure needs data} \ref{fig:grow}.
-		Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
-		Plants tended by \textit{Crem.} ants had higher growth rates than tree cholla tended by any other ant or vacant. 
-		The next highest growth rate was seen in plants tended by other ants, followed by \textit{Liom.} ants, then vacant plants. 
-		The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
+\section*{Acknowledgments}
 
-		\ali{ }{I’m really unsure about what else I should say about this. }
-			\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{year_ant_timeseries.png}
-			\caption{asdasdf}
-			\label{fig:year_ant}
+OEC would like to thank Madlen Wilmes, Gyuri Barab\'{a}s, Flo D\'{e}barre, Vlastimil K\v{r}ivan, and Greg Dwyer for their comments and suggestions on this template.
+
+%%%%%%%%%%%%%%%%%%%%%
+% Statement of Authorship
+%%%%%%%%%%%%%%%%%%%%%
+% This section should also be commented out while your MS is undergoing
+% double-blind review. The specifics should of course be adapted to
+% your paper, but the paragraph below gives some hints of possible
+% contributions.
+
+\section*{Statement of Authorship}
+
+OEC conceived the experiments, collected the data, and wrote the original draft.
+GHC provided specimens and analyzed the model.
+AQE oversaw data analysis and developed the code. 
+All authors reviewed and edited the writing at all stages of revision.
+
+\section*{Data and Code Availability}
+
+On initial submission, you may use this section to provide a URL for editors and reviewers that is `private for peer review'.
+After acceptance, this section must be updated with correct, working DOIs for data deposits (such as on the Dryad Digital Repository, \citealt{CookEtAl2015}) and code deposits (such as in Zenodo).
+
+\section*{Appendix A: Additional Methods and Parameters}
+
+% In most cases, authors should typeset supplementary material in a separate,
+% author-supplied PDF. For author-supplied PDFs, please consult the
+% AmNat_supp_template.tex document, available from
+% https://www.journals.uchicago.edu/journals/an/instruct 
+%
+% By contrast, the Appendix instructions below apply to cases in which
+% a brief appendix is to appear in print after the main body of the article.
+% That notably includes descriptions of methods, tables defining parameters,
+% and other material necessary for reproducing the MS's results.
+%
+% Please reset counters for the appendix (thus normally figure A1, 
+% figure A2, table A1, etc.).
+%
+% Most AmNat articles have no more than one print appendix. If your article
+% has more than one, counters for each appendix should match the letter of
+% that appendix. For example, tables in Appendix B should be numbered table B1, % table C2, etc. This applies to tables, equations, and figures.
+%
+% It's better not to use the \appendix command, because we have some
+% formatting peculiarities that \appendix conflicts with.
+
+\renewcommand{\theequation}{A\arabic{equation}}
+% redefine the command that creates the equation number.
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{equation}{0}  % reset counter 
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
+\subsection*{Fox--dog encounters through the ages}
+
+The quick red fox jumps over the lazy brown dog.
+The quick red fox has always jumped over the lazy brown dog.
+The quick red fox began jumping over the lazy brown dog in the 19th century and has never ceased from so jumping, as we shall see in figure~\ref{Fig:Jumps}.
+But there can be surprises (figure~\ref{Fig:JumpsOk}).
+
+If the order and location of figures is not otherwise clear, feel free to include explanatory dummy text like this:
+
+[Figure A1 goes here.]
+
+[Figure A2 goes here.]
+
+\subsection*{Further insights}
+
+Tables in the appendices can appear in the appendix text (see table~\ref{Table:Rivers} for an example), unlike appendix figure legends which should be grouped at the end of the document together with the other figure legends.
+
+\begin{table}[h]
+\caption{Various rivers, cities, and animals}
+\label{Table:Rivers}
+\centering
+\begin{tabular}{lll}\hline
+River        & City        & Animal            \\ \hline
+Chicago      & Chicago     & Raccoon           \\
+Des Plaines  & Joliet      & Coyote            \\
+Illinois     & Peoria      & Cardinal          \\
+Kankakee     & Bourbonnais & White-tailed deer \\
+Mississippi  & Galena      & Bald eagle        \\ \hline
+\end{tabular}
+\bigskip{}
+\\
+{\footnotesize Note: See table~\ref{Table:Founders} below for further table formatting hints.}
+\end{table}
+
+Lorem ipsum dolor sit amet, as we have seen in figures~\ref{Fig:Jumps} and \ref{Fig:JumpsOk}.
+
+%%%%%%%%%%%%%%%%%%%%%
+% Bibliography
+%%%%%%%%%%%%%%%%%%%%%
+% You can either type your references following the examples below, or
+% compile your BiBTeX database and paste the contents of your .bbl file
+% here. The amnatnat.bst style file should work for this---but please
+% let us know if you run into any hitches with it!
+%
+% If you upload a .bib file with your submission, please upload the .bbl
+% file as well; this will be required for typesetting.
+%
+% The list below includes sample journal articles, book chapters, and
+% Dryad references.
+
+\begin{thebibliography}{}
+
+\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
+Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
+\newblock Data from: Template and guidelines for using \LaTeX{} in \textit{The American Naturalist}.
+\newblock American Naturalist, Dryad Digital Repository, https://dx.doi.org/10.5061/dryad.XYZAB123.
+
+\bibitem[{Darwin(1859)Darwin}]{OriginSpec}
+Darwin, C. 1859.
+\newblock On the origin of species by means of natural selection, or the preservation of favoured races in the struggle for life.
+\newblock J.~Murray, London.
+
+\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
+Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
+\newblock Evolution of ruminant headgear: a review.
+\newblock Proceedings of the Royal Society~B 278:2857--2865.
+
+\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
+Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
+\newblock Spite and the scale of competition in \textit{Pseudomonas
+  aeruginosa}.
+\newblock American Naturalist 178:276--285.
+
+\bibitem[{Fastovsky(2009)Fastovsky}]{LemKapEx07}
+Fastovsky, D.~E. 2009.
+\newblock Ideas in dinosaur paleontology: resonating to social and political context.
+\newblock Pages 239--253 \emph{in} D. Sepkoski and M. Ruse, eds. The Paleobiological Revolution. University of Chicago Press, Chicago~IL.
+
+\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
+Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
+\newblock A strong test of the maximum entropy theory of ecology.
+\newblock American Naturalist 185:E705--E80.
+
+\section*{References Cited Only in the Online Enhancements}
+
+% This section is for works cited in your supplemental PDF but
+% not in the main text.
+
+\bibitem[{Tytler(1759)Tytler}]{tytler-mqos}
+Tytler, W. 1759.
+\newblock The Inquiry, Historical and Critical, into the Evidence against Mary Queen of Scots, and an Examination of the Histories of Dr. Robertson and David Hume with respect to that Evidence.
+\newblock W.~Creech, Edinburgh.
+
+\end{thebibliography}
+
+\newpage{}
+
+\section*{Tables}
+\renewcommand{\thetable}{\arabic{table}}
+\setcounter{table}{0}
+
+\begin{table}[h]
+\caption{Founders of \textit{The~American Naturalist}}
+\label{Table:Founders}
+\centering
+\begin{tabular}{lll}\hline
+Early editor            & Years with the journal \\ \hline
+Alpheus S. Packard Jr.  & 1867--1886 \\
+Frederick W. Putnam     & 1867--1874 \\ 
+Edward S. Morse         & 1867--1871 \\ 
+Alpheus Hyatt           & 1867--1871 \\
+Edward Drinker Cope$^a$ & 1878--1897 \\
+J.~S. Kingsley          & 1887--1896 \\ \hline 
+\end{tabular}
+\bigskip{}
+\\
+{\footnotesize Note: Table titles should be short. Further details should go in a `notes' area after the tabular environment, like this. $^a$ Published the first description of \textit{Dimetrodon}.}
+\end{table}
+
+\newpage{}
+
+\section*{Figure legends}
+
+\begin{figure}[h!]
+%\includegraphics{horn-of-okapi}
+\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long---in most cases, they should be no more than 100 words each.}
+\label{Fig:OkapiHorn}
 \end{figure}
 
-		\tom{We broke down the impacts of ant partner on the growth rate by year and found that the effects of each ant on the growth rates of the cacti don’t always align, as shown in Figure \ref{fig:year-ant}.
-		In 2004, 2005, and 2007, the only ant which has a positive effect on the growth rate of the tree cholla. 
-		In 2006, \textit{Crem.} has the most positive effect on the growth rate of cacti, followed by Vacant, Other, and \textit{Liom.}. 
-		In 2007, 2008, and 2016, vacant had the most negative effect on the growth rates of cacti, while in 2012, vacant plants had experienced the most positive effects on their growth rates. 
-		In 2010 and 2012, the only ant which has a negative effect on the growth rate of the cacti. 
-		From 2013-2019, the effects of all ant partners are quite tightly coupled (all closely positive or closely negative). }{This is confusing and I think incorrect.}
-
-		\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{surv_panels.png}
-			\caption{asdasdf}
-			\label{fig:surv}
+\begin{figure}[h!]
+%\includegraphics{elegance}
+\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
+Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
+\label{Fig:AnotherFigure}
 \end{figure}
 
-		Our analyses of the bayesian hierarchical survival model showed that large cacti have higher survival rates than small, regardless of the ant partner. 
-		It also showed that small cacti had significantly different survival rates depending on the ant partner. 
-		Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:surv}.
-		As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last.
-		\textit{Crem.} tended plants have survival rates \tom{ranging}{These ranges are not meaningful because survival is so dependent on size.} from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.
-		\textit{Liom.} tended plants have survival rates ranging from $35.997\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
-		Other tended plants have survival rates ranging from $15.078\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
-		Vacant plants have survival rates ranging from $22.031\%$ to $99.647\%$, with the rates increasing with the size of the cacti.
+% Figure legends for any print appendices you might have.
 
-		We broke down the survival rates by year to determine the differences in ant effects across time. 
-		In 2004, 2011, 2014, and 2014, vacant tree cholla experienced more positive effects on their survival rates than any other cacti, while vacant cacti experienced the most negative effects on survival rates in 2005 and 2010.
-		In 2007 and 2013, tree cholla tended by \textit{Liom.} experienced significantly more positive effects on the survival rate than any other cacti, whereas in 2018 and 2019 \textit{Liom.} tended cacti experienced the most negative effects on the survival rates.
-		In 2004 and 2009, plants tended by ants in the category of other experienced the most negative effects on survival rates, while in 2017 these plants experienced the most positive effects on the survival rates.
-		\textit{Crem.} tended tree cholla experienced more negative effects on the survival rates than any other cacti. 
-		\textit{Crem.}, \textit{Liom.}, and Other-tended plants all experience similar patterns of positive and negative effects on survival rates through most years, with exceptions between the years of 2009-2010, 2011-2012, and 2017-2019.
+\renewcommand{\thefigure}{A\arabic{figure}}
+\setcounter{figure}{0}
 
-		\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{Viability.png}
-			\caption{asdasdf}
-			\label{fig:viab}
+\begin{figure}[h!]
+%\includegraphics{jumps20m}
+\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs.
+\textit{B}, the quick red fox landing gracefully despite the skepticism of naysayers.}
+\label{Fig:Jumps}
 \end{figure}
 
-		Tree cholla tended by \textit{Liom.} ants had significantly higher viability rates, from $79.83\%$ to $91.81\%$ with a mean of $87.047\%$, as compared to those of cacti tended by \textit{Crem.}, from $69.19\%$ to $87.60\%$ with a mean of $79,784\%$, and other, from $76.98\%$ to $86.20\%$ with a mean of $76.977\%$, as shown in Figure \ref{fig:viab}.
-		The lowest observed viability rate of tree cholla flower buds ranged from $62.83\%$ to $83.78\%$ with a mean of $74.604\%$ when there were no ant partners. 
-		Using a \tom{chi squared test}{I did not know you did this but i don't think you should} I determined that there is a $13.6\%$ chance that this difference between the mean viability rates of \textit{Liom.} tended plants and vacant plants.
+\begin{figure}[h!]
+%\includegraphics{jumps20m}
+\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details, see \citet{LemKapEx07}.}
+\label{Fig:JumpsOk}
+\end{figure}
 
-		\tom{We}{I think you need to re-think these paragraphs describing the time series, because I don't think you have given readers enough info to interpret these.} broke the effects of ant partners on the viability rates of cacti down by year and found that in some years the effects of different ant partners on the viability rates of the cacti are coupled while in others they differ significantly.
-		IN 2004 vacant cacti experienced the most positive effects on their viability rates, whereas in 2005, vacant cacti experienced the most negative effects on their viability rates compared to other cacti.
-		In 2006 and 2017, \textit{Crem.} tended cacti experienced the most positive effects on viability rates, while in 2012, 2014, 2016, and 2018, \textit{Crem.} tended cacti experienced the most negative effects on viability rates.
-		In 2005,2013, 2015, and 2019, cacti tended by ants in the other category experienced the most positive effects on the viability rates, while in 2006 they experienced the most negative effects. 
-		In 2012, \textit{Liom} tended cacti experienced the most positive effects on their viability rates, while in 2019 they experienced the most negative. 
-		From the years 2013 to 2019, the effects of all ant partners on the viability rates of cacti are tightly coupled in patterns from positive to negative.
+%%%%%%%%%%%%%%%%%%%%%
+% Videos
+%%%%%%%%%%%%%%%%%%%%%
+% If you have videos, journal style for them is generally similar to that for
+% figures.
 
-		\subsection*{Ant Transition Rates}
-			\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{Ant_Size_Multi.png}
-			\caption{asdasdf}
-			\label{fig:ant}
+%%%%% Include the text below if you have videos
+
+\renewcommand{\figurename}{Video} 
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[h!]
+%\includegraphics{VideoScreengrab.jpg}
+\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
+\label{VideoExample}
 \end{figure}
-		All very small cacti \tom{are}{your tense is inconsistent} vacant with the probability of an ant partner increasing as the cacti grow larger, as shown in Figure \ref{fig:ant}. 
-		Most large tree cholla are tended by \textit{Liom.} ants even if they had a different previous partner.
-		The largest vacant cacti have a \tom{$42.875\%$}{we should talk about this but I don't think these point values are very effectivein the text} probability of being tended by \textit{Liom.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, $28.571\%$ probability of being tended by \textit{Crem.}, and $21.429\%$ probability of being vacant in the next season.
-		Previously vacant cacti are most likely to stay vacant until the cacti reach about $10 log(m)^3$, at which point they are more likely to be tended by \textit{Liom.} ants in the next season. 
-		Large cacti tended by \textit{Liom.} ants are likely to be tended by \textit{Liom.} ants again ($90.476\%$) in the next season.
-		They have a $2.372\%$ probability of being tended by \textit{Crem.} ants in the next season, $7.143\%$ probability of being tended by other ants in the next season, and $asdfasdf$ probability of being vacant.
-		Previously \textit{Liom.} tended cacti are most likely to be vacant until they reach the size of about $7 log(m^3)$, at which point they are most likely to be tended by \textit{Liom.} in the next season.
-		Only large tree cholla previously tended by \textit{Crem.} are more likely to be tended by \textit{Crem.} again than be tended by \textit{Liom.} ants in the next year. 
-		Large tree cholla tended by \textit{Crem.} have a $47.059\%$ probability of being tended by \textit{Crem.} in the next season, $33.333\%$ probability of being tended by \textit{Liom.} in the next season, $30\%$ probability of being tended by other ants in the next season, and $33.333\%$ probability of being vacant.
-		Previously \textit{Crem.} tended plants are most likely to become vacant in the next season until they reach the size of about $15 log(m^3)$, after which they are more likely to remain tended by \textit{Crem.} ants.
-		Cacti previously tended by other ants have a $24.138\%$ probability of being tended by \textit{Crem.} in the next season, $69.231\%$ probability of being tended by \textit{Liom.} in the next season, $7.692\%$ probability of being tended by other ants in the next season, and $23.076\%$ probability of being vacant.
-		Medium cacti follow the same patterns as large cacti in partner transitions.
 
-		\tom{}{What about the other vital rates?}
+\renewcommand{\figurename}{Figure}
+\setcounter{figure}{1}
 
-		\subsection*{Demographic Modeling}
-		\begin{figure}[h]
-			\includegraphics[width=0.58\linewidth]{lambda_det_full.png}
-			\caption{asdasdf}
-			\label{fig:lambda_det}
-		\end{figure}
-		\begin{figure}[h]
-		\includegraphics[width=0.58\linewidth]{lambda_st_full3.png}
-		\caption{asdasdf}
-		\label{fig:lambda_stoch}
-	\end{figure}
-		\tom{We considered both a deterministic Integral Projection Model and a stochastic one to contrast the differences.}{Just emphasizing again that the methods does not convey this.} 
-		In the deterministic model, we found that \tom{simulations}{you did not provide reproducible methods for these simulations} with all ant partners (and vacancy) present resulted in the highest mean population growth rate while populations with no ant partners had the lowest mean population growth rate, as shown in Figure \ref{fig:lambda_det}.
-		The estimated mean population fitness of tree cholla when all ants were present was higher than the mean population fitness of any other scenario.
-		Similarly, only a few other scenarios had means within the interquartile range of this high partner diversity scenario. 
-		\tom{Honestly have no idea how much detail I should go into in this?
-}{It's less a matter of detail (though more would be good) and more about having a point.
-Your text seems to describe the results in a quasi-random, surface-level way.
-Bring more intention to your decisions about what to write here.
-What does a reader need to know to understand the story you are trying to tell?}
-
-		\section*{Discussion}
-		\tom{\subsection*{Vital Rates}}{I don't think these are the right subsection headings because they are simply redescribing results.}
-		\textbf{Ant partners have contrasting effects on vital rate processes of tree cholla.}
-		Regression analyses showed that the different ant partners had significantly different impacts on the various vital rate processes of the tree cholla cacti. 
-		Specifically, \textit{Crem.} tended cacti have advantages, at small sizes, over cacti tended by any other ant or vacant. 
-		\textit{Crem.} tended tree cholla had the highest survival rates (Figure \ref{fig:surv})  at small sizes and the highest growth rates (Figure \ref{fig:grow}), two of the most important vital rates for small cacti which are not yet reproducing.
-		On the other hand, reproducing cacti which have \textit{Liom.} partners  have advantages with the highest viability rates of flower buds (Figure \ref{fig:viab}). 
-		Together this indicates that the best partner may change as the cacti grow and begin to reproduce, when the most vulnerable part of the plant is the flower with the seeds. 
-		This reflects the changes in the resource use of tree cholla as they begin to use their resources for reproduction rather than growth. 
-		The fact that different ant partners have significantly different effects on the various vital rates of tree cholla indicates that none of them are the “perfect” partner, and that the “best” partner may in fact change over the lifespan of the cacti.
-		As the tree cholla grew, the best partner changed from \textit{Crem.} ants, partners known for ***, to \textit{Liom.} ants, partners best known for defensive benefits to the cacti, particularly against the seed predators which most impact viability.
-		The difference in ant partners made a significant difference in the observed vital rates of the cacti, indicating that \tom{considering the interaction between the tree cholla and any individual partner would fail to capture the extent of the benefits}{great point and well made -- I am going to stop commenting on the rest of the discussion because I think it needs to be re-worked.
-We can discuss more.} to the cacti.
-		Like many other systems, pairwise perspectives do not fully encompass the complex impacts that multispecies mutualisms have on the focal mutualists vital rates.
-
-		\subsection*{Ant Transition Rates}
-		\textbf{Small cacti remain vacant, while large cacti are most likely to be tended by \textit{Liom.} ants.}
-		Small cacti are unlikely to be tended because most do not produce EFN, of if they do it is a very small amount. 
-		Medium-sized to large cacti are much more likely to be tended than vacant as they begin producing EFN, with many factors that determine the most likely partner. 
-		Some of the ant partners appear to have high turnover rates, meaning they are unlikely to tend the same plant multiple years in a row. 
-		Ants in the other category have high turnover rates, shown by the fact that medium and large cacti tended by other are unlikely to be tended by other ants again in the next season. 
-		The reason for these turnovers could be due to the inability to defend their territories, an ability to find and colonize new resources, or an ability to colonize resources that have been left unclaimed.
-		This discovery-dominancy tradeoff is a well-studied hypothesis in ant literature \cite{lach2010} that could explain the remaining presence of ants in the other category despite high turnover rates.
-
-		On the other hand, \textit{Crem.} ants appear to have lower turnover rates, since large cacti have up to $47.059\%$ probability of being tended by \textit{Crem.} in the next year.
-		In addition to turnover rates, there are also colonization rates, the probability of a species taking over a cactus that was previously tended by different ants. 
-		\textit{Liom.} are the only ant partners we observed with high colonization rates. 
-		Most cacti tended by non-\textit{Liom.} ants have a high probability of being taken over by \textit{Liom} ants in the next season, as shown in Figure \ref{fig:ant}.
-		This pattern could be due to the well-known high levels of aggression displayed by \textit{Liom.} ants in this system, however there are other possible explanations, such as nectar composition.
-		The exception to this rule are plants tended by \textit{Crem.} ants which are more likely to remain tended by \textit{Crem.} than be taken over by \textit{Liom.} in the next season. 
-		The trend that most large ants are likely to be tended by \textit{Liom.} ants reflects the findings that large cacti benefit most from \textit{Liom.} as ant partners. 
-		As explained in the introduction, there are many factors that determine the colonization rates of different cacti by their ant partners, including EFN quantity and quality, ability to seek out cacti, aggression, and more.
-		These patterns could be explained by the \textit{Liom.} ants ability to dominate at a resource site and therefore takeover from different ants which originally found the site.
-
-		An alternative, or parallel, explanation for these ant transitions could be the changes in EFN composition across ontogeny of the cacti.
-		As cacti grow the chemistry of EFN produced changes ***, changing from more *** composition to *** composition. 
-		Different ants have preferences for different nectar compositions\cite{Lach2010}, specifically, ***.
-		This could provide a potential route for the cacti to select their own ideal partners
-		This indicates a potential future avenue of research into the correlations between ant partners and the chemistry of the EFN produced by the tree cholla.
-
-		\subsection*{Demographic Modeling}
+%%%%% Include the above if you have videos
 
 \end{document}
-
-\typeout{\bibliography{References.bib}}

### 2023-05-29 skewnormal_test.R
Delete skewnormal_test.R

@@ -1,11 +1,0 @@
-library(sn)
-library(rstan)
-library(bayesplot)
-
-y<-rsn(n=10000,xi=0,omega=1,alpha=-5)
-mean(y);hist(y)
-
-stan_data <- list(y=y,N=length(y))
-fit_skew <- stan(file="Data Analysis/STAN Models/skewnormal_test.stan", 
-                      data = stan_data, warmup = 150, iter = 1000, chains = 3, cores = 3, thin = 2)
-mcmc_trace(fit_skew)

### 2023-05-30 Manuscript/Manuscript_Am_Nat.tex
Update Manuscript_Am_Nat.tex

@@
 \section*{Abstract}
+
   Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
@@
 \section*{Introduction}
+
   Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions\cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing\cite{Willmer1997}, pollination and dispersal for food\cite{Sakai2002,Burns2004}, resource uptake for housing\cite{Holland2010}, etc.
-Mutualisms are typically considered more context dependent than other species interactions\cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
-Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.
-Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.\ali{In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host}{Is this the best way to introduce this or should I talk about quality of benefits compared to cost of rewards?
-AKA some partners offer a net benefit when the benefits outweigh the cost of rewards to the partner ...
-etc.} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
-In cases where the partner offers benefits, \ali{the function, or type of benefits, offered by the partner can vary}{Is this enough detail or do I need to go into examples?} \cite{Stanton2003}.
+They are among the most widespread species interactions \cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
+Mutualisms are typically considered more context dependent than other species interactions \cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
 
-  \begin{figure}[h]
-    \includegraphics[width=0.58\linewidth]{Complementarity copy.png}
-    \includegraphics[width=0.40\linewidth]{Sampling_Effect copy.png}\caption{asdasdf}\label{fig:comp-samp}
-  \end{figure}
-  \begin{figure}[h]
-    \includegraphics[width=0.60\linewidth]{Portfolio_Effect copy.png}
-    \caption{asdasdf}
-    \label{fig:portfolio}
-  \end{figure}
+  Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions \cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
+The differences in mutualistic outcomes are due to differences in partner quality \cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered \cite{Kiers2003,Afkhami2014}, and direct interactions between partners \cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
+For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms \cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
+Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms \cite{Bascompte2019}. 
+Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
+In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
+In cases where the partner offers benefits, the function, or type of benefits, offered by the partner can vary \cite{Stanton2003}.
 
-  This variation in partners can lead to \ali{net benefits, no net benefits, or net costs for the focal mutualist}{Is this the best way to start this paragraph?
-If so should I include a figure panel?}.In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below.
-When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}.
-Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}.
-Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}.
-Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years.
-The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.\ali{The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.}{Not quite sure what to say here.
-} The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualists life\cite{Fonseca2003}.For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.\ali{Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns}{...
-Is this a better transition?
-I tried to tie it in a bit more conceptually.}, making these interactions great systems to study the effects of partner diversity.A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}.
-While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates\cite{Miller2009}.
-Ant partners defend the cholla from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}.
-The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
+  This variation in partners can lead to net benefits, no net benefits, or net costs for the focal mutualist.
+In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
+When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit) \cite{Frederickson2013}.
+When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, sampling effect can explain the positive effects of partner diversity \cite{Batstone2018}.
+Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist \cite{Stachowicz2005,Bronstein2006,Stanton2003}.
+When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity \cite{Batstone2018}. 
+Partners can also have asynchronous responses to the environment, either spatially \cite{Ollerton2006} or temporally \cite{Alarcon2008}.Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, through the portfolio effect \cite{Batstone2018,Lazaro2022}.
 
-  \begin{figure}[h]	
-    \includegraphics[width=.8\linewidth]{herb_ant_only_flow.png}	
-    \caption{asdasdf}	
-    \label{fig:herb}
-  \end{figure}
+  Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
+Turnover can happen at different timescales \cite{Oliveira1999,Horvitz1986} from minutes to years. 
+The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
+The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.
+The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualists life \cite{Fonseca2003}.
+For example, susceptibility to enemies can change across life stages \cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.
+Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns, making these interactions great systems to study the effects of partner diversity.
+A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores \cite{Bronstein1998, Bronstein2006}.
+While these interactions have been well studied \cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds \cite{Stanton2013} or temporal fluctuations in partner interactions \cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
+
+  This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates \cite{Miller2009}.
+Ant partners defend the cholla from these insect herbivores, reducing the negative effects of herbivory \cite{Miller2007}. 
+The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.
+While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
 
@@
     \end{enumerate}
-    To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.We used vital rate functions to evaluate the first two questions.
+    To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
+We used vital rate functions to evaluate the first two questions. 
 We then used the composite integral projection model (IPM) to evaluate how vital rate differences affect intrinsic population growth rates and identify mechanisms at play.
@@
   \subsection*{Study System}
+
   This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
 This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
-Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
+Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \cite{Benson1982}. 
 These arborescent cholla produce cylindrical segments with large spines. 
@@
 Tree cholla generally reach at least 9 years of age before beginning to reproduce \cite{Ohm2014}.
-Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures\cite{Ness2006,Oliveira1999}.
+Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \cite{Ness2006,Oliveira1999}.
 
   This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, a \tom{\textit{Phenogaster spp.}}{Incorrect spelling.}, and \textit{Camponotus spp.}.
+At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
 \textit{L.
 apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $0\% - 20\%$ of cacti\cite{Donald2022} depending on the year.
+opuntiae} visiting between $0\% - 20\%$ of cacti \cite{Donald2022} depending on the year.
 Up to $80\%$ of cacti remain vacant in any given year.
 
-  These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}.
+  These ants rarely co-occur on a plant, probably due to interspecific competition \cite{Miller2007}.
 Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
 At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset\cite{Ohm2014}. 
-Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended\cite{Miller2014}. 
+They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \cite{Ohm2014}. 
+Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \cite{Miller2014}. 
 In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-  There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments\cite{Mann1969}. 
+  There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \cite{Mann1969}. 
 An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
@@
 A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cholla and depress population growth\cite{Miller2009}.
-There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators\cite{Miller2007}.
+These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \cite{Miller2009}.
+There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \cite{Miller2007}.
 
 	\subsection*{Data}
+
   The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
@@
 	\subsection*{Statistical Modeling -- Construction}
+
   With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
@@
   \subsection*{Integral Projection Model Construction}
+
   Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
@@
   \subsection*{Deterministic IPM Analysis}
+
   Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
@@
   \subsection*{Stochastic IPM Analysis}
+
   The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
@@
 Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
-
-% I am creating a table here to include all parameter estimations and descriptions
-  \begin{table}[]
-  \begin{tabular}{l|l|l}
-    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
-    \hline
-    %% Growth Parameters
-    growth xi intercept vacant $\Beta_01^g$ & -5.210899 (-5.686865, -5.491787) & \\
-    growth xi intercept other $\Beta_02^g$ & -5.8288 (-5.956217, 1.766021) & \\
-    growth xi intercept \textit{C. opuntiae} $\Beta_03^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
-    growth xi intercept \textit{L. apiculatum} $\Beta_04^g$ -5.106802 (-5.4499944, 0.5453901) & \\
-    growth xi size dependent vacant $\Beta_11^g$ & & \\
-    growth xi size dependent other $\Beta_12^g$ & & \\
-    growth xi size dependent \textit{C. opuntiae} $\Beta_13^g$ & & \\
-    growth xi size dependent \textit{L. apiculatum} $\Beta_14^g$ & & \\
-    growth omega intercept $\omega_0^g$ & & \\
-    growth omega size dependent $\omega_1^g$ & & \\
-    growth alpha intercept $\alpha_0^g$ & & \\
-    growth alpha size dependent $\alpha_1^g$ & & \\
-    \hline
-    %% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
-    \hline
-    %% Survival Parameters
-    survival intercept vacant $\Beta_01^s$ & & \\
-    survival intercept other $\Beta_02^s$ & & \\
-    survival intercept \textit{C.opuntiae} $\Beta_03^s$ & & \\
-    survival intercept \textit{L. apiculatum} $\Beta_04^s$ & & \\
-    survival size dependent vacant $\Beta_11^s$ & & \\
-    survival size dependent other $\Beta_12^s$ & & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_13^s$ & & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_14^s$ & & \\
-    \hline
-    %% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & & \\
-    flowering size dependent $\Beta_1^f$ & & \\
-    \hline
-    %% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & & \\
-    viability intercept other $\Beta02^v$ & & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
-  \end{tabular}
-  \end{table}
 
@@
 
-OEC would like to thank Madlen Wilmes, Gyuri Barab\'{a}s, Flo D\'{e}barre, Vlastimil K\v{r}ivan, and Greg Dwyer for their comments and suggestions on this template.
-
 %%%%%%%%%%%%%%%%%%%%%
@@
 
-OEC conceived the experiments, collected the data, and wrote the original draft.
-GHC provided specimens and analyzed the model.
-AQE oversaw data analysis and developed the code. 
-All authors reviewed and edited the writing at all stages of revision.
-
 \section*{Data and Code Availability}
-
-On initial submission, you may use this section to provide a URL for editors and reviewers that is `private for peer review'.
-After acceptance, this section must be updated with correct, working DOIs for data deposits (such as on the Dryad Digital Repository, \citealt{CookEtAl2015}) and code deposits (such as in Zenodo).
 
@@
 
-\renewcommand{\theequation}{A\arabic{equation}}
-% redefine the command that creates the equation number.
-\renewcommand{\thetable}{A\arabic{table}}
-\setcounter{equation}{0}  % reset counter 
-\setcounter{figure}{0}
-\setcounter{table}{0}
-
-\subsection*{Fox--dog encounters through the ages}
-
-The quick red fox jumps over the lazy brown dog.
-The quick red fox has always jumped over the lazy brown dog.
-The quick red fox began jumping over the lazy brown dog in the 19th century and has never ceased from so jumping, as we shall see in figure~\ref{Fig:Jumps}.
-But there can be surprises (figure~\ref{Fig:JumpsOk}).
-
-If the order and location of figures is not otherwise clear, feel free to include explanatory dummy text like this:
-
-[Figure A1 goes here.]
-
-[Figure A2 goes here.]
-
-\subsection*{Further insights}
-
-Tables in the appendices can appear in the appendix text (see table~\ref{Table:Rivers} for an example), unlike appendix figure legends which should be grouped at the end of the document together with the other figure legends.
-
-\begin{table}[h]
-\caption{Various rivers, cities, and animals}
-\label{Table:Rivers}
-\centering
-\begin{tabular}{lll}\hline
-River        & City        & Animal            \\ \hline
-Chicago      & Chicago     & Raccoon           \\
-Des Plaines  & Joliet      & Coyote            \\
-Illinois     & Peoria      & Cardinal          \\
-Kankakee     & Bourbonnais & White-tailed deer \\
-Mississippi  & Galena      & Bald eagle        \\ \hline
-\end{tabular}
-\bigskip{}
-\\
-{\footnotesize Note: See table~\ref{Table:Founders} below for further table formatting hints.}
-\end{table}
-
-Lorem ipsum dolor sit amet, as we have seen in figures~\ref{Fig:Jumps} and \ref{Fig:JumpsOk}.
-
 %%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{thebibliography}{}
-
-\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
-Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
-\newblock Data from: Template and guidelines for using \LaTeX{} in \textit{The American Naturalist}.
-\newblock American Naturalist, Dryad Digital Repository, https://dx.doi.org/10.5061/dryad.XYZAB123.
-
-\bibitem[{Darwin(1859)Darwin}]{OriginSpec}
-Darwin, C. 1859.
-\newblock On the origin of species by means of natural selection, or the preservation of favoured races in the struggle for life.
-\newblock J.~Murray, London.
-
-\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
-Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
-\newblock Evolution of ruminant headgear: a review.
-\newblock Proceedings of the Royal Society~B 278:2857--2865.
-
-\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
-Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
-\newblock Spite and the scale of competition in \textit{Pseudomonas
-  aeruginosa}.
-\newblock American Naturalist 178:276--285.
-
-\bibitem[{Fastovsky(2009)Fastovsky}]{LemKapEx07}
-Fastovsky, D.~E. 2009.
-\newblock Ideas in dinosaur paleontology: resonating to social and political context.
-\newblock Pages 239--253 \emph{in} D. Sepkoski and M. Ruse, eds. The Paleobiological Revolution. University of Chicago Press, Chicago~IL.
-
-\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
-Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
-\newblock A strong test of the maximum entropy theory of ecology.
-\newblock American Naturalist 185:E705--E80.
-
-\section*{References Cited Only in the Online Enhancements}
-
-% This section is for works cited in your supplemental PDF but
-% not in the main text.
-
-\bibitem[{Tytler(1759)Tytler}]{tytler-mqos}
-Tytler, W. 1759.
-\newblock The Inquiry, Historical and Critical, into the Evidence against Mary Queen of Scots, and an Examination of the Histories of Dr. Robertson and David Hume with respect to that Evidence.
-\newblock W.~Creech, Edinburgh.
-
-\end{thebibliography}
+\typeout{\bibliography{References.bib}}
 
@@
 
-\begin{table}[h]
-\caption{Founders of \textit{The~American Naturalist}}
-\label{Table:Founders}
-\centering
-\begin{tabular}{lll}\hline
-Early editor            & Years with the journal \\ \hline
-Alpheus S. Packard Jr.  & 1867--1886 \\
-Frederick W. Putnam     & 1867--1874 \\ 
-Edward S. Morse         & 1867--1871 \\ 
-Alpheus Hyatt           & 1867--1871 \\
-Edward Drinker Cope$^a$ & 1878--1897 \\
-J.~S. Kingsley          & 1887--1896 \\ \hline 
+% I am creating a table here to include all parameter estimations and descriptions
+  \begin{table}[]
+  \begin{tabular}{l|l|l}
+    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
+    \hline
+    %% Growth Parameters
+    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
+    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
+    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
+    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
+    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
+    growth xi size dependent other $\Beta_{12}^g$ & & \\
+    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
+    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+    growth omega intercept $\omega_0^g$ & & \\
+    growth omega size dependent $\omega_1^g$ & & \\
+    growth alpha intercept $\alpha_0^g$ & & \\
+    growth alpha size dependent $\alpha_1^g$ & & \\
+    \hline
+    %% Germination Parameters
+    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
+    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
+    \hline
+    %% Survival Parameters
+    survival intercept vacant $\Beta_{01}^s$ & & \\
+    survival intercept other $\Beta_{02}^s$ & & \\
+    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & & \\
+    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & & \\
+    survival size dependent vacant $\Beta_{11}^s$ & & \\
+    survival size dependent other $\Beta_{12}^s$ & & \\
+    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & & \\
+    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & & \\
+    \hline
+    %% Probability of Flowering Parameters
+    flowering intercept $\Beta_0^f$ & & \\
+    flowering size dependent $\Beta_1^f$ & & \\
+    \hline
+    %% Floral Viability Parameters
+    viability intercept vacant $\Beta_01^v$ & & \\
+    viability intercept other $\Beta02^v$ & & \\
+    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
+    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
   \end{tabular}
+  \end{table}
 \bigskip{}

### 2023-06-08 Manuscript/Manuscript_Am_Nat.tex
Update Manuscript_Am_Nat.tex
I started commenting, then just started a new draft

@@
   Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
-Mutualisms are typically considered more context dependent than other species interactions \cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
+They are among the most widespread species interactions \cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into commensalism or parasitism,  \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+%One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
+Mutualisms are considered more context dependent than other species interactions \cite{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by the surrounding environment and species' identities.
 
-  Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions \cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
+  Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of \tom{multi-species mutualistic interactions}{It is very important to give this an operational definition early in the paper.} \cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
 The differences in mutualistic outcomes are due to differences in partner quality \cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered \cite{Kiers2003,Afkhami2014}, and direct interactions between partners \cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms \cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
+\tom{For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms \cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.}{This re-states something you said two sentences ago.}
 Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms \cite{Bascompte2019}. 
 Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
-In cases where the partner offers benefits, the function, or type of benefits, offered by the partner can vary \cite{Stanton2003}.
+In some cases, the quality of the benefits offered varies leading to \tom{some partners which benefit the individual}{what is implicit here is that your definition of multi-species mutualism is where a focal species has multiple partners -- that is not at all clear.}, some which have neither a positive or negative impact on the individual, and some which negatively impact the host \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
+In cases where the partner offers benefits, the function, or type of benefits, offered by the partner can vary \cite{Stanton2003}.\tom{}{The structure of this paragraph is disorienting because you have a list of mechanisms in sentence two, then you introduce new ideas, then you return to these same mechanisms here.}
 
-  This variation in partners can lead to net benefits, no net benefits, or net costs for the focal mutualist.
-In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
+  \tom{This variation in partners can lead to net benefits, no net benefits, or net costs for the focal mutualist.}{Are there examples of all of these?}
+\tom{In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below.}{``I'' is not the right pronoun for a multi-autho paper, and the tone of this sentence makes your focus on positive effects seem arbitrary and not well supported.}
 When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit) \cite{Frederickson2013}.
@@
   Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Turnover can happen at different timescales \cite{Oliveira1999,Horvitz1986} from minutes to years. 
+Turnover can happen at different timescales \cite{Oliveira1999,Horvitz1986}, from minutes to years. 
 The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
 The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.
-The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualists life \cite{Fonseca2003}.
+The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist \cite{Fonseca2003}.
 For example, susceptibility to enemies can change across life stages \cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.
@@
   Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix \as shown in figure ***. 
+Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***. 
 The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.

### 2023-06-08 Manuscript/ant_cactus_AmNat.tex
created a new ms draft

--- 2023-05-30 Manuscript/Manuscript_Am_Nat.tex
@@
-\documentclass[11pt]{article}\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
+\documentclass[11pt]{article}
+\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
 \usepackage{fullpage}
-\usepackage[authoryear,sectionbib,sort]{natbib}\linespread{1.7}
+\usepackage[authoryear,sectionbib,sort]{natbib}
+\linespread{1.7}
 \usepackage[utf8]{inputenc}
-\usepackage{lineno}\usepackage{titlesec}
+\usepackage{lineno}
+\usepackage{titlesec}
 \titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
@@
 \titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}
-\usepackage{biblatex}
-\addbibresource{References.bib}
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
 \newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
-%% Following is used only for line numbers for pre-acceptance submissions
-\usepackage{lineno}
+
 %%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms }
+\title{Thank you, next: partner turnover elevates benefits of mutualism for an ant-tended plant}
 
@@
 \author{Alexandra Campbell$^{1,\ast}$ \\ 
-Tom E.X. Miller$^{1,\dag}$}
+	Tom E.X. Miller$^{1}$}
 
@@
 
-\noindent{} 1. Rice University, Houston, Texas 77004;
+	\noindent{} 1. Program in Ecology and Evolutionary Biology, Department of BioSciences, Rice University, Houston, Texas 77005;
 
@@
 
-\noindent{} $\dag$ Corresponding author; e-mail: tom.miller@rice.edu.
-
-\textit{Manuscript elements}: Figure~1, figure~2, table~1, appendix~A (for print; including figure~A1, figure~A2, and table~A1), supplemental PDF. Figure~2 is to print in color.
+	\textit{Manuscript elements}:
 
@@
 
-\textit{Keywords}: Examples, model, template, guidelines.
+	\textit{Keywords}:
 
@@
 
-%% Comment out the following two lines after acceptance
 \linenumbers{}
@@
 \newpage{}
+
 \section*{Abstract}
 
-  Mutualisms are among the most widespread species interactions with diverse and dynamic consequences.
-They are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity.
-Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies.
-It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions.
-The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms, portfolio effect, complementarity, and sampling effect.
-In this paper, I will focus on defensive ant-plant mutualisms.
-These involve plants which provide extra-floral nectar and/or “housing” to ants which in turn defend them from herbivores.
-While these interactions have been well studied in the literature, few have considered how diversity within partner guilds affect the overall benefits of mutualism for the plant partner.
-  I use the plant, Cylindropuntia imbracata (tree cholla), and ant, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), and more, multispecies mutualism in which the cacti provide extrafloral nectar in exchange for defense from various herbivores and seed predators.
-I used 18 years of data collected from demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, ant partner, and herbivory for 8 30x30 m plots, in New Mexico.
-With this data I parameterize a series of Bayesian generalized linear vital rate models to determine the impacts of different partners on the focal mutualists.
-I found that different ant partners did have different impacts on the vital rates of the tree cholla.
-Specifically, Crem.
-tended plants had advantages in both growth and survival when small, and Liom.
-tended plants had advantages in colonizing partners as well as floral viability.
-With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest fitness, and which of the above mechanisms may explain the effects of partner diversity in this system.
-I found that the real-life scenario (all possible ants are present) lead to the highest fitness for the tree cholla, indicating that partner diversity is beneficial in this system.
-It also shows that complementarity is at play in this system, meaning different partners offer different benefits leading to synergistic benefits for the tree cholla.
+\newpage{}
 
-\newpage{}
 \section*{Introduction}
 
+% The journal does not have numbered sections in the main portion of
+% articles. Please refrain from using section references (à la
+% section~\ref{section:CountingOwlEggs}), and refer to sections by name
+% (e.g. section ``Counting Owl Eggs'').
+
 Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \cite{Chamberlain2014, BoucherDouglasH.1985} with diverse and dynamic consequences \cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
-Mutualisms are typically considered more context dependent than other species interactions \cite{Chamberlain2014,Frederickson2013}, meaning strength of the interaction is often determined by the surrounding environment and species.
+They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013}, but can deteriorate into commensalism or parasitism  \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+%One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
+Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities.
 
-  Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions \cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
-The differences in mutualistic outcomes are due to differences in partner quality \cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered \cite{Kiers2003,Afkhami2014}, and direct interactions between partners \cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
-For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms \cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
-Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms \cite{Bascompte2019}. 
-Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
-In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}.
-In cases where the partner offers benefits, the function, or type of benefits, offered by the partner can vary \cite{Stanton2003}.
+Mutualism is defined at the level of a species pair (+/+), but these interactions are embedded within multi-species communities. 
+Growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \cite{Afkhami2014,Palmer2010}. 
+A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
+Even within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \cite{Stanton2003}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings such as agriculture \citep{rogers2014bee}, restoration (cite), and pest management (cite).
 
-  This variation in partners can lead to net benefits, no net benefits, or net costs for the focal mutualist.
-In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
-When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit) \cite{Frederickson2013}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (BEF chapter).
+When there is a consistent hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
 When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, sampling effect can explain the positive effects of partner diversity \cite{Batstone2018}.
-Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist \cite{Stachowicz2005,Bronstein2006,Stanton2003}.
-When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity \cite{Batstone2018}. 
-Partners can also have asynchronous responses to the environment, either spatially \cite{Ollerton2006} or temporally \cite{Alarcon2008}.Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, through the portfolio effect \cite{Batstone2018,Lazaro2022}.
-
-  Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Turnover can happen at different timescales \cite{Oliveira1999,Horvitz1986} from minutes to years. 
-The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
-The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.
-The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualists life \cite{Fonseca2003}.
-For example, susceptibility to enemies can change across life stages \cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.
-Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns, making these interactions great systems to study the effects of partner diversity.
-A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores \cite{Bronstein1998, Bronstein2006}.
-While these interactions have been well studied \cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds \cite{Stanton2013} or temporal fluctuations in partner interactions \cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.
-
-  This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.Insect herbivory negatively affects the growth of cholla, in turn reducing population growth rates \cite{Miller2009}.
-Ant partners defend the cholla from these insect herbivores, reducing the negative effects of herbivory \cite{Miller2007}. 
-The multiple ant partners do not co-occur on individual plants, but a single cholla may interact with multiple types of partners over its lifetime.
-While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these tree cholla.
-
-  In this study we will answer three questions about the demographic effects of partner diversity across host lifetimes in multi-species mutualisms:
-  \begin{enumerate}	
-    \item{What are the contrasting demographic effects of multiple partners?}	
-    \item{How do host size and partner identity impact the directionality of turnover between multiple           partners?}	
-    \item{What mechanisms explain the effects of partner diversity in a multispecies mutualism?}
-    \end{enumerate}
-    To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
-We used vital rate functions to evaluate the first two questions. 
-We then used the composite integral projection model (IPM) to evaluate how vital rate differences affect intrinsic population growth rates and identify mechanisms at play.
 
 \section*{Methods}
-  \subsection*{Study System}
 
-  This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
-This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
-Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \cite{Benson1982}. 
-These arborescent cholla produce cylindrical segments with large spines. 
-In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
-While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
-Tree cholla generally reach at least 9 years of age before beginning to reproduce \cite{Ohm2014}.
-Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \cite{Ness2006,Oliveira1999}.
+\section*{Results}
 
-  This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
-\textit{L.
-apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $0\% - 20\%$ of cacti \cite{Donald2022} depending on the year.
-Up to $80\%$ of cacti remain vacant in any given year.
+\section*{Discussion}
 
-  These ants rarely co-occur on a plant, probably due to interspecific competition \cite{Miller2007}.
-Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
-At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \cite{Ohm2014}. 
-Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \cite{Miller2014}. 
-In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
-
-  There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \cite{Mann1969}. 
-An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
-A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \cite{Miller2006}.
-A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \cite{Miller2009}.
-There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \cite{Miller2007}.
-
-	\subsection*{Data}
-
-  The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
-The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
-Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
-The final two plots were added to this census from 2011 onwards. 
-Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
-For each cholla, we recorded plant survival from the last survey to the current survey. 
-We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
-We also recorded the species of herbivore and the number present on the plant.
-
-	\subsection*{Statistical Modeling -- Construction}
-
-  With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
-Many of the vital rates are estimated as a function of plant size, ant partner, or both.
-Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
-The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
-Unless otherwise mentioned, all models use vague priors. 
-The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
-Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
-The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
-Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these. 
-The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
-The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
-The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
-Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion. 
-The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
-Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
-The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
-With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
-With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
-With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
-Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds.
-
-  To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
-For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
-The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
-We did not thin the chains, thus all samples were retained. 
-To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
-To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
-Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-All estimated parameters are described in table 1.
-Data and code for all vital rate models is included in the supplemental information.
-
-  \subsection*{Integral Projection Model Construction}
-
-  Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
-We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
-Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
-There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
-This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
-
-  Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
-The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
-
-  \begin{linenomath*}
-		$$
-		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
-		$$
-		$$
-		B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
-		$$
-  \end{linenomath*}
-
-  The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
-The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
-The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
-Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
-Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
-Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
-
-  The size dynamics of the cholla are given by:
-
-	\begin{linenomath*}
-		$$
-		n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
-		$$
-		$$
-		\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
-		$$
-	\end{linenomath*}
-
-  The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
-The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
-$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
-Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
-
-  \subsection*{Deterministic IPM Analysis}
-
-  Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix \as shown in figure ***. 
-The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
-We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
-We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
-apiculatum} and vacancy; \textit{C.
-opuntiae} and vacancy; other and vacancy; \textit{L.
-apiculatum}, \textit{C.
-opuntiae}, and vacancy, \textit{L.
-apiculatum}, other, and vacancy; \textit{C.
-opuntiae}, other, and vacancy; and all ant partners and vacancy.
-The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
-Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
-Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner.
-
-  \subsection*{Stochastic IPM Analysis}
-
-  The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
-The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
-In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
-For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
-With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
-This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
-We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
-These two relative differences allow us to determine if portfolio effect is at play.
-Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
-
-%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgments
-%%%%%%%%%%%%%%%%%%%%%
-% You may wish to remove the Acknowledgments section while your paper 
-% is under review (unless you wish to waive your anonymity under
-% double-blind review) if the Acknowledgments reveal your identity.
-% If you remove this section, you will need to add it back in to your
-% final files after acceptance.
+\section*{Conclusion}
 
@@
 % contributions.
-
-\section*{Statement of Authorship}
 
@@
 
+\renewcommand{\theequation}{A\arabic{equation}}
+% redefine the command that creates the equation number.
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{equation}{0}  % reset counter 
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
 %%%%%%%%%%%%%%%%%%%%%
@@
 % Dryad references.
-
-\typeout{\bibliography{References.bib}}
+\bibliographystyle{apalike}
+\bibliography{References.bib}
 
@@
 
-% I am creating a table here to include all parameter estimations and descriptions
-  \begin{table}[]
-  \begin{tabular}{l|l|l}
-    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
-    \hline
-    %% Growth Parameters
-    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
-    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
-    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
-    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
-    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
-    growth xi size dependent other $\Beta_{12}^g$ & & \\
-    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
-    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
-    growth omega intercept $\omega_0^g$ & & \\
-    growth omega size dependent $\omega_1^g$ & & \\
-    growth alpha intercept $\alpha_0^g$ & & \\
-    growth alpha size dependent $\alpha_1^g$ & & \\
-    \hline
-    %% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
-    \hline
-    %% Survival Parameters
-    survival intercept vacant $\Beta_{01}^s$ & & \\
-    survival intercept other $\Beta_{02}^s$ & & \\
-    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & & \\
-    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & & \\
-    survival size dependent vacant $\Beta_{11}^s$ & & \\
-    survival size dependent other $\Beta_{12}^s$ & & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & & \\
-    \hline
-    %% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & & \\
-    flowering size dependent $\Beta_1^f$ & & \\
-    \hline
-    %% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & & \\
-    viability intercept other $\Beta02^v$ & & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
-  \end{tabular}
-  \end{table}
-\bigskip{}
-\\
-{\footnotesize Note: Table titles should be short. Further details should go in a `notes' area after the tabular environment, like this. $^a$ Published the first description of \textit{Dimetrodon}.}
-\end{table}
-
 \newpage{}
@@
 \section*{Figure legends}
-
-\begin{figure}[h!]
-%\includegraphics{horn-of-okapi}
-\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long---in most cases, they should be no more than 100 words each.}
-\label{Fig:OkapiHorn}
-\end{figure}
-
-\begin{figure}[h!]
-%\includegraphics{elegance}
-\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
-Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
-\label{Fig:AnotherFigure}
-\end{figure}
-
-% Figure legends for any print appendices you might have.
-
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-
-\begin{figure}[h!]
-%\includegraphics{jumps20m}
-\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs.
-\textit{B}, the quick red fox landing gracefully despite the skepticism of naysayers.}
-\label{Fig:Jumps}
-\end{figure}
-
-\begin{figure}[h!]
-%\includegraphics{jumps20m}
-\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details, see \citet{LemKapEx07}.}
-\label{Fig:JumpsOk}
-\end{figure}
 
@@
 
-\renewcommand{\figurename}{Video} 
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-
-\begin{figure}[h!]
-%\includegraphics{VideoScreengrab.jpg}
-\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
-\label{VideoExample}
-\end{figure}
-
-\renewcommand{\figurename}{Figure}
-\setcounter{figure}{1}
-
 %%%%% Include the above if you have videos

### 2023-06-09 Manuscript/ant_cactus_AmNat.tex
done with edits to Intro

@@
 \usepackage{titlesec}
+\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
+\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+
 \titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
@@
 \titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
-\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}
-\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
-\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[.]\renewcommand{\refname}{Literature Cited}
 
@@
 % (e.g. section ``Counting Owl Eggs'').
-
 Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
@@
 
-Mutualism is defined at the level of a species pair (+/+), but these interactions are embedded within multi-species communities. 
-Growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \cite{Afkhami2014,Palmer2010}. 
+Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010}.
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Even within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \cite{Stanton2003}.
+Even within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
 Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings such as agriculture \citep{rogers2014bee}, restoration (cite), and pest management (cite).
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (BEF chapter).
-When there is a consistent hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
-When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, sampling effect can explain the positive effects of partner diversity \cite{Batstone2018}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (BEF chapter).
+First, when there is a hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
+This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
+However, if partner associations are mutually exclusive then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the best partner \citep{Miller2007}.
+Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
+For example, [give example].
+This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
+Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions (cite). 
+Third, partner species can have species-specific responses to the environment, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across environmental stochasticity or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
+
+Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014}. 
+Sequential associations are likely when alternative partners engage in interference comptetition for access to a shared mutualist (cite examples, including non-ant-plant examples). 
+Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
+The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004}.
+Directionality of turnover can also influence diversity effects, particularly if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003}.
+For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005}.
+
+Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
+Ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist (examples), including dynamic turnover patterns between partners that may differ in quantity and quality of defensive services (cite).
+While these interactions have been well studied \citep{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds \citep{Stanton2013} or temporal fluctuations in partner interactions \citep{Tgaard2015} affect the overall benefits of mutualism for the plant partner.
+\footnote{It is noticeable to anyone who knows this literature that you do not cite Palmer et al.
+2010 in this paragraph.
+I think this paragraph gives a nice intro to ant-plant studies but I think you need a more thorough summary of what is known about effects of partner diversity in these systems (including Palmer), and what outstanding gaps in knowledge this study will fill.}
+
+This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindriopuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
+Previous studies have shown that herbivory by specialized insect herbivores negatively affects plant fitness \cite{Miller2009}, and ant defense reduces herbivore damage \cite{Miller2007}. 
+Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
+These ant species locally co-occur at the scale of meters, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022does}.
+Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} actually having net negative fitness effects though deterrence of pollinators \citep{Miller2007,Ohm2014}.
+However, those studies did not consider variation in ant defense across the plant life cycle, nor did they account for inter-annual fluctuations, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
+
+In this study we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species and how the demographic effects of alternative partner species varied across the range of plant size structure and nearly 20 years of inter-annual fluctuations.
+We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of mutualism for the focal plant partner.
+Specifically, we asked:
+\begin{enumerate}	
+	\item{What are the demographic effects of association with alternative partners and how do these effects fluctuate across years?}
+	\item{What are the frequency and direction of partner turnover across the plant life cycle?}	
+	\item{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
+\end{enumerate}
+We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy, and to quantify state-dependent partner turnover. 
+We then used a multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
 

### 2023-07-06 Data Analysis/Setup_Script.R
Update Setup_Script.R

@@ -29,6 +29,7 @@ library(moments)
 library("posterior")
 library(gridExtra)
 library(grid)
+library(MASS)
 # data management
 library(tidyverse)
 library(dplyr)
@@ -39,23 +40,7 @@ library(corrplot)
 library(ggpubr)
 
 
-#library(shinystan)
-#library(devtools)
-#library(tidyr)
-#library(lattice)
-#library(fixest)
-#library(mlogit)
-#library(dfidx)
-#library(effects)
-#library(bbmle)
-#library(Gmisc)
-#library(ggeffects)
-#library(heplots)
-#library(Rlab)
-#library(gbm)
-#library(dismo)
-#library(popbio)
-#library(sjPlot)
+
 knitr::opts_chunk$set(echo = TRUE)
 options(mc.cores = parallel::detectCores())
 
@@ -63,7 +48,8 @@ options(mc.cores = parallel::detectCores())
 ################        import the data -- Cacti (main)        #######################################################
 #######################################################################################################
 #cactus_uncleaned <- read.csv("cholla_demography_20042021.csv", header = TRUE,stringsAsFactors=T)
-cactus <- read.csv("Data Analysis/cholla_demography_20042021.csv", header = TRUE,stringsAsFactors=T)
+cactus <- read.csv("Data Analysis/cholla_demography_20042023.csv", header = TRUE,stringsAsFactors=T)
+cactus_old <- read.csv("Data Analysis/cholla_demography_20042021.csv", header = TRUE,stringsAsFactors=T)
 str(cactus) ##<- problem: antcount is a factor
 #levels(cactus$Antcount_t);levels(cactus$Antcount_t1)
 ## function for the volume of a cone
@@ -76,84 +62,10 @@ cactus$logsize_t <- log(volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t))
 cactus$logsize_t1 <- log(volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1))
 
 
-
-# ####################################################################################################
-# ########            Cactus 2019 data                    ############################################
-# ####################################################################################################
-# ## Change ant counts
[…]

### 2023-07-06 Manuscript/ant_cactus_AmNat.tex
Manuscripts Updates

@@
 They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013}, but can deteriorate into commensalism or parasitism  \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-%One reason they are so ubiquitous is they take many forms in nature, including defense for food or housing \cite{Willmer1997}, pollination and dispersal for food \cite{Sakai2002,Burns2004}, resource uptake for housing \cite{Holland2010}, etc.
 Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities.
@@
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
-For example, [give example].
 This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
@@
 Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
-Ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist (examples), including dynamic turnover patterns between partners that may differ in quantity and quality of defensive services (cite).
-While these interactions have been well studied \citep{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds \citep{Stanton2013} or temporal fluctuations in partner interactions \citep{Tgaard2015} affect the overall benefits of mutualism for the plant partner.
-\footnote{It is noticeable to anyone who knows this literature that you do not cite Palmer et al.
-2010 in this paragraph.
-I think this paragraph gives a nice intro to ant-plant studies but I think you need a more thorough summary of what is known about effects of partner diversity in these systems (including Palmer), and what outstanding gaps in knowledge this study will fill.}
+Extrafloral nectar (EFN) producing plants can serve as dietary resources for for ant species which can promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
+Presence of defensive ant partners is linked to herbivory reduction \citep{Trager2010, Rudgers2010} and increased plant growth and survival \citep{Baez2016}.
+These interactions are almost entirely studied from the perspective of plant fitness \citep{Bronstein1994,Bronstein1998}, with little recorded about the impacts on ant fitness \citep{Lanan2013}.
+Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
+Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
+Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
+Recently many studies have investigated how partner diversity in these guilds has been shown to lead to either increased plant fitness \citep{Palmer2010, Afkhami2014} or decreased plant fitness (often in more highly specialized interactions) \citep{Barrett2015, Ushio2020}, stressing the importance of considering all ant partners as unique within these guilds.
 
@@
 \section*{Methods}
+\subsection*{Study System}
+
+This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
+This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
+Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \citep{Benson1982}. 
+These arborescent cholla produce cylindrical segments with large spines. 
+In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
+Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
+Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \citep{Ness2006,Oliveira1999}.
+
+This EFN is harvested by various ant species in return for defense. 
+At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
+\textit{L.
+apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
+opuntiae} visiting between $0\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
+Up to $80\%$ of cacti remain vacant in any given year.
+
+These ants rarely co-occur on a plant, probably due to interspecific competition \citep{Miller2007}.
+Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
+At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
+They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \citep{Ohm2014}. 
+Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \citep{Miller2014}. 
+In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
+
+There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \citep{Mann1969}. 
+An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
+A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
+A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
+These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \citep{Miller2009}.
+There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \citep{Miller2007}.
+
+\subsection*{Data}
+
+The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
+Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
+The final two plots were added to this census from 2011 onwards. 
+Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
+For each cholla, we recorded plant survival from the last survey to the current survey. 
+We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
+We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
+We also recorded the species of herbivore and the number present on the plant.
+
+\subsection*{Statistical Modeling -- Construction}
+
+With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
+Many of the vital rates are estimated as a function of plant size, ant partner, or both.
+Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
+The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
+Unless otherwise mentioned, all models use vague priors. 
+The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
+The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
+Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these. 
+The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
+The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
+The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
+Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion. 
+The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
+Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
+The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
+With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
+With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
+With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds.
+
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
+For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
+The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
+We did not thin the chains, thus all samples were retained. 
+To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
+To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
+Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
+All estimated parameters are described in table 1.
+Data and code for all vital rate models is included in the supplemental information.
+
+\subsection*{Integral Projection Model Construction}
+
+Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
+We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
+Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
+There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
+This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
+
+Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
+The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
+
+  \begin{linenomath*}
+		$$
+		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
+		$$
+		$$
+		B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
+		$$
+  \end{linenomath*}
+
+The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
+The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
+The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
+Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
+Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
+
+The size dynamics of the cholla are given by:
+
+	\begin{linenomath*}
+		$$
+		n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
+		$$
+		$$
+		\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
+		$$
+	\end{linenomath*}
+
+The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
+The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
+$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
+Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
+
+\subsection*{Deterministic IPM Analysis}
+
+Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
+Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***. 
+The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
+We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
+We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
+apiculatum} and vacancy; \textit{C.
+opuntiae} and vacancy; other and vacancy; \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and vacancy, \textit{L.
+apiculatum}, other, and vacancy; \textit{C.
+opuntiae}, other, and vacancy; and all ant partners and vacancy.
+The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
+Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
+Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner.
+
+\subsection*{Stochastic IPM Analysis}
+
+The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
+The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
+In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
+For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
+With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
+This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
+We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
+These two relative differences allow us to determine if portfolio effect is at play.
+Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
 
@@
 
+\renewcommand{\thetable}{\arabic{table}}
+\setcounter{table}{0}
+
+% I am creating a table here to include all parameter estimations and descriptions
+  \begin{table}[]
+  \begin{tabular}{l|l|l}
+    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
+    \hline
+    %% Growth Parameters
+    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
+    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
+    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
+    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
+    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
+    growth xi size dependent other $\Beta_{12}^g$ & & \\
+    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
+    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+    growth omega intercept $\omega_0^g$ & & \\
+    growth omega size dependent $\omega_1^g$ & & \\
+    growth alpha intercept $\alpha_0^g$ & & \\
+    growth alpha size dependent $\alpha_1^g$ & & \\
+    \hline
+    %% Germination Parameters
+    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
+    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
+    \hline
+    %% Survival Parameters
+    survival intercept vacant $\Beta_{01}^s$ & & \\
+    survival intercept other $\Beta_{02}^s$ & & \\
+    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & & \\
+    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & & \\
+    survival size dependent vacant $\Beta_{11}^s$ & & \\
+    survival size dependent other $\Beta_{12}^s$ & & \\
+    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & & \\
+    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & & \\
+    \hline
+    %% Probability of Flowering Parameters
+    flowering intercept $\Beta_0^f$ & & \\
+    flowering size dependent $\Beta_1^f$ & & \\
+    \hline
+    %% Floral Viability Parameters
+    viability intercept vacant $\Beta_01^v$ & & \\
+    viability intercept other $\Beta02^v$ & & \\
+    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
+    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
+  \end{tabular}
+  \end{table}
+\bigskip{}
+\\
+\end{table}
+
 \newpage{}

### 2023-07-06 Figures/Cholla_Analysis_Visualization_Vital_Rates.R
Figures Updates

@@ -308,21 +308,21 @@ grow_out <- read.csv("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutua
 ## Formulas
 ## Other
 
-y_other_mean_grow <- quantile(grow_out$beta0.2,0.5) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.5)
-y_other_low_grow <- quantile(grow_out$beta0.2,0.05) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.05)
-y_other_high_grow <- quantile(grow_out$beta0.2,0.95) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.2,0.95)
+y_other_mean_grow <- quantile(grow_out$beta0.3,0.5) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.5)
+y_other_low_grow <- quantile(grow_out$beta0.3,0.05) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.05)
+y_other_high_grow <- quantile(grow_out$beta0.3,0.95) + (size_other+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.95)
 ## Crem
-y_crem_mean_grow <- quantile(grow_out$beta0.3,0.5) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.5)
-y_crem_low_grow <- quantile(grow_out$beta0.3,0.05) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.05)
-y_crem_high_grow <- quantile(grow_out$beta0.3,0.95) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.3,0.95)
+y_crem_mean_grow <- quantile(grow_out$beta0.1,0.5) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.1,0.5)
+y_crem_low_grow <- quantile(grow_out$beta0.1,0.05) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.1,0.05)
+y_crem_high_grow <- quantile(grow_out$beta0.1,0.95) + (size_crem+mean(growth_data$logsize_t)) * quantile(grow_out$beta1.1,0.95)
 ## Liom
[…]

### 2023-08-02 IPM Code/Params.R
Update Params.R

@@ -3,7 +3,7 @@ setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
 
 
 ## load in the cleaned data product created by Create_Clean_Data_Script.R
-cactus<-read.csv("cholla_demography_20042023_cleaned.csv")
+cactus<-read.csv("cholla_demography_20042021_cleaned.csv")
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/IPM Code")
 
 cholla_min<- min((cactus$logsize_t), na.rm = TRUE)  ## minsize 
@@ -24,18 +24,30 @@ upper<- cholla_max + ceiling.extend
 mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 #Tom
 #mcmc_dir <- "C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
+# grow.params <- read.csv(paste0(mcmc_dir,"grow.params.csv"), header = TRUE,stringsAsFactors=T)    
+# surv.params <- read.csv("surv.params.csv", header = TRUE,stringsAsFactors=T)    
+# flow.params <- read.csv("flow.params.csv", header = TRUE,stringsAsFactors=T)    
+# flow.phi <- read.csv("flow.phi.csv", header = TRUE,stringsAsFactors=T) 
+# viab.params <- read.csv("viab.params.csv", header = TRUE,stringsAsFactors=T)    
+# repro.params <- read.csv("repro.params.csv", header = TRUE,stringsAsFactors=T)    
+# seed.params <- read.csv("seed.params.csv", header = TRUE,stringsAsFactors=T)    
+# pre.seed.params <- read.csv(paste0(mcmc_dir,"seed.surv.params.csv"), header = TRUE,stringsAsFactors=T)    
+# germ1.params <- read.csv(paste0(mcmc_dir,"germ1.params.csv"), header = TRUE,stringsAsFactors=T)    
+# germ2.params <- read.csv(paste0(mcmc_dir,"germ2.params.csv"), header = TRUE,stringsAsFactors=T)    
+# rec.params <- read.csv(paste0(mcmc_dir,"rec.params.csv"), header = TRUE,stringsAsFactors=T)    
 
 ##These files contain all draws from the posterior distributions of all parameters
[…]

### 2023-08-02 Manuscript/ant_cactus_AmNat_Ali.tex
manuscript

--- 2023-07-06 Manuscript/ant_cactus_AmNat.tex
@@
 \section*{Results}
+\subsection*{Statistical Modeling}
+
+%% Growth Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{grow_contour_lines_color.png}
+%	\caption{asdasdf}
+%	\label{fig:growth}
+%\end{figure}
+Plants tended by other ants experiened positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}a.
+They experienced significantly higher growth than vacant plants ($p = 2.2 \times 10^{-16}$), and similar mean growth to \textit{L.
+apiculatum} tended plants ($p = 0.9135$) and \textit{C.
+opuntiae} tended plants ($p = 0.4864$).
+Plants tended by \textit{C. opuntiae} ants experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}b.
+They experienced significantly higher mean growth than vacant plants ($p = 4.88 \times 10^{-14}$).
+Plants tended by \textit{L. apiculatum} experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}c.
+Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d.
+
+%% Survival Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{surv_panels.png}
+%	\caption{asdasdf}
+%	\label{fig:surv}
+%\end{figure}
+Plants tended by \texit{C.
+opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
+apiculatum} tended plants ($p = 2.908 \times 10^{-14}$), plants tended by other ants ($p = 4.488 \times 10^{-14}$), and vacant plants ($p = 2.2 \times 10^{-16}$) \cite{fig:surv}.
+These \textit{C. opuntiae} plants have the highest mean survival at small sizes before reaching a mean survival rate of about 100\% when medium to large.
+Plants tended by \textit{L. apiculatum} experienced similar mean survival rates to plants tended by other ants and vacant plants.
+\textit{L. apiculatum} tended plants and other tended plants reached a mean survival rate of nearly 100\% by medium sizes, with other tended plants reaching this rate slower. 
+Vacant plants never reach mean survival rates of 100\%.
+
+%% Flowers Model
+There is a clear size effect on the number of flowers produced. 
+The mean number of flowers produced by a plant remains at 0 until the plant reachesmedium sizes after which the mean number of flower produced increases eponentially to about 40 at large sizes.
+
+%% Viability Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{Viability.png}
+%	\caption{asdasdf}
+%	\label{fig:comp-samp}
+%\end{figure}
+Plants tended by \textit{L.
+apiculatum} ants experienced the highest viability rates, significantly higher than \textit{C.
+opuntiae} tended plants ($p = 2.2 \times 10^{-16}$), other tended plants ($p = 2.2 \times 10^{-16}$), and vacant plants ($p = 2.2 \times 10^{-16}$).
+The mean viability rate of \textit{L. apiculatum} tended plants was 84.888\%.
+Plants tended by \textit{C. opuntiae} ants experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$), but not than other tended plants ($p = 0.1168$).
+The mean viability rate of \textit{C. opuntiae} plants was 75.934\%.
+Plants tended by other ants also experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$).
+The mean viability rate of other tended plants was 76.367\%.
+The mean viability rate of vacant plants was 70.442\%.
+
+%% Reproduction Model
+The probability of a plant reproducing in a given year is highly size dependent. 
+The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+
+%% Seeds Produced
+The mean number of seeds produced by \textit{C.
+opuntiae} tended plants is significantly lower than the mean number of seeds produced by \textit{L.
+apiculatum} plants ($p = 2.2 \times 10^{-16}$) and vacant plants ($p = 2.2 \times 10^{-16}$).
+\textit{C. opuntiae} tended plants produce a mean of 2.561 seeds per flower. 
+The mean number of seeds produced by \textit{L. apiculatum} tended plants is slightly higher than the mean number produced by vacant plants ($p = 0.02521$). 
+\textit{L. apiculatum} tended plants produce a mean of 3.563 seeds per flower. 
+Vacant plants produce a mean of 3.435 seeds per flower.
+
+%% Precensus Survival
+The mean pre-census seed survival rate is 16.99\% with 90\% of model iterations landing between 1.13\% and 40.72\%.
+
+%% Germination
+Seeds have a significantly higher probability of germinating in year one than in year two ($p = 2.2 \times 10^{-16}$).
+Seeds in year one have a mean germination rate of 0.578\% and seeds in year two have a mean germination rate of 0.221\%.
+
+%% Recruit size distribution
+ The mean size of new recruits is -1.88 log($cm^3$) with 90\% of new recruits falling between the sizes of -2.12 log($cm^3$) and -1.64 log($cm^3$).
+
+%% Ant Transition Rates
+% previously vacant
+Plants that were previously vacant will most likely remain vacant until large. 
+These plants have significantly higher probabilities of being vacant rather than becoming tended  by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+The probability of previously vacant plants becoming tended by \texit{L.
+apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
+opuntiae} or other ants (respecitvely, $p = 3.83 \times 10^{-14}$, $p = 2.2 \times 10^{-16}$).
+The probability of previously vacant plants becoming tended by other ants is significantly lower than the probability of becoming tended by \textit{C. opuntiae} ($p = 0.00227$).
+
+% previously crem
+Plants that were previously tended by \textit{C. opuntiae} will most likely remain vacant until large. 
+These plants have significantly higher probabilities of becoming vacant rather than remaining tended by \textit{C.
+opuntiae} or becoming tended by \textit{L.
+apiculatum} or other ants (repsectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$ ).
+The probability of these plants remaining tended by \textit{C.
+opuntiae} ants is significantly higher than the probability of becoming tended by \textit{L.
+apiculatum} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 3.797 \times 10^{-9}$).
+The probability of these plants becoming tended by \textit{L. apiculatum} is significantly higher than the probability of becoming tended by other ants ($p = 3.797 \times 10^{-9}$).
+
+% previously liom
+Plants that were previously tended by \textit{L. apiculatum} ants will most likely remain vacant until large. 
+These plants have significantly higher probabilities of becoming vacant than remaining tended by \textit{L.
+apiculatum} or becoming tended by \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+These plants have significantly higher probabilities of remaining tended by \textit{L.
+apiculatum} ants than becoming tended by  \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+The probability of these plants becoming tended by \textit{C. opuntiae} ants is significantly higher than the probability of them becoming tended by other ants ($p = 0.007925$).
+
+% previously other
+Plants that were previously tended by other ants will most likely remain vacant until they are large. 
+These plants have significantly higher probabilities of becoming vacant rather than remaining tended by other ants or becoming tended by \textit{C.
+opuntiae} or \textit{L.
+apiculatum} ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+The probability of these plants becoming tended by \textit{L.
+apiculatum} ants is significantly higher than the probability of remaining tended by other ants or becomnig tended by \textit{C.
+opuntiae} (respectively, $p = 6.309 \times 10^{-12}$, $p = 6.256 \times 10^{-13}$).
+The probability of these plants remaining tended by other ants is not higher than the probability of becoming tended by \textit{C. opuntiae} ants ($p = 0.954$).
+
+%% All cases/trends
+In all plants ,the most likely scenario is vacancy.
+In most cases, the second most likely next ant partner is \textit{L.
+apiculatum} with the exception of plants which were previously tended by \textit{C.
+opuntiae} who are second most likely to retain \textit{C.
+opuntiae} as their partners in successive years.
+The only case where two next partners do not have statistically significantly different probabilities is when previously other tended plants either remain tended by other ants or become tended by \textit{C.
+opuntiae}.
+
+\subsection*{Deterministic IPM}
+
+\subsection*{Stochastic IPM}
 
@@
     %% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
+    1-year germination intercept $\alpha^{\gamma_1}$ & -5.144627 (-5.370185, -4.966587) & \\
+    2-year germination intercept $\alpha^{\gamma_2}$ & -6.155088 (-6.417991, -5.786996) & \\
     \hline
     %% Survival Parameters
-    survival intercept vacant $\Beta_{01}^s$ & & \\
-    survival intercept other $\Beta_{02}^s$ & & \\
-    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & & \\
-    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & & \\
-    survival size dependent vacant $\Beta_{11}^s$ & & \\
-    survival size dependent other $\Beta_{12}^s$ & & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & & \\
+    survival intercept vacant $\Beta_{01}^s$ & 0.46375959 (0.07778875, 0.86267803) & \\
+    survival intercept other $\Beta_{02}^s$ & 0.2677732 (-1.7939375, 2.5857164) & \\
+    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & 2.2690061 (0.7347587, 3.9967755) & \\
+    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & 0.0538502 (-1.0054015, 1.3153753) & \\
+    survival size dependent vacant $\Beta_{11}^s$ & 0.86267803 (0.2696006, 0.3217293) & \\
+    survival size dependent other $\Beta_{12}^s$ & 0.4387909 (0.1953732, 0.7325813) & \\
+    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & 0.28392994 (0.08742771, 0.47423484) & \\
+    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & 0.5150193 (0.5150193, 0.6492092) & \\
     \hline
     %% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & & \\
-    flowering size dependent $\Beta_1^f$ & & \\
+    flowering intercept $\Beta_0^f$ & -4.611414 (-5.119275, -4.082511) & \\
+    flowering size dependent $\Beta_1^f$ & 0.5693555 (0.5239602, 0.6104472) & \\
     \hline
     %% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & & \\
-    viability intercept other $\Beta02^v$ & & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
+    viability intercept vacant $\Beta_01^v$ & 1.0930266 (0.5002784, 1.6629479) & \\
+    viability intercept other $\Beta02^v$ & 1.1102242 (0.5073163, 1.7414692) & \\
+    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & 1.3605548 (0.7372043, 1.9359609) & \\
+    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & 2.023599 (1.461603, 2.534776) & 
   \end{tabular}

### 2023-08-14 Manuscript/ant_cactus_AmNat.tex
tom's edits on intro and methods

@@
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Even within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings such as agriculture \citep{rogers2014bee}, restoration (cite), and pest management (cite).
+Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014bee}.
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (BEF chapter).
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (cite BEF chapter).
 First, when there is a hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
 This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
-However, if partner associations are mutually exclusive then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the best partner \citep{Miller2007}.
+However, if partner associations are mutually exclusive then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
@@
 Third, partner species can have species-specific responses to the environment, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across environmental stochasticity or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
 
 Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Sequential associations are likely when alternative partners engage in interference comptetition for access to a shared mutualist (cite examples, including non-ant-plant examples). 
+Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist (cite examples, including non-ant-plant examples). 
 Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
@@
 Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
-Extrafloral nectar (EFN) producing plants can serve as dietary resources for for ant species which can promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
-Presence of defensive ant partners is linked to herbivory reduction \citep{Trager2010, Rudgers2010} and increased plant growth and survival \citep{Baez2016}.
-These interactions are almost entirely studied from the perspective of plant fitness \citep{Bronstein1994,Bronstein1998}, with little recorded about the impacts on ant fitness \citep{Lanan2013}.
+Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
+Presence of defensive ant partners is often linked to herbivory reduction \citep{Trager2010, Rudgers2010} and demographic advantages for the plant partner \citep{Baez2016}.
+\tom{These interactions are almost entirely studied from the perspective of plant fitness \citep{Bronstein1994,Bronstein1998}, with little recorded about the impacts on ant fitness \citep{Lanan2013}.}{Here is a good example of something to avoid.
+You are highlighting a gap that your paper actually does not fill.
+Never do this in an Intro, sometimes it can be useful in a discussion as a future direction.}
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
 Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
-Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
-Recently many studies have investigated how partner diversity in these guilds has been shown to lead to either increased plant fitness \citep{Palmer2010, Afkhami2014} or decreased plant fitness (often in more highly specialized interactions) \citep{Barrett2015, Ushio2020}, stressing the importance of considering all ant partners as unique within these guilds.
+\tom{Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
+Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.}{I think you are mixing two ideas here: ontogenetic variation in ant-plant associations (eg directional turnover) and temporal stochasticity that is independent of plant ontogeny.
+Your study considers both and I think this needs to be set up more clearly.}
+\tom{Recently many studies have investigated how partner diversity in these guilds has been shown to lead to either increased plant fitness \citep{Palmer2010, Afkhami2014} or decreased plant fitness (often in more highly specialized interactions) \citep{Barrett2015, Ushio2020}, stressing the importance of considering all ant partners as unique within these guilds.
+}{I think you need a more effective argument for why your study is necessary given that there are previous studies that have yielded results in both directions.}
 
@@
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} actually having net negative fitness effects though deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not consider variation in ant defense across the plant life cycle, nor did they account for inter-annual fluctuations, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the benefits provided by alternative partners, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
 
-In this study we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species and how the demographic effects of alternative partner species varied across the range of plant size structure and nearly 20 years of inter-annual fluctuations.
-We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of mutualism for the focal plant partner.
+We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
+We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
 Specifically, we asked:
@@
 We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy, and to quantify state-dependent partner turnover. 
-We then used a multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
+We then used a stochastic, multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
 
@@
 
-This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
-This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
+This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research site (SEV-LTER) in central New Mexico, USA.
+This is an area characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B. gracilis}), yuccas, cacti, and junipers. 
 Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \citep{Benson1982}. 
-These arborescent cholla produce cylindrical segments with large spines. 
-In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+These arborescent plants produce cylindrical segments with large spines. 
+In the growing season (May to August in New Mexico), the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
 While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
-Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
-Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \citep{Ness2006,Oliveira1999}.
+%Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
+Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative segments and flower buds \citep{Ness2006,Oliveira1999}. 
+Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners.
 
-This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
+Tree cholla EFN is harvested by various ant species. 
+At SEV-LTER, cholla are visited primarily by two species of formicoid, ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster}, and a \textit{Camponotus}.
 \textit{L.
 apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $0\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
-Up to $80\%$ of cacti remain vacant in any given year.
-
-These ants rarely co-occur on a plant, probably due to interspecific competition \citep{Miller2007}.
+opuntiae} visiting between \tom{$0\% - 20\%$}{Are there years with zero CREM?} of cacti \citep{Donald2022} depending on the year.
+\tom{Up to $80\%$ of cacti remain vacant in any given year.
+}{This is probably the extreme high end of vacancy.
+I would give the ranges as you did the others.}
+These ants rarely co-occur on a plant, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
+opuntiae} to \textit{L.
+apiculatum}-tended plants, and vice versa, provoke aggressive responses by resident ants (A.
+Cambpell, \textit{personal observation}).
 Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
@@
 They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \citep{Ohm2014}. 
-Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \citep{Miller2014}. 
+Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla, especially flowering individuals, are generally more highly tended \citep{Miller2014}. 
 In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \citep{Mann1969}. 
+There are a variety of insect herbivores and seed predators which attack the cholla \citep{Mann1969}. 
 An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
@@
 These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \citep{Miller2009}.
-There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \citep{Miller2007}.
+There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory than plants from which ants were excluded \citep{Miller2007}.
 
 \subsection*{Data}
-
-The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+\tom{}{Give this section a more informative title.}	
+The \tom{data collected are from a long-term dataset}{This grammar does not really make sense.} spanning 2004 to 2023 \tom{taken from 30 $\times$ 30 meter plots}{These plots only started in 2009.} at SEV-LTER.
 The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
@@
 The final two plots were added to this census from 2011 onwards. 
-Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
+Annually, in May we surveyed all individuals in these plots, taking demographic and partner data. 
 For each cholla, we recorded plant survival from the last survey to the current survey. 
-We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
-We also recorded the species of herbivore and the number present on the plant.
+For surviving plants, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm), which are used to calculate plant volume ($cm^3$) based on the \tom{volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter}{Check your volume function but I think it is actually the volume of an elliptical cone.}.
+We recorded the total number of flower buds, including how many were aborted and how many were not. 
+\tom{We recorded all ant species present and the number of ants we could count in 30 seconds.}{Need to describe how plants were assigned to an ant state in the case of multiple ant species (rare, I know.)}
+We also recorded the numbers and species of herbivores on each plant.
+\tom{}{Also need to describe recruit data, seed banks, seed counts per fruit.}
 
-\subsection*{Statistical Modeling -- Construction}
+\subsection*{Statistical Modeling}
 
-With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
+With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework to serve as vital rate sub-models of the IPM.
 Many of the vital rates are estimated as a function of plant size, ant partner, or both.
-Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
-The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
+\tom{Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. }{This needs to be explained.}
+The biological sources of variance (including \tom{individual}{But you do not have an individual random effect.}, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models.
 Unless otherwise mentioned, all models use vague priors.
-The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+
+\tom{The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+}{I think you need to explain this more thoroughly, and also explain why you used a Skewed Normal.
+Omega and alpha need to be defined and explained, and this is best done showing the full notation of the model.}
 Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
@@
 Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds. 
+\tom{}{General comment here is that each of these models needs to be described in greater detail and shown in reproducible notation.
+You also need to describe the idea of transitions in size x,y and ant state i,j since this sets the stage for the structure of the IPM.}
 
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \tom{}{You need to cite R and RStan}
 For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
@@
 Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-All estimated parameters are described in table 1.
+\tom{All estimated parameters are described in table 1.}{I think I suggested this years ago but there are so many parameters in your model that I don't think it is particularly useful to report all of these in a table.
+Including publicly available data and code is good enough.}
 Data and code for all vital rate models is included in the supplemental information.
@@
 
-Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
+Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables. 
+While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
 We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
-Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
-There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
-This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
+%Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
+%There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
+This novel structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
 
-Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
-The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
+Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where $x$ is the size of the cholla and $i$ is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
+The dynamics of these 1 and 2-year old seedbanks are given by the following equations: \tom{}{Rather than ``hard-code'' 4 in the summation, I would create a parameter for number of possible ant states, maybe $n_{A}$.}
 
@@
 
-The size dynamics of the cholla are given by:
+The \tom{size dynamics}{It is not just size dynamics, it is size and ant dynamics. That needs to be communicated more clearly.} of the cholla are given by:
 
@@
 
-The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
+\tom{The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
 The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
 $G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
-Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
+Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.}{This is a good start but I think you need a more thorough description, since this model will be unfamiliar to many readers.}
 
 \subsection*{Deterministic IPM Analysis}
-
+  \tom{}{The model you show above is the deterministic model.
+The stochastic model is a different model with additional parameters that should be represented symbolically.
+So my point is that I would no introduce the model first and then describe deterministic and stochastic analysis.
+Each of those thngs corresponds to a different model.}
 Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***. 
+\tom{Each component of the IPM (each set of ant partner specific vital rates) }{I don't think readers will understand what you are talking about here} can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***.
 The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
@@
 opuntiae}, other, and vacancy; and all ant partners and vacancy.
-The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
-Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
+The \tom{relative distributions}{Unclear what this means} allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
+\tom{Sampling Effect requires}{I don't think this is the right grammar (``requires'')} that the $r$ when all possible partners are present is \tom{equal to}{What if it is less than?} the $r$ of the cholla population when only the best partner is present.
 Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner. 
+\tom{}{These last two sentences are very important -- laying out predictions under alternative hypotheses -- so it is work making them laser sharp. }
 
@@
 The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
-In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
+In order to include stochasticity in the analysis,\tom{ we randomly sampled from the year-random effect intercepts 1,000 times}{I don't think you have communicated, in either words or notation, that the random intercepts are unique to ant state.}.
 For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
@@
 These two relative differences allow us to determine if portfolio effect is at play.
-Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
+\tom{Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.}{This is good and would make a good figure (posteriors of delta r with and without annual variation.)}
 
@@
 \section*{Discussion}
-
-\section*{Conclusion}
 

### 2023-08-15 Manuscript/ant_cactus_AmNat.tex
updates

@@
 
-\subsection*{Data}
-\tom{}{Give this section a more informative title.}	
-The \tom{data collected are from a long-term dataset}{This grammar does not really make sense.} spanning 2004 to 2023 \tom{taken from 30 $\times$ 30 meter plots}{These plots only started in 2009.} at SEV-LTER.
+\subsection*{Data Collection}
+We used a longterm dataset spanning 2004 to 2023 taken from natural plants (2004 -- 2008) and 30 $\times$ 30 meter plots (established in 2009) at SEV-LTER. 
 The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
@@
 For each cholla, we recorded plant survival from the last survey to the current survey. 
-For surviving plants, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm), which are used to calculate plant volume ($cm^3$) based on the \tom{volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter}{Check your volume function but I think it is actually the volume of an elliptical cone.}.
+For surviving plants, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm), which are used to calculate plant volume ($cm^3$) based on the  elliptical volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
 We recorded the total number of flower buds, including how many were aborted and how many were not. 
-\tom{We recorded all ant species present and the number of ants we could count in 30 seconds.}{Need to describe how plants were assigned to an ant state in the case of multiple ant species (rare, I know.)}
+We recorded all ant species present and the number of ants we could count in 30 seconds.
+In the rare case that more than one species of ant was present, we recorded the ant state as the more prevalent of the species, and recorded the identity and number of the secondary species present in the comments.
 We also recorded the numbers and species of herbivores on each plant.
+We conducted recruit and new plant surveys by splitting up each plot and walking the entire area to search for untagged plants. 
+Plants less than 1 cm tall were marked as recruits and larger plants as new plants, these were all tagged, GPS located, and recorded. 
 \tom{}{Also need to describe recruit data, seed banks, seed counts per fruit.}
@@
 
-\subsection*{Integral Projection Model Construction}
+\subsection*{Deterministic Integral Projection Model Construction and Analysis}
 
@@
 
-\subsection*{Deterministic IPM Analysis}
-  \tom{}{The model you show above is the deterministic model.
-The stochastic model is a different model with additional parameters that should be represented symbolically.
-So my point is that I would no introduce the model first and then describe deterministic and stochastic analysis.
-Each of those thngs corresponds to a different model.}
 Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
@@
 
-\subsection*{Stochastic IPM Analysis}
+\subsection*{Stochastic Integral Projection Model Construction and Analysis}
 

### 2023-08-15 Cholla_Analysis_Vital_Rates.R

@@ -123,9 +123,9 @@ for(i in 1:1000){
 }
 ## Plot the simulated data over the real data (separated by ant partners)
 ## This looks pretty good
-png(file = "grow_post.png")
+png(file = "grow_post1.png")
 bayesplot::color_scheme_set(scheme = "pink")
-bayesplot::ppc_dens_overlay_grouped(y, y_rep, group = ant) + xlim(-50,50)
+bayesplot::ppc_dens_overlay_grouped(y, y_sim, group = ant) + xlim(-50,50)
 dev.off()
 ## Plot the convergence of all chains for parameters
 ## They all converge
@@ -201,8 +201,8 @@ points(x = sim_bins2$bin_mean, y = sim_bins2$kurt_sim, col = "pink", pch = 20,ce
 points(x = bins2$bin_mean, y = bins2$kurt_t1, col = "grey")
 dev.off()
 
-png("grow_moments2.png")
-size_moments_ppc(growth_data, "logsize_t1",y_rep,10, title = NA)
+png("grow_moments_ysim.png")
+size_moments_ppc(growth_data, "logsize_t1",y_sim,10, title = NA)
 dev.off()
 
 

### 2023-08-15 Data Analysis/Setup_Script.R

@@ -40,6 +40,8 @@ library(magrittr)
 library(RColorBrewer)
 library(corrplot)
 library(ggpubr)
+library(reshape)
+library(reshape2)
 
 
 

### 2023-08-15 Figures/Cholla_Analysis_Visualization_Vital_Rates.R

@@ -1197,49 +1197,55 @@ dev.off()
 ######################################################################################################################
 ###################### Ant Partner Transitions #######################################################################
 ######################################################################################################################
+#### Create new column that is a unique ID for every
+#### plot and plant then select needed columns
 cactus$individual_ID <- paste(cactus$Plot,cactus$TagID)
 cactus_ants <- cactus[,c("Recruit","individual_ID","Year_t","ant_t1","Newplant","logsize_t1")]
 dim(cactus_ants)
 dim(cactus)
+## remove all NAs
 cactus_ants <- na.omit(cactus_ants)
+## Creat a new column that identifies the year the plant was first recorded
 cactus_ants$obs_yr <- NA
 
 #### Find the earliest year of observation for each individual plant
 cactus_ants %>%
   group_by(individual_ID)  %>%
   summarise(min_yr = min(Year_t)) -> order
-## and record this year as obs_yr in the dataset
 for(i in 1:nrow(cactus_ants)){
-  if(cactus_ants$Recruit[i] == 1| cactus_ants$Newplant[i] == 1 )
-    for(j in 1:nrow(order)){
-      if(cactus_ants$individual_ID[i] == order$individual_ID[j]) cactus_ants$obs_yr[i] <- order$min_yr[j]
-    }
-}
-
-
-for(i in 1:nrow(cactus_ants)){
-  #print(i)
-  #if(cactus_ants$Newplant[i] == 1 | cactus_ants$Recruit[i] == 1) print(i)
   for(j in 1:nrow(order)){
     if(cactus_ants$individual_ID[i] == order$individual_ID[j]) cactus_ants$obs_yr[i] <- order$min_yr[j]
   }
 }
-view(cactus_ants)
-cactus_ants <- subset(cactus_ants, Recruit == 1)
+#### Pull out a subset of plants that we have followed since they were recruits
+## first find the plant IDs of all those we have recruit info for
+cactus_rec <- subset(cactus_ants, Recruit == 1)
+cactus_ants$from_rec <- 0
+## Then match each ID to the actual dataset and mark as from recruit
+for(i in 1:nrow(cactus_rec)){
+  for(j in 1:nrow(cactus_ants)){
[…]

### 2023-08-15 Manuscript/ant_cactus_AmNat.tex
manuscript

@@
 Presence of defensive ant partners is often linked to herbivory reduction \citep{Trager2010, Rudgers2010} and demographic advantages for the plant partner \citep{Baez2016}.
-\tom{These interactions are almost entirely studied from the perspective of plant fitness \citep{Bronstein1994,Bronstein1998}, with little recorded about the impacts on ant fitness \citep{Lanan2013}.}{Here is a good example of something to avoid.
-You are highlighting a gap that your paper actually does not fill.
-Never do this in an Intro, sometimes it can be useful in a discussion as a future direction.}
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
 Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-\tom{Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
-Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.}{I think you are mixing two ideas here: ontogenetic variation in ant-plant associations (eg directional turnover) and temporal stochasticity that is independent of plant ontogeny.
-Your study considers both and I think this needs to be set up more clearly.}
-\tom{Recently many studies have investigated how partner diversity in these guilds has been shown to lead to either increased plant fitness \citep{Palmer2010, Afkhami2014} or decreased plant fitness (often in more highly specialized interactions) \citep{Barrett2015, Ushio2020}, stressing the importance of considering all ant partners as unique within these guilds.
-}{I think you need a more effective argument for why your study is necessary given that there are previous studies that have yielded results in both directions.}
+Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
+Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
+\tom{Recently many studies have investigated how partner diversity in these guilds has been shown to impact plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}, however, none have considered the combined effects of partner identity, directional partner turnover, and temporal stochasticity on the fitness of plants with a longterm dataset.
 

### 2023-08-28 Cholla_Analysis_Vital_Rates.R
tom's growth model updates

@@ -5,6 +5,17 @@
 ##                  TM: the header here should make it clear which scripts need to be run before
 ##                      running this one (for example, there are packages needed here but none are loaded)
 #######################################################################################################
+library(quantreg)
+library(qgam)
+## quantile-based moments
+Q.mean<-function(q.25,q.50,q.75){(q.25+q.50+q.75)/3}
+Q.sd<-function(q.25,q.75){(q.75-q.25)/1.35}
+Q.skewness<-function(q.10,q.50,q.90){(q.10 + q.90 - 2*q.50)/(q.90 - q.10)}
+Q.kurtosis<-function(q.05,q.25,q.75,q.95){
+  qN = qnorm(c(0.05,0.25,0.75,0.95))
+  KG = (qN[4]-qN[1])/(qN[3]-qN[2])
+  return(((q.95-q.05)/(q.75-q.25))/KG - 1)
+}
 
 ## First read the data in 
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
@@ -54,47 +65,103 @@ fit_grow_skew<-sampling(grow_skew_model,data = stan_data_grow_skew,chains=3,
                         control = list(adapt_delta=0.99,stepsize=0.1),
                         iter=10000,cores=3,thin=2,
                         pars = c("u","w",          # plot and year random effects
-                                  "beta0","beta1", #location coefficients
+                                  "beta0","beta1","beta2", #location coefficients
                                  "d_0","d_size", #scale coefficiences
                                  "a_0","a_size"), #shape coefficients
                         save_warmup=F)
+#saveRDS(fit_grow_skew, "C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew.rds")
 #for control parameters see:
 #https://github.com/stan-dev/stan/issues/1504#issuecomment-114685444
[…]

### 2023-08-28 sgt_growth.R

@@ -1,0 +1,55 @@
+library(sgt)
+library(cmdstanr)
+library(bayesplot)
+
+
+cactus <- read.csv("cholla_demography_20042023_cleaned.csv", header = TRUE,stringsAsFactors=T)
+## re-assign the seedling plots ("HT1B1" etc) to transects 1-3
+levels(cactus$Plot)<-c(levels(cactus$Plot),"T4")
+cactus$Plot[cactus$Plot=="HT1B1"]<-"T1"
+cactus$Plot[cactus$Plot=="HT2B3"]<-"T2"
+cactus$Plot[cactus$Plot=="HT3B1" | cactus$Plot=="HT3B2" | cactus$Plot=="HT3B3"]<-"T3"
+cactus$Plot[cactus$Plot=="HT4B1" | cactus$Plot=="HT4B2"]<-"T4"
+
+##############################################################################################
+##
+##   Skew Growth Model -- What size will the cacti be next time step?
+##
+##############################################################################################
+## Pull all necessary variables together, remove NAs, and put them into a list so they are
+## ready to feed into the stan model
+growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
+growth_data <- na.omit(growth_data_orig)
+
+
+## Lose 2032 rows (due to plant death & recruit status)
+nrow(growth_data_orig)
+nrow(growth_data)
+# check that you are happy with the subsetting by plotting the original and cleaned data
+plot(growth_data$logsize_t, growth_data$logsize_t1)
+points((cactus$logsize_t), (cactus$logsize_t1), col = "red")
+## Make a list of all necessary variables so they are properly formatted to feed into the stan model
+stan_data_grow_skew <- list(N = nrow(growth_data),                                ## number of observations
+                            vol = (growth_data$logsize_t), ## predictor volume year t
+                            #vol2 = (growth_data$logsize_t)^2,
+                            y = (growth_data$logsize_t1),                              ## response volume next year
+                            ant = as.integer(as.factor(growth_data$ant_t)),            ## predictor ant state
+                            K = 4,                                                     ## number of ant states
+                            N_Year = max(as.integer(as.factor(growth_data$Year_t))),   ## number of years
+                            N_Plot = max(as.integer(growth_data$Plot)),     ## number of plots
+                            plot = as.integer(growth_data$Plot),            ## predictor plots
[…]

### 2023-08-28 Manuscript/ant_cactus_AmNat.tex

@@
 \title{Thank you, next: partner turnover elevates benefits of mutualism for an ant-tended plant}
-
-% This version of the LaTeX template was last updated on
-% May 11, 2023.
-
-%%%%%%%%%%%%%%%%%%%%%
-% Authorship
-%%%%%%%%%%%%%%%%%%%%%
-% Please remove authorship information while your paper is under review,
-% unless you wish to waive your anonymity under double-blind review. You
-% will need to add this information back in to your final files after
-% acceptance.
 
@@
 
-% The journal does not have numbered sections in the main portion of
-% articles. Please refrain from using section references (à la
-% section~\ref{section:CountingOwlEggs}), and refer to sections by name
-% (e.g. section ``Counting Owl Eggs'').
 Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
@@
 Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
-\tom{Recently many studies have investigated how partner diversity in these guilds has been shown to impact plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}, however, none have considered the combined effects of partner identity, directional partner turnover, and temporal stochasticity on the fitness of plants with a longterm dataset.
+Recently many studies have investigated how partner diversity in these guilds has been shown to impact plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}, however, none have considered the combined effects of partner identity, directional partner turnover, and temporal stochasticity on the fitness of plants with a longterm dataset.
 
@@
 apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between \tom{$0\% - 20\%$}{Are there years with zero CREM?} of cacti \citep{Donald2022} depending on the year.
-\tom{Up to $80\%$ of cacti remain vacant in any given year.
-}{This is probably the extreme high end of vacancy.
-I would give the ranges as you did the others.}
+opuntiae} visiting between $5\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
+Between $ 30\% 80\%$ of cacti remain vacant in any given year.
 These ants rarely co-occur on a plant, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
@@
 \subsection*{Data Collection}
+
 We used a longterm dataset spanning 2004 to 2023 taken from natural plants (2004 -- 2008) and 30 $\times$ 30 meter plots (established in 2009) at SEV-LTER. 
@@
 Many of the vital rates are estimated as a function of plant size, ant partner, or both.
-\tom{Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. }{This needs to be explained.}
-The biological sources of variance (including \tom{individual}{But you do not have an individual random effect.}, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models.
+Ant partner type is included as a predictor in the growth model, survival model, and floral viability model, where there are biological pathways through which ants could impact the outcome of that process.
+The ant partners defend the cacti from herbivores, meaning their presence can lead to higher survival rates and less vegetative structure loss, and seed predators, meaning their presence can lead to fewer attacked and aborted flowers.
+The biological sources of variance (including spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
 Unless otherwise mentioned, all models use vague priors.
 
-\tom{The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+\tom{The growth model ($G(y,x,i)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
 }{I think you need to explain this more thoroughly, and also explain why you used a Skewed Normal.
 Omega and alpha need to be defined and explained, and this is best done showing the full notation of the model.}
+A skew normal distribution was chosen here because the data was skewed.
+$\omega$ denotes the scale of the data in a Skew Normal distribution while $\alpha$ denotes the slant, when $\alpha$ is 0, the $\omega$ is equivalent to the sd of a Normal distribution. 
+The $\omega$ and $\alpha$ estimates vary across size because the scale and skew of the data varied across size. 
 Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
-The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
+
+The survival model ($S(i,x)$) estimates the probability of survival from year $t$ to year $t1$, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution:
+$$y^S ~ Bern(\hat{S})$$
+$$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times i + \beta_{3}^{S} \times x \times i + u + w$$
+where $y_S$ is the probability of survival in year $t1$, $x$ is the size of the plant in the previous year $t$, $i$ is the ant partner of the plant in year $t$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+A Bernoulli distribution was chosen here because the only two possible outcomes are survival or death. 
 Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these.
-The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
-The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
-The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
+
+The reproduction model ($P(y)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution:
+$$y^P ~ Bern(\hat{P})$$
+$$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times y + u + w$$
+where $y^P$ is the probability of reproducing in year $t1$, $y$ is the size of the plant in the year $t1$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+
+The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution:
+$$y^{F} ~ Negative Binom(\hat{F})$$
+$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x + u + w$$
+where $y^{F}$ is the number of flowers produced by a plant in year $t1$, $x$ is the size of the plant in year $t$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+
+The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution:
+$$y^{V} ~ Binom(\hat{V})$$
+$$logit(\hat{V}) = \beta_{0}^{V} \times i + u + w$$
+where $y^{V}$ is the proportion of flowerbuds produced that remain viable in year $t1$, $i$ is the ant partner of the plant in year $t$,  $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion.
+
 The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
 Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
+
 The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
 With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
+
 With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
+
 With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
@@
 \section*{Results}
+\subsection*{Statistical Modeling}
+
+%% Growth Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{grow_contour_lines_color.png}
+%	\caption{asdasdf}
+%	\label{fig:growth}
+%\end{figure}
+Plants tended by other ants experiened positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}a.
+They experienced significantly higher growth than vacant plants ($p = 2.2 \times 10^{-16}$), and similar mean growth to \textit{L.
+apiculatum} tended plants ($p = 0.9135$) and \textit{C.
+opuntiae} tended plants ($p = 0.4864$).
+Plants tended by \textit{C. opuntiae} ants experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}b.
+They experienced significantly higher mean growth than vacant plants ($p = 4.88 \times 10^{-14}$).
+Plants tended by \textit{L. apiculatum} experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}c.
+Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d.
 

### 2023-08-30 Manuscript/ant_cactus_AmNat.tex
Manuscript work

@@
 
-\tom{The growth model ($G(y,x,i)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
-}{I think you need to explain this more thoroughly, and also explain why you used a Skewed Normal.
-Omega and alpha need to be defined and explained, and this is best done showing the full notation of the model.}
-A skew normal distribution was chosen here because the data was skewed.
+The growth model \ali{($G(x',x,i)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
+Should this all be noted like this?} in year $t+1$, with fixed effects of the previous size $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+A skew normal distribution was chosen here because the data had evidence of skew and previous efforts using a gaussian distribution poorly predicted the sizes of these plants, overestimating the next sizes for large plants.
 $\omega$ denotes the scale of the data in a Skew Normal distribution while $\alpha$ denotes the slant, when $\alpha$ is 0, the $\omega$ is equivalent to the sd of a Normal distribution. 
 The $\omega$ and $\alpha$ estimates vary across size because the scale and skew of the data varied across size. 
+$$y^G ~ Skew N(\hat{G},\hat{\omega},\hat{\alpha})$$
+$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times i + \beta_{4}^{G} \times x \times i + u + w$$
+$$\hat{\omega} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x$$
+$$\hat{\alpha} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x + \beta_{2}^{\omega} \times x^2$$
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
 
-The survival model ($S(i,x)$) estimates the probability of survival from year $t$ to year $t1$, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution:
+The survival model ($S(i,x)$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
 $$y^S ~ Bern(\hat{S})$$
 $$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times i + \beta_{3}^{S} \times x \times i + u + w$$
-where $y_S$ is the probability of survival in year $t1$, $x$ is the size of the plant in the previous year $t$, $i$ is the ant partner of the plant in year $t$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 A Bernoulli distribution was chosen here because the only two possible outcomes are survival or death. 
@@
 
-The reproduction model ($P(y)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution:
+The reproduction model ($P(x')$) estimates the probability of reproducing $y^P$ in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
 $$y^P ~ Bern(\hat{P})$$
-$$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times y + u + w$$
-where $y^P$ is the probability of reproducing in year $t1$, $y$ is the size of the plant in the year $t1$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+$$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w$$
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 
-The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution:
-$$y^{F} ~ Negative Binom(\hat{F})$$
+The total flowers model ($F(x)$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x$ in year $t$ and random effects of plot $w$ and year $u$, using a Negative Binomial distribution:
+$$y^{F} ~ Negative Binom(\hat{F},\hat{\phi})$$
 $$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x + u + w$$
-where $y^{F}$ is the number of flowers produced by a plant in year $t1$, $x$ is the size of the plant in year $t$, $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+$$log(\hat{\phi}) = \beta_{0}^{\phi}$$
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 
-The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution:
+The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted) $y^V$ in year $t+1$, with fixed effects of the ant partner of the cactus $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
 $$y^{V} ~ Binom(\hat{V})$$
 $$logit(\hat{V}) = \beta_{0}^{V} \times i + u + w$$
-where $y^{V}$ is the proportion of flowerbuds produced that remain viable in year $t1$, $i$ is the ant partner of the plant in year $t$,  $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion.
 
-The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
+The ant transition rates model ($\tau_(x,i,j)$) estimates the probability of a cactus being visited by an ant partner $j$ $y^{\tau}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $i$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
+$$y^{\tau} ~ Multinomial(\hat{\tau})$$
+$$logit(\tau) = \beta_{0}^{\tau} + \beta_{1}^{\tau} \times x + \beta_{2}^{\tau} \times i + \beta_{3}^{\tau} \times x \times i + u + w$$
+where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
 
-The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
-With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
+The recruit size model ($n_j(x)$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
+$$y^{\eta} ~ N(\hat{\eta},\hat{\sigma})$$
+$$\hat{eta} = \beta_{0}^{\eta}$$
+where $\hat{\sigma}$ is estimated with a non-informative prior.
 
-With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
+With germination data from Miller et al., 2009****, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) in year $t+1$, with no fixed or random effects, using a Binomial distribution:
+$$y^{\gamma_1} ~ Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} ~ Binomial(\hat{\gamma_2})$$
+$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
 
-With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
+$$y^{\delta} ~ Bern(\hat{\delta})$$
+$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+ where $m ~ N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
+With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) in year $t+1$ based on the ant partner $i$ in year $t+1$, using a Negative Binomial distribution:
+$$y^{\kappa} ~ Negative Binomial(\hat{L},\hat{\phi})$$
+$$\hat{L} = \beta_{0}^{L} \times j$$
+$$\hat{\phi} = \beta_{0}^{\phi}$$
 Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds.
-\tom{}{General comment here is that each of these models needs to be described in greater detail and shown in reproducible notation.
-You also need to describe the idea of transitions in size x,y and ant state i,j since this sets the stage for the structure of the IPM.}
 
@@
 Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-\tom{All estimated parameters are described in table 1.}{I think I suggested this years ago but there are so many parameters in your model that I don't think it is particularly useful to report all of these in a table.
-Including publicly available data and code is good enough.}
+All estimated parameters are available in the data. 
 Data and code for all vital rate models is included in the supplemental information.
@@
 While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
-We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
-%Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
-%There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
-This novel structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
+We constructed a deterministic composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
+This novel structure allows us to determine the individual effects of each ant species and the composite effects of several partners across the cholla population.
 
-Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where $x$ is the size of the cholla and $i$ is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
-The dynamics of these 1 and 2-year old seedbanks are given by the following equations: \tom{}{Rather than ``hard-code'' 4 in the summation, I would create a parameter for number of possible ant states, maybe $n_{A}$.}
+Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n(x',j,t+1)$, where $x'$ is the size of the cholla in year $t+1$ and $j$ is the ant partner in year $t+1$, and two discrete seed banks ($B_{1}(t+1)$ and $B_{2}(t+1)$) corresponding to 1 and 2-year old seeds.
+The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
 
@@
 		$$
-		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
+		B_1(t+1) = \delta \sum_{i}^{K} \int_L^U  \kappa(i) P(x') V(i) F(x) n(x,i,t) dx \\
 		$$
 		$$
-		B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
+		B_2(t+1) =  (1 - \gamma_1)B_1(t)\\
 		$$
@@
 
-The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
-The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
-The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+The functions $P(y)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $y$ in year $t+1$ and $x$ in year $t$ respectively. 
+The proportion of flowerbuds which will produce seeds ($V(i)$) is dependent on the ant species present on the plant $i$ in year $t$. 
+The function for the number of seeds per fruit ($\kappa(i)$)  is dependent on the ant partner present $i$ in year $t$.
+These functions are integrated across size, from the smallest observed size $L$ to the largest observed size $U$.
+\ali{}{I know that technically L is smaller than the smallest size and U is larger, but I am not sure if here is the place to put it.
+}
+The integral is multiplied by the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
 Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
@@
 
-The \tom{size dynamics}{It is not just size dynamics, it is size and ant dynamics. That needs to be communicated more clearly.} of the cholla are given by:
+The size and ant dynamics of the cholla are given by:
 
@@
 		$$
-		n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
+		n(y,j,t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(y) \omega \rho_j  + \\
 		$$
 		$$
-		\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
+		\sum_{j}^{K} \int_L^U S(x,i) G(y,x,i) \tau(x,i,j) n(x,i,t) dx \\
 		$$
@@
 
-\tom{The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
-The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
-$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
-Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.}{This is a good start but I think you need a more thorough description, since this model will be unfamiliar to many readers.}
+$n(y,j,t+1)$  explains the dynamics of all plants of size $y$ with ant partner $j$ at time $t+1$. 
+The first term in the equation shows how many of these plants come from the seedbank, either year one or year two seed banks, while the second term describes how many of these plants come from previously existing plants.
+The larger size $y$ is the less likely these plants were recently in the seedbanks. 
+The number of recruits generated from year-one seedbanks is given by the number of seeds in the seedbank $B_1(t)$ multiplied by the year-one seed bank germination rate $\gamma_1$.
+The number of recruits generated from year-two seedbanks is calculated in a similar way with year-two specific estimates. 
+The first term estimates the number of individuals recruited from a one or two-year seed bank to a plant of size $y$ as the number of individuals germinated from a first year seed bank $\gamma_1 B_1(t)$ and a second year seed bank $\gamma_2 B_2(t)$ multiplied by the recruit size distribution term $\eta(y)$, the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
+Finally, this term is multiplied by $\rho_j$, which is an ant specific term which is 1 when $j$ indicates vacancy and 0 when $j$ indicates any ant presence, to ensure all new recruits are vacant as observed in nature.
+Plants not coming from the seedbank to size $y$ and ant state $j$ were previously existing plants $n(x,i,t)$ of size $x$ and ant state $i$.
+The number of these plants can be described by the proportion of plants of size $x$ and ant state $i$ in year $t$ ($n(x,i,t)$)which survive to year $t+1$ ($S(x,i)$), from from size $x$ to size $y$ ($G(y,x,i)$), and transition from ant partner $i$ to $j$ ($\tau(x,i,j)$).
 
 Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-\tom{Each component of the IPM (each set of ant partner specific vital rates) }{I don't think readers will understand what you are talking about here} can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***.
-The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
-We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
-We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
-apiculatum} and vacancy; \textit{C.
+In a traditional IPM $x$ is discretized into $b$ bins, replacing the continuous kernel into a $b \times b$ matrix, in this case there is additional complexity in the form of transitions between ant partners.
+Each combination of previous ant state $i$ and next ant state $j$ represents a unique set of plants in the popualtion and therefore must be discritized individually, leading to a matrix size of $m b \times m b$ where $m$ is the number of unique combinations of $i$ and $j$ (how many possible ant transitions there are).
+In this model we have two additional discrete states (year one and year two seed banks) leading to a final matrix size of $m(b+2) \times m(b+2)$.
+We used $b = 200$ bins.
+We extend the integration limits $L$ and $U$ to avoid unintentional eviction \cite{Williams2012}.
+We estimate the asymptotic population growth rate $\lambda$ as the dominant eigenvalue of the discrete kernel for each combination of ant partners: complete vacancy; \textit{C.
 opuntiae} and vacancy; other and vacancy; \textit{L.
@@
 opuntiae}, other, and vacancy; and all ant partners and vacancy.
-The \tom{relative distributions}{Unclear what this means} allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
-\tom{Sampling Effect requires}{I don't think this is the right grammar (``requires'')} that the $r$ when all possible partners are present is \tom{equal to}{What if it is less than?} the $r$ of the cholla population when only the best partner is present.
-Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner. 
-\tom{}{These last two sentences are very important -- laying out predictions under alternative hypotheses -- so it is work making them laser sharp. }
+Rather than estimating a single $\lambda$ value for each possible population, we used 1,000 random parameter estimations from iterations of our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution.
+We compare the distributions of $\lambda$ across each combination of ant partners to whether sampling effect or complementarity is at play in the system.
+To compare the distributions of two hypothetical populations, A and B, we subtract the vector of 1,000 $\lambda_A$ estimations of one population from the 1,000 $\lambda_B$ estimations of the other ($\lambda_A - \lambda_B$).
+If the average of these differences is positive, population A has a higher $\lambda$ than B, if the average is 0, they are equal, and if the average is negative, population B has a higher $\lambda$ than A.
+
+To determine if sampling effect is at play in a system, we must first determine if there is a "best" partner, by determining which single ant association is correlated with the highest $\lambda$ estimation.
+We can do this by comparing the distributions of each $\lambda$ and finding the one which is larger than all others.
+Then we must show that the $\lambda$ of a population with all possible partners is equal to that of $\lambda$ of a population with only the "best" partner, by comparing the distributions as described earlier.
+If it is significantly larger then complementarity is potentially at play (more on this in the next paragraph).
+If it is significantly smaller, this indicates that rather than positive effects of partner diversity, there are actually important costs of partner diversity that dampen the population growth rate.
+
+To determine if complementarity is at play in a system, we must determine if the partner scenario which leads to the highest $\lambda$ is the most partner diverse scenario.
+We can do this by comparing the distribution of $\lambda$ of a population with all possible partners to the distribution of $\lambda$ of all other populations.
+If the $\lambda$ of a population with all possible partners is the largest, it indicates that complementarity is at play.
+If neither the conditions for sampling effect or complimentarity, it indicates that rather than positive effects of partner diversity, there are actuallly costs of diversity.
 
@@
 The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
-The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
-In order to include stochasticity in the analysis,\tom{ we randomly sampled from the year-random effect intercepts 1,000 times}{I don't think you have communicated, in either words or notation, that the random intercepts are unique to ant state.}.
-For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
-With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
-This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
-We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
-These two relative differences allow us to determine if portfolio effect is at play.
-\tom{Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.}{This is good and would make a good figure (posteriors of delta r with and without annual variation.)}
+This stochastic IPM draws from the same statistical models as the deterministic version, and is described below:
+
+ \begin{linenomath*}
+	$$
+	B_1(t+1) = \delta \sum_{i}^{K} \int_L^U  \kappa(i) P(y,u,w) V(i,u,w) F(x,u,w) n(x,i,t) dx \\
+	$$
+	$$
+	B_2(t+1) =  (1 - \gamma_1)B_1(t)\\
+	$$
+	$$
+	n(y,j,t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(y) \omega \rho_j  + \\
+	$$
+	$$
+	\sum_{j}^{K} \int_L^U S(x,i,u,w) G(y,x,i,u,w) \tau(x,i,j,u,w) n(x,i,t) dx \\
+	$$
+\end{linenomath*}
+
+This system of equations still describes the contribution of seedbanks and existing plants to the group of plants of size $y$ and ant partner $j$ in year $t+1$, but now includes the random effects of plot $u$ and year $w$.
+These effects are included in the form of linear scales on each statistical model where they were biologically relevant.
+Year effects $w$ are both year and ant speciifc, meaning for each year there are different ant and year-specific linear scales for each year of data. 
+When calculating stochastic $\lambda$, every one of the 1,000 parameter-iteration associated $\lambda$s actually come from the mean of 1,000 year-specific estimations. 
+For each parameter estimate taken from the bayesian models we calculate 1,000 estimations of $\lambda$, each associated with a year random effect which was randomly selected from the *******18 years of data.
+The mean of these estimations become the parameter-iteraction specific $\lambda$ estimation. 
+This is repeated 1,000 times for each iteration of parameters to get the stochastic posterior distribution of $\lambda$.
+We compare the stochastic distributions to the deterministic ones to determine if portfolio effect is at play in this system.
+
+To determine if portfolio effect is at play we must show that partner diversity buffers the population from environmental variation (in our case annual variation assessed by the inclusion of year random effects).
+We can estimate the effects of partner diversity on the focal mutualist by calculating the difference in $\lambda$ distributions between populations with no partners and populations with all partners present.
+Then we compare the difference in the effects of partner diversity on the focal mutualist with and without stochasticity.
+If the stochastic difference in the effects of partner diversity is greater than the deterministic difference, this indicates that partner diversity is more beneficial under varying environments and therefore portfolio effect is at play.
 
@@
 Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d.
+
+%% Survival Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{surv_panels.png}
+%	\caption{asdasdf}
+%	\label{fig:surv}
+%\end{figure}
+Plants tended by \textit{C.
+opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
+apiculatum} tended plants ($p = 2.908 \times 10^{-14}$), plants tended by other ants ($p = 4.488 \times 10^{-14}$), and vacant plants ($p = 2.2 \times 10^{-16}$) \cite{fig:surv}.
+These \textit{C. opuntiae} plants have the highest mean survival at small sizes before reaching a mean survival rate of about 100\% when medium to large.
+Plants tended by \textit{L. apiculatum} experienced similar mean survival rates to plants tended by other ants and vacant plants.
+\textit{L. apiculatum} tended plants and other tended plants reached a mean survival rate of nearly 100\% by medium sizes, with other tended plants reaching this rate slower. 
+Vacant plants never reach mean survival rates of 100\%.
+
+%% Flowers Model
+There is a clear size effect on the number of flowers produced. 
+The mean number of flowers produced by a plant remains at 0 until the plant reachesmedium sizes after which the mean number of flower produced increases eponentially to about 40 at large sizes.
+
+%% Viability Model
+%\begin{figure}[h]
+%	\includegraphics[width=0.95\linewidth]{Viability.png}
+%	\caption{asdasdf}
+%	\label{fig:comp-samp}
+%\end{figure}
+Plants tended by \textit{L.
+apiculatum} ants experienced the highest viability rates, significantly higher than \textit{C.
+opuntiae} tended plants ($p = 2.2 \times 10^{-16}$), other tended plants ($p = 2.2 \times 10^{-16}$), and vacant plants ($p = 2.2 \times 10^{-16}$).
+The mean viability rate of \textit{L. apiculatum} tended plants was 84.888\%.
+Plants tended by \textit{C. opuntiae} ants experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$), but not than other tended plants ($p = 0.1168$).
+The mean viability rate of \textit{C. opuntiae} plants was 75.934\%.
+Plants tended by other ants also experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$).
+The mean viability rate of other tended plants was 76.367\%.
+The mean viability rate of vacant plants was 70.442\%.
+
+%% Reproduction Model
+The probability of a plant reproducing in a given year is highly size dependent. 
+The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+
+%% Seeds Produced
+The mean number of seeds produced by \textit{C.
+opuntiae} tended plants is significantly lower than the mean number of seeds produced by \textit{L.
+apiculatum} plants ($p = 2.2 \times 10^{-16}$) and vacant plants ($p = 2.2 \times 10^{-16}$).
+\textit{C. opuntiae} tended plants produce a mean of 2.561 seeds per flower. 
+The mean number of seeds produced by \textit{L. apiculatum} tended plants is slightly higher than the mean number produced by vacant plants ($p = 0.02521$). 
+\textit{L. apiculatum} tended plants produce a mean of 3.563 seeds per flower. 
+Vacant plants produce a mean of 3.435 seeds per flower.
+
+%% Precensus Survival
+The mean pre-census seed survival rate is 16.99\% with 90\% of model iterations landing between 1.13\% and 40.72\%.
+
+%% Germination
+Seeds have a significantly higher probability of germinating in year one than in year two ($p = 2.2 \times 10^{-16}$).
+Seeds in year one have a mean germination rate of 0.578\% and seeds in year two have a mean germination rate of 0.221\%.
+
+%% Recruit size distribution
+The mean size of new recruits is -1.88 log($cm^3$) with 90\% of new recruits falling between the sizes of -2.12 log($cm^3$) and -1.64 log($cm^3$).
+
+%% Ant Transition Rates
+% previously vacant
+Plants that were previously vacant will most likely remain vacant until large. 
+These plants have significantly higher probabilities of being vacant rather than becoming tended  by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+The probability of previously vacant plants becoming tended by \textit{L.
+apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
+opuntiae} or other ants (respecitvely, $p = 3.83 \times 10^{-14}$, $p = 2.2 \times 10^{-16}$).
+The probability of previously vacant plants becoming tended by other ants is significantly lower than the probability of becoming tended by \textit{C. opuntiae} ($p = 0.00227$).
+
+% previously crem
+Plants that were previously tended by \textit{C. opuntiae} will most likely remain vacant until large. 
+These plants have significantly higher probabilities of becoming vacant rather than remaining tended by \textit{C.
+opuntiae} or becoming tended by \textit{L.
+apiculatum} or other ants (repsectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$ ).
+The probability of these plants remaining tended by \textit{C.
+opuntiae} ants is significantly higher than the probability of becoming tended by \textit{L.
+apiculatum} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 3.797 \times 10^{-9}$).
+The probability of these plants becoming tended by \textit{L. apiculatum} is significantly higher than the probability of becoming tended by other ants ($p = 3.797 \times 10^{-9}$).
+
+% previously liom
+Plants that were previously tended by \textit{L. apiculatum} ants will most likely remain vacant until large. 
+These plants have significantly higher probabilities of becoming vacant than remaining tended by \textit{L.
+apiculatum} or becoming tended by \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+These plants have significantly higher probabilities of remaining tended by \textit{L.
+apiculatum} ants than becoming tended by  \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+The probability of these plants becoming tended by \textit{C. opuntiae} ants is significantly higher than the probability of them becoming tended by other ants ($p = 0.007925$).
+
+% previously other
+Plants that were previously tended by other ants will most likely remain vacant until they are large. 
+These plants have significantly higher probabilities of becoming vacant rather than remaining tended by other ants or becoming tended by \textit{C.
+opuntiae} or \textit{L.
+apiculatum} ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+The probability of these plants becoming tended by \textit{L.
+apiculatum} ants is significantly higher than the probability of remaining tended by other ants or becomnig tended by \textit{C.
+opuntiae} (respectively, $p = 6.309 \times 10^{-12}$, $p = 6.256 \times 10^{-13}$).
+The probability of these plants remaining tended by other ants is not higher than the probability of becoming tended by \textit{C. opuntiae} ants ($p = 0.954$).
+
+%% All cases/trends
+In all plants ,the most likely scenario is vacancy.
+In most cases, the second most likely next ant partner is \textit{L.
+apiculatum} with the exception of plants which were previously tended by \textit{C.
+opuntiae} who are second most likely to retain \textit{C.
+opuntiae} as their partners in successive years.
+The only case where two next partners do not have statistically significantly different probabilities is when previously other tended plants either remain tended by other ants or become tended by \textit{C.
+opuntiae}.
+
+\subsection*{Deterministic IPM}
+
+\subsection*{Stochastic IPM}
 

### 2023-09-01 Manuscript/ant_cactus_AmNat.tex
Update ant_cactus_AmNat.tex

@@
 
-The growth model \ali{($G(x',x,i)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
+The growth model \ali{($G(x',x,a)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
 Should this all be noted like this?} in year $t+1$, with fixed effects of the previous size $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
@@
 $$y^G ~ Skew N(\hat{G},\hat{\omega},\hat{\alpha})$$
-$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times i + \beta_{4}^{G} \times x \times i + u + w$$
+$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times a + \beta_{4}^{G} \times x \times a + u + w$$
 $$\hat{\omega} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x$$
@@
 
-The survival model ($S(i,x)$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+The survival model ($S(a,x)$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
 $$y^S ~ Bern(\hat{S})$$
-$$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times i + \beta_{3}^{S} \times x \times i + u + w$$
+$$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times a + \beta_{3}^{S} \times x \times a + u + w$$
 where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
@@
 $$y^{V} ~ Binom(\hat{V})$$
-$$logit(\hat{V}) = \beta_{0}^{V} \times i + u + w$$
+$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
 where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
@@
 
-The ant transition rates model ($\tau_(x,i,j)$) estimates the probability of a cactus being visited by an ant partner $j$ $y^{\tau}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $i$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
+The ant transition rates model ($\tau_(x,a,a')$) estimates the probability of a cactus being visited by an ant partner $a'$ $y^{\tau}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
 $$y^{\tau} ~ Multinomial(\hat{\tau})$$
-$$logit(\tau) = \beta_{0}^{\tau} + \beta_{1}^{\tau} \times x + \beta_{2}^{\tau} \times i + \beta_{3}^{\tau} \times x \times i + u + w$$
+$$logit(\tau) = \beta_{0}^{\tau} + \beta_{1}^{\tau} \times x + \beta_{2}^{\tau} \times a + \beta_{3}^{\tau} \times x \times a + u + w$$
 where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
@@
 
-The recruit size model ($n_j(x)$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
+The recruit size model ($n(x,a')$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
 $$y^{\eta} ~ N(\hat{\eta},\hat{\sigma})$$
@@
 
-Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n(x',j,t+1)$, where $x'$ is the size of the cholla in year $t+1$ and $j$ is the ant partner in year $t+1$, and two discrete seed banks ($B_{1}(t+1)$ and $B_{2}(t+1)$) corresponding to 1 and 2-year old seeds.
+Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n(x',a',t+1)$, where $x'$ is the size of the cholla in year $t+1$ and $a'$ is the ant partner in year $t+1$, and two discrete seed banks ($B_{1}(t+1)$ and $B_{2}(t+1)$) corresponding to 1 and 2-year old seeds.
 The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
@@
 		$$
-		B_1(t+1) = \delta \sum_{i}^{K} \int_L^U  \kappa(i) P(x') V(i) F(x) n(x,i,t) dx \\
+		B_1(t+1) = \delta \sum_{a}^{A} \int_L^U  \kappa(a) P(x') V(a) F(x) n(x,a,t) dx \\
 		$$
@@
 
-The functions $P(y)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $y$ in year $t+1$ and $x$ in year $t$ respectively. 
-The proportion of flowerbuds which will produce seeds ($V(i)$) is dependent on the ant species present on the plant $i$ in year $t$. 
-The function for the number of seeds per fruit ($\kappa(i)$)  is dependent on the ant partner present $i$ in year $t$.
+The functions $P(x')$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x'$ in year $t+1$ and $x$ in year $t$ respectively. 
+The proportion of flowerbuds which will produce seeds ($V(a)$) is dependent on the ant species present on the plant $i$ in year $t$. 
+The function for the number of seeds per fruit ($\kappa(a)$)  is dependent on the ant partner present $a$ in year $t$.
 These functions are integrated across size, from the smallest observed size $L$ to the largest observed size $U$.
@@
 		$$
-		n(y,j,t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(y) \omega \rho_j  + \\
+		n(x',a',t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(x') \omega \rho_{a'}  + \\
 		$$
 		$$
-		\sum_{j}^{K} \int_L^U S(x,i) G(y,x,i) \tau(x,i,j) n(x,i,t) dx \\
+		\sum_{a'}^{A} \int_L^U S(x,a) G(x',x,a) \tau(x,a,a') n(x,a,t) dx \\
 		$$
@@
 
-$n(y,j,t+1)$  explains the dynamics of all plants of size $y$ with ant partner $j$ at time $t+1$. 
+$n(x',a',t+1)$  explains the dynamics of all plants of size $x'$ with ant partner $x'$ at time $t+1$. 
 The first term in the equation shows how many of these plants come from the seedbank, either year one or year two seed banks, while the second term describes how many of these plants come from previously existing plants.
-The larger size $y$ is the less likely these plants were recently in the seedbanks. 
+The larger size $x'$ is the less likely these plants were recently in the seedbanks. 
 The number of recruits generated from year-one seedbanks is given by the number of seeds in the seedbank $B_1(t)$ multiplied by the year-one seed bank germination rate $\gamma_1$.
 The number of recruits generated from year-two seedbanks is calculated in a similar way with year-two specific estimates. 
-The first term estimates the number of individuals recruited from a one or two-year seed bank to a plant of size $y$ as the number of individuals germinated from a first year seed bank $\gamma_1 B_1(t)$ and a second year seed bank $\gamma_2 B_2(t)$ multiplied by the recruit size distribution term $\eta(y)$, the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
-Finally, this term is multiplied by $\rho_j$, which is an ant specific term which is 1 when $j$ indicates vacancy and 0 when $j$ indicates any ant presence, to ensure all new recruits are vacant as observed in nature.
-Plants not coming from the seedbank to size $y$ and ant state $j$ were previously existing plants $n(x,i,t)$ of size $x$ and ant state $i$.
-The number of these plants can be described by the proportion of plants of size $x$ and ant state $i$ in year $t$ ($n(x,i,t)$)which survive to year $t+1$ ($S(x,i)$), from from size $x$ to size $y$ ($G(y,x,i)$), and transition from ant partner $i$ to $j$ ($\tau(x,i,j)$).
+The first term estimates the number of individuals recruited from a one or two-year seed bank to a plant of size $x'$ as the number of individuals germinated from a first year seed bank $\gamma_1 B_1(t)$ and a second year seed bank $\gamma_2 B_2(t)$ multiplied by the recruit size distribution term $\eta(x')$, the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
+Finally, this term is multiplied by $\rho_{a'}$, which is an ant specific term which is 1 when $a'$ indicates vacancy and 0 when $a'$ indicates any ant presence, to ensure all new recruits are vacant as observed in nature.
+Plants not coming from the seedbank to size $x'$ and ant state $a'$ were previously existing plants $n(x,a,t)$ of size $x$ and ant state $a$.
+The number of these plants can be described by the proportion of plants of size $x$ and ant state $a$ in year $t$ ($n(x,a,t)$)which survive to year $t+1$ ($S(x,a)$), from from size $x$ to size $x'$ ($G(x',x,a)$), and transition from ant partner $a$ to $a'$ ($\tau(x,a,a')$).
 
@@
 In a traditional IPM $x$ is discretized into $b$ bins, replacing the continuous kernel into a $b \times b$ matrix, in this case there is additional complexity in the form of transitions between ant partners.
-Each combination of previous ant state $i$ and next ant state $j$ represents a unique set of plants in the popualtion and therefore must be discritized individually, leading to a matrix size of $m b \times m b$ where $m$ is the number of unique combinations of $i$ and $j$ (how many possible ant transitions there are).
+Each combination of previous ant state $a$ and next ant state $a'$ represents a unique set of plants in the popualtion and therefore must be discritized individually, leading to a matrix size of $m b \times m b$ where $m$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
 In this model we have two additional discrete states (year one and year two seed banks) leading to a final matrix size of $m(b+2) \times m(b+2)$.
@@
 	$$
-	B_1(t+1) = \delta \sum_{i}^{K} \int_L^U  \kappa(i) P(y,u,w) V(i,u,w) F(x,u,w) n(x,i,t) dx \\
+	B_1(t+1) = \delta \sum_{a}^{A} \int_L^U  \kappa(a) P(x',u,w) V(a,u,w) F(x,u,w) n(x,a,t) dx \\
 	$$
@@
 	$$
-	n(y,j,t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(y) \omega \rho_j  + \\
+	n(x',a',t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(x') \omega \rho_{a'}  + \\
 	$$
 	$$
-	\sum_{j}^{K} \int_L^U S(x,i,u,w) G(y,x,i,u,w) \tau(x,i,j,u,w) n(x,i,t) dx \\
+	\sum_{a}^{A} \int_L^U S(x,a,u,w) G(x',x,a,u,w) \tau(x,a,a',u,w) n(x,a,t) dx \\
 	$$
@@
 
-This system of equations still describes the contribution of seedbanks and existing plants to the group of plants of size $y$ and ant partner $j$ in year $t+1$, but now includes the random effects of plot $u$ and year $w$.
+This system of equations still describes the contribution of seedbanks and existing plants to the group of plants of size $x'$ and ant partner $a'$ in year $t+1$, but now includes the random effects of plot $u$ and year $w$.
 These effects are included in the form of linear scales on each statistical model where they were biologically relevant.

### 2023-09-01 Cholla_Analysis_Vital_Rates.R
Update Cholla_Analysis_Vital_Rates.R

@@ -76,8 +76,9 @@ saveRDS(fit_grow_skew, "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutuali
 #https://mc-stan.org/rstanarm/reference/adapt_delta.html
 fit_grow_skew<-readRDS("C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew.rds")
 fit_grow_skew<-readRDS("/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew.rds")
-
-bayesplot::mcmc_trace(fit_grow_skew,pars=c("d_0","d_size","a_0","a_size"))
+fit_grow_skew<-readRDS("/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew.rds")
+fit_grow_skew@model_pars
+bayesplot::mcmc_trace(fit_grow_skew,pars=c("d_0","d_size","d_size2","a_0","a_size","a_size2"))
 bayesplot::mcmc_trace(fit_grow_skew,pars=c("beta0[1]","beta0[2]","beta0[3]","beta0[4]"))
 bayesplot::mcmc_trace(fit_grow_skew,pars=c("beta1[1]","beta1[2]","beta1[3]","beta1[4]"))
 bayesplot::mcmc_trace(fit_grow_skew,pars=c("beta2[1]","beta2[2]","beta2[3]","beta2[4]"))

### 2023-09-14 Manuscript/ant_cactus_AmNat.tex
manuscript pushes

@@
 Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014bee}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014}.
 
@@
 Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
-Presence of defensive ant partners is often linked to herbivory reduction \citep{Trager2010, Rudgers2010} and demographic advantages for the plant partner \citep{Baez2016}.
+Presence of defensive ant partners is often linked to herbivory reduction \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
@@
 Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
-These ant species locally co-occur at the scale of meters, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022does}.
+These ant species locally co-occur at the scale of meters, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
@@
     %% Growth Parameters
-    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
-    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
-    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
-    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
-    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
-    growth xi size dependent other $\Beta_{12}^g$ & & \\
-    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
-    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+    growth xi intercept vacant $\beta_{01}^g$ & $-5.210899 (-5.686865, -5.491787)$ & sDE\\
+    growth xi intercept other $\beta_{02}^g$ & $-5.8288 (-5.956217, 1.766021) $&asdf \\
+    growth xi intercept \textit{C. opuntiae} $\beta_{03}^g$ & $-4.529523 (-6.0770390, 0.1222112)$ & asdf\\
+    growth xi intercept \textit{L. apiculatum} $\beta_{04}^g$ & $-5.106802 (-5.4499944, 0.5453901)$ & asdf\\
+    growth xi size dependent vacant $\beta_{11}^g$ & asdf&asdf \\
+    growth xi size dependent other $\beta_{12}^g$ & asdf&asdf \\
+    growth xi size dependent \textit{C. opuntiae} $\beta_{13}^g$ & asdf&asdf \\
+    growth xi size dependent \textit{L. apiculatum} $\beta_{14}^g$ &sadf &asdf \\
     growth omega intercept $\omega_0^g$ & & \\
@@
     %% Survival Parameters
-    survival intercept vacant $\Beta_{01}^s$ & & \\
-    survival intercept other $\Beta_{02}^s$ & & \\
-    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & & \\
-    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & & \\
-    survival size dependent vacant $\Beta_{11}^s$ & & \\
-    survival size dependent other $\Beta_{12}^s$ & & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & & \\
+    survival intercept vacant $\beta_{01}^s$ & & \\
+    survival intercept other $\beta_{02}^s$ & & \\
+    survival intercept \textit{C.opuntiae} $\beta_{03}^s$ & & \\
+    survival intercept \textit{L. apiculatum} $\beta_{04}^s$ & & \\
+    survival size dependent vacant $\beta_{11}^s$ & & \\
+    survival size dependent other $\beta_{12}^s$ & & \\
+    survival size dependent \textit{C. opuntiae} $\beta_{13}^s$ & & \\
+    survival size dependent \textit{L. apiculatum} $\beta_{14}^s$ & & \\
     \hline
     %% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & & \\
-    flowering size dependent $\Beta_1^f$ & & \\
+    flowering intercept $\beta_0^f$ & & \\
+    flowering size dependent $\beta_1^f$ & & \\
     \hline
     %% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & & \\
-    viability intercept other $\Beta02^v$ & & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & & 
+    viability intercept vacant $\beta_01^v$ & & \\
+    viability intercept other $\beta02^v$ & & \\
+    viability intercept \textit{C. opuntiae} $\beta_03^v$ & & \\
+    viability intercept \textit{L. apiculatum} $\beta_04^v$ & & 
   \end{tabular}
@@
 
-%%%%%%%%%%%%%%%%%%%%%
-% Videos
-%%%%%%%%%%%%%%%%%%%%%
-% If you have videos, journal style for them is generally similar to that for
-% figures.
+%%%%%%%%%%%%%%%%%%%%
+ Videos
+%%%%%%%%%%%%%%%%%%%%
+ If you have videos, journal style for them is generally similar to that for
+ figures.
 
-%%%%% Include the text below if you have videos
+%%%% Include the text below if you have videos
 
-%%%%% Include the above if you have videos
+%%%% Include the above if you have videos
 

### 2023-09-14 Manuscript/ant_cactus_AmNat_Ali.tex
manuscript pushes

--- 2023-06-09 Manuscript/ant_cactus_AmNat.tex
@@
 
-\section*{Methods}
-
-\section*{Results}
-
-\section*{Discussion}
-
-\section*{Conclusion}
-
-\section*{Acknowledgments}
+%\section*{Methods}
+%\subsection*{Study System}
+%  
+%This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
+%This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
+%Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \citep{Benson1982}. 
+%These arborescent cholla produce cylindrical segments with large spines. 
+%In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+%While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
+%Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
+%Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \citep{Ness2006,Oliveira1999}.
+%
+%This EFN is harvested by various ant species in return for defense. 
+%At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
+%\textit{L.
+apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
+opuntiae} visiting between $0\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
+Up to $80\%$ of cacti remain vacant in any given year.
+%
+%These ants rarely co-occur on a plant, probably due to interspecific competition \citep{Miller2007}.
+%Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
+%At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
+%They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \citep{Ohm2014}. 
+%Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \citep{Miller2014}. 
+%In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season. 
+%
+%There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \citep{Mann1969}. 
+%An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
+%A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
+%A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
+%These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \citep{Miller2009}.
+%There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \citep{Miller2007}. 
+%
+%\subsection*{Data}
+%	
+%The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
+%The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
+%Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
+%The final two plots were added to this census from 2011 onwards. 
+%Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
+%For each cholla, we recorded plant survival from the last survey to the current survey. 
+%We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
+%We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
+%We also recorded the species of herbivore and the number present on the plant.
+%		
+%\subsection*{Statistical Modeling -- Construction}
+%	
+%With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
+%Many of the vital rates are estimated as a function of plant size, ant partner, or both.
+%Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
+%The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
+%Unless otherwise mentioned, all models use vague priors. 
+%The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+%Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
+%The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
+%Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these. 
+%The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
+%The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
+%The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
+%Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion. 
+%The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
+%Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
+%The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
+%With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
+%With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
+%With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
+%Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds. 
+%
+%To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
+%For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
+%The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
+%We did not thin the chains, thus all samples were retained. 
+%To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
+%To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
+%Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
+%All estimated parameters are described in table 1.
+%Data and code for all vital rate models is included in the supplemental information.
+%
+%\subsection*{Integral Projection Model Construction}
+%  
+%Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
+%We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
+%Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
+%There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
+%This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population. 
+%
+%Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
+%The dynamics of these 1 and 2-year old seedbanks are given by the following equations: 
+%
+%  \begin{linenomath*}
+%		$$
+%		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
+%		$$
+%		$$
+%		B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
+%		$$
+%  \end{linenomath*}
+%
+%The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
+%The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
+%The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+%Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
+%Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
+%Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die. 
+%		
+%The size dynamics of the cholla are given by:
+%		
+%	\begin{linenomath*}
+%		$$
+%		n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
+%		$$
+%		$$
+%		\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
+%		$$
+%	\end{linenomath*}
+%	
+%The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
+%The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
+%$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
+%Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
+%
+%\subsection*{Deterministic IPM Analysis}
+%  
+%Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
+%Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***. 
+%The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
+%We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
+%We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
+apiculatum} and vacancy; \textit{C.
+opuntiae} and vacancy; other and vacancy; \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and vacancy, \textit{L.
+apiculatum}, other, and vacancy; \textit{C.
+opuntiae}, other, and vacancy; and all ant partners and vacancy.
+%The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
+%Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
+%Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner. 
+%
+%
+%\subsection*{Stochastic IPM Analysis}
+%  
+%The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
+%The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
+%In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
+%For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
+%With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
+%This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
+%We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
+%These two relative differences allow us to determine if portfolio effect is at play.
+%Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation. 
+%			
+%    
+%  
+%\section*{Results}
+%\subsection*{Statistical Modeling}
+%
+%%% Growth Model
+%%\begin{figure}[h]
+%%	\includegraphics[width=0.95\linewidth]{grow_contour_lines_color.png}
+%%	\caption{asdasdf}
+%%	\label{fig:growth}
+%%\end{figure}
+%Plants tended by other ants experiened positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}a.
+%They experienced significantly higher growth than vacant plants ($p = 2.2 \times 10^{-16}$), and similar mean growth to \textit{L.
+apiculatum} tended plants ($p = 0.9135$) and \textit{C.
+opuntiae} tended plants ($p = 0.4864$).
+%Plants tended by \textit{C. opuntiae} ants experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}b.
+%They experienced significantly higher mean growth than vacant plants ($p = 4.88 \times 10^{-14}$).
+%Plants tended by \textit{L. apiculatum} experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}c.
+%Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d. 
+%
+%%% Survival Model
+%%\begin{figure}[h]
+%%	\includegraphics[width=0.95\linewidth]{surv_panels.png}
+%%	\caption{asdasdf}
+%%	\label{fig:surv}
+%%\end{figure}
+%Plants tended by \textit{C.
+opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
+apiculatum} tended plants ($p = 2.908 \times 10^{-14}$), plants tended by other ants ($p = 4.488 \times 10^{-14}$), and vacant plants ($p = 2.2 \times 10^{-16}$) \cite{fig:surv}.
+%These \textit{C. opuntiae} plants have the highest mean survival at small sizes before reaching a mean survival rate of about 100\% when medium to large.
+%Plants tended by \textit{L. apiculatum} experienced similar mean survival rates to plants tended by other ants and vacant plants.
+%\textit{L. apiculatum} tended plants and other tended plants reached a mean survival rate of nearly 100\% by medium sizes, with other tended plants reaching this rate slower. 
+%Vacant plants never reach mean survival rates of 100\%. 
+%
+%%% Flowers Model
+%There is a clear size effect on the number of flowers produced. 
+%The mean number of flowers produced by a plant remains at 0 until the plant reachesmedium sizes after which the mean number of flower produced increases eponentially to about 40 at large sizes.
+%
+%
+%%% Viability Model
+%%\begin{figure}[h]
+%%	\includegraphics[width=0.95\linewidth]{Viability.png}
+%%	\caption{asdasdf}
+%%	\label{fig:comp-samp}
+%%\end{figure}
+%Plants tended by \textit{L.
+apiculatum} ants experienced the highest viability rates, significantly higher than \textit{C.
+opuntiae} tended plants ($p = 2.2 \times 10^{-16}$), other tended plants ($p = 2.2 \times 10^{-16}$), and vacant plants ($p = 2.2 \times 10^{-16}$).
+%The mean viability rate of \textit{L. apiculatum} tended plants was 84.888\%.
+%Plants tended by \textit{C. opuntiae} ants experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$), but not than other tended plants ($p = 0.1168$).
+%The mean viability rate of \textit{C. opuntiae} plants was 75.934\%.
+%Plants tended by other ants also experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$).
+%The mean viability rate of other tended plants was 76.367\%.
+%The mean viability rate of vacant plants was 70.442\%.
+%
+%%% Reproduction Model
+%The probability of a plant reproducing in a given year is highly size dependent. 
+%The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+%
+%%% Seeds Produced
+%The mean number of seeds produced by \textit{C.
+opuntiae} tended plants is significantly lower than the mean number of seeds produced by \textit{L.
+apiculatum} plants ($p = 2.2 \times 10^{-16}$) and vacant plants ($p = 2.2 \times 10^{-16}$).
+%\textit{C. opuntiae} tended plants produce a mean of 2.561 seeds per flower. 
+%The mean number of seeds produced by \textit{L. apiculatum} tended plants is slightly higher than the mean number produced by vacant plants ($p = 0.02521$). 
+%\textit{L. apiculatum} tended plants produce a mean of 3.563 seeds per flower. 
+%Vacant plants produce a mean of 3.435 seeds per flower. 
+%
+%%% Precensus Survival
+%The mean pre-census seed survival rate is 16.99\% with 90\% of model iterations landing between 1.13\% and 40.72\%. 
+%
+%%% Germination
+%Seeds have a significantly higher probability of germinating in year one than in year two ($p = 2.2 \times 10^{-16}$).
+%Seeds in year one have a mean germination rate of 0.578\% and seeds in year two have a mean germination rate of 0.221\%. 
+%
+%%% Recruit size distribution
+% The mean size of new recruits is -1.88 log($cm^3$) with 90\% of new recruits falling between the sizes of -2.12 log($cm^3$) and -1.64 log($cm^3$).
+% 
+%%% Ant Transition Rates
+%% previously vacant
+%Plants that were previously vacant will most likely remain vacant until large. 
+%These plants have significantly higher probabilities of being vacant rather than becoming tended  by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+%The probability of previously vacant plants becoming tended by \textit{L.
+apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
+opuntiae} or other ants (respecitvely, $p = 3.83 \times 10^{-14}$, $p = 2.2 \times 10^{-16}$).
+%The probability of previously vacant plants becoming tended by other ants is significantly lower than the probability of becoming tended by \textit{C. opuntiae} ($p = 0.00227$).
+%
+%% previously crem
+%Plants that were previously tended by \textit{C. opuntiae} will most likely remain vacant until large. 
+%These plants have significantly higher probabilities of becoming vacant rather than remaining tended by \textit{C.
+opuntiae} or becoming tended by \textit{L.
+apiculatum} or other ants (repsectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$ ).
+%The probability of these plants remaining tended by \textit{C.
+opuntiae} ants is significantly higher than the probability of becoming tended by \textit{L.
+apiculatum} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 3.797 \times 10^{-9}$).
+%The probability of these plants becoming tended by \textit{L. apiculatum} is significantly higher than the probability of becoming tended by other ants ($p = 3.797 \times 10^{-9}$).
+%
+%% previously liom
+%Plants that were previously tended by \textit{L. apiculatum} ants will most likely remain vacant until large. 
+%These plants have significantly higher probabilities of becoming vacant than remaining tended by \textit{L.
+apiculatum} or becoming tended by \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
+%These plants have significantly higher probabilities of remaining tended by \textit{L.
+apiculatum} ants than becoming tended by  \textit{C.
+opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+%The probability of these plants becoming tended by \textit{C. opuntiae} ants is significantly higher than the probability of them becoming tended by other ants ($p = 0.007925$).
+%
+%% previously other
+%Plants that were previously tended by other ants will most likely remain vacant until they are large. 
+%These plants have significantly higher probabilities of becoming vacant rather than remaining tended by other ants or becoming tended by \textit{C.
+opuntiae} or \textit{L.
+apiculatum} ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
+%The probability of these plants becoming tended by \textit{L.
+apiculatum} ants is significantly higher than the probability of remaining tended by other ants or becomnig tended by \textit{C.
+opuntiae} (respectively, $p = 6.309 \times 10^{-12}$, $p = 6.256 \times 10^{-13}$).
+%The probability of these plants remaining tended by other ants is not higher than the probability of becoming tended by \textit{C. opuntiae} ants ($p = 0.954$). 
+%
+%%% All cases/trends
+%In all plants ,the most likely scenario is vacancy.
+%In most cases, the second most likely next ant partner is \textit{L.
+apiculatum} with the exception of plants which were previously tended by \textit{C.
+opuntiae} who are second most likely to retain \textit{C.
+opuntiae} as their partners in successive years.
+%The only case where two next partners do not have statistically significantly different probabilities is when previously other tended plants either remain tended by other ants or become tended by \textit{C.
+opuntiae}.
+%
+%\subsection*{Deterministic IPM}
+%
+%\subsection*{Stochastic IPM}
+%
+%
+%\section*{Discussion}
+%
+%
+%\section*{Conclusion}
+%
+%\section*{Acknowledgments}
 
@@
 
+\renewcommand{\thetable}{\arabic{table}}
+\setcounter{table}{0}
+
+%% I am creating a table here to include all parameter estimations and descriptions
+%  \begin{table}[]
+%  \begin{tabular}{l|l|l}
+%    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
+%    \hline
+%    %% Growth Parameters
+%    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
+%    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
+%    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
+%    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
+%    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
+%    growth xi size dependent other $\Beta_{12}^g$ & & \\
+%    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
+%    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+%    growth omega intercept $\omega_0^g$ & & \\
+%    growth omega size dependent $\omega_1^g$ & & \\
+%    growth alpha intercept $\alpha_0^g$ & & \\
+%    growth alpha size dependent $\alpha_1^g$ & & \\
+%    \hline
+%    %% Germination Parameters
+%    1-year germination intercept $\alpha^{\gamma_1}$ & -5.144627 (-5.370185, -4.966587) & \\
+%    2-year germination intercept $\alpha^{\gamma_2}$ & -6.155088 (-6.417991, -5.786996) & \\
+%    \hline
+%    %% Survival Parameters
+%    survival intercept vacant $\Beta_{01}^s$ & 0.46375959 (0.07778875, 0.86267803) & \\
+%    survival intercept other $\Beta_{02}^s$ & 0.2677732 (-1.7939375, 2.5857164) & \\
+%    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & 2.2690061 (0.7347587, 3.9967755) & \\
+%    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & 0.0538502 (-1.0054015, 1.3153753) & \\
+%    survival size dependent vacant $\Beta_{11}^s$ & 0.86267803 (0.2696006, 0.3217293) & \\
+%    survival size dependent other $\Beta_{12}^s$ & 0.4387909 (0.1953732, 0.7325813) & \\
+%    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & 0.28392994 (0.08742771, 0.47423484) & \\
+%    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & 0.5150193 (0.5150193, 0.6492092) & \\
+%    \hline
+%    %% Probability of Flowering Parameters
+%    flowering intercept $\Beta_0^f$ & -4.611414 (-5.119275, -4.082511) & \\
+%    flowering size dependent $\Beta_1^f$ & 0.5693555 (0.5239602, 0.6104472) & \\
+%    \hline
+%    %% Floral Viability Parameters
+%    viability intercept vacant $\Beta_01^v$ & 1.0930266 (0.5002784, 1.6629479) & \\
+%    viability intercept other $\Beta02^v$ & 1.1102242 (0.5073163, 1.7414692) & \\
+%    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & 1.3605548 (0.7372043, 1.9359609) & \\
+%    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & 2.023599 (1.461603, 2.534776) & 
+%  \end{tabular}
+%  \end{table}
+%\bigskip{}
+%\\
+%\end{table}
+
 \newpage{}

### 2023-09-14 Manuscript/ant_cactus_AmNat.tex
Update ant_cactus_AmNat.tex

@@
 \usepackage{fullpage}
-\usepackage[authoryear,sectionbib,sort]{natbib}
+%\usepackage[authoryear,sectionbib,sort]{natbib}
+\usepackage[square,sort,comma,numbers]{natbib}
 \linespread{1.7}
@@
 \usepackage{titlesec}
-\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
-\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+%\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
+%\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
 
@@
 Plants less than 1 cm tall were marked as recruits and larger plants as new plants, these were all tagged, GPS located, and recorded. 
-\tom{}{Also need to describe recruit data, seed banks, seed counts per fruit.}
+%\tom{}{Also need to describe recruit data, seed banks, seed counts per fruit.}
 
@@
 
-The growth model \ali{($G(x',x,a)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
+The growth model %\ali{($G(x',x,a)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
 Should this all be noted like this?} in year $t+1$, with fixed effects of the previous size $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
@@
 
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \tom{}{You need to cite R and RStan}
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R %\tom{}{You need to cite R and RStan}
 For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
@@
 These functions are integrated across size, from the smallest observed size $L$ to the largest observed size $U$.
-\ali{}{I know that technically L is smaller than the smallest size and U is larger, but I am not sure if here is the place to put it.
+%\ali{}{I know that technically L is smaller than the smallest size and U is larger, but I am not sure if here is the place to put it.
 }
@@
   \end{table}
-\bigskip{}
-\\
-\end{table}
-
-\newpage{}
 

### 2023-09-29 sgt_growth.R
Remove outdated files

@@ -1,55 +1,0 @@
-library(sgt)
-library(cmdstanr)
-library(bayesplot)
-
-
-cactus <- read.csv("cholla_demography_20042023_cleaned.csv", header = TRUE,stringsAsFactors=T)
-## re-assign the seedling plots ("HT1B1" etc) to transects 1-3
-levels(cactus$Plot)<-c(levels(cactus$Plot),"T4")
-cactus$Plot[cactus$Plot=="HT1B1"]<-"T1"
-cactus$Plot[cactus$Plot=="HT2B3"]<-"T2"
-cactus$Plot[cactus$Plot=="HT3B1" | cactus$Plot=="HT3B2" | cactus$Plot=="HT3B3"]<-"T3"
-cactus$Plot[cactus$Plot=="HT4B1" | cactus$Plot=="HT4B2"]<-"T4"
-
-##############################################################################################
-##
-##   Skew Growth Model -- What size will the cacti be next time step?
-##
-##############################################################################################
-## Pull all necessary variables together, remove NAs, and put them into a list so they are
-## ready to feed into the stan model
-growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
-growth_data <- na.omit(growth_data_orig)
-
-
-## Lose 2032 rows (due to plant death & recruit status)
-nrow(growth_data_orig)
-nrow(growth_data)
-# check that you are happy with the subsetting by plotting the original and cleaned data
-plot(growth_data$logsize_t, growth_data$logsize_t1)
-points((cactus$logsize_t), (cactus$logsize_t1), col = "red")
-## Make a list of all necessary variables so they are properly formatted to feed into the stan model
-stan_data_grow_skew <- list(N = nrow(growth_data),                                ## number of observations
-                            vol = (growth_data$logsize_t), ## predictor volume year t
-                            #vol2 = (growth_data$logsize_t)^2,
-                            y = (growth_data$logsize_t1),                              ## response volume next year
-                            ant = as.integer(as.factor(growth_data$ant_t)),            ## predictor ant state
-                            K = 4,                                                     ## number of ant states
-                            N_Year = max(as.integer(as.factor(growth_data$Year_t))),   ## number of years
-                            N_Plot = max(as.integer(growth_data$Plot)),     ## number of plots
-                            plot = as.integer(growth_data$Plot),            ## predictor plots
[…]

### 2023-09-29 skew_normal_chisq_mixture.R

@@ -1,70 +1,0 @@
-#see https://discourse.mc-stan.org/t/skewed-distribution/1383/15
-
-library(rstan)
-scode <- "
-data {
-  real<lower=1> nu;
-  real alpha;
-}
-transformed data {
-  real sqrt_nu = sqrt(nu);
-}
-parameters {
-  real<lower=0> V;
-  real Z;
-}
-transformed parameters {
-  real T = Z * sqrt(V) * sqrt_nu;
-}
-model {
-  V ~ inv_chi_square(nu); 
-  Z ~ skew_normal(0, 1, alpha);
- }
-"
-
-foo_data <- list(nu = 4, alpha = 1)
-
-foo <- stan(model_code = scode, data = foo_data, chains = 1, iter = 4000, control=list(adapt_delta=0.8))
-T <- extract(foo)$T
-
-###############################################
-library(extraDistr)
-x<-runif(1000,0,100)
-beta0<--2
-beta1<-0.8
-y<-rlst(n=1000,df = 6,mu=beta0+beta1*x,sigma = 8)
-hist(y)
-plot(x,y)
-
-toycode <- "
-data {
-  int <lower = 1> N;
-  vector[N] x;
-  vector[N] y;
-}
-parameters {
-  real alpha;
-  real beta0;
-  real beta1;
-  real<lower=0> tausq;
-  real<lower=1> nu;
-}
-transformed parameters {
-  vector[N] xi;
-  vector[N] omega;
-
-  for(i in 1:N){
-  xi[i] = beta0 + beta1 * x[i];
-  }
-}
-model {
-  beta0 ~ normal(0,5);
-  beta1 ~ normal(0,5);
-  alpha ~ normal(0,5);
-
-  omega ~ scaled_inv_chi_square(nu, tausq);
-  y ~ skew_normal(xi, omega, alpha);
- }
-"
-toy_data <- list(x = x, y = y, N = length(y))
-toy_fit <- stan(model_code = toycode, data = toy_data, chains = 1, iter = 400, control=list(adapt_delta=0.99))

### 2023-09-29 toy_sim_skew_norm.R

@@ -1,90 +1,0 @@
-n_sim<-1000
-x<-runif(n_sim,-4,16)
-beta_xi0<-4
-beta_xi1<-2.6
-beta_omega0<- -0.5
-beta_omega1<- 0.2
-beta_alpha0<- -5
-beta_alpha1<- -2
-
-y<-rsn(n=length(x),
-       xi = beta_xi0+beta_xi1*x,
-       omega = exp(beta_omega0 +beta_omega1*x),
-       alpha = beta_alpha0+beta_alpha1*x)
-
-hist(y)
-plot(x,y)
-lines(x,beta_xi0+beta_xi1*x,col="red")
-
-stan_data_skew_sim <- list(N = n_sim,  
-                            x = x,
-                            y = y)
-########## growth model with a skew normal distribution -- fixed effects: previous size and ant state, ##############
-########## random effects: plot and year, size variation is included for both the omega and alpha estimates #########
-# fit_grow_skew <- stan(file = "Data Analysis/STAN Models/grow_skew_toy_sim.stan", data = stan_data_skew_sim, 
-#                       warmup = 1000, iter = 5000, chains = 3, cores = 3, thin = 2)
-# bayesplot::mcmc_trace(fit_grow_skew,pars=c("beta0","beta1","d_0","d_size","a_0","a_size"))
-# 
-# sim_params<-rstan::extract(fit_grow_skew,c("beta0","beta1","d_0","d_size","a_0","a_size"))
-# samples<-sample.int(100,n=3000,replace=F)
-# 
-y_sim<-matrix(NA,100,n_sim)
-for(i in 1:100){
-  y_sim[i,]<-rsn(n=length(x),
-         xi = sim_params$beta0[i]+sim_params$beta1[i]*x,
-         omega = exp(sim_params$d_0[i] +sim_params$d_size[i]*x),
-         alpha = sim_params$a_0[i]+sim_params$a_size[i]*x)
-}
-n_bins<-10
-sim_dat<-data.frame(x=x,y=y)
-  require(tidyverse)
-  require(patchwork)
-  bins <- sim_dat %>%
-    ungroup() %>% 
-    arrange(x) %>% 
-    mutate(size_bin = cut_number(x, n_bins)) %>% 
-    group_by(size_bin)  %>% 
-    dplyr::summarize(mean_t1 = mean(y),
-                     sd_t1 = sd(y),
-                     skew_t1 = skewness(y),
-                     kurt_t1 = Lkurtosis(y),
[…]

### 2023-09-29 Cholla_Analysis_Vital_Rates.R
Update Cholla_Analysis_Vital_Rates.R

@@ -442,7 +442,9 @@ t.test(estimates[,2],estimates[,4], alternative = "two.sided")
 t.test(estimates[,3],estimates[,4], alternative = "two.sided")
 
 #####################################################################################################
-##### Reproductive State Model -- Prob of reproducing at next time step #############################
+##### 
+#####        Reproductive State Model -- Prob of reproducing at next time step           #############################
+##### 
 #####################################################################################################
 ## Repro Data Set
 reproductive_data_orig <- cactus[ , c("flower1_YN","logsize_t","Year_t","Plot", "logsize_t1")]

### 2023-09-29 Manuscript/ant_cactus_AmNat_Ali.tex
Update ant_cactus_AmNat_Ali.tex

--- 2023-08-02 Manuscript/ant_cactus_AmNat_Ali.tex
@@
 
-\section*{Methods}
+\textsc{\textsc{%\section*{Methods}
 \subsection*{Study System}
@@
 %\end{figure}
-Plants tended by \texit{C.
+Plants tended by \textit{C.
 opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
@@
 apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
-The probability of previously vacant plants becoming tended by \texit{L.
+The probability of previously vacant plants becoming tended by \textit{L.
 apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
@@
 
+Populations of tree cholla that have no ant partners present have a mean fitness of $\lambda = 0.9895$.
+Populations with only one ant partner present (\textit{L.
+apiculatum}, \textit{C.
+opuntiae}, or other ants) have a mean fitness of $\lambda = 0.9993$, $\lambda = 0.9853$, and $\lambda = 0.9903$ respectively.
+Populations with two ant partners present (\textit{L.apiculatum} and \textit{C.
+opuntiae}, \textit{L.apiculatum} and other, or \textit{C.
+opuntiae} and other) have a mean fitness of $\lambda = 1.0035$, $\lambda = 1.004$, and $\lambda = 0.9962$ respectively.
+Populations with all ant partners present have a mean fitness of $\lambda = 1.0039$.
+The only partner scenarios where $\lambda > 1$ are when \textit{L. apiculatum} ants are present with any other ant partner.
+
+The means are not the only important result.
+We used bayesian modeling, so each of these $\lambda$ estimations has a distribution rather than a single estimate.
+By subtracting one of these distributions from the other, the difference from 0 tells us how much larger or smaller the distributions are with a percent confidence. 
+% single partners
+We are 52\% confident that having \textit{C.
+opuntiae} partners only leads to a lower $\lambda$ than having no partners, indicating these are actually parasites rather than mutualists as has long been suggested.
+We are 56\% and 84\% confident that having other ants or \textit{L. apiculatum} ants respectively as the only partner leads to a higher $\lambda$ than having no partners.
+\textit{L. apiculatum} ants are the best single partner for tree cholla populations, as they lead to the highest population fitness with a single partner.
+% two partners
+
+% all partners
+
 \subsection*{Stochastic IPM}
@@
 
-%%%%%%%%%%%%%%%%%%%%%
-% Statement of Authorship
-%%%%%%%%%%%%%%%%%%%%%
-% This section should also be commented out while your MS is undergoing
-% double-blind review. The specifics should of course be adapted to
-% your paper, but the paragraph below gives some hints of possible
-% contributions.
+%%%%%%%%%%%%%%%%%%%%
+ Statement of Authorship
+%%%%%%%%%%%%%%%%%%%%
+ This section should also be commented out while your MS is undergoing
+ double-blind review. The specifics should of course be adapted to
+ your paper, but the paragraph below gives some hints of possible
+ contributions.}}
 
@@
 
-% I am creating a table here to include all parameter estimations and descriptions
-  \begin{table}[]
-  \begin{tabular}{l|l|l}
-    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
-    \hline
-    %% Growth Parameters
-    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
-    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
-    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
-    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
-    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
-    growth xi size dependent other $\Beta_{12}^g$ & & \\
-    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
-    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
-    growth omega intercept $\omega_0^g$ & & \\
-    growth omega size dependent $\omega_1^g$ & & \\
-    growth alpha intercept $\alpha_0^g$ & & \\
-    growth alpha size dependent $\alpha_1^g$ & & \\
-    \hline
-    %% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & -5.144627 (-5.370185, -4.966587) & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & -6.155088 (-6.417991, -5.786996) & \\
-    \hline
-    %% Survival Parameters
-    survival intercept vacant $\Beta_{01}^s$ & 0.46375959 (0.07778875, 0.86267803) & \\
-    survival intercept other $\Beta_{02}^s$ & 0.2677732 (-1.7939375, 2.5857164) & \\
-    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & 2.2690061 (0.7347587, 3.9967755) & \\
-    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & 0.0538502 (-1.0054015, 1.3153753) & \\
-    survival size dependent vacant $\Beta_{11}^s$ & 0.86267803 (0.2696006, 0.3217293) & \\
-    survival size dependent other $\Beta_{12}^s$ & 0.4387909 (0.1953732, 0.7325813) & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & 0.28392994 (0.08742771, 0.47423484) & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & 0.5150193 (0.5150193, 0.6492092) & \\
-    \hline
-    %% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & -4.611414 (-5.119275, -4.082511) & \\
-    flowering size dependent $\Beta_1^f$ & 0.5693555 (0.5239602, 0.6104472) & \\
-    \hline
-    %% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & 1.0930266 (0.5002784, 1.6629479) & \\
-    viability intercept other $\Beta02^v$ & 1.1102242 (0.5073163, 1.7414692) & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & 1.3605548 (0.7372043, 1.9359609) & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & 2.023599 (1.461603, 2.534776) & 
-  \end{tabular}
-  \end{table}
-\bigskip{}
-\\
-\end{table}
+%% I am creating a table here to include all parameter estimations and descriptions
+%  \begin{table}[]
+%  \begin{tabular}{l|l|l}
+%    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
+%    \hline
+%    %% Growth Parameters
+%    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
+%    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
+%    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
+%    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
+%    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
+%    growth xi size dependent other $\Beta_{12}^g$ & & \\
+%    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
+%    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+%    growth omega intercept $\omega_0^g$ & & \\
+%    growth omega size dependent $\omega_1^g$ & & \\
+%    growth alpha intercept $\alpha_0^g$ & & \\
+%    growth alpha size dependent $\alpha_1^g$ & & \\
+%    \hline
+%    %% Germination Parameters
+%    1-year germination intercept $\alpha^{\gamma_1}$ & -5.144627 (-5.370185, -4.966587) & \\
+%    2-year germination intercept $\alpha^{\gamma_2}$ & -6.155088 (-6.417991, -5.786996) & \\
+%    \hline
+%    %% Survival Parameters
+%    survival intercept vacant $\Beta_{01}^s$ & 0.46375959 (0.07778875, 0.86267803) & \\
+%    survival intercept other $\Beta_{02}^s$ & 0.2677732 (-1.7939375, 2.5857164) & \\
+%    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & 2.2690061 (0.7347587, 3.9967755) & \\
+%    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & 0.0538502 (-1.0054015, 1.3153753) & \\
+%    survival size dependent vacant $\Beta_{11}^s$ & 0.86267803 (0.2696006, 0.3217293) & \\
+%    survival size dependent other $\Beta_{12}^s$ & 0.4387909 (0.1953732, 0.7325813) & \\
+%    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & 0.28392994 (0.08742771, 0.47423484) & \\
+%    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & 0.5150193 (0.5150193, 0.6492092) & \\
+%    \hline
+%    %% Probability of Flowering Parameters
+%    flowering intercept $\Beta_0^f$ & -4.611414 (-5.119275, -4.082511) & \\
+%    flowering size dependent $\Beta_1^f$ & 0.5693555 (0.5239602, 0.6104472) & \\
+%    \hline
+%    %% Floral Viability Parameters
+%    viability intercept vacant $\Beta_01^v$ & 1.0930266 (0.5002784, 1.6629479) & \\
+%    viability intercept other $\Beta02^v$ & 1.1102242 (0.5073163, 1.7414692) & \\
+%    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & 1.3605548 (0.7372043, 1.9359609) & \\
+%    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & 2.023599 (1.461603, 2.534776) & 
+%  \end{tabular}
+%  \end{table}
+%\bigskip{}
+%\\
+%\end{table}
 

### 2023-10-12 Manuscript/ant_cactus_AmNat.tex
ms edits

@@
 \documentclass[11pt]{article}
+\usepackage{amsmath}
 \usepackage[sc]{mathpazo} %Like Palatino with extensive math support
@@
 \usepackage{titlesec}
-%\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
-%\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+\usepackage{xcolor}
+\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}} 
+\newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}
 
@@
 \newpage{}
-
 \section*{Introduction}
-
-Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013}, but can deteriorate into commensalism or parasitism  \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
+They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
 Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities.
@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (cite BEF chapter).
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (\textbf{cite BEF chapter}).
 First, when there is a hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
 This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
-However, if partner associations are mutually exclusive then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time) then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
 This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
-Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions (cite). 
-Third, partner species can have species-specific responses to the environment, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions (\textbf{cite}). 
+Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
 Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
@@
 Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist (cite examples, including non-ant-plant examples). 
+Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist (\tom{cite examples, including non-ant-plant examples}{Flagging this.}).
 Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
-The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004}.
-Directionality of turnover can also influence diversity effects, particularly if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003}.
+The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate with successive turnover (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004}.
+Directionality of turnover can also influence effects of partner diversity, \tom{particularly}{Two `particularly' sentences in a row.} if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003}.
 For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005}.
@@
 Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
-Presence of defensive ant partners is often linked to herbivory reduction \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
+Presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
 Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
-Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
-Recently many studies have investigated how partner diversity in these guilds has been shown to impact plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}, however, none have considered the combined effects of partner identity, directional partner turnover, and temporal stochasticity on the fitness of plants with a longterm dataset.
+\tom{Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.}{Some grammatical errors or typos in this sentence, hard to follow.}
+Temporal dynamics of ant visitation therefore can have important impacts on the fitness of the plant partners \citep{Barton2010, Boege2005, Fonseca2003}.
+\tom{Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}.}{Only one of the studies you cite here is actually ant-plant.
+I think you are missing some important literature.}
+However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data required to detect portfolio effects are rarely available.
 
@@
 Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
-These ant species locally co-occur at the scale of meters, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022}.
+These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
 Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the benefits provided by alternative partners, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
+However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the benefits provided by alternative partners \tom{(which may arise from fluctuations in herbivore pressure or identity)}{This is awkwardly inserted here but I think we need to explain that the reason the portfolio effect may arise is the threat to plants can vary from year to year}, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
 
@@
 	\end{enumerate}
-We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy, and to quantify state-dependent partner turnover. 
+We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
 We then used a stochastic, multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
@@
 Tree cholla EFN is harvested by various ant species. 
-At SEV-LTER, cholla are visited primarily by two species of formicoid, ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster}, and a \textit{Camponotus}.
+At SEV-LTER, cholla are visited primarily by two species of formicoid, ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster} and \textit{Camponotus}.
 \textit{L.
@@
 \subsection*{Data Collection}
+This study is based a long-term demographic data set spanning 2004 to 2023 at SEV-LTER. 
+From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
+This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots has since been censused annually through 2023 (with the exception of 2020 due to the covid shutdown). 
+For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
+Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
+We recorded reproductive effort as counts of viable and aborted flowerbuds. 
+We recorded the ant species present (or vacancy if no ants present) \tom{and the number of ants we could count in 30 seconds}{As long as we are not using this there is no need to report it.}.
+Occurrences of more than one ant species on one plant were rare (QUANTITY), and in these cases we classified the plant as being occupied by the more abundant species. 
+\tom{We also recorded the numbers and identities of herbivores on each plant.}{Again, if this is not used then do not report this here.}
+Plots were searched for new recruits each year, and these were added to the census.
+In total, the data set included \# unique individuals and \# plant-year observations. 
+These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
 
-We used a longterm dataset spanning 2004 to 2023 taken from natural plants (2004 -- 2008) and 30 $\times$ 30 meter plots (established in 2009) at SEV-LTER. 
-The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
-Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
-The final two plots were added to this census from 2011 onwards. 
-Annually, in May we surveyed all individuals in these plots, taking demographic and partner data. 
-For each cholla, we recorded plant survival from the last survey to the current survey. 
-For surviving plants, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm), which are used to calculate plant volume ($cm^3$) based on the  elliptical volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-We recorded the total number of flower buds, including how many were aborted and how many were not. 
-We recorded all ant species present and the number of ants we could count in 30 seconds.
-In the rare case that more than one species of ant was present, we recorded the ant state as the more prevalent of the species, and recorded the identity and number of the secondary species present in the comments.
-We also recorded the numbers and species of herbivores on each plant.
-We conducted recruit and new plant surveys by splitting up each plot and walking the entire area to search for untagged plants. 
-Plants less than 1 cm tall were marked as recruits and larger plants as new plants, these were all tagged, GPS located, and recorded. 
-%\tom{}{Also need to describe recruit data, seed banks, seed counts per fruit.}
+We used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
+Ohm et al. \citep{Ohm2014} provide data on the number of seeds per fruit for plants tended by \textit{L. apiculatum}, \textit{C. opuntiae}, or no ants (experimental exclusion). 
+Miller et al. \citep{Miller2009} provide data on seed entry to the seed bank and seedling germination and survival rates.
 
-\subsection*{Statistical Modeling}
+\subsection*{Multi-state Integral Projection Model}
+Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables. 
+While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
+We constructed a multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with their transition dynamics modeled explictly.
 
-With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework to serve as vital rate sub-models of the IPM.
-Many of the vital rates are estimated as a function of plant size, ant partner, or both.
-Ant partner type is included as a predictor in the growth model, survival model, and floral viability model, where there are biological pathways through which ants could impact the outcome of that process.
-The ant partners defend the cacti from herbivores, meaning their presence can lead to higher survival rates and less vegetative structure loss, and seed predators, meaning their presence can lead to fewer attacked and aborted flowers.
-The biological sources of variance (including spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
+Given the low frequency of ant species other than \textit{L.
+apiculatum} and \textit{C.
+opuntiae} (QUANTIFY THIS) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and other.
+\tom{}{Need a sentence here describing the composition of ``other''.}
+Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
+Ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
+Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
+In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independent of its influence on plant size. 
+Therefore, these sub-models included plant size but not ant state as predictor variables.
+
+Following previous studies, we modeled the tree cholla life cycle using continuously size-structured plants where $n(x,a)_{t}$ gives the number of plants of size $x$ and ant state $a$ in year $t$, plus two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds.
+Seed bank dynamics are given by:
+
+\begin{linenomath*}
+	$$
+	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau^P_{a}}) F(x;\pmb{\tau^F_{a}}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
+	$$
+	$$
+	B^2_{t+1} =  (1 - \gamma_1)B^1_{t}\\
+	$$
+\end{linenomath*}
+
+Functions $P(x;\pmb{\tau^P_{a}})$ and $F(x;\pmb{\tau^F_{a}})$ give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x$ in year $t$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$ but not size. 
+The vector $\pmb{\tau_{a}}$ gives year-specific deviates (with mean zero) that modify the intercepts of the functions in which it appears. 
+This vector appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
+Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
+Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.
+
+For the above-ground part of the life cycle, the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) is given by survival/growth transitions from size $x$ and ant state $a$ in year $t$, plus germination out of the seed banks:
+\begin{linenomath*}
+	$$
+	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{a'}  + \\
+	$$
+	$$
+	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \tau(x,a,a') n(x,a,t) dx \\
+	$$
+\end{linenomath*}
+
+The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
+This term is multiplied by $\rho_{a'}$, which gives the probability that a new recruit has ant state $a'$ at its first appears in our census ($\sum_{a'}^{A}\rho_{a'}=1$). 
+The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
+Survival ($S(x,a;\pmb{\tau^S_{a}})$) and growth ($G(x',x,a;\pmb{\tau^G_{a}})$) are both dependent on initial size and ant state. 
+As above, these functions accommodate inter-annual variability through year-specific intercepts that can vary by ant state ($\pmb{\tau_{a}}$). 
+Ant transition function \tom{$\tau(a',a,x)$}{Are there random intercepts in the ant transition function, and are they ant-specific?} gives the probability that a plant with size $x$ and ant state $a$ transitions to ant state $a'$ in the next census.
+
+\subsection*{Statistical modeling and parameter estimation}
+We parameterized the IPM using a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework to serve as vital rate sub-models. 
+Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively, and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
+In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transition between ant states conditional on plant size and previous ant state. 
+All models were fit using Stan and Rstan (CITE). 
 Unless otherwise mentioned, all models use vague priors.
 
-The growth model %\ali{($G(x',x,a)$) estimates the size of cholla $y^G$}{This is where I get a little confused between the x and x' sizes and the ys.
-Should this all be noted like this?} in year $t+1$, with fixed effects of the previous size $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
+\paragraph{Growth}
+The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given previous size $x$ and previous ant partner $a$. 
+We fit this model to size data at the end of the transition year $y^G$ using the location-scale parameterization of the student $t$ distribution, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
+Specifically, the model was:
+
+\begin{align}
+y^G ~ Skew N(\hat{G},\hat{\omega},\hat{\alpha}) \\
+\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times a + \beta_{4}^{G} \times x \times a + u + w \\
+\hat{\omega} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x \\
+\hat{\alpha} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x + \beta_{2}^{\omega} \times x^2
+\end{align}
+
+\tom{in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size. 
 A skew normal distribution was chosen here because the data had evidence of skew and previous efforts using a gaussian distribution poorly predicted the sizes of these plants, overestimating the next sizes for large plants.
@@
 The $\omega$ and $\alpha$ estimates vary across size because the scale and skew of the data varied across size. 
-$$y^G ~ Skew N(\hat{G},\hat{\omega},\hat{\alpha})$$
-$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times a + \beta_{4}^{G} \times x \times a + u + w$$
-$$\hat{\omega} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x$$
-$$\hat{\alpha} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x + \beta_{2}^{\omega} \times x^2$$
 where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
+Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.}{I am skipping all of this for now so that you can update with new growth model, then I will circle back.}
 
@@
 
-\subsection*{Deterministic Integral Projection Model Construction and Analysis}
-
-Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables. 
-While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
-We constructed a deterministic composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
-This novel structure allows us to determine the individual effects of each ant species and the composite effects of several partners across the cholla population.
-
-Following previous studies of this population, we modeled the life cycle of cholla using continuously size-structured plants, $n(x',a',t+1)$, where $x'$ is the size of the cholla in year $t+1$ and $a'$ is the ant partner in year $t+1$, and two discrete seed banks ($B_{1}(t+1)$ and $B_{2}(t+1)$) corresponding to 1 and 2-year old seeds.
-The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
-
-  \begin{linenomath*}
-		$$
-		B_1(t+1) = \delta \sum_{a}^{A} \int_L^U  \kappa(a) P(x') V(a) F(x) n(x,a,t) dx \\
-		$$
-		$$
-		B_2(t+1) =  (1 - \gamma_1)B_1(t)\\
-		$$
-  \end{linenomath*}
-
-The functions $P(x')$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x'$ in year $t+1$ and $x$ in year $t$ respectively. 
-The proportion of flowerbuds which will produce seeds ($V(a)$) is dependent on the ant species present on the plant $i$ in year $t$. 
-The function for the number of seeds per fruit ($\kappa(a)$)  is dependent on the ant partner present $a$ in year $t$.
-These functions are integrated across size, from the smallest observed size $L$ to the largest observed size $U$.
-%\ali{}{I know that technically L is smaller than the smallest size and U is larger, but I am not sure if here is the place to put it.
-}
-The integral is multiplied by the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
-Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
-Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
-Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
-
-The size and ant dynamics of the cholla are given by:
-
-	\begin{linenomath*}
-		$$
-		n(x',a',t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(x') \omega \rho_{a'}  + \\
-		$$
-		$$
-		\sum_{a'}^{A} \int_L^U S(x,a) G(x',x,a) \tau(x,a,a') n(x,a,t) dx \\
-		$$
-	\end{linenomath*}
-
-$n(x',a',t+1)$  explains the dynamics of all plants of size $x'$ with ant partner $x'$ at time $t+1$. 
-The first term in the equation shows how many of these plants come from the seedbank, either year one or year two seed banks, while the second term describes how many of these plants come from previously existing plants.
-The larger size $x'$ is the less likely these plants were recently in the seedbanks. 
-The number of recruits generated from year-one seedbanks is given by the number of seeds in the seedbank $B_1(t)$ multiplied by the year-one seed bank germination rate $\gamma_1$.
-The number of recruits generated from year-two seedbanks is calculated in a similar way with year-two specific estimates. 
-The first term estimates the number of individuals recruited from a one or two-year seed bank to a plant of size $x'$ as the number of individuals germinated from a first year seed bank $\gamma_1 B_1(t)$ and a second year seed bank $\gamma_2 B_2(t)$ multiplied by the recruit size distribution term $\eta(x')$, the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
-Finally, this term is multiplied by $\rho_{a'}$, which is an ant specific term which is 1 when $a'$ indicates vacancy and 0 when $a'$ indicates any ant presence, to ensure all new recruits are vacant as observed in nature.
-Plants not coming from the seedbank to size $x'$ and ant state $a'$ were previously existing plants $n(x,a,t)$ of size $x$ and ant state $a$.
-The number of these plants can be described by the proportion of plants of size $x$ and ant state $a$ in year $t$ ($n(x,a,t)$)which survive to year $t+1$ ($S(x,a)$), from from size $x$ to size $x'$ ($G(x',x,a)$), and transition from ant partner $a$ to $a'$ ($\tau(x,a,a')$).
+\subsection*{IPM Analysis}
 

### 2023-10-18 IPM Code/Analysis.R
Updates

@@ -153,17 +153,17 @@ lams_stoch_null <- lams_stoch_null[,-c(1)]
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
 
 ## Plot the means of the deterministic and stochastic distributions together
-png("lambda_means.png")
-plot(c(1,3,5,7,9,11,13,15),colMeans(lams_dpost),
-     col = cols, pch = 20, cex = 5,
-     xlim = c(0,16), ylim = c(0.985,1.01),
+png("lambda_portfolios.png")
+plot(c(1,3,5,7,9,11,13,15),colMeans(lams_stoch),
+     col = cols, pch = 1, cex = 5,
+     xlim = c(0,16), ylim = c(0.985,1.0),
      xaxt = "n",cex.lab = 2,
      xlab = "Ant Scenario", ylab = "Mean Lambda Value", main = "Full Partner Diversity Leads to \n Highest Fitness")
-     text(x = c(1,3,5,7,9,11,13,15)-0.2, y = colMeans(lams_dpost)+0.002,cex = 2, labels = lams$scenario_abv,srt = 35)
+     text(x = c(1,3,5,7,9,11,13,15)-0.2, y = colMeans(lams_stoch)+0.002,cex = 2, labels = lams$scenario_abv,srt = 35)
      legend("topleft",legend = c("L = Liom.","C = Crem.","O = Other"),cex = 1.5)
-points(c(1.5,3.5,5.5,7.5,9.5,11.5,13.5,15.5),colMeans(lams_stoch),
-       col = cols, cex = 5)
-points(c(1.5,3.5,5.5,7.5,9.5,11.5,13.5,15.5),colMeans(lams_stoch_null), cex = 5, pch = 20)
+#points(c(1.5,3.5,5.5,7.5,9.5,11.5,13.5,15.5),colMeans(lams_stoch),
+       #col = cols, cex = 5)
+points(c(1.5,3.5,5.5,7.5,9.5,11.5,13.5,15.5),colMeans(lams_stoch_null), cex = 5, pch = 18, col = cols)
 dev.off()
 
 ## plot the distributions of the deterministic distribution 
@@ -258,6 +258,18 @@ mtext("Lambda",side=1,line=-2,outer=TRUE,cex=1.3)
 mtext("Density",side=2,line=-2,outer=TRUE,cex=1.3,las=0)
 legend("topright",legend = c("All Ants"),fill = c(acol),
[…]

### 2024-01-16 Manuscript/References_Cacti.tex
Updates

@@
 }
+@article{Wetzel2023,
+	abstract = {Plants and herbivores are remarkably variable in space and time, and variability has been considered a defining feature of their interactions.
+Empirical research, however, has traditionally focused on understanding differences in means and overlooked the theoretically significant ecological and evolutionary roles of variability itself.
+We review the literature with the goal of showing how variability-explicit research expands our perspective on plant-herbivore ecology and evolution.
+We first clarify terminology for describing variation and then review patterns, causes, and consequences of variation in herbivory across scales of space, time, and biological organization.
+We consider how incorporating variability improves existing hypotheses and leads to new ones.
+We conclude by suggesting future work that reports full distributions, integrates effects of variation across scales, describes nonlin-earities, and considers how stochastic and deterministic variation combine to determine herbivory distributions.
+451},
+	author = {William C Wetzel and Brian D Inouye and Philip G Hahn and Susan R Whitehead and Nora Underwood},
+	doi = {10.1146/annurev-ecolsys-102221},
+	journal = {Annu. Rev. Ecol. Evol. Syst. 2023},
+	keywords = {defense,ecological variability,herbivore,herbivory,plant,species interactions},
+	pages = {2023},
+	title = {Annual Review of Ecology, Evolution, and Systematics Variability in Plant-Herbivore Interactions},
+	volume = {54},
+	url = {https://doi.org/10.1146/annurev-ecolsys-102221-},
+	year = {2023},
+}

### 2024-01-16 Manuscript/ant_cactus_AmNat.tex
Updates

@@
 \usepackage{xcolor}
+\usepackage{graphicx}
+\usepackage{placeins}
+\usepackage{float}
+\usepackage{subfigure}
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}} 
@@
 
-\author{Alexandra Campbell$^{1,\ast}$ \\ 
-	Tom E.X. Miller$^{1}$}
+\author{Alexandra Campbell$^{1,\dagger}$ \\ 
+	Tom E.X. Miller$^{1,\ast}$}
 
@@
 
-	\noindent{} $\ast$ Corresponding author; e-mail: amc49@rice.edu.
-
-	\textit{Manuscript elements}:
+	\noindent{} $\dagger$ e-mail: amc49@rice.edu\\
+	\noindent{} $\ast$ e-mail: tom.miller@rice.edu
 
@@
 
-	\textit{Keywords}:
+	\textit{Keywords}:  Integral Projection Model, \textit{Cylindropuntia imbricata}, population fitness, multi-species mutualism, complementarity, sampling effect, portfolio effect
 
@@
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities.
+They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \cite{Noe1994,Leigh2010}.
 
-Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010}.
+Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}.
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (\textbf{cite BEF chapter}).
-First, when there is a hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \cite{Yeung2006,Barrett2015,Ushio2020}.
+First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
 This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
-However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time) then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
 This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
-Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions (\textbf{cite}). 
-Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
+Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \cite{Afkhami2014}. 
+Third, partner species and herbivores can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits and protection across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
 
-Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist (\tom{cite examples, including non-ant-plant examples}{Flagging this.}).
+Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
+Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \cite{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
 Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
-The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate with successive turnover (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004}.
-Directionality of turnover can also influence effects of partner diversity, \tom{particularly}{Two `particularly' sentences in a row.} if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003}.
-For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005}.
+The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate with successive turnover (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004,Fiala1994}.
+Directionality of turnover can also influence effects of partner diversity if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003,Noe1994,Dejean2008}.
+For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
 
@@
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
-Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-\tom{Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.}{Some grammatical errors or typos in this sentence, hard to follow.}
+Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
 Temporal dynamics of ant visitation therefore can have important impacts on the fitness of the plant partners \citep{Barton2010, Boege2005, Fonseca2003}.
-\tom{Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010, Afkhami2014,Barrett2015, Ushio2020}.}{Only one of the studies you cite here is actually ant-plant.
-I think you are missing some important literature.}
+Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face varies just as much as the protection they recieve due to temporal stochasticity.
+Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data required to detect portfolio effects are rarely available.
@@
 Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
-These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022}.
+These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014, Donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
 Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the benefits provided by alternative partners \tom{(which may arise from fluctuations in herbivore pressure or identity)}{This is awkwardly inserted here but I think we need to explain that the reason the portfolio effect may arise is the threat to plants can vary from year to year}, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
+However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the herbivore populations.
 
@@
 We recorded reproductive effort as counts of viable and aborted flowerbuds. 
-We recorded the ant species present (or vacancy if no ants present) \tom{and the number of ants we could count in 30 seconds}{As long as we are not using this there is no need to report it.}.
+We recorded the ant species present (or vacancy if no ants present).
 Occurrences of more than one ant species on one plant were rare (QUANTITY), and in these cases we classified the plant as being occupied by the more abundant species. 
-\tom{We also recorded the numbers and identities of herbivores on each plant.}{Again, if this is not used then do not report this here.}
 Plots were searched for new recruits each year, and these were added to the census.
@@
 apiculatum}, \textit{C.
-opuntiae}, and other.
-\tom{}{Need a sentence here describing the composition of ``other''.}
+opuntiae}, and ``other''.
+The ``other'' category was made of unidentified ant species and ant species which occurred at relatively low frequencies compared to \textit{C.
+opuntiae} and \textit{L.
+apiculatum} (such as unknown species belonging to the genus \textit{Aphenogaster} and unknown species belonging to the genus \textit{Camponotus}).
 Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
@@
 	$$
-	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau^P_{a}}) F(x;\pmb{\tau^F_{a}}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
+	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x';\pmb{\tau^P}) F(x';\pmb{\tau^F}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
 	$$
@@
 
-Functions $P(x;\pmb{\tau^P_{a}})$ and $F(x;\pmb{\tau^F_{a}})$ give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x$ in year $t$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$ but not size. 
-The vector $\pmb{\tau_{a}}$ gives year-specific deviates (with mean zero) that modify the intercepts of the functions in which it appears. 
+Functions $P(x';\pmb{\tau^P})$ and $F(x';\pmb{\tau^F})$ give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x'$ in year $t$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a')$) are dependent on ant state $a$ but not size. 
+The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) that modify the intercepts of the functions in which it appears. 
 This vector appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
-Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A$) giving total seed production. 
 Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
@@
 	$$
-	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{a'}  + \\
+	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho(a')  + \\
 	$$
 	$$
-	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \tau(x,a,a') n(x,a,t) dx \\
+	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \epsilon(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
 	$$
@@
 The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
-This term is multiplied by $\rho_{a'}$, which gives the probability that a new recruit has ant state $a'$ at its first appears in our census ($\sum_{a'}^{A}\rho_{a'}=1$). 
+This term is multiplied by $\rho(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appears in our census ($\sum_{a'}^{A}\rho(a')=1$). 
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
@@
 As above, these functions accommodate inter-annual variability through year-specific intercepts that can vary by ant state ($\pmb{\tau_{a}}$). 
-Ant transition function \tom{$\tau(a',a,x)$}{Are there random intercepts in the ant transition function, and are they ant-specific?} gives the probability that a plant with size $x$ and ant state $a$ transitions to ant state $a'$ in the next census.
+Ant transition function $\epsilon(a',a,x;\pmb{\tau^{\epsilon}})$ gives the probability that a plant with size $x$ and ant state $a$ transitions to ant state $a'$ in the next census. 
+This function accomodates inter-annual variability through year-specific intercepts which are stable across ant states ($\pmb{\tau}$).
 
@@
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given previous size $x$ and previous ant partner $a$. 
+The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$.
 We fit this model to size data at the end of the transition year $y^G$ using the location-scale parameterization of the student $t$ distribution, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
 Specifically, the model was: 
+$$y^G \sim Student T(\hat{nu},\hat{G},\hat{\sigma}) $$
+$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x  + \beta_{2}^{G} \times a + \beta_{3}^{G} \times x \times a  + \beta_{4}^{G} \times x^2 + \beta_{5}^{G} \times a \times x^2 + u + w $$
+$$\hat{\sigma} = \beta_{0}^{\sigma} + \beta_{1}^{\sigma} \times x $$
+$$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu} \times x $$
+where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ with ant $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+Ants are included as predictors here because ant partners defend cholla from herbivores, preventing the loss of limbs and therefore shrinking.
 
-\begin{align}
-y^G ~ Skew N(\hat{G},\hat{\omega},\hat{\alpha}) \\
-\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x + \beta_{2}^{G} \times a + \beta_{4}^{G} \times x \times a + u + w \\
-\hat{\omega} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x \\
-\hat{\alpha} = \beta_{0}^{\omega} + \beta_{1}^{\omega} \times x + \beta_{2}^{\omega} \times x^2
-\end{align}
-
-\tom{in year $t$ and random effects of plot $w$ and year $u$, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size. 
-A skew normal distribution was chosen here because the data had evidence of skew and previous efforts using a gaussian distribution poorly predicted the sizes of these plants, overestimating the next sizes for large plants.
-$\omega$ denotes the scale of the data in a Skew Normal distribution while $\alpha$ denotes the slant, when $\alpha$ is 0, the $\omega$ is equivalent to the sd of a Normal distribution. 
-The $\omega$ and $\alpha$ estimates vary across size because the scale and skew of the data varied across size. 
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.}{I am skipping all of this for now so that you can update with new growth model, then I will circle back.}
-
-The survival model ($S(a,x)$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
-$$y^S ~ Bern(\hat{S})$$
+\paragraph{Survival}
+The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+$$y^S \sim Bern(\hat{S})$$
 $$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times a + \beta_{3}^{S} \times x \times a + u + w$$
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 A Bernoulli distribution was chosen here because the only two possible outcomes are survival or death. 
@@
 
-The reproduction model ($P(x')$) estimates the probability of reproducing $y^P$ in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
-$$y^P ~ Bern(\hat{P})$$
+\paragraph{Reproduction}
+The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing $y^P$ in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+$$y^P \sim Bern(\hat{P})$$
 $$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w$$
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 
-The total flowers model ($F(x)$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x$ in year $t$ and random effects of plot $w$ and year $u$, using a Negative Binomial distribution:
-$$y^{F} ~ Negative Binom(\hat{F},\hat{\phi})$$
-$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x + u + w$$
+\paragraph{Flowers}
+The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Negative Binomial distribution:
+$$y^{F} \sim 0 Truncated Negative Binom(\hat{F},\hat{\phi})$$
+$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u + w$$
 $$log(\hat{\phi}) = \beta_{0}^{\phi}$$
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 
-The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted) $y^V$ in year $t+1$, with fixed effects of the ant partner of the cactus $i$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
-$$y^{V} ~ Binom(\hat{V})$$
+\paragraph{Viability}
+The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates the proportion of flowers produced by a plant which are viable (not aborted) $y^V$ in year $t+1$, with fixed effects of the ant partner of the cactus $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
+$$y^{V} \sim Binom(\hat{V})$$
 $$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion.
 
-The ant transition rates model ($\tau_(x,a,a')$) estimates the probability of a cactus being visited by an ant partner $a'$ $y^{\tau}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
-$$y^{\tau} ~ Multinomial(\hat{\tau})$$
-$$logit(\tau) = \beta_{0}^{\tau} + \beta_{1}^{\tau} \times x + \beta_{2}^{\tau} \times a + \beta_{3}^{\tau} \times x \times a + u + w$$
-where $u ~ N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ in year $t$, and $w ~ N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+\paragraph{Ant Transitions}
+The ant transition rates model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$ $y^{\epsilon}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
+$$y^{\epsilon} ~ Multinomial(\hat{\epsilon})$$
+$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
+where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$.
 Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
 
+\paragraph{Recruit Size Distribution}
 The recruit size model ($n(x,a')$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
-$$y^{\eta} ~ N(\hat{\eta},\hat{\sigma})$$
-$$\hat{eta} = \beta_{0}^{\eta}$$
+$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
+$$\hat{\eta} = \beta_{0}^{\eta}$$
 where $\hat{\sigma}$ is estimated with a non-informative prior.
 
-With germination data from Miller et al., 2009****, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$) in year $t+1$, with no fixed or random effects, using a Binomial distribution:
-$$y^{\gamma_1} ~ Binomial(\hat{\gamma_1})$$
-$$y^{\gamma_2} ~ Binomial(\hat{\gamma_2})$$
+\paragraph{Germination}
+With germination data from \tom{Miller et al., 2007}{CITE}, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($\gamma_2$) in year $t+1$, with no fixed or random effects, using a Binomial distribution:
+$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
 $$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
@@
 
-With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
+\paragraph{Pre-Census Survival}
+With data collected in a \tom{2005-2006}{FIX}********** recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
 $$y^{\delta} ~ Bern(\hat{\delta})$$
 $$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
- where $m ~ N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+ where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
 
-With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) in year $t+1$ based on the ant partner $i$ in year $t+1$, using a Negative Binomial distribution:
-$$y^{\kappa} ~ Negative Binomial(\hat{L},\hat{\phi})$$
-$$\hat{L} = \beta_{0}^{L} \times j$$
+\paragraph{Seeds Per Flower}
+With data from \tom{Miller 2007}{CITE}********, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$, using a Negative Binomial distribution:
+$$y^{\kappa} \sim 0 Truncated Negative Binomial(\hat{\kappa},\hat{\phi})$$
+$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
 $$\hat{\phi} = \beta_{0}^{\phi}$$
@@
 
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R %\tom{}{You need to cite R and RStan}
+\paragraph{MCMC Simulations}
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite} 
 For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
@@
 We extend the integration limits $L$ and $U$ to avoid unintentional eviction \cite{Williams2012}.
-We estimate the asymptotic population growth rate $\lambda$ as the dominant eigenvalue of the discrete kernel for each combination of ant partners: complete vacancy; \textit{C.
+Traditionally, in a deterministic IPM the asympotic population growth rate $\lambda$ is estimated as the dominant eigenvalue of the discrete kernel.
+In a deterministic bayesian IPM, we create 1,000 discrete kernels with a unique set of parameters from our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution. 
+In a deterministic version of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
 opuntiae} and vacancy; other and vacancy; \textit{L.
@@
 opuntiae}, other, and vacancy; and all ant partners and vacancy.
-Rather than estimating a single $\lambda$ value for each possible population, we used 1,000 random parameter estimations from iterations of our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution.
+In order to calculate stochastic $\lambda$ distributions for each ant scenario, every one of the 1,000 parameter-iteration associated $\lambda$s actually comes from the mean of 5,000 year-specific estimations each associated with a year random effech which was randomly selected from the 18 years of data.
+
 We compare the distributions of $\lambda$ across each combination of ant partners to whether sampling effect or complementarity is at play in the system.
@@
 
-\subsection*{Stochastic Integral Projection Model Construction and Analysis}
-
-The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
-This stochastic IPM draws from the same statistical models as the deterministic version, and is described below:
-
- \begin{linenomath*}
-	$$
-	B_1(t+1) = \delta \sum_{a}^{A} \int_L^U  \kappa(a) P(x',u,w) V(a,u,w) F(x,u,w) n(x,a,t) dx \\
-	$$
-	$$
-	B_2(t+1) =  (1 - \gamma_1)B_1(t)\\
-	$$
-	$$
-	n(x',a',t+1) = (\gamma_1 B_1(t) + \gamma_2 B_2(t)) \eta(x') \omega \rho_{a'}  + \\
-	$$
-	$$
-	\sum_{a}^{A} \int_L^U S(x,a,u,w) G(x',x,a,u,w) \tau(x,a,a',u,w) n(x,a,t) dx \\
-	$$
-\end{linenomath*}
-
-This system of equations still describes the contribution of seedbanks and existing plants to the group of plants of size $x'$ and ant partner $a'$ in year $t+1$, but now includes the random effects of plot $u$ and year $w$.
-These effects are included in the form of linear scales on each statistical model where they were biologically relevant.
-Year effects $w$ are both year and ant speciifc, meaning for each year there are different ant and year-specific linear scales for each year of data. 
-When calculating stochastic $\lambda$, every one of the 1,000 parameter-iteration associated $\lambda$s actually come from the mean of 1,000 year-specific estimations. 
-For each parameter estimate taken from the bayesian models we calculate 1,000 estimations of $\lambda$, each associated with a year random effect which was randomly selected from the *******18 years of data.
-The mean of these estimations become the parameter-iteraction specific $\lambda$ estimation. 
-This is repeated 1,000 times for each iteration of parameters to get the stochastic posterior distribution of $\lambda$.
-We compare the stochastic distributions to the deterministic ones to determine if portfolio effect is at play in this system.
-
-To determine if portfolio effect is at play we must show that partner diversity buffers the population from environmental variation (in our case annual variation assessed by the inclusion of year random effects).
+To determine if portfolio effect is at play we must show that partner diversity buffers the population from environmental variation.
+To do this we parameterized a new set of bayesian statistical models in which year random effects were not ant specific, meaning the effects of ant partners were not able to covary with year effects (the scenario in which all ant partners must respond to annual variation the same way and portfolio effect is not possible).
+The new estimations of year random effects are $u \sim N(0,\sigma_{yr}^2)$ with year specific variance of $\sigma_{yr}$. 
+With this verision of the IPM (now referred to as the stochastic null IPM) we followed the same approach to calculate distributions of stochastif $\lambda$s for each ant scenario.
 We can estimate the effects of partner diversity on the focal mutualist by calculating the difference in $\lambda$ distributions between populations with no partners and populations with all partners present.
-Then we compare the difference in the effects of partner diversity on the focal mutualist with and without stochasticity.
-If the stochastic difference in the effects of partner diversity is greater than the deterministic difference, this indicates that partner diversity is more beneficial under varying environments and therefore portfolio effect is at play.
+Then we can compare the difference in the effects of partner diversity on the focal mutualist calculated from the stochastic and stochastic null IPMs respectively.
+If the stochastic difference is greater than the stochastic null difference, this indicates that  partner diversity is more beneficial under varying environments and therefore portfolio effect is at play.
 
 \section*{Results}
+\subsection*{IPM Analysis}
+%%%% Start with stochastic model and comparing means to vacancy
+% means
+Using the results of the stochastic IPM, we determined if complementarity or sampling effect was at play within this system. 
+We found that populations of tree cholla which have no partners have a mean fitness of $\lambda = 0.9861$ \ref{}.
+Populations with only one ant partner present (\textit{L.
+apiculatum}, \textit{C.
+opuntiae}, or other ants) have a mean fitness of $\lambda = 0.9944$, $\lambda = 0.9893$, and $\lambda =0.9896$ respectively.
+Populations with two ant partners present (\textit{L.apiculatum} and \textit{C.
+opuntiae}, \textit{L.apiculatum} and other, or \textit{C.
+opuntiae} and other) have a mean fitness of $\lambda = 0.9942$, $\lambda = 0.9943$, and $\lambda = 0.9913$ respectively.
+Populations with all ant partners present have a mean fitness of $\lambda = 0.9940$.
+The only partner scenarios where $\lambda > 0.99$ are when \textit{L. apiculatum} ants are present with any other ant partner. 
+The means are not the only important result.
+We used bayesian modeling, so each of these $\lambda$ estimations has a distribution rather than a single estimate.
+By subtracting one of these distributions from the other, the difference from 0 tells us how much larger or smaller the distributions are with a percent confidence. 
+%% All to vacant
+% single partners confidence
+%We are 82\%, 100\%, and 84\% confident that having \textit{C.
+opuntiae} ants, other ants, or \textit{L.
+apiculatum} ants respectively as the only partner leads to a higher $\lambda$ than having no partners.
+%\textit{L. apiculatum} ants are the best single partner for tree cholla populations, as they lead to the highest population fitness with a single partner.
+% two partners confidence
+%We are 96\%, 98\%, and 92\% confident that having two ant partners present (\textit{L.apiculatum} and \textit{C.
+opuntiae}, \textit{L.apiculatum} and other, or \textit{C.
+opuntiae} and other) respecitvely leads to a higher $\lambda$ than having no ant partners.
+% all partners confidence
+%We are 94\% confident that having all ant partners present leads to a higher $\lambda$ than having no ant partners present. 
+Any ant partner is shown to lead to an increase in fitness of the tree cholla in these results. 
+We are between 82\% to 100\% confident that having any ant parnter leads to a higher $\lambda$ than having no partners.
+
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/LambdaMeans.png}
+\end{figure}
+
+%% Liom to all
+%We are 100\%, 88\%, 90\%, and 76\% certain that having only \textit{L.
+apiculatum} ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
+opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
+opuntiae} and other ants as partners respecitvely.
+%We are 96\%, 92\%, 92\%, and 84\% confident that having \textit{L.
+apiculatum} and \textit{C.
+opuntiae} ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
+opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
+opuntiae} and other ants as partners respecitvely.
+%We are 98\%, 92\%, 96\%, and 92\% confident that having \textit{L.
+apiculatum} and other ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
+opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
+opuntiae} and other ants as partners respecitvely.
+%We are 94\%, 84\%, 94\%, and 84\% confident that having only \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and other ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
+opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
+opuntiae} and other ants as partners respecitvely.
+We are between 84\% and 100\% confident that any diversity scenario where \textit{L.
+apiculatum} ants are included as partners leads to a higher $\lambda$ for the tree cholla than any diversity scenario without these parnters.
+These results indicate that sampling effect, not complementarity, is at play in this system.
+
+%%%% Stochastic null follows similar patterns but dampened
+Using the stochastic null IPM results, we found that when ant partners responded the same way to inter-annual variation similar patterns were found, though the magnitude of the patterns were somewhat different.
+We used the results of the stochastic null IPM to determine if portfolio effect is at play. 
+We found that both the stochastic model and stochastic null model resulted in higher fitness when all ants are present compared to no partners 94\% of the time. 
+These results differed in the magnitude of fitness boost recieved from partners.
+In the stochastic IPM, tree cholla with all ants present resulted in a $\lambda$ that was 0.008 greater than tree cholla with no ants present; in the stochsatic null IPM, tree cholla with all ants rpesent resulted in a $\lambda$ that was 0.013 greater than tree cholla with no ants present.
+We are 52\% confidence that the difference is greater when inter-annual effects were not ant dependent (in the stochastic null model).
+This indicates that portfolio effect is at play, if only weakly evident.
+
 \subsection*{Statistical Modeling}
 
-%% Growth Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{grow_contour_lines_color.png}
-%	\caption{asdasdf}
-%	\label{fig:growth}
-%\end{figure}
-Plants tended by other ants experiened positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}a.
-They experienced significantly higher growth than vacant plants ($p = 2.2 \times 10^{-16}$), and similar mean growth to \textit{L.
-apiculatum} tended plants ($p = 0.9135$) and \textit{C.
-opuntiae} tended plants ($p = 0.4864$).
-Plants tended by \textit{C. opuntiae} ants experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}b.
-They experienced significantly higher mean growth than vacant plants ($p = 4.88 \times 10^{-14}$).
-Plants tended by \textit{L. apiculatum} experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}c.
-Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d.
+\paragraph{Growth Model}
+% Growth Model
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/GrowContourLinesColor.png}
+	\caption{asdasdf}
+	\label{fig:growth}
+\end{figure}
+Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes \ref{fig:growth}e.
+The mean growth rate expected for plants that are 10 $cm^3$ or smaller in year t is 10.01 $cm^3/yr$, 4.65 $cm^3/yr$, 7.44 $cm^3/yr$, and 5.14 $cm^3/yr$ when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+The mean growth rate expected for plants that are between 10 $cm^3$ and 150 $cm^3$ in year t is 2.81 $cm^3/yr$, 2.11 $cm^3/yr$, 2.49 $cm^3/yr$, and 1.75 $cm^3/yr$ when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+The mean growth rate expected for plants that are larger than 150 $cm^3$ in year t is 1.28 $cm^3/yr$, 1.21 $cm^3/yr$, 1.29 $cm^3/yr$, and 1.10 $cm^3/yr$ when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+Plants with \textit{C.
+opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants \ref{fig:growth}a.
+Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes \ref{fig:growth}b. 
+Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates\ref{fig:growth}c.
+Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates \ref{fig:growth}d.
 
-%% Survival Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{surv_panels.png}
-%	\caption{asdasdf}
-%	\label{fig:surv}
-%\end{figure}
-Plants tended by \textit{C.
-opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
-apiculatum} tended plants ($p = 2.908 \times 10^{-14}$), plants tended by other ants ($p = 4.488 \times 10^{-14}$), and vacant plants ($p = 2.2 \times 10^{-16}$) \cite{fig:surv}.
-These \textit{C. opuntiae} plants have the highest mean survival at small sizes before reaching a mean survival rate of about 100\% when medium to large.
-Plants tended by \textit{L. apiculatum} experienced similar mean survival rates to plants tended by other ants and vacant plants.
-\textit{L. apiculatum} tended plants and other tended plants reached a mean survival rate of nearly 100\% by medium sizes, with other tended plants reaching this rate slower. 
-Vacant plants never reach mean survival rates of 100\%.
+We are 89\%, 88\%, and 70\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean growth rates across sizes than plants tended by no parnters, \textit{L.
+spiculatum} ants, or other ants respectively.
+We are 89\%, 65\%, and 94\% confident that plants with no partners experience lower mean growth rates across all sizes than plants tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants respectively.
+
+\paragraph{Survival Model}
+% Survival Model
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/SurvivalPlot.png}
+	\caption{asdasdf}
+	\label{fig:surv}
+\end{figure}
+Tree cholla experience between 7.7\% and 99.9\% survival rate depending on their size and ant partner \ref{fig:surv}e.
+Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
+The mean survival rate expected for plants that are 10 $cm^3$ or smaller in year t is 48\%, 37\%, 37\%, and 47\% when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+The mean survival rate expected for plants that are between 10 $cm^3$ and 150 $cm^3$ in year t is 90\%, 87\%, 84\%, and 80\% when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+The mean survival rate expected for plants that are larger than 150 $cm^3$ in year t is 99\%, 99\%, 98\%, and 95\% when tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, other ants, or no ants respectively.
+Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes \ref{fig:surv}a.
+Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes \ref{fig:growth}b. 
+Plants with other ants experience the second lowest mean survival rates across all sizes \ref{fig:growth}c.
+Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates \ref{fig:growth}d.
+
+We are 82\%, 63\%, and 100\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean survival rates across all sizes than plants tended by no parnters, \textit{L.
+spiculatum} ants, or other ants respectively.
+We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants respectively.
 
 %% Flowers Model
+\paragraph{Flowering Model}
 There is a clear size effect on the number of flowers produced. 
-The mean number of flowers produced by a plant remains at 0 until the plant reachesmedium sizes after which the mean number of flower produced increases eponentially to about 40 at large sizes.
+The mean number of flowers produced by a plant remains at 0 until the plant reaches medium sizes after which the mean number of flower produced increases eponentially to about 40 flowers per plant per year at large sizes.
+The mean number of flowers produced for a plant that is 150 $cm^3$ or smaller in year $t$ is <1 flower per plant.
+The mean number of flowers produced for a plant that is larger than 150 $cm^3$ in year $t$ is 8.7 flowers per plant.
 
-%% Viability Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{Viability.png}
-%	\caption{asdasdf}
-%	\label{fig:comp-samp}
-%\end{figure}
-Plants tended by \textit{L.
-apiculatum} ants experienced the highest viability rates, significantly higher than \textit{C.
-opuntiae} tended plants ($p = 2.2 \times 10^{-16}$), other tended plants ($p = 2.2 \times 10^{-16}$), and vacant plants ($p = 2.2 \times 10^{-16}$).
-The mean viability rate of \textit{L. apiculatum} tended plants was 84.888\%.
-Plants tended by \textit{C. opuntiae} ants experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$), but not than other tended plants ($p = 0.1168$).
-The mean viability rate of \textit{C. opuntiae} plants was 75.934\%.
-Plants tended by other ants also experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$).
-The mean viability rate of other tended plants was 76.367\%.
-The mean viability rate of vacant plants was 70.442\%.
+% Viability Model
+\paragraph{Viability Model}
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/ViabHist.png}
+	\caption{asdasdf}
+	\label{fig:viab}
+\end{figure}
+Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers \ref{fig:viab}.
+The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
+apiculatum} tended plants experiencing the highest mean viability rate (at 86\%), followed by other tended plants (at 75\%), \textit{C.
+opuntiae} tended plants (at 74\%) and vacant plants (at 71\%).
+We are 99\%, 98\%, and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by no partners, \textit{C. opuntiae}, or other ants respectively.
+We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
 
 %% Reproduction Model
+\paragraph{Reproduction Model}
 The probability of a plant reproducing in a given year is highly size dependent. 
@@
 %% Seeds Produced
-The mean number of seeds produced by \textit{C.
-opuntiae} tended plants is significantly lower than the mean number of seeds produced by \textit{L.
-apiculatum} plants ($p = 2.2 \times 10^{-16}$) and vacant plants ($p = 2.2 \times 10^{-16}$).
-\textit{C. opuntiae} tended plants produce a mean of 2.561 seeds per flower. 
-The mean number of seeds produced by \textit{L. apiculatum} tended plants is slightly higher than the mean number produced by vacant plants ($p = 0.02521$). 
-\textit{L. apiculatum} tended plants produce a mean of 3.563 seeds per flower. 
-Vacant plants produce a mean of 3.435 seeds per flower.
+\paragraph{Seeds Per Flower Model}
+Each viable flower on a plant produces between 97 and 257 seeds.
+This number is affected by the ant partner present. 
+\textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
+\textit{L. apiculatum} tended plants produce a mean of 149 seeds per flower. 
+Vacant plants produce a mean of 160 seeds per flower. 
+We are 73\% and 78\% confident that vacant plants produce more seeds per flower on average than plants tended by \textit{C. opuntiae} and \textit{L. apiculatum} ants respectively.
 
-%% Precensus Survival
-The mean pre-census seed survival rate is 16.99\% with 90\% of model iterations landing between 1.13\% and 40.72\%.
+%% Precensus Survival\
+\paragraph{Precensus Survival Model}
+Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.
 
 %% Germination
-Seeds have a significantly higher probability of germinating in year one than in year two ($p = 2.2 \times 10^{-16}$).
-Seeds in year one have a mean germination rate of 0.578\% and seeds in year two have a mean germination rate of 0.221\%.
+\paragraph{Germination Model}
+Seeds have a significantly higher probability of germinating in year one than in year two.
+Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
+Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.
 
 %% Recruit size distribution
-The mean size of new recruits is -1.88 log($cm^3$) with 90\% of new recruits falling between the sizes of -2.12 log($cm^3$) and -1.64 log($cm^3$).
+New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.
 
 %% Ant Transition Rates
-% previously vacant
-Plants that were previously vacant will most likely remain vacant until large. 
-These plants have significantly higher probabilities of being vacant rather than becoming tended  by \textit{C.
+\paragraph{Ant Transition Model}
+\begin{figure}
+	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+\end{figure}
+All small plants are most likely to be vacant, while large plants are much more likely to be occuplied by one ant or another. 
+\textit{L.
+apiculatum} ants are the most likely tender in year $t+1$ for all large plants that were not previously tended by \textit{C.
+opuntiae} ants in year $t$, while \textit{C.
+opuntiae ants} tended plants in year $t$ are most likely to remain tended by \textit{C.
+opuntiae} ants in year $t+1$.
+We are 93\%, 89\%, 86\%, and 93\% confident that plants that are smaller than 150 $cm^3$ in year $t$ will be vacant in year $t+1$ when they were tended by no partners, \textit{C.
 opuntiae}, \textit{L.
-apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
-The probability of previously vacant plants becoming tended by \textit{L.
-apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
-opuntiae} or other ants (respecitvely, $p = 3.83 \times 10^{-14}$, $p = 2.2 \times 10^{-16}$).
-The probability of previously vacant plants becoming tended by other ants is significantly lower than the probability of becoming tended by \textit{C. opuntiae} ($p = 0.00227$).
+apiculatum}, or other ants respectively in year $t$.
 
-% previously crem
-Plants that were previously tended by \textit{C. opuntiae} will most likely remain vacant until large. 
-These plants have significantly higher probabilities of becoming vacant rather than remaining tended by \textit{C.
-opuntiae} or becoming tended by \textit{L.
-apiculatum} or other ants (repsectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$ ).
-The probability of these plants remaining tended by \textit{C.
-opuntiae} ants is significantly higher than the probability of becoming tended by \textit{L.
-apiculatum} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 3.797 \times 10^{-9}$).
-The probability of these plants becoming tended by \textit{L. apiculatum} is significantly higher than the probability of becoming tended by other ants ($p = 3.797 \times 10^{-9}$).
+% Liom 
+We are 75\%, 100\%, and 100\% confident that plants which were previously tended by \textit{L.
+apiculatum} and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
+apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
+opuntiae}, or other ants in year $t+1$.
+We are 35\%, 100\%, and 100\% confident that plants which were previously vacant and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
+apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
+opuntiae}, or other ants in year $t+1$.
+We are 32\%, 100\%, and 100\% confident that plants which were previously tended by other ants and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
+apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
+opuntiae}, or other ants in year $t+1$.
 
-% previously liom
-Plants that were previously tended by \textit{L. apiculatum} ants will most likely remain vacant until large. 
-These plants have significantly higher probabilities of becoming vacant than remaining tended by \textit{L.
-apiculatum} or becoming tended by \textit{C.
-opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
-These plants have significantly higher probabilities of remaining tended by \textit{L.
-apiculatum} ants than becoming tended by  \textit{C.
-opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
-The probability of these plants becoming tended by \textit{C. opuntiae} ants is significantly higher than the probability of them becoming tended by other ants ($p = 0.007925$).
+% Crem
+We are 37\%, 100\%, and 100\% confident that plants which were previously tended by \textit{C.
+opuntiae} and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{C.
+opuntiae} ants in year $t+1$ than be vacant, tended by \textit{L.
+apiculatum}, or other ants in year $t+1$.
 
-% previously other
-Plants that were previously tended by other ants will most likely remain vacant until they are large. 
-These plants have significantly higher probabilities of becoming vacant rather than remaining tended by other ants or becoming tended by \textit{C.
-opuntiae} or \textit{L.
-apiculatum} ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
-The probability of these plants becoming tended by \textit{L.
-apiculatum} ants is significantly higher than the probability of remaining tended by other ants or becomnig tended by \textit{C.
-opuntiae} (respectively, $p = 6.309 \times 10^{-12}$, $p = 6.256 \times 10^{-13}$).
-The probability of these plants remaining tended by other ants is not higher than the probability of becoming tended by \textit{C. opuntiae} ants ($p = 0.954$).
+\ali{}{Or I can do this section like this:}
+% Prev Vacant
+Plants which were previously vacant will most likely remain vacant until large. 
+We expect that between 81\% and 98\% of tree cholla which are smaller than 150 $cm^3$ with no partners in year $t$ are going to be vacant in year $t+1$.
+We expect that between 11\% and 55\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
+We expect that between 6\% and 17\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
+We expect that between 2\% and 10\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by other ants in year $t+1$.
 
-%% All cases/trends
-In all plants ,the most likely scenario is vacancy.
-In most cases, the second most likely next ant partner is \textit{L.
-apiculatum} with the exception of plants which were previously tended by \textit{C.
-opuntiae} who are second most likely to retain \textit{C.
-opuntiae} as their partners in successive years.
-The only case where two next partners do not have statistically significantly different probabilities is when previously other tended plants either remain tended by other ants or become tended by \textit{C.
-opuntiae}.
+% Prev Crem
+Plants which were previously tended by \textit{C. opuntiae} will most likely be vacant or tended by \textit{C. opuntiae} ants in the next year depending on size. 
+We expect that between 74\% and 97\% of tree cholla which are smaller than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be vacant in year $t+1$.
+We expect that between 8\% and 34\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
+We expect that between 15\% and 38\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
+We expect that between 3\% and 13\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by other ants in year $t+1$.
 
-\subsection*{Deterministic IPM}
+% Prev Liom
+Plants which were previously tended by \textit{L. apiculatum} will most likely be vacant or tended by \textit{L. apiculatum} in the next year depending on size. 
+We expect that between 64\% and 97\% of tree cholla which are smaller than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be vacant in year $t+1$.
+We expect that between 30\% and 82\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L.
+apiculatum} in year $t$ are going to be tended by \textit{L.
+apiculatum} ants in year $t+1$.
+We expect that between 4\% and 7\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
+We expect that between 1\% and 3\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be tended by other ants in year $t+1$.
 
-\subsection*{Stochastic IPM}
+% Prev Other
+Plants which were previously tended by other ants will most likely be vacant or tended by other ants in the next year depending on size. 
+We expect that between 82\% and 99\% of tree cholla which are smaller than 150 $cm^3$ and tended byother ants in year $t$ are going to be vacant in year $t+1$.
+We expect that between 10\% and 53\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
+We expect that between 4\% and 13\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
+We expect that between 3\% and 15\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by other ants in year $t+1$.
 
 \section*{Discussion}
+The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
+Many studies have looked at multispecies mutualisms and the how having a variety of partners leads to variation in demographic effects \cite{Palmer2010, Bascompte2019, Stachowicz2005, Ford2015, Baez2016}.
+Because these tree cholla interact with only one ant partner at a time, it is a unique system in which to parse out the individual effects of each ant partners, both in isolated settings or in combinations we cannot test in the real world.
+
+%\subsection*{Demographic effects of association with alternative partners}
+We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, and more) have on the vital rates of tree cholla. 
+Using a system of heirarchical bayesian models we found that there were discernable differences in the effects that each partner had on vital processes of the focal mutualist. 
+The different vital rates vary in importance across tree cholla ontogeny.
+Several of them are negatively impacted by the presence and pressure of herbivores and seed predators \cite{Miller2009, Miller2006} and positively impacted by the presence of ant partners \cite{Miller2007}.
+The predators and herbivores target new growth and flowers, leading to negative impacts on the growth rates, survival rates, and floral viability rates of tree cholla \cite{Louda1995, Agrawal2004}.
+The presence of the ant partners can reduce those negative effects.
+
+It has been previously hypothesized that their is a heirarchy of partners due to the ability for \textit{L. apiculatum} ants to defend the cacti from seed predators and herbivores \cite{Miller2007}.
+These prior results would suggest that all vital rates that are affected by ant partners would be boosted the most by the presence of \textit{L. apiculatum} ants, this is not what we found. 
+Our results suggest that different partners differ significantly in their effects on vital rates.
+
+Prior to reproduction, the tree cholla experience only growth and survival.
+\textit{C. opuntiae} tended ants are associated with the highest growth rates and survival rates of plants. 
+This indicates that \textit{C. opuntiae} ants may be good ants for pre-reproductive tree cholla plants. 
+Reproducing plants experience a probability of reproducing, flower production, and floral abortion.
+Floral abortion is heavily affected by seed predators \cite{Miller2008}, which the ants defend the cacti, leading to increased floral viability.
+We specifically found that tree cholla experienced the highest floral viability rates when tended by \textit{L. apiculatum} ants. 
+This indicates that \textit{L. apiculatum} ants may be good partners for reproducing plants.
+
+These results together would suggest that complementarity may be the underlying mechanism that explains why partner diversity is beneficial in this system.
+
+%\subsection*{Frequency and direction of partner turnover}
+We have shown that the identity of partners is important to the processes that define tree cholla fitness. 
+Now we need to analyze the dynamics of partner turnover which dictate the identity of tree cholla partners and therefore the effects of vital rates on the tree cholla. 
+With our models we were able to identify both the direction, frequency, and distinct patterns of partner turnover.
+
+In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
+In some systems high freqency of turnover is necessary to resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
+While the purpose of this paper is not to establish which would be most beneficial in this system, we were able to identify the pattern. 
+Small plants are almost entirely vacant in this system until they grow large enough to begin producing significant amounts of EFN.
+Our model shows that once they do produce EFN, plants experience a relatively significant amount of turnover.
+Mid-sized and large plants which were either vacant or tended by other ants are most likely to become tended by \textit{L. apiculatum} ants in the next year, thereby experiencing partner turnover. 
+Plants which were tended by \textit{L. apiculatum} or \textit{C. opuntiae} ants are most likely to remain tended by the same partners multiple years in a row. 
+This indicates that \textit{C. opuntiae} ants and \textit{L. apiculatum} ants are loyal partners which retain the same plants year after year with regularity.
+
+As established in previous studies, the direction of partner turnover is important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
+In our study we found that there are distinct patterns to the direction of partner turnover. 
+Specifically, most plants are tended by \textit{L.
+apiculatum} ants at some point, because vacant plants, other tended plants, and \textit{L.
+apiculatum} tended plants are all most likely to be tended by \textit{L.
+apiculatum} ants.
+This indicates that while \textit{L. apiculatum} ants are loyal to their own plants and return multiple years in a row to the same ones, they are also strong colonizers.
+
+%\subsection*{Effect of partner diversity on plant fitness and what mechanisms explain this}
+The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (as some have done before \cite{Palmer2010}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
+We found that the combination of accurate partner transitions with partner identity affected the fitness of the tree cholla in interesting and dynamic ways. 
+Namely, a best partner emerged in this analysis, which was surprising given the nature of our vital rate findings. 
+The variation in best partner for each vital rate suggested the potential that the different ant partners had some level of unique specialty in what they offered, which would support complementarity as the mechanism which explained the effects of partner diveristy \cite{Stachowicz2005, Stanton2003}.
+The results of our IPM however differ from this prediction.
+
+Using the stochastic IPM we developed, we found evidence of sampling effect rather than complementarity.
+We found that \textit{L. apiculatum} was the single best partner, and that all diversity scenarios where \textit{L. apiculatum} was present resulted in the highest possible fitness of tree cholla. 
+This indicates that despite the fact that \textit{L. apiculatum} partnership does not result in the highest growth and survival rates, it is still the overall best partner.
+
+Using the stochastic null IPM and the stochastic IPM we compared the fitness boost recieved by all ant partners when ants effects varied separately across years and when they did not. 
+When all ants responded to inter-annual variability the same way (shown in the stochastic null IPM) we found that the fitness boost recieved from partners was larger than the fitness boost recieved when ants responded to inter-annual variability differently.
+This indicates that having multiple possible partners benefits the tree cholla by buffering the potentially negative effects of inter-annual variation.
+
+%\subsection*{Synthesis????? All together}
+\ali{}{What this all means more broadly?? I'm currently not sure what to do with this}
+This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
+Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
+The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
+
+%\subsection*{Future Directions}
+This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
+We revealed the dynamics of partner turnover and showed that different ant partners are correlated with different fitness benefits.
+As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work.
 

### 2024-01-16 03_cholla_ant_IPM_params_functions.R

@@ -27,7 +27,7 @@ years <- unique(cactus$Year_t)
 ## Choose your pathway to pull from 
 #Ali
 mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
-mcmc_dir <- "/Users/alicampbell/Desktop/"
+#mcmc_dir <- "/Users/alicampbell/Desktop/"
 #Tom
 #mcmc_dir <- "C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 #Lab

### 2024-01-16 04_cholla_ant_IPM_analysis.R

@@ -112,9 +112,9 @@ write.csv(lams_dpost,"det_post_lambda_mean.csv")
 scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
 max_scenario = length(scenario)
 # Choose the number of parameter iterations 
-max_rep = 1000 ## Posterior Draws from vital rate models
+max_rep = 50 ## Posterior Draws from vital rate models
 # Choose the number of years for stochasticity
-max_yrs = 1000 ## Years of randomly sampled annual effects
+max_yrs = 200 ## Years of randomly sampled annual effects
 # Create an empty matrix to fill with the lambda estimations using functions defined in IPM_Stochastic_Post.R
 # Rows correspond to parameter iterations
 # Columns correspond to partner scenarios
@@ -122,7 +122,7 @@ lams_stoch <- matrix(nrow = max_rep, ncol = max_scenario)
 for(n in 1:max_scenario){
   print(scenario[n])
   for(m in 1:max_rep){
-  lams_stoch[m,n] <- lambdaSim(params=mean_params,
+  lams_stoch[m,n] <- lambdaSim(params=params[m,],
                                lower=lower,
                                upper=upper,
                                scenario = scenario[n],
@@ -162,17 +162,17 @@ write.csv(lams_stoch,"stoch_post_lambda.csv")
 scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
 max_scenario = length(scenario)
 # Choose the number of parameter iterations 
-max_rep = 1 ## Posterior Draws from vital rate models
+max_rep = 50 ## Posterior Draws from vital rate models
 # Choose the number of years for stochasticity
-max_yrs = 1000 ## Years of randomly sampled annual effects
+max_yrs = 200 ## Years of randomly sampled annual effects
 # Create an empty matrix to fill with the lambda estimations using functions defined in IPM_Stochastic_Post.R
 # Rows correspond to parameter iterations
 # Columns correspond to partner scenarios
-lams_stoch_null_mean <- matrix(nrow = 1, ncol = max_scenario)
+lams_stoch_null <- matrix(nrow = max_rep, ncol = max_scenario)
 for(n in 1:max_scenario){
   print(scenario[n])
   for(m in 1:max_rep){
-    lams_stoch_null[m,n] <- lambdaSim(params=mean_params,
+    lams_stoch_null[m,n] <- lambdaSim(params=params[m,],
                                       lower=lower,
[…]

### 2024-01-16 05_cholla_ant_IPM_figures_tables.R

@@ -5,7 +5,7 @@
 ##################################################################################################################
 ##################################################################################################################
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
-source( "/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Cholla_Analysis_Vital_Rates.R")
+source( "/Users/alicampbell/Documents/GitHub/ant_cactus_demography/04_cholla_ant_IPM_analysis.R")
 
 ## Color Codes
 ## Retro bright
@@ -13,15 +13,15 @@ cremcol <- "#9239F6"
 liomcol <- "#00A08A"
 othercol <- "#FF0076"
 vaccol <- "#F8B660"
-  vcol <- "#ad90ec"
-    lcol <- "#084f98"
-      ccol <- "#e9a67a"
-        ocol <- "#93022f"
-          lccol <- "#5dc9cf"
-            locol <- "#cf3545"
-              cocol <- "#ab59c8"
-                acol <- "#5d906b"
-                  cols <- c(vcol, ccol, lcol, ocol, lccol, locol, cocol, acol)
+vcol <- "#ad90ec"
+lcol <- "#084f98"
+ccol <- "#e9a67a"
+ocol <- "#93022f"
+lccol <- "#5dc9cf"
+locol <- "#cf3545"
+cocol <- "#ab59c8"
+acol <- "#5d906b"
+cols <- c(vcol, ccol, lcol, ocol, lccol, locol, cocol, acol)
 str(cactus)
 ##### Size variable used in most visualizations
 size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm = TRUE), by = 0.1)
@@ -1315,49 +1315,13 @@ setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
 scenario_abv <- c("V","CV","LV","OV","LCV","LOV","OCV","LOCV")
 ## Plot the means of the deterministic and stochastic distributions together
 png("lambda_means.png")
-plot(c(1,3,5,7,9,11,13,15),(lams_dpost), pch = 20, cex = 5,col = cols,
-     xlim = c(0,16), ylim = c(0.98,1.01),
+plot(c(1,3,5,7,9,11,13,15),colMeans(lams_stoch), pch = 20, cex = 5,col = cols,
+     xlim = c(0,16), ylim = c(0.97,1.007),
      xaxt = "n",cex.lab = 2,
[…]

### 2024-01-16 Visuals Not simplified.R

@@ -1,4 +1,4 @@
-setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography/Figures")
+setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
 source( "/Users/alicampbell/Documents/GitHub/ant_cactus_demography/03_cholla_ant_IPM_params_functions.R")
 size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm = TRUE), by = 0.1)
 
@@ -156,6 +156,61 @@ mtext("Log(Volume Year t)",side=1,line=0,outer=TRUE,cex=2)
 mtext("Log(Volume Year t+1)",side=2,line=0,outer=TRUE,cex=2)
 dev.off()
 
+## Mean growth rate of plants < 0 
+small <- seq(min(cactus$logsize_t, na.rm = T), 2.3, length = 205)
+y_other_mean_grow <- quantile(grow.params$beta0[,3],0.5) + (small) * quantile(grow.params$beta1[,3],0.5) + (small)^2 * quantile(grow.params$beta2[,3],0.5)
+# Crem
+y_crem_mean_grow <- quantile(grow.params$beta0[,1],0.5) + (small) * quantile(grow.params$beta1[,1],0.5) + (small)^2 * quantile(grow.params$beta2[,1],0.5)
+# Liom
+y_liom_mean_grow <- quantile(grow.params$beta0[,2],0.5) + (small) * quantile(grow.params$beta1[,2],0.5) + (small)^2 * quantile(grow.params$beta2[,2],0.5)
+# Vac
+y_vac_mean_grow <-  quantile(grow.params$beta0[,4],0.5) + (small) * quantile(grow.params$beta1[,4],0.5) + (small)^2 * quantile(grow.params$beta2[,4],0.5)
+exp(mean(y_other_mean_grow - small))
+exp(mean(y_crem_mean_grow - small))
+exp(mean(y_liom_mean_grow - small))
+exp(mean(y_vac_mean_grow - small))
+## Mean growth rate of plants 0<x<10
+small <- seq(2.3, 5.01, length = 205)
+y_other_mean_grow <- quantile(grow.params$beta0[,3],0.5) + (small) * quantile(grow.params$beta1[,3],0.5) + (small)^2 * quantile(grow.params$beta2[,3],0.5)
+# Crem
[…]

### 2024-01-18 Manuscript/ant_cactus_AmNat.tex
manuscript updates

@@
 \usepackage{fullpage}
-%\usepackage[authoryear,sectionbib,sort]{natbib}
-\usepackage[square,sort,comma,numbers]{natbib}
+\usepackage[authoryear,sectionbib,sort]{natbib}
+%\usepackage[square,sort,comma,numbers]{natbib}
 \linespread{1.7}
@@
 
-\title{Thank you, next: partner turnover elevates benefits of mutualism for an ant-tended plant}
+\title{An Integral Projection Modeling Approach to Understanding Demographic Effects of Multispecies Mutualisms}
 
@@
 \section*{Abstract}
+Mutualisms are widespread species interactions with diverse and dynamic consequences. 
+They are considered more context dependent than other species interactions, meaning there are many different factors which change the outcomes of interactions between mutualists, including partner diversity.
+Partner diversity has become a central focus in the field of mutualisms, expanding previous work from primarily pairwise to multispecies mutualisms. 
+It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. 
+The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
+
+I use the plant-ant multi-species mutualism in which, the cactus \textit{Cylindropuntia imbracata} (tree cholla) pro- duce extrafloral nectar and the ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, \textit{Forelius pruinosus}, and rarer species, provide defense from various herbivores and seed predators.
+I used 18 years of data collected from plant demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, and ant partner for all plants in 8 30$\times$30 m plots at the Sevilleta National Wildlife Refuge in central New Mexico.
+With this data I parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
+I found that different ant partners had different impacts on the vital rates of the tree cholla. 
+Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
+With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest plant fitness, and which mechanism(s) may explain the effects of partner diversity.
+I found that the all scenarios which included the partner \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
+Results further suggest that diversity benefits in this system are driven by sampling effect , meaning \textit{L. apiculatum} ants are the "best`` partner.
+I also found that partner diversity benefits the focal mutualist in this system in the form of portfolio effect by buffering the tree cholla from the effects of inter-annual variation. 
+This study highlights how partner diversity can increase the overall benefits a focal mutualist receives. 
+It also highlights the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
 

### 2024-01-19 Manuscript/ant_cactus_AmNat.tex
thursday

@@
 \section*{Abstract}
+\tom{}{I think this is too long for Am Nat requirements. Also, use ``we''.}
 Mutualisms are widespread species interactions with diverse and dynamic consequences. 
@@
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
+Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of context-dependence in mutualism \citep{Stanton2003}.
 Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014}.
@@
 However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
-Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
+Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
 This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
 Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \cite{Afkhami2014}. 
-Third, partner species and herbivores can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+Third, partner species and \tom{herbivores}{notice that ``herbiviores'' does not fit here, at least not yet.
+you have not said anything about ants and plants, you are talking very generally.
+put yourself in the mind of a reader as you write, this will help ensure that ideas appear in an order that can be digested.} can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}.
 Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits and protection across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
@@
 Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
-Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
+Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
 Presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
@@
 Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
-Temporal dynamics of ant visitation therefore can have important impacts on the fitness of the plant partners \citep{Barton2010, Boege2005, Fonseca2003}.
-Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face varies just as much as the protection they recieve due to temporal stochasticity.
+Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
+Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
 Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
-However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data required to detect portfolio effects are rarely available.
+However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
 
@@
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the herbivore populations.
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits through anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the \tom{herbivore populations}{I would be careful here -- unless you are using the herbivore counts I don't think you have data on this.}.
 
-We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
+We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity at the individual level, and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
 We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
@@
 Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners. 
+Smaller cholla produce little to no EFN, so larger cholla, especially flowering individuals, are generally more highly tended \citep{Miller2014}.
 
 Tree cholla EFN is harvested by various ant species. 
-At SEV-LTER, cholla are visited primarily by two species of formicoid, ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster} and \textit{Camponotus}.
+At SEV-LTER, cholla are visited primarily by two species of ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster} and \textit{Camponotus}.
 \textit{L.
 apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $5\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
+opuntiae} visiting between $5\% - 20\%$ of cacti depending on the year \citep{Donald2022}.
 Between $ 30\% 80\%$ of cacti remain vacant in any given year. 
@@
 Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
-At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \citep{Ohm2014}. 
-Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla, especially flowering individuals, are generally more highly tended \citep{Miller2014}. 
 In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-There are a variety of insect herbivores and seed predators which attack the cholla \citep{Mann1969}. 
-An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
+There are a variety of insect herbivores and seed predators which specialize on tree cholla \citep{Mann1969}. 
+A weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
 A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
 A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \citep{Miller2009}.
+These predators can have significant negative impacts on plant fitness of and depress population growth \citep{Miller2009}.
 There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory than plants from which ants were excluded \citep{Miller2007}.
@@
 We recorded the ant species present (or vacancy if no ants present).
-Occurrences of more than one ant species on one plant were rare (QUANTITY), and in these cases we classified the plant as being occupied by the more abundant species. 
+Occurrences of more than one ant species on one plant were rare (\tom{}{quantify}), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species. 
 Plots were searched for new recruits each year, and these were added to the census.
-In total, the data set included \# unique individuals and \# plant-year observations. 
+\tom{In total, the data set included \# unique individuals and \# plant-year observations. }{update numbers}
 These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
@@
 apiculatum} and \textit{C.
-opuntiae} (QUANTIFY THIS) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
+opuntiae} (\tom{}{Quantify this.}) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
 opuntiae}, and ``other''.
-The ``other'' category was made of unidentified ant species and ant species which occurred at relatively low frequencies compared to \textit{C.
+\tom{The ``other'' category was made of unidentified ant species and ant species which occurred at relatively low frequencies compared to \textit{C.
 opuntiae} and \textit{L.
-apiculatum} (such as unknown species belonging to the genus \textit{Aphenogaster} and unknown species belonging to the genus \textit{Camponotus}).
+apiculatum} (such as unknown species belonging to the genus \textit{Aphenogaster} and unknown species belonging to the genus \textit{Camponotus}).}{I think you can do better than this because we actually have the designated species in the data.
+I would report here the actual breakdown (percentages) of species in the ``other'' group.
+I expect Forelious will be a big chunk of it.}
 Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
-Ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
+For example, ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
 Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
-In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independent of its influence on plant size. 
+In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independently of its influence on plant size. 
 Therefore, these sub-models included plant size but not ant state as predictor variables.
@@
 
-Functions $P(x';\pmb{\tau^P})$ and $F(x';\pmb{\tau^F})$ give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x'$ in year $t$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a')$) are dependent on ant state $a$ but not size. 
-The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) that modify the intercepts of the functions in which it appears. 
-This vector appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
+\noindent Functions \tom{$P(x';\pmb{\tau^P})$ and $F(x';\pmb{\tau^F})$}{These should be $x$ not $x'$ because we modeled flowering and fertility in year t based on size in year t.} give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x'$ in year $t$.
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit (\tom{$\kappa(a')$}{Why is this $a'$?
+I think it should be $a$.
+More generally, I think you need to explain your use of primes and how readers should interpret them at the start of model exposition.}) are dependent on ant state $a$ but not size.
+The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that \tom{deviates are specific to plants in ant state $a$}{Why does the tau vector for viability have the $a$ subscript but the others do not?}.
 Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A$) giving total seed production. 
@@
 	$$
-	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho(a')  + \\
+	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + \\
 	$$
 	$$
-	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \epsilon(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
+	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
 	$$
@@
 
-The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the proportion of seedlings which survive from germination (late summer) to the census (May) $\omega$.
-This term is multiplied by $\rho(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appears in our census ($\sum_{a'}^{A}\rho(a')=1$). 
+\noindent The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of seedling survival ($\omega$) from germination (late summer) to the census (May).
+This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appearance in our census ($\sum_{a'}^{A}\rho_{0}(a')=1$). 
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
-Survival ($S(x,a;\pmb{\tau^S_{a}})$) and growth ($G(x',x,a;\pmb{\tau^G_{a}})$) are both dependent on initial size and ant state. 
-As above, these functions accommodate inter-annual variability through year-specific intercepts that can vary by ant state ($\pmb{\tau_{a}}$). 
-Ant transition function $\epsilon(a',a,x;\pmb{\tau^{\epsilon}})$ gives the probability that a plant with size $x$ and ant state $a$ transitions to ant state $a'$ in the next census. 
-This function accomodates inter-annual variability through year-specific intercepts which are stable across ant states ($\pmb{\tau}$).
+Survival from initial size $x$ ($S(x,a;\pmb{\tau^S_{a}})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau^G_{a}})$) are both dependent on initial size and ant state. 
+As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau_{a}}$). 
+Ant transition function $\rho(a',a,x;\pmb{\tau^{\rho}})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
+This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau^\rho}$).
 
@@
 In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transition between ant states conditional on plant size and previous ant state. 
-All models were fit using Stan and Rstan (CITE). 
+All models were fit using Stan and Rstan (\tom{}{Cite Stan and RStan.}). 
 Unless otherwise mentioned, all models use vague priors.
@@
 Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds. 
+\tom{}{This paragraph needs to address how we assigned seed number to the ``other'' category.}
 

### 2024-01-21 Manuscript/ant_cactus_AmNat.tex
manuscript comments

@@
 \section*{Abstract}
-\tom{}{I think this is too long for Am Nat requirements. Also, use ``we''.}
+\tom{}{I think this is too long for Am Nat requirements.
+Also, use ``we''.
+I think you just pasted in an abstract that you used elsewhere, so I will work on this once you write a real abstract for the Am Nat paper.}
 Mutualisms are widespread species interactions with diverse and dynamic consequences. 
@@
 The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that \tom{deviates are specific to plants in ant state $a$}{Why does the tau vector for viability have the $a$ subscript but the others do not?}.
-Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A$) giving total seed production. 
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
 Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
@@
 All models were fit using Stan and Rstan (\tom{}{Cite Stan and RStan.}). 
-Unless otherwise mentioned, all models use vague priors.
+Unless otherwise mentioned, all used vague priors.
 
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$.
-We fit this model to size data at the end of the transition year $y^G$ using the location-scale parameterization of the student $t$ distribution, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
+The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given \tom{the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$.
+}{You need to go a little farther than this, because the fixed effects are arranged as a second-order polynomial with size-ant interactions.
+That needs to be explained and the choice of polynomial should be justified.
+You also do not describe the size-dependence in variance and kurtosis.
+Also, I would subscript the $u$ as $u_{a}$, then explain what that means.
+The $u_{a}$ is important to describe in a biologically meaningful way because this is how the portfolio effect is tested.
+Finally, notice that the tau vector does not actually appear in the model.
+This is a problem.
+You can solve it by changing u to tau, or vice versa.}
+We fit this model to size data at the end of the transition year $y^G$ using the location-scale parameterization of the \tom{student $t$ distribution}{Don't kill me but I think we should either figure out a skewed t or go back to the skew normal.
+But just to comment on the text as is, you need to describe the three parameters of the t (e.g., G-hat is the mean) in order for the equations to make sense.}, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
 Specifically, the model was: 
@@
 $$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu} \times x $$
-where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ with ant $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as predictors here because ant partners defend cholla from herbivores, preventing the loss of limbs and therefore shrinking.
+\tom{}{I think the equations will be more readable if you do not use the $\times$ symbol for multiplication.}
+where \tom{$u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ with ant $a$ in year $t$}{This description is not correct.
+The variance is not ant-specific, the deviates are.}, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+\tom{Ants are included as predictors here because ant partners defend cholla from herbivores, preventing vegetative tissue loss.}{I think this is unnecessary since you provide the rationale for ant effects in the beginning of this section.}
 
 \paragraph{Survival}
-The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the \tom{probability of survival $y^S$}{$y^S$ is not the probability of survival} from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
 $$y^S \sim Bern(\hat{S})$$
 $$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times a + \beta_{3}^{S} \times x \times a + u + w$$
-where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-A Bernoulli distribution was chosen here because the only two possible outcomes are survival or death. 
-Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these.
+\tom{where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.}{Much of this is redundant with the growth model, and likely redundant with models below.
+Think about efficient ways to economize text so that you are not re-explaining the same things across sub-models.}
 
 \paragraph{Reproduction}
-The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing $y^P$ in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the \tom{probability of reproducing $y^P$}{Same comment: $y^P$ is not the probability of reproducing.} in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
 $$y^P \sim Bern(\hat{P})$$
-$$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w$$
+$$logit(\hat{P} = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w)$$
 where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
@@
 \paragraph{Flowers}
-The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Negative Binomial distribution:
+The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a zero-truncated negative binomial distribution:
 $$y^{F} \sim 0 Truncated Negative Binom(\hat{F},\hat{\phi})$$
@@
 
-\paragraph{Viability}
-The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates the proportion of flowers produced by a plant which are viable (not aborted) $y^V$ in year $t+1$, with fixed effects of the ant partner of the cactus $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
+\paragraph{Flowerbud viability}
+The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates the \tom{proportion of flowers produced by a plant which are viable (not aborted) $y^V$}{$y^V$ is not the proportion of viable buds.} in year $t+1$, with fixed effects of ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
 $$y^{V} \sim Binom(\hat{V})$$
@@
 where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion.
 
 \paragraph{Ant Transitions}
-The ant transition rates model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$ $y^{\epsilon}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$ \tom{$y^{\epsilon}$}{This is not the probability of being visited.}, with fixed effects of the previous size \tom{of the cholla}{I will not make this comment everywhere but notice that ``of the cholla'' does basically no work in this sentence.
+It is very clear to readers at this point that ``size'' refers to size of the cholla, so these words can be cut.
+More generally, edit your own writing to remove words (or sentences) like this that take up space but provide no new information.} $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a multinomial distribution:
 $$y^{\epsilon} ~ Multinomial(\hat{\epsilon})$$
@@
 where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$.
-Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
 
 \paragraph{Recruit Size Distribution}
-The recruit size model ($n(x,a')$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
+The recruit size model \tom{($n(x,a')$)}{Not the correct notation.} estimates the size distribution of all recruits \tom{$y^{\eta}$}{Think deeply about what the $y$'s represent and how you can better describe them.} from a given year $t+1$, with no fixed or random effects, using a Normal distribution:
 $$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
@@
 \paragraph{Germination}
-With germination data from \tom{Miller et al., 2007}{CITE}, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($\gamma_2$) in year $t+1$, with no fixed or random effects, using a Binomial distribution:
+With germination data from \tom{Miller et al., 2007}{CITE}, we fit two \tom{Bayesian generalized linear models}{they are all Bayesian glms so no need to say this here.} for the probability of germinating from \tom{a seed in the first year ($\gamma_1$) or the second year ($\gamma_2$)}{This can be described more accurately by referring to the 1 and 2 yo seed banks.} in year $t+1$, with no fixed or random effects, using a binomial distribution:
 $$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
@@
 \paragraph{Pre-Census Survival}
-With data collected in a \tom{2005-2006}{FIX}********** recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
+With data collected in a \tom{2005-2006}{FIX}********** recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (\tom{accounting for missed mortality events}{This could be explained more clearly, including the natural history that recruitment occurs months earlier than the census, so we need to account for mortality during that time.}), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
 $$y^{\delta} ~ Bern(\hat{\delta})$$
@@
 \paragraph{Seeds Per Flower}
-With data from \tom{Miller 2007}{CITE}********, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$, using a Negative Binomial distribution:
+With data from \tom{Miller 2007}{CITE}********, we fit a Bayesian generalized linear model for the number of \tom{seeds produced by every flower}{This is not seeds per flower, it is seeds per fruit, since the viability model accounts for the probability that a flower becomes a fruit.
+I would actually move this earlier to follow that order of events.} on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$, using a \tom{Negative Binomial}{Not sure why this is written as zero-truncated.} distribution:
 $$y^{\kappa} \sim 0 Truncated Negative Binomial(\hat{\kappa},\hat{\phi})$$
@@
 
-\paragraph{MCMC Simulations}
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite} 
-For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
-The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
-We did not thin the chains, thus all samples were retained. 
-To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
-To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
-Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-All estimated parameters are available in the data. 
-Data and code for all vital rate models is included in the supplemental information.
+\paragraph{Parameter estimation}
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+For each model, we obtained \tom{3}{Spell out numbers less than ten.} chains of 10,000 iterations, each with randomly chosen initial conditions. 
+The first 1,500 iterations were discarded. 
+We did not thin the chains, thus all samples were retained following burn-in. 
+We assessed parameter convergence between and within chains (\tom{Fig.}{reference figure}). 
+To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (\tom{Fig.}{reference figure}). 
+\tom{All estimated parameters are available in the data. }{What does this mean?}
+\tom{Data and code for all vital rate models is included in the supplemental information.}{I do not know what this refers to.
+Am Nat (and most journals) requires a data accessibility statement that will need to include details of data and code availability.}
 
 \subsection*{IPM Analysis}
+\tom{}{I think this section needs to start with an overview paragraph descibing in plain English what you are trying to do with the fitted model, before getting into any technical details.}
+Analyzing an IPM requires discretizing the continuous IPM kernel into an approximating matrix \tom{to calculate the dominant eigenvalue}{The dominant eigenvalue is but one of many, MANY forms of ``analysis'' of an IPM.
+I would cut this bit that I highlighted.}.
+Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
+In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of \tom{$m b \times m b$ where $m$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).}{I think what you call $m$ here is actually just $A$, the total number of ant states.}
+Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $m(b+2) \times m(b+2)$.
+\tom{We used $b = 200$ bins.}{It is also useful to briefly acknowledge \emph{why} 200.}
+\tom{We extend the integration limits $L$ and $U$ to avoid unintentional eviction \cite{Williams2012}.}{You have not said how you estimate $L$ and $U$ from the data.}
 
-Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-In a traditional IPM $x$ is discretized into $b$ bins, replacing the continuous kernel into a $b \times b$ matrix, in this case there is additional complexity in the form of transitions between ant partners.
-Each combination of previous ant state $a$ and next ant state $a'$ represents a unique set of plants in the popualtion and therefore must be discritized individually, leading to a matrix size of $m b \times m b$ where $m$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
-In this model we have two additional discrete states (year one and year two seed banks) leading to a final matrix size of $m(b+2) \times m(b+2)$.
-We used $b = 200$ bins.
-We extend the integration limits $L$ and $U$ to avoid unintentional eviction \cite{Williams2012}.
 Traditionally, in a deterministic IPM the asympotic population growth rate $\lambda$ is estimated as the dominant eigenvalue of the discrete kernel.
-In a deterministic bayesian IPM, we create 1,000 discrete kernels with a unique set of parameters from our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution. 
-In a deterministic version of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
+\tom{In a deterministic bayesian IPM,}{I thought you were only using the two types of stochastic models (and I have not read ahead).} we create \tom{1,000 discrete kernels with a unique set of parameters}{This should explicitly refer to samples from the posterior distribution of parameter values} from our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution.
+\tom{In a deterministic version of our IPM}{At this point, it is not clear to readers how many ``versions'' you have and why, which is why an overview paragraph providing a ``roadmap'' will be an important addition.}, \tom{this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
 opuntiae} and vacancy; other and vacancy; \textit{L.
@@
 apiculatum}, other, and vacancy; \textit{C.
-opuntiae}, other, and vacancy; and all ant partners and vacancy.
-In order to calculate stochastic $\lambda$ distributions for each ant scenario, every one of the 1,000 parameter-iteration associated $\lambda$s actually comes from the mean of 5,000 year-specific estimations each associated with a year random effech which was randomly selected from the 18 years of data.
+opuntiae}, other, and vacancy; and all ant partners and vacancy.}{Rationale for why single ant scenarios are coupled with vacancy should be better developed.}
+In order to calculate stochastic $\lambda$ distributions for each ant scenario, \tom{every one of the 1,000 parameter-iteration associated $\lambda$s comes from the mean of 5,000 year-specific estimations each associated with a year random effech which was randomly selected from the 18 years of data.
+}{Even having worked on this with you for a long time, I cannot follow what you are saying here.
+This will need careful editing so that the methods and rationale for those methods (just as important!) are crystal-clear.}
 
-We compare the distributions of $\lambda$ across each combination of ant partners to whether sampling effect or complementarity is at play in the system.
-To compare the distributions of two hypothetical populations, A and B, we subtract the vector of 1,000 $\lambda_A$ estimations of one population from the 1,000 $\lambda_B$ estimations of the other ($\lambda_A - \lambda_B$).
-If the average of these differences is positive, population A has a higher $\lambda$ than B, if the average is 0, they are equal, and if the average is negative, population B has a higher $\lambda$ than A.
+\tom{We compare the distributions of $\lambda$ across each combination of ant partners to whether sampling effect or complementarity is at play in the system.}{I do not think you have sufficiently developed the logic for how these comparisons inform hypotheses about different mechanisms.
+Also, and this is a matter of taste, I find the language ``at play in the system'' to be not very scientific and would prefer to phrase this more precisely.}
+To compare the distributions of two hypothetical \tom{populations, A and B}{It is not clear what this toy example stands in for in the real model.}, we subtract the vector of 1,000 $\lambda_A$ estimations of one population from the 1,000 $\lambda_B$ estimations of the other ($\lambda_A - \lambda_B$).
+\tom{If the average of these differences is positive, population A has a higher $\lambda$ than B, if the average is 0, they are equal, and if the average is negative, population B has a higher $\lambda$ than A.}{I think this could be edited to more effectively communicate what is gained from these contrasts.
+The average will never be exactly zero.
+The important element, IMO, is that we gain a probabilistic estimate of which combinations have higher or lower fitness, given the data.}
 
-To determine if sampling effect is at play in a system, we must first determine if there is a "best" partner, by determining which single ant association is correlated with the highest $\lambda$ estimation.
-We can do this by comparing the distributions of each $\lambda$ and finding the one which is larger than all others.
-Then we must show that the $\lambda$ of a population with all possible partners is equal to that of $\lambda$ of a population with only the "best" partner, by comparing the distributions as described earlier.
+\tom{To determine if sampling effect is at play in a system}{Ditto previous comment re: ``at play''.
+But it is also important to recognize, as I have emphasized before, that these are mechanisms that could explain positive diversity effects.
+Your writing here presumes a positive diversity effect.}, we must first determine if there is a "best" partner, by determining which single ant association is correlated with the highest \tom{$\lambda$}{Here and elsewhere, it is important to use correct notation for fitness.
+$\lambda$ is population mean fitness in a constant environment.
+$\lambda_{S}$ is population mean fitness in a stochastic environment.
+If you are talking about stochastic environments, you should be using $\lambda_{S}$.
+I also think it would be helpful to notate either $\lambda$ with ant partner, e.g.
+$\lambda^{C}$ is fitness corresponding to exclusive association with \textit{Crematogsaster}.
+I think that would help communicate methods.} estimation.
+\tom{We can do this by comparing the distributions of each $\lambda$ and finding the one which is larger than all others.}{Distributions are not larger than one another.
+They have properties -- moments -- that can differ, like the mean or standard deviation.
+Be more precise here.}
+Then we \tom{must show}{I don't like ``must show'' -- sounds like we are hunting for something.} that the $\lambda$ of a population with all possible partners is equal to that of $\lambda$ of a population with only the "best" partner, by comparing the distributions as described earlier.
 If it is significantly larger then complementarity is potentially at play (more on this in the next paragraph).
@@
 To determine if portfolio effect is at play we must show that partner diversity buffers the population from environmental variation.
-To do this we parameterized a new set of bayesian statistical models in which year random effects were not ant specific, meaning the effects of ant partners were not able to covary with year effects (the scenario in which all ant partners must respond to annual variation the same way and portfolio effect is not possible).
-The new estimations of year random effects are $u \sim N(0,\sigma_{yr}^2)$ with year specific variance of $\sigma_{yr}$. 
-With this verision of the IPM (now referred to as the stochastic null IPM) we followed the same approach to calculate distributions of stochastif $\lambda$s for each ant scenario.
+\tom{To do this we parameterized a new set of bayesian statistical models in which year random effects were not ant specific, meaning the effects of ant partners were not able to covary with year effects (the scenario in which all ant partners must respond to annual variation the same way and portfolio effect is not possible).
+The new estimations of year random effects are $u \sim N(0,\sigma_{yr}^2)$ with year specific variance of $\sigma_{yr}$.}{This needs to be explained earlier, in the statistical estimation sections, so that this section can focus on teasing out mechanisms from the model.
+I also think you need to better explain what the random year effects represent, biologically.
+We do not have any data on how the ants are actually fluctuating, but we do know something about how the effects of the ants on plants fluctuates.
+That is a critical nuance and it does not currently come through.
+I am also wondering whether it would be preferable to have the $a$ coefficients temporally fluctuate rather than making the $u$ ant-specific.
+Or are these effectively the same thing?
+(we could probably figure it out on a whiteboard)}
+With this verision of the IPM (now referred to as the stochastic null IPM) \tom{we followed the same approach to calculate distributions of stochastic $\lambda$s for each ant scenario.}{It is still not clear if what you described above was based on deterministic or stochastic $\lambda$.}
 We can estimate the effects of partner diversity on the focal mutualist by calculating the difference in $\lambda$ distributions between populations with no partners and populations with all partners present.
@@
 
+\tom{}{General comment: the methods for this analysis are a little complicated, which is an inevitable consequence of asking a complex question.
+In these cases, a writer needs to really step up to communicate something complicated in a way that is sufficiently simple that it can be easily understood and, ideally, replicated.
+I think with focused effort you can significantly improve this section.
+Again: put yourself in the shoes of the reader and think about what they need from you in order to ``get it''.}
+
 \section*{Results}
+\tom{}{I strongly suggest re-structuring the Results section such that each sub-section is titled with a question from the end of the Intro, and the content of the sub-section answers that question.}
 \subsection*{IPM Analysis}
@@
 Using the results of the stochastic IPM, we determined if complementarity or sampling effect was at play within this system. 
-We found that populations of tree cholla which have no partners have a mean fitness of $\lambda = 0.9861$ \ref{}.
+\tom{We found that populations of tree cholla which have no partners have a mean fitness of $\lambda = 0.9861$ \ref{}.
 Populations with only one ant partner present (\textit{L.
@@
 opuntiae} and other) have a mean fitness of $\lambda = 0.9942$, $\lambda = 0.9943$, and $\lambda = 0.9913$ respectively.
-Populations with all ant partners present have a mean fitness of $\lambda = 0.9940$.
+Populations with all ant partners present have a mean fitness of $\lambda = 0.9940$.}{This is not an effective use of Results text. You are just reporting values -- this is why we have figures!}
 The only partner scenarios where $\lambda > 0.99$ are when \textit{L. apiculatum} ants are present with any other ant partner. 
 The means are not the only important result.
-We used bayesian modeling, so each of these $\lambda$ estimations has a distribution rather than a single estimate.
+We \tom{used bayesian modeling}{I recommend less emphasis on this.
+It is not the only way to get a distribution of lambda values, I don't think it is the most important thing to emphasize about the study, and you risk rankling reviewers who might think you are trying to sound impressive with ``fancy statistics''.}, so each of these $\lambda$ estimations has a distribution rather than a single estimate.
 By subtracting one of these distributions from the other, the difference from 0 tells us how much larger or smaller the distributions are with a percent confidence. 
@@
 Any ant partner is shown to lead to an increase in fitness of the tree cholla in these results. 
-We are between 82\% to 100\% confident that having any ant parnter leads to a higher $\lambda$ than having no partners.
+\tom{We are between 82\% to 100\% confident that having any ant parnter leads to a higher $\lambda$ than having no partners.}{Where can readers see this result?
+You say there are distributions lambdas and differences but you do not show them, and that will need to be remedied.
+ALso, in terms of fitness contrasts, I think it is more important to start with the Vacant -- All Ant contrast.
+``all ants'' is our best representation of what is happening in nature, so it is relevant to first ask if there is evidence that these interactions, collectively, are beneficial, and then use the hypothetical scenarios to unpack how those benefits arise.
+}
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/LambdaMeans.png}
+	\caption{THIS FIGURE HAS NO LEGEND. Are you sure this is not deterministic and stochastic? I am surprised that both stochastic models would be this different.}
 \end{figure}
@@
 apiculatum} ants are included as partners leads to a higher $\lambda$ for the tree cholla than any diversity scenario without these parnters.
-These results indicate that sampling effect, not complementarity, is at play in this system.
+\tom{These results indicate that sampling effect, not complementarity, is at play in this system.}{I suspect we may get some push-back on this interpretation, because the sampling effect is a mechanisms to explain benefits of diversity.
+Based on your results, there are no benefits of diversity, there is just a best partner.
+I am not suggesting to entirely discard the sampling effect idea, but it is important to recognize that when you discuss these mechanisms, they are mechanisms for positive diversity effects.
+So the absence of any diversity effect seems like an important result to contextualize how you interpret evidence for a sampling effect.}
 
 %%%% Stochastic null follows similar patterns but dampened
+\tom{}{Throughout this paragraph and much of the results section, you are not explicitly referencing figures, so it is not always clear what is the basis for some of your assertions.
+Also, the figures need legends.}
 Using the stochastic null IPM results, we found that when ant partners responded the same way to inter-annual variation similar patterns were found, though the magnitude of the patterns were somewhat different.
 We used the results of the stochastic null IPM to determine if portfolio effect is at play. 
-We found that both the stochastic model and stochastic null model resulted in higher fitness when all ants are present compared to no partners 94\% of the time. 
-These results differed in the magnitude of fitness boost recieved from partners.
+We found that both the stochastic model and stochastic null model resulted in higher fitness when all ants are present compared to no partners \tom{94\% of the time}{This is not a measure of time, it is a measure of confidence based on the posterior probability of the parameters given the data.}.
+These results differed in the magnitude of fitness boost received from partners.
 In the stochastic IPM, tree cholla with all ants present resulted in a $\lambda$ that was 0.008 greater than tree cholla with no ants present; in the stochsatic null IPM, tree cholla with all ants rpesent resulted in a $\lambda$ that was 0.013 greater than tree cholla with no ants present.
-We are 52\% confidence that the difference is greater when inter-annual effects were not ant dependent (in the stochastic null model).
-This indicates that portfolio effect is at play, if only weakly evident.
+\tom{We are 52\% confident that the difference is greater when inter-annual effects were not ant dependent (in the stochastic null model).}{If I understand this correctly (and I am not sure because you are not showing the results) then it is equally likely that fitness is either higher or lower when ant effects do or do not fluctuate synchronously -- this tells me there is no evidence for a portfolio effect.}
+\tom{This indicates that portfolio effect is at play, if only weakly evident. }{This is not a very scientific statement.}
 
 \subsection*{Statistical Modeling}
-
+\tom{}{In my opinion it does not make sense to have the statistical modeling results follow the IPM results.
+But I also think the results would benefit from restructuring around the motivating questions (which would put vital rates first).}
 \paragraph{Growth Model}
@@
 	\includegraphics[width=0.95\linewidth]{Figures/GrowContourLinesColor.png}
-	\caption{asdasdf}
+	\caption{THIS FIGURE NEEDS A CAPTION}
 	\label{fig:growth}
 \end{figure}
-Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes \ref{fig:growth}e.
-The mean growth rate expected for plants that are 10 $cm^3$ or smaller in year t is 10.01 $cm^3/yr$, 4.65 $cm^3/yr$, 7.44 $cm^3/yr$, and 5.14 $cm^3/yr$ when tended by \textit{C.
+\tom{}{Here and elsewhere, begin paragraphs with the main result that you want readers to see.
+In this case I think you want a topic sentence to emphasize that there is evidence that ant visitation enhances cactus growth, partner identity can influence the growth trajectory.
+That is a simple statement but it captures the essential reason why you did this analysis in the first place.
+That connection to your motivating questions is often lacking.}Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes \tom{\ref{fig:growth}e}{When you reference figures like this the ref function only returns the number, so you need to type ``Figure''.}.
+\tom{The mean growth rate expected for plants that are 10 $cm^3$ or smaller in year t is 10.01 $cm^3/yr$, 4.65 $cm^3/yr$, 7.44 $cm^3/yr$, and 5.14 $cm^3/yr$ when tended by \textit{C.
 opuntiae}, \textit{L.
@@
 apiculatum}, other ants, or no ants respectively.
+}{Similar to my comment above, this type of recapitulation of the figure is not a very effective use of space.}
 Plants with \textit{C.
@@
 
-We are 89\%, 88\%, and 70\% confident that plants tended by \textit{C.
+\tom{We are 89\%, 88\%, and 70\% confident that plants tended by \textit{C.
 opuntiae} ants experience higher mean growth rates across sizes than plants tended by no parnters, \textit{L.
@@
 opuntiae}, \textit{L.
-apiculatum}, or other ants respectively.
+apiculatum}, or other ants respectively.}{The empirical basis for these assertions is not clear.
+I am also a little confused because the text above makes it sounds like the ranks are dependent on size.}
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/SurvivalPlot.png}
-	\caption{asdasdf}
+	\caption{THIS FIGURE NEEDS A CAPTION, AND IT SHOULD EXPLAIN THE DASHED LINES AMONG OTHER THINGS.}
 	\label{fig:surv}
@@
 opuntiae}, \textit{L.
-apiculatum}, or other ants respectively.
+apiculatum}, or other ants respectively.\tom{}{Most of my comments from previous sections apply here, and it is uncear why you select specific contrasts to report statistical confidence but not others.}
 
@@
 There is a clear size effect on the number of flowers produced. 
-The mean number of flowers produced by a plant remains at 0 until the plant reaches medium sizes after which the mean number of flower produced increases eponentially to about 40 flowers per plant per year at large sizes.
+The mean number of flowers produced by a plant remains at 0 until the plant reaches medium sizes after which the mean number of flower produced increases to about 40 flowers per plant per year at large sizes.
 The mean number of flowers produced for a plant that is 150 $cm^3$ or smaller in year $t$ is <1 flower per plant.
-The mean number of flowers produced for a plant that is larger than 150 $cm^3$ in year $t$ is 8.7 flowers per plant.
+The mean number of flowers produced for a plant that is larger than 150 $cm^3$ in year $t$ is 8.7 flowers per plant.\tom{}{If you are not showing these figures in the main text that you should provide them in an appendix, and reference them there.
+In fact, anywhere there is no role for ants I would move all those results to an appendix section.
+It is not clear to me whether you text here is referring to the prob flowering model, the number of flowerbuds model, or both.}
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/ViabHist.png}
-	\caption{asdasdf}
+	\caption{THIS FIGURE NEEDS A CAPTION.}
 	\label{fig:viab}
@@
 Each viable flower on a plant produces between 97 and 257 seeds.
-This number is affected by the ant partner present. 
+\tom{This number is affected by the ant partner present. 
 \textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
-\textit{L. apiculatum} tended plants produce a mean of 149 seeds per flower. 
+\textit{L.
+apiculatum} tended plants produce a mean of 149 seeds per flower.
+}{These results are not consistent with Ohm and Miller, where Crem had lower seeds than Liom.
+I would check this.
+This section should also reference that paper because these are not new results.}
 Vacant plants produce a mean of 160 seeds per flower. 
@@
 
-%% Precensus Survival\
+\tom{%% Precensus Survival\
 \paragraph{Precensus Survival Model}
@@
 %% Recruit size distribution
-New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.
+New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.}{Move to an appendix. These results are not relevant for the questions at hand.}
 
@@
 \paragraph{Ant Transition Model}
+\tom{}{This section is hard to read, and I think it needs to be restructured in a way that is guided by what you want readers to know as it relates to your questions.
+You provide lots and lots of statistics and contrasts without much context for why these matter and how they should be interpreted.
+I suggest you think about the patterns that emerge for you in the figure, and re-write this section with that in mind.
+I do not think this is much value in breaking it down into the four ant states.
+You should be able to describe transition dynamics in an engaging and succinct way in just a paragraph or two.}
 \begin{figure}
 	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+	\caption{THIS FIGURE NEEDS A CAPTION.}
 \end{figure}
@@
 \textit{L.
-apiculatum} ants are the most likely tender in year $t+1$ for all large plants that were not previously tended by \textit{C.
+apiculatum} ants are the most likely \tom{tender}{Not sure I would use this as a noun.} in year $t+1$ for all large plants that were not previously tended by \textit{C.
 opuntiae} ants in year $t$, while \textit{C.
@@
 opuntiae} ants in year $t+1$.
-We are 93\%, 89\%, 86\%, and 93\% confident that plants that are smaller than 150 $cm^3$ in year $t$ will be vacant in year $t+1$ when they were tended by no partners, \textit{C.
+\tom{We are 93\%, 89\%, 86\%, and 93\% confident that plants that are smaller than 150 $cm^3$ in year $t$ will be vacant in year $t+1$ when they were tended by no partners, \textit{C.
 opuntiae}, \textit{L.
-apiculatum}, or other ants respectively in year $t$.
+apiculatum}, or other ants respectively in year $t$.}{Statements like these do not feel strongly motivated, and I am not sure what you want me to know as it relates to the motivating questions.}
 
@@
 \section*{Discussion}
+\tom{}{I have not commented too heavily here because I would like to talk through what we want this section to achieve.
+Currently, most of this section is actually describing results, whereas the pupose is to interpret and contextualize results, and connect them to the broader literature.
+Some of your content here is actually better at describing results than you had in the Results section, because it includes that element of ``how is this connected to the question'' that was missing from the Results section.}
 The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
@@
 Because these tree cholla interact with only one ant partner at a time, it is a unique system in which to parse out the individual effects of each ant partners, both in isolated settings or in combinations we cannot test in the real world.
+\tom{}{This is a good start.
+In general, I suggest opening the Discussion section with a brief summary of what you were trying to learn in this study, what you found, and its broader significance.
+This paragraph does a little of that, but I think it can be stronger.}
 
 %\subsection*{Demographic effects of association with alternative partners}
-We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, and more) have on the vital rates of tree cholla. 
+We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, \tom{and more}{I would not say this.}) have on the vital rates of tree cholla. 
 Using a system of heirarchical bayesian models we found that there were discernable differences in the effects that each partner had on vital processes of the focal mutualist. 
@@
 Reproducing plants experience a probability of reproducing, flower production, and floral abortion.
-Floral abortion is heavily affected by seed predators \cite{Miller2008}, which the ants defend the cacti, leading to increased floral viability.
+\tom{Floral abortion is heavily affected by seed predators \cite{Miller2008}}{Seed predators do not influence floral abortion.}, which the ants defend the cacti, leading to increased floral viability.
 We specifically found that tree cholla experienced the highest floral viability rates when tended by \textit{L. apiculatum} ants. 
@@
 %\subsection*{Effect of partner diversity on plant fitness and what mechanisms explain this}
-The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (as some have done before \cite{Palmer2010}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
-We found that the combination of accurate partner transitions with partner identity affected the fitness of the tree cholla in interesting and dynamic ways. 
+The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (\tom{as some have done before \cite{Palmer2010}}{There is more than just the Palmer pape.
+Be sure that you are comprehensive in your use of the literaure, include non-ant-plant studies,}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
+We found that the combination of \tom{accurate partner transitions with partner identity}{Not sure what this means.} affected the fitness of the tree cholla in interesting and dynamic ways. 
 Namely, a best partner emerged in this analysis, which was surprising given the nature of our vital rate findings. 
@@
 %\subsection*{Synthesis????? All together}
-\ali{}{What this all means more broadly?? I'm currently not sure what to do with this}
+\ali{}{What this all means more broadly?? I'm currently not sure what to do with this. Tom: that's the entire Discussion!}
 This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
-Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+\tom{Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
 Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
+The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is really good. More of this!}
 
 %\subsection*{Future Directions}
-This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
+\tom{This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
 We revealed the dynamics of partner turnover and showed that different ant partners are correlated with different fitness benefits.
-As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work.
+As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work. }{This feels a little weak and incomplete.}
 
 \section*{Acknowledgments}
-
+This should be drafted.
 %%%%%%%%%%%%%%%%%%%%%
@@
 \section*{Data and Code Availability}
+This should be drafted.
 
 \section*{Appendix A: Additional Methods and Parameters}
-
+This is not referenced in the paper, to my knowledge, and I think you need to think more deeply about what content should go into appendices and why. 
 % In most cases, authors should typeset supplementary material in a separate,

### 2024-02-22 01_cholla_ant_IPM_setup.R
Figures Updates

@@ -77,6 +77,34 @@ cactus$Plot[cactus$Plot=="HT4B1" | cactus$Plot=="HT4B2"]<-"T4"
 cactus$logsize_t <- log(volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t))
 cactus$logsize_t1 <- log(volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1))
 
+## Get the frequencies of ant species
+Antsp_t_levels <- levels(cactus$Antsp_t)
+# liom
+a <- subset(cactus, cactus$Antsp_t == "Liom" | cactus$Antsp_t == "LIOM" | cactus$Antsp_t == "liom" | cactus$Antsp_t == "L" | cactus$Antsp_t == "LIOM " |  cactus$Antsp_t == "liom ")
+nrow(a)
+# crem
+b <- subset(cactus, cactus$Antsp_t == "CREN" | cactus$Antsp_t == "LCREM" | cactus$Antsp_t == "crem" | cactus$Antsp_t == "Crem" | cactus$Antsp_t == "C" | cactus$Antsp_t == "CREM" | cactus$Antsp_t == "VCREM")
+nrow(b)
+# camp
+c <- subset(cactus, cactus$Antsp_t == "camp" | cactus$Antsp_t == "CAMP" | cactus$Antsp_t == "CAMP" | cactus$Antsp_t == "large black shiny" | cactus$Antsp_t == "LARGE BLACK SHINY" | cactus$Antsp_t == "drpoff" | cactus$Antsp_t == "dropoff")
+nrow(c)
+# honeypot
+d <- subset(cactus, cactus$Antsp_t ==  "HNEYPOT" | cactus$Antsp_t ==  "honeypot" | cactus$Antsp_t == "HONEYPOT")
+nrow(d)
+# phen
+e <- subset(cactus, cactus$Antsp_t == "phen" | cactus$Antsp_t == "PHEN" | cactus$Antsp_t == "aph" | cactus$Antsp_t == "unk (Aphaeno?)")
+nrow(e)
+# tetra
+h <- subset(cactus, cactus$Antsp_t == "tetra")
+nrow(h)
+# brachy
+j <- subset(cactus, cactus$Antsp_t == "brachy")
+nrow(j)
+# unknown
[…]

### 2024-02-22 05_cholla_ant_IPM_figures_tables.R

@@ -1489,11 +1489,11 @@ prop
 all_vac_stoch_null <- lams_stoch_null$all - lams_stoch_null$none
 all_vac_stoch <- lams_stoch$all - lams_stoch$none
 # Plot the boost offered by the real ant scenario based on stochastic and deterministic lambda estimates
-png("portfolio_effect.png")
-plot(density(all_vac_stoch_null), lwd = 3, col = "chartreuse4", ylim = c(0,60))
+png("Figures/portfolio_effect.png")
+plot(density(all_vac_stoch_null), lwd = 3, col = "chartreuse4", ylim = c(0,100), xlab = "Effect of Partner Presence on Fitness", main = "", cex.lab = 2)
 lines(density(all_vac_stoch), lwd = 3, col = "violet")
 abline(v = 0, lty = 2, lwd = 3)
-legend("topleft",legend = c("Synchronicity Possible","Synchronicity Excluded"), fill = c("violet","chartreuse4"))
+legend("topright",legend = c("Synchronicity Possible","Synchronicity Excluded"), fill = c("violet","chartreuse4"))
 dev.off()
 # check the mean density 
 mean(all_vac_stoch_null>0)
@@ -1508,7 +1508,21 @@ length(subset((all_vac_stoch-all_vac_stoch_null), (all_vac_stoch-all_vac_stoch_n
 mean(lams_stoch$all) - mean(lams_stoch$none)
 mean(lams_stoch_null$all) - mean(lams_stoch_null$none)
 
+#################################################################################################
+#################################################################################################
+######## VISUALIZE THE MEAN LAMBDA BY NUMBER OF ANT PARTNERS RATHER THAN IDENTITY ###############
+#################################################################################################
+#################################################################################################
+## No partners mean
+zero_part <- mean(lams_stoch$none)
+## One partners mean
+one_part <- mean(colMeans(cbind(lams_stoch$cremvac,lams_stoch$liomvac, lams_stoch$othervac)))
+## Two partners mean
+two_part <- mean(colMeans(cbind(lams_stoch$liomcremvac,lams_stoch$liomvacother, lams_stoch$othercremvac)))
+## All partners mean
+all_part <- mean(lams_stoch$all)
 
-
-
-
+## Plot these means
+png("Figures/Lambda_Num_Partners.png")
[…]

### 2024-02-22 Visuals Not simplified.R

@@ -212,6 +212,57 @@ length(subset(v_c, v_c>0))/205
 length(subset(v_l, v_l>0))/205
 length(subset(v_o, v_o>0))/205
 
+
+## Annual variation of growth models 
+## Show the correlation between ant and year -- from growth model random effects
+g_crem <- colMeans((grow_rfx1))
+g_liom <- colMeans((grow_rfx2))
+g_other <- colMeans((grow_rfx3))
+g_vac <- colMeans((grow_rfx4))
+g_years <- seq(2004,2022, by = 1)
+## Show the correlation between ant and year -- from survival model random effects
+s_crem <- colMeans((surv_rfx1))
+s_liom <- colMeans((surv_rfx2))
+s_other <- colMeans((surv_rfx3))
+s_vac <- colMeans((surv_rfx4))
+s_years <- seq(2004,2022, by = 1)
+## Show the correlation between ant and year -- from growth model random effects
+v_crem <- colMeans((viab_rfx1))
+v_liom <- colMeans((viab_rfx2))
+v_other <- colMeans((viab_rfx3))
+v_vac <- colMeans((viab_rfx4))
+v_years <- seq(2004,2023, by = 1)
+unique(viability_data$Year_t)
+png("Figures/year_ant_timeseries.png")
+par(mar=c(4,2,2,1),oma=c(2,2,0,0))
+layout(matrix(c(1,2,3),
+              ncol = 3, byrow = TRUE), heights = c(1), widths = c(4,4,4))
+## Growth Ant EEffects
+plot(g_years,g_liom,col = liomcol, cex.main = 2,type = "b",lwd = 4, pch = 16,cex = 2,
+     main = "a)                               ",
+     ylim = c(-1.5,2.1), xlab = " ",ylab = " ",cex.lab = 2)
+lines(g_years, g_crem, type = "b", col = cremcol, lwd = 4, pch = 16, cex = 2)
+lines(g_years, g_vac, type = "b", col = vaccol, lwd = 4, pch = 16, cex = 2)
+lines(g_years, g_other, type = "b", col = othercol, lwd = 4, pch = 16, cex = 2)
+legend("topleft",legend = c("Liom.","Crem.","Other","Vacant"),fill = c(liomcol,cremcol,othercol,vaccol),cex=1.8)
+## Survival ant effects
+plot(s_years,s_liom,col = liomcol, cex.main = 2,type = "b",lwd = 4, pch = 16,cex = 2,
+     main = "b)                                 ",
[…]

### 2024-02-22 Manuscript/ant_cactus_AmNat.tex
Manuscript Updates

--- 2024-01-19 Manuscript/ant_cactus_AmNat.tex
@@
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \cite{Noe1994,Leigh2010}.
+They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
 
@@
 Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \cite{Afkhami2014}. 
-Third, partner species and \tom{herbivores}{notice that ``herbiviores'' does not fit here, at least not yet.
-you have not said anything about ants and plants, you are talking very generally.
-put yourself in the mind of a reader as you write, this will help ensure that ideas appear in an order that can be digested.} can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}.
+Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
 Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits and protection across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
@@
 Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits through anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the \tom{herbivore populations}{I would be careful here -- unless you are using the herbivore counts I don't think you have data on this.}.
+However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the herbivore populations (data on herbivory is included in supplementary documents).
 
@@
 We recorded the ant species present (or vacancy if no ants present).
-Occurrences of more than one ant species on one plant were rare (\tom{}{quantify}), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species. 
+Occurrences of more than one ant species on one plant were rare (less than 5\% of observed ant species), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
 Plots were searched for new recruits each year, and these were added to the census.
-\tom{In total, the data set included \# unique individuals and \# plant-year observations. }{update numbers}
+In total, the data set included 1141 unique individuals and 19 plant-year observations. 
 These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
@@
 apiculatum} and \textit{C.
-opuntiae} (\tom{}{Quantify this.}) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
+opuntiae} (7.78\% of observed ants) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
 opuntiae}, and ``other''.
-\tom{The ``other'' category was made of unidentified ant species and ant species which occurred at relatively low frequencies compared to \textit{C.
-opuntiae} and \textit{L.
-apiculatum} (such as unknown species belonging to the genus \textit{Aphenogaster} and unknown species belonging to the genus \textit{Camponotus}).}{I think you can do better than this because we actually have the designated species in the data.
-I would report here the actual breakdown (percentages) of species in the ``other'' group.
-I expect Forelious will be a big chunk of it.}
+The ``other'' category was made of unidentified ant species (2.8\% of observed ants), unknown species belonging to the geni \textit{Aphenogaster} (0.4\%),  \textit{Camponotus} (0.9\%),  \textit{Tetramorium} (0.02\%), \textit{Brachymyrmer} (0.02\%), a honeypot ant (0.08\%), and \textit{Forelius pruinosus} (3.5\%).
 Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
@@
 	$$
-	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x';\pmb{\tau^P}) F(x';\pmb{\tau^F}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
+	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x;\pmb{\tau^P}) F(x;\pmb{\tau^F}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
 	$$
@@
 
-\noindent Functions \tom{$P(x';\pmb{\tau^P})$ and $F(x';\pmb{\tau^F})$}{These should be $x$ not $x'$ because we modeled flowering and fertility in year t based on size in year t.} give the probability of flowering and the number of flowerbuds produced, respectively, by plants of size $x'$ in year $t$.
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit (\tom{$\kappa(a')$}{Why is this $a'$?
-I think it should be $a$.
-More generally, I think you need to explain your use of primes and how readers should interpret them at the start of model exposition.}) are dependent on ant state $a$ but not size.
-The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that \tom{deviates are specific to plants in ant state $a$}{Why does the tau vector for viability have the $a$ subscript but the others do not?}.
-Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A$) giving total seed production. 
+\noindent In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
+Functions $P(x;\pmb{\tau^P})$ and $F(x;\pmb{\tau^F})$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$ but not size. 
+The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$ and year.
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
 Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
@@
 In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transition between ant states conditional on plant size and previous ant state. 
-All models were fit using Stan and Rstan (\tom{}{Cite Stan and RStan.}). 
-Unless otherwise mentioned, all models use vague priors.
+All models were fit using R version 4.2.2 and Rstan package (\cite{Rstancite, Rcite}).
+Unless otherwise mentioned, all used vague priors.
+
+The vital rate models take the form of the function:
+$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
+\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+Some models deviate from this general model, which will be specified below.
 
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$.
-We fit this model to size data at the end of the transition year $y^G$ using the location-scale parameterization of the student $t$ distribution, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
-Specifically, the model was: 
-$$y^G \sim Student T(\hat{nu},\hat{G},\hat{\sigma}) $$
-$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G} \times x  + \beta_{2}^{G} \times a + \beta_{3}^{G} \times x \times a  + \beta_{4}^{G} \times x^2 + \beta_{5}^{G} \times a \times x^2 + u + w $$
-$$\hat{\sigma} = \beta_{0}^{\sigma} + \beta_{1}^{\sigma} \times x $$
-$$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu} \times x $$
-where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ with ant $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as predictors here because ant partners defend cholla from herbivores, preventing the loss of limbs and therefore shrinking.
+The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+We fit this model to the size in year $t+1$ ($y^G$) using the location-scale parameterization of the student $t$ distribution and the identity link function, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
+$$y^G \sim Student T(\hat{\nu},\hat{G},\hat{\sigma}) $$
+Student $t$ distributions are dependent on three parameters, $\hat{\nu}$,  $\hat{G}$, and $\hat{\sigma}$.
+$\hat{G}$ is the mean of the distribution, and in our case is expressed as a second-order polynomial with ant-size interactions because  preliminary analysis found this was a better fit to slightly concave data.
+$\hat{\nu}$ is the shape parameter which determines how thin the tails of the distribution are.
+$\hat{\sigma}$ is the scale parameter which determines how wide the distribution is. 
+Both $\hat{\nu}$ and $\hat{\nu}$ are size dependent because preliminary analyses revealed that there was size dependence in both variance and kurtosis.
+%$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G}  x  + \beta_{2}^{G}  a + \beta_{3}^{G}  x  a  + \beta_{4}^{G}  x^2 + \beta_{5}^{G}  a  x^2 + u + w $$
+$$\hat{\sigma} = \beta_{0}^{\sigma} + \beta_{1}^{\sigma}  x $$
+$$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu}  x $$
 
 \paragraph{Survival}
-The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival $y^S$ from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
-$$y^S \sim Bern(\hat{S})$$
-$$logit(\hat{S} = \beta_{0}^{S}) + \beta_{1}^{S} \times x + \beta_{2}^{S} \times a + \beta_{3}^{S} \times x \times a + u + w$$
-where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-A Bernoulli distribution was chosen here because the only two possible outcomes are survival or death. 
-Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these.
+The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function
+$y^S \sim Bern(\hat{S})$.
+%$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u + w$$
 
 \paragraph{Reproduction}
-The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing $y^P$ in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Bernoulli distribution:
+The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
+We fit this model to the reproductive data $y^R$ using a Bernoulli distribution and a logit link function:
 $$y^P \sim Bern(\hat{P})$$
-$$logit(\hat{P}) = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w$$
-where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+%$$logit(\hat{P} = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w)$$
 
-\paragraph{Flowers}
-The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant $y^F$ in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$, using a Negative Binomial distribution:
+\paragraph{No. Flowers}
+The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
+We fit this model to flowerbud count data $y^F$ using a zero-truncated negative binomial distribution with a log transformation:
 $$y^{F} \sim 0 Truncated Negative Binom(\hat{F},\hat{\phi})$$
-$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u + w$$
+%$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u + w$$
 $$log(\hat{\phi}) = \beta_{0}^{\phi}$$
-where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
 
-\paragraph{Viability}
-The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates the proportion of flowers produced by a plant which are viable (not aborted) $y^V$ in year $t+1$, with fixed effects of the ant partner of the cactus $a$ in year $t$ and random effects of plot $w$ and year $u$, using a Binomial distribution:
+\paragraph{Flowerbud viability}
+The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates theproportion of flowers produced by a plant which are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+We fit this model to floral abortion data $y^V$ using a Binomial distribution and a logit link function:
 $$y^{V} \sim Binom(\hat{V})$$
-$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
-where $u \sim N(0,\sigma_{yr \times a}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$ and ant state $a$ in year $t$, and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion.
+%$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
+
+\paragraph{Seeds Per Fruit}
+With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
+$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
+%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
+$$\hat{\phi} = \beta_{0}^{\phi}$$
+The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
 
 \paragraph{Ant Transitions}
-The ant transition rates model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$ $y^{\epsilon}$, with fixed effects of the previous size of the cholla $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$, using a Multinomial distribution:
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
+We fit this model to ant partner data from year $t+1$ $a'$ using a multinomial distribution with a logit link function: 
 $$y^{\epsilon} ~ Multinomial(\hat{\epsilon})$$
-$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
-where $u \sim N(0,\sigma_{yr}^{2})$ is the year random effect with year specific variance of $\sigma_{yr}$.
-Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
+%$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
 
 \paragraph{Recruit Size Distribution}
-The recruit size model ($n(x,a')$) estimates the size distribution of all recruits $y^{\eta}$ from a given year $t+1$, with no fixed or random effects, using a Normal distribution: 
+%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
+We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
 $$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
-$$\hat{\eta} = \beta_{0}^{\eta}$$
+%$$\hat{\eta} = \beta_{0}^{\eta}$$
 where $\hat{\sigma}$ is estimated with a non-informative prior.
@@
 \paragraph{Germination}
-With germination data from \tom{Miller et al., 2007}{CITE}, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($\gamma_2$) in year $t+1$, with no fixed or random effects, using a Binomial distribution:
+With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
 $$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
 $$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
-$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
-$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
+%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
 
 \paragraph{Pre-Census Survival}
-With data collected in a \tom{2005-2006}{FIX}********** recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$, using a Bernoulli distribution:
+With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
+We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
 $$y^{\delta} ~ Bern(\hat{\delta})$$
-$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
  where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
 
-\paragraph{Seeds Per Flower}
-With data from \tom{Miller 2007}{CITE}********, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$, using a Negative Binomial distribution:
-$$y^{\kappa} \sim 0 Truncated Negative Binomial(\hat{\kappa},\hat{\phi})$$
-$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
-$$\hat{\phi} = \beta_{0}^{\phi}$$
-Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds. 
-\tom{}{This paragraph needs to address how we assigned seed number to the ``other'' category.}
-
-\paragraph{MCMC Simulations}
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite} 
-For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
-The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
-We did not thin the chains, thus all samples were retained. 
-To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
-To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
-Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-All estimated parameters are available in the data. 
-Data and code for all vital rate models is included in the supplemental information.
+\paragraph{Parameter estimation}
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+For each model, we obtained three chains of 10,000 iterations, each with randomly chosen initial conditions. 
+The first 1,500 iterations were discarded. 
+We did not thin the chains, thus all samples were retained following burn-in. 
+We assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b). 
+To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
+All estimated parameters are available in Table \ref{tab:Params}.
 
 \subsection*{IPM Analysis}
+We constructed two stochastic IPMs in order to simulate ant combinations that don’t exist in real life with two different versions of environmental variation.
+The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
+The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
+By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
+By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
+In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
 
-Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-In a traditional IPM $x$ is discretized into $b$ bins, replacing the continuous kernel into a $b \times b$ matrix, in this case there is additional complexity in the form of transitions between ant partners.
-Each combination of previous ant state $a$ and next ant state $a'$ represents a unique set of plants in the popualtion and therefore must be discritized individually, leading to a matrix size of $m b \times m b$ where $m$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
-In this model we have two additional discrete states (year one and year two seed banks) leading to a final matrix size of $m(b+2) \times m(b+2)$.
-We used $b = 200$ bins.
-We extend the integration limits $L$ and $U$ to avoid unintentional eviction \cite{Williams2012}.
-Traditionally, in a deterministic IPM the asympotic population growth rate $\lambda$ is estimated as the dominant eigenvalue of the discrete kernel.
-In a deterministic bayesian IPM, we create 1,000 discrete kernels with a unique set of parameters from our bayesian statistical models to estimate 1,000 $\lambda$ values as a distribution. 
-In a deterministic version of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
+Analyzing an IPM requires discretizing the continuous IPM kernel into an approximating matrix. 
+Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
+In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$ where $A$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
+Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $A(b+2) \times A(b+2)$.
+We used $b = 200$ bins after stability analysis to determine how many bins were needed to consistently estimate the same lambda.
+We extend the integration limits $L$ and $U$ (the lowest and highest size values observed) to avoid unintentional eviction \cite{Williams2012}.
+
+Traditionally, in a bayesian IPM, rather than having one set of parameters to populate an approximating matrix and estimate population fitness ($\lambda$), we use 1,000 draws from the posterior distributions of our bayesian statistical models to populate 1,000 approximating matrix and estimate a 1,000 element distribution for $\lambda$.
+After preliminary stability analysis we found that 1,000 random posterior draws leads to the same estimate of $\lambda$ as any greater number of posterior draws.
+In a stochastic bayesian IPM, the estimation of each individual $\lambda$ in the distribution comes from a geometric mean of many values. 
+Rather than a single approximating matrix for each value of $\lambda$, we take the geometric mean of 500 approximating matricies with year-ant-specific effects, resulting in a stochastic version of populationfitness ($\lambda_{S}$).
+These year-ant-specific effects are calculated as random effects for each unique combination of year and ant species included in our data. 
+While we do not have data on how the ant species actually fluctuate annually, we are able to measure how the effects of the ant species on plants fluctuate annually.
+Each of these 500 approximating matricies have the same draw from the posterior distribution, but are assigned the year-ant-specific effects of a random year of our data (year-ant-specific effects are sampled randomly with replacement).
+We chose to use 500 year-specific values, because our $\lambda_{S}$ calculations stablilized after 500 year-specific values. 
+To calculate a single $\lambda_{S}$ for these 500 year-ant-specific estimates, we take the geometric mean of the \ali{500 dominant eigenvalues}{OK, I am not positive this is correct, but I have forgotten how to correctly describe it.
+Can we talk through this sentence?} for a single stochastic $\lambda_{S}$.
+This process is repeated for each of the 1,000 random posterior draws, leading to a distribution of 1,000 stochastic $\lambda_{S}$ values. 
+To calculate the stochastic null bayesian IPM we use the exact same process, but the year-specific values used to calculate the stochastic null $\lambda_{SN}$ are not ant specific, forcing all ant species to have the same effects on their cactus partners in response to environmental variation.
+
+In the both versions of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
 opuntiae} and vacancy; other and vacancy; \textit{L.
@@
 opuntiae}, other, and vacancy; and all ant partners and vacancy.
-In order to calculate stochastic $\lambda$ distributions for each ant scenario, every one of the 1,000 parameter-iteration associated $\lambda$s actually comes from the mean of 5,000 year-specific estimations each associated with a year random effech which was randomly selected from the 18 years of data.
+Each partner scenario includes vacancy in addition to the combination of partners being analyzed for added realism because we have never observed a population of cacti in which all are occupied.
 
-We compare the distributions of $\lambda$ across each combination of ant partners to whether sampling effect or complementarity is at play in the system.
-To compare the distributions of two hypothetical populations, A and B, we subtract the vector of 1,000 $\lambda_A$ estimations of one population from the 1,000 $\lambda_B$ estimations of the other ($\lambda_A - \lambda_B$).
-If the average of these differences is positive, population A has a higher $\lambda$ than B, if the average is 0, they are equal, and if the average is negative, population B has a higher $\lambda$ than A.
+We can compare the distributions of $\lambda_{S}$ in many ways, but have chosen two moments in particular: the means and the proportion of each distribution that is greater than another.
+To calculate this second moment, we can subtract each $\lambda_{S}$ distribution from the others (\ali{e.g.}{This notation may need to be revised somewhat} $\lambda_{S,Vacant}$ - $\lambda_{S,C.
+opuntiae \& Vacant}$) to determine which partner scenario leads to a higher fitness for the cactus population.
+The proportion of the resulting vector that is greater than zero allows us to give a probabilistic estimate of which combination results in the greatest fitness. 
+The magnitude of this proportion is how we determine the magnitude of difference between the resulting population fitnesses.
 
-To determine if sampling effect is at play in a system, we must first determine if there is a "best" partner, by determining which single ant association is correlated with the highest $\lambda$ estimation.
-We can do this by comparing the distributions of each $\lambda$ and finding the one which is larger than all others.
-Then we must show that the $\lambda$ of a population with all possible partners is equal to that of $\lambda$ of a population with only the "best" partner, by comparing the distributions as described earlier.
-If it is significantly larger then complementarity is potentially at play (more on this in the next paragraph).
-If it is significantly smaller, this indicates that rather than positive effects of partner diversity, there are actually important costs of partner diversity that dampen the population growth rate.
+Using only the $\lambda_S$ distributions from the stochastic IPM we can determine which scenario leads to the highest fitness, which will tell us if there are benefits to diversity in this system, any other finding indicates that there is quantitative evidence that diversity benefits this system.
+If the partner scenario which leads to the highest possible population fitness is vacancy, this indicates that there are no benefits to partner diversity.
+If all of the highest scenarios include a common ant partner, this indicates that there is a single best partner. 
+This quantitative finding that one particular partner leads to all highest possible population fitnesses is evidence that sampling effect explains the benefits of partner diversity.
+If the highest scenario includes all possible ant partners, this indicates there is a synergy which occurs when all partners interact.
+This quantitative finding indicates that complementarity is the mechanism which explains the benefits of partner diversity in this system.
 
-To determine if complementarity is at play in a system, we must determine if the partner scenario which leads to the highest $\lambda$ is the most partner diverse scenario.
-We can do this by comparing the distribution of $\lambda$ of a population with all possible partners to the distribution of $\lambda$ of all other populations.
-If the $\lambda$ of a population with all possible partners is the largest, it indicates that complementarity is at play.
-If neither the conditions for sampling effect or complimentarity, it indicates that rather than positive effects of partner diversity, there are actuallly costs of diversity.
+Using the $\lambda_S$ distribution and the $\lambda_{SN}$ distribution, from both the stochastic IPM and the stochastic null IPM, we can determine if having multiple partners can buffer the cactus population from environmental variation.
+The difference in the full partner scenario and vacancy ($\lambda_{S,All}$ - $\lambda_{S,Vacant}$) can be interpreted as the effect of having all partners when each ant species has unique effects on the cacti in response to annual variation.
+The difference in the full partner scenario and vacancy ($\lambda_{SN,All}$ - $\lambda_{SN,Vacant}$) can be interpreted as the effect of having all partners when each ant species has the same effects on the cacti in response to annual variation.
+The proportion of the resulting vectors that are greater than 0 allows us to give  probabilistic estimates of the effects of partner diversity on the fitness of the cacti in a varying environment under two ant scenarios.
+These two distributions can again be subtracted ($\lambda_{S} - \lambda_{SN}$), allowing us to determine the difference in effects of partner diversity in two annually varying scenarios.
+The proportion of the resulting vector that is greater than 0 gives us a probablistic estimate of how strongly partner diversity buffers the population from environmental variation when the effects of each partner fluctuate uniquely.
+This quantitative finding that the effects of partner diversity in the stochastic IPM are greater than those in the stochastic null IPM indicate that portfolio effect can explain some benefits of partner diversity in this system.
 
-To determine if portfolio effect is at play we must show that partner diversity buffers the population from environmental variation.
-To do this we parameterized a new set of bayesian statistical models in which year random effects were not ant specific, meaning the effects of ant partners were not able to covary with year effects (the scenario in which all ant partners must respond to annual variation the same way and portfolio effect is not possible).
-The new estimations of year random effects are $u \sim N(0,\sigma_{yr}^2)$ with year specific variance of $\sigma_{yr}$. 
-With this verision of the IPM (now referred to as the stochastic null IPM) we followed the same approach to calculate distributions of stochastif $\lambda$s for each ant scenario.
-We can estimate the effects of partner diversity on the focal mutualist by calculating the difference in $\lambda$ distributions between populations with no partners and populations with all partners present.
-Then we can compare the difference in the effects of partner diversity on the focal mutualist calculated from the stochastic and stochastic null IPMs respectively.
-If the stochastic difference is greater than the stochastic null difference, this indicates that  partner diversity is more beneficial under varying environments and therefore portfolio effect is at play.
+\ali{}{I focused on several things throughout this rewrite of the methods: I tried to go straight into talking about the stochastic IPM, I focused on explaining rationale for specific numbers better, I tried to update notation to be clearer, and I tried to explain more clearly...
+not sure if I've accomplished these, let me know}
 
 \section*{Results}
-\subsection*{IPM Analysis}
-%%%% Start with stochastic model and comparing means to vacancy
-% means
-Using the results of the stochastic IPM, we determined if complementarity or sampling effect was at play within this system. 
-We found that populations of tree cholla which have no partners have a mean fitness of $\lambda = 0.9861$ \ref{}.
-Populations with only one ant partner present (\textit{L.
-apiculatum}, \textit{C.
-opuntiae}, or other ants) have a mean fitness of $\lambda = 0.9944$, $\lambda = 0.9893$, and $\lambda =0.9896$ respectively.
-Populations with two ant partners present (\textit{L.apiculatum} and \textit{C.
-opuntiae}, \textit{L.apiculatum} and other, or \textit{C.
-opuntiae} and other) have a mean fitness of $\lambda = 0.9942$, $\lambda = 0.9943$, and $\lambda = 0.9913$ respectively.
-Populations with all ant partners present have a mean fitness of $\lambda = 0.9940$.
-The only partner scenarios where $\lambda > 0.99$ are when \textit{L. apiculatum} ants are present with any other ant partner. 
-The means are not the only important result.
-We used bayesian modeling, so each of these $\lambda$ estimations has a distribution rather than a single estimate.
-By subtracting one of these distributions from the other, the difference from 0 tells us how much larger or smaller the distributions are with a percent confidence. 
-%% All to vacant
-% single partners confidence
-%We are 82\%, 100\%, and 84\% confident that having \textit{C.
-opuntiae} ants, other ants, or \textit{L.
-apiculatum} ants respectively as the only partner leads to a higher $\lambda$ than having no partners.
-%\textit{L. apiculatum} ants are the best single partner for tree cholla populations, as they lead to the highest population fitness with a single partner.
-% two partners confidence
-%We are 96\%, 98\%, and 92\% confident that having two ant partners present (\textit{L.apiculatum} and \textit{C.
-opuntiae}, \textit{L.apiculatum} and other, or \textit{C.
-opuntiae} and other) respecitvely leads to a higher $\lambda$ than having no ant partners.
-% all partners confidence
-%We are 94\% confident that having all ant partners present leads to a higher $\lambda$ than having no ant partners present. 
-Any ant partner is shown to lead to an increase in fitness of the tree cholla in these results. 
-We are between 82\% to 100\% confident that having any ant parnter leads to a higher $\lambda$ than having no partners.
+\subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
+\subsubsection*{The demographic effects of alternative ant partners}
+\paragraph{Growth Model} We found evidence that ant visitation enhances cactus growth and that partner identity can influence the growth trajectory of cholla (Figure \ref{fig:Grow}). 
+Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes (Figure \ref{fig:Grow}e).
+Plants with \textit{C.
+opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants (Figure \ref{fig:Grow}a).
+Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes (Figure \ref{fig:Grow}b). 
+Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates (Figure \ref{fig:Grow}c.)
+Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates (Figure \ref{fig:Grow}d).
 
-\begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/LambdaMeans.png}
-\end{figure}
-
-%% Liom to all
-%We are 100\%, 88\%, 90\%, and 76\% certain that having only \textit{L.
-apiculatum} ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
-opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
-opuntiae} and other ants as partners respecitvely.
-%We are 96\%, 92\%, 92\%, and 84\% confident that having \textit{L.
-apiculatum} and \textit{C.
-opuntiae} ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
-opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
-opuntiae} and other ants as partners respecitvely.
-%We are 98\%, 92\%, 96\%, and 92\% confident that having \textit{L.
-apiculatum} and other ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
-opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
-opuntiae} and other ants as partners respecitvely.
-%We are 94\%, 84\%, 94\%, and 84\% confident that having only \textit{L.
-apiculatum}, \textit{C.
-opuntiae}, and other ants as partners leads to a higher $\lambda$ than having no partners, only having \textit{C.
-opuntiae} ants as partners, only having other ants as partners, and having \textit{C.
-opuntiae} and other ants as partners respecitvely.
-We are between 84\% and 100\% confident that any diversity scenario where \textit{L.
-apiculatum} ants are included as partners leads to a higher $\lambda$ for the tree cholla than any diversity scenario without these parnters.
-These results indicate that sampling effect, not complementarity, is at play in this system.
-
-%%%% Stochastic null follows similar patterns but dampened
-Using the stochastic null IPM results, we found that when ant partners responded the same way to inter-annual variation similar patterns were found, though the magnitude of the patterns were somewhat different.
-We used the results of the stochastic null IPM to determine if portfolio effect is at play. 
-We found that both the stochastic model and stochastic null model resulted in higher fitness when all ants are present compared to no partners 94\% of the time. 
-These results differed in the magnitude of fitness boost recieved from partners.
-In the stochastic IPM, tree cholla with all ants present resulted in a $\lambda$ that was 0.008 greater than tree cholla with no ants present; in the stochsatic null IPM, tree cholla with all ants rpesent resulted in a $\lambda$ that was 0.013 greater than tree cholla with no ants present.
-We are 52\% confidence that the difference is greater when inter-annual effects were not ant dependent (in the stochastic null model).
-This indicates that portfolio effect is at play, if only weakly evident.
-
-\subsection*{Statistical Modeling}
-
-\paragraph{Growth Model}
-% Growth Model
-\begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/GrowContourLinesColor.png}
-	\caption{asdasdf}
-	\label{fig:growth}
-\end{figure}
-Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes \ref{fig:growth}e.
-The mean growth rate expected for plants that are 10 $cm^3$ or smaller in year t is 10.01 $cm^3/yr$, 4.65 $cm^3/yr$, 7.44 $cm^3/yr$, and 5.14 $cm^3/yr$ when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-The mean growth rate expected for plants that are between 10 $cm^3$ and 150 $cm^3$ in year t is 2.81 $cm^3/yr$, 2.11 $cm^3/yr$, 2.49 $cm^3/yr$, and 1.75 $cm^3/yr$ when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-The mean growth rate expected for plants that are larger than 150 $cm^3$ in year t is 1.28 $cm^3/yr$, 1.21 $cm^3/yr$, 1.29 $cm^3/yr$, and 1.10 $cm^3/yr$ when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-Plants with \textit{C.
-opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants \ref{fig:growth}a.
-Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes \ref{fig:growth}b. 
-Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates\ref{fig:growth}c.
-Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates \ref{fig:growth}d.
-
-We are 89\%, 88\%, and 70\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean growth rates across sizes than plants tended by no parnters, \textit{L.
-spiculatum} ants, or other ants respectively.
+Using the method of subtracting one size distribution from another we were able to determine how confident we are that the size of cholla visited by one combination of ant partners is greater than another.
+ $G_{Crematogaster} - G_{Vacant}$ returns a vector which is 89\% positive, meaning we are 89\% confident that when \textit{C.
+opuntiae} ants are partners the size of the cholla is greater than the size of the cholla with no ant partners.
+Given the evidence that plants visited by \textit{C.
+opuntiae} experience the highest mean growth rates we have reported our confidence that this partner is associated with the highest growth rates below.
+We are 88\% and 70\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean growth rates across sizes than plants tended by \textit{L.
+spiculatum} ants or other ants respectively.
 We are 89\%, 65\%, and 94\% confident that plants with no partners experience lower mean growth rates across all sizes than plants tended by \textit{C.
@@
 
+\begin{figure}
+	\includegraphics[width = 0.95\linewidth]{Figures/GrowContourLinesColor.png}
+	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
+The solid colored lines (seen in all panels) are the next mean predicted size of cholla.
+The points (seen in panels a-d) are the observed data which informs these estimates.
+The black countour lines (seen in a-d) appear at 5\% increments showing where 5\%, 10\%, etc.
+of the data is expected to fall.
+They grey dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).
+}
+	\label{fig:Grow}
+\end{figure}
+
 \paragraph{Survival Model}
-% Survival Model
+We found evidence that ant visitation enhances the survival of medium to large plants, and that partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
+Tree cholla experience between 7.7\% and 99.9\% survival rates depending on 
+their size and ant partner (Figure \ref{fig:Surv}e).
+Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
+Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes (Figure \ref{fig:Surv}a).
+Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes (Figure \ref{fig:Surv}b). 
+Plants with other ants experience the second lowest mean survival rates across all sizes (Figure \ref{fig:Surv}c).
+Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates (Figure \ref{fig:Surv}d).
+
+Using the method of subtracting one survival distribution from another we were able to determine how confident we are that the survival of cholla visited by one combination of ant partners is more likely than another.
+$S_{Crematogaster} - S_{Vacant}$ returns a vector which is 82\% positive, meaning we are 82\% confident that when \textit{C.
+opuntiae} ants are partners the survival of the cholla is more likely than the survival of the cholla with no ant partners.
+Given the evidence that plants visited by \textit{C.
+opuntiae} experience the highest mean survival rates we have reported our confidence that this partner is associated with the highest survival rates below.
+We are 63\% and 100\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean survival rates across all sizes than plants tended by \textit{L.
+spiculatum} ants or other ants respectively.
+We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by
+
 \begin{figure}[H]
 \includegraphics[width=0.95\linewidth]{Figures/SurvivalPlot.png}
-	\caption{asdasdf}
-	\label{fig:surv}
+\caption{This figure shows the estimated survival rates based on the size of the cactus with each individual ant partner.
+The solid colored lines (shown on all panels) indicate the mean estimated survival rates.
+The dashed lines (shown in panel e) indicate extrapolations beyond existing data (where we estimated survival for plants tended by ants where we had never seen a tended cactus of that size).
+The grey area around the solid lines (shown in panels a-d) show the 90\% confidence interval for the estimates.
+The colored dots are the real data binned by size to show how our estimates align with real survival observations.
+A larger circle means we had more data on survival of plants of this size with this partner.}
+\label{fig:Surv}
 \end{figure}
-Tree cholla experience between 7.7\% and 99.9\% survival rate depending on their size and ant partner \ref{fig:surv}e.
-Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
-The mean survival rate expected for plants that are 10 $cm^3$ or smaller in year t is 48\%, 37\%, 37\%, and 47\% when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-The mean survival rate expected for plants that are between 10 $cm^3$ and 150 $cm^3$ in year t is 90\%, 87\%, 84\%, and 80\% when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-The mean survival rate expected for plants that are larger than 150 $cm^3$ in year t is 99\%, 99\%, 98\%, and 95\% when tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, other ants, or no ants respectively.
-Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes \ref{fig:surv}a.
-Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes \ref{fig:growth}b. 
-Plants with other ants experience the second lowest mean survival rates across all sizes \ref{fig:growth}c.
-Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates \ref{fig:growth}d.
 
-We are 82\%, 63\%, and 100\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean survival rates across all sizes than plants tended by no parnters, \textit{L.
-spiculatum} ants, or other ants respectively.
-We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, or other ants respectively.
+\paragraph{Viability Model}
+We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability.
+Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers (Figure \ref{fig:viab}).
+The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
+apiculatum} tended plants experiencing the highest mean viability rate (at 86\%, Figure \ref{fig:Viab}b), followed by other tended plants (at 75\%, Figure \ref{fig:Viab}c), \textit{C.
+opuntiae} tended plants (at 74\%, Figure \ref{fig:Viab}a) and vacant plants (at 71\%, Figure \ref{fig:Viab}d).
 
-%% Flowers Model
-\paragraph{Flowering Model}
-There is a clear size effect on the number of flowers produced. 
-The mean number of flowers produced by a plant remains at 0 until the plant reaches medium sizes after which the mean number of flower produced increases eponentially to about 40 flowers per plant per year at large sizes.
-The mean number of flowers produced for a plant that is 150 $cm^3$ or smaller in year $t$ is <1 flower per plant.
-The mean number of flowers produced for a plant that is larger than 150 $cm^3$ in year $t$ is 8.7 flowers per plant.
+Using the method of subtracting one viability distribution from another we were able to determine how confident we are that the floral viability of cholla visited by one combination of ant partners is greater than another.
+$V_{Liomatopum} - V_{Vacant}$ returns a vector which is 99\% positive, meaning we are 99\% confident that when \textit{L.
+apiculatum} ants are partners the floral viability of the cholla is greater than the cholla with no ant partners.
+Given the evidence that plants visited by \textit{L.
+apiculatum} experience the highest viability rates we have reported our confidence that this partner is associated with the highest viability rates below.
+We are 98\% and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by \textit{C. opuntiae} or other ants respectively.
+We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
 
-% Viability Model
-\paragraph{Viability Model}
 \begin{figure}[H]
 	\includegraphics[width=0.95\linewidth]{Figures/ViabHist.png}
-	\caption{asdasdf}
-	\label{fig:viab}
+	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
+The solid lines indicate the estimated viability distribution.
+The colored histograms represent the observed viability rates of plants with that partner.
+}
+	\label{fig:Viab}
 \end{figure}
-Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers \ref{fig:viab}.
-The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
-apiculatum} tended plants experiencing the highest mean viability rate (at 86\%), followed by other tended plants (at 75\%), \textit{C.
-opuntiae} tended plants (at 74\%) and vacant plants (at 71\%).
-We are 99\%, 98\%, and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by no partners, \textit{C. opuntiae}, or other ants respectively.
-We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
 
-%% Reproduction Model
-\paragraph{Reproduction Model}
-The probability of a plant reproducing in a given year is highly size dependent. 
-The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+\ali{}{Should I include other model results here or just indicate that the rest are reported in an appendix/supplementary materials?}
 
-%% Seeds Produced
-\paragraph{Seeds Per Flower Model}
-Each viable flower on a plant produces between 97 and 257 seeds.
-This number is affected by the ant partner present. 
-\textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
-\textit{L. apiculatum} tended plants produce a mean of 149 seeds per flower. 
-Vacant plants produce a mean of 160 seeds per flower. 
-We are 73\% and 78\% confident that vacant plants produce more seeds per flower on average than plants tended by \textit{C. opuntiae} and \textit{L. apiculatum} ants respectively.
+\subsubsection*{The role of annual fluctuations in demographic effects of ant partners}
+We found evidence that annual variation impacts the effect of ant partner on various demographic traits (Figure \ref{fig:Annual_Ant}).
+The effects of each partner on demographic measure vary uniquely across temporal fluctuations, making partner identity important to track in conjunction with annual variation. 
+Specifically, where some ants offered greater than average benefits in one year, other species offered reduced benefits in the same year, indicating that each ant partner reacts differently to the fluctuating environment.
 
-%% Precensus Survival\
-\paragraph{Precensus Survival Model}
-Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.
+Each demographic trait was not affected equally by the intersection of ant partner and annual fluctuations.
+We looked at the mean effect of each ant partner on growth (figure \ref{fig:Annual_Ant}a), survival (Figure \ref{fig:Annual_Ant}b), and viability (Figure \ref{fig:Annual_Ant}c) rates across all the years included in our study.
+Where the mean effect is exactly 0 there is missing ant data due to variation in censusing focus.
+This is an indirect way to analyze how temporal fluctuation impacts our system without attributing the effects to a specific climate variables. 
+Where the mean effect is positive, ant presence increases the respective estimated growth, survival, or viability rates in response to environmental fluctuations, and where the mean effect is negative, the opposite occurs.
+We found the general magnitude of variation was the smallest for the growth rate, meaning annual variation affected the growth of plants the least, followed by survival then viability.
 
-%% Germination
-\paragraph{Germination Model}
-Seeds have a significantly higher probability of germinating in year one than in year two.
-Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
-Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/year_ant_timeseries.png}
+	\caption{This figure shows the mean affect of each ant partner on a) the estimated next size, b) the estimated survival, and c) the floral viability of cacti across every year of our study.
+These values are estimated from the fitted random effects of ant and year in our models.
+Each point represents the mean of the random effect of the identified model, ant, and year (e.g.
+the lowest dot in panel b) represents the mean effect of vacancy on survival rates in year 2011).}
+	\label{fig:Annual_Ant}
+\end{figure}
 
-%% Recruit size distribution
-New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.
+\subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
+We found that there is a high frequency of partner turnover observed in this system with very distinct directional patterns. 
+Small plants are almost always vacant (Figure \ref{fig:Ant_Transition}b-d), if they were previously tended by a partner, they are likely to become vacant again (Figure \ref{fig:Ant_Transition}a).
+As the size of the plants increases, the probability of becoming tended increases as well, though it is not equally likely to be tended by all partners.
 
-%% Ant Transition Rates
-\paragraph{Ant Transition Model}
+\textit{L. apiculatum} ants become the most likely next partner in the case of most large plants, with the exception of a large plant which was previously tended by \textit{C. opuntiae}.
+Plants which were previously tended by \textit{C. opuntiae} ants are most likely to remain tended by \textit{C. opuntiae} ants.
+This indicates that these species may follow the well documented retention-discovery trade off exhibited  by many ants, with \textit{L.
+apiculatum} ants excelling at discovery and colonization of new plants and \textit{C.
+opuntiae} ants excelling at plant retention season after season.
+The overall frequency of these partners, however is a potential alternative explanation to this discovery-retention trade off.
+\textit{L.
+apiculatum} ants are by far the most frequent ant partners, accounting for 75\% of the occupied cacti, while \textit{C.
+opuntiae} and other ants account for 17\% and 8\% of occupied cacti respectively.
+This high frequency of \textit{L. apiculatum} ants may lead to the inflation of turnover to \textit{L. apiculatum} from other ants.
+
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+	\caption{This figure shows the probability of being tended by each ant partner or vacant based on the size of the plant.
+Each panel shows these probabilities for a different previous ant state.
+The solid lines represent the mean probability of being tended by a specific partner.
+The colored points are the real data binned by size to show how our estimates align with real visitation observations.
+A larger circle means we had more data on visitation of plants of this size with this previous partner.}
+	\label{fig:Ant_Transition}
+\end{figure}
+
+\subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
+We found that partner diversity was beneficial in this system. 
+The lowest mean fitness was $\lambda_{S,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
+By subtracting the distributions $\lambda_{S, any} - \lambda_{S,Vacant}$, we found that we are between 82\% and 100\% confident that having any partner leads to a higher population fitness than having no partner.
+If you consider just the number of partners (ignoring the identity), what you find is actually that the more partners are present, the higher the fitness of the cacti (Figure \ref{fig:LambdaMeans}b), though this increase is primarily driven by the presence of \textit{L.
+apiculatum}.
+
 \begin{figure}
-	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+	\includegraphics[width=0.61\linewidth]{Figures/LambdaMeans.png}
+	\includegraphics[width=0.39\linewidth]{Figures/Lambda_Num_Partners.png}
+	\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) and $\lambda_{SN}$ (empty circles with an X) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+	\label{fig:LambdaMeans}
 \end{figure}
-All small plants are most likely to be vacant, while large plants are much more likely to be occuplied by one ant or another. 
-\textit{L.
-apiculatum} ants are the most likely tender in year $t+1$ for all large plants that were not previously tended by \textit{C.
-opuntiae} ants in year $t$, while \textit{C.
-opuntiae ants} tended plants in year $t$ are most likely to remain tended by \textit{C.
-opuntiae} ants in year $t+1$.
-We are 93\%, 89\%, 86\%, and 93\% confident that plants that are smaller than 150 $cm^3$ in year $t$ will be vacant in year $t+1$ when they were tended by no partners, \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, or other ants respectively in year $t$.
 
-% Liom 
-We are 75\%, 100\%, and 100\% confident that plants which were previously tended by \textit{L.
-apiculatum} and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
-apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
-opuntiae}, or other ants in year $t+1$.
-We are 35\%, 100\%, and 100\% confident that plants which were previously vacant and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
-apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
-opuntiae}, or other ants in year $t+1$.
-We are 32\%, 100\%, and 100\% confident that plants which were previously tended by other ants and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{L.
-apiculatum} ants in year $t+1$ than be vacant, tended by \textit{C.
-opuntiae}, or other ants in year $t+1$.
+Despite this apparent synergy, when partner identity was considered, we found the benefits of partner diversity could be explained by Sampling Effect.
+We believe the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
+All simulated combinations of ant partners which included \textit{L. apiculatum} were nearly equal and the highest possible fitness estimated for the cholla. 
+This indicates that \textit{L. apiculatum} are the single best partner for the cholla under existing conditions. 
+Based on the definitions of Sampling Effect and Complementarity we use in this study \citep{Batstone2018}, it is clear that Sampling Effect can explain the benefits of partner diversity in the cholla system.
 
-% Crem
-We are 37\%, 100\%, and 100\% confident that plants which were previously tended by \textit{C.
-opuntiae} and larger than 150 $cm^3$ in year $t$ are more likely to be tended by \textit{C.
-opuntiae} ants in year $t+1$ than be vacant, tended by \textit{L.
-apiculatum}, or other ants in year $t+1$.
+It is possible that this, like the frequency of partner turnover to \textit{L. apiculatum}, is driven by the extreme frequency of \textit{L. apiculatum} ants in comparison to others. 
+With this in mind, we simulated the population fitness with equal probability for transitioning to any ant state.
+We found ...... from the simulations with different transition probabilities.
+ \ali{}{This feels like the most natural progression to me.}
 
-\ali{}{Or I can do this section like this:}
-% Prev Vacant
-Plants which were previously vacant will most likely remain vacant until large. 
-We expect that between 81\% and 98\% of tree cholla which are smaller than 150 $cm^3$ with no partners in year $t$ are going to be vacant in year $t+1$.
-We expect that between 11\% and 55\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
-We expect that between 6\% and 17\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
-We expect that between 2\% and 10\% of tree cholla which are larger than 150 $cm^3$ with no partners in year $t$ are going to be tended by other ants in year $t+1$.
+We found evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
+We are are 94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the stochastic and stochastic null model. 
+When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
+\ali{There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.}{I am not sure if I have explained enough here honestly.}
 
-% Prev Crem
-Plants which were previously tended by \textit{C. opuntiae} will most likely be vacant or tended by \textit{C. opuntiae} ants in the next year depending on size. 
-We expect that between 74\% and 97\% of tree cholla which are smaller than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be vacant in year $t+1$.
-We expect that between 8\% and 34\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
-We expect that between 15\% and 38\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
-We expect that between 3\% and 13\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{C. opuntiae} in year $t$ are going to be tended by other ants in year $t+1$.
-
-% Prev Liom
-Plants which were previously tended by \textit{L. apiculatum} will most likely be vacant or tended by \textit{L. apiculatum} in the next year depending on size. 
-We expect that between 64\% and 97\% of tree cholla which are smaller than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be vacant in year $t+1$.
-We expect that between 30\% and 82\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L.
-apiculatum} in year $t$ are going to be tended by \textit{L.
-apiculatum} ants in year $t+1$.
-We expect that between 4\% and 7\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
-We expect that between 1\% and 3\% of tree cholla which are larger than 150 $cm^3$ and tended by \textit{L. apiculatum} in year $t$ are going to be tended by other ants in year $t+1$.
-
-% Prev Other
-Plants which were previously tended by other ants will most likely be vacant or tended by other ants in the next year depending on size. 
-We expect that between 82\% and 99\% of tree cholla which are smaller than 150 $cm^3$ and tended byother ants in year $t$ are going to be vacant in year $t+1$.
-We expect that between 10\% and 53\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by \textit{L. apiculatum} ants in year $t+1$.
-We expect that between 4\% and 13\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by \textit{C. opuntiae} ants in year $t+1$.
-We expect that between 3\% and 15\% of tree cholla which are larger than 150 $cm^3$ and tended by other ants in year $t$ are going to be tended by other ants in year $t+1$.
+\begin{figure}
+	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
+	\caption{This figure shows the distribution of $\lambda_{S,All}-\lambda_{S,Vacant}$ in pink and $\lambda_{SN,All}-\lambda_{SN,Vacant}$ in green.
+The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
+	\label{fig:Portfolio}
+\end{figure}
 
 \section*{Discussion}
+\tom{}{I have not commented too heavily here because I would like to talk through what we want this section to achieve.
+Currently, most of this section is actually describing results, whereas the pupose is to interpret and contextualize results, and connect them to the broader literature.
+Some of your content here is actually better at describing results than you had in the Results section, because it includes that element of ``how is this connected to the question'' that was missing from the Results section.}
 The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
@@
     viability intercept \textit{L. apiculatum} $\beta_04^v$ & &
+
   \end{tabular}
+  \caption{This table includes the median estimates, the 95$\%$ confidence intervals, and the prior distribution for each parameter in each model.}
+  \label{tab:Params}
   \end{table}
@@
 \section*{Figure legends}
+
+\section*{Supplementary Materials}
+\subsection*{Herbivory Data}
+\subsection*{Model Checks}
+For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 
+First, we checked the convergence of each parameter.
+Below we show the convergence of all $\beta$ terms listed in the Statistical Modeling subsection of Methods.
+Second, we checked the posterior fit, comparing the estimated values of each model to the $y$ values of the actual data.
+We show these posterior checks below, split by ant partner where relevant.
+%% Growth Figure
+%% Survival Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/surv_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/surv_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the survival model and the b) posterior distribution of survival estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean survival distribution (black line) of the real data.}
+	\label{fig:Surv_post}
+\end{figure}
+%% Reproduction Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/repro_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/repro_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the reproduction model and the b) posterior distribution of reproductive status estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean reproductive status distribution (black line) of the real data.}
+	\label{fig:Repro_post}
+\end{figure}
+%% Flowering Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/flow_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/flow_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the number of flowers model and the b) posterior distribution of the number of flowers estimated (pink lines) compared to the mean distribution of observed flowers (black line).}
+	\label{fig:Flow_post}
+\end{figure}
+%% Viability Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/viab_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/viab_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the viability model and the b) posterior distributions of floral viability estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean floral viability distribution (black line) of the real data.}
+	\label{fig:Viab_post}
+\end{figure}
+ %% Seeds Per Fruit
+ \begin{figure}
+ 	\includegraphics[width = 0.45\linewidth]{Figures/seed_conv.png}
+ 	\includegraphics[width=0.45\linewidth]{Figures/seed_ant_post.png}
+ 	\caption{The a) posterior convergence of the parameters estimated by the seeds per fruit model and the b) posterior distributions of seeds per fruit estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = vacant) compared to the mean seeds per fruit distribution (black line) of the real data.}
+ 	\label{fig:seed_post}
+ \end{figure}
+ %% Fruit survival 
+% \begin{figure}
+%% 	\includegraphics[width = 0.45\linewidth]{Figures/                .png}
+%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
+% 	\caption{The a) posterior convergence of the parameters estimated by the fruit survival model and the b) posterior distributions of fruit survival estimates (pink lines) compared to the mean fruit survival distribution (black line) of the real data.}
+% 	\label{fig:fruit_surv_post}
+% \end{figure}
+% %% Ant Transitions
+% \begin{figure}
+%% 	\includegraphics[width = 0.45\linewidth]{Figures/             .png}
+%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
+% 	\caption{The a) posterior convergence of the parameters estimated by the ant transitions model and the b) posterior distributions of next ant partners estimates (pink lines) for each previous ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean next ant partner distribution (black line) of the real data.}
+% 	\label{fig:Transitions_post}
+% %% Germination
+ \begin{figure}
+ 	\includegraphics[width = 0.45\linewidth]{Figures/germ1_conv.png}
+ 	\includegraphics[width=0.45\linewidth]{Figures/germ2_conv.png}\\
+ 	\includegraphics[width=0.45\linewidth]{Figures/germ1_post.png} 	\includegraphics[width=0.45\linewidth]{Figures/germ2_post.png}
+ \caption{The a-b) posterior convergence of the parameters estimated by the germination from year one seedbank and germination from year two seedbank models respectively.
+The c-d) posterior distributions of floral viability estimates (pink lines) compared to the mean germination distribution (black line) of the real data for first year germinants and second year germinants respectively.}
+ 	\label{fig:Germ_post}
+ \end{figure}
+
+ %% Pre-Census Survival
+ \begin{figure}
+ 	\includegraphics[width = 0.45\linewidth]{Figures/seed_surv_conv.png}
+ 	\includegraphics[width=0.45\linewidth]{Figures/seed_surv_post.png}
+ 	\caption{The a) posterior convergence of the parameters estimated by the pre-census survival model and the b) posterior distribution of the pre-census survival estimated (pink lines) compared to the mean distribution of observed pre-census survival (black line).}
+ 	\label{fig:Pre_Surv_post}
+ \end{figure}
+ %%
+
+ \subsection*{Statistical Models -- Results}
+ Below are the results reorted of all statistical models not described in the main body of the text.
+
+ %% Reproduction Model
+ \paragraph{Reproduction Model}
+ The probability of a plant reproducing in a given year is highly size dependent. 
+ The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+
+ %% Seeds Produced
+ \paragraph{Seeds Per Flower Model}
+ Each viable flower on a plant produces between 97 and 257 seeds.
+ \tom{This number is affected by the ant partner present. 
+ 	\textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
+ 	\textit{L.
+apiculatum} tended plants produce a mean of 149 seeds per flower.
+}{These results are not consistent with Ohm and Miller, where Crem had lower seeds than Liom.
+I would check this.
+This section should also reference that paper because these are not new results.}
+ Vacant plants produce a mean of 160 seeds per flower. 
+ We are 73\% and 78\% confident that vacant plants produce more seeds per flower on average than plants tended by \textit{C. opuntiae} and \textit{L. apiculatum} ants respectively.
+
+ \tom{%% Precensus Survival\
+ 	\paragraph{Precensus Survival Model}
+ 	Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.
+
+ 	%% Germination
+ 	\paragraph{Germination Model}
+ 	Seeds have a significantly higher probability of germinating in year one than in year two.
+ 	Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
+ 	Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.
+
+ 	%% Recruit size distribution
+ 	New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.}{Move to an appendix. These results are not relevant for the questions at hand.}
 

### 2024-02-28 Manuscript/ant_cactus_AmNat.tex
Wednesday afternoon

@@
 
-\title{An Integral Projection Modeling Approach to Understanding Demographic Effects of Multispecies Mutualisms}
+\title{Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism}
+% title suggestion, trying to move away from focus on IPM
 
@@
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of context-dependence in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014}.
+Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in \tom{applied settings}{would be good if you could find another applied example to cite here} \citep{rogers2014}.
 
@@
 First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
-This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
-However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018} -- but there is no benefit of diversity \emph{per se}, only better and worse partners. 
+If partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
@@
 Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits and protection across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
 
@@
 Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
-Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
+Extrafloral nectar (EFN) -bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
 Presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
@@
 Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
+Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
@@
 Previous studies have shown that herbivory by specialized insect herbivores negatively affects plant fitness \cite{Miller2009}, and ant defense reduces herbivore damage \cite{Miller2007}. 
-Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
+Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
 These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014, Donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits through anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the herbivore populations (data on herbivory is included in supplementary documents).
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects \citep[e.g.,][]{Palmer2010}.
+To our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
 
 We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity at the individual level, and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
-We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
+We used this observational data set of ant-plant associations, informed by previous ant exclusion experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
 Specifically, we asked:
@@
 Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative segments and flower buds \citep{Ness2006,Oliveira1999}. 
-Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners. 
-Smaller cholla produce little to no EFN, so larger cholla, especially flowering individuals, are generally more highly tended \citep{Miller2014}.
+Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
+%Smaller cholla produce little to no EFN, so larger cholla, especially flowering individuals, are generally more highly tended.
 
 Tree cholla EFN is harvested by various ant species. 
-At SEV-LTER, cholla are visited primarily by two species of ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster} and \textit{Camponotus}.
+At SEV-LTER, cholla are visited primarily by two species of ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species in the genera \textit{Aphaenogaster} and \textit{Camponotus}.
 \textit{L.
@@
 opuntiae} visiting between $5\% - 20\%$ of cacti depending on the year \citep{Donald2022}.
-Between $ 30\% 80\%$ of cacti remain vacant in any given year. 
-These ants rarely co-occur on a plant, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
+Between $ 30\% - 80\%$ of cacti remain vacant in any given year. 
+Workers of different species rarely co-occur on individual plants, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
 opuntiae} to \textit{L.
-apiculatum}-tended plants, and vice versa, provoke aggressive responses by resident ants (A.
+apiculatum}-tended plants, and vice versa, provoke aggressive responses by residents (A.
 Cambpell, \textit{personal observation}).
-Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
-In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
+%Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
+In Fall, tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-There are a variety of insect herbivores and seed predators which specialize on tree cholla \citep{Mann1969}. 
-A weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
-A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
-A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
+Multiple insect herbivores and seed predators specialize on tree cholla \citep{Mann1969}. 
+The Cerambycid beetle \textit{Moneilema appressum} and a weevil (Curculionidae) of the genus \textit{Gerstaekeria} feed on vegetative and reproductive structures as adults and their larvae feed internally.
+A cactus bug, \textit{Narnia pallidicornis} (Hemiptera: Coreidae), feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
+A seed predator, \textit{Cahela ponderosella} (Lepidoptera: Pyralidae), oviposits in open flowers and larvae eat seeds in developing fruits. 
 These predators can have significant negative impacts on plant fitness of and depress population growth \citep{Miller2009}.
-There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory than plants from which ants were excluded \citep{Miller2007}.
+Prior experiments showed that ant-tended tree cholla experience less herbivory and seed predation than plants from which ants were excluded \citep{Miller2007,Ohm2014}.
 
 \subsection*{Data Collection}
-This study is based a long-term demographic data set spanning 2004 to 2023 at SEV-LTER. 
+This study is based on long-term demographic data spanning 2004 to 2023 at SEV-LTER. 
 From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
 This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
-Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots has since been censused annually through 2023 (with the exception of 2020 due to the covid shutdown). 
-For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown). 
+For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
@@
 We recorded the ant species present (or vacancy if no ants present).
-Occurrences of more than one ant species on one plant were rare (less than 5\% of observed ant species), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
+Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
 Plots were searched for new recruits each year, and these were added to the census.
-In total, the data set included 1141 unique individuals and 19 plant-year observations. 
-These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
+In total, the data set included 1141 unique individuals and \tom{19 plant-year observations}{It is 19 years but the number of plant-year observations should be the total number of rows in the data set}.
+%These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
 

### 2024-03-05 Manuscript/ant_cactus_AmNat.tex
ipm methods

@@
 Multiple insect herbivores and seed predators specialize on tree cholla \citep{Mann1969}. 
-The Cerambycid beetle \textit{Moneilema appressum} and a weevil (Curculionidae) of the genus \textit{Gerstaekeria} feed on vegetative and reproductive structures as adults and their larvae feed internally.
+The Cerambycid beetle \textit{Moneilema appressum} and a weevil (Coleoptera: Curculionidae) of the genus \textit{Gerstaekeria} feed on vegetative and reproductive structures as adults and their larvae feed internally.
 A cactus bug, \textit{Narnia pallidicornis} (Hemiptera: Coreidae), feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
@@
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We recorded reproductive effort as counts of viable and aborted flowerbuds. 
+We measured reproduction as counts of viable and aborted flowerbuds. 
 We recorded the ant species present (or vacancy if no ants present).
@@
 We used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
-Ohm et al. \citep{Ohm2014} provide data on the number of seeds per fruit for plants tended by \textit{L. apiculatum}, \textit{C. opuntiae}, or no ants (experimental exclusion). 
-Miller et al. \citep{Miller2009} provide data on seed entry to the seed bank and seedling germination and survival rates.
+Ohm et al.
+\citeyear{Ohm2014} provide data on the number of seeds per fruit for plants tended by \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, or no ants (experimental exclusion), accounting for their effects on pollinator visitation.
+Elderd and Miller \citeyear{elderd2016quantifying} provide data on seed entry to the seed bank and seedling germination and survival rates.
 
 \subsection*{Multi-state Integral Projection Model}
-Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables. 
+Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables \citep{ellnerbook}. 
 While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
-We constructed a multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with their transition dynamics modeled explictly.
+We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explictly.
 
-Given the low frequency of ant species other than \textit{L.
+Given the low frequency of ant occupancy states other than \textit{L.
 apiculatum} and \textit{C.
-opuntiae} (7.78\% of observed ants) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
+opuntiae} (\textless8\% of observation) we combined all other ants into an ``other'' category, such that our multi-state IPM included four ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
-opuntiae}, and ``other''.
-The ``other'' category was made of unidentified ant species (2.8\% of observed ants), unknown species belonging to the geni \textit{Aphenogaster} (0.4\%),  \textit{Camponotus} (0.9\%),  \textit{Tetramorium} (0.02\%), \textit{Brachymyrmer} (0.02\%), a honeypot ant (0.08\%), and \textit{Forelius pruinosus} (3.5\%).
-Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
+opuntiae}, and other.
+The ``other'' category was made of \textit{Forelius pruinosus} (3.5\% of observations), unidentified species belonging to the genera \textit{Camponotus} (0.9\%), \textit{Aphaenogaster} (0.4\%), \textit{Myrmecocystus} (0.08\%), \textit{Tetramorium} (0.02\%), \textit{Brachymyrmex} (0.02\%), and additional ants not identified to genus or species (2.8\%).
+
+Ant state is included as a predictor variable in IPM sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
 For example, ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
 Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
-In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independently of its influence on plant size. 
-Therefore, these sub-models included plant size but not ant state as predictor variables.
+In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independently of its influence on plant size, so these sub-models included plant size but not ant state as predictor variables.
 
-Following previous studies, we modeled the tree cholla life cycle using continuously size-structured plants where $n(x,a)_{t}$ gives the number of plants of size $x$  and ant state $a$ in year $t$, plus two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds.
+We modeled the tree cholla life cycle using continuously size-structured plants where $n(x,a)_{t}$ gives the number of plants of size $x$  and ant state $a$ in year $t$, plus two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds.
 Seed bank dynamics are given by:
@@
 	$$
-	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x;\pmb{\tau^P}) F(x;\pmb{\tau^F}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
+	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
 	$$
@@
 	$$
+	\label{eqn:IPM1}
 \end{linenomath*}
 
-\noindent In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
-Functions $P(x;\pmb{\tau^P})$ and $F(x;\pmb{\tau^F})$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$ but not size. 
-The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$ and year.
+\noindent %In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
+Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit \tom{($\kappa(a)$)}{This is correct but in the equation above it is shown as $\kappa(a')$ -- I am flagging that so you can update the equation and check that the code is consistent.} is dependent on ant state $a$.
+The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
 Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
@@
 	$$
+	\label{eqn:IPM2}
 \end{linenomath*}
 
-\noindent The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of seedling survival ($\omega$) from germination (late summer) to the census (May).
-This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appearance in our census ($\sum_{a'}^{A}\rho_{0}(a')=1$). 
+\noindent The first term in Eq.
+\tom{\ref{eqn:IPM2}}{We should label equations.
+I am not sure why the equation label is not working here and I did not try to figure it out.
+It is probably something with the linenomath formatting.} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of seedling survival ($\omega$) from germination (late summer) to the census (May).
+This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appearance in our census ($\sum\rho_{0}(a')=1$). 
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
-Survival from initial size $x$ ($S(x,a;\pmb{\tau^S_{a}})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau^G_{a}})$) are both dependent on initial size and ant state. 
-As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau_{a}}$). 
-Ant transition function $\rho(a',a,x;\pmb{\tau^{\rho}})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
-This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau^\rho}$).
+Survival from initial size $x$ ($S(x,a;\pmb{\tau}^S_{a})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau}^G_{a})$) are both dependent on initial size and ant state. 
+As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau}_{a}$). 
+Ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
+This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau}^\rho$).
 
 \subsection*{Statistical modeling and parameter estimation}
-We parameterized the IPM using a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework to serve as vital rate sub-models. 
+We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework to serve as vital rate sub-models. 
 Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively, and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
-In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transition between ant states conditional on plant size and previous ant state. 
+In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
 All models were fit using R version 4.2.2 and Rstan package (\cite{Rstancite, Rcite}).
-Unless otherwise mentioned, all used vague priors.
 
-The vital rate models take the form of the function:
+Vital rate models for survival, growth, probability of flowering,  take the form of the function:
 $$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
@@
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
-We fit this model to the size in year $t+1$ ($y^G$) using the location-scale parameterization of the student $t$ distribution and the identity link function, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
-$$y^G \sim Student T(\hat{\nu},\hat{G},\hat{\sigma}) $$
-Student $t$ distributions are dependent on three parameters, $\hat{\nu}$,  $\hat{G}$, and $\hat{\sigma}$.
-$\hat{G}$ is the mean of the distribution, and in our case is expressed as a second-order polynomial with ant-size interactions because  preliminary analysis found this was a better fit to slightly concave data.
-$\hat{\nu}$ is the shape parameter which determines how thin the tails of the distribution are.
-$\hat{\sigma}$ is the scale parameter which determines how wide the distribution is. 
-Both $\hat{\nu}$ and $\hat{\nu}$ are size dependent because preliminary analyses revealed that there was size dependence in both variance and kurtosis.
-%$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G}  x  + \beta_{2}^{G}  a + \beta_{3}^{G}  x  a  + \beta_{4}^{G}  x^2 + \beta_{5}^{G}  a  x^2 + u + w $$
-$$\hat{\sigma} = \beta_{0}^{\sigma} + \beta_{1}^{\sigma}  x $$
-$$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu}  x $$
+The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$. 
+We fit this model to the size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
+$$y^G \sim Skewed Normal(\hat{G},\sigma,\alpha) $$
+$$\hat{G} = \beta^G_0 + \beta^G_1 x + \beta^G_2 a + \beta^G_3 ax + \beta^G_4 x^2 + \beta^G_5 x^2a + u^G_{a} + w^G $$
+The \tom{mean}{I rewrote this for the skewed normal but I cannot remember if the location parameter is the mean, so you should double check this.} $\hat{G}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
+$u^G_{a} \sim N(0,\sigma_{year}^{2})$ is the year random effect for each ant state (which parameterizes the $\pmb{\tau}^G_{a}$ vector), and $w^G \sim N(0,\sigma_{plot}^{2})$ is the plot random effect.
+Parameters $\hat{\sigma}$ and $\hat{\alpha}$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x$.
 
 \paragraph{Survival}
-The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
-We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function
-$y^S \sim Bern(\hat{S})$.
-%$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u + w$$
+The survival sub-model ($S(a,x;\pmb{\tau}_{a}^{S})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of size $x$ and ant partner $a$ in year $t$.
+We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function, with year and plot random effects as above. 
+$$y^S \sim Bern(\hat{S})$$
+$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u^S_{a} + w^S$$
 
 \paragraph{Reproduction}
-The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
-We fit this model to the reproductive data $y^R$ using a Bernoulli distribution and a logit link function:
-$$y^P \sim Bern(\hat{P})$$
-%$$logit(\hat{P} = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w)$$
+The flowering sub-model ($P(x';\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot and year.
+We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects. 
+The flower bud function $F(x';\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$. 
+We fit this model to flowerbud count data $y^F$ (sum of viable and aborted) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects as above:
+$$y^{F} \sim NegBinom(\hat{F},\hat{\phi})$$
+$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u^F + w^F$$
 
-\paragraph{No. Flowers}
-The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
-We fit this model to flowerbud count data $y^F$ using a zero-truncated negative binomial distribution with a log transformation:
-$$y^{F} \sim 0 Truncated Negative Binom(\hat{F},\hat{\phi})$$
-%$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u + w$$
-$$log(\hat{\phi}) = \beta_{0}^{\phi}$$
+The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the \tom{proportion of flowers produced by a plant that are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$}{The way this is written here is not correct.
+Viability in year t should be a function of ants in year t.
+I would double check that this is what the model is actually doing.}.
+We fit this model to \tom{floral abortion data $y^V$}{The way this is written it's unclear whether you are modeling probability of viability or probability of abortion} using a binomial distribution with $N$ trials (total number of flower buds) and a logit link function:
+$$y^{V} \sim Binom(\hat{V},N)$$
+$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
 
-\paragraph{Flowerbud viability}
-The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates theproportion of flowers produced by a plant which are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
-We fit this model to floral abortion data $y^V$ using a Binomial distribution and a logit link function:
-$$y^{V} \sim Binom(\hat{V})$$
-%$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
-
-\paragraph{Seeds Per Fruit}
-With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
-We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
-$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
-%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
-$$\hat{\phi} = \beta_{0}^{\phi}$$
-The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
 
 \paragraph{Ant Transitions}
-The ant transition model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
-We fit this model to ant partner data from year $t+1$ $a'$ using a multinomial distribution with a logit link function: 
-$$y^{\epsilon} ~ Multinomial(\hat{\epsilon})$$
-%$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
-
-\paragraph{Recruit Size Distribution}
-%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
-We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
-$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
-%$$\hat{\eta} = \beta_{0}^{\eta}$$
-where $\hat{\sigma}$ is estimated with a non-informative prior.
-
-\paragraph{Germination}
-With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
-These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
-$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
-$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
-%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
-%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
-
-\paragraph{Pre-Census Survival}
-With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
-We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
-$$y^{\delta} ~ Bern(\hat{\delta})$$
-%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
- where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
+We fit this model to ant partner data $y^{\epsilon}$ using a multinomial distribution with a logit link function: 
+$$y^{\epsilon} \sim Multinomial(\hat{\epsilon})$$
+$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
 
 \paragraph{Parameter estimation}
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
-For each model, we obtained three chains of 10,000 iterations, each with randomly chosen initial conditions. 
-The first 1,500 iterations were discarded. 
-We did not thin the chains, thus all samples were retained following burn-in. 
-We assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b). 
-To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
-All estimated parameters are available in Table \ref{tab:Params}.
+We fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+We used vague priors for all parameters. 
+For each model, we obtained three chains of 10,000 iterations, discarding the first 1,500 iterations. 
+%We did not thin the chains, thus all samples were retained following burn-in. 
+We visually assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
+%All estimated parameters are available in Table \ref{tab:Params}.
 
 \subsection*{IPM Analysis}
-We constructed two stochastic IPMs in order to simulate ant combinations that don’t exist in real life with two different versions of environmental variation.
-The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
-The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
-By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
-By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
-In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
-
 Analyzing an IPM requires discretizing the continuous IPM kernel into an approximating matrix. 
 Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
-In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$ where $A$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
+In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$.
 Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $A(b+2) \times A(b+2)$.
 We used $b = 200$ bins after stability analysis to determine how many bins were needed to consistently estimate the same lambda.
-We extend the integration limits $L$ and $U$ (the lowest and highest size values observed) to avoid unintentional eviction \cite{Williams2012}.
+We extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
 
-Traditionally, in a bayesian IPM, rather than having one set of parameters to populate an approximating matrix and estimate population fitness ($\lambda$), we use 1,000 draws from the posterior distributions of our bayesian statistical models to populate 1,000 approximating matrix and estimate a 1,000 element distribution for $\lambda$.
-After preliminary stability analysis we found that 1,000 random posterior draws leads to the same estimate of $\lambda$ as any greater number of posterior draws.
-In a stochastic bayesian IPM, the estimation of each individual $\lambda$ in the distribution comes from a geometric mean of many values. 
-Rather than a single approximating matrix for each value of $\lambda$, we take the geometric mean of 500 approximating matricies with year-ant-specific effects, resulting in a stochastic version of populationfitness ($\lambda_{S}$).
-These year-ant-specific effects are calculated as random effects for each unique combination of year and ant species included in our data. 
-While we do not have data on how the ant species actually fluctuate annually, we are able to measure how the effects of the ant species on plants fluctuate annually.
-Each of these 500 approximating matricies have the same draw from the posterior distribution, but are assigned the year-ant-specific effects of a random year of our data (year-ant-specific effects are sampled randomly with replacement).
-We chose to use 500 year-specific values, because our $\lambda_{S}$ calculations stablilized after 500 year-specific values. 
-To calculate a single $\lambda_{S}$ for these 500 year-ant-specific estimates, we take the geometric mean of the \ali{500 dominant eigenvalues}{OK, I am not positive this is correct, but I have forgotten how to correctly describe it.
-Can we talk through this sentence?} for a single stochastic $\lambda_{S}$.
-This process is repeated for each of the 1,000 random posterior draws, leading to a distribution of 1,000 stochastic $\lambda_{S}$ values. 
-To calculate the stochastic null bayesian IPM we use the exact same process, but the year-specific values used to calculate the stochastic null $\lambda_{SN}$ are not ant specific, forcing all ant species to have the same effects on their cactus partners in response to environmental variation.
+For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
+To estimated population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{19}{Is this right?
+The number of transition matrices should be one fewer than then number of years, but I don't remember what happens with 2020.} annual transition matrices, discarding the first 100 years to minimize the influence of initial conditions.
+Sampling observed transition matrices (rather than independently sampling regression coefficients) produces models that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
+We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \cite{rees2009integral}.
+We propagated uncertainty from the vital rate models using 1000 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
+
+Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to test how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
+From the fullest version of the model corresponding to the observed assemblage of partners, we created model versions
 
@@
 Each partner scenario includes vacancy in addition to the combination of partners being analyzed for added realism because we have never observed a population of cacti in which all are occupied.
+
+We constructed two versions of the stochastic, multi-state IPMs that allowed us to explore two different scenarios of environmental variation. 
+Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation.
+
+The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
+The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
+By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
+By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
+In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
 
@@
 \section*{Appendix A: Additional Methods and Parameters}
+
+\paragraph{Seeds Per Fruit}
+With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
+$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
+%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
+$$\hat{\phi} = \beta_{0}^{\phi}$$
+The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+
+\paragraph{Recruit Size Distribution}
+%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
+We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
+$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
+%$$\hat{\eta} = \beta_{0}^{\eta}$$
+where $\hat{\sigma}$ is estimated with a non-informative prior.
+
+\paragraph{Germination}
+With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
+$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
+%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
+
+\paragraph{Pre-Census Survival}
+With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
+We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
+$$y^{\delta} ~ Bern(\hat{\delta})$$
+%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
 This is not referenced in the paper, to my knowledge, and I think you need to think more deeply about what content should go into appendices and why. 

### 2024-03-14 Manuscript/ant_cactus_AmNat_v2.tex
done editing through methods
I also created a new _v2 since the changes amount to major surgery.

--- 2024-03-05 Manuscript/ant_cactus_AmNat.tex
@@
 
-\title{Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism}
+\title{Thank you, next: demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism}
 % title suggestion, trying to move away from focus on IPM
@@
 apiculatum} and \textit{C.
-opuntiae} (\textless8\% of observation) we combined all other ants into an ``other'' category, such that our multi-state IPM included four ant states: vacant, \textit{L.
+opuntiae} (\textless8\% of observations) we combined all other ants into an ``other'' category, such that our multi-state IPM included four ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
-opuntiae}, and other.
-The ``other'' category was made of \textit{Forelius pruinosus} (3.5\% of observations), unidentified species belonging to the genera \textit{Camponotus} (0.9\%), \textit{Aphaenogaster} (0.4\%), \textit{Myrmecocystus} (0.08\%), \textit{Tetramorium} (0.02\%), \textit{Brachymyrmex} (0.02\%), and additional ants not identified to genus or species (2.8\%).
+opuntiae}, and Other.
+The ``Other'' category was made up of \textit{Forelius pruinosus} (3.5\% of observations), unidentified species belonging to the genera \textit{Camponotus} (0.9\%), \textit{Aphaenogaster} (0.4\%), \textit{Myrmecocystus} (0.08\%), \textit{Tetramorium} (0.02\%), \textit{Brachymyrmex} (0.02\%), and additional ants not identified to genus or species (2.8\%).
 
@@
 Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
-In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independently of its influence on plant size, so these sub-models included plant size but not ant state as predictor variables.
+In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering or flowerbud production independently of its influence on plant size, so these sub-models do not include ant state as a predictor variable.
 
@@
 The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit \tom{($\kappa(a)$)}{This is correct but in the equation above it is shown as $\kappa(a')$ -- I am flagging that so you can update the equation and check that the code is consistent.} is dependent on ant state $a$.
-The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
+The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
+For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability and plants in different ant states can fluctuate independently (good years for \textit{L.
+apiculatum} -occupied plants may not be good years for \textit{C.
+opuntiae}-occupied plants, for example).
 Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
@@
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
-Survival from initial size $x$ ($S(x,a;\pmb{\tau}^S_{a})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau}^G_{a})$) are both dependent on initial size and ant state. 
+Survival ($S(x,a;\pmb{\tau}^S_{a})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau}^G_{a})$) are both dependent on initial size and ant state. 
 As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau}_{a}$). 
-Ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
+Finally, ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
 This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau}^\rho$).
@@
 \subsection*{Statistical modeling and parameter estimation}
-We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework to serve as vital rate sub-models. 
-Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively, and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
+We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework. 
+Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively (only year variation is used in the IPM), and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
 In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
-All models were fit using R version 4.2.2 and Rstan package (\cite{Rstancite, Rcite}).
+As in the IPM, our statistical modeling assumed that demographic effects of ant occupancy are limited to survival, growth, and flowerbud viability.
 
-Vital rate models for survival, growth, probability of flowering,  take the form of the function:
-$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
-\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Some models deviate from this general model, which will be specified below.
+%Vital rate models for survival, growth, probability of flowering,  take the form of the function:
+%$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
+%\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+%Some models deviate from this general model, which will be specified below.
 
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$. 
-We fit this model to the size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
-$$y^G \sim Skewed Normal(\hat{G},\sigma,\alpha) $$
-$$\hat{G} = \beta^G_0 + \beta^G_1 x + \beta^G_2 a + \beta^G_3 ax + \beta^G_4 x^2 + \beta^G_5 x^2a + u^G_{a} + w^G $$
-The \tom{mean}{I rewrote this for the skewed normal but I cannot remember if the location parameter is the mean, so you should double check this.} $\hat{G}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
-$u^G_{a} \sim N(0,\sigma_{year}^{2})$ is the year random effect for each ant state (which parameterizes the $\pmb{\tau}^G_{a}$ vector), and $w^G \sim N(0,\sigma_{plot}^{2})$ is the plot random effect.
-Parameters $\hat{\sigma}$ and $\hat{\alpha}$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x$.
+We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
+$$y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) $$
+$$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
+The \tom{mean}{I rewrote this for the skewed normal but I cannot remember if the location parameter is the mean, so you should double check this.} of the $i$th observation $\hat{G_i}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
+The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vector) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
+Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
 
@@
 The survival sub-model ($S(a,x;\pmb{\tau}_{a}^{S})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of size $x$ and ant partner $a$ in year $t$.
-We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function, with year and plot random effects as above. 
-$$y^S \sim Bern(\hat{S})$$
-$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u^S_{a} + w^S$$
+We fit this model to the survival data (alive or dead) using a Bernoulli distribution and the logit link function, with a similar linear predictor as the growth model but without the second-order influence of size.
 
 \paragraph{Reproduction}
-The flowering sub-model ($P(x';\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot and year.
+The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects for the size $x$ and random effects of plot and year.
 We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects.  
-The flower bud function $F(x';\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$. 
-We fit this model to flowerbud count data $y^F$ (sum of viable and aborted) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects as above:
-$$y^{F} \sim NegBinom(\hat{F},\hat{\phi})$$
-$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u^F + w^F$$
+The flower bud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$. 
+We fit this model to flowerbud count data (sum of viable and aborted buds) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects.
 
@@
 I would double check that this is what the model is actually doing.}.
-We fit this model to \tom{floral abortion data $y^V$}{The way this is written it's unclear whether you are modeling probability of viability or probability of abortion} using a binomial distribution with $N$ trials (total number of flower buds) and a logit link function:
-$$y^{V} \sim Binom(\hat{V},N)$$
-$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
+We fit this model to \tom{floral abortion data}{The way this is written it's unclear whether you are modeling probability of viability or probability of abortion} using a binomial distribution where trials and successes are given by the total number of flower buds and the number that are viable, respectively.
+This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models.
 
-Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
+Additional reproductive parameters for the \tom{number of seeds per fruit}{We need to briefly explain that this was any-specific, where those values come from, and how we deal with Other.}, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
 
 \paragraph{Ant Transitions}
-The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
-We fit this model to ant partner data $y^{\epsilon}$ using a multinomial distribution with a logit link function: 
-$$y^{\epsilon} \sim Multinomial(\hat{\epsilon})$$
-$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$.
+We fit this model to ant partner data using a multinomial distribution with a logit link function and random effects for plot and year.
 
@@
 Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
-In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$.
-Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $A(b+2) \times A(b+2)$.
-We used $b = 200$ bins after stability analysis to determine how many bins were needed to consistently estimate the same lambda.
-We extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
+In our model there is additional complexity in the form of transitions between $A$ ant partners and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
+We used $b = 200$ bins (sufficient for numerically stable model outputs) and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
 
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimated population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{19}{Is this right?
-The number of transition matrices should be one fewer than then number of years, but I don't remember what happens with 2020.} annual transition matrices, discarding the first 100 years to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{19}{Is this right?
+The number of transition matrices should be one fewer than then number of years, but I don't remember what happens with 2020.} annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces models that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
@@
 
-Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to test how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
-From the fullest version of the model corresponding to the observed assemblage of partners, we created model versions
+\subsubsection*{Partner diversity experiments}
+Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to quantify how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
+From the fullest version of the model corresponding to the observed assemblage of partners, we created subsets corresponding to all eight possible scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
+opuntiae} only, \textit{L.
+apiculatum} only, Other only); two partners (all pairwise combinations of \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, and Other); and three partners (observed scenario of all ant states).
+These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed.
+For example, the no-partner simulation scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
 
-In the both versions of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
-opuntiae} and vacancy; other and vacancy; \textit{L.
-apiculatum}, \textit{C.
-opuntiae}, and vacancy, \textit{L.
-apiculatum}, other, and vacancy; \textit{C.
-opuntiae}, other, and vacancy; and all ant partners and vacancy.
-Each partner scenario includes vacancy in addition to the combination of partners being analyzed for added realism because we have never observed a population of cacti in which all are occupied.
+In all scenarios that included any ant partners, we preserved the observed pattern of size-dependent vacancy/occupancy (estimated through the ant transition sub-model) and manipulated partner identity conditional on occupancy.
+This means, for example, that the \textit{C. opuntiae}-only scenario included two possible states, vacancy and occupied by \textit{C. opuntiae}. 
+Small, non-reproductive plants are typically vacant because they do not produce extrafloral nectar, and once plants begin producing nectar they are nearly always ant-tended \citep{Miller2014}. 
+Our simulation experiments preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature.
 
-We constructed two versions of the stochastic, multi-state IPMs that allowed us to explore two different scenarios of environmental variation. 
-Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation.
+\subsubsection*{Temporal stochasticity experiments}
+Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation, yielding an advantage of a diverse ``portfolio'' of partners when the environment fluctuates.
+Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
+We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two different scenarios of environmental variation. 
+First, we evaluated the model using empirical estimates for the $\pmb{\tau}_{a}$ vectors that describe ant-specific year deviates. 
+In this scenario, good years and bad years can differ between ant states, according to the empirical parameter estimates. 
+\tom{We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.}{Worth doing!} 
+Second, we \tom{re-fit}{I think this is correct?
+The alternative would be to choose one ant state to define the year effects, or perhaps average them.} the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
+We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
+Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above.
 
-The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
-The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
-By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
-By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
-In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
+\subsubsection*{Statistical inference on fitness consequences of partner identity and diversity}
+The range of models we created generated many outputs; we focus our inference on the following specific contrasts. 
+First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+Second, to determine whether each partner, in isolation, confers a fitness advantage and to rank alternative partners, we contrasted the fitness of each single partner scenario (\textit{C.
+opuntiae} only, \textit{L.
+apiculatum} only, Other only) against vacancy (zero partners).
+Third, to determine whether apparent benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario.
+If the best multi-partner scenarios exceed the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
+Alternatively, a sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
+It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with opportunity costs of diversity. 
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (three-partner) scenario for synchronized and non-synchronized responses to temporal stochasticity.
+If the portfolio effect confers a benefit of diversity, $\lambda_{S}$ should be higher under non-synchronized temporal fluctuations.\tom{}{I am actually not sure whether we should look at this as stated, or whether we should compare the difference between all partners and vacancy for synchronous and non-synchronous (which is what I think you currently do.)}
 
-We can compare the distributions of $\lambda_{S}$ in many ways, but have chosen two moments in particular: the means and the proportion of each distribution that is greater than another.
-To calculate this second moment, we can subtract each $\lambda_{S}$ distribution from the others (\ali{e.g.}{This notation may need to be revised somewhat} $\lambda_{S,Vacant}$ - $\lambda_{S,C.
-opuntiae \& Vacant}$) to determine which partner scenario leads to a higher fitness for the cactus population.
-The proportion of the resulting vector that is greater than zero allows us to give a probabilistic estimate of which combination results in the greatest fitness. 
-The magnitude of this proportion is how we determine the magnitude of difference between the resulting population fitnesses.
-
-Using only the $\lambda_S$ distributions from the stochastic IPM we can determine which scenario leads to the highest fitness, which will tell us if there are benefits to diversity in this system, any other finding indicates that there is quantitative evidence that diversity benefits this system.
-If the partner scenario which leads to the highest possible population fitness is vacancy, this indicates that there are no benefits to partner diversity.
-If all of the highest scenarios include a common ant partner, this indicates that there is a single best partner. 
-This quantitative finding that one particular partner leads to all highest possible population fitnesses is evidence that sampling effect explains the benefits of partner diversity.
-If the highest scenario includes all possible ant partners, this indicates there is a synergy which occurs when all partners interact.
-This quantitative finding indicates that complementarity is the mechanism which explains the benefits of partner diversity in this system.
-
-Using the $\lambda_S$ distribution and the $\lambda_{SN}$ distribution, from both the stochastic IPM and the stochastic null IPM, we can determine if having multiple partners can buffer the cactus population from environmental variation.
-The difference in the full partner scenario and vacancy ($\lambda_{S,All}$ - $\lambda_{S,Vacant}$) can be interpreted as the effect of having all partners when each ant species has unique effects on the cacti in response to annual variation.
-The difference in the full partner scenario and vacancy ($\lambda_{SN,All}$ - $\lambda_{SN,Vacant}$) can be interpreted as the effect of having all partners when each ant species has the same effects on the cacti in response to annual variation.
-The proportion of the resulting vectors that are greater than 0 allows us to give  probabilistic estimates of the effects of partner diversity on the fitness of the cacti in a varying environment under two ant scenarios.
-These two distributions can again be subtracted ($\lambda_{S} - \lambda_{SN}$), allowing us to determine the difference in effects of partner diversity in two annually varying scenarios.
-The proportion of the resulting vector that is greater than 0 gives us a probablistic estimate of how strongly partner diversity buffers the population from environmental variation when the effects of each partner fluctuate uniquely.
-This quantitative finding that the effects of partner diversity in the stochastic IPM are greater than those in the stochastic null IPM indicate that portfolio effect can explain some benefits of partner diversity in this system.
-
-\ali{}{I focused on several things throughout this rewrite of the methods: I tried to go straight into talking about the stochastic IPM, I focused on explaining rationale for specific numbers better, I tried to update notation to be clearer, and I tried to explain more clearly...
-not sure if I've accomplished these, let me know}
+We base our statistical inferences on the posterior probability distributions of the contrasts described above. 
+For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{S}$ ($\Delta\lambda_{C-V}$). 
+We can quantify from this distribution our certainty in the effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
+We apply similar logic to other contrasts described above.
 
@@
 
+Discussion brainstorming:
+
+1. Mini abstract paragraph
+- synthesis of question plus answers, and overall significance
+Topic sentence: Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. Here we show...
+
+2. COnnecting to broader patterns in the literature
+- address how complimentarity or synergies are commonly reported -- why did we not see this, especially since the vital rates suggest some complimentarity
+- may involve the simulation experiments
+- some contrast with other systems where benefits of diversity have been found
+Topic sentence: similar studies have reported complementarity while we found sampling effect explains the benefits of diversity in our system.
+This may have to do with the sheer number of "liom" in the system.
+
+3. If there is one best partner then diversity should impose opportunity costs -- why do we not see this?
+- should address how some systems can host multiple partners at a time while others (ours) can exclusive associations; for the latter diversity should only be costly if there is one best partner?
+
+4. Turnover
+- what's going on with turnover? what processes may drive the patters we see
+- what is seen elsewhere in the lit (invertase papers may be an interesting reference? showing plant ``choice'')
+- patterns could be consistent with ant and or plant choice (currently a ``black box'')
+- this might address both frequency and direction (direction may lead to some discussion of competition)
+
+5.  Portfolio effect
+- what is it, what does it predict, what did we find
+- what have other studies found? 
+Topic sentence: Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits. Yet, we find...
+
+6. Value of long-term data
+- includes turnover, portfolio effect
+- also address how the full demographic data provides the correct insight into ant effects (eg We thought Crem was a parasite based on less complete data)
+
+7. Herbivory
+- this should underlie why these interactions are beneficial, so we need to address this
+- this connects to portfolio effect, because fluctuations in herbivores (in conjunction with ant-specific defenses against herbivores) would be why porfolio effect could occur
+
+8. Limitations
+- confounds fundamental to observational data regarding ant effects on plant demography (correlation/causation problem) -- maybe plants that attract Liom through nectar have a different pattern of allocation o growth and survival
+- cite experiments to support causal interpretation of observational patterns (ants really do stuff)
+- extrapolating effects of ants beyond the interactions that we can actually see (e.g., new recruits with Liom); and the counterfactual of a Liom-only population is another form of extrapolation
+
+9. Future directions -- maybe embed these into paragraphs above where we see opportunity to address limitations or extend the results -- probably lots we can do with herbivory
+
+Conclusion.
+
+Appendices:
+1. Herbivory
+2. Additional methods and results
+3. Posterior predictive checks and model convergence
+4. Maybe ant transition simulations
+
 \end{document}

### 2024-03-15 Manuscript/ant_cactus_AmNat_v2.tex
edits through end of Results

@@
 Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
+We assume growth variance and skewness were not dependent on ant occupancy state.
 
@@
 apiculatum}, and Other); and three partners (observed scenario of all ant states).
-These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed.
-For example, the no-partner simulation scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
+These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely or never observed.
+For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
 
@@
 Third, to determine whether apparent benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario.
-If the best multi-partner scenarios exceed the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
-Alternatively, a sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
+If the best multi-partner scenario exceeds the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
+Alternatively, the sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with opportunity costs of diversity. 
@@
 For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{S}$ ($\Delta\lambda_{C-V}$). 
-We can quantify from this distribution our certainty in the effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
+We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
 We apply similar logic to other contrasts described above.
@@
 \subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
-\subsubsection*{The demographic effects of alternative ant partners}
-\paragraph{Growth Model} We found evidence that ant visitation enhances cactus growth and that partner identity can influence the growth trajectory of cholla (Figure \ref{fig:Grow}). 
-Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes (Figure \ref{fig:Grow}e).
-Plants with \textit{C.
-opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants (Figure \ref{fig:Grow}a).
-Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes (Figure \ref{fig:Grow}b). 
-Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates (Figure \ref{fig:Grow}c.)
-Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates (Figure \ref{fig:Grow}d).
-
-Using the method of subtracting one size distribution from another we were able to determine how confident we are that the size of cholla visited by one combination of ant partners is greater than another.
- $G_{Crematogaster} - G_{Vacant}$ returns a vector which is 89\% positive, meaning we are 89\% confident that when \textit{C.
-opuntiae} ants are partners the size of the cholla is greater than the size of the cholla with no ant partners.
-Given the evidence that plants visited by \textit{C.
-opuntiae} experience the highest mean growth rates we have reported our confidence that this partner is associated with the highest growth rates below.
-We are 88\% and 70\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean growth rates across sizes than plants tended by \textit{L.
-spiculatum} ants or other ants respectively.
-We are 89\%, 65\%, and 94\% confident that plants with no partners experience lower mean growth rates across all sizes than plants tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, or other ants respectively.
+Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demoghraphic effects across host vital rates. 
+\tom{Plants tended by \textit{C.
+opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
+apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).}{Your draft had posterior probabilities associated with growth differences.
+I think it would be possble to keep the vital rate results very simple -- like what I have here -- and keep the posterior probability estimates for the IPM results.
+I am open to including posterior estimates for the vital rates but it is not clear to me how those values account for size, since we are comparing functions not single values.
+Let's discuss.}
+For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean. 
+Similarly, ant visitation enhances cactus survival, and ant partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
+Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
+\textit{C. opuntiae}-occupied plants had a survival advantage over other ant states, particularly at smaller sizes, consistent with the positive effects on growth. 
+At larger sizes, plants in any state of ant occupancy had a survival advantage over vacant plants.
 
@@
 
-\paragraph{Survival Model}
-We found evidence that ant visitation enhances the survival of medium to large plants, and that partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
-Tree cholla experience between 7.7\% and 99.9\% survival rates depending on 
-their size and ant partner (Figure \ref{fig:Surv}e).
-Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
-Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes (Figure \ref{fig:Surv}a).
-Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes (Figure \ref{fig:Surv}b). 
-Plants with other ants experience the second lowest mean survival rates across all sizes (Figure \ref{fig:Surv}c).
-Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates (Figure \ref{fig:Surv}d).
-
-Using the method of subtracting one survival distribution from another we were able to determine how confident we are that the survival of cholla visited by one combination of ant partners is more likely than another.
-$S_{Crematogaster} - S_{Vacant}$ returns a vector which is 82\% positive, meaning we are 82\% confident that when \textit{C.
-opuntiae} ants are partners the survival of the cholla is more likely than the survival of the cholla with no ant partners.
-Given the evidence that plants visited by \textit{C.
-opuntiae} experience the highest mean survival rates we have reported our confidence that this partner is associated with the highest survival rates below.
-We are 63\% and 100\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean survival rates across all sizes than plants tended by \textit{L.
-spiculatum} ants or other ants respectively.
-We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by
-
 \begin{figure}[H]
@@
 
-\paragraph{Viability Model}
-We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability.
-Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers (Figure \ref{fig:viab}).
-The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
-apiculatum} tended plants experiencing the highest mean viability rate (at 86\%, Figure \ref{fig:Viab}b), followed by other tended plants (at 75\%, Figure \ref{fig:Viab}c), \textit{C.
-opuntiae} tended plants (at 74\%, Figure \ref{fig:Viab}a) and vacant plants (at 71\%, Figure \ref{fig:Viab}d).
-
-Using the method of subtracting one viability distribution from another we were able to determine how confident we are that the floral viability of cholla visited by one combination of ant partners is greater than another.
-$V_{Liomatopum} - V_{Vacant}$ returns a vector which is 99\% positive, meaning we are 99\% confident that when \textit{L.
-apiculatum} ants are partners the floral viability of the cholla is greater than the cholla with no ant partners.
-Given the evidence that plants visited by \textit{L.
-apiculatum} experience the highest viability rates we have reported our confidence that this partner is associated with the highest viability rates below.
-We are 98\% and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by \textit{C. opuntiae} or other ants respectively.
+\tom{We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.}{Great example of an effective topic sentence for a Results paragraph.
+Sets the stage with the main result, and leaves the details for below.}
+Reproductive cacti experience between \tom{39\% and 96\%}{Unclear what this range indicates. The min to max would be 0 to 100\%.} viability rates of flowers (Figure \ref{fig:viab}).
+Ant partners influence the mean viability rate of flowers, with \textit{L.
+apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by \textit{C.
+opuntiae} and Other tended plants (at 74-75\%, Figure \ref{fig:Viab}b,c), and vacant plants had the lowest floral viability rate (71\%, Figure \ref{fig:Viab}d).
+Given the uncertainty of the parameter estimates, we are 98\% and 97\% confident that \textit{L.
+apiculatum} tended plants experience higher viability rates than plants tended by \textit{C.
+opuntiae} or other ants respectively.
 We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
@@
 \ali{}{Should I include other model results here or just indicate that the rest are reported in an appendix/supplementary materials?}
+\tom{}{I think you can include a short paragraph that briefly mentions that other vital rates are in the appendix.}
 
-\subsubsection*{The role of annual fluctuations in demographic effects of ant partners}
-We found evidence that annual variation impacts the effect of ant partner on various demographic traits (Figure \ref{fig:Annual_Ant}).
-The effects of each partner on demographic measure vary uniquely across temporal fluctuations, making partner identity important to track in conjunction with annual variation. 
-Specifically, where some ants offered greater than average benefits in one year, other species offered reduced benefits in the same year, indicating that each ant partner reacts differently to the fluctuating environment.
-
-Each demographic trait was not affected equally by the intersection of ant partner and annual fluctuations.
-We looked at the mean effect of each ant partner on growth (figure \ref{fig:Annual_Ant}a), survival (Figure \ref{fig:Annual_Ant}b), and viability (Figure \ref{fig:Annual_Ant}c) rates across all the years included in our study.
-Where the mean effect is exactly 0 there is missing ant data due to variation in censusing focus.
-This is an indirect way to analyze how temporal fluctuation impacts our system without attributing the effects to a specific climate variables. 
-Where the mean effect is positive, ant presence increases the respective estimated growth, survival, or viability rates in response to environmental fluctuations, and where the mean effect is negative, the opposite occurs.
-We found the general magnitude of variation was the smallest for the growth rate, meaning annual variation affected the growth of plants the least, followed by survival then viability.
-
+\tom{Inter-annual variation results}{I think the text you had for this did not directly address the synchrony hypothesis, and it may be more informative to look at correlation coefficients rather than the time series, which are hard to make heads or tails of.}
 \begin{figure}[H]
@@
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-We found that there is a high frequency of partner turnover observed in this system with very distinct directional patterns. 
-Small plants are almost always vacant (Figure \ref{fig:Ant_Transition}b-d), if they were previously tended by a partner, they are likely to become vacant again (Figure \ref{fig:Ant_Transition}a).
-As the size of the plants increases, the probability of becoming tended increases as well, though it is not equally likely to be tended by all partners.
-
-\textit{L. apiculatum} ants become the most likely next partner in the case of most large plants, with the exception of a large plant which was previously tended by \textit{C. opuntiae}.
-Plants which were previously tended by \textit{C. opuntiae} ants are most likely to remain tended by \textit{C. opuntiae} ants.
-This indicates that these species may follow the well documented retention-discovery trade off exhibited  by many ants, with \textit{L.
-apiculatum} ants excelling at discovery and colonization of new plants and \textit{C.
-opuntiae} ants excelling at plant retention season after season.
-The overall frequency of these partners, however is a potential alternative explanation to this discovery-retention trade off.
-\textit{L.
-apiculatum} ants are by far the most frequent ant partners, accounting for 75\% of the occupied cacti, while \textit{C.
-opuntiae} and other ants account for 17\% and 8\% of occupied cacti respectively.
-This high frequency of \textit{L. apiculatum} ants may lead to the inflation of turnover to \textit{L. apiculatum} from other ants.
+We found a \tom{high frequency of partner turnover across transition years}{could we estimate the global average turnover rate?
+Across all observed transitions, what fraction of plants switched ant states.
+I think that would be a good place to begin, then unpacking that into the effects of initial size and ants.}, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
+Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
+Even when small plants are ant-tended at the start of the transition year, they are most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
+The probability of becoming ant-tended increases with size, though it is not equally likely to be tended by all partners. 
+For large plants that are initially vacant or tended by \textit{L.
+apiculatum} or Other ants, \textit{L.
+apiculatum} is the most likely next partner, suggesting that this partner species is able to colonize plants that were previously vacant or occupied by Other ants, and effectively retain plants that it previously occupied.
+\textit{C.
+opuntiae} were also able to retain plants they previously occupied, but not as well as \textit{L.
+apiculatum}: for plants that begin the transition year with \textit{C.
+opuntiae}, the probability that those plants remain occupied by \textit{C.
+opuntiae} at the end of the transition year is only slightly greater than the probability of take-over by \textit{L.
+apiculatum}, while take-over in the other directino is extremely rare.
+It is also notable that transitions away from the initial state of  \textit{L.
+apiculatum} were almost always transitions to vacancy (Figure \ref{fig:Ant_Transition}d), while transitions away from the initial states of \textit{C.
+opuntiae} and Other  were often transitions to other ants.
+This suggests a competitive hierarchy whereby \textit{L. apiculatum} may abandon low-value plants with litte nectar production but is almost never displaced from high-value plants.
 
@@
 \subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
-We found that partner diversity was beneficial in this system. 
+By integrating vital rate results and ant transition dynamics into the multi-state IPM we can evaluate the fitness implications of different scenarios of partner diversity and identity. 
+First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
 The lowest mean fitness was $\lambda_{S,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-By subtracting the distributions $\lambda_{S, any} - \lambda_{S,Vacant}$, we found that we are between 82\% and 100\% confident that having any partner leads to a higher population fitness than having no partner.
-If you consider just the number of partners (ignoring the identity), what you find is actually that the more partners are present, the higher the fitness of the cacti (Figure \ref{fig:LambdaMeans}b), though this increase is primarily driven by the presence of \textit{L.
-apiculatum}.
+Across all 1+ partner scenarios, we are 82--100\% confident that any scenario of ant visitation elevates fitness. 
+Furthermore, we find an apparent positive effect of partner diversity, with cactus fitness postive related to partner richness, averaging over partner identities (Figure \ref{fig:LambdaMeans}b).
 
@@
 
-We found evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+\tom{We found evidence}{I think this should say ``no evidence''??} of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are 94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the stochastic and stochastic null model. 
+We are are 94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized and non-synchronized model scnearios. 
 When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
 \ali{There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.}{I am not sure if I have explained enough here honestly.}
+\tom{}{I agree with your interpretation, but let's come back to editing this after we talk through the best way to look at portfolio effect.}
 

### 2024-03-20 Manuscript/ant_cactus_AmNat_Ali.tex
Updates
Visuals Updates

--- 2023-08-02 Manuscript/ant_cactus_AmNat_Ali.tex
@@
 \usepackage[authoryear,sectionbib,sort]{natbib}
+%\usepackage[square,sort,comma,numbers]{natbib}
 \linespread{1.7}
@@
 \usepackage{titlesec}
+\usepackage{xcolor}
+\usepackage{graphicx}
+\usepackage{placeins}
+\usepackage{float}
+\usepackage{subfigure}
 \newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}} 
-\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}
+\newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}
 
@@
 
-\title{Thank you, next: partner turnover elevates benefits of mutualism for an ant-tended plant}
+\title{An Integral Projection Modeling Approach to Understanding Demographic Effects of Multispecies Mutualisms}
 
-% This version of the LaTeX template was last updated on
-% May 11, 2023.
-
-%%%%%%%%%%%%%%%%%%%%%
-% Authorship
-%%%%%%%%%%%%%%%%%%%%%
-% Please remove authorship information while your paper is under review,
-% unless you wish to waive your anonymity under double-blind review. You
-% will need to add this information back in to your final files after
-% acceptance.
-
-\author{Alexandra Campbell$^{1,\ast}$ \\ 
-	Tom E.X. Miller$^{1}$}
+\author{Alexandra Campbell$^{1,\dagger}$ \\ 
+	Tom E.X. Miller$^{1,\ast}$}
 
@@
 	\section*{Abstract}
+	% 200 words max
+	The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
+	\textit{Cylindropuntia imbracata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
+	We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
+	We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in the wild.
+	The hierarchical models revealed that different ant partners had different impacts on the cholla vital rates. 
+	Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
+	The IPM results revealed that scenarios which included \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
+	This suggests that diversity benefits in this system are explained by sampling effect.
+	This study highlights that partner diversity can increase the overall benefits a focal mutualist receives and the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
 
 	\newpage{}
+	\section*{Introduction}
+	Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
+	They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+	Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
 
-\section*{Introduction}
+	Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}.
+	A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
+	Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of context-dependence in mutualism \citep{Stanton2003}.
+	Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014}.
 
-% The journal does not have numbered sections in the main portion of
-% articles. Please refrain from using section references (à la
-% section~\ref{section:CountingOwlEggs}), and refer to sections by name
-% (e.g. section ``Counting Owl Eggs'').
-Mutualisms are species interactions where all participants benefit, leading to higher individual fitness and increased population growth rates. 
-They are among the most widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013}, but can deteriorate into commensalism or parasitism  \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
-Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities.
+	There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \cite{Yeung2006,Barrett2015,Ushio2020}.
+	First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
+	This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
+	However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+	Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
+	This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
+	Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \cite{Afkhami2014}. 
+	Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+	Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits and protection across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
 
-Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010}.
-A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Even within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings such as agriculture \citep{rogers2014bee}, restoration (cite), and pest management (cite).
-
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist -- mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function (BEF chapter).
-First, when there is a hierarchy of fitness effects -- a consistent ranking of best to worst mutualists -- a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
-This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018}. 
-However, if partner associations are mutually exclusive then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the best partner \citep{Miller2007}.
-Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type, and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
-This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
-Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions (cite). 
-Third, partner species can have species-specific responses to the environment, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across environmental stochasticity or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022}.
-
-Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014}. 
-Sequential associations are likely when alternative partners engage in interference comptetition for access to a shared mutualist (cite examples, including non-ant-plant examples). 
+	Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
+	Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \cite{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
 	Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
-The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004}.
-Directionality of turnover can also influence diversity effects, particularly if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003}.
-For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005}.
+	The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate with successive turnover (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004,Fiala1994}.
+	Directionality of turnover can also influence effects of partner diversity if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003,Noe1994,Dejean2008}.
+	For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
 
 	Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
-Extrafloral nectar (EFN) producing plants can serve as dietary resources for for ant species which can promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006}.
-Presence of defensive ant partners is linked to herbivory reduction \citep{Trager2010, Rudgers2010} and increased plant growth and survival \citep{Baez2016}.
-These interactions are almost entirely studied from the perspective of plant fitness \citep{Bronstein1994,Bronstein1998}, with little recorded about the impacts on ant fitness \citep{Lanan2013}.
+	Extrafloral nectar (EFN) bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
+	Presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
 	Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
-Ant partners can vary in their ability to deter herbivores \citep{bruna2004}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
-Another source of temporal variation is succeptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the effectiveness of ant partners.
-Temporal dynamics of partner visitation therefore have important impacts on the fitness of the plant partners in these interactions \citep{Barton2010, Boege2005, Fonseca2003}.
-Recently many studies have investigated how partner diversity in these guilds has been shown to lead to either increased plant fitness \citep{Palmer2010, Afkhami2014} or decreased plant fitness (often in more highly specialized interactions) \citep{Barrett2015, Ushio2020}, stressing the importance of considering all ant partners as unique within these guilds.
+	Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+	Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
+	Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
+	Recent studies have begun to investigate how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
+	However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
 
@@
 	Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several infrequent species, all of which are ground-nesting. 
-These ant species locally co-occur at the scale of meters, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014,donald2022does}.
+	These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014, Donald2022}.
 	Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits and \textit{Crematogaster opuntiae} actually having net negative fitness effects though deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, those studies did not consider variation in ant defense across the plant life cycle, nor did they account for inter-annual fluctuations, and therefore may have missed important mechanisms through which different partner species, and their combination, may be beneficial.
+	Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong fitness benefits through anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative fitness effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+	However, those studies did not integrate the demographic effects of ant defense across the plant life cycle, nor did they account for inter-annual fluctuations in the herbivore populations (data on herbivory is included in supplementary documents).
 
-In this study we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species and how the demographic effects of alternative partner species varied across the range of plant size structure and nearly 20 years of inter-annual fluctuations.
-We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of mutualism for the focal plant partner.
+	We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity at the individual level, and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
+	We used this observational data set, contextualized by previous experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
 	Specifically, we asked:
@@
 	\end{enumerate}
-We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy, and to quantify state-dependent partner turnover. 
-We then used a multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
+	We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
+	We then used a stochastic, multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
 
@@
 
-This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico.
-This is an area characterized by steep, rocky slopes, and perennial vegetation like cacti and junipers. 
+	This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research site (SEV-LTER) in central New Mexico, USA.
+	This is an area characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B. gracilis}), yuccas, cacti, and junipers. 
 	Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \citep{Benson1982}. 
-These arborescent cholla produce cylindrical segments with large spines. 
-In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
+	These arborescent plants produce cylindrical segments with large spines. 
+	In the growing season (May to August in New Mexico), the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
 	While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
-Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
-Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative and reproductive structures \citep{Ness2006,Oliveira1999}.
+	%Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
+	Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative segments and flower buds \citep{Ness2006,Oliveira1999}. 
+	Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners. 
+	Smaller cholla produce little to no EFN, so larger cholla, especially flowering individuals, are generally more highly tended \citep{Miller2014}.
 
-This EFN is harvested by various ant species in return for defense. 
-At the Sevilleta, the cholla are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus}, an \textit{Aphaenogaster spp.}, and a \textit{Camponotus spp.}.
+	Tree cholla EFN is harvested by various ant species. 
+	At SEV-LTER, cholla are visited primarily by two species of ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species of \textit{Aphaenogaster} and \textit{Camponotus}.
 	\textit{L.
 apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $0\% - 20\%$ of cacti \citep{Donald2022} depending on the year.
-Up to $80\%$ of cacti remain vacant in any given year.
-
-These ants rarely co-occur on a plant, probably due to interspecific competition \citep{Miller2007}.
+opuntiae} visiting between $5\% - 20\%$ of cacti depending on the year \citep{Donald2022}.
+	Between $ 30\% 80\%$ of cacti remain vacant in any given year. 
+	These ants rarely co-occur on a plant, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
+opuntiae} to \textit{L.
+apiculatum}-tended plants, and vice versa, provoke aggressive responses by resident ants (A.
+Cambpell, \textit{personal observation}).
 	Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
-At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
-They will visit the cholla every day during the season around the clock, with the most acticity around sunrise or sunset \citep{Ohm2014}. 
-Smaller cholla are less likely to be visited because they produce very little EFN, so larger cholla are generally more highly tended \citep{Miller2014}. 
 	In late August, the tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-There are a variety of insect herbivores and seed predators which attack the cholla, focusing either on the vegetative segments and the reproductive segments \citep{Mann1969}. 
-An unidentified weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
+	There are a variety of insect herbivores and seed predators which specialize on tree cholla \citep{Mann1969}. 
+	A weevil of the genus \textit{Gerstaekeria} feeds on vegetative and reproductive structures and implants their larvae within the plant tissue for the winter. 
 	A cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
 	A seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attacks developing fruits pre-dispersal and oviposits in open flowers mid-growing season where larvae burrow into the ripening ovary.
-These predators can have significant negative impacts on the fitness of individual cholla and depress population growth \citep{Miller2009}.
-There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory from all of the mentioned insect predators \citep{Miller2007}.
+	These predators can have significant negative impacts on plant fitness of and depress population growth \citep{Miller2009}.
+	There is experimental evidence that tree cholla tended by \textit{L. apiculatum} and \textit{C. opuntiae} experience less herbivory than plants from which ants were excluded \citep{Miller2007}.
 
-\subsection*{Data}
+	\subsection*{Data Collection}
+	This study is based a long-term demographic data set spanning 2004 to 2023 at SEV-LTER. 
+	From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
+	This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
+	Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots has since been censused annually through 2023 (with the exception of 2020 due to the covid shutdown). 
+	For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded the height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
+	Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
+	We recorded reproductive effort as counts of viable and aborted flowerbuds. 
+	We recorded the ant species present (or vacancy if no ants present).
+	Occurrences of more than one ant species on one plant were rare (less than 5\% of observed ant species), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
+	Plots were searched for new recruits each year, and these were added to the census.
+	In total, the data set included 1141 unique individuals and 19 plant-year observations. 
+	These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
 
-The data collected are from a long-term dataset spanning 2004 to 2023 taken from 30 $\times$ 30 meter plots at the Sevilleta LTER. 
-The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
-Six of the plots were established in 2009 by tagging all existing plants within a 30 $\times$ 30 meter area. 
-The final two plots were added to this census from 2011 onwards. 
-Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
-For each cholla, we recorded plant survival from the last survey to the current survey. 
-We recorded the height (cm), maximum crown width, and crown width perpendicular to the maximum, which are used to calculate plant volume ($cm^3$) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter.
-We recorded the total number of flower buds, including how many were aborted and how many were not. We recorded all ant species present and the number of ants we could count in 30 seconds. 
-We also recorded the species of herbivore and the number present on the plant.
+	We used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
+	Ohm et al. \citep{Ohm2014} provide data on the number of seeds per fruit for plants tended by \textit{L. apiculatum}, \textit{C. opuntiae}, or no ants (experimental exclusion). 
+	Miller et al. \citep{Miller2009} provide data on seed entry to the seed bank and seedling germination and survival rates.
 
-\subsection*{Statistical Modeling -- Construction}
+	\subsection*{Multi-state Integral Projection Model}
+	Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables. 
+	While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
+	We constructed a multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with their transition dynamics modeled explictly.
 
-With the data described above we fit a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework with both fixed and random effects.
-Many of the vital rates are estimated as a function of plant size, ant partner, or both.
-Ant partner type is included as a predictor only where there are biological pathways through which ants could impact the outcome of that process. 
-The biological sources of variance (including individual, spatial, and annual variance) are accounted for by including year-to-year and plot-to-plot random effects in the models. 
-Unless otherwise mentioned, all models use vague priors. 
-The growth model ($G_j(y,x)$) estimates the size of cholla, with fixed effects of the previous size and ant partner and random effects of plot and year, using a Skew Normal distribution, with $\omega$ and $\alpha$ varying with the previous size.
-Ants are included as a predictor here because ant partners defend plants from herbivory, therefore decreasing the likelihood of segment loss.
-The survival model ($S_j(x)$) estimates the probability of survival, with fixed effects of the previous size of the cholla and ant partner and random effects of plot and year, using a Bernoulli distribution.
-Ants are included as predictors here because ant partners defend cholla from herbivores and predators, decreasing the likelihood of mortality due to either of these. 
-The reproduction model ($P(x)$) estimates the probability of reproducing each year, with fixed effects for the size and random effects of plot and year, using a Bernoulli distribution. 
-The total flowers model ($F(x)$) estimates the total flowers produced by a plant, with fixed effects of size and random effects of plot and year, using a Negative Binomial distribution. 
-The viability model ($V_i(x)$) estimates the proportion of flowers produced by a plant which are viable (not aborted), with fixed effects of the ant partner of the cactus and random effects of plot and year, using a Binomial distribution.
-Ants are included as predictors here because they defend the cacti from seed predation which can lead to floral abortion. 
-The ant transition rates model ($\tau_{i,j}(x)$) estimates the probability of a cactus being visited by an ant partner, with fixed effects of the previous size of the cholla and the previous ant partner and random effects of plot and year, using a Multinomial distribution.
-Ant partners are included as predictors here because partners may choose to return to the same cholla repeatedly or choose new ones, therefore the previous partner may be a good indicator of the next partner.
-The recruit size model ($n_j(x)$) estimates the size distribution of all recruits from a given year, with no fixed or random effects, using a Normal distribution. 
-With germination data from Miller et al., 2009, we fit two Bayesian generalized linear models for the probability of germinating from a seed in the first year ($\gamma_1$) or the second year ($1 - \gamma_1$), with no fixed or random effects, using a Binomial distribution.
-With data collected in a 2005-2006 recruit census, we fit a Bayesian generalized linear model for the probability of a seedling surviving to May ($\delta$) (accounting for missed mortality events), with fixed effects of the previous size and random effects of the transect, using a Bernoulli distribution.
-With data from Miller 2007, we fit a Bayesian generalized linear model for the number of seeds produced by every flower on a cholla ($\kappa$) based on the ant partner, using a Negative Binomial distribution.
-Ant partners are included as predictors here because they reduce floral abortion rates and therefore may lead to higher numbers of seeds.
+	Given the low frequency of ant species other than \textit{L.
+apiculatum} and \textit{C.
+opuntiae} (7.78\% of observed ants) we combined observations of all other ants into an ``other'' category, such that our models included four possible ant states: vacant, \textit{L.
+apiculatum}, \textit{C.
+opuntiae}, and ``other''.
+	The ``other'' category was made of unidentified ant species (2.8\% of observed ants), unknown species belonging to the geni \textit{Aphenogaster} (0.4\%),  \textit{Camponotus} (0.9\%),  \textit{Tetramorium} (0.02\%), \textit{Brachymyrmer} (0.02\%), a honeypot ant (0.08\%), and \textit{Forelius pruinosus} (3.5\%).
+	Ant state is included as a predictor variable in sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
+	For example, ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
+	Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
+	In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering and flowerbud production, independently of its influence on plant size. 
+	Therefore, these sub-models included plant size but not ant state as predictor variables.
 
-To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R 
-For each model, we obtained 3 chains of 10,000 iterations, each with randomly chosen initial conditions. 
-The first 1,500 iterations were discarded as burn-in to eliminate transience associated with initial conditions. 
-We did not thin the chains, thus all samples were retained. 
-To assess the convergence of our models we assessed between and within chain convergence, the resulting figures are included in supplemental documents. 
-To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
-Large differences in the two indicate a poor model fit and can be assessed visually (figures included in supplemental documents). 
-All estimated parameters are described in table 1.
-Data and code for all vital rate models is included in the supplemental information.
-
-\subsection*{Integral Projection Model Construction}
-
-Integral Projection Models describe population dynamics in discrete time, with continuous functions that relate vital rates to continuous and discrete state variables. 
-We constructed a composite IPM which allows us to analyze the long term population growth rate of cholla with ant transition dynamics explicitly included.
-Many of the vital rate models are dependent not only on continuous size variables, but also on discrete variables representing the ant partner present.
-There is also a transition rate which determines what proportion of cholla are tended by each ant in a given year based on the previous ant partner and the size of the plants. 
-This unique composite structure allows us to determine the individual effects of each ant species as well as the composite effects of several partners across the cholla population.
-
-Following previous studies, we modeled the life cycle of cholla using continuously size-structured plants, $n_i(x)$, where x is the size of the cholla and i is the ant partner, and two discrete seed banks ($B_{1,t}$ and $B_{2,t}$) corresponding to 1 and 2-year old seeds.
-The dynamics of these 1 and 2-year old seedbanks are given by the following equations:
+	Following previous studies, we modeled the tree cholla life cycle using continuously size-structured plants where $n(x,a)_{t}$ gives the number of plants of size $x$  and ant state $a$ in year $t$, plus two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds.
+	Seed bank dynamics are given by:
 
@@
 		$$
-		B_{1, t+1} = \kappa \delta \sum_{i}^{4} \int_L^U P(x) V_i(x) F(x) n_i(x) dx \\
+		B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x;\pmb{\tau^P}) F(x;\pmb{\tau^F}) V(a;\pmb{\tau^V_{a}}) n(x,a)_{t} dx \\
 		$$
 		$$
-		B_{2,t+1} =  (1 - \gamma_1)B_{1,t}\\
+		B^2_{t+1} =  (1 - \gamma_1)B^1_{t}\\
 		$$
@@
 
-The functions $P(x)$ and $F(x)$ give the probability of flowering, the number of flowerbuds produced based on the plant size $x$ and the year $t$. 
-The proportion of flowerbuds which will produce seeds ($V_i(x)$) is dependent on the plant size $x$ and the ant species present on the plant $i$ in year $t$. 
-The integral is multiplied by the number of seeds per fruit ($\kappa$) and the probability of seed dispersal/survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
-Parameters $U$ and $L$ are the upper and lower bounds, respectively, of the plant size distribution. 
-Plants can recruit out of the one-year seed bank with the probability of $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
-Seeds in the two-year seed bank are assumed to either germinate with a probability of $\gamma_2$ or die.
+	\noindent In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
+	Functions $P(x;\pmb{\tau^P})$ and $F(x;\pmb{\tau^F})$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
+	The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau^V_{a}})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$ but not size. 
+	The vector $\pmb{\tau}$ gives year-specific deviates (with mean zero) and appears in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$ and year.
+	Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
+	Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+	Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
+	Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.
 
-The size dynamics of the cholla are given by:
-
+	For the above-ground part of the life cycle, the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) is given by survival/growth transitions from size $x$ and ant state $a$ in year $t$, plus germination out of the seed banks:
 	\begin{linenomath*}
 		$$
-		n(y,i)_{t+1} = (\gamma_1 B_{1,t} + \gamma_2 B_{2,t}) \eta(y) \omega \beta_i  + \\
+		n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + \\
 		$$
 		$$
-		\sum_{j}^{4} \int_L^U S_j(x) G_j(y,x) \tau_{ij}(x) n_j(x) dx \\
+		\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
 		$$
@@
 
-The first term gives the recruitment from one and two-year seed banks to a plant of size $y$, where $\eta(y) ~ N(***)$ gives the seedling size distribution and $\omega ~ ***$ gives the proportion of seedlings which surive from germination (late summer) to the census (May).
-The second term reflects the changes in the population of the cholla which are not recruits, where $S_j(x)$ gives the probability of a plant of size $x$ in year $t$ surviving to year $t+1$ with partner $j$.
-$G_j(y,x)$ gives the probability of growing from size $x$ in year $t$ to size $y$ in year $t+1$, respectively with partner $j$. 
-Finally, $\tau_{ij}$ is the probability of a cholla which is size $x$ with ant partner $i$ in year $t$ being tended by ant partner $j$ in year $t+1$.
+	\noindent The first term estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of seedling survival ($\omega$) from germination (late summer) to the census (May).
+	This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appearance in our census ($\sum_{a'}^{A}\rho_{0}(a')=1$). 
+	The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
+	Survival from initial size $x$ ($S(x,a;\pmb{\tau^S_{a}})$) and growth from size $x$ to $x'$ ($G(x',x,a;\pmb{\tau^G_{a}})$) are both dependent on initial size and ant state. 
+	As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau_{a}}$). 
+	Ant transition function $\rho(a',a,x;\pmb{\tau^{\rho}})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
+	This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau^\rho}$).
 
-\subsection*{Deterministic IPM Analysis}
+	\subsection*{Statistical modeling and parameter estimation}
+	We parameterized the IPM using a series of generalized linear mixed models (GLMMs) in a hierarchical Bayesian framework to serve as vital rate sub-models. 
+	Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively, and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
+	In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transition between ant states conditional on plant size and previous ant state. 
+	All models were fit using R version 4.2.2 and Rstan package (\cite{Rstancite, Rcite}).
+	Unless otherwise mentioned, all used vague priors.
 
-Analyzing an IPM requires discretizing the composite IPM into a matrix to calculate the dominant eigenvalue. 
-Each component of the IPM (each set of ant partner specific vital rates) can be discretized into its own $ 200 \times 200$ matrix as shown in figure ***. 
-The possibility of eviction, when individuals are predicted to grow outside of the possible size classes, was avoided by adding probabilities of growing smaller or larger than the existing boundaries, as done by many others***.
-We used the composite IPM to quantify the effects of partner diversity on the intrinsic growth rate of cholla, $r$ $ln(\lambda)$. 
-We calculated the $r$ for each combination of ant partners: complete vacancy; \textit{L.
-apiculatum} and vacancy; \textit{C.
+	The vital rate models take the form of the function:
+	$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
+	\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+	Some models deviate from this general model, which will be specified below.
+
+	\paragraph{Growth}
+	The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+	We fit this model to the size in year $t+1$ ($y^G$) using the location-scale parameterization of the student $t$ distribution and the identity link function, because in preliminary analyses we found that size transition data were more fat-tailed than a Gaussian distribution could accommodate.
+	$$y^G \sim Student T(\hat{\nu},\hat{G},\hat{\sigma}) $$
+	Student $t$ distributions are dependent on three parameters, $\hat{\nu}$,  $\hat{G}$, and $\hat{\sigma}$.
+	$\hat{G}$ is the mean of the distribution, and in our case is expressed as a second-order polynomial with ant-size interactions because  preliminary analysis found this was a better fit to slightly concave data.
+	$\hat{\nu}$ is the shape parameter which determines how thin the tails of the distribution are.
+	$\hat{\sigma}$ is the scale parameter which determines how wide the distribution is. 
+	Both $\hat{\nu}$ and $\hat{\nu}$ are size dependent because preliminary analyses revealed that there was size dependence in both variance and kurtosis.
+	%$$\hat{G} = \beta_{0}^{G} + \beta_{1}^{G}  x  + \beta_{2}^{G}  a + \beta_{3}^{G}  x  a  + \beta_{4}^{G}  x^2 + \beta_{5}^{G}  a  x^2 + u + w $$
+	$$\hat{\sigma} = \beta_{0}^{\sigma} + \beta_{1}^{\sigma}  x $$
+	$$\hat{\nu} = \beta_{0}^{\nu} + \beta_{1}^{\nu}  x $$
+
+	\paragraph{Survival}
+	The survival model ($S(a,x;\pmb{\tau_{a}^{S}})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of the previous size of the cholla $x$ and ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+	We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function
+	$y^S \sim Bern(\hat{S})$.
+	%$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u + w$$
+
+	\paragraph{Reproduction}
+	The reproduction model ($P(x';\pmb{\tau^{P}})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
+	We fit this model to the reproductive data $y^R$ using a Bernoulli distribution and a logit link function:
+	$$y^P \sim Bern(\hat{P})$$
+	%$$logit(\hat{P} = \beta_{0}^{P} + \beta_{1}^{P} \times x' + u + w)$$
+
+	\paragraph{No. Flowers}
+	The total flowers model ($F(x';\pmb{\tau^{F}})$) estimates the total flowers produced by a plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$ and random effects of plot $w$ and year $u$.
+	We fit this model to flowerbud count data $y^F$ using a zero-truncated negative binomial distribution with a log transformation:
+	$$y^{F} \sim 0 Truncated Negative Binom(\hat{F},\hat{\phi})$$
+	%$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u + w$$
+	$$log(\hat{\phi}) = \beta_{0}^{\phi}$$
+
+	\paragraph{Flowerbud viability}
+	The viability model ($V(a;\pmb{\tau^{V}_{a}})$) estimates theproportion of flowers produced by a plant which are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$ and random effects of plot $w$ and year $u$ which interacts with ant state $a$.
+	We fit this model to floral abortion data $y^V$ using a Binomial distribution and a logit link function:
+	$$y^{V} \sim Binom(\hat{V})$$
+	%$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
+
+	\paragraph{Seeds Per Fruit}
+	With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+	We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
+	$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
+	%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
+	$$\hat{\phi} = \beta_{0}^{\phi}$$
+	The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+
+	\paragraph{Ant Transitions}
+	The ant transition model ($\epsilon(x,a,a';\pmb{\tau^{\epsilon}})$) estimates the probability of a cactus being visited by an ant partner $a'$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
+	We fit this model to ant partner data from year $t+1$ $a'$ using a multinomial distribution with a logit link function: 
+	$$y^{\epsilon} ~ Multinomial(\hat{\epsilon})$$
+	%$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
+
+	\paragraph{Recruit Size Distribution}
+	%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
+We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
+$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
+%$$\hat{\eta} = \beta_{0}^{\eta}$$
+where $\hat{\sigma}$ is estimated with a non-informative prior.
+
+\paragraph{Germination}
+With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
+$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
+%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
+
+\paragraph{Pre-Census Survival}
+With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
+We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
+$$y^{\delta} ~ Bern(\hat{\delta})$$
+%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
+\paragraph{Parameter estimation}
+To obtain posterior estimates of the demographic parameters, we fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+For each model, we obtained three chains of 10,000 iterations, each with randomly chosen initial conditions. 
+The first 1,500 iterations were discarded. 
+We did not thin the chains, thus all samples were retained following burn-in. 
+We assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b). 
+To assess the overall model fit we carried out posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
+All estimated parameters are available in Table \ref{tab:Params}.
+
+\subsection*{IPM Analysis}
+We constructed two stochastic IPMs in order to simulate ant combinations that don’t exist in real life with two different versions of environmental variation.
+The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
+The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
+By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
+By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
+In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
+
+Analyzing an IPM requires discretizing the continuous IPM kernel into an approximating matrix. 
+Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
+In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$ where $A$ is the number of unique combinations of $a$ and $a'$ (how many possible ant transitions there are).
+Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $A(b+2) \times A(b+2)$.
+We used $b = 200$ bins after stability analysis to determine how many bins were needed to consistently estimate the same lambda.
+We extend the integration limits $L$ and $U$ (the lowest and highest size values observed) to avoid unintentional eviction \cite{Williams2012}.
+
+Traditionally, in a bayesian IPM, rather than having one set of parameters to populate an approximating matrix and estimate population fitness ($\lambda$), we use 1,000 draws from the posterior distributions of our bayesian statistical models to populate 1,000 approximating matrix and estimate a 1,000 element distribution for $\lambda$.
+After preliminary stability analysis we found that 1,000 random posterior draws leads to the same estimate of $\lambda$ as any greater number of posterior draws.
+In a stochastic bayesian IPM, the estimation of each individual $\lambda$ in the distribution comes from a geometric mean of many values. 
+Rather than a single approximating matrix for each value of $\lambda$, we take the geometric mean of 500 approximating matricies with year-ant-specific effects, resulting in a stochastic version of populationfitness ($\lambda_{S}$).
+These year-ant-specific effects are calculated as random effects for each unique combination of year and ant species included in our data. 
+While we do not have data on how the ant species actually fluctuate annually, we are able to measure how the effects of the ant species on plants fluctuate annually.
+Each of these 500 approximating matricies have the same draw from the posterior distribution, but are assigned the year-ant-specific effects of a random year of our data (year-ant-specific effects are sampled randomly with replacement).
+We chose to use 500 year-specific values, because our $\lambda_{S}$ calculations stablilized after 500 year-specific values. 
+To calculate a single $\lambda_{S}$ for these 500 year-ant-specific estimates, we take the geometric mean of the \ali{500 dominant eigenvalues}{OK, I am not positive this is correct, but I have forgotten how to correctly describe it.
+Can we talk through this sentence?} for a single stochastic $\lambda_{S}$.
+This process is repeated for each of the 1,000 random posterior draws, leading to a distribution of 1,000 stochastic $\lambda_{S}$ values. 
+To calculate the stochastic null bayesian IPM we use the exact same process, but the year-specific values used to calculate the stochastic null $\lambda_{SN}$ are not ant specific, forcing all ant species to have the same effects on their cactus partners in response to environmental variation.
+
+In the both versions of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
 opuntiae} and vacancy; other and vacancy; \textit{L.
@@
 opuntiae}, other, and vacancy; and all ant partners and vacancy.
-The relative distributions allowed us to determine if either complementarity or sampling effect were at play in the tree cholla-ant defense system.
-Sampling Effect requires that the $r$ when all possible partners are present is equal to the $r$ of the cholla population when only the best partner is present.
-Complementarity requires that the $r$ when all possible partners are present is greater than the $r$ of any other combination of partners or any single partner.
+Each partner scenario includes vacancy in addition to the combination of partners being analyzed for added realism because we have never observed a population of cacti in which all are occupied.
 
-\subsection*{Stochastic IPM Analysis}
+We can compare the distributions of $\lambda_{S}$ in many ways, but have chosen two moments in particular: the means and the proportion of each distribution that is greater than another.
+To calculate this second moment, we can subtract each $\lambda_{S}$ distribution from the others (\ali{e.g.}{This notation may need to be revised somewhat} $\lambda_{S,Vacant}$ - $\lambda_{S,C.
+opuntiae \& Vacant}$) to determine which partner scenario leads to a higher fitness for the cactus population.
+The proportion of the resulting vector that is greater than zero allows us to give a probabilistic estimate of which combination results in the greatest fitness. 
+The magnitude of this proportion is how we determine the magnitude of difference between the resulting population fitnesses.
 
-The third mechanism considered in this study requires annual variation to be explicitly considered in the IPM so we constructed a stochastic version of the IPM described above. 
-The year of record was a random effect in all vital rate models meaning we were able to include an intercept for specific years. 
-In order to include stochasticity in the analysis, we randomly sampled from the year-random effect intercepts 1,000 times.
-For each of these 1,000 iterations we calculated the $r$ for every possible combination of ant partners. 
-With these values we calculated $\delta r$ between the scenario with all possible partners and the scenario with no partners.
-This $\delta r$ measures the net benefits offered by partner diversity with annual variation.
-We also calculated the $\delta r$ of the deterministic IPM to measure the net benefits offered by partner diversity without annual variation. 
-These two relative differences allow us to determine if portfolio effect is at play.
-Portfolio effect is demonstrated when the net benefit with annual variation is larger than the net benefit without annual variation.
+Using only the $\lambda_S$ distributions from the stochastic IPM we can determine which scenario leads to the highest fitness, which will tell us if there are benefits to diversity in this system, any other finding indicates that there is quantitative evidence that diversity benefits this system.
+If the partner scenario which leads to the highest possible population fitness is vacancy, this indicates that there are no benefits to partner diversity.
+If all of the highest scenarios include a common ant partner, this indicates that there is a single best partner. 
+This quantitative finding that one particular partner leads to all highest possible population fitnesses is evidence that sampling effect explains the benefits of partner diversity.
+If the highest scenario includes all possible ant partners, this indicates there is a synergy which occurs when all partners interact.
+This quantitative finding indicates that complementarity is the mechanism which explains the benefits of partner diversity in this system.
+
+Using the $\lambda_S$ distribution and the $\lambda_{SN}$ distribution, from both the stochastic IPM and the stochastic null IPM, we can determine if having multiple partners can buffer the cactus population from environmental variation.
+The difference in the full partner scenario and vacancy ($\lambda_{S,All}$ - $\lambda_{S,Vacant}$) can be interpreted as the effect of having all partners when each ant species has unique effects on the cacti in response to annual variation.
+The difference in the full partner scenario and vacancy ($\lambda_{SN,All}$ - $\lambda_{SN,Vacant}$) can be interpreted as the effect of having all partners when each ant species has the same effects on the cacti in response to annual variation.
+The proportion of the resulting vectors that are greater than 0 allows us to give  probabilistic estimates of the effects of partner diversity on the fitness of the cacti in a varying environment under two ant scenarios.
+These two distributions can again be subtracted ($\lambda_{S} - \lambda_{SN}$), allowing us to determine the difference in effects of partner diversity in two annually varying scenarios.
+The proportion of the resulting vector that is greater than 0 gives us a probablistic estimate of how strongly partner diversity buffers the population from environmental variation when the effects of each partner fluctuate uniquely.
+This quantitative finding that the effects of partner diversity in the stochastic IPM are greater than those in the stochastic null IPM indicate that portfolio effect can explain some benefits of partner diversity in this system.
+
+\ali{}{I focused on several things throughout this rewrite of the methods: I tried to go straight into talking about the stochastic IPM, I focused on explaining rationale for specific numbers better, I tried to update notation to be clearer, and I tried to explain more clearly...
+not sure if I've accomplished these, let me know}
 
 \section*{Results}
-\subsection*{Statistical Modeling}
+\subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
+\subsubsection*{The demographic effects of alternative ant partners}
+\paragraph{Growth Model} We found evidence that ant visitation enhances cactus growth and that partner identity can influence the growth trajectory of cholla (Figure \ref{fig:Grow}). 
+Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes (Figure \ref{fig:Grow}e).
+Plants with \textit{C.
+opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants (Figure \ref{fig:Grow}a).
+Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes (Figure \ref{fig:Grow}b). 
+Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates (Figure \ref{fig:Grow}c.)
+Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates (Figure \ref{fig:Grow}d).
 
-%% Growth Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{grow_contour_lines_color.png}
-%	\caption{asdasdf}
-%	\label{fig:growth}
-%\end{figure}
-Plants tended by other ants experiened positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}a.
-They experienced significantly higher growth than vacant plants ($p = 2.2 \times 10^{-16}$), and similar mean growth to \textit{L.
-apiculatum} tended plants ($p = 0.9135$) and \textit{C.
-opuntiae} tended plants ($p = 0.4864$).
-Plants tended by \textit{C. opuntiae} ants experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}b.
-They experienced significantly higher mean growth than vacant plants ($p = 4.88 \times 10^{-14}$).
-Plants tended by \textit{L. apiculatum} experienced positive growth rates across all sizes, with the highest chances of shrinking occurring at the largest sizes \cite{fig:growth}c.
-Vacant plants experienced the lowest observed growth, except at the largest sizes where they are less likely to shrink than tended plants \cite{fig:growth}d.
+Using the method of subtracting one size distribution from another we were able to determine how confident we are that the size of cholla visited by one combination of ant partners is greater than another.
+$G_{Crematogaster} - G_{Vacant}$ returns a vector which is 89\% positive, meaning we are 89\% confident that when \textit{C.
+opuntiae} ants are partners the size of the cholla is greater than the size of the cholla with no ant partners.
+Given the evidence that plants visited by \textit{C.
+opuntiae} experience the highest mean growth rates we have reported our confidence that this partner is associated with the highest growth rates below.
+We are 88\% and 70\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean growth rates across sizes than plants tended by \textit{L.
+spiculatum} ants or other ants respectively.
+We are 89\%, 65\%, and 94\% confident that plants with no partners experience lower mean growth rates across all sizes than plants tended by \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, or other ants respectively.
 
-%% Survival Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{surv_panels.png}
-%	\caption{asdasdf}
-%	\label{fig:surv}
-%\end{figure}
-Plants tended by \texit{C.
-opuntiae} ants experienced significantly higher mean survival rates than \textit{L.
-apiculatum} tended plants ($p = 2.908 \times 10^{-14}$), plants tended by other ants ($p = 4.488 \times 10^{-14}$), and vacant plants ($p = 2.2 \times 10^{-16}$) \cite{fig:surv}.
-These \textit{C. opuntiae} plants have the highest mean survival at small sizes before reaching a mean survival rate of about 100\% when medium to large.
-Plants tended by \textit{L. apiculatum} experienced similar mean survival rates to plants tended by other ants and vacant plants.
-\textit{L. apiculatum} tended plants and other tended plants reached a mean survival rate of nearly 100\% by medium sizes, with other tended plants reaching this rate slower. 
-Vacant plants never reach mean survival rates of 100\%.
+\begin{figure}
+	\includegraphics[width = 0.95\linewidth]{Figures/GrowContourLinesColor.png}
+	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
+The solid colored lines (seen in all panels) are the next mean predicted size of cholla.
+The points (seen in panels a-d) are the observed data which informs these estimates.
+The black countour lines (seen in a-d) appear at 5\% increments showing where 5\%, 10\%, etc.
+of the data is expected to fall.
+They grey dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).
+}
+	\label{fig:Grow}
+\end{figure}
 
-%% Flowers Model
-There is a clear size effect on the number of flowers produced. 
-The mean number of flowers produced by a plant remains at 0 until the plant reachesmedium sizes after which the mean number of flower produced increases eponentially to about 40 at large sizes.
+\paragraph{Survival Model}
+We found evidence that ant visitation enhances the survival of medium to large plants, and that partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
+Tree cholla experience between 7.7\% and 99.9\% survival rates depending on 
+their size and ant partner (Figure \ref{fig:Surv}e).
+Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
+Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes (Figure \ref{fig:Surv}a).
+Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes (Figure \ref{fig:Surv}b). 
+Plants with other ants experience the second lowest mean survival rates across all sizes (Figure \ref{fig:Surv}c).
+Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates (Figure \ref{fig:Surv}d).
 
-%% Viability Model
-%\begin{figure}[h]
-%	\includegraphics[width=0.95\linewidth]{Viability.png}
-%	\caption{asdasdf}
-%	\label{fig:comp-samp}
-%\end{figure}
-Plants tended by \textit{L.
-apiculatum} ants experienced the highest viability rates, significantly higher than \textit{C.
-opuntiae} tended plants ($p = 2.2 \times 10^{-16}$), other tended plants ($p = 2.2 \times 10^{-16}$), and vacant plants ($p = 2.2 \times 10^{-16}$).
-The mean viability rate of \textit{L. apiculatum} tended plants was 84.888\%.
-Plants tended by \textit{C. opuntiae} ants experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$), but not than other tended plants ($p = 0.1168$).
-The mean viability rate of \textit{C. opuntiae} plants was 75.934\%.
-Plants tended by other ants also experienced significantly higher viability rates than vacant plants ($p = 2.2 \times 10^{-16}$).
-The mean viability rate of other tended plants was 76.367\%.
-The mean viability rate of vacant plants was 70.442\%.
+Using the method of subtracting one survival distribution from another we were able to determine how confident we are that the survival of cholla visited by one combination of ant partners is more likely than another.
+$S_{Crematogaster} - S_{Vacant}$ returns a vector which is 82\% positive, meaning we are 82\% confident that when \textit{C.
+opuntiae} ants are partners the survival of the cholla is more likely than the survival of the cholla with no ant partners.
+Given the evidence that plants visited by \textit{C.
+opuntiae} experience the highest mean survival rates we have reported our confidence that this partner is associated with the highest survival rates below.
+We are 63\% and 100\% confident that plants tended by \textit{C.
+opuntiae} ants experience higher mean survival rates across all sizes than plants tended by \textit{L.
+spiculatum} ants or other ants respectively.
+We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by
 
-%% Reproduction Model
-The probability of a plant reproducing in a given year is highly size dependent. 
-The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/SurvivalPlot.png}
+	\caption{This figure shows the estimated survival rates based on the size of the cactus with each individual ant partner.
+The solid colored lines (shown on all panels) indicate the mean estimated survival rates.
+The dashed lines (shown in panel e) indicate extrapolations beyond existing data (where we estimated survival for plants tended by ants where we had never seen a tended cactus of that size).
+The grey area around the solid lines (shown in panels a-d) show the 90\% confidence interval for the estimates.
+The colored dots are the real data binned by size to show how our estimates align with real survival observations.
+A larger circle means we had more data on survival of plants of this size with this partner.}
+	\label{fig:Surv}
+\end{figure}
 
-%% Seeds Produced
-The mean number of seeds produced by \textit{C.
-opuntiae} tended plants is significantly lower than the mean number of seeds produced by \textit{L.
-apiculatum} plants ($p = 2.2 \times 10^{-16}$) and vacant plants ($p = 2.2 \times 10^{-16}$).
-\textit{C. opuntiae} tended plants produce a mean of 2.561 seeds per flower. 
-The mean number of seeds produced by \textit{L. apiculatum} tended plants is slightly higher than the mean number produced by vacant plants ($p = 0.02521$). 
-\textit{L. apiculatum} tended plants produce a mean of 3.563 seeds per flower. 
-Vacant plants produce a mean of 3.435 seeds per flower.
+\paragraph{Viability Model}
+We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability.
+Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers (Figure \ref{fig:viab}).
+The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
+apiculatum} tended plants experiencing the highest mean viability rate (at 86\%, Figure \ref{fig:Viab}b), followed by other tended plants (at 75\%, Figure \ref{fig:Viab}c), \textit{C.
+opuntiae} tended plants (at 74\%, Figure \ref{fig:Viab}a) and vacant plants (at 71\%, Figure \ref{fig:Viab}d).
 
-%% Precensus Survival
-The mean pre-census seed survival rate is 16.99\% with 90\% of model iterations landing between 1.13\% and 40.72\%.
+Using the method of subtracting one viability distribution from another we were able to determine how confident we are that the floral viability of cholla visited by one combination of ant partners is greater than another.
+$V_{Liomatopum} - V_{Vacant}$ returns a vector which is 99\% positive, meaning we are 99\% confident that when \textit{L.
+apiculatum} ants are partners the floral viability of the cholla is greater than the cholla with no ant partners.
+Given the evidence that plants visited by \textit{L.
+apiculatum} experience the highest viability rates we have reported our confidence that this partner is associated with the highest viability rates below.
+We are 98\% and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by \textit{C. opuntiae} or other ants respectively.
+We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
 
-%% Germination
-Seeds have a significantly higher probability of germinating in year one than in year two ($p = 2.2 \times 10^{-16}$).
-Seeds in year one have a mean germination rate of 0.578\% and seeds in year two have a mean germination rate of 0.221\%.
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/ViabHist.png}
+	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
+The solid lines indicate the estimated viability distribution.
+The colored histograms represent the observed viability rates of plants with that partner.
+}
+	\label{fig:Viab}
+\end{figure}
 
-%% Recruit size distribution
- The mean size of new recruits is -1.88 log($cm^3$) with 90\% of new recruits falling between the sizes of -2.12 log($cm^3$) and -1.64 log($cm^3$).
+\ali{}{Should I include other model results here or just indicate that the rest are reported in an appendix/supplementary materials?}
 
-%% Ant Transition Rates
-% previously vacant
-Plants that were previously vacant will most likely remain vacant until large. 
-These plants have significantly higher probabilities of being vacant rather than becoming tended  by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
-The probability of previously vacant plants becoming tended by \texit{L.
-apiculatum} ants is significantly higher than the probabilities of becoming tended by \textit{C.
-opuntiae} or other ants (respecitvely, $p = 3.83 \times 10^{-14}$, $p = 2.2 \times 10^{-16}$).
-The probability of previously vacant plants becoming tended by other ants is significantly lower than the probability of becoming tended by \textit{C. opuntiae} ($p = 0.00227$).
+\subsubsection*{The role of annual fluctuations in demographic effects of ant partners}
+We found evidence that annual variation impacts the effect of ant partner on various demographic traits (Figure \ref{fig:Annual_Ant}).
+The effects of each partner on demographic measure vary uniquely across temporal fluctuations, making partner identity important to track in conjunction with annual variation. 
+Specifically, where some ants offered greater than average benefits in one year, other species offered reduced benefits in the same year, indicating that each ant partner reacts differently to the fluctuating environment.
 
-% previously crem
-Plants that were previously tended by \textit{C. opuntiae} will most likely remain vacant until large. 
-These plants have significantly higher probabilities of becoming vacant rather than remaining tended by \textit{C.
-opuntiae} or becoming tended by \textit{L.
-apiculatum} or other ants (repsectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$ ).
-The probability of these plants remaining tended by \textit{C.
-opuntiae} ants is significantly higher than the probability of becoming tended by \textit{L.
-apiculatum} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 3.797 \times 10^{-9}$).
-The probability of these plants becoming tended by \textit{L. apiculatum} is significantly higher than the probability of becoming tended by other ants ($p = 3.797 \times 10^{-9}$).
+Each demographic trait was not affected equally by the intersection of ant partner and annual fluctuations.
+We looked at the mean effect of each ant partner on growth (figure \ref{fig:Annual_Ant}a), survival (Figure \ref{fig:Annual_Ant}b), and viability (Figure \ref{fig:Annual_Ant}c) rates across all the years included in our study.
+Where the mean effect is exactly 0 there is missing ant data due to variation in censusing focus.
+This is an indirect way to analyze how temporal fluctuation impacts our system without attributing the effects to a specific climate variables. 
+Where the mean effect is positive, ant presence increases the respective estimated growth, survival, or viability rates in response to environmental fluctuations, and where the mean effect is negative, the opposite occurs.
+We found the general magnitude of variation was the smallest for the growth rate, meaning annual variation affected the growth of plants the least, followed by survival then viability.
 
-% previously liom
-Plants that were previously tended by \textit{L. apiculatum} ants will most likely remain vacant until large. 
-These plants have significantly higher probabilities of becoming vacant than remaining tended by \textit{L.
-apiculatum} or becoming tended by \textit{C.
-opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, and $p = 2.2 \times 10^{-16}$).
-These plants have significantly higher probabilities of remaining tended by \textit{L.
-apiculatum} ants than becoming tended by  \textit{C.
-opuntiae} or other ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
-The probability of these plants becoming tended by \textit{C. opuntiae} ants is significantly higher than the probability of them becoming tended by other ants ($p = 0.007925$).
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/year_ant_timeseries.png}
+	\caption{This figure shows the mean affect of each ant partner on a) the estimated next size, b) the estimated survival, and c) the floral viability of cacti across every year of our study.
+These values are estimated from the fitted random effects of ant and year in our models.
+Each point represents the mean of the random effect of the identified model, ant, and year (e.g.
+the lowest dot in panel b) represents the mean effect of vacancy on survival rates in year 2011).}
+	\label{fig:Annual_Ant}
+\end{figure}
 
-% previously other
-Plants that were previously tended by other ants will most likely remain vacant until they are large. 
-These plants have significantly higher probabilities of becoming vacant rather than remaining tended by other ants or becoming tended by \textit{C.
-opuntiae} or \textit{L.
-apiculatum} ants (respectively, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$, $p = 2.2 \times 10^{-16}$).
-The probability of these plants becoming tended by \textit{L.
-apiculatum} ants is significantly higher than the probability of remaining tended by other ants or becomnig tended by \textit{C.
-opuntiae} (respectively, $p = 6.309 \times 10^{-12}$, $p = 6.256 \times 10^{-13}$).
-The probability of these plants remaining tended by other ants is not higher than the probability of becoming tended by \textit{C. opuntiae} ants ($p = 0.954$).
+\subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
+We found that there is a high frequency of partner turnover observed in this system with very distinct directional patterns. 
+Small plants are almost always vacant (Figure \ref{fig:Ant_Transition}b-d), if they were previously tended by a partner, they are likely to become vacant again (Figure \ref{fig:Ant_Transition}a).
+As the size of the plants increases, the probability of becoming tended increases as well, though it is not equally likely to be tended by all partners.
 
-%% All cases/trends
-In all plants ,the most likely scenario is vacancy.
-In most cases, the second most likely next ant partner is \textit{L.
-apiculatum} with the exception of plants which were previously tended by \textit{C.
-opuntiae} who are second most likely to retain \textit{C.
-opuntiae} as their partners in successive years.
-The only case where two next partners do not have statistically significantly different probabilities is when previously other tended plants either remain tended by other ants or become tended by \textit{C.
-opuntiae}.
+\textit{L. apiculatum} ants become the most likely next partner in the case of most large plants, with the exception of a large plant which was previously tended by \textit{C. opuntiae}.
+Plants which were previously tended by \textit{C. opuntiae} ants are most likely to remain tended by \textit{C. opuntiae} ants.
+This indicates that these species may follow the well documented retention-discovery trade off exhibited  by many ants, with \textit{L.
+apiculatum} ants excelling at discovery and colonization of new plants and \textit{C.
+opuntiae} ants excelling at plant retention season after season.
+The overall frequency of these partners, however is a potential alternative explanation to this discovery-retention trade off.
+\textit{L.
+apiculatum} ants are by far the most frequent ant partners, accounting for 75\% of the occupied cacti, while \textit{C.
+opuntiae} and other ants account for 17\% and 8\% of occupied cacti respectively.
+This high frequency of \textit{L. apiculatum} ants may lead to the inflation of turnover to \textit{L. apiculatum} from other ants.
 
-\subsection*{Deterministic IPM}
+\begin{figure}[H]
+	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+	\caption{This figure shows the probability of being tended by each ant partner or vacant based on the size of the plant.
+Each panel shows these probabilities for a different previous ant state.
+The solid lines represent the mean probability of being tended by a specific partner.
+The colored points are the real data binned by size to show how our estimates align with real visitation observations.
+A larger circle means we had more data on visitation of plants of this size with this previous partner.}
+	\label{fig:Ant_Transition}
+\end{figure}
 
-\subsection*{Stochastic IPM}
+\subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
+We found that partner diversity was beneficial in this system. 
+The lowest mean fitness was $\lambda_{S,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
+By subtracting the distributions $\lambda_{S, any} - \lambda_{S,Vacant}$, we found that we are between 82\% and 100\% confident that having any partner leads to a higher population fitness than having no partner.
+If you consider just the number of partners (ignoring the identity), what you find is actually that the more partners are present, the higher the fitness of the cacti (Figure \ref{fig:LambdaMeans}b), though this increase is primarily driven by the presence of \textit{L.
+apiculatum}.
+
+\begin{figure}
+	\includegraphics[width=0.61\linewidth]{Figures/LambdaMeans.png}
+	\includegraphics[width=0.39\linewidth]{Figures/Lambda_Num_Partners.png}
+	\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) and $\lambda_{SN}$ (empty circles with an X) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+	\label{fig:LambdaMeans}
+\end{figure}
+
+Despite this apparent synergy, when partner identity was considered, we found the benefits of partner diversity could be explained by Sampling Effect.
+We believe the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
+All simulated combinations of ant partners which included \textit{L. apiculatum} were nearly equal and the highest possible fitness estimated for the cholla. 
+This indicates that \textit{L. apiculatum} are the single best partner for the cholla under existing conditions. 
+Based on the definitions of Sampling Effect and Complementarity we use in this study \citep{Batstone2018}, it is clear that Sampling Effect can explain the benefits of partner diversity in the cholla system.
+
+It is possible that this, like the frequency of partner turnover to \textit{L. apiculatum}, is driven by the extreme frequency of \textit{L. apiculatum} ants in comparison to others. 
+With this in mind, we simulated the population fitness with equal probability for transitioning to any ant state.
+We found ...... from the simulations with different transition probabilities.
+\ali{}{This feels like the most natural progression to me.}
+
+We found evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
+We are are 94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the stochastic and stochastic null model. 
+When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
+\ali{There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.}{I am not sure if I have explained enough here honestly.}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
+	\caption{This figure shows the distribution of $\lambda_{S,All}-\lambda_{S,Vacant}$ in pink and $\lambda_{SN,All}-\lambda_{SN,Vacant}$ in green.
+The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
+	\label{fig:Portfolio}
+\end{figure}
 
 \section*{Discussion}
+\tom{}{I have not commented too heavily here because I would like to talk through what we want this section to achieve.
+Currently, most of this section is actually describing results, whereas the pupose is to interpret and contextualize results, and connect them to the broader literature.
+Some of your content here is actually better at describing results than you had in the Results section, because it includes that element of ``how is this connected to the question'' that was missing from the Results section.}
+The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
+Many studies have looked at multispecies mutualisms and the how having a variety of partners leads to variation in demographic effects \cite{Palmer2010, Bascompte2019, Stachowicz2005, Ford2015, Baez2016}.
+Because these tree cholla interact with only one ant partner at a time, it is a unique system in which to parse out the individual effects of each ant partners, both in isolated settings or in combinations we cannot test in the real world.
+\tom{}{This is a good start.
+In general, I suggest opening the Discussion section with a brief summary of what you were trying to learn in this study, what you found, and its broader significance.
+This paragraph does a little of that, but I think it can be stronger.}
 
-\section*{Conclusion}
+%\subsection*{Demographic effects of association with alternative partners}
+We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, \tom{and more}{I would not say this.}) have on the vital rates of tree cholla. 
+Using a system of heirarchical bayesian models we found that there were discernable differences in the effects that each partner had on vital processes of the focal mutualist. 
+The different vital rates vary in importance across tree cholla ontogeny.
+Several of them are negatively impacted by the presence and pressure of herbivores and seed predators \cite{Miller2009, Miller2006} and positively impacted by the presence of ant partners \cite{Miller2007}.
+The predators and herbivores target new growth and flowers, leading to negative impacts on the growth rates, survival rates, and floral viability rates of tree cholla \cite{Louda1995, Agrawal2004}.
+The presence of the ant partners can reduce those negative effects.
+
+It has been previously hypothesized that their is a heirarchy of partners due to the ability for \textit{L. apiculatum} ants to defend the cacti from seed predators and herbivores \cite{Miller2007}.
+These prior results would suggest that all vital rates that are affected by ant partners would be boosted the most by the presence of \textit{L. apiculatum} ants, this is not what we found. 
+Our results suggest that different partners differ significantly in their effects on vital rates.
+
+Prior to reproduction, the tree cholla experience only growth and survival.
+\textit{C. opuntiae} tended ants are associated with the highest growth rates and survival rates of plants. 
+This indicates that \textit{C. opuntiae} ants may be good ants for pre-reproductive tree cholla plants. 
+Reproducing plants experience a probability of reproducing, flower production, and floral abortion.
+\tom{Floral abortion is heavily affected by seed predators \cite{Miller2008}}{Seed predators do not influence floral abortion.}, which the ants defend the cacti, leading to increased floral viability.
+We specifically found that tree cholla experienced the highest floral viability rates when tended by \textit{L. apiculatum} ants. 
+This indicates that \textit{L. apiculatum} ants may be good partners for reproducing plants.
+
+These results together would suggest that complementarity may be the underlying mechanism that explains why partner diversity is beneficial in this system.
+
+%\subsection*{Frequency and direction of partner turnover}
+We have shown that the identity of partners is important to the processes that define tree cholla fitness. 
+Now we need to analyze the dynamics of partner turnover which dictate the identity of tree cholla partners and therefore the effects of vital rates on the tree cholla. 
+With our models we were able to identify both the direction, frequency, and distinct patterns of partner turnover.
+
+In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
+In some systems high freqency of turnover is necessary to resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
+While the purpose of this paper is not to establish which would be most beneficial in this system, we were able to identify the pattern. 
+Small plants are almost entirely vacant in this system until they grow large enough to begin producing significant amounts of EFN.
+Our model shows that once they do produce EFN, plants experience a relatively significant amount of turnover.
+Mid-sized and large plants which were either vacant or tended by other ants are most likely to become tended by \textit{L. apiculatum} ants in the next year, thereby experiencing partner turnover. 
+Plants which were tended by \textit{L. apiculatum} or \textit{C. opuntiae} ants are most likely to remain tended by the same partners multiple years in a row. 
+This indicates that \textit{C. opuntiae} ants and \textit{L. apiculatum} ants are loyal partners which retain the same plants year after year with regularity.
+
+As established in previous studies, the direction of partner turnover is important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
+In our study we found that there are distinct patterns to the direction of partner turnover. 
+Specifically, most plants are tended by \textit{L.
+apiculatum} ants at some point, because vacant plants, other tended plants, and \textit{L.
+apiculatum} tended plants are all most likely to be tended by \textit{L.
+apiculatum} ants.
+This indicates that while \textit{L. apiculatum} ants are loyal to their own plants and return multiple years in a row to the same ones, they are also strong colonizers.
+
+%\subsection*{Effect of partner diversity on plant fitness and what mechanisms explain this}
+The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (\tom{as some have done before \cite{Palmer2010}}{There is more than just the Palmer pape.
+Be sure that you are comprehensive in your use of the literaure, include non-ant-plant studies,}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
+We found that the combination of \tom{accurate partner transitions with partner identity}{Not sure what this means.} affected the fitness of the tree cholla in interesting and dynamic ways. 
+Namely, a best partner emerged in this analysis, which was surprising given the nature of our vital rate findings. 
+The variation in best partner for each vital rate suggested the potential that the different ant partners had some level of unique specialty in what they offered, which would support complementarity as the mechanism which explained the effects of partner diveristy \cite{Stachowicz2005, Stanton2003}.
+The results of our IPM however differ from this prediction.
+
+Using the stochastic IPM we developed, we found evidence of sampling effect rather than complementarity.
+We found that \textit{L. apiculatum} was the single best partner, and that all diversity scenarios where \textit{L. apiculatum} was present resulted in the highest possible fitness of tree cholla. 
+This indicates that despite the fact that \textit{L. apiculatum} partnership does not result in the highest growth and survival rates, it is still the overall best partner.
+
+Using the stochastic null IPM and the stochastic IPM we compared the fitness boost recieved by all ant partners when ants effects varied separately across years and when they did not. 
+When all ants responded to inter-annual variability the same way (shown in the stochastic null IPM) we found that the fitness boost recieved from partners was larger than the fitness boost recieved when ants responded to inter-annual variability differently.
+This indicates that having multiple possible partners benefits the tree cholla by buffering the potentially negative effects of inter-annual variation.
+
+%\subsection*{Synthesis????? All together}
+\ali{}{What this all means more broadly?? I'm currently not sure what to do with this. Tom: that's the entire Discussion!}
+This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
+\tom{Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+	Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
+	The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is really good. More of this!}
+
+%\subsection*{Future Directions}
+\tom{This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
+	We revealed the dynamics of partner turnover and showed that different ant partners are correlated with different fitness benefits.
+	As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work. }{This feels a little weak and incomplete.}
 
 \section*{Acknowledgments}
-
+This should be drafted.
 %%%%%%%%%%%%%%%%%%%%%
@@
 		%% Growth Parameters
-    growth xi intercept vacant $\Beta_{01}^g$ & -5.210899 (-5.686865, -5.491787) & \\
-    growth xi intercept other $\Beta_{02}^g$ & -5.8288 (-5.956217, 1.766021) & \\
-    growth xi intercept \textit{C. opuntiae} $\Beta_{03}^g$ & -4.529523 (-6.0770390, 0.1222112) & \\
-    growth xi intercept \textit{L. apiculatum} $\Beta_{04}^g$ & -5.106802 (-5.4499944, 0.5453901) & \\
-    growth xi size dependent vacant $\Beta_{11}^g$ & & \\
-    growth xi size dependent other $\Beta_{12}^g$ & & \\
-    growth xi size dependent \textit{C. opuntiae} $\Beta_{13}^g$ & & \\
-    growth xi size dependent \textit{L. apiculatum} $\Beta_{14}^g$ & & \\
+		growth xi intercept vacant $\beta_{01}^g$ & $-5.210899 (-5.686865, -5.491787)$ & sDE\\
+		growth xi intercept other $\beta_{02}^g$ & $-5.8288 (-5.956217, 1.766021) $&asdf \\
+		growth xi intercept \textit{C. opuntiae} $\beta_{03}^g$ & $-4.529523 (-6.0770390, 0.1222112)$ & asdf\\
+		growth xi intercept \textit{L. apiculatum} $\beta_{04}^g$ & $-5.106802 (-5.4499944, 0.5453901)$ & asdf\\
+		growth xi size dependent vacant $\beta_{11}^g$ & asdf&asdf \\
+		growth xi size dependent other $\beta_{12}^g$ & asdf&asdf \\
+		growth xi size dependent \textit{C. opuntiae} $\beta_{13}^g$ & asdf&asdf \\
+		growth xi size dependent \textit{L. apiculatum} $\beta_{14}^g$ &sadf &asdf \\
 		growth omega intercept $\omega_0^g$ & & \\
@@
 		%% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & -5.144627 (-5.370185, -4.966587) & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & -6.155088 (-6.417991, -5.786996) & \\
+		1-year germination intercept $\alpha^{\gamma_1}$ & & \\
+		2-year germination intercept $\alpha^{\gamma_2}$ & & \\
 		\hline
 		%% Survival Parameters
-    survival intercept vacant $\Beta_{01}^s$ & 0.46375959 (0.07778875, 0.86267803) & \\
-    survival intercept other $\Beta_{02}^s$ & 0.2677732 (-1.7939375, 2.5857164) & \\
-    survival intercept \textit{C.opuntiae} $\Beta_{03}^s$ & 2.2690061 (0.7347587, 3.9967755) & \\
-    survival intercept \textit{L. apiculatum} $\Beta_{04}^s$ & 0.0538502 (-1.0054015, 1.3153753) & \\
-    survival size dependent vacant $\Beta_{11}^s$ & 0.86267803 (0.2696006, 0.3217293) & \\
-    survival size dependent other $\Beta_{12}^s$ & 0.4387909 (0.1953732, 0.7325813) & \\
-    survival size dependent \textit{C. opuntiae} $\Beta_{13}^s$ & 0.28392994 (0.08742771, 0.47423484) & \\
-    survival size dependent \textit{L. apiculatum} $\Beta_{14}^s$ & 0.5150193 (0.5150193, 0.6492092) & \\
+		survival intercept vacant $\beta_{01}^s$ & & \\
+		survival intercept other $\beta_{02}^s$ & & \\
+		survival intercept \textit{C.opuntiae} $\beta_{03}^s$ & & \\
+		survival intercept \textit{L. apiculatum} $\beta_{04}^s$ & & \\
+		survival size dependent vacant $\beta_{11}^s$ & & \\
+		survival size dependent other $\beta_{12}^s$ & & \\
+		survival size dependent \textit{C. opuntiae} $\beta_{13}^s$ & & \\
+		survival size dependent \textit{L. apiculatum} $\beta_{14}^s$ & & \\
 		\hline
 		%% Probability of Flowering Parameters
-    flowering intercept $\Beta_0^f$ & -4.611414 (-5.119275, -4.082511) & \\
-    flowering size dependent $\Beta_1^f$ & 0.5693555 (0.5239602, 0.6104472) & \\
+		flowering intercept $\beta_0^f$ & & \\
+		flowering size dependent $\beta_1^f$ & & \\
 		\hline
 		%% Floral Viability Parameters
-    viability intercept vacant $\Beta_01^v$ & 1.0930266 (0.5002784, 1.6629479) & \\
-    viability intercept other $\Beta02^v$ & 1.1102242 (0.5073163, 1.7414692) & \\
-    viability intercept \textit{C. opuntiae} $\Beta_03^v$ & 1.3605548 (0.7372043, 1.9359609) & \\
-    viability intercept \textit{L. apiculatum} $\Beta_04^v$ & 2.023599 (1.461603, 2.534776) & 
+		viability intercept vacant $\beta_01^v$ & & \\
+		viability intercept other $\beta02^v$ & & \\
+		viability intercept \textit{C. opuntiae} $\beta_03^v$ & & \\
+		viability intercept \textit{L. apiculatum} $\beta_04^v$ & &
+
 	\end{tabular}
+	\caption{This table includes the median estimates, the 95$\%$ confidence intervals, and the prior distribution for each parameter in each model.}
+	\label{tab:Params}
 \end{table}
-\bigskip{}
-\\
-\end{table}
-
-\newpage{}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%
-% Videos
-%%%%%%%%%%%%%%%%%%%%%
-% If you have videos, journal style for them is generally similar to that for
-% figures.
+\section*{Supplementary Materials}
+\subsection*{Herbivory Data}
+\subsection*{Model Checks}
+For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 
+First, we checked the convergence of each parameter.
+Below we show the convergence of all $\beta$ terms listed in the Statistical Modeling subsection of Methods.
+Second, we checked the posterior fit, comparing the estimated values of each model to the $y$ values of the actual data.
+We show these posterior checks below, split by ant partner where relevant.
+%% Growth Figure
+%% Survival Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/surv_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/surv_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the survival model and the b) posterior distribution of survival estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean survival distribution (black line) of the real data.}
+	\label{fig:Surv_post}
+\end{figure}
+%% Reproduction Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/repro_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/repro_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the reproduction model and the b) posterior distribution of reproductive status estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean reproductive status distribution (black line) of the real data.}
+	\label{fig:Repro_post}
+\end{figure}
+%% Flowering Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/flow_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/flow_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the number of flowers model and the b) posterior distribution of the number of flowers estimated (pink lines) compared to the mean distribution of observed flowers (black line).}
+	\label{fig:Flow_post}
+\end{figure}
+%% Viability Figure
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/viab_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/viab_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the viability model and the b) posterior distributions of floral viability estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean floral viability distribution (black line) of the real data.}
+	\label{fig:Viab_post}
+\end{figure}
+%% Seeds Per Fruit
+\begin{figure}
+	\includegraphics[width = 0.45\linewidth]{Figures/seed_conv.png}
+	\includegraphics[width=0.45\linewidth]{Figures/seed_ant_post.png}
+	\caption{The a) posterior convergence of the parameters estimated by the seeds per fruit model and the b) posterior distributions of seeds per fruit estimates (pink lines) for each ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = vacant) compared to the mean seeds per fruit distribution (black line) of the real data.}
+	\label{fig:seed_post}
+\end{figure}
+%% Fruit survival 
+% \begin{figure}
+	%% 	\includegraphics[width = 0.45\linewidth]{Figures/                .png}
+	%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
+	% 	\caption{The a) posterior convergence of the parameters estimated by the fruit survival model and the b) posterior distributions of fruit survival estimates (pink lines) compared to the mean fruit survival distribution (black line) of the real data.}
+	% 	\label{fig:fruit_surv_post}
+	% \end{figure}
+% %% Ant Transitions
+% \begin{figure}
+	%% 	\includegraphics[width = 0.45\linewidth]{Figures/             .png}
+	%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
+	% 	\caption{The a) posterior convergence of the parameters estimated by the ant transitions model and the b) posterior distributions of next ant partners estimates (pink lines) for each previous ant species (1 = \textit{C.
+opuntiae}, 2 = \textit{L.
+apiculatum}, 3 = other, 4 = vacant) compared to the mean next ant partner distribution (black line) of the real data.}
+	% 	\label{fig:Transitions_post}
+	% %% Germination
+	\begin{figure}
+		\includegraphics[width = 0.45\linewidth]{Figures/germ1_conv.png}
+		\includegraphics[width=0.45\linewidth]{Figures/germ2_conv.png}\\
+		\includegraphics[width=0.45\linewidth]{Figures/germ1_post.png} 	\includegraphics[width=0.45\linewidth]{Figures/germ2_post.png}
+		\caption{The a-b) posterior convergence of the parameters estimated by the germination from year one seedbank and germination from year two seedbank models respectively.
+The c-d) posterior distributions of floral viability estimates (pink lines) compared to the mean germination distribution (black line) of the real data for first year germinants and second year germinants respectively.}
+		\label{fig:Germ_post}
+	\end{figure}
 
-%%%%% Include the text below if you have videos
+	%% Pre-Census Survival
+	\begin{figure}
+		\includegraphics[width = 0.45\linewidth]{Figures/seed_surv_conv.png}
+		\includegraphics[width=0.45\linewidth]{Figures/seed_surv_post.png}
+		\caption{The a) posterior convergence of the parameters estimated by the pre-census survival model and the b) posterior distribution of the pre-census survival estimated (pink lines) compared to the mean distribution of observed pre-census survival (black line).}
+		\label{fig:Pre_Surv_post}
+	\end{figure}
+	%%
 
-%%%%% Include the above if you have videos
+	\subsection*{Statistical Models -- Results}
+	Below are the results reorted of all statistical models not described in the main body of the text.
+
+	%% Reproduction Model
+	\paragraph{Reproduction Model}
+	The probability of a plant reproducing in a given year is highly size dependent. 
+	The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+
+	%% Seeds Produced
+	\paragraph{Seeds Per Flower Model}
+	Each viable flower on a plant produces between 97 and 257 seeds.
+	\tom{This number is affected by the ant partner present. 
+		\textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
+		\textit{L.
+apiculatum} tended plants produce a mean of 149 seeds per flower.
+}{These results are not consistent with Ohm and Miller, where Crem had lower seeds than Liom.
+I would check this.
+This section should also reference that paper because these are not new results.}
+	Vacant plants produce a mean of 160 seeds per flower. 
+	We are 73\% and 78\% confident that vacant plants produce more seeds per flower on average than plants tended by \textit{C. opuntiae} and \textit{L. apiculatum} ants respectively.
+
+	\tom{%% Precensus Survival\
+		\paragraph{Precensus Survival Model}
+		Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.
+
+		%% Germination
+		\paragraph{Germination Model}
+		Seeds have a significantly higher probability of germinating in year one than in year two.
+		Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
+		Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.
+
+		%% Recruit size distribution
+		New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.}{Move to an appendix. These results are not relevant for the questions at hand.}
+
+	%%%%%%%%%%%%%%%%%%%%
+	Videos
+	%%%%%%%%%%%%%%%%%%%%
+	If you have videos, journal style for them is generally similar to that for
+	figures.
+
+	%%%% Include the text below if you have videos
+
+	%%%% Include the above if you have videos
 

### 2024-03-20 Visuals Not simplified.R

@@ -6,13 +6,6 @@ size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm 
 ## Growth model
 ################################################################################
 ################################################################################
-## Visualize the outputs of the model -- trace plots to check conve4rgence, data moments to check fit
-png("grow_conv.png")
-bayesplot::mcmc_trace(fit_grow_stud,pars=c("d_0","d_size","a_0","a_size",
-                                           "beta0[1]","beta0[2]","beta0[3]","beta0[4]",
-                                           "beta1[1]","beta1[2]","beta1[3]","beta1[4]",
-                                           "beta2[1]","beta2[2]","beta2[3]","beta2[4]"))
-dev.off()
 # # Check the different quantile fits of the model to make sure not only the mean but also other quantiles fit well with the real data
 # # real data moments
 # q.fit<-matrix(NA,7,length(stan_data_grow_stud$vol))
@@ -68,499 +61,21 @@ dev.off()
 
 
 
-## Format the original data
-y_subset <- growth_data[,c("logsize_t1","ant_t", "logsize_t")]
-y_crem_subset_grow <- subset(y_subset, ant_t == "crem")
-y_liom_subset_grow <- subset(y_subset, ant_t == "liom")
-y_vac_subset_grow <- subset(y_subset, ant_t == "vacant")
-y_other_subset_grow <- subset(y_subset, ant_t == "other")
-## Size dummies for each subset
-size_crem <- seq(min(y_crem_subset_grow$logsize_t, na.rm = TRUE), max(y_crem_subset_grow$logsize_t, na.rm = TRUE), by = 0.1)
-size_liom <- seq(min(y_liom_subset_grow$logsize_t, na.rm = TRUE), max(y_liom_subset_grow$logsize_t, na.rm = TRUE), by = 0.1)
-size_other <- seq(min(y_other_subset_grow$logsize_t, na.rm = TRUE), max(y_other_subset_grow$logsize_t, na.rm = TRUE), by = 0.1)
[…]

### 2024-03-22 Manuscript/ant_cactus_AmNat_v2.tex
Manuscript Updates

@@
 	\section*{Abstract}
-\tom{}{I think this is too long for Am Nat requirements.
-Also, use ``we''.
-I think you just pasted in an abstract that you used elsewhere, so I will work on this once you write a real abstract for the Am Nat paper.}
-Mutualisms are widespread species interactions with diverse and dynamic consequences. 
-They are considered more context dependent than other species interactions, meaning there are many different factors which change the outcomes of interactions between mutualists, including partner diversity.
-Partner diversity has become a central focus in the field of mutualisms, expanding previous work from primarily pairwise to multispecies mutualisms. 
-It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. 
+% 200 words max
 The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
-
-I use the plant-ant multi-species mutualism in which, the cactus \textit{Cylindropuntia imbracata} (tree cholla) pro- duce extrafloral nectar and the ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, \textit{Forelius pruinosus}, and rarer species, provide defense from various herbivores and seed predators.
-I used 18 years of data collected from plant demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, and ant partner for all plants in 8 30$\times$30 m plots at the Sevilleta National Wildlife Refuge in central New Mexico.
-With this data I parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-I found that different ant partners had different impacts on the vital rates of the tree cholla. 
+\textit{Cylindropuntia imbracata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
+We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
+We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in the wild.
+The hierarchical models revealed that different ant partners had different impacts on the cholla vital rates. 
 Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
-With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest plant fitness, and which mechanism(s) may explain the effects of partner diversity.
-I found that the all scenarios which included the partner \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
-Results further suggest that diversity benefits in this system are driven by sampling effect , meaning \textit{L. apiculatum} ants are the "best`` partner.
-I also found that partner diversity benefits the focal mutualist in this system in the form of portfolio effect by buffering the tree cholla from the effects of inter-annual variation. 
-This study highlights how partner diversity can increase the overall benefits a focal mutualist receives. 
-It also highlights the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
+The IPM results revealed that scenarios which included \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
+This suggests that diversity benefits in this system are explained by sampling effect.
+This study highlights that partner diversity can increase the overall benefits a focal mutualist receives and the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
 
@@
 Plots were searched for new recruits each year, and these were added to the census.
-In total, the data set included 1141 unique individuals and \tom{19 plant-year observations}{It is 19 years but the number of plant-year observations should be the total number of rows in the data set}.
+In total, the data set included 1141 unique individuals and 19 year observations. 
 %These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
@@
 Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit \tom{($\kappa(a)$)}{This is correct but in the equation above it is shown as $\kappa(a')$ -- I am flagging that so you can update the equation and check that the code is consistent.} is dependent on ant state $a$.
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is dependent on ant state $a$. 
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
@@
 
-The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the \tom{proportion of flowers produced by a plant that are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$}{The way this is written here is not correct.
-Viability in year t should be a function of ants in year t.
-I would double check that this is what the model is actually doing.}.
-We fit this model to \tom{floral abortion data}{The way this is written it's unclear whether you are modeling probability of viability or probability of abortion} using a binomial distribution where trials and successes are given by the total number of flower buds and the number that are viable, respectively.
+The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the proportion of flowers produced by a plant that are viable (not aborted) in year $t$, with fixed effects of ant partner $a$ in year $t$.
+We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flower buds and the number that are viable, respectively.
 This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models.
@@
 \tom{We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.}{Worth doing!} 
-Second, we \tom{re-fit}{I think this is correct?
-The alternative would be to choose one ant state to define the year effects, or perhaps average them.} the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
+Second, we re-fit the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
 We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
@@
 \end{figure}
+
+\section*{Discussion -- Real}
+% mini abstract paragraph
+Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. 
+Here we show that partner identity can play an important role in determining both vital rates and population fitness. 
+The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
+opuniae} tended plants experiencing advantages in growth and survival when small, and \textit{L.
+apiculatum} tended plants experiencing floral viability advantages.
+The results of our stochastic IPM revealed that all partner diversity scenarios which included \textit{L.
+apiculatum} resulted in the highest possible fitness for tree cholla, suggesting that diversity benefits here are explained by sampling effect.
+The results of our stochastic null IPM revealed that there is no evidence of portfolio effect in our system. 
+These results highlight that partner diversity can increase the overall benefits of a focal mutualism and the importance of a mechanistic understanding to explain the benefits of diversity across systems.
+
+% Connect to broader patterns in the lit
+Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found sampling effect explains the benefits of diversity in our system. 
+This may have to do with the sheer number of \textit{L. apiculatum} in the system.
+None of these cited studies reported a single species with overwhelming frequency within their systems.
+In most cases of complementarity, there are relatively similar frequencies of each partner, which is not the case in our system. 
+75\% of observed ants are \textit{L. apiculatum} in our system, while only 17\% are \textit{C. opuntiae}.
+Our vital rate results suggested we may find evidence of complementarity, with \textit{C.
+opuntiae} appearing to boost growth and survial and \textit{L.
+apiculatum} appearing to boost floral viability.
+We believe there may simply not be enough \textit{C. opuntiae} in the system to have complementarity, leading instead to sampling effect driven by the overwhelming frequency of one ant.
+
+% If there is one best partner then diversity should impose opportunity costs
+In systems where there is one best partner, theory suggests that diversity should impose an opportunity cost of the form of interacting with "inferior" partners, meaning scenarios where more than just the best partner is available should result in decreased fitness\ali{}{get a citation here}.
+We don't see this pattern emerge in our results, likely due to the nature of interactions.
+
+%Turnover
+Partner turnover is likely a significant driver behind the fitness we see within cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
+In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
+In some systems high freqency of turnover is necessary for resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
+We found that some ants had higher frequency of turnover than others, specifically the other category appeared to have high frequency of turnover, rarely returning to the same plant, while \textit{C.
+opuntiae} and \textit{L.
+apiculatum} often returned.
+The direction of partner turnover is also important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
+In our study we found that there are distinct patterns to the direction of partner turnover. 
+Specifically, most plants are tended by \textit{L.
+apiculatum} ants at some point, because vacant plants, other tended plants, and \textit{L.
+apiculatum} tended plants are all most likely to be tended by \textit{L.
+apiculatum} ants.
+This indicates that while \textit{L. apiculatum} ants are loyal to their own plants and return multiple years in a row to the same ones, they are also strong colonizers.
+
+What now remains is to determine the underlying processes that lead to the emergence of these characterized patterns. 
+\ali{}{Find papers about the processes which drive these patterns in other papers -- invertase lit}
+
+% Portfolio Effect
+Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits.
+Yet, we find no evidence of portfolio effect within our system.
+When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
+In our system, we do not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure the effects of each partner on the cholla across time, allowing us to statistically quantify how the benefits recieved from each partner vary across time.
+We found that each ant partner had different effects on different vital rates across time,  indicating they exhibit different reactions to varying environments, leading us to believe that portfolio effect could be possible in this system.
+We compared the fitness of cholla under simulations where the variation in benefits across time reflected what we found in nature to siulations where each ant was forced to react the same way to temporal variation (excluding the possibility of portfolio effect).
+We found the benefits were not stronger across time when the ants were allowed to vary naturally than when they were forced to react synchronously, indicating that portfolio effect does not explain the benefits of diversity in our system.
+Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found 
+One study \cite{Dallas2021} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in real systems.
+
+% Value of Long-term data
+This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
+Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
+The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
+
+First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that the frequency of \textit{L.
+apiculatum} may be primarily responsible for the fitness of the cholla rather than any other ant partner observed.
+Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
+Second, our long-term data set allowed us to test the portfolio effect.
+This data is required, because we must show that the partners react asynchronously to temporal environmental stochasticity for portfolio effect to even be possible.
+This is a great example of how long-term data may offer more power to test underlying mechanisms in similar systems.
+Finallly, previous studies by \cite{Miller2007} hypothesized that \textit{C. opuntiae} may be a parasite based on the effects on the cholla cacti.
+Using more years of consecutive data in this system, we were able to show that \textit{C.
+opuntiae}, while the worst individual partner in terms of the overall cholla fitness (Figure \ref{fig:LambdaMeans}), increases the growth and survival rates of cacti prior to reproduction.
+The length of our data set has allowed us to dive into individual vital rates as well as overall fitness with more clarity.
+
+% Herbivory
+Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti.
+Despite this, we do not directly account for herbivory in our system.
+
+In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
+
+% Limitations
+As with any study, there are limitations to consider when interpreting the results reported here.
+These results are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
+We are able to report evidence that plants experience higher growth when associating with a particular partner, but we are not able to determine if the presence of the ant is the cause of the higher growth observed.
+It is possible, for example, that plants which attract \textit{L.
+apiculatum} through nectar composition have a different pattern of resource allocation to growth and survival and this is what we are observing rather than a direct affect of \textit{L.
+apiculatum} on the plant.
+This is a typical limitation of observational studies, but does not reduce the power or importance of our results, just the interpretation.
+However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find ***** \ali{}{Find causal studies to show that ants really do things.}.
+In the future, it would be interesting to verify the causation of our results through experimental manipulations.
+One further expansion could include nectar analysis of these cacti in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+
+Another limitation of our study is that we are simulating populations through an IPM to estimate fitness, meaning we have to extrapolate beyond observed interaction.
+In our study, small plants are rarely seen with partners, yet we report on the expected effects of each ant on the growth and survival of plants from the size of new recruits to the largest plants. 
+While we are confident in our model, there remains the small possibility of interactions which have never been observed occurring (such as a seedling tended by \textit{L. apiculatum}).
+This means that our fitness estimations are based on extrapolations of our observed data, and while we have done what we can to be confident in our model, it is possible that our assumptions that observed patterns would continue beyond observed interactions could be incorrect.
+Another example of this includes our single partner simulations.
+We have no data on what the fitness of cholla would look like if the only ant partner possible was \textit{C.
+opuntiae}, yet we offer an estimation based on the effects of this partner in a diverse system.
+The many years of data and 10467 unique plant-year observations included in our analysis allow us to have relatively high confidence in our outputs despite extrapolations.
+
+% Conclusion
+
+Discussion brainstorming:
+
+1. Mini abstract paragraph
+- synthesis of question plus answers, and overall significance
+Topic sentence:
+
+2. COnnecting to broader patterns in the literature
+- address how complimentarity or synergies are commonly reported -- why did we not see this, especially since the vital rates suggest some complimentarity
+- may involve the simulation experiments
+- some contrast with other systems where benefits of diversity have been found
+Topic sentence:
+
+3. If there is one best partner then diversity should impose opportunity costs -- why do we not see this?
+- should address how some systems can host multiple partners at a time while others (ours) can exclusive associations; for the latter diversity should only be costly if there is one best partner?
+
+4. Turnover
+- what's going on with turnover? what processes may drive the patters we see
+- what is seen elsewhere in the lit (invertase papers may be an interesting reference? showing plant ``choice'')
+- patterns could be consistent with ant and or plant choice (currently a ``black box'')
+- this might address both frequency and direction (direction may lead to some discussion of competition)
+
+5.  Portfolio effect
+- what is it, what does it predict, what did we find
+- what have other studies found? 
+Topic sentence:
+
+6. Value of long-term data
+- includes turnover, portfolio effect
+- also address how the full demographic data provides the correct insight into ant effects (eg We thought Crem was a parasite based on less complete data)
+
+7. Herbivory
+- this should underlie why these interactions are beneficial, so we need to address this
+- this connects to portfolio effect, because fluctuations in herbivores (in conjunction with ant-specific defenses against herbivores) would be why porfolio effect could occur
+
+8. Limitations
+- confounds fundamental to observational data regarding ant effects on plant demography (correlation/causation problem) -- maybe plants that attract Liom through nectar have a different pattern of allocation o growth and survival
+- cite experiments to support causal interpretation of observational patterns (ants really do stuff)
+- extrapolating effects of ants beyond the interactions that we can actually see (e.g., new recruits with Liom); and the counterfactual of a Liom-only population is another form of extrapolation
+
+9. Future directions -- maybe embed these into paragraphs above where we see opportunity to address limitations or extend the results -- probably lots we can do with herbivory
+
+Conclusion.
+
+Appendices:
+1. Herbivory
+2. Additional methods and results
+3. Posterior predictive checks and model convergence
+4. Maybe ant transition simulations
 
@@
 
-%%%%%%%%%%%%%%%%%%%%
- Videos
-%%%%%%%%%%%%%%%%%%%%
- If you have videos, journal style for them is generally similar to that for
- figures.
-
-%%%% Include the text below if you have videos
-
-%%%% Include the above if you have videos
-
-Discussion brainstorming:
-
-1. Mini abstract paragraph
-- synthesis of question plus answers, and overall significance
-Topic sentence: Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. Here we show...
-
-2. COnnecting to broader patterns in the literature
-- address how complimentarity or synergies are commonly reported -- why did we not see this, especially since the vital rates suggest some complimentarity
-- may involve the simulation experiments
-- some contrast with other systems where benefits of diversity have been found
-Topic sentence: similar studies have reported complementarity while we found sampling effect explains the benefits of diversity in our system.
-This may have to do with the sheer number of "liom" in the system.
-
-3. If there is one best partner then diversity should impose opportunity costs -- why do we not see this?
-- should address how some systems can host multiple partners at a time while others (ours) can exclusive associations; for the latter diversity should only be costly if there is one best partner?
-
-4. Turnover
-- what's going on with turnover? what processes may drive the patters we see
-- what is seen elsewhere in the lit (invertase papers may be an interesting reference? showing plant ``choice'')
-- patterns could be consistent with ant and or plant choice (currently a ``black box'')
-- this might address both frequency and direction (direction may lead to some discussion of competition)
-
-5.  Portfolio effect
-- what is it, what does it predict, what did we find
-- what have other studies found? 
-Topic sentence: Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits. Yet, we find...
-
-6. Value of long-term data
-- includes turnover, portfolio effect
-- also address how the full demographic data provides the correct insight into ant effects (eg We thought Crem was a parasite based on less complete data)
-
-7. Herbivory
-- this should underlie why these interactions are beneficial, so we need to address this
-- this connects to portfolio effect, because fluctuations in herbivores (in conjunction with ant-specific defenses against herbivores) would be why porfolio effect could occur
-
-8. Limitations
-- confounds fundamental to observational data regarding ant effects on plant demography (correlation/causation problem) -- maybe plants that attract Liom through nectar have a different pattern of allocation o growth and survival
-- cite experiments to support causal interpretation of observational patterns (ants really do stuff)
-- extrapolating effects of ants beyond the interactions that we can actually see (e.g., new recruits with Liom); and the counterfactual of a Liom-only population is another form of extrapolation
-
-9. Future directions -- maybe embed these into paragraphs above where we see opportunity to address limitations or extend the results -- probably lots we can do with herbivory
-
-Conclusion.
-
-Appendices:
-1. Herbivory
-2. Additional methods and results
-3. Posterior predictive checks and model convergence
-4. Maybe ant transition simulations
-
 \end{document}

### 2024-03-22 03_cholla_ant_IPM_params_functions.R
Additional Analyses

@@ -111,35 +111,35 @@ params$grow_beta12<-grow.params$beta1[draws,2]				## growth slope
 params$grow_beta22<-grow.params$beta2[draws,2]				## growth slope
 ## Year Random Effects
 # Ant 1 (prev crem)
-# grow_rfx1 <- cbind(grow.params$w[draws,1,1],grow.params$w[draws,1,2],grow.params$w[draws,1,3],rep(0,N_draws),rep(0,N_draws),
-#                    grow.params$w[draws,1,4],grow.params$w[draws,1,5],grow.params$w[draws,1,6],grow.params$w[draws,1,7],
-#                    grow.params$w[draws,1,8],grow.params$w[draws,1,9],grow.params$w[draws,1,10],grow.params$w[draws,1,11],
-#                    grow.params$w[draws,1,12],grow.params$w[draws,1,13],grow.params$w[draws,1,14],rep(0,N_draws),
-#                    grow.params$w[draws,1,15],grow.params$w[draws,1,16])
-# # Ant 2 (prev liom)
-# grow_rfx2 <- cbind(grow.params$w[draws,2,1],grow.params$w[draws,2,2],grow.params$w[draws,2,3],rep(0,N_draws),rep(0,N_draws),
-#                    grow.params$w[draws,2,4],grow.params$w[draws,2,5],grow.params$w[draws,2,6],grow.params$w[draws,2,7],
-#                    grow.params$w[draws,2,8],grow.params$w[draws,2,9],grow.params$w[draws,2,10],grow.params$w[draws,2,11],
-#                    grow.params$w[draws,2,12],grow.params$w[draws,2,13],grow.params$w[draws,2,14],rep(0,N_draws),
-#                    grow.params$w[draws,2,15],grow.params$w[draws,2,16])
-# # Ant 3 (prev other)
-# grow_rfx3 <- cbind(grow.params$w[draws,3,1],grow.params$w[draws,3,2],grow.params$w[draws,3,3],rep(0,N_draws),rep(0,N_draws),
-#                    grow.params$w[draws,3,4],grow.params$w[draws,3,5],grow.params$w[draws,3,6],grow.params$w[draws,3,7],
-#                    grow.params$w[draws,3,8],grow.params$w[draws,3,9],grow.params$w[draws,3,10],grow.params$w[draws,3,11],
-#                    grow.params$w[draws,3,12],grow.params$w[draws,3,13],grow.params$w[draws,3,14],rep(0,N_draws),
[…]

### 2024-03-22 05_cholla_ant_IPM_figures_tables.R

@@ -228,6 +228,93 @@ mtext("Ant-Year Effects",side=2,line=0,outer=TRUE,cex=1.7,las=0)
 dev.off()
 
 
+##### CREM FIGS 
+# Corr between crem and crem
+ccg <- cov(grow_rfx1,grow_rfx1)
+ccs <- cov(surv_rfx1,surv_rfx1)
+ccv <- cov(viab_rfx1,viab_rfx1)
+# Corr between crem and liom
+clg <- cov(grow_rfx1,grow_rfx2)
+cls <- cov(surv_rfx1,surv_rfx2)
+clv <- cov(viab_rfx1,viab_rfx2)
+# Corr between liom and liom
+llg <- cov(grow_rfx2,grow_rfx2)
+lls <- cov(surv_rfx2,surv_rfx2)
+llv <- cov(viab_rfx2,viab_rfx2)
+# Corr between crem and other
+cog <- cov(grow_rfx1,grow_rfx3)
+cos <- cov(surv_rfx1,surv_rfx3)
+cov <- cov(viab_rfx1,viab_rfx3)
+# Corr between liom and other
+log <- cov(grow_rfx2,grow_rfx3)
+los <- cov(surv_rfx2,surv_rfx3)
+lov <- cov(viab_rfx2,viab_rfx3)
+# Corr between other and other
+oog <- cov(grow_rfx3,grow_rfx3)
+oos <- cov(surv_rfx3,surv_rfx3)
+oov <- cov(viab_rfx3,viab_rfx3)
+# Corr between crem and vacant
+cvg <- cov(grow_rfx1,grow_rfx4)
+cvs <- cov(surv_rfx1,surv_rfx4)
+cvv <- cov(viab_rfx1,viab_rfx4)
+# Corr between liom and vacant
+lvg <- cov(grow_rfx2,grow_rfx4)
+lvs <- cov(surv_rfx2,surv_rfx4)
+lvv <- cov(viab_rfx2,viab_rfx4)
+# Corr between other and vacant
+ovg <- cov(grow_rfx3,grow_rfx4)
+ovs <- cov(surv_rfx3,surv_rfx4)
+ovv <- cov(viab_rfx3,viab_rfx4)
+# Corr between vacant and vacant
+vvg <- cov(grow_rfx4,grow_rfx4)
+vvs <- cov(surv_rfx4,surv_rfx4)
+vvv <- cov(viab_rfx4,viab_rfx4)
[…]

### 2024-03-27 Manuscript/ant_cactus_AmNat_v2.tex
Updates

@@
 
-\section*{Discussion}
-\tom{}{I have not commented too heavily here because I would like to talk through what we want this section to achieve.
-Currently, most of this section is actually describing results, whereas the pupose is to interpret and contextualize results, and connect them to the broader literature.
-Some of your content here is actually better at describing results than you had in the Results section, because it includes that element of ``how is this connected to the question'' that was missing from the Results section.}
-The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
-Many studies have looked at multispecies mutualisms and the how having a variety of partners leads to variation in demographic effects \cite{Palmer2010, Bascompte2019, Stachowicz2005, Ford2015, Baez2016}.
-Because these tree cholla interact with only one ant partner at a time, it is a unique system in which to parse out the individual effects of each ant partners, both in isolated settings or in combinations we cannot test in the real world.
-\tom{}{This is a good start.
-In general, I suggest opening the Discussion section with a brief summary of what you were trying to learn in this study, what you found, and its broader significance.
-This paragraph does a little of that, but I think it can be stronger.}
-
-%\subsection*{Demographic effects of association with alternative partners}
-We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, \tom{and more}{I would not say this.}) have on the vital rates of tree cholla. 
-Using a system of heirarchical bayesian models we found that there were discernable differences in the effects that each partner had on vital processes of the focal mutualist. 
-The different vital rates vary in importance across tree cholla ontogeny.
-Several of them are negatively impacted by the presence and pressure of herbivores and seed predators \cite{Miller2009, Miller2006} and positively impacted by the presence of ant partners \cite{Miller2007}.
-The predators and herbivores target new growth and flowers, leading to negative impacts on the growth rates, survival rates, and floral viability rates of tree cholla \cite{Louda1995, Agrawal2004}.
-The presence of the ant partners can reduce those negative effects.
-
-It has been previously hypothesized that their is a heirarchy of partners due to the ability for \textit{L. apiculatum} ants to defend the cacti from seed predators and herbivores \cite{Miller2007}.
-These prior results would suggest that all vital rates that are affected by ant partners would be boosted the most by the presence of \textit{L. apiculatum} ants, this is not what we found. 
-Our results suggest that different partners differ significantly in their effects on vital rates.
-
-Prior to reproduction, the tree cholla experience only growth and survival.
-\textit{C. opuntiae} tended ants are associated with the highest growth rates and survival rates of plants. 
-This indicates that \textit{C. opuntiae} ants may be good ants for pre-reproductive tree cholla plants. 
-Reproducing plants experience a probability of reproducing, flower production, and floral abortion.
-\tom{Floral abortion is heavily affected by seed predators \cite{Miller2008}}{Seed predators do not influence floral abortion.}, which the ants defend the cacti, leading to increased floral viability.
-We specifically found that tree cholla experienced the highest floral viability rates when tended by \textit{L. apiculatum} ants. 
-This indicates that \textit{L. apiculatum} ants may be good partners for reproducing plants.
-
-These results together would suggest that complementarity may be the underlying mechanism that explains why partner diversity is beneficial in this system.
-
-%\subsection*{Frequency and direction of partner turnover}
-We have shown that the identity of partners is important to the processes that define tree cholla fitness. 
-Now we need to analyze the dynamics of partner turnover which dictate the identity of tree cholla partners and therefore the effects of vital rates on the tree cholla. 
-With our models we were able to identify both the direction, frequency, and distinct patterns of partner turnover.
-
-In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
-In some systems high freqency of turnover is necessary to resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
-While the purpose of this paper is not to establish which would be most beneficial in this system, we were able to identify the pattern. 
-Small plants are almost entirely vacant in this system until they grow large enough to begin producing significant amounts of EFN.
-Our model shows that once they do produce EFN, plants experience a relatively significant amount of turnover.
-Mid-sized and large plants which were either vacant or tended by other ants are most likely to become tended by \textit{L. apiculatum} ants in the next year, thereby experiencing partner turnover. 
-Plants which were tended by \textit{L. apiculatum} or \textit{C. opuntiae} ants are most likely to remain tended by the same partners multiple years in a row. 
-This indicates that \textit{C. opuntiae} ants and \textit{L. apiculatum} ants are loyal partners which retain the same plants year after year with regularity.
-
-As established in previous studies, the direction of partner turnover is important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
-In our study we found that there are distinct patterns to the direction of partner turnover. 
-Specifically, most plants are tended by \textit{L.
-apiculatum} ants at some point, because vacant plants, other tended plants, and \textit{L.
-apiculatum} tended plants are all most likely to be tended by \textit{L.
-apiculatum} ants.
-This indicates that while \textit{L. apiculatum} ants are loyal to their own plants and return multiple years in a row to the same ones, they are also strong colonizers.
-
-%\subsection*{Effect of partner diversity on plant fitness and what mechanisms explain this}
-The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (\tom{as some have done before \cite{Palmer2010}}{There is more than just the Palmer pape.
-Be sure that you are comprehensive in your use of the literaure, include non-ant-plant studies,}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
-We found that the combination of \tom{accurate partner transitions with partner identity}{Not sure what this means.} affected the fitness of the tree cholla in interesting and dynamic ways. 
-Namely, a best partner emerged in this analysis, which was surprising given the nature of our vital rate findings. 
-The variation in best partner for each vital rate suggested the potential that the different ant partners had some level of unique specialty in what they offered, which would support complementarity as the mechanism which explained the effects of partner diveristy \cite{Stachowicz2005, Stanton2003}.
-The results of our IPM however differ from this prediction.
-
-Using the stochastic IPM we developed, we found evidence of sampling effect rather than complementarity.
-We found that \textit{L. apiculatum} was the single best partner, and that all diversity scenarios where \textit{L. apiculatum} was present resulted in the highest possible fitness of tree cholla. 
-This indicates that despite the fact that \textit{L. apiculatum} partnership does not result in the highest growth and survival rates, it is still the overall best partner.
-
-Using the stochastic null IPM and the stochastic IPM we compared the fitness boost recieved by all ant partners when ants effects varied separately across years and when they did not. 
-When all ants responded to inter-annual variability the same way (shown in the stochastic null IPM) we found that the fitness boost recieved from partners was larger than the fitness boost recieved when ants responded to inter-annual variability differently.
-This indicates that having multiple possible partners benefits the tree cholla by buffering the potentially negative effects of inter-annual variation.
-
-%\subsection*{Synthesis????? All together}
-\ali{}{What this all means more broadly?? I'm currently not sure what to do with this. Tom: that's the entire Discussion!}
-This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
-\tom{Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
-Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is really good. More of this!}
-
-%\subsection*{Future Directions}
-\tom{This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
-We revealed the dynamics of partner turnover and showed that different ant partners are correlated with different fitness benefits.
-As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work. }{This feels a little weak and incomplete.}
-
 \section*{Acknowledgments}

### 2024-03-27 02_cholla_ant_IPM_vital_rates.R

@@ -471,12 +471,16 @@ stan_data_rec <- list(N = length(seedling.dat$logsize_t1),
 ##############################################################################
 cactus_real <- cactus[,c("ant_t","ant_t1","logsize_t","Year_t","Plot")]
 cactus_real <- na.omit(cactus_real)
-cactus_real$ant_t1_relevel <- relevel(cactus_real$ant_t1,ref = "vacant")
-cactus_real$ant_t_relevel <- relevel(cactus_real$ant_t, ref = "vacant")
-cactus_real <- cactus_real[,c("ant_t_relevel","ant_t1_relevel","logsize_t", "ant_t", "ant_t1","Year_t","Plot")]
+cactus_real$ant_t1 <- relevel(cactus_real$ant_t1,ref = "vacant")
+cactus_real$ant_t <- relevel(cactus_real$ant_t, ref = "vacant")
+cactus_real <- cactus_real[,c("logsize_t", "ant_t", "ant_t1","Year_t","Plot")]
+
+unique(cactus_real$Year_t)
+cactus_real$ant_t1 <- factor(cactus_real$ant_t1, levels = c("crem","liom","other","vacant"))
+cactus_real$ant_t <- factor(cactus_real$ant_t, levels = c("crem","liom","other","vacant"))
 levels(cactus_real$ant_t)
 levels(cactus_real$ant_t1)
-unique(cactus_real$Year_t)
+(stan_data_multi$x)
 cactus_real <- subset(cactus_real, cactus_real$Year_t != 2022 & cactus_real$Year_t != 2021)
 ## make stan data set
 stan_data_multi <- list(K = length(unique(cactus_real$ant_t1)), #number of possible ant species
@@ -492,12 +496,12 @@ stan_data_multi <- list(K = length(unique(cactus_real$ant_t1)), #number of possi
 multi_model <- stan_model("Data Analysis/STAN Models/multi_mixed.stan")
 fit_multi<-sampling(multi_model, data = stan_data_multi,chains=3,
                         control = list(adapt_delta=0.99,stepsize=0.1),
-                        iter=9000,cores=3,thin=3,
+                        iter=7500,cores=3,thin=3,
                         pars = c("beta"   #location coefficients
                         ),save_warmup=F)
 ## Save the RDS file which saves all parameters, draws, and other information
[…]

### 2024-03-27 03_cholla_ant_IPM_params_functions.R

@@ -35,14 +35,14 @@ mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism proj
 #mcmc_dir <- "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 ## These files contain all draws from the posterior distributions of all parameters
 # growth model
-# fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew.rds"))
-# grow.params <- rstan::extract(fit_grow_skew)
+fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew.rds"))
+grow.params <- rstan::extract(fit_grow_skew)
 #fit_grow_skew_null<-readRDS(paste0(mcmc_dir,"fit_grow_skew_null.rds"))
 #grow.params.null <- rstan::extract(fit_grow_skew_null)
-fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
-grow.params <- rstan::extract(fit_grow_stud)
-fit_grow_stud_null<-readRDS(paste0(mcmc_dir,"fit_grow_student_t_null.rds"))
-grow.params.null <- rstan::extract(fit_grow_stud_null)
+# fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
+# grow.params <- rstan::extract(fit_grow_stud)
+# fit_grow_stud_null<-readRDS(paste0(mcmc_dir,"fit_grow_student_t_null.rds"))
+# grow.params.null <- rstan::extract(fit_grow_stud_null)
 # survival model
 fit_surv<-readRDS(paste0(mcmc_dir,"fit_surv.rds"))
 surv.params <- rstan::extract(fit_surv)
@@ -78,13 +78,13 @@ rec.params <- rstan::extract(fit_rec)
 fit_fruit<-readRDS(paste0(mcmc_dir,"fit_fruit.rds"))
 fruit.params <- rstan::extract(fit_fruit)
 # ant transitions model
-fit_multi<-readRDS(paste0(mcmc_dir,"fit_multi1.rds"))
+fit_multi<-readRDS(paste0(mcmc_dir,"fit_multi.rds"))
 multi.params <- rstan::extract(fit_multi)
 ## 'params' is a matrix where rows are vital rate coefficients and columns are posterior draws
[…]

### 2024-03-27 04_cholla_ant_IPM_analysis.R

@@ -132,7 +132,7 @@ all_stable_cut <- all_stable[3:2002]
 all_stable_stan <- all_stable_cut/sum(all_stable_cut)
 sum(all_stable_stan)
 png("Figures/stable_all_test.png")
-plot(all_y,(all_stable_stan[1:500]/sum(all_stable_stan)), col = cremcol, type = "l")#, xlim = c(-5,15), ylim = c(0,0.025))
+plot(all_y,(all_stable_stan[1:500]/sum(all_stable_stan)), col = cremcol, type = "l", xlim = c(-5,15), ylim = c(0,0.025))
 lines(all_y, all_stable_stan[501:1000]/sum(all_stable_stan), col = liomcol)
 lines(all_y, all_stable_stan[1001:1500]/sum(all_stable_stan), col = othercol)
 lines(all_y, all_stable_stan[1501:2000]/sum(all_stable_stan), col = vaccol)
@@ -168,8 +168,10 @@ vac_stable_cut <- vac_stable[3:502]
 vac_stable_stan <- vac_stable_cut/sum(vac_stable_cut)
 sum(vac_stable_stan)
 png("Figures/stable_vac_test.png")
-plot(vac_IPM$y,vac_stable_stan, type= "l", xlim = c(-5,15), col = vaccol)
+plot(vac_IPM$y,vac_stable_stan, type= "l", col = vaccol)
 legend("topleft", legend = c("Crem.","Liom.","Other","Vac."), fill = c(cremcol,liomcol,othercol,vaccol))
+abline(v = lower, col = "black", lty = 2)
+abline(v = upper, col = "black", lty = 2)
 dev.off()
 
 

### 2024-03-27 05_cholla_ant_IPM_figures_tables.R

@@ -27,7 +27,7 @@ size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm 
 ## Growth Model Visuals
 ################################################################################
 # Visualize the outputs of the model -- trace plots to check conve4rgence, data moments to check fit
-png("grow_conv.png")
+png("grow_conv_skew.png")
 bayesplot::mcmc_trace(fit_grow_skew,pars=c("d_0","d_size","a_0","a_size",
                                            "beta0[1]","beta0[2]","beta0[3]","beta0[4]",
                                            "beta1[1]","beta1[2]","beta1[3]","beta1[4]",
@@ -110,36 +110,66 @@ y_vac_mean_grow <-  quantile(grow.params$beta0[,4],0.5) + (size_dummy) * quantil
 ## Create a contour plot which shows the full fit of the growth model rather than just the mean
 x <- seq(min(cactus$logsize_t, na.rm = T),max(cactus$logsize_t,na.rm = T), length = 25); # three columns
 y <- seq(min(cactus$logsize_t1, na.rm = T),max(cactus$logsize_t1,na.rm = T), length = 25); # five rows
+# Student T Kernels
+# other <- outer (
+#   y,     # First dimension:  the columns (y)
+#   x,     # Second dimension: the rows    (x)
+#   function (x, y)   dlst(y,mu=quantile(grow.params$beta0[,3],0.5) + quantile(grow.params$beta1[,3],0.5)*x + quantile(grow.params$beta2[,3],0.5)*x^2, 
+#                          sigma = exp(quantile(grow.params$d_0,0.5) + x * quantile(grow.params$d_size,0.5)), 
+#                          df = exp(quantile(grow.params$a_0,0.5) + x * quantile(grow.params$a_size,0.5)))
+# );
+# vacant <- outer (
+#   y,     # First dimension:  the columns (y)
+#   x,     # Second dimension: the rows    (x)
+#   function (x, y)   dlst(y,mu=quantile(grow.params$beta0[,4],0.5) + quantile(grow.params$beta1[,4],0.5)*x + quantile(grow.params$beta2[,4],0.5)*x^2, 
[…]

### 2024-04-02 01_cholla_ant_IPM_setup.R
RFX Visualization

@@ -275,4 +275,6 @@ sum(camp_prop,honey_prop,phen_prop,tetra_prop,brach_prop,unk_prop,for_prop)
 
 nrow(cactus)
 
+length(unique(cactus$Year_t))
+
 

### 2024-04-02 02_cholla_ant_IPM_vital_rates.R

@@ -105,18 +105,18 @@ stan_data_grow_skew <- list(N = nrow(growth_data),                              
                             year = as.integer(as.factor(growth_data$Year_t))           ## predictor years
 )
 ## Run the growth model with a student t distribution -- fixed effects: previous size and a non linear previous size variable and ant state; random effects: plot and year; size variation is included for both the omega and alpha estimates
-grow_skew_model <- stan_model("Data Analysis/STAN Models/grow_skew.stan")
-fit_grow_skew<-sampling(grow_skew_model,data = stan_data_grow_skew,chains=3,
-                       control = list(adapt_delta=0.99,stepsize=0.1),
-                       iter=10000,cores=3,thin=2,
-                       pars = c("u","w",          # plot and year random effects
-                                "beta0","beta1", #location coefficients
-                                "d_0","d_size","d_size2", #scale coefficiences
-                                "a_0","a_size","a_size2"), #shape coefficients
-                       save_warmup=F)
-## Save the RDS file which saves all parameters, draws, and other information
-saveRDS(fit_grow_skew, "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew.rds")
-saveRDS(fit_grow_skew,"/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_grow_skew1.rds")
+# grow_skew_model <- stan_model("Data Analysis/STAN Models/grow_skew.stan")
+# fit_grow_skew<-sampling(grow_skew_model,data = stan_data_grow_skew,chains=3,
+#                        control = list(adapt_delta=0.99,stepsize=0.1),
+#                        iter=10000,cores=3,thin=2,
+#                        pars = c("u","w",          # plot and year random effects
+#                                 "beta0","beta1", #location coefficients
+#                                 "d_0","d_size","d_size2", #scale coefficiences
+#                                 "a_0","a_size","a_size2"), #shape coefficients
+#                        save_warmup=F)
+# ## Save the RDS file which saves all parameters, draws, and other information
[…]

### 2024-04-02 03_cholla_ant_IPM_params_functions.R

@@ -35,14 +35,14 @@ mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism proj
 #mcmc_dir <- "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 ## These files contain all draws from the posterior distributions of all parameters
 # growth model
-fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew1.rds"))
-grow.params <- rstan::extract(fit_grow_skew)
+# fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew1.rds"))
+# grow.params <- rstan::extract(fit_grow_skew)
 #fit_grow_skew_null<-readRDS(paste0(mcmc_dir,"fit_grow_skew_null.rds"))
 #grow.params.null <- rstan::extract(fit_grow_skew_null)
-# fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
-# grow.params <- rstan::extract(fit_grow_stud)
-# fit_grow_stud_null<-readRDS(paste0(mcmc_dir,"fit_grow_student_t_null.rds"))
-# grow.params.null <- rstan::extract(fit_grow_stud_null)
+fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
+grow.params <- rstan::extract(fit_grow_stud)
+fit_grow_stud_null<-readRDS(paste0(mcmc_dir,"fit_grow_student_t_null.rds"))
+grow.params.null <- rstan::extract(fit_grow_stud_null)
 # survival model
 fit_surv<-readRDS(paste0(mcmc_dir,"fit_surv.rds"))
 surv.params <- rstan::extract(fit_surv)

### 2024-04-02 05_cholla_ant_IPM_figures_tables.R

@@ -275,91 +275,136 @@ dev.off()
 
 ##### CREM FIGS 
 # Corr between crem and crem
-ccg <- cov(grow_rfx1,grow_rfx1)
-ccs <- cov(surv_rfx1,surv_rfx1)
-ccv <- cov(viab_rfx1,viab_rfx1)
+ccg <- cor.test(grow_rfx1,grow_rfx1)
+ccs <- cor.test(surv_rfx1,surv_rfx1)
+ccv <- cor.test(viab_rfx1,viab_rfx1)
 # Corr between crem and liom
-clg <- cov(grow_rfx1,grow_rfx2)
-cls <- cov(surv_rfx1,surv_rfx2)
-clv <- cov(viab_rfx1,viab_rfx2)
+clg <- cor.test(grow_rfx1,grow_rfx2)
+cls <- cor.test(surv_rfx1,surv_rfx2)
+clv <- cor.test(viab_rfx1,viab_rfx2)
 # Corr between liom and liom
-llg <- cov(grow_rfx2,grow_rfx2)
-lls <- cov(surv_rfx2,surv_rfx2)
-llv <- cov(viab_rfx2,viab_rfx2)
+llg <- cor.test(grow_rfx2,grow_rfx2)
+lls <- cor.test(surv_rfx2,surv_rfx2)
+llv <- cor.test(viab_rfx2,viab_rfx2)
 # Corr between crem and other
-cog <- cov(grow_rfx1,grow_rfx3)
-cos <- cov(surv_rfx1,surv_rfx3)
-cov <- cov(viab_rfx1,viab_rfx3)
+cog <- cor.test(grow_rfx1,grow_rfx3)
+cos <- cor.test(surv_rfx1,surv_rfx3)
+cov <- cor.test(viab_rfx1,viab_rfx3)
 # Corr between liom and other
-log <- cov(grow_rfx2,grow_rfx3)
-los <- cov(surv_rfx2,surv_rfx3)
-lov <- cov(viab_rfx2,viab_rfx3)
+log <- cor.test(grow_rfx2,grow_rfx3)
+los <- cor.test(surv_rfx2,surv_rfx3)
+lov <- cor.test(viab_rfx2,viab_rfx3)
 # Corr between other and other
-oog <- cov(grow_rfx3,grow_rfx3)
-oos <- cov(surv_rfx3,surv_rfx3)
-oov <- cov(viab_rfx3,viab_rfx3)
[…]

### 2024-04-02 Manuscript/ant_cactus_AmNat_v2.tex
Manuscript Edits

@@
 \usepackage{fullpage}
-\usepackage[authoryear,sectionbib,sort]{natbib}
-%\usepackage[square,sort,comma,numbers]{natbib}
+%\usepackage[authoryear,sectionbib,sort]{natbib}
+\usepackage[square,sort,comma,numbers]{natbib}
 \linespread{1.7}
@@
 Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in \tom{applied settings}{would be good if you could find another applied example to cite here} \citep{rogers2014}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
 
@@
 $$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
-The \tom{mean}{I rewrote this for the skewed normal but I cannot remember if the location parameter is the mean, so you should double check this.} of the $i$th observation $\hat{G_i}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
+The mean of the $i$th observation $\hat{G_i}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship. 
 The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vector) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
@@
 
-Additional reproductive parameters for the \tom{number of seeds per fruit}{We need to briefly explain that this was any-specific, where those values come from, and how we deal with Other.}, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
+Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
+Estimates for the number of seeds per fruit were ant specific, but the data (cite tom's data) only included \textit{L. apiculatum} and \textit{C. opuntiae} ants.
+We used the vacancy estimates for plants tended by other ants. 
+The rest of these parameters were not ant specific.
 
@@
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{19}{Is this right?
-The number of transition matrices should be one fewer than then number of years, but I don't remember what happens with 2020.} annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the 17 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces models that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
-We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \cite{rees2009integral}.
+We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \citep{Mark 2009}.
 We propagated uncertainty from the vital rate models using 1000 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
@@
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with opportunity costs of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (three-partner) scenario for synchronized and non-synchronized responses to temporal stochasticity.
-If the portfolio effect confers a benefit of diversity, $\lambda_{S}$ should be higher under non-synchronized temporal fluctuations.\tom{}{I am actually not sure whether we should look at this as stated, or whether we should compare the difference between all partners and vacancy for synchronous and non-synchronous (which is what I think you currently do.)}
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (three-partner) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
+If the portfolio effect confers a benefit of diversity, $\lambda_{S,All} - \lambda_{S_vac}$ should be higher under non-synchronized temporal fluctuations.
 
@@
 
-\tom{We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.}{Great example of an effective topic sentence for a Results paragraph.
-Sets the stage with the main result, and leaves the details for below.}
-Reproductive cacti experience between \tom{39\% and 96\%}{Unclear what this range indicates. The min to max would be 0 to 100\%.} viability rates of flowers (Figure \ref{fig:viab}).
+We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
+We observed mean viability rates of flowers between 39\% and 96\% (Figure \ref{fig:viab}).
 Ant partners influence the mean viability rate of flowers, with \textit{L.
@@
 opuntiae} and Other tended plants (at 74-75\%, Figure \ref{fig:Viab}b,c), and vacant plants had the lowest floral viability rate (71\%, Figure \ref{fig:Viab}d).
-Given the uncertainty of the parameter estimates, we are 98\% and 97\% confident that \textit{L.
-apiculatum} tended plants experience higher viability rates than plants tended by \textit{C.
-opuntiae} or other ants respectively.
-We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
+Plants tended by \textit{L. apiculatum} had floral viability advantages, while plants in states of \textit{C. opuntiae}, Other ants, and vacancy had very similar floral viability rates.
 
@@
 
-\ali{}{Should I include other model results here or just indicate that the rest are reported in an appendix/supplementary materials?}
-\tom{}{I think you can include a short paragraph that briefly mentions that other vital rates are in the appendix.}
+We analyzed all other vital rates mentioned in the methods as well.
+The results of each of these as well as the posterior predictive checks are all included in the Apendix XX.
 
-\tom{Inter-annual variation results}{I think the text you had for this did not directly address the synchrony hypothesis, and it may be more informative to look at correlation coefficients rather than the time series, which are hard to make heads or tails of.}
+We analyzed the correlation coefficients of all models which included ant state as a predictor and found that the annually varying effects of each ant on survival were the least correlated, and that the effects on growth were the most correlated.
+Low correlation indicates synchronicity, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
+Across growth, survival, and viability models, the annally varying random effects of Other ants were the most asynchronous (the least correlated with the effects of other partners), while vacant random effects were the most synchronous.
+This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed low synchronicity, particularly in the survival model.
+
 \begin{figure}[H]
@@
 \end{figure}
+\ali{}{Honestly not sure if I should include an image for this one or just report some values? We should discuss. }
 
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-We found a \tom{high frequency of partner turnover across transition years}{could we estimate the global average turnover rate?
-Across all observed transitions, what fraction of plants switched ant states.
-I think that would be a good place to begin, then unpacking that into the effects of initial size and ants.}, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
+We found that 40\% of plants experienced an ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}). 
 Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
@@
 We found ...... from the simulations with different transition probabilities.
- \ali{}{This feels like the most natural progression to me.}
 
-\tom{We found evidence}{I think this should say ``no evidence''??} of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+We found no evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
@@
 Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found sampling effect explains the benefits of diversity in our system. 
+In our first round of results, focused on individual ant effects on vital rates, we found that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
+This indicated to us that we may be dealing with complementarity, and that each ant partner offered a unique form of herbivore protection across ontogeny and or across different processes. 
+However, our IPM results showed that we were actually observing sampling effect rather than complementarity, leaving us with questions about how these results could fit together. 
 This may have to do with the sheer number of \textit{L. apiculatum} in the system.
@@
 % If there is one best partner then diversity should impose opportunity costs
-In systems where there is one best partner, theory suggests that diversity should impose an opportunity cost of the form of interacting with "inferior" partners, meaning scenarios where more than just the best partner is available should result in decreased fitness\ali{}{get a citation here}.
-We don't see this pattern emerge in our results, likely due to the nature of interactions.
+In systems where there is a single "best" partner and concurrent partner interactions, theory suggests that diversity should lead to more interactions with the best partner, thereby increasing the fitness of the focal mutualist \citep{Batstone2018}.
+In systems with mutually exclusive partner interactions and a "best`` partner,  theory suggests that diversity should impose an opportunity cost of the form of interacting with "inferior" partners, meaning scenarios where more than just the best partner is available should result in decreased fitness \citep{Noe1994}.
+While out system is aligned with this second expectation, we did not see any decreased fitness from the single "best" partner scenario to scenarios with the "best" partner and other ants included. 
+We do not see this reflected in our IPM results, potentially because of the high frequency of \textit{L. apiculatum}, the "best" partner, across the population.
+In the mutually exclusive interaction scenario, the fitness of the population should essentially be an average of the fitness conferred by each partner weighted by their frequency. 
+Due to the much higher frequency of \textit{L. apiculatum} ants, we believe the weighted average is almost equivalent to that of the single ant scenario with \textit{L. apiculatum} present.
 
@@
 What now remains is to determine the underlying processes that lead to the emergence of these characterized patterns. 
-\ali{}{Find papers about the processes which drive these patterns in other papers -- invertase lit}
+Throughout the literature, it is reported that non-symbiotic ants and non-mutualist ants are highly attracted to monosaccharide heavy EFN \citep{Heil2004,Kautz2009} and that plants often produce monosaccharide EFN to avoid exploitation \citep{Heil2010}.
+It has been previously reported that when reproducing, the nectar of stem segments has a greater disaccharide vs monosaccharide ratio, while flowerbuds have a lower ratio \citep{Miller2008}. 
+This could potentially have a significant impact on the partners who tend a plant in a given season, though determining the impacts of nectar on directional turnover requires further study.
 
@@
 Despite this, we do not directly account for herbivory in our system. 
-
+Rather, we assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
 In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
+This is particularly true in light of our consideration of portfolio effect. 
+In the same way that different species of ants likely react to temporal environmental variation differently, the same can be said for herbivores \citep{Thibaut2012}.
+Herbivores have variations in their population reflective of the environment changing temporally, meaning there are "good" and "bad" years for herbivores just like for ant defenders \citep{Batstone2018}, so the strength of their negative effect varies.
+The fluctuations in herbivore populations in conjunction with ant-specific defenses agains herbivores provide a solid foundation to study portfolio effect in this system. 
+We found no evidence of this mechanism (Figure \ref{fig:Portfolio}), but a more herbivore explicit model in the future may reveal more about driving mechanisms.
 
@@
 This is a typical limitation of observational studies, but does not reduce the power or importance of our results, just the interpretation.
-However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find ***** \ali{}{Find causal studies to show that ants really do things.}.
+However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find the presence of ants can lead to direct increases in survival \citep{Beattie1985}, reproductive efforts \citep{Ford2015}, growth \citep{Frederickson2005}, etc.
 In the future, it would be interesting to verify the causation of our results through experimental manipulations.

### 2024-04-26 Manuscript/ant_cactus_AmNat_v2.tex
Manuscript

@@
 Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
-The IPM results revealed that scenarios which included \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
-This suggests that diversity benefits in this system are explained by sampling effect.
-This study highlights that partner diversity can increase the overall benefits a focal mutualist receives and the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
+The IPM results revealed that scenario which included \textit{L. apiculatum} only resulted in the highest possible fitness for the tree cholla. 
+This suggests that there are no benefits of diversity in this system.
+This study highlights that partner diversity does not always increase the benefits a focal mutualist receives.
 
@@
 
+We used this model to estimate probabilities which are counterfactual (such as the probability of being \textit{L. apiculatum} tended or vacant with no other options). 
+To do this we assumed that the level of vacancy was constant across all states and that whatever ants were present in our simulation would tend the gaps which would otherwise be tended by another ant species.
+This assumes that competitive exclusion is the mechanism which determines the ants we see tending cacti, other potential approaches are outlined in Apendix XX.
+
 \paragraph{Parameter estimation}
@@
 
+\subsubsection*{SIMULATION IDK WHAT TO CALL THIS RN }
+Using an alternative transition model to parameterize the IPM, we conducted simulation experiments to determine if the sheer number of \textit{L.
+apiculatum} ants might be driving the outcomes of our partner diversity experiments.
+We followed the same protocol as the partner diversity experiments, except the transition model equalized the frequency of each ant partner.
+We still preserved the observed pattern of size-dependent vacancy/occupancy, but occupancy was manipulated to be equally likely for all partner identities.
+
 \subsubsection*{Temporal stochasticity experiments}
@@
 Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demoghraphic effects across host vital rates. 
-\tom{Plants tended by \textit{C.
+Plants tended by \textit{C.
 opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
-apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).}{Your draft had posterior probabilities associated with growth differences.
-I think it would be possble to keep the vital rate results very simple -- like what I have here -- and keep the posterior probability estimates for the IPM results.
-I am open to including posterior estimates for the vital rates but it is not clear to me how those values account for size, since we are comparing functions not single values.
-Let's discuss.}
+apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).
 For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean. 
@@
 
-Despite this apparent synergy, when partner identity was considered, we found the benefits of partner diversity could be explained by Sampling Effect.
-We believe the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
-All simulated combinations of ant partners which included \textit{L. apiculatum} were nearly equal and the highest possible fitness estimated for the cholla. 
-This indicates that \textit{L. apiculatum} are the single best partner for the cholla under existing conditions. 
-Based on the definitions of Sampling Effect and Complementarity we use in this study \citep{Batstone2018}, it is clear that Sampling Effect can explain the benefits of partner diversity in the cholla system.
+Despite the positive effects of ant visitation in any combination of partners, when identity of partners is considered, we found that there is an opportunity cost to partner diversity.
+There is no fitness scenario better than \textit{L. apiculatum} ants alone.
+All other scenarios lead to diminished fitness of tree cholla, indicating that there is not a benefit of diversity in this system.
+We originally believed the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
 
-It is possible that this, like the frequency of partner turnover to \textit{L. apiculatum}, is driven by the extreme frequency of \textit{L. apiculatum} ants in comparison to others. 
-With this in mind, we simulated the population fitness with equal probability for transitioning to any ant state.
-We found ...... from the simulations with different transition probabilities.
+After further simulations, we found that this lack of diversity benefits are not driven by frequency of a single ant. 
+Using the simulations where all ants had equal frequencies across sizes, we found that scenarios with \textit{C.
+opuntiae} and Other ants present experienced fitness boosts under this simulation, though not eough to offer better fitness than \textit{L.
+apiculatum}.
+Equal probability for transitioning into any ant state meant that the numbers of these ants were boosted significantly.
 
@@
 This could potentially have a significant impact on the partners who tend a plant in a given season, though determining the impacts of nectar on directional turnover requires further study. 
+We conducted alternative simulations where the driving mechanism for ant occupation was not competitive exclusion, these results are included in Apendix XX.
 
@@
 It is possible, for example, that plants which attract \textit{L.
-apiculatum} through nectar composition have a different pattern of resource allocation to growth and survival and this is what we are observing rather than a direct affect of \textit{L.
+apiculatum} through osition have a different pattern of resource allocation to growth and survival and this is what we are observing rather than a direct affect of \textit{L.
 apiculatum} on the plant.
@@
 
-% Conclusion
-
-Discussion brainstorming:
-
-1. Mini abstract paragraph
-- synthesis of question plus answers, and overall significance
-Topic sentence:
-
-2. COnnecting to broader patterns in the literature
-- address how complimentarity or synergies are commonly reported -- why did we not see this, especially since the vital rates suggest some complimentarity
-- may involve the simulation experiments
-- some contrast with other systems where benefits of diversity have been found
-Topic sentence:
-
-3. If there is one best partner then diversity should impose opportunity costs -- why do we not see this?
-- should address how some systems can host multiple partners at a time while others (ours) can exclusive associations; for the latter diversity should only be costly if there is one best partner?
-
-4. Turnover
-- what's going on with turnover? what processes may drive the patters we see
-- what is seen elsewhere in the lit (invertase papers may be an interesting reference? showing plant ``choice'')
-- patterns could be consistent with ant and or plant choice (currently a ``black box'')
-- this might address both frequency and direction (direction may lead to some discussion of competition)
-
-5.  Portfolio effect
-- what is it, what does it predict, what did we find
-- what have other studies found? 
-Topic sentence:
-
-6. Value of long-term data
-- includes turnover, portfolio effect
-- also address how the full demographic data provides the correct insight into ant effects (eg We thought Crem was a parasite based on less complete data)
-
-7. Herbivory
-- this should underlie why these interactions are beneficial, so we need to address this
-- this connects to portfolio effect, because fluctuations in herbivores (in conjunction with ant-specific defenses against herbivores) would be why porfolio effect could occur
-
-8. Limitations
-- confounds fundamental to observational data regarding ant effects on plant demography (correlation/causation problem) -- maybe plants that attract Liom through nectar have a different pattern of allocation o growth and survival
-- cite experiments to support causal interpretation of observational patterns (ants really do stuff)
-- extrapolating effects of ants beyond the interactions that we can actually see (e.g., new recruits with Liom); and the counterfactual of a Liom-only population is another form of extrapolation
-
-9. Future directions -- maybe embed these into paragraphs above where we see opportunity to address limitations or extend the results -- probably lots we can do with herbivory
-
-Conclusion.
-
-Appendices:
-1. Herbivory
-2. Additional methods and results
-3. Posterior predictive checks and model convergence
-4. Maybe ant transition simulations
-
 \section*{Acknowledgments}
@@
 This should be drafted.
-
-\section*{Appendix A: Additional Methods and Parameters}
-
-\paragraph{Seeds Per Fruit}
-With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
-We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
-$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
-%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
-$$\hat{\phi} = \beta_{0}^{\phi}$$
-The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
-
-\paragraph{Recruit Size Distribution}
-%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
-We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
-$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
-%$$\hat{\eta} = \beta_{0}^{\eta}$$
-where $\hat{\sigma}$ is estimated with a non-informative prior.
-
-\paragraph{Germination}
-With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
-These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
-$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
-$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
-%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
-%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
-
-\paragraph{Pre-Census Survival}
-With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
-We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
-$$y^{\delta} ~ Bern(\hat{\delta})$$
-%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
-where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
-
-This is not referenced in the paper, to my knowledge, and I think you need to think more deeply about what content should go into appendices and why. 
-% In most cases, authors should typeset supplementary material in a separate,
-% author-supplied PDF. For author-supplied PDFs, please consult the
-% AmNat_supp_template.tex document, available from
-% https://www.journals.uchicago.edu/journals/an/instruct 
-%
-% By contrast, the Appendix instructions below apply to cases in which
-% a brief appendix is to appear in print after the main body of the article.
-% That notably includes descriptions of methods, tables defining parameters,
-% and other material necessary for reproducing the MS's results.
-%
-% Please reset counters for the appendix (thus normally figure A1, 
-% figure A2, table A1, etc.).
-%
-% Most AmNat articles have no more than one print appendix. If your article
-% has more than one, counters for each appendix should match the letter of
-% that appendix. For example, tables in Appendix B should be numbered table B1, % table C2, etc. This applies to tables, equations, and figures.
-%
-% It's better not to use the \appendix command, because we have some
-% formatting peculiarities that \appendix conflicts with.
 
@@
 
-\section*{Supplementary Materials}
-\subsection*{Herbivory Data}
-\subsection*{Model Checks}
+\section*{Appendix A: Additional Methods and Parameters}
+In addition to the models described in the body of the paper, we fit several other simpler models using data from previous studies.
+These models are described below.
+
+\paragraph{Seeds Per Fruit}
+With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
+$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
+%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
+$$\hat{\phi} = \beta_{0}^{\phi}$$
+The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+
+\paragraph{Recruit Size Distribution}
+%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
+We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
+$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
+%$$\hat{\eta} = \beta_{0}^{\eta}$$
+where $\hat{\sigma}$ is estimated with a non-informative prior.
+
+\paragraph{Germination}
+With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
+$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
+%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
+
+\paragraph{Pre-Census Survival}
+With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
+We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
+$$y^{\delta} ~ Bern(\hat{\delta})$$
+%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
+\section*{Appendix B: Observed Herbivory Data}
+One type of data we took every year is herbivory data, which involved looking at the plant for evidence of recent herbivory (fresh bites) and noting the type and quantity of herbivores observed.
+
+\section*{Appendix C: Alternative Ant Transition Simulations}
+In addition to the competitive exclusion model and the equal probability (null) model defined and analyzed in the body of the paper, we simulated results from several other potential models. 
+We chose to include competitive exclusion as our primary results in the paper because we believe it to be the most biologically realistic.
+However, in the building and testing of alternative models we found that the method of ants occupying plants significantly impacts the fitness of the population.
+
+\paragraph{Frequency Model}
+The first alternative hypothesis we tested was what we called the frequncy model.
+In this model rather than the proportion of vacant cacti being maintained, the proportion of cacti occupied by each species is maintained and when one is removed it is replaced with vacancy.
+This version of the model assumes that the frequency of each ant we see is reflective of the real frequency of populations rather than some other mechanism.
+With this model we found that both \textit{C. opuntiae} and Other ants are parasites when they are the only partner present (Figure \ref{fig:FreqLambdaMeans}a).
+This indicates that at low frequencies, these ants are not able to provide strong benefits.
+Diversity is beneficial through a sampling effect in this simulation (Figure \ref{fig:FreqLambdaMeans}b).
+
+\begin{figure}
+\includegraphics[width=0.61\linewidth]{Figures/lambda_means_freq.png}
+\includegraphics[width=0.39\linewidth]{Figures/Lambda_Num_Partners_freq.png}
+\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+\label{fig:FreqLambdaMeans}
+\end{figure}
+
+\paragraph{In-Between Model}
+The first alternative hypothesis we tested was what we called the in-between model.
+In this model rather than only competitive exclusion or frequency based occupancy, the occupancy is in-between these two models. 
+We once again found that both \textit{C. opuntiae} and Other ants are parasites when they are the only partner present (Figure \ref{fig:BetweenLambdaMeans}a).
+Diversity is beneficial through a sampling effect in this simulation (Figure \ref{fig:BetweenLambdaMeans}b).
+
+\begin{figure}
+\includegraphics[width=0.61\linewidth]{Figures/lambda_means_between.png}
+\includegraphics[width=0.39\linewidth]{Figures/Lambda_Num_Partners_between.png}
+\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+\label{fig:BetweenLambdaMeans}
+\end{figure}
+
+\section*{Appendix D: Posterior Checks and Model Validation}
 For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 

### 2024-08-03 Manuscript/ant_cactus_AmNat_v2.tex
small edits to intro

@@
 
-Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
+Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for understanding the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
 Extrafloral nectar (EFN) -bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
@@
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
-Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
 Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
 Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
+Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}.
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
@@
 Previous studies have shown that herbivory by specialized insect herbivores negatively affects plant fitness \cite{Miller2009}, and ant defense reduces herbivore damage \cite{Miller2007}. 
-Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
-These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014, Donald2022}.
+In our study region in central New Mexico, US, tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
+These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, typically for an entire growing season \citep{Ohm2014, Donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
@@
 However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects \citep[e.g.,][]{Palmer2010}.
-To our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
+Furthermore, to our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
 
 We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity at the individual level, and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
-We used this observational data set of ant-plant associations, informed by previous ant exclusion experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
+We used this observational data set of ant-plant associations, contextualized by previous ant exclusion experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
 Specifically, we asked:

### 2024-08-05 Manuscript/ant_cactus_AmNat.tex
Updates

@@
 	\section*{Abstract}
-\tom{}{I think this is too long for Am Nat requirements.
-Also, use ``we''.
-I think you just pasted in an abstract that you used elsewhere, so I will work on this once you write a real abstract for the Am Nat paper.}
-Mutualisms are widespread species interactions with diverse and dynamic consequences. 
-They are considered more context dependent than other species interactions, meaning there are many different factors which change the outcomes of interactions between mutualists, including partner diversity.
-Partner diversity has become a central focus in the field of mutualisms, expanding previous work from primarily pairwise to multispecies mutualisms. 
-It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. 
+% 200 words max
 The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
-
-I use the plant-ant multi-species mutualism in which, the cactus \textit{Cylindropuntia imbracata} (tree cholla) pro- duce extrafloral nectar and the ants, \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, \textit{Forelius pruinosus}, and rarer species, provide defense from various herbivores and seed predators.
-I used 18 years of data collected from plant demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, and ant partner for all plants in 8 30$\times$30 m plots at the Sevilleta National Wildlife Refuge in central New Mexico.
-With this data I parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-I found that different ant partners had different impacts on the vital rates of the tree cholla. 
+\textit{Cylindropuntia imbracata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
+We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
+We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in the wild.
+The hierarchical models revealed that different ant partners had different impacts on the cholla vital rates. 
 Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
-With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest plant fitness, and which mechanism(s) may explain the effects of partner diversity.
-I found that the all scenarios which included the partner \textit{L. apiculatum} resulted in the highest possible fitness for the tree cholla. 
-Results further suggest that diversity benefits in this system are driven by sampling effect , meaning \textit{L. apiculatum} ants are the "best`` partner.
-I also found that partner diversity benefits the focal mutualist in this system in the form of portfolio effect by buffering the tree cholla from the effects of inter-annual variation. 
-This study highlights how partner diversity can increase the overall benefits a focal mutualist receives. 
-It also highlights the importance of a mechanistic understanding to explain the benefits of this diversity across systems.
+The IPM results revealed that scenario which included \textit{L. apiculatum} only resulted in the highest possible fitness for the tree cholla. 
+This suggests that there are no benefits of diversity in this system.
+This study highlights that partner diversity does not always increase the benefits a focal mutualist receives.
 
@@
 Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit \tom{($\kappa(a)$)}{This is correct but in the equation above it is shown as $\kappa(a')$ -- I am flagging that so you can update the equation and check that the code is consistent.} is dependent on ant state $a$.
-The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and subscripts indicate that deviates are specific to plants in ant state $a$.
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is dependent on ant state $a$. 
+The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
+For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability and plants in different ant states can fluctuate independently (good years for \textit{L.
+apiculatum} -occupied plants may not be good years for \textit{C.
+opuntiae}-occupied plants, for example).
 Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
@@
 \subsection*{Statistical modeling and parameter estimation}
-We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework to serve as vital rate sub-models. 
-Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively, and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
+We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework. 
+Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively (only year variation is used in the IPM), and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
 In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
-All models were fit using R version 4.2.2 and Rstan package (\cite{Rstancite, Rcite}).
+As in the IPM, our statistical modeling assumed that demographic effects of ant occupancy are limited to survival, growth, and flowerbud viability.
 
-Vital rate models for survival, growth, probability of flowering,  take the form of the function:
-$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
-\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-Some models deviate from this general model, which will be specified below.
+%Vital rate models for survival, growth, probability of flowering,  take the form of the function:
+%$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
+%\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
+%Some models deviate from this general model, which will be specified below.
 
 \paragraph{Growth}
-The growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) gives the probability of future size given the fixed effects of previous size $x$ and previous ant partner $a$. 
-We fit this model to the size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
-$$y^G \sim Skewed Normal(\hat{G},\sigma,\alpha) $$
-$$\hat{G} = \beta^G_0 + \beta^G_1 x + \beta^G_2 a + \beta^G_3 ax + \beta^G_4 x^2 + \beta^G_5 x^2a + u^G_{a} + w^G $$
-The \tom{mean}{I rewrote this for the skewed normal but I cannot remember if the location parameter is the mean, so you should double check this.} $\hat{G}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
-$u^G_{a} \sim N(0,\sigma_{year}^{2})$ is the year random effect for each ant state (which parameterizes the $\pmb{\tau}^G_{a}$ vector), and $w^G \sim N(0,\sigma_{plot}^{2})$ is the plot random effect.
-Parameters $\hat{\sigma}$ and $\hat{\alpha}$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x$.
+We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
+$$y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) $$
+$$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
+The mean of the $i$th observation $\hat{G_i}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship. 
+The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vector) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
+Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
+We assume growth variance and skewness were not dependent on ant occupancy state.
 
@@
 The survival sub-model ($S(a,x;\pmb{\tau}_{a}^{S})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of size $x$ and ant partner $a$ in year $t$.
-We fit this model to the survival data $y^S$  using a Bernoulli distribution and the logit link function, with year and plot random effects as above. 
-$$y^S \sim Bern(\hat{S})$$
-$$logit(\hat{S}) = \beta_{0}^{S} + \beta_{1}^{S}  x + \beta_{2}^{S}  a + \beta_{3}^{S}  x  a + u^S_{a} + w^S$$
+We fit this model to the survival data (alive or dead) using a Bernoulli distribution and the logit link function, with a similar linear predictor as the growth model but without the second-order influence of size.
 
 \paragraph{Reproduction}
-The flowering sub-model ($P(x';\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t+1$, with fixed effects for the size $x'$ in year $t+1$ and random effects of plot and year.
+The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects for the size $x$ and random effects of plot and year.
 We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects.  
-The flower bud function $F(x';\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t+1$, with fixed effects of size $x'$ in year $t+1$. 
-We fit this model to flowerbud count data $y^F$ (sum of viable and aborted) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects as above:
-$$y^{F} \sim NegBinom(\hat{F},\hat{\phi})$$
-$$log(\hat{F}) = \beta_{0}^{F} + \beta_{1}^{F} \times x' + u^F + w^F$$
+The flower bud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$. 
+We fit this model to flowerbud count data (sum of viable and aborted buds) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects.
 
-The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the \tom{proportion of flowers produced by a plant that are viable (not aborted) in year $t+1$, with fixed effects of ant partner $a$ in year $t$}{The way this is written here is not correct.
-Viability in year t should be a function of ants in year t.
-I would double check that this is what the model is actually doing.}.
-We fit this model to \tom{floral abortion data $y^V$}{The way this is written it's unclear whether you are modeling probability of viability or probability of abortion} using a binomial distribution with $N$ trials (total number of flower buds) and a logit link function:
-$$y^{V} \sim Binom(\hat{V},N)$$
-$$logit(\hat{V}) = \beta_{0}^{V} \times a + u + w$$
+The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the proportion of flowers produced by a plant that are viable (not aborted) in year $t$, with fixed effects of ant partner $a$ in year $t$.
+We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flower buds and the number that are viable, respectively.
+This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models.
 
 Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
+Estimates for the number of seeds per fruit were ant specific, but the data (cite tom's data) only included \textit{L. apiculatum} and \textit{C. opuntiae} ants.
+We used the vacancy estimates for plants tended by other ants. 
+The rest of these parameters were not ant specific.
 
 \paragraph{Ant Transitions}
-The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$ and random effects of plot $w$ and year $u$.
-We fit this model to ant partner data $y^{\epsilon}$ using a multinomial distribution with a logit link function: 
-$$y^{\epsilon} \sim Multinomial(\hat{\epsilon})$$
-$$logit(\epsilon) = \beta_{0}^{\epsilon} + \beta_{1}^{\epsilon} \times x + \beta_{2}^{\epsilon} \times a + u$$
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$.
+We fit this model to ant partner data using a multinomial distribution with a logit link function and random effects for plot and year.
+
+We used this model to estimate probabilities which are counterfactual (such as the probability of being \textit{L. apiculatum} tended or vacant with no other options). 
+To do this we assumed that the level of vacancy was constant across all states and that whatever ants were present in our simulation would tend the gaps which would otherwise be tended by another ant species.
+This assumes that competitive exclusion is the mechanism which determines the ants we see tending cacti, other potential approaches are outlined in Apendix XX.
 
@@
 Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
-In our model there is additional complexity in the form of transitions between ant partners, leading to a matrix size of $A b \times A b$.
-Our model also has two additional discrete states (year one and year two seed banks) leading to a final matrix size of $A(b+2) \times A(b+2)$.
-We used $b = 200$ bins after stability analysis to determine how many bins were needed to consistently estimate the same lambda.
-We extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
+In our model there is additional complexity in the form of transitions between $A$ ant partners and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
+We used $b = 200$ bins (sufficient for numerically stable model outputs) and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
 
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimated population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{19}{Is this right?
-The number of transition matrices should be one fewer than then number of years, but I don't remember what happens with 2020.} annual transition matrices, discarding the first 100 years to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the 17 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces models that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
-We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \cite{rees2009integral}.
+We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \citep{Mark 2009}.
 We propagated uncertainty from the vital rate models using 1000 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
 
-Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to test how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
-From the fullest version of the model corresponding to the observed assemblage of partners, we created model versions
+\subsubsection*{Partner diversity experiments}
+Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to quantify how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
+From the fullest version of the model corresponding to the observed assemblage of partners, we created subsets corresponding to all eight possible scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
+opuntiae} only, \textit{L.
+apiculatum} only, Other only); two partners (all pairwise combinations of \textit{C.
+opuntiae}, \textit{L.
+apiculatum}, and Other); and three partners (observed scenario of all ant states).
+These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely or never observed.
+For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
 
-In the both versions of our IPM, this process is repeated separately for every combination of ant partners: complete vacancy; \textit{C.
-opuntiae} and vacancy; other and vacancy; \textit{L.
-apiculatum}, \textit{C.
-opuntiae}, and vacancy, \textit{L.
-apiculatum}, other, and vacancy; \textit{C.
-opuntiae}, other, and vacancy; and all ant partners and vacancy.
-Each partner scenario includes vacancy in addition to the combination of partners being analyzed for added realism because we have never observed a population of cacti in which all are occupied.
+In all scenarios that included any ant partners, we preserved the observed pattern of size-dependent vacancy/occupancy (estimated through the ant transition sub-model) and manipulated partner identity conditional on occupancy.
+This means, for example, that the \textit{C. opuntiae}-only scenario included two possible states, vacancy and occupied by \textit{C. opuntiae}. 
+Small, non-reproductive plants are typically vacant because they do not produce extrafloral nectar, and once plants begin producing nectar they are nearly always ant-tended \citep{Miller2014}. 
+Our simulation experiments preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature.
 
-We constructed two versions of the stochastic, multi-state IPMs that allowed us to explore two different scenarios of environmental variation. 
-Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation.
+\subsubsection*{Temporal stochasticity experiments}
+Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation, yielding an advantage of a diverse ``portfolio'' of partners when the environment fluctuates.
+Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
+We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two different scenarios of environmental variation. 
+First, we evaluated the model using empirical estimates for the $\pmb{\tau}_{a}$ vectors that describe ant-specific year deviates. 
+In this scenario, good years and bad years can differ between ant states, according to the empirical parameter estimates. 
+\tom{We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.}{Worth doing!} 
+Second, we re-fit the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
+We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
+Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above.
 
-The first is referred to as the stochastic IPM, in which we include random annual variation in the model and allow each ant partner to respond distinctly to annual variation. 
-The second is referred to as the stochastic null IPM, in which we include random annual variation in the model, but force each ant partner to respond in the same way to annual variation. 
-By analyzing the stochastic model we can estimate the fitness of the cholla under these simulated combinations.
-By comparing the stochastic model to the stochastic null model we can identify what mechanisms may explain differences in cholla fitness with different ant partners. 
-In this section I first explain how a simple deterministic IPM is developed before diving into how we created our more complex stochastic and stochastic null models.
+\subsubsection*{Statistical inference on fitness consequences of partner identity and diversity}
+The range of models we created generated many outputs; we focus our inference on the following specific contrasts. 
+First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+Second, to determine whether each partner, in isolation, confers a fitness advantage and to rank alternative partners, we contrasted the fitness of each single partner scenario (\textit{C.
+opuntiae} only, \textit{L.
+apiculatum} only, Other only) against vacancy (zero partners).
+Third, to determine whether apparent benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario.
+If the best multi-partner scenario exceeds the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
+Alternatively, the sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
+It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with opportunity costs of diversity. 
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (three-partner) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
+If the portfolio effect confers a benefit of diversity, $\lambda_{S,All} - \lambda_{S_vac}$ should be higher under non-synchronized temporal fluctuations.
 
-We can compare the distributions of $\lambda_{S}$ in many ways, but have chosen two moments in particular: the means and the proportion of each distribution that is greater than another.
-To calculate this second moment, we can subtract each $\lambda_{S}$ distribution from the others (\ali{e.g.}{This notation may need to be revised somewhat} $\lambda_{S,Vacant}$ - $\lambda_{S,C.
-opuntiae \& Vacant}$) to determine which partner scenario leads to a higher fitness for the cactus population.
-The proportion of the resulting vector that is greater than zero allows us to give a probabilistic estimate of which combination results in the greatest fitness. 
-The magnitude of this proportion is how we determine the magnitude of difference between the resulting population fitnesses.
-
-Using only the $\lambda_S$ distributions from the stochastic IPM we can determine which scenario leads to the highest fitness, which will tell us if there are benefits to diversity in this system, any other finding indicates that there is quantitative evidence that diversity benefits this system.
-If the partner scenario which leads to the highest possible population fitness is vacancy, this indicates that there are no benefits to partner diversity.
-If all of the highest scenarios include a common ant partner, this indicates that there is a single best partner. 
-This quantitative finding that one particular partner leads to all highest possible population fitnesses is evidence that sampling effect explains the benefits of partner diversity.
-If the highest scenario includes all possible ant partners, this indicates there is a synergy which occurs when all partners interact.
-This quantitative finding indicates that complementarity is the mechanism which explains the benefits of partner diversity in this system.
-
-Using the $\lambda_S$ distribution and the $\lambda_{SN}$ distribution, from both the stochastic IPM and the stochastic null IPM, we can determine if having multiple partners can buffer the cactus population from environmental variation.
-The difference in the full partner scenario and vacancy ($\lambda_{S,All}$ - $\lambda_{S,Vacant}$) can be interpreted as the effect of having all partners when each ant species has unique effects on the cacti in response to annual variation.
-The difference in the full partner scenario and vacancy ($\lambda_{SN,All}$ - $\lambda_{SN,Vacant}$) can be interpreted as the effect of having all partners when each ant species has the same effects on the cacti in response to annual variation.
-The proportion of the resulting vectors that are greater than 0 allows us to give  probabilistic estimates of the effects of partner diversity on the fitness of the cacti in a varying environment under two ant scenarios.
-These two distributions can again be subtracted ($\lambda_{S} - \lambda_{SN}$), allowing us to determine the difference in effects of partner diversity in two annually varying scenarios.
-The proportion of the resulting vector that is greater than 0 gives us a probablistic estimate of how strongly partner diversity buffers the population from environmental variation when the effects of each partner fluctuate uniquely.
-This quantitative finding that the effects of partner diversity in the stochastic IPM are greater than those in the stochastic null IPM indicate that portfolio effect can explain some benefits of partner diversity in this system.
-
-\ali{}{I focused on several things throughout this rewrite of the methods: I tried to go straight into talking about the stochastic IPM, I focused on explaining rationale for specific numbers better, I tried to update notation to be clearer, and I tried to explain more clearly...
-not sure if I've accomplished these, let me know}
+We base our statistical inferences on the posterior probability distributions of the contrasts described above. 
+For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{S}$ ($\Delta\lambda_{C-V}$). 
+We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
+We apply similar logic to other contrasts described above.
 
@@
 \subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
-\subsubsection*{The demographic effects of alternative ant partners}
-\paragraph{Growth Model} We found evidence that ant visitation enhances cactus growth and that partner identity can influence the growth trajectory of cholla (Figure \ref{fig:Grow}). 
-Tree cholla experience positive mean growth rates across all partners at small to medium sizes with the highest chances of shrinking occurring at the largest sizes (Figure \ref{fig:Grow}e).
-Plants with \textit{C.
-opuntiae} ant partners experience the highest mean growth rates across all but the largest sizes, where they experience comprable growth to the other tended plants (Figure \ref{fig:Grow}a).
-Plants with \textit{L. apiculatum} ants experience the lowest mean growth rates at the smallest sizes and the second lowest at all other sizes (Figure \ref{fig:Grow}b). 
-Plants with other ants experience the second highest mean growth rates across all but the largest sizes, were they experience the highest growth rates (Figure \ref{fig:Grow}c.)
-Plants with no partners experience the second lowest growth rates at small sizes, after which they experience the lowest growth rates (Figure \ref{fig:Grow}d).
-
-Using the method of subtracting one size distribution from another we were able to determine how confident we are that the size of cholla visited by one combination of ant partners is greater than another.
- $G_{Crematogaster} - G_{Vacant}$ returns a vector which is 89\% positive, meaning we are 89\% confident that when \textit{C.
-opuntiae} ants are partners the size of the cholla is greater than the size of the cholla with no ant partners.
-Given the evidence that plants visited by \textit{C.
-opuntiae} experience the highest mean growth rates we have reported our confidence that this partner is associated with the highest growth rates below.
-We are 88\% and 70\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean growth rates across sizes than plants tended by \textit{L.
-spiculatum} ants or other ants respectively.
-We are 89\%, 65\%, and 94\% confident that plants with no partners experience lower mean growth rates across all sizes than plants tended by \textit{C.
-opuntiae}, \textit{L.
-apiculatum}, or other ants respectively.
+Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demoghraphic effects across host vital rates. 
+Plants tended by \textit{C.
+opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
+apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).
+For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean. 
+Similarly, ant visitation enhances cactus survival, and ant partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
+Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
+\textit{C. opuntiae}-occupied plants had a survival advantage over other ant states, particularly at smaller sizes, consistent with the positive effects on growth. 
+At larger sizes, plants in any state of ant occupancy had a survival advantage over vacant plants.
 
 \begin{figure}
-	\includegraphics[width = 0.95\linewidth]{Figures/GrowContourLinesColor.png}
+	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour.png}
 	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
@@
 
-\paragraph{Survival Model}
-We found evidence that ant visitation enhances the survival of medium to large plants, and that partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
-Tree cholla experience between 7.7\% and 99.9\% survival rates depending on 
-their size and ant partner (Figure \ref{fig:Surv}e).
-Smaller cacti all have lower survival rates, while larger cacti have higher survival rates, all nearing 100\% when they reach their largest observed sizes.
-Plants with \textit{C. opuntiae} ants experience the highest mean survival rates across all sizes (Figure \ref{fig:Surv}a).
-Plants with \textit{L. apiculatum} ants experience the lowest mean survival rates when small and the second highest mean survival rates across all other sizes (Figure \ref{fig:Surv}b). 
-Plants with other ants experience the second lowest mean survival rates across all sizes (Figure \ref{fig:Surv}c).
-Plants with no partners experience the second highest survival rates at small sizes, after which they experience the lowest survival rates (Figure \ref{fig:Surv}d).
-
-Using the method of subtracting one survival distribution from another we were able to determine how confident we are that the survival of cholla visited by one combination of ant partners is more likely than another.
-$S_{Crematogaster} - S_{Vacant}$ returns a vector which is 82\% positive, meaning we are 82\% confident that when \textit{C.
-opuntiae} ants are partners the survival of the cholla is more likely than the survival of the cholla with no ant partners.
-Given the evidence that plants visited by \textit{C.
-opuntiae} experience the highest mean survival rates we have reported our confidence that this partner is associated with the highest survival rates below.
-We are 63\% and 100\% confident that plants tended by \textit{C.
-opuntiae} ants experience higher mean survival rates across all sizes than plants tended by \textit{L.
-spiculatum} ants or other ants respectively.
-We are 82\%, 68\%, and 64\% confident that plants with no partners experience lower mean survival rates across all sizes than plants tended by
-
 \begin{figure}[H]
-\includegraphics[width=0.95\linewidth]{Figures/SurvivalPlot.png}
+\includegraphics[width=0.95\linewidth]{Figures/survival_plot.png}
 \caption{This figure shows the estimated survival rates based on the size of the cactus with each individual ant partner.
@@
 
-\paragraph{Viability Model}
-We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability.
-Tree cholla that are reproducing in year $t$ experience between 39\% and 96\% viability rates of flowers (Figure \ref{fig:viab}).
-The ant partners make a difference in the mean viability rate of flowers, with \textit{L.
-apiculatum} tended plants experiencing the highest mean viability rate (at 86\%, Figure \ref{fig:Viab}b), followed by other tended plants (at 75\%, Figure \ref{fig:Viab}c), \textit{C.
-opuntiae} tended plants (at 74\%, Figure \ref{fig:Viab}a) and vacant plants (at 71\%, Figure \ref{fig:Viab}d).
-
-Using the method of subtracting one viability distribution from another we were able to determine how confident we are that the floral viability of cholla visited by one combination of ant partners is greater than another.
-$V_{Liomatopum} - V_{Vacant}$ returns a vector which is 99\% positive, meaning we are 99\% confident that when \textit{L.
-apiculatum} ants are partners the floral viability of the cholla is greater than the cholla with no ant partners.
-Given the evidence that plants visited by \textit{L.
-apiculatum} experience the highest viability rates we have reported our confidence that this partner is associated with the highest viability rates below.
-We are 98\% and 97\% confident that \textit{L. apiculatum} tended plants experience higher viability rates than plants tended by \textit{C. opuntiae} or other ants respectively.
-We are 95\% and 69\% confident that vacant plants experience lower viability rates than plants tended by \textit{C. opuntiae} ants or other ants respectively.
+We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
+We observed mean viability rates of flowers between 39\% and 96\% (Figure \ref{fig:viab}).
+Ant partners influence the mean viability rate of flowers, with \textit{L.
+apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by \textit{C.
+opuntiae} and Other tended plants (at 74-75\%, Figure \ref{fig:Viab}b,c), and vacant plants had the lowest floral viability rate (71\%, Figure \ref{fig:Viab}d).
+Plants tended by \textit{L. apiculatum} had floral viability advantages, while plants in states of \textit{C. opuntiae}, Other ants, and vacancy had very similar floral viability rates.
 
 \begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/ViabHist.png}
+	\includegraphics[width=0.95\linewidth]{Figures/viab.png}
 	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
@@
 
-\ali{}{Should I include other model results here or just indicate that the rest are reported in an appendix/supplementary materials?}
+We analyzed all other vital rates mentioned in the methods as well.
+The results of each of these as well as the posterior predictive checks are all included in the Apendix A.
 
-\subsubsection*{The role of annual fluctuations in demographic effects of ant partners}
-We found evidence that annual variation impacts the effect of ant partner on various demographic traits (Figure \ref{fig:Annual_Ant}).
-The effects of each partner on demographic measure vary uniquely across temporal fluctuations, making partner identity important to track in conjunction with annual variation. 
-Specifically, where some ants offered greater than average benefits in one year, other species offered reduced benefits in the same year, indicating that each ant partner reacts differently to the fluctuating environment.
-
-Each demographic trait was not affected equally by the intersection of ant partner and annual fluctuations.
-We looked at the mean effect of each ant partner on growth (figure \ref{fig:Annual_Ant}a), survival (Figure \ref{fig:Annual_Ant}b), and viability (Figure \ref{fig:Annual_Ant}c) rates across all the years included in our study.
-Where the mean effect is exactly 0 there is missing ant data due to variation in censusing focus.
-This is an indirect way to analyze how temporal fluctuation impacts our system without attributing the effects to a specific climate variables. 
-Where the mean effect is positive, ant presence increases the respective estimated growth, survival, or viability rates in response to environmental fluctuations, and where the mean effect is negative, the opposite occurs.
-We found the general magnitude of variation was the smallest for the growth rate, meaning annual variation affected the growth of plants the least, followed by survival then viability.
+We analyzed the correlation coefficients of all models which included ant state as a predictor and found that the annually varying effects of each ant on survival were the least correlated, and that the effects on growth were the most correlated.
+Low correlation indicates synchronicity, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
+Across growth, survival, and viability models, the annally varying random effects of Other ants were the most asynchronous (the least correlated with the effects of other partners), while vacant random effects were the most synchronous.
+This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed low synchronicity, particularly in the survival model.
 
@@
 \end{figure}
+\ali{}{Honestly not sure if I should include an image for this one or just report some values? We should discuss. }
 
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-We found that there is a high frequency of partner turnover observed in this system with very distinct directional patterns. 
-Small plants are almost always vacant (Figure \ref{fig:Ant_Transition}b-d), if they were previously tended by a partner, they are likely to become vacant again (Figure \ref{fig:Ant_Transition}a).
-As the size of the plants increases, the probability of becoming tended increases as well, though it is not equally likely to be tended by all partners.
-
-\textit{L. apiculatum} ants become the most likely next partner in the case of most large plants, with the exception of a large plant which was previously tended by \textit{C. opuntiae}.
-Plants which were previously tended by \textit{C. opuntiae} ants are most likely to remain tended by \textit{C. opuntiae} ants.
-This indicates that these species may follow the well documented retention-discovery trade off exhibited  by many ants, with \textit{L.
-apiculatum} ants excelling at discovery and colonization of new plants and \textit{C.
-opuntiae} ants excelling at plant retention season after season.
-The overall frequency of these partners, however is a potential alternative explanation to this discovery-retention trade off.
-\textit{L.
-apiculatum} ants are by far the most frequent ant partners, accounting for 75\% of the occupied cacti, while \textit{C.
-opuntiae} and other ants account for 17\% and 8\% of occupied cacti respectively.
-This high frequency of \textit{L. apiculatum} ants may lead to the inflation of turnover to \textit{L. apiculatum} from other ants.
+We found that 40\% of plants experienced an ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}). 
+Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
+Even when small plants are ant-tended at the start of the transition year, they are most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
+The probability of becoming ant-tended increases with size, though it is not equally likely to be tended by all partners. 
+For large plants that are initially vacant or tended by \textit{L.
+apiculatum} or Other ants, \textit{L.
+apiculatum} is the most likely next partner, suggesting that this partner species is able to colonize plants that were previously vacant or occupied by Other ants, and effectively retain plants that it previously occupied.
+\textit{C.
+opuntiae} were also able to retain plants they previously occupied, but not as well as \textit{L.
+apiculatum}: for plants that begin the transition year with \textit{C.
+opuntiae}, the probability that those plants remain occupied by \textit{C.
+opuntiae} at the end of the transition year is only slightly greater than the probability of take-over by \textit{L.
+apiculatum}, while take-over in the other directino is extremely rare.
+It is also notable that transitions away from the initial state of  \textit{L.
+apiculatum} were almost always transitions to vacancy (Figure \ref{fig:Ant_Transition}d), while transitions away from the initial states of \textit{C.
+opuntiae} and Other  were often transitions to other ants.
+This suggests a competitive hierarchy whereby \textit{L. apiculatum} may abandon low-value plants with litte nectar production but is almost never displaced from high-value plants.
 
 \begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/AntSizeMulti.png}
+	\includegraphics[width=0.95\linewidth]{Figures/transition.png}
 	\caption{This figure shows the probability of being tended by each ant partner or vacant based on the size of the plant.
@@
 \subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
-We found that partner diversity was beneficial in this system. 
+By integrating vital rate results and ant transition dynamics into the multi-state IPM we can evaluate the fitness implications of different scenarios of partner diversity and identity. 
+First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
 The lowest mean fitness was $\lambda_{S,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-By subtracting the distributions $\lambda_{S, any} - \lambda_{S,Vacant}$, we found that we are between 82\% and 100\% confident that having any partner leads to a higher population fitness than having no partner.
-If you consider just the number of partners (ignoring the identity), what you find is actually that the more partners are present, the higher the fitness of the cacti (Figure \ref{fig:LambdaMeans}b), though this increase is primarily driven by the presence of \textit{L.
-apiculatum}.
+Across all 1+ partner scenarios, we are *****(double check after final runs)82--100\% confident that any scenario of ant visitation elevates fitness. 
+Furthermore, we find an apparent positive effect of partner diversity, with cactus fitness postive related to partner richness, averaging over partner identities (Figure \ref{fig:LambdaMeans}b).
 
 \begin{figure}
-	\includegraphics[width=0.61\linewidth]{Figures/LambdaMeans.png}
-	\includegraphics[width=0.39\linewidth]{Figures/Lambda_Num_Partners.png}
+	\includegraphics[width=0.61\linewidth]{Figures/comp_outputmeans_s_sn.png}
+	\includegraphics[width=0.39\linewidth]{Figures/comp_num_partners_s.png}
 	\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) and $\lambda_{SN}$ (empty circles with an X) for each simulated combination of ant partners.
@@
 
-Despite this apparent synergy, when partner identity was considered, we found the benefits of partner diversity could be explained by Sampling Effect.
-We believe the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
-All simulated combinations of ant partners which included \textit{L. apiculatum} were nearly equal and the highest possible fitness estimated for the cholla. 
-This indicates that \textit{L. apiculatum} are the single best partner for the cholla under existing conditions. 
-Based on the definitions of Sampling Effect and Complementarity we use in this study \citep{Batstone2018}, it is clear that Sampling Effect can explain the benefits of partner diversity in the cholla system.
+We originally believed the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
+However, we found that there was little benefit of any specific partner over another.
+The identity and number of partners does not appear to significantly affect the fitness of the cacti. 
+This indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.
 
-It is possible that this, like the frequency of partner turnover to \textit{L. apiculatum}, is driven by the extreme frequency of \textit{L. apiculatum} ants in comparison to others. 
-With this in mind, we simulated the population fitness with equal probability for transitioning to any ant state.
-We found ...... from the simulations with different transition probabilities.
- \ali{}{This feels like the most natural progression to me.}
+After further simulations, we found that this lack of diversity benefits are not driven by frequency of a single ant. 
+Using the simulations where all ants had equal frequencies across sizes (further explained and analyzed in Appendix C), we found the same fitness patterns as in the competitive exclusion model described above.
+Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
+Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.
 
-We found evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+We found no evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are 94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the stochastic and stochastic null model. 
-When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
-\ali{There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.}{I am not sure if I have explained enough here honestly.}
+We are are *****(double check after final runs)94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized and non-synchronized model scnearios.
+When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only *****(double check after final runs) 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
+There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.
 
@@
 
-\section*{Discussion}
-\tom{}{I have not commented too heavily here because I would like to talk through what we want this section to achieve.
-Currently, most of this section is actually describing results, whereas the pupose is to interpret and contextualize results, and connect them to the broader literature.
-Some of your content here is actually better at describing results than you had in the Results section, because it includes that element of ``how is this connected to the question'' that was missing from the Results section.}
-The large, long-lived tree cholla produce EFN which tempts several species of ant partners to protect them from herbivores and seed predators. 
-Many studies have looked at multispecies mutualisms and the how having a variety of partners leads to variation in demographic effects \cite{Palmer2010, Bascompte2019, Stachowicz2005, Ford2015, Baez2016}.
-Because these tree cholla interact with only one ant partner at a time, it is a unique system in which to parse out the individual effects of each ant partners, both in isolated settings or in combinations we cannot test in the real world.
-\tom{}{This is a good start.
-In general, I suggest opening the Discussion section with a brief summary of what you were trying to learn in this study, what you found, and its broader significance.
-This paragraph does a little of that, but I think it can be stronger.}
+\section*{Discussion}\ali{}{I am struggling reframing this part.
+I think I may need to start with a new outline, because a lot of what we had doesn't apply to our new conclusions that mutualism is good but we don't see a benefit of partner diversity?}
+% mini abstract paragraph
+Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. 
+Here we show that partner identity can play an important role in determining both vital rates and population fitness. 
+The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
+opuniae} tended plants experiencing advantages in growth and survival when small, and \textit{L.
+apiculatum} tended plants experiencing floral viability advantages.
+The results of our stochastic IPM revealed that all diversity scenarios which included partners resulted in the highest possible fitness for tree cholla, suggesting that partner diversity may not be beneficial or negative in this system.
+The results of our stochastic null IPM revealed that there is no evidence of portfolio effect in our system. 
+These results highlight that partner presence can increase the overall benefits of a focal mutualism without synergistic benefits and the importance of a mechanistic understanding to explain the benefits of diversity across systems.
+\ali{}{This may not be great, but I am struggling a bit with this framing}
 
-%\subsection*{Demographic effects of association with alternative partners}
-We asked what effects the partners which interact with tree cholla (\textit{C. opuntiae},\textit{L. apiculatum}, \tom{and more}{I would not say this.}) have on the vital rates of tree cholla. 
-Using a system of heirarchical bayesian models we found that there were discernable differences in the effects that each partner had on vital processes of the focal mutualist. 
-The different vital rates vary in importance across tree cholla ontogeny.
-Several of them are negatively impacted by the presence and pressure of herbivores and seed predators \cite{Miller2009, Miller2006} and positively impacted by the presence of ant partners \cite{Miller2007}.
-The predators and herbivores target new growth and flowers, leading to negative impacts on the growth rates, survival rates, and floral viability rates of tree cholla \cite{Louda1995, Agrawal2004}.
-The presence of the ant partners can reduce those negative effects.
+% Connect to broader patterns in the lit
+Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that neither complementarity nor sampling effect explains the benefits of diversity in our system.
+In our first round of results, focused on individual ant effects on vital rates, we found that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
+This indicated to us that we may be dealing with complementarity, and that each ant partner offered a unique form of herbivore protection across ontogeny and or across different processes. 
+However, our IPM results showed that we were actually observing a lack of beneficial diversity, leaving us with questions about how these results could fit together.
 
-It has been previously hypothesized that their is a heirarchy of partners due to the ability for \textit{L. apiculatum} ants to defend the cacti from seed predators and herbivores \cite{Miller2007}.
-These prior results would suggest that all vital rates that are affected by ant partners would be boosted the most by the presence of \textit{L. apiculatum} ants, this is not what we found. 
-Our results suggest that different partners differ significantly in their effects on vital rates.
+%This may have to do with the sheer number of \textit{L. apiculatum} in the system.
+%None of these cited studies reported a single species with overwhelming frequency within their systems.
+%In most cases of complementarity, there are relatively similar frequencies of each partner, which is not the case in our system. 
+%75\% of observed ants are \textit{L. apiculatum} in our system, while only 17\% are \textit{C. opuntiae}.
+%Our vital rate results suggested we may find evidence of complementarity, with \textit{C.
+opuntiae} appearing to boost growth and survial and \textit{L.
+apiculatum} appearing to boost floral viability.
+%We believe there may simply not be enough \textit{C. opuntiae} in the system to have complementarity, leading instead to sampling effect driven by the overwhelming frequency of one ant.
 
-Prior to reproduction, the tree cholla experience only growth and survival.
-\textit{C. opuntiae} tended ants are associated with the highest growth rates and survival rates of plants. 
-This indicates that \textit{C. opuntiae} ants may be good ants for pre-reproductive tree cholla plants. 
-Reproducing plants experience a probability of reproducing, flower production, and floral abortion.
-\tom{Floral abortion is heavily affected by seed predators \cite{Miller2008}}{Seed predators do not influence floral abortion.}, which the ants defend the cacti, leading to increased floral viability.
-We specifically found that tree cholla experienced the highest floral viability rates when tended by \textit{L. apiculatum} ants. 
-This indicates that \textit{L. apiculatum} ants may be good partners for reproducing plants.
+% If there is one best partner then diversity should impose opportunity costs
+%In systems where there is a single "best" partner and concurrent partner interactions, theory suggests that diversity should lead to more interactions with the best partner, thereby increasing the fitness of the focal mutualist \citep{Batstone2018}.
+%In systems with mutually exclusive partner interactions and a "best`` partner,  theory suggests that diversity should impose an opportunity cost of the form of interacting with "inferior" partners, meaning scenarios where more than just the best partner is available should result in decreased fitness \citep{Noe1994}.
+%While out system is aligned with this second expectation, we did not see any decreased fitness from the single "best" partner scenario to scenarios with the "best" partner and other ants included. 
+%We do not see this reflected in our IPM results, potentially because of the high frequency of \textit{L. apiculatum}, the "best" partner, across the population.
+%In the mutually exclusive interaction scenario, the fitness of the population should essentially be an average of the fitness conferred by each partner weighted by their frequency. 
+%Due to the much higher frequency of \textit{L. apiculatum} ants, we believe the weighted average is almost equivalent to that of the single ant scenario with \textit{L. apiculatum} present.
 
-These results together would suggest that complementarity may be the underlying mechanism that explains why partner diversity is beneficial in this system.
-
-%\subsection*{Frequency and direction of partner turnover}
-We have shown that the identity of partners is important to the processes that define tree cholla fitness. 
-Now we need to analyze the dynamics of partner turnover which dictate the identity of tree cholla partners and therefore the effects of vital rates on the tree cholla. 
-With our models we were able to identify both the direction, frequency, and distinct patterns of partner turnover.
-
+%Turnover
+Partner turnover is likely a significant driver behind the fitness we see within cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
 In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
-In some systems high freqency of turnover is necessary to resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
-While the purpose of this paper is not to establish which would be most beneficial in this system, we were able to identify the pattern. 
-Small plants are almost entirely vacant in this system until they grow large enough to begin producing significant amounts of EFN.
-Our model shows that once they do produce EFN, plants experience a relatively significant amount of turnover.
-Mid-sized and large plants which were either vacant or tended by other ants are most likely to become tended by \textit{L. apiculatum} ants in the next year, thereby experiencing partner turnover. 
-Plants which were tended by \textit{L. apiculatum} or \textit{C. opuntiae} ants are most likely to remain tended by the same partners multiple years in a row. 
-This indicates that \textit{C. opuntiae} ants and \textit{L. apiculatum} ants are loyal partners which retain the same plants year after year with regularity.
-
-As established in previous studies, the direction of partner turnover is important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
+In some systems high freqency of turnover is necessary for resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
+We found that some ants had higher frequency of turnover than others, specifically the other category appeared to have high frequency of turnover, rarely returning to the same plant, while \textit{C.
+opuntiae} and \textit{L.
+apiculatum} often returned.
+The direction of partner turnover is also important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
 In our study we found that there are distinct patterns to the direction of partner turnover. 
@@
 
-%\subsection*{Effect of partner diversity on plant fitness and what mechanisms explain this}
-The combination of partner identity, partner turnover, and temporal stochasticity gives us the unique power to consider both the fitness of the tree cholla under different partner scenarios (\tom{as some have done before \cite{Palmer2010}}{There is more than just the Palmer pape.
-Be sure that you are comprehensive in your use of the literaure, include non-ant-plant studies,}) and a unique set of mechanisms (\cite{Batstone2018}) which explain how the multi-partner interactions lead to fitness differences.
-We found that the combination of \tom{accurate partner transitions with partner identity}{Not sure what this means.} affected the fitness of the tree cholla in interesting and dynamic ways. 
-Namely, a best partner emerged in this analysis, which was surprising given the nature of our vital rate findings. 
-The variation in best partner for each vital rate suggested the potential that the different ant partners had some level of unique specialty in what they offered, which would support complementarity as the mechanism which explained the effects of partner diveristy \cite{Stachowicz2005, Stanton2003}.
-The results of our IPM however differ from this prediction.
+What now remains is to determine the underlying processes that lead to the emergence of these characterized patterns. 
+Throughout the literature, it is reported that non-symbiotic ants and non-mutualist ants are highly attracted to monosaccharide heavy EFN \citep{Heil2004,Kautz2009} and that plants often produce monosaccharide EFN to avoid exploitation \citep{Heil2010}.
+It has been previously reported that when reproducing, the nectar of stem segments has a greater disaccharide vs monosaccharide ratio, while flowerbuds have a lower ratio \citep{Miller2008}. 
+This could potentially have a significant impact on the partners who tend a plant in a given season, though determining the impacts of nectar on directional turnover requires further study. 
+We conducted alternative simulations where the driving mechanism for ant occupation was not competitive exclusion, these results are included in Apendix XX.
 
-Using the stochastic IPM we developed, we found evidence of sampling effect rather than complementarity.
-We found that \textit{L. apiculatum} was the single best partner, and that all diversity scenarios where \textit{L. apiculatum} was present resulted in the highest possible fitness of tree cholla. 
-This indicates that despite the fact that \textit{L. apiculatum} partnership does not result in the highest growth and survival rates, it is still the overall best partner.
+% Portfolio Effect
+Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits.
+Yet, we find no evidence of portfolio effect within our system.
+When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
+In our system, we do not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure the effects of each partner on the cholla across time, allowing us to statistically quantify how the benefits recieved from each partner vary across time.
+We found that each ant partner had different effects on different vital rates across time,  indicating they exhibit different reactions to varying environments, leading us to believe that portfolio effect could be possible in this system.
+We compared the fitness of cholla under simulations where the variation in benefits across time reflected what we found in nature to siulations where each ant was forced to react the same way to temporal variation (excluding the possibility of portfolio effect).
+We found the benefits were not stronger across time when the ants were allowed to vary naturally than when they were forced to react synchronously, indicating that portfolio effect does not explain the benefits of diversity in our system.
+Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found 
+One study \cite{Dallas2021} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in real systems.
 
-Using the stochastic null IPM and the stochastic IPM we compared the fitness boost recieved by all ant partners when ants effects varied separately across years and when they did not. 
-When all ants responded to inter-annual variability the same way (shown in the stochastic null IPM) we found that the fitness boost recieved from partners was larger than the fitness boost recieved when ants responded to inter-annual variability differently.
-This indicates that having multiple possible partners benefits the tree cholla by buffering the potentially negative effects of inter-annual variation.
+% Value of Long-term data
+This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
+Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
+The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
 
-%\subsection*{Synthesis????? All together}
-\ali{}{What this all means more broadly?? I'm currently not sure what to do with this. Tom: that's the entire Discussion!}
-This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
-\tom{Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
-Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is really good. More of this!}
+First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that the frequency of \textit{L.
+apiculatum} may be primarily responsible for the fitness of the cholla rather than any other ant partner observed.
+Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
+Second, our long-term data set allowed us to test the portfolio effect.
+This data is required, because we must show that the partners react asynchronously to temporal environmental stochasticity for portfolio effect to even be possible.
+This is a great example of how long-term data may offer more power to test underlying mechanisms in similar systems.
+Finallly, previous studies by \cite{Miller2007} hypothesized that \textit{C. opuntiae} may be a parasite based on the effects on the cholla cacti.
+Using more years of consecutive data in this system, we were able to show that \textit{C.
+opuntiae}, while the worst individual partner in terms of the overall cholla fitness (Figure \ref{fig:LambdaMeans}), increases the growth and survival rates of cacti prior to reproduction.
+The length of our data set has allowed us to dive into individual vital rates as well as overall fitness with more clarity.
 
-%\subsection*{Future Directions}
-\tom{This paper has limitations, specifically surrounding the driving forces behind the ant-plant interactions.
-We revealed the dynamics of partner turnover and showed that different ant partners are correlated with different fitness benefits.
-As of now, the driving mechanisms behind how ant species come to interact with individual plants is still unknown and could be subject to future work. }{This feels a little weak and incomplete.}
+% Herbivory
+Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti.
+Despite this, we do not directly account for herbivory in our system. 
+Rather, we assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
+In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
+This is particularly true in light of our consideration of portfolio effect. 
+In the same way that different species of ants likely react to temporal environmental variation differently, the same can be said for herbivores \citep{Thibaut2012}.
+Herbivores have variations in their population reflective of the environment changing temporally, meaning there are "good" and "bad" years for herbivores just like for ant defenders \citep{Batstone2018}, so the strength of their negative effect varies.
+The fluctuations in herbivore populations in conjunction with ant-specific defenses agains herbivores provide a solid foundation to study portfolio effect in this system. 
+We found no evidence of this mechanism (Figure \ref{fig:Portfolio}), but a more herbivore explicit model in the future may reveal more about driving mechanisms.
+
+% Limitations
+As with any study, there are limitations to consider when interpreting the results reported here.
+These results are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
+We are able to report evidence that plants experience higher growth when associating with a particular partner, but we are not able to determine if the presence of the ant is the cause of the higher growth observed.
+It is possible, for example, that plants which attract \textit{L.
+apiculatum} through osition have a different pattern of resource allocation to growth and survival and this is what we are observing rather than a direct affect of \textit{L.
+apiculatum} on the plant.
+This is a typical limitation of observational studies, but does not reduce the power or importance of our results, just the interpretation.
+However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find the presence of ants can lead to direct increases in survival \citep{Beattie1985}, reproductive efforts \citep{Ford2015}, growth \citep{Frederickson2005}, etc.
+In the future, it would be interesting to verify the causation of our results through experimental manipulations.
+One further expansion could include nectar analysis of these cacti in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+
+Another limitation of our study is that we are simulating populations through an IPM to estimate fitness, meaning we have to extrapolate beyond observed interaction.
+In our study, small plants are rarely seen with partners, yet we report on the expected effects of each ant on the growth and survival of plants from the size of new recruits to the largest plants. 
+While we are confident in our model, there remains the small possibility of interactions which have never been observed occurring (such as a seedling tended by \textit{L. apiculatum}).
+This means that our fitness estimations are based on extrapolations of our observed data, and while we have done what we can to be confident in our model, it is possible that our assumptions that observed patterns would continue beyond observed interactions could be incorrect.
+Another example of this includes our single partner simulations.
+We have no data on what the fitness of cholla would look like if the only ant partner possible was \textit{C.
+opuntiae}, yet we offer an estimation based on the effects of this partner in a diverse system.
+The many years of data and 10467 unique plant-year observations included in our analysis allow us to have relatively high confidence in our outputs despite extrapolations.
 
@@
 
-\section*{Supplementary Materials}
-\subsection*{Herbivory Data}
-\subsection*{Model Checks}
+\section*{Appendix A: Additional Methods and Parameters}
+In addition to the models described in the body of the paper, we fit several other simpler models using data from previous studies.
+These models are described below.
+
+\paragraph{Seeds Per Fruit}
+With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
+$$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
+%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
+$$\hat{\phi} = \beta_{0}^{\phi}$$
+The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+
+\paragraph{Recruit Size Distribution}
+%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
+We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
+$$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
+%$$\hat{\eta} = \beta_{0}^{\eta}$$
+where $\hat{\sigma}$ is estimated with a non-informative prior.
+
+\paragraph{Germination}
+With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
+$$y^{\gamma_1} \sim Binomial(\hat{\gamma_1})$$
+$$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
+%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
+%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
+
+\paragraph{Pre-Census Survival}
+With recruit census data \cite{Miller2006}, we fit a model for the probability of a seedling (which germinates in early Fall) surviving to when we census in May ($\delta$) of year $t+1$ (accounting for missed mortality events), with fixed effects of the previous size $x$ and random effects of the transect $m$.
+We fit this model to pre-census survival data $y^{\delta}$ using a Bernoulli distribution with a logit link function: 
+$$y^{\delta} ~ Bern(\hat{\delta})$$
+%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
+where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
+\section*{Appendix B: Observed Herbivory Data}
+One type of data we took every year is herbivory data, which involved looking at the plant for evidence of recent herbivory (fresh bites) and noting the type and quantity of herbivores observed.
+
+\section*{Appendix C: Alternative Ant Transition Simulations}
+In addition to the competitive exclusion model defined and analyzed in the body of the paper, we simulated results from several other potential models. 
+We chose to include competitive eclusion as our primary results in the paper because we believe it to be the most biologically realistic.
+However, in building and testing of alternative models we found that the method of ants occupying plants significantly impacts the fitness of the population. 
+We tested two alternative transition models, one called the frequency based model and one called the equal likelihood model.
+
+\paragraph{Frequency Based Model}
+The first alternative hypothesis we tested was what we called the frequncy based model.
+In this model rather than the proportion of vacant cacti being maintained, the proportion of cacti occupied by each species is maintained and when one is removed it is replaced with vacancy.
+This version of the model assumes that the frequency of each ant we see is reflective of the real frequency of populations rather than some other mechanism.
+With this model we found very clear evidence of Sampling Effect in the system. 
+When only  \textit{C. opuntiae}, Other ants, or both ants are present, there is very little difference in the fitness of the cacti from when no partners are present. 
+Only when \textit{L. apiculatum} ants are present do we see an increase in the fitness of the focal mutualist (Figure \ref{fig:FreqLambdaMeans}a).
+In this simulation, the more partners that are present the higher the fitness of the focal mutualist is, confirming that partner diversity would be beneficial through sampling effect if this transition model were correct.
+(Figure \ref{fig:FreqLambdaMeans}b).
+
+\begin{figure}
+\includegraphics[width=0.61\linewidth]{Figures/freq_outputmeans_s_sn.png}
+\includegraphics[width=0.39\linewidth]{Figures/freq_num_partners_s.png}
+\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+\label{fig:FreqLambdaMeans}
+\end{figure}
+
+\paragraph{Equal Likelihood Model}
+The second alternative hypothesis we tested was what we called the equal likelihood model.
+In this model we preserved the observed pattern of size-dependent vacancy/occupancy, but occupancy was manipulated to be equally likely for all partner identities. 
+This was designed to remove the effect overwhelming numbers of \textit{L. apiculatum} ants may have. 
+Despite very different proportions, we found very similar outcomes to the competitive exclusion model analyzed in the paper. 
+All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:BetweenLambdaMeans}a).
+Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:BetweenLambdaMeans}b).
+
+\begin{figure}
+\includegraphics[width=0.61\linewidth]{Figures/equal_outputmeans_s_sn.png}
+\includegraphics[width=0.39\linewidth]{Figures/equal_num_partners_s.png}
+\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
+The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+opuntiae}, L = \textit{L.
+apiculatum}, O = other).
+Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+\label{fig:BetweenLambdaMeans}
+\end{figure}
+
+\section*{Appendix D: Posterior Checks and Model Validation}
 For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 
@@
 
-%%%%%%%%%%%%%%%%%%%%
- Videos
-%%%%%%%%%%%%%%%%%%%%
- If you have videos, journal style for them is generally similar to that for
- figures.
-
-%%%% Include the text below if you have videos
-
-%%%% Include the above if you have videos
-
 \end{document}

### 2024-08-05 03_cholla_ant_IPM_params_functions.R

@@ -32,15 +32,15 @@ mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism proj
 #Tom
 #mcmc_dir <- "C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 #Lab
-mcmc_dir <- "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
+# mcmc_dir <- "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 ## These files contain all draws from the posterior distributions of all parameters
 # growth model
-mcmc_dir <- "/Users/Labuser/Desktop/Model_Reads/"
+# mcmc_dir <- "/Users/Labuser/Desktop/Model_Reads/"
 fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew.rds"))
 grow.params <- rstan::extract(fit_grow_skew, permuted = F)
 grow.params <- grow.params[,c(2),]
-fit_grow_skew_null<-readRDS(paste0(mcmc_dir,"fit_grow_skew_null.rds"))
-grow.params.null <- rstan::extract(fit_grow_skew_null)
+# fit_grow_skew_null<-readRDS(paste0(mcmc_dir,"fit_grow_skew_null.rds"))
+# grow.params.null <- rstan::extract(fit_grow_skew_null)
 # fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
 # grow.params <- rstan::extract(fit_grow_stud)
 # fit_grow_stud_null<-readRDS(paste0(mcmc_dir,"fit_grow_student_t_null.rds"))
@@ -185,16 +185,16 @@ grow_rfx4 <- cbind(grow.params[draws,"w[4,1]"],grow.params[draws,"w[4,2]"],
                    grow.params[draws,"w[4,11]"],grow.params[draws,"w[4,12]"],
                    grow.params[draws,"w[4,13]"],grow.params[draws,"w[4,14]"],
                    rep(0,N_draws),grow.params[draws,"w[4,15]"],grow.params[draws,"w[4,16]"])
-# Non ant specific -- Only run for the Stochastic Null Version
[…]

### 2024-08-05 04_cholla_ant_IPM_analysis.R

@@ -20,9 +20,10 @@ acol <- "#5d906b"
 cols <- c(vcol, ccol, lcol, ocol, lccol, locol, cocol, acol)
 
 ################################################################################
-#### Check the stable stage distribution of sizes
+#### Get Deterministic Estimates of Lambda for each ant scenario for each 
+#### ant colonizing scenario
 ################################################################################
-########## all partner scenario
+####### Set up parameters for all models
 # Pull out the mean matrix
 params_mean <- matrix(rep(0,ncol(params)), nrow = 1)
 for(i in 1:ncol(params)){
@@ -31,212 +32,573 @@ for(i in 1:ncol(params)){
 colMeans(params)
 colnames(params_mean) <- colnames(params)
 params_mean <- as.data.frame(params_mean)
+scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
+scenario_abv <- c("V","C","L","O","LC","LO","OC","LOC")
 m = 1
-all_IPM <- bigmatrix(params = (params_mean[m,]),
-                     scenario = "all",
-                     lower = cholla_min - 5,
-                     upper = cholla_max + 5,
-                     floor = floor,
-                     ceiling = ceiling,
-                     matsize = matsize,
-                     mean(grow_rfx1),mean(grow_rfx2),mean(grow_rfx3),mean(grow_rfx4),
-                     mean(surv_rfx1),mean(surv_rfx2),mean(surv_rfx3),mean(surv_rfx4),
-                     mean(flow_rfx),
-                     mean(repro_rfx),
-                     mean(viab_rfx1),mean(viab_rfx2),mean(viab_rfx3),mean(viab_rfx4))
-##View(all_IPM$Tmat)
-# Pull out the exact size distribution from the matrix above
-all_y <- all_IPM$y
-## Check that the lambda makes sense before going any further
-all_mean <- lambda(all_IPM$IPMmat) # Looks correct!
-# Pull out the IPM from the matrix above
-# all_stable <- stable.stage(all_IPM$IPMmat)
-# # Remove the seed banks
-# all_stable_cut <- all_stable[3:2002]
-# # Standardize -- NOTE THIS DOES NOT APPEAR TO HAVE ANY AFFECT ON THE PLOT
-# all_stable_stan <- all_stable_cut/sum(all_stable_cut)
-# sum(all_stable_stan)
-# png("Figures/stable_all_test.png")
[…]

### 2024-08-05 05_cholla_ant_IPM_figures_tables.R

@@ -198,7 +198,7 @@ crem <- outer (
 # dev.off()
 
 ## Plot the countour lines of the studetn t growth model with the mean fit of the model and the real data
-png("Manuscript/Figures/grow_skew.png")
+png("Manuscript/Figures/grow_contour.png")
 par(mar=c(3,3,3,1),oma=c(2,2,0,0))
 layout(matrix(c(1,2,3,4,5,5),
               ncol = 3, byrow = TRUE), heights = c(1.4,1.4), widths = c(3.9,3.9,3.9))
@@ -258,7 +258,7 @@ v_other <- colMeans((viab_rfx3))
 v_vac <- colMeans((viab_rfx4))
 v_years <- seq(2004,2022, by = 1)
 unique(viability_data$Year_t)
-png("Figures/year_ant_timeseries.png")
+png("Manuscript/Figures/year_ant_timeseries.png")
 par(mar=c(4,2,2,1),oma=c(2,2,0,0))
 layout(matrix(c(1,2,3),
               ncol = 3, byrow = TRUE), heights = c(1), widths = c(4,4,4))
@@ -355,7 +355,7 @@ a[4,3] <- ovg$estimate
 a[4,4] <- vvg$estimate
 colnames(a) <- c("Crem.","Liom.","Other","Vacant")
 rownames(a) <- c("Crem.","Liom.","Other","Vacant")
-png("Figures/corr_GRFX.png")
+png("Manuscript/Figures/corr_GRFX.png")
 corrplot(a, method = "shade", col.lim = c(0,1), addCoef.col = T)
 dev.off()
 
@@ -384,7 +384,7 @@ b[4,3] <- ovs$estimate
 b[4,4] <- vvs$estimate
 colnames(b) <- c("Crem.","Liom.","Other","Vacant")
 rownames(b) <- c("Crem.","Liom.","Other","Vacant")
-png("Figures/corr_SRFX.png")
+png("Manuscript/Figures/corr_SRFX.png")
 corrplot(b, method = "shade", col.lim = c(0,1), addCoef.col = T)
 dev.off()
 
@@ -412,7 +412,7 @@ c[4,3] <- ovv$estimate
 c[4,4] <- vvv$estimate
 colnames(c) <- c("Crem.","Liom.","Other","Vacant")
 rownames(c) <- c("Crem.","Liom.","Other","Vacant")
-png("Figures/corr_VRFX.png")
[…]

### 2024-08-08 Manuscript/ant_cactus_AmNat.tex
tiny edits to intro

@@
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
-They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
+\tom{They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.}{I suggest condensing the two sets of citations into one at the end of the sentence.
+The states that they are widespread species interactions is so broad and vague that it is not really worth citing.}
 Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
@@
 
-Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
+Partner diversity can have different effects depending on whether partners are present simultaneously or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
 Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \cite{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
@@
 Directionality of turnover can also influence effects of partner diversity if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003,Noe1994,Dejean2008}.
-For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of defensive mutualism with ants are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
+For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of a diverse guild of defensive mutualists are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
 
@@
 Extrafloral nectar (EFN) -bearing plants can serve as dietary resources that promote ant abundance and colony size \citep{Byk2011, Ness2009, Ness2006, Donald2022}.
-Presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
+In turn, presence of defensive ant partners is often linked to reductions in herbivory  \citep{Trager2010, Rudgers2004} and demographic advantages for the plant partner \citep{Baez2016}.
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
@@
 Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
-Herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
+Finally, herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
 Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
@@
 This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindriopuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
-Previous studies have shown that herbivory by specialized insect herbivores negatively affects plant fitness \cite{Miller2009}, and ant defense reduces herbivore damage \cite{Miller2007}. 
+Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \cite{Miller2009}, and that ant defense reduces herbivore damage \cite{Miller2007}. 
 Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
-These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources for an entire growing season \citep{Ohm2014, Donald2022}.
+These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects \citep[e.g.,][]{Palmer2010}.
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{L.
+apiculatum} providing strong anti-herbivore defense and \textit{C.
+opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+However, all of our previous studies in this system have focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects of a diverse partner guild \citep[e.g.,][]{Palmer2010}.
 To our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
 
-We used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity at the individual level, and how the demographic effects of alternative partner species varied across plant size structure and nearly 20 years of inter-annual fluctuations.
-We used this observational data set of ant-plant associations, informed by previous ant exclusion experiments, to ask whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation for the focal plant partner.
+Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity, and how the demographic effects of alternative partners varies across plant size structure and nearly 20 years of inter-annual fluctuations.
+We used this observational data set of plant demography and ant-plant associations, contextualized by previous ant exclusion experiments, to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
 Specifically, we asked:
@@
 	\end{enumerate}
-We used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
+To answer these questions we used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
 We then used a stochastic, multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.

### 2024-08-09 Manuscript/ant_cactus_AmNat.tex
edits through end of methods

@@
 This is an area characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B. gracilis}), yuccas, cacti, and junipers. 
-Tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA \citep{Benson1982}. 
+Tree cholla cacti are common in high Chihuahuan desert habitats of the southwestern USA \citep{Benson1982}. 
 These arborescent plants produce cylindrical segments with large spines. 
@@
 Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
-%Smaller cholla produce little to no EFN, so larger cholla, especially flowering individuals, are generally more highly tended.
+Smaller, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation at the time of our census) \citep{Miller2014}.
 
-Tree cholla EFN is harvested by various ant species. 
-At SEV-LTER, cholla are visited primarily by two species of ground-nesting ants, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species in the genera \textit{Aphaenogaster} and \textit{Camponotus}.
-\textit{L.
-apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants, followed by \textit{C.
-opuntiae} visiting between $5\% - 20\%$ of cacti depending on the year \citep{Donald2022}.
-Between $ 30\% - 80\%$ of cacti remain vacant in any given year. 
-Workers of different species rarely co-occur on individual plants, likely due to interspecific competition \citep{Miller2007}: staged introductions of \textit{C.
+Tree cholla EFN is harvested by various ground-nesting ant species. 
+At SEV-LTER, cholla are visited primarily by two species, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species in the genera \textit{Aphaenogaster} and \textit{Camponotus}.
+\textit{L. apiculatum} are the most frequent visitors with $25\% - 60\%$ of tree cholla tended by these ants in a given year, followed by \textit{C. opuntiae} ($5\% - 20\%$) \citep{Donald2022}. 
+Between $ 30\% - 80\%$ of cacti are vacant in any given year. 
+Workers of different species rarely co-occur on individual plants, likely due to interspecific competition. 
+For example, staged introductions of \textit{C.
 opuntiae} to \textit{L.
@@
 %Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
-In Fall, tree cholla stop producing EFN and the ants vacate until the next growing season.
+%In Fall, tree cholla stop producing EFN and the ants vacate until the next growing season.
 
-Multiple insect herbivores and seed predators specialize on tree cholla \citep{Mann1969}. 
-The Cerambycid beetle \textit{Moneilema appressum} and a weevil (Coleoptera: Curculionidae) of the genus \textit{Gerstaekeria} feed on vegetative and reproductive structures as adults and their larvae feed internally.
-A cactus bug, \textit{Narnia pallidicornis} (Hemiptera: Coreidae), feeds on all cholla parts with a preference for the reproductive structures \citep{Miller2006}.
+Several insect herbivores and seed predators specialize on tree cholla \citep{Mann1969}, and defense against these enemies is the main pathway by which ant visitation affects plant demography. 
+The Cerambycid beetle \textit{Moneilema appressum} and an unidentified weevil (Coleoptera: Curculionidae) of the genus \textit{Gerstaekeria} feed on vegetative and reproductive structures as adults and their larvae feed internally.
+Two species of cactus bugs, \textit{Narnia pallidicornis} and \textit{Chelinidea vittiger} (Hemiptera: Coreidae), feed on all cholla parts with a preference for flower buds; their damage can induce floral abortion \citep{Miller2006}.
 A seed predator, \textit{Cahela ponderosella} (Lepidoptera: Pyralidae), oviposits in open flowers and larvae eat seeds in developing fruits. 
-These predators can have significant negative impacts on plant fitness of and depress population growth \citep{Miller2009}.
+These consumers can have significant negative impacts on plant fitness and depress population growth \citep{Miller2009}.
 Prior experiments showed that ant-tended tree cholla experience less herbivory and seed predation than plants from which ants were excluded \citep{Miller2007,Ohm2014}.
@@
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We measured reproduction as counts of viable and aborted flowerbuds. 
+We measured reproduction by counting flowerbuds, and we distinguished between flowerbuds that were viable and aborted. 
 We recorded the ant species present (or vacancy if no ants present).
@@
 Plots were searched for new recruits each year, and these were added to the census.
-In total, the data set included 1141 unique individuals and 19 year observations. 
+These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
+In total, the data set includes a total of 9,787 observations of 1141 unique individuals across \tom{19}{There are actually 16 years (you can see my tally in the 01 script).
+In your code you took length(unique(Yeart1)).
+But one of these years is 2024 (for which there is no t1 data entered), one is NA, and 2019-2021 does not count.} complete transition years (spanning May/June of year $t$ to May/June of year $t+1$).
+\tom{}{You should cite the data package here.
+The published package only goes up to 2018, so we should talk about updating that package with you as lead author, which should be easy to do as long as the data are QAQC'd.}
 %These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
@@
 \subsection*{Multi-state Integral Projection Model}
-Integral Projection Models describe population dynamics in discrete time, with functions that relate vital rates to continuous state variables \citep{ellnerbook}. 
+The demographic data were used to parameterize a multi-state Integral Projection Model (IPM).
+IPMs describe population dynamics in discrete time with functions that relate vital rates to continuous state variables, typically size \citep{ellnerbook}. 
 While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
@@
 For example, ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
-Therefore, ant state was included in sub-models for survival, growth, and flowerbud viability. 
+Therefore, ant state was included in sub-models for survival, growth, flowerbud viability, and seed number per flowerbud. 
 In contrast, we have no reason to expect that ant tending can directly influence the probability of flowering or flowerbud production independently of its influence on plant size, so these sub-models do not include ant state as a predictor variable.
 
-We modeled the tree cholla life cycle using continuously size-structured plants where $n(x,a)_{t}$ gives the number of plants of size $x$  and ant state $a$ in year $t$, plus two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds.
+We modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates.
+The models also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
 Seed bank dynamics are given by:
@@
 \noindent %In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
-Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is dependent on ant state $a$. 
+Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is \tom{dependent on ant state $a$}{The text says ant state $a$ but the notation here and in the equation shows ant state $a'$.
+Which is it?
+(Should be $a$.)}.
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
@@
 opuntiae}-occupied plants, for example).
-Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
-Seeds are multiplied by the probability of seed dispersal and survival ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible initial ant states ($A=4$) giving total seed production. 
+Seeds are multiplied by the probability of escaping post-dispersal seed predation ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
 Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
@@
 
-For the above-ground part of the life cycle, the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) is given by survival/growth transitions from size $x$ and ant state $a$ in year $t$, plus germination out of the seed banks:
+For the above-ground part of the life cycle, the number of plants of size $x'$ and ant state $a'$ in year $t+1$ is given by survival/growth transitions from size $x$ and ant state $a$ in year $t$, plus germination out of the seed banks:
 \begin{linenomath*}
@@
 I am not sure why the equation label is not working here and I did not try to figure it out.
-It is probably something with the linenomath formatting.} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of seedling survival ($\omega$) from germination (late summer) to the census (May).
-This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ at its first appearance in our census ($\sum\rho_{0}(a')=1$). 
+It is probably something with the linenomath formatting.} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
+This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ ($\sum\rho_{0}(a')=1$). 
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
@@
 Finally, ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
-This function includes inter-annual variability through year-specific intercepts which are consistent across ant states ($\pmb{\tau}^\rho$).
+This function includes inter-annual variability through year-specific intercepts which are consistent across initial ant states ($\pmb{\tau}^\rho$).
 
@@
 We parameterized the IPM using a series of generalized linear mixed models in a hierarchical Bayesian framework. 
-Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively (only year variation is used in the IPM), and included plant size ($log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
-In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
+Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively (only year variation is used in the IPM), and included plant size (the natural logarithm of volume, $log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
+%In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
 As in the IPM, our statistical modeling assumed that demographic effects of ant occupancy are limited to survival, growth, and flowerbud viability.
@@
 \paragraph{Growth}
-We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account left-skewed size transitions:
+We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account for left-skewed size transitions (at some initial sizes, transitions below the expected future size were more common than transitions above it).
+The skew-normal has three parameters corresponding to location ($\hat{G}$), shape ($\sigma$), and skew ($\alpha$):
 $$y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) $$
 $$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
-The mean of the $i$th observation $\hat{G_i}$ is a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship. 
-The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vector) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
+Here, the location parameter for the $i$th observation $\hat{G_i}$ is defined as a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship (note that the location parameter of the skew-normal is not the mean).
+The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vectors) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
 Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
@@
 The survival sub-model ($S(a,x;\pmb{\tau}_{a}^{S})$) estimates the probability of survival from year $t$ to year $t+1$, with fixed effects of size $x$ and ant partner $a$ in year $t$.
-We fit this model to the survival data (alive or dead) using a Bernoulli distribution and the logit link function, with a similar linear predictor as the growth model but without the second-order influence of size.
+We fit this model to the survival data (alive or dead) using a Bernoulli distribution with a similar linear predictor for the probability of survival as in the growth model but with a logit link function and without the second-order influence of size.
 
 \paragraph{Reproduction}
-The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects for the size $x$ and random effects of plot and year.
+The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects size $x$ in year $t$ and random effects of plot and year.
 We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects.  
-The flower bud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$. 
+The flower bud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$ in year $t$. 
 We fit this model to flowerbud count data (sum of viable and aborted buds) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects.
 
-The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the proportion of flowers produced by a plant that are viable (not aborted) in year $t$, with fixed effects of ant partner $a$ in year $t$.
-We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flower buds and the number that are viable, respectively.
-This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models.
+The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the probability that flowerbuds produced in year $t$ remain viable (not aborted), with fixed effects of ant partner $a$ in year $t$.
+We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of \tom{flower buds}{Sometimes this is flowerbuds and sometimes it is flower buds.
+I would tighten this up.} and the number that are viable, respectively.
+This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models, with ant-specific year random effects.
 
@@
 Estimates for the number of seeds per fruit were ant specific, but the data (cite tom's data) only included \textit{L. apiculatum} and \textit{C. opuntiae} ants.
-We used the vacancy estimates for plants tended by other ants. 
-The rest of these parameters were not ant specific.
+\tom{We used the vacancy estimates for plants tended by other ants. }{This is an assumption that needs to be explained or justified. Reviewers will ask if results are sensitive to this assumption.}
+\tom{The rest of these parameters were not ant specific. }{Because the seed number model is ant-specific, I suggest moving it from the appendix to the main text.}
 
 \paragraph{Ant Transitions}
-The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$, with fixed effects of the previous size $x$  and the previous ant partner $a$  in year $t$.
+The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$ given that it was occupied by initial ant partner $a$  in year $t$, with fixed effect of initial size $x$.
 We fit this model to ant partner data using a multinomial distribution with a logit link function and random effects for plot and year.
 
-We used this model to estimate probabilities which are counterfactual (such as the probability of being \textit{L. apiculatum} tended or vacant with no other options). 
-To do this we assumed that the level of vacancy was constant across all states and that whatever ants were present in our simulation would tend the gaps which would otherwise be tended by another ant species.
-This assumes that competitive exclusion is the mechanism which determines the ants we see tending cacti, other potential approaches are outlined in Apendix XX.
-
 \paragraph{Parameter estimation}
-We fit models using Markov chain Monte Carlo (MCMC) simulations via STAN run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+We fit models using Stan run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
 We used vague priors for all parameters. 
@@
 %We did not thin the chains, thus all samples were retained following burn-in. 
-We visually assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
+\tom{We visually assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).}{These figures need to be labeled so that it is clear that they are in the Appendix.
+Am Nat's template should allow you to do this automatically.}
 %All estimated parameters are available in Table \ref{tab:Params}.
@@
 \subsection*{IPM Analysis}
-Analyzing an IPM requires discretizing the continuous IPM kernel into an approximating matrix. 
-Size variable $x$ is discretized into $b$ bins, resulting in a $b \times b$ matrix.
-In our model there is additional complexity in the form of transitions between $A$ ant partners and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
-We used $b = 200$ bins (sufficient for numerically stable model outputs) and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction \cite{Williams2012}.
+Analyzing an IPM requires that we discretize the continuous IPM kernel into an approximating matrix. 
+Size variable $x$ was discretized into $b$ bins, resulting in a $b \times b$ matrix.
+In our model there is additional complexity in the form of transitions between $A$ ant states and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
+We used $b = 200$ bins, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
 
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the 17 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
-Sampling observed transition matrices (rather than independently sampling regression coefficients) produces models that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
-We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t}}{N_{t+1}})]$ \citep{Mark 2009}.
-We propagated uncertainty from the vital rate models using 1000 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
+To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{17 annual transition matrices}{Unless you add 2024 I think this is 16.}, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
+Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
+We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$ \citep{Mark2009}.
+We propagated uncertainty from the vital rate models using \tom{1000 draws}{I did not realize you were doing this many.
+I think you could use as few as 100, which would make your numerical work much faster.} from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
 
 \subsubsection*{Partner diversity experiments}
-Using the fully parameterized multi-state IPM, we conducted simulation ``experiments'' to quantify how diversity and identity of ant partners influenced plant fitness ($\lambda_{S}$). 
-From the fullest version of the model corresponding to the observed assemblage of partners, we created subsets corresponding to all eight possible scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
+Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness in a stochastic environment ($\lambda_{S}$).
+From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created treatments corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
 opuntiae} only, \textit{L.
@@
 apiculatum}, and Other); and three partners (observed scenario of all ant states).
-These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely or never observed.
+These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed.
 For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
@@
 This means, for example, that the \textit{C. opuntiae}-only scenario included two possible states, vacancy and occupied by \textit{C. opuntiae}. 
+While our statistical models allow us to extrapolate the demographic performance of ant-tended plants to small sizes that are typically vacant, the natural history of this system tells us that this is not biologically sensible.
 Small, non-reproductive plants are typically vacant because they do not produce extrafloral nectar, and once plants begin producing nectar they are nearly always ant-tended \citep{Miller2014}. 
 Our simulation experiments preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature.
+
+The partner diversity treatment scenarios required additional assumptions about the mechanisms that give rise to observed occupancy patterns. 
+Based on evidence that EFN-bearing cacti are nearly always ant-occupied \citep{Miller2014}, we assume that ant partners competitively exclude one another from EFN-bearing cacti and that competition is zero-sum.
+This means that, in scenarios that remove species from the partner community, remaining species gain access to plants that the removed species would have tended. 
+In Appendix XX, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
+
+%\tom{We used this model to estimate probabilities which are counterfactual (such as the probability of being \textit{L. apiculatum} tended or vacant with no other options). 
+%To do this we assumed that the level of vacancy was constant across all states and that whatever ants were present in our simulation would tend the gaps which would otherwise be tended by another ant species.
+%This assumes that competitive exclusion is the mechanism which determines the ants we see tending cacti, other potential approaches are outlined in Apendix XX.}{I would move this material to the model analysis section.}
 
@@
 Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
-We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two different scenarios of environmental variation. 
+We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two scenarios of environmental variation. 
 First, we evaluated the model using empirical estimates for the $\pmb{\tau}_{a}$ vectors that describe ant-specific year deviates. 
@@
 First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+If cactus fitness increases with partner richness, this would be interpreted as evidence for benefits of partner diversity. 
 Second, to determine whether each partner, in isolation, confers a fitness advantage and to rank alternative partners, we contrasted the fitness of each single partner scenario (\textit{C.
@@
 apiculatum} only, Other only) against vacancy (zero partners).
-Third, to determine whether apparent benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario.
+Third, to determine whether any benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario. 
 If the best multi-partner scenario exceeds the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
 Alternatively, the sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
-It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with opportunity costs of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (three-partner) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
-If the portfolio effect confers a benefit of diversity, $\lambda_{S,All} - \lambda_{S_vac}$ should be higher under non-synchronized temporal fluctuations.
+It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with an opportunity cost of diversity. 
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
+If the portfolio effect confers a benefit of diversity, the fitness advantage of having all vs. no partners should be greater when temporal fluctuations are not synchronized across ant states.
 

### 2024-08-11 Manuscript/ant_cactus_AmNat.tex
refit growth function

@@
 \tom{They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.}{I suggest condensing the two sets of citations into one at the end of the sentence.
-The states that they are widespread species interactions is so broad and vague that it is not really worth citing.}
+The statement that they are widespread species interactions is so broad and vague that it is not really worth citing.}
 Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1141 unique individuals across \tom{19}{There are actually 16 years (you can see my tally in the 01 script).
-In your code you took length(unique(Yeart1)).
-But one of these years is 2024 (for which there is no t1 data entered), one is NA, and 2019-2021 does not count.} complete transition years (spanning May/June of year $t$ to May/June of year $t+1$).
+In total, the data set includes a total of 9,787 observations of 1141 unique individuals across \tom{19}{There are actually 15 complete transition years.} complete transition years (spanning May/June of year $t$ to May/June of year $t+1$).
 \tom{}{You should cite the data package here.
@@
 We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account for left-skewed size transitions (at some initial sizes, transitions below the expected future size were more common than transitions above it).
-The skew-normal has three parameters corresponding to location ($\hat{G}$), shape ($\sigma$), and skew ($\alpha$):
+The skew-normal has three parameters corresponding to location ($\hat{G}$), shape ($\sigma$), and scale ($\alpha$):
+\begin{linenomath*}
 	$$y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) $$
 	$$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
-Here, the location parameter for the $i$th observation $\hat{G_i}$ is defined as a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship (note that the location parameter of the skew-normal is not the mean).
+	$$log(\sigma_i)  = \beta^3 + \beta^4 x_i $$
+	$$\alpha_i = \beta^5 + \beta^6 x_i$$
+	\label{eqn:growth}
+\end{linenomath*}
+Here, the location parameter for the $i$th observation $\hat{G_i}$ is defined as a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship (note that the location parameter of the skew-normal is not the mean, but the mean can be derived as $\hat{G} + \frac{\sigma\alpha}{\sqrt{1+\alpha^2}} \sqrt{\frac{2}{\pi}}$).
 The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vectors) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
@@
 The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$ given that it was occupied by initial ant partner $a$  in year $t$, with fixed effect of initial size $x$.
-We fit this model to ant partner data using a multinomial distribution with a logit link function and random effects for plot and year.
+We fit this model to ant partner data using a multinomial distribution with a logit link function and \tom{random effects for plot and year}{As far as I can tell there are no plot or year random effects in the model you fit.}.
 
@@
 In our model there is additional complexity in the form of transitions between $A$ ant states and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
-We used $b = 200$ bins, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
+\tom{We used $b = 200$ bins}{In your script matsize is 500.}, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
 
@@
 \subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
-Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demoghraphic effects across host vital rates. 
+Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
 Plants tended by \textit{C.
@@
 Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found 
-One study \cite{Dallas2021} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
 This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in real systems.

### 2024-08-11 01_cholla_ant_IPM_setup.R

@@ -10,6 +10,8 @@
 ## Load All Necessary Packages Here
 # stan models
 library(rstan)
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
 library(sn)
 library(sgt)
 library(qgam)
@@ -39,7 +41,6 @@ library(ggpubr)
 library(reshape)
 library(reshape2)
 knitr::opts_chunk$set(echo = TRUE)
-options(mc.cores = parallel::detectCores())
 # check the workign directory and make sure it is set to the right location
 #getwd()
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
@@ -285,9 +286,8 @@ length(unique(cactus$Year_t1))
 cactus %>% filter(Year_t1 == Year_t +1) %>% group_by(Year_t1) %>% summarise(n())
 ## get total number of transition-year observations, excluding the last, incomplete transition
 cactus %>% filter(Year_t!=2023) %>% summarise(n())
-## it looks like there is a 2019-2021 transition year -- which is a problem!
-cactus %>% filter(Year_t==2019)
-#
+
+
 
 ## Get the proportions of each species
 # how many ant observations are there

### 2024-08-11 02_cholla_ant_IPM_vital_rates.R

@@ -87,6 +87,10 @@ stan_data_grow_stud <- list(N = nrow(growth_data),                              
 ## Pull all necessary variables together, remove NAs, and put them into a list so they are ready to feed into the stan model
 growth_data_orig <- cactus[,c("Plot","Year_t","logsize_t","logsize_t1","ant_t")]
 growth_data <- na.omit(growth_data_orig) # Lose 2032 rows (due to plant death & recruit status)
+#check the years and plots we are working with
+unique(growth_data$Year_t);table(growth_data$Year_t)
+unique(growth_data$Plot);table(growth_data$Plot)
+
 # nrow(growth_data_orig)
 # nrow(growth_data)
 # check that you are happy with the subsetting by plotting the original and cleaned data
@@ -105,22 +109,23 @@ stan_data_grow_skew <- list(N = nrow(growth_data),                              
                             year = as.integer(as.factor(growth_data$Year_t))           ## predictor years
 )
 # Run the growth model with a student t distribution -- fixed effects: previous size and a non linear previous size variable and ant state; random effects: plot and year; size variation is included for both the omega and alpha estimates
-# grow_skew_model <- stan_model("Data Analysis/STAN Models/grow_skew.stan")
-# fit_grow_skew<-sampling(grow_skew_model,data = stan_data_grow_skew,chains=3,
-#                       control = list(adapt_delta=0.99,stepsize=0.1),
-#                       iter=5000,cores=3,thin=2,
-#                       pars = c("u","w",          # plot and year random effects
-#                                "beta0","beta1","beta2", #location coefficients
-#                                "d_0","d_size", #scale coefficiences
-#                                "a_0","a_size"), #shape coefficients
-#                       save_warmup=F)
+ grow_skew_model <- stan_model("Data Analysis/STAN Models/grow_skew.stan")
+ fit_grow_skew<-sampling(grow_skew_model,data = stan_data_grow_skew,chains=3,
+                       control = list(adapt_delta=0.99,stepsize=0.1),
+                       iter=5000,cores=3,thin=2,
[…]

### 2024-08-11 03_cholla_ant_IPM_params_functions.R

@@ -30,15 +30,16 @@ years <- unique(cactus$Year_t)
 mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 #mcmc_dir <- "/Users/alicampbell/Desktop/"
 #Tom
-#mcmc_dir <- "C:/Users/tm9/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
+mcmc_dir <- "G:/Shared drives/Miller Lab/Sevilleta/Cholla/Model Outputs/"
+
 #Lab
 # mcmc_dir <- "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 ## These files contain all draws from the posterior distributions of all parameters
 # growth model
 # mcmc_dir <- "/Users/Labuser/Desktop/Model_Reads/"
 fit_grow_skew<-readRDS(paste0(mcmc_dir,"fit_grow_skew.rds"))
-grow.params <- rstan::extract(fit_grow_skew, permuted = F)
-grow.params <- grow.params[,c(2),]
+grow.params <- rstan::extract(fit_grow_skew)
+# grow.params <- grow.params[,c(2),]
 # fit_grow_skew_null<-readRDS(paste0(mcmc_dir,"fit_grow_skew_null.rds"))
 # grow.params.null <- rstan::extract(fit_grow_skew_null)
 # fit_grow_stud<-readRDS(paste0(mcmc_dir,"fit_grow_student_t.rds"))
@@ -88,103 +89,111 @@ params <- data.frame(matrix(NA,nrow=N_draws,ncol=1))
 params<-params[,-1]
 ##----------------------Growth Parameters Student T----------------## 
 # # No specific ant
-# params$grow_sig0 <- grow.params$d_0[draws]
-# params$grow_sig1 <- grow.params$d_size[draws]
-# params$grow_alp0 <- grow.params$a_0[draws]
-# params$grow_alp1 <- grow.params$a_size[draws]
-# # Ant 4 (vacant)
-# params$grow_beta04<-grow.params$beta0[draws,4]     	  ## growth intercept
-# params$grow_beta14<-grow.params$beta1[draws,4]				## growth slope
-# params$grow_beta24<-grow.params$beta2[draws,4]				## growth slope
-# # Ant 3 (other)
-# params$grow_beta03<-grow.params$beta0[draws,3]     	  ## growth intercept
[…]

### 2024-08-11 05_cholla_ant_IPM_figures_tables.R

@@ -197,8 +197,82 @@ crem <- outer (
 # lines(size_dummy,size_dummy, lty = 2, col = "grey")
 # dev.off()
 
+##plot showing future size distribution conditional on initial size
+##small size is the 5th percentile of tended plants
+initsmall<-quantile(cactus$logsize_t[cactus$ant_t!="vacant"],probs=0.01,na.rm=T)
+##large size is the 95th percentile of tended plants
+initlarge<-quantile(cactus$logsize_t[cactus$ant_t!="vacant"],probs=0.99,na.rm=T)
+
 ## Plot the countour lines of the studetn t growth model with the mean fit of the model and the real data
-png("Manuscript/Figures/grow_contour.png")
+png("Manuscript/Figures/grow_contour_v1.png")
+par(mar=c(3,5,3,1),oma=c(2,2,0,0))
+layout(matrix(c(1,2,3,4,5,5),
+              ncol = 3, byrow = TRUE), heights = c(1.4,1.4), widths = c(3.9,3.9,3.9))
+# Crem
+contour(x,y,crem, nlevels = 20,  xlim = c(-5,15), ylim = c(-2,15),
+        main = "a)       Crem.               ", cex.main = 2,lwd=1.5,col="black")
+points(y_crem_subset_grow$logsize_t, y_crem_subset_grow$logsize_t1,col=alpha(cremcol,0.5),pch=16,cex=0.75)
+lines(size_dummy, crem_mean, col = cremcol, lwd = 2)
+# Liom
+contour(x,y,liom, nlevels = 20, col = "black", xlim = c(-5,15), ylim = c(-2,15),
+        main = "b)      Liom.                ", cex.main = 2, lwd = 1.5)
+points(y_liom_subset_grow$logsize_t, y_liom_subset_grow$logsize_t1,col=alpha(liomcol,0.5),pch=16,cex=0.75)
+lines(size_dummy, liom_mean, col = liomcol, lwd = 2)
+# Other
+contour(x,y,other, nlevels = 30, col = "black", xlim = c(-5,15), ylim = c(-2,15),
+        main = "c)       Other                ", cex.main = 2, lwd = 1.5)
+points(y_other_subset_grow$logsize_t, y_other_subset_grow$logsize_t1,col=alpha(othercol,0.5),pch=16,cex=0.75)
+lines(size_dummy,other_mean, type = "l", col = othercol,lwd = 2)
+# Vacant
[…]

### 2024-08-19 04_cholla_ant_IPM_analysis.R
Update 04_cholla_ant_IPM_analysis.R

@@ -25,41 +25,187 @@ cols <- c(vcol, ccol, lcol, ocol, lccol, locol, cocol, acol)
 ################################################################################
 ####### Set up parameters for all models
 # Pull out the mean matrix
-params_mean <- matrix(rep(0,ncol(params)), nrow = 1)
-for(i in 1:ncol(params)){
-  params_mean[,i] <- mean(params[,i])
-}
-colMeans(params)
-colnames(params_mean) <- colnames(params)
-params_mean <- as.data.frame(params_mean)
+params_mean<-data.frame(matrix(colMeans(params),1,ncol(params)))
+names(params_mean)<-names(params)
 scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
 scenario_abv <- c("V","C","L","O","LC","LO","OC","LOC")
-m = 1
 colors_lambdas <- c(vcol,ccol,lcol,ocol,lccol,locol,cocol,acol)
 x_vals <- c(1,3,5,7,9,11,13,15)
+
+
+# tom ---------------------------------------------------------------------
+## as you can see below, these extension values work well
+lower.extension<-4
+upper.extension<-0.5
+vacant_mat<-bigmatrix(params = params_mean,
+                      scenario = "none",
+                      lower = cholla_min,
+                      upper = cholla_max,
+                      floor = lower.extension,
+                      ceiling = upper.extension,
+                      matsize = matsize)
+full_mat<-bigmatrix(params = params_mean,
+                    scenario = "all",
+                    lower = cholla_min,
+                    upper = cholla_max,
+                    floor = lower.extension,
+                    ceiling = upper.extension,
+                    matsize = matsize)
+par(mfrow=c(1,2))
+plot(vacant_mat$y,vacant_mat$p,col=vaccol,ylim=c(0.75,1.05))
+plot(full_mat$y,full_mat$p_v,col=vaccol,ylim=c(0.75,1.05))
+points(full_mat$y,full_mat$p_l,col=liomcol)
+points(full_mat$y,full_mat$p_o,col=othercol)
+points(full_mat$y,full_mat$p_c,col=cremcol)
+
+## find the right matrix dimension
+dims<-c(10,25,50,100,150,200,250,300)
+vacant_lam<-full_lam<-c()
+for(i in 1:length(dims)){
+  vacant_lam[i]<-lambda(bigmatrix(params = params_mean,
[…]

### 2024-08-19 Manuscript/ant_cactus_AmNat.tex
Update ant_cactus_AmNat.tex

@@
 Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
-Estimates for the number of seeds per fruit were ant specific, but the data (cite tom's data) only included \textit{L. apiculatum} and \textit{C. opuntiae} ants.
+\tom{Estimates for the number of seeds per fruit were ant specific}{I think this should be expanded.
+The data come from an experiment where we excluded ants, so this is a real strength (relative to the rest of the observational data) that we should not undersell.}, but the data (cite tom's data) only included \textit{L.
+apiculatum} and \textit{C.
+opuntiae} ants.
 \tom{We used the vacancy estimates for plants tended by other ants. }{This is an assumption that needs to be explained or justified. Reviewers will ask if results are sensitive to this assumption.}

### 2024-09-19 Manuscript/ant_cactus_AmNat.tex
Updates

--- 2024-08-11 Manuscript/ant_cactus_AmNat.tex
@@
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
-\tom{They are widespread species interactions \citep{Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010} but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.}{I suggest condensing the two sets of citations into one at the end of the sentence.
-The statement that they are widespread species interactions is so broad and vague that it is not really worth citing.}
+They are widespread species interactions but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022,Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010}.
 Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
@@
 These arborescent plants produce cylindrical segments with large spines. 
-In the growing season (May to August in New Mexico), the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
-While most plants produce new segments every season, only those which are reproductively mature produce flower buds. 
+In the growing season (May to August in New Mexico), the plants initiate new vegetative segments and flowerbuds at the ends of existing segments. 
+While most plants produce new segments every season, only those which are reproductively mature produce flowerbuds. 
 %Tree cholla generally reach at least 9 years of age before beginning to reproduce \citep{Ohm2014}.
-Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative segments and flower buds \citep{Ness2006,Oliveira1999}. 
-Flower buds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
+Like other EFN-bearing cacti, tree cholla secrete nectar from specialized glands on young vegetative segments and flowerbuds \citep{Ness2006,Oliveira1999}. 
+Flowerbuds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
 Smaller, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation at the time of our census) \citep{Miller2014}.
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1141 unique individuals across \tom{19}{There are actually 15 complete transition years.} complete transition years (spanning May/June of year $t$ to May/June of year $t+1$).
+In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$). 
 \tom{}{You should cite the data package here.
@@
 	$$
-	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a') P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
+	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
 	$$
@@
 Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is \tom{dependent on ant state $a$}{The text says ant state $a$ but the notation here and in the equation shows ant state $a'$.
-Which is it?
-(Should be $a$.)}.
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is dependent on ant state $a$. 
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
@@
 We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects.  
-The flower bud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$ in year $t$. 
+The flowerbud function $F(x;\pmb{\tau}^{F})$ estimates the total flowers produced by a reproducing plant in year $t$, with fixed effects of size $x$ in year $t$. 
 We fit this model to flowerbud count data (sum of viable and aborted buds) using a zero-truncated negative binomial distribution with a log link and normally distributed year and plot random effects.
@@
 The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the probability that flowerbuds produced in year $t$ remain viable (not aborted), with fixed effects of ant partner $a$ in year $t$.
-We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of \tom{flower buds}{Sometimes this is flowerbuds and sometimes it is flower buds.
-I would tighten this up.} and the number that are viable, respectively.
+We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
 This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models, with ant-specific year random effects.
 
+Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants (cite Tom's data).
+This data only included  \textit{L. apiculatum} and \textit{C. opuntiae} ants, so we had to make an assumption about the seeds per fruit on plants tended by Other ants.
+We chose to use the vacancy estimates for plants tended by other ants, a decision which does not have a significant impact on the final results. 
 Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
-Estimates for the number of seeds per fruit were ant specific, but the data (cite tom's data) only included \textit{L. apiculatum} and \textit{C. opuntiae} ants.
-\tom{We used the vacancy estimates for plants tended by other ants. }{This is an assumption that needs to be explained or justified. Reviewers will ask if results are sensitive to this assumption.}
-\tom{The rest of these parameters were not ant specific. }{Because the seed number model is ant-specific, I suggest moving it from the appendix to the main text.}
 
@@
 The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$ given that it was occupied by initial ant partner $a$  in year $t$, with fixed effect of initial size $x$.
-We fit this model to ant partner data using a multinomial distribution with a logit link function and \tom{random effects for plot and year}{As far as I can tell there are no plot or year random effects in the model you fit.}.
+We fit this model to ant partner data using a multinomial distribution with a logit link function.
 
@@
 \tom{We visually assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).}{These figures need to be labeled so that it is clear that they are in the Appendix.
-Am Nat's template should allow you to do this automatically.}
+Am Nat's template should allow you to do this automatically.}\ali{}{They are referencing the figures, do you mean another way?
+I am not quite sure what this means.}
 %All estimated parameters are available in Table \ref{tab:Params}.
@@
 In our model there is additional complexity in the form of transitions between $A$ ant states and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
-\tom{We used $b = 200$ bins}{In your script matsize is 500.}, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
+We used $b = 500$ bins, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
 
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the \tom{17 annual transition matrices}{Unless you add 2024 I think this is 16.}, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment where ants responded to annual variation uniquely ($\lambda_{NS}$) we simulated population dynamics for 500 years by randomly sampling among the 16 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
 We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$ \citep{Mark2009}.
-We propagated uncertainty from the vital rate models using \tom{1000 draws}{I did not realize you were doing this many.
-I think you could use as few as 100, which would make your numerical work much faster.} from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
+We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{NS}$ and other derived quantities.
 
 \subsubsection*{Partner diversity experiments}
-Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness in a stochastic environment ($\lambda_{S}$).
+Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness in a stochastic environment ($\lambda_{NS}$).
 From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created treatments corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
@@
 We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two scenarios of environmental variation.
+The 'non-synchronous' (NS) version included ants effects that varied uniquely across time and the 'synchronous' (S) version included  ant effects that were forced to be the same across species. 
 First, we evaluated the model using empirical estimates for the $\pmb{\tau}_{a}$ vectors that describe ant-specific year deviates. 
@@
 The range of models we created generated many outputs; we focus our inference on the following specific contrasts. 
-First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{NS}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
 If cactus fitness increases with partner richness, this would be interpreted as evidence for benefits of partner diversity. 
@@
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with an opportunity cost of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{NS}$ of the full (four-state) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
 If the portfolio effect confers a benefit of diversity, the fitness advantage of having all vs. no partners should be greater when temporal fluctuations are not synchronized across ant states.
@@
 We base our statistical inferences on the posterior probability distributions of the contrasts described above. 
-For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{S}$ ($\Delta\lambda_{C-V}$). 
+For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{NS}$ ($\Delta\lambda_{C-V}$). 
 We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
@@
 We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
-We observed mean viability rates of flowers between 39\% and 96\% (Figure \ref{fig:viab}).
+We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:viab}).
 Ant partners influence the mean viability rate of flowers, with \textit{L.
-apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by \textit{C.
-opuntiae} and Other tended plants (at 74-75\%, Figure \ref{fig:Viab}b,c), and vacant plants had the lowest floral viability rate (71\%, Figure \ref{fig:Viab}d).
+apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by vacant and Other tended plants (at 60-61\%, Figure \ref{fig:Viab}b,c), and \textit{C.
+opuntiae} tended plants had the lowest floral viability rate (57\%, Figure \ref{fig:Viab}d).
 Plants tended by \textit{L. apiculatum} had floral viability advantages, while plants in states of \textit{C. opuntiae}, Other ants, and vacancy had very similar floral viability rates.
@@
 We analyzed all other vital rates mentioned in the methods as well.
-The results of each of these as well as the posterior predictive checks are all included in the Apendix A.
+The results of each of these as well as the posterior predictive checks are all included in the Apendix A (Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}).
 
@@
 First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
-The lowest mean fitness was $\lambda_{S,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-Across all 1+ partner scenarios, we are *****(double check after final runs)82--100\% confident that any scenario of ant visitation elevates fitness. 
-Furthermore, we find an apparent positive effect of partner diversity, with cactus fitness postive related to partner richness, averaging over partner identities (Figure \ref{fig:LambdaMeans}b).
+The lowest mean fitness was $\lambda_{NS,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
+Across all 1+ partner scenarios, we are 100\% confident that any scenario of ant visitation elevates fitness. 
+Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans}).
 
 \begin{figure}
-	\includegraphics[width=0.61\linewidth]{Figures/comp_outputmeans_s_sn.png}
-	\includegraphics[width=0.39\linewidth]{Figures/comp_num_partners_s.png}
-	\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) and $\lambda_{SN}$ (empty circles with an X) for each simulated combination of ant partners.
+	\includegraphics[width=0.61\linewidth]{Figures/Lambdas_comp.png}
+	\includegraphics[width=0.39\linewidth]{Figures/comp_num_partners.png}
+	\caption{Panel a) shows the mean values of the estimated $\lambda_{NS}$ (filled in circles) and $\lambda_{S}$ (empty circles with an X) for each simulated combination of ant partners.
 The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
@@
 apiculatum}, O = other).
-Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+Panel b) shows the mean values of the estimated $\lambda_{NS}$ for }
 	\label{fig:LambdaMeans}
@@
 
-We originally believed the benefits of partner diversity are heavily driven by the presence of a single best partner rather than overall synergy.
-However, we found that there was little benefit of any specific partner over another.
+We found that there was little benefit of any specific partner over another.
 The identity and number of partners does not appear to significantly affect the fitness of the cacti. 
@@
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are *****(double check after final runs)94\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized and non-synchronized model scnearios.
-When subtracting these two resulting vectors from each other (($\lambda_{S,All} - \lambda_{S,Vacant}$) - ($\lambda_{SN,All} - \lambda_{SN,Vacant}$)), we found that we are only *****(double check after final runs) 52\% confident that partners offer higher benefits when able to respond uniquely to a fluctuating environment.
-There is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.
+We are are 98\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scnearios.
+When subtracting these two resulting vectors from each other (($\lambda_{NS,All} - \lambda_{NS,Vacant}$) - ($\lambda_{S,All} - \lambda_{S,Vacant}$)), we found that there is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.
 
@@
 	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
-	\caption{This figure shows the distribution of $\lambda_{S,All}-\lambda_{S,Vacant}$ in pink and $\lambda_{SN,All}-\lambda_{SN,Vacant}$ in green.
+	\caption{This figure shows the distribution of $\lambda_{NS,All}-\lambda_{NS,Vacant}$ in pink and $\lambda_{S,All}-\lambda_{S,Vacant}$ in green.
 The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
@@
 
+We found that vacant plants produced the most mean seeds (147.2 per fruit), followed by \textit{L.
+apiculatum} tended plants (142.4 per fruit), and finally, \textit{C.
+opuntiae} tended plants (115.0 per fruit) (Figure \ref{app:AppA_Seeds_Per_Fruit}).
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
+	\caption{Shows the distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
+	\label{app:AppA_Seeds_Per_Fruit}
+\end{figure}
+
 \paragraph{Recruit Size Distribution}
@@
 where $\hat{\sigma}$ is estimated with a non-informative prior.
+
+We found that the mean size of recruits is $log(-2.097) m^3$ with an interquartile range from $log(-2.173) m^3$ to $log(-1.712) m^3$ (Figure \ref{app:AppA_Recruit_Dist}).
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/rec_size.png}
+	\caption{Shows the distribution of recruit sizes.
+The mean  size of a recruit is marked by the dark line in the green box, the green box shows the interquartile range of recruit sizes, and the dashed lines show the minimum and maximum of estimated recruit sizes.
+The points beyond these dashed lines are considered outliers.
+}
+	\label{app:AppA_Recruit_Dist}
+\end{figure}
 
@@
 
+We found that the mean germination rates for seeds in the seedbank for one year  is 0\%, with an interquartile range of 0\% and 1\%.
+We found that the mean germination rates for seeds in the seedbank for a second year is 0\%, with an interquartile range of 0\% to 0.4\% (Figure \ref{app:AppA_Germ}).
+Seeds are more likely to germinate in their first year in the seedbank, but most seeds will never germinate.
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/germination.png}
+	\caption{Shows the probability of a plant germinating during its first year in the seedbank (left) or during the second year in the seedbank (right).
+The mean germination rates are shown by the dark line near the bottom of the green box.
+The green boxes show the interquartile ranges of the germination rates for year 1 and 2 plants.
+The dashed line shows the maximum germination rates observed.
+The points above that show the outliers of our data.
+}
+	\label{app:AppA_Germ}
+\end{figure}
+
 \paragraph{Pre-Census Survival}
@@
 where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
+
+We found that plants have a 16.2\% probability of surviving from germination to the next census.
+Our model estimated this very well, expecting a 16.3\% probability (Figure \ref{app:AppA_Pre_Census_Surv}).
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/seed_surv.png}
+	\caption{Shows the distribution of the probability of plants which germinated between censuses surviving to the next census to be counted.
+The green line shows the predicted pre-census survival and the grey dashed line shows the actual observed pre-census survival.
+}
+	\label{app:AppA_Pre_Census_Surv}
+\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=0.61\linewidth]{Figures/freq_outputmeans_s_sn.png}
-\includegraphics[width=0.39\linewidth]{Figures/freq_num_partners_s.png}
+\includegraphics[width=0.61\linewidth]{Figures/Lambdas_freq.png}
+\includegraphics[width=0.39\linewidth]{Figures/freq_num_partners.png}
 \caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
@@
 \begin{figure}
-\includegraphics[width=0.61\linewidth]{Figures/equal_outputmeans_s_sn.png}
-\includegraphics[width=0.39\linewidth]{Figures/equal_num_partners_s.png}
+\includegraphics[width=0.61\linewidth]{Figures/Lambdas_equal.png}
+\includegraphics[width=0.39\linewidth]{Figures/equal_num_partners.png}
 \caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.

### 2024-09-19 04_cholla_ant_IPM_analysis.R

@@ -33,84 +33,83 @@ colors_lambdas <- c(vcol,ccol,lcol,ocol,lccol,locol,cocol,acol)
 x_vals <- c(1,3,5,7,9,11,13,15)
 
 
-# tom ---------------------------------------------------------------------
-## as you can see below, these extension values work well
+#### check the extension values
 lower.extension<-4
 upper.extension<-0.5
-vacant_mat<-bigmatrix(params = params_mean,
-                      scenario = "none",
-                      lower = cholla_min,
-                      upper = cholla_max,
-                      floor = lower.extension,
-                      ceiling = upper.extension,
-                      matsize = matsize)
-full_mat<-bigmatrix(params = params_mean,
-                    scenario = "all",
-                    lower = cholla_min,
-                    upper = cholla_max,
-                    floor = lower.extension,
-                    ceiling = upper.extension,
-                    matsize = matsize)
-par(mfrow=c(1,2))
-plot(vacant_mat$y,vacant_mat$p,col=vaccol,ylim=c(0.75,1.05))
-plot(full_mat$y,full_mat$p_v,col=vaccol,ylim=c(0.75,1.05))
-points(full_mat$y,full_mat$p_l,col=liomcol)
-points(full_mat$y,full_mat$p_o,col=othercol)
-points(full_mat$y,full_mat$p_c,col=cremcol)
+# vacant_mat<-bigmatrix(params = params_mean,
+#                       scenario = "none",
+#                       lower = cholla_min,
+#                       upper = cholla_max,
+#                       floor = lower.extension,
+#                       ceiling = upper.extension,
+#                       matsize = matsize)
+# full_mat<-bigmatrix(params = params_mean,
+#                     scenario = "all",
+#                     lower = cholla_min,
+#                     upper = cholla_max,
+#                     floor = lower.extension,
+#                     ceiling = upper.extension,
+#                     matsize = matsize)
+# par(mfrow=c(1,2))
+# plot(vacant_mat$y,vacant_mat$p,col=vaccol,ylim=c(0.75,1.05))
+# plot(full_mat$y,full_mat$p_v,col=vaccol,ylim=c(0.75,1.05))
+# points(full_mat$y,full_mat$p_l,col=liomcol)
[…]

### 2024-09-19 05_cholla_ant_IPM_figures_tables.R

@@ -547,7 +547,7 @@ percentiles <- function(ant,percent){
   y_surv <- quantile(surv.params$beta0[,a],percent) + size_dummy * quantile(surv.params$beta1[,a],percent)
   return(y_surv)
 }
-mean(invlogit(percentiles("vacant",0.95)))
+mean(invlogit(percentiles("vacant",0.05)))
 ## Bin the size data
 # Crem
 surv_plot_crem <- y_crem_subset_surv %>% 
@@ -793,11 +793,13 @@ dev.off()
 ################################################################################
 ## Seed number visuals
 ################################################################################
+png("Manuscript/Figures/Seeds_Per_Fruit.png")
 plot(density(exp(seed.params$beta0[,3])),lwd=3,col=vaccol,
      xlab="Mean seeds per fruit",main=" ",xlim=c(50,250))
 lines(density(exp(seed.params$beta0[,2])),lwd=3,col=liomcol)
 lines(density(exp(seed.params$beta0[,1])),lwd=3,col=cremcol)
-
+legend("topright",legend = c("Vacant","Crem.","Liom."), fill = c(vaccol, cremcol, liomcol))
+dev.off()
 ################################################################################
 ## Probability of Reproducing Model Visuals
 ################################################################################
@@ -944,7 +946,6 @@ dev.off()
 
 
 
-
 ################################################################################
 ## Germination Model Visuals
 ################################################################################
@@ -961,8 +962,8 @@ dev.off()
 
 
 ## Format the original data
-y_germ1 <- stan_data_germ1$y_germ/stan_data_germ1$trials
-y_germ2 <- stan_data_germ2$y_germ/stan_data_germ2$trials
+y_germ1 <- stan_data_germ$y_germ[stan_data_germ$year==1]/stan_data_germ$trials[stan_data_germ$year==1]
[…]

### 2024-09-19 05_cholla_ant_IPM_figures_tables.R
Updates

@@ -1486,6 +1486,62 @@ scenario_abv <- c("V","CV","LV","OV","LCV","LOV","OCV","LOCV")
 # axis(side=1,at=c(1,3,5,7),labels=c("0","1","2","3"))
 # dev.off()
 
+###### Manuscript Visual
+
+## Plot the means for the competitive exclusion model
+png("Manuscript/Figures/Lambdas_Comp_Full.png")
+par(mar=c(4,4,1,1))
+layout(matrix(c(1,2,3),
+              ncol = 3, nrow = 1), heights = c(1,1,1))
+plot(x  = c(1,10,11,9,20,21,19,30), y = colMeans(lams_comp_stoch), pch = 19, cex = 3,col = cols,
+     xlim = c(0,31), ylim = c(0.962,0.995),xaxt = "n",cex.lab = 2,
+     xlab = " ", ylab = " ",
+     main = "Non-Synchronous")
+axis(side=1,at=c(1,10,20,31),labels=c("0","1","2","3"))
+mtext("Mean Lambda Value", side = 2, cex = 2.2, line = 2)
+
+plot(x  = c(1,10,11,9,20,21,19,30), y = colMeans(lams_comp_stoch_null), pch = 19, cex = 3,col = cols,
+     xlim = c(0,31), ylim = c(0.962,0.995),xaxt = "n",cex.lab = 2,
+     xlab = " ", ylab = " ",
+     main = "Synchronous")
+axis(side=1,at=c(1,10,20,31),labels=c("0","1","2","3"))
+legend("bottomright", legend = scenario_abv, fill = cols)
+mtext("Number of Ant Partners", side = 1, cex = 2.2, line = 3, adj = 1.25)
+
+no_part <- colMeans(lams_comp_stoch)[1]
+one_part <- mean(colMeans(lams_comp_stoch)[2:4])
+two_part <- mean(colMeans(lams_comp_stoch)[5:7])
+three_part <- colMeans(lams_comp_stoch)[8]
+no_part_sync <- colMeans(lams_comp_stoch_null)[1]
+one_part_sync <- mean(colMeans(lams_comp_stoch_null)[2:4])
+two_part_sync <- mean(colMeans(lams_comp_stoch_null)[5:7])
+three_part_sync <- colMeans(lams_comp_stoch_null)[8]
+plot(c(1,2,3,4),c(no_part, one_part, two_part, three_part), pch = 16, cex = 3, 
[…]

### 2024-09-23 Manuscript/ant_cactus_AmNat.tex
Updates
manuscript discussion

@@
 \begin{figure}
-	\includegraphics[width=0.61\linewidth]{Figures/Lambdas_comp.png}
-	\includegraphics[width=0.39\linewidth]{Figures/comp_num_partners.png}
+	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_Full.png}
 	\caption{Panel a) shows the mean values of the estimated $\lambda_{NS}$ (filled in circles) and $\lambda_{S}$ (empty circles with an X) for each simulated combination of ant partners.
@@
 
-\section*{Discussion}\ali{}{I am struggling reframing this part.
-I think I may need to start with a new outline, because a lot of what we had doesn't apply to our new conclusions that mutualism is good but we don't see a benefit of partner diversity?}
+\section*{Discussion}
 % mini abstract paragraph
 Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. 
-Here we show that partner identity can play an important role in determining both vital rates and population fitness. 
+Here we show that while partners can remain ecologically different, their identity can still be effectively equivalent.
 The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
@@
 apiculatum} tended plants experiencing floral viability advantages.
-The results of our stochastic IPM revealed that all diversity scenarios which included partners resulted in the highest possible fitness for tree cholla, suggesting that partner diversity may not be beneficial or negative in this system.
-The results of our stochastic null IPM revealed that there is no evidence of portfolio effect in our system. 
-These results highlight that partner presence can increase the overall benefits of a focal mutualism without synergistic benefits and the importance of a mechanistic understanding to explain the benefits of diversity across systems.
-\ali{}{This may not be great, but I am struggling a bit with this framing}
+The results of our synchronous and non-synchronous stochastic IPM revealed that all diversity scenarios which included partners resulted in the highest possible fitness for tree cholla, suggesting that partner diversity is not beneficial in this system.
+These results highlight that partners can differ meaningfully while still resulting in the same fitness benefits for a focal mutualist.
 
 % Connect to broader patterns in the lit
-Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that neither complementarity nor sampling effect explains the benefits of diversity in our system.
+Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that there are no benefits of diversity in our system. 
 In our first round of results, focused on individual ant effects on vital rates, we found that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
@@
 
-%This may have to do with the sheer number of \textit{L. apiculatum} in the system.
-%None of these cited studies reported a single species with overwhelming frequency within their systems.
-%In most cases of complementarity, there are relatively similar frequencies of each partner, which is not the case in our system. 
-%75\% of observed ants are \textit{L. apiculatum} in our system, while only 17\% are \textit{C. opuntiae}.
-%Our vital rate results suggested we may find evidence of complementarity, with \textit{C.
-opuntiae} appearing to boost growth and survial and \textit{L.
-apiculatum} appearing to boost floral viability.
-%We believe there may simply not be enough \textit{C. opuntiae} in the system to have complementarity, leading instead to sampling effect driven by the overwhelming frequency of one ant.
-
-% If there is one best partner then diversity should impose opportunity costs
-%In systems where there is a single "best" partner and concurrent partner interactions, theory suggests that diversity should lead to more interactions with the best partner, thereby increasing the fitness of the focal mutualist \citep{Batstone2018}.
-%In systems with mutually exclusive partner interactions and a "best`` partner,  theory suggests that diversity should impose an opportunity cost of the form of interacting with "inferior" partners, meaning scenarios where more than just the best partner is available should result in decreased fitness \citep{Noe1994}.
-%While out system is aligned with this second expectation, we did not see any decreased fitness from the single "best" partner scenario to scenarios with the "best" partner and other ants included. 
-%We do not see this reflected in our IPM results, potentially because of the high frequency of \textit{L. apiculatum}, the "best" partner, across the population.
-%In the mutually exclusive interaction scenario, the fitness of the population should essentially be an average of the fitness conferred by each partner weighted by their frequency. 
-%Due to the much higher frequency of \textit{L. apiculatum} ants, we believe the weighted average is almost equivalent to that of the single ant scenario with \textit{L. apiculatum} present.
+% why we don't see the vital rate differences in real life
+Despite the apparent benefits for small \textit{C. opuntiae} tended cacti, these benefits are almost never experienced due to chronic vacancy of small cacti (Figure \ref{fig:Ant_Transition}).
+We estimated that cacti between the approximate sizes of -$5 - 5 log(cm^3)$($0.01 - 148 cm^3$) would have important advantages in growth and survival if tended by \textit{C. opuntiae} ants.
+At these sizes plants are very rarely tended ($< 10\%$ occupancy) due to lack of EFN production.
+This means, while our vital rate models predict benefits for small ants, the assumptions we put in place for ant tending in our IPM means that these benefits are extremely rare and therefore have next to no impact on the fitness of the cacti.
+This is not to say that growth and survival are not important vital rates, in fact they are likely to be the two most important in the model. 
+In the literature, growth and survival for established plants are the vital rates that fitness estimates are most sensitive to \cite{Eacker2017}.
+This indicates that the boosts recieved from ant tending of small plants would not be likely to make a large impact on fitness estimations even if this were allowed to occur in this model.
 
 %Turnover
-Partner turnover is likely a significant driver behind the fitness we see within cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
+\ali{}{I could always reframe this to be We see little difference numerically based on ant partner identity, but ants have interesting turnover dynamics or something.
+Aka do I need to reshape this paragraph more?}
+The processes which drive the actual turnover frequency and directions remain a mystery in this system.
 In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
@@
 
+% rephrase this to be more relevant
 What now remains is to determine the underlying processes that lead to the emergence of these characterized patterns. 
@@
 This could potentially have a significant impact on the partners who tend a plant in a given season, though determining the impacts of nectar on directional turnover requires further study. 
-We conducted alternative simulations where the driving mechanism for ant occupation was not competitive exclusion, these results are included in Apendix XX.
+We conducted alternative simulations where the driving mechanism for ant occupation was not competitive exclusion, these results are included in Apendix C (Figures \ref{app:FreqLambdaMeans} - \ref{app:EqualLambdaMeans}).
 
@@
 
-First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that the frequency of \textit{L.
-apiculatum} may be primarily responsible for the fitness of the cholla rather than any other ant partner observed.
+First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that despite the overwhelming  frequency of \textit{L.
+apiculatum} they do not impact the fitness of the cholla more than any other ant observed.
 Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
@@
 Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
-\label{fig:FreqLambdaMeans}
+\label{app:FreqLambdaMeans}
 \end{figure}
@@
 Despite very different proportions, we found very similar outcomes to the competitive exclusion model analyzed in the paper. 
-All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:BetweenLambdaMeans}a).
-Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:BetweenLambdaMeans}b).
+All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:EqualLambdaMeans}a).
+Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:EqualLambdaMeans}b).
 
@@
 Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
-\label{fig:BetweenLambdaMeans}
+\label{app:EqualLambdaMeans}
 \end{figure}

### 2024-09-26 Manuscript/ant_cactus_AmNat.tex
Sept 26 edits

@@
 %\usepackage[authoryear,sectionbib,sort]{natbib}
-\usepackage[square,sort,comma,numbers]{natbib}
+\usepackage[authoryear]{natbib}
 \linespread{1.7}
@@
 \paragraph{Ant Transitions}
-The ant transition model ($\epsilon(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$ given that it was occupied by initial ant partner $a$  in year $t$, with fixed effect of initial size $x$.
+The ant transition model ($\rho(x,a,a';\pmb{\tau}^{\epsilon})$) estimates the probability of a cactus being occupied by ant partner $a'$ in year $t+1$ given that it was occupied by initial ant partner $a$  in year $t$, with fixed effect of initial size $x$.
 We fit this model to ant partner data using a multinomial distribution with a logit link function.
@@
 \subsection*{IPM Analysis}
-Analyzing an IPM requires that we discretize the continuous IPM kernel into an approximating matrix. 
+Analyzing the IPM required that we discretize the continuous IPM kernel into an approximating matrix. 
 Size variable $x$ was discretized into $b$ bins, resulting in a $b \times b$ matrix.
@@
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimate population mean fitness in a stochastic environment where ants responded to annual variation uniquely ($\lambda_{NS}$) we simulated population dynamics for 500 years by randomly sampling among the 16 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment where ants responded to annual variation uniquely \tom{($\lambda_{NS}$)}{Let's discuss this symbology.} we simulated population dynamics for 500 years by randomly sampling among the 16 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
@@
 
-\subsubsection*{Partner diversity experiments}
+\subsubsection*{Partner diversity simulation experiments}
 Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness in a stochastic environment ($\lambda_{NS}$).
@@
 These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed.
-For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants (which were mostly small and non-reproductive) across the size distribution.
+For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants across the size distribution.
 
@@
 \tom{We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.}{Worth doing!} 
-Second, we re-fit the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
+\tom{Second, we re-fit the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
+}{I think this is not what you are doing -- you are averaging them.}
 We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
@@
 \section*{Results}
+\tom{}{All throughout the results section, your assertions are not sufficiently supported with evidence. This will be an important thing to improve, and it should not be hard to do.}
 \subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
+\tom{}{For growth, I suggest using one of my figure options, v1 or v2.
+Also, why do all of the vital rate and turnover figures have different x axis ranges?
+They should all show the entire size distribution.}
 Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
-Plants tended by \textit{C.
+\tom{Plants tended by \textit{C.
 opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
-apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).
+apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).}{You cannot actually see this in the figure.}
 For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean. 
-Similarly, ant visitation enhances cactus survival, and ant partner identity has a significant impact on survival for smaller plants (Figure \ref{fig:Surv}).
+Similarly, ant visitation enhanced cactus survival, and ant partner identity had \tom{a significant impact}{What is the basis for this being ``significant''?} on survival for smaller plants (Figure \ref{fig:Surv}).
 Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
@@
 Ant partners influence the mean viability rate of flowers, with \textit{L.
-apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by vacant and Other tended plants (at 60-61\%, Figure \ref{fig:Viab}b,c), and \textit{C.
-opuntiae} tended plants had the lowest floral viability rate (57\%, Figure \ref{fig:Viab}d).
-Plants tended by \textit{L. apiculatum} had floral viability advantages, while plants in states of \textit{C. opuntiae}, Other ants, and vacancy had very similar floral viability rates.
+apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by vacant and Other tended plants (at 60-61\%, Figure \ref{fig:Viab}b,c), \tom{and \textit{C.
+opuntiae} tended plants had the lowest floral viability rate (57\%, Figure \ref{fig:Viab}d).}{This contradicts the first sentence of the paragraph.
+This figure legend needs to be updated.}
+\tom{Plants tended by \textit{L.
+apiculatum} had floral viability advantages, while plants in states of \textit{C.
+opuntiae}, Other ants, and vacancy had very similar floral viability rates.
+}{This sentence is restating the previous one.}
 
@@
 
-We analyzed the correlation coefficients of all models which included ant state as a predictor and found that the annually varying effects of each ant on survival were the least correlated, and that the effects on growth were the most correlated.
-Low correlation indicates synchronicity, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
-Across growth, survival, and viability models, the annally varying random effects of Other ants were the most asynchronous (the least correlated with the effects of other partners), while vacant random effects were the most synchronous.
-This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed low synchronicity, particularly in the survival model.
+We analyzed the \tom{correlation coefficients}{You are not actually providing any correlation coefficients.} of all models which included ant state as a predictor and found that the annually varying effects of each ant on survival were the least correlated, and that the effects on growth were the most correlated.
+\tom{Low correlation indicates synchronicity}{This is not correct.}, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
+Across growth, survival, and viability models, the annually varying random effects of Other ants were the most asynchronous (the least correlated with the effects of other partners), while vacant random effects were the most synchronous.
+This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed \tom{low synchronicity}{You have not actualy quantified what you mean by low synchronicity.}, particularly in the survival model.
 
@@
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-We found that 40\% of plants experienced an ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}). 
+\tom{We found that 40\% of plants experienced an ant state transition on average}{I am not sure what exactly this is quantifying -- 40\% of individuals in the long-term data had at least one ant transition, is that right?
+I would have expected higher.}, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
 Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
@@
 opuntiae} at the end of the transition year is only slightly greater than the probability of take-over by \textit{L.
-apiculatum}, while take-over in the other directino is extremely rare.
+apiculatum}, while take-over in the other direction is extremely rare.
 It is also notable that transitions away from the initial state of  \textit{L.
@@
 The lowest mean fitness was $\lambda_{NS,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-Across all 1+ partner scenarios, we are 100\% confident that any scenario of ant visitation elevates fitness. 
-Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans}).
+Across all 1+ partner scenarios, \tom{we are 100\% confident}{I know what you are saying here but it would be better to rephrase this in the language of the contrasts described in the methods.} that any scenario of ant visitation elevates fitness.
+Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \tom{\ref{fig:LambdaMeans}).}{All multi-panel figures should be labeled with letters, and they should be referenced specifically.
+Here you are actually referencing the third panel, not the whole figure.
+I suggest re-ordering to make this panel first, since you describe it first.
+I also thought we decided to use the uncertainty version of this figure.}
 
@@
 
-We found that there was little benefit of any specific partner over another.
+\tom{We found that there was little benefit of any specific partner over another.
 The identity and number of partners does not appear to significantly affect the fitness of the cacti. 
-This indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.
-
-After further simulations, we found that this lack of diversity benefits are not driven by frequency of a single ant. 
+This indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.}{You are not referencing any figures or providing any results in the text, so there is nothing backing up your assertions here.
+The methods does a nice job of describing the contrasts that are the focus of the analysis but you do not ever provide those results here.
+You never address the apparent reduction in fitness at the full partner scenario.}
+
+The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
 Using the simulations where all ants had equal frequencies across sizes (further explained and analyzed in Appendix C), we found the same fitness patterns as in the competitive exclusion model described above.
@@
 
-We found no evidence of portfolio effect, meaning the presence of multiple partners did not buffer against the potentially negative effects of annual fluctuations.
+We found no evidence of portfolio effect, meaning the presence of \tom{multiple partners did not buffer against the potentially negative effects of annual fluctuations}{The portfoilio effect is not about buffering negative effects of fluctuations.}.
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are 98\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scnearios.
+We are are 98\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scenarios.
 When subtracting these two resulting vectors from each other (($\lambda_{NS,All} - \lambda_{NS,Vacant}$) - ($\lambda_{S,All} - \lambda_{S,Vacant}$)), we found that there is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.
@@
 Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. 
-Here we show that while partners can remain ecologically different, their identity can still be effectively equivalent.
+Here we show that while alternative partners may be ecologically different, their fitness effects on a shared mutualist can be effectively equivalent and interchangeable.
 The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
-opuniae} tended plants experiencing advantages in growth and survival when small, and \textit{L.
-apiculatum} tended plants experiencing floral viability advantages.
-The results of our synchronous and non-synchronous stochastic IPM revealed that all diversity scenarios which included partners resulted in the highest possible fitness for tree cholla, suggesting that partner diversity is not beneficial in this system.
+opuniae} tended plants \tom{experiencing advantages in growth}{I don't think this is true if you look at my v1 figure.} and survival when small, and \textit{L.
+apiculatum}-tended plants experiencing floral viability advantages.
+The results of our synchronous and non-synchronous stochastic IPM revealed that all diversity scenarios which included any partners resulted in the highest possible fitness for tree cholla, suggesting that while ant visitation is beneficial, partner identity and diversity are inconsequential in this system.
 These results highlight that partners can differ meaningfully while still resulting in the same fitness benefits for a focal mutualist.
@@
 % Connect to broader patterns in the lit
-Similar studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that there are no benefits of diversity in our system. 
-In our first round of results, focused on individual ant effects on vital rates, we found that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
-This indicated to us that we may be dealing with complementarity, and that each ant partner offered a unique form of herbivore protection across ontogeny and or across different processes. 
-However, our IPM results showed that we were actually observing a lack of beneficial diversity, leaving us with questions about how these results could fit together.
+\tom{Previous studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that there are no benefits of diversity in our system.
+}{This senetence cites mutualism and non-mutualism studies.
+It is important that your discussion be focused on effects of partner diversity in mutualisms -- that is the frame of reference against which you are interpreting your results.}
+Vital rate results indicated that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
+On their own, these results suggested a likely benefit of diversity through complementarity, such that each ant partner offered a unique form of herbivore protection across ontogeny and or across different life history processes.
+However, our multi-state IPM results, which synthesize all life history processes and accouns for partner identity and turnover, revealed an overall positive effect of ant visitation but no apparent benefit of diversity, raising questions about how to resolve the tension between these two sets of results.
+
+We speculate that our results are driven by the vital rate sensitivity structure of this population. 
+Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
+Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in survival at small sizes, where mortality risk was high (Figure \ref{}) and the probability of ant tending was low (Figure \ref{}).
+At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of ant identity, and that result dominates integrated measures of fitness because of the high sensitvity of established plant survival.
+The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness.
 
 % why we don't see the vital rate differences in real life
-Despite the apparent benefits for small \textit{C. opuntiae} tended cacti, these benefits are almost never experienced due to chronic vacancy of small cacti (Figure \ref{fig:Ant_Transition}).
-We estimated that cacti between the approximate sizes of -$5 - 5 log(cm^3)$($0.01 - 148 cm^3$) would have important advantages in growth and survival if tended by \textit{C. opuntiae} ants.
-At these sizes plants are very rarely tended ($< 10\%$ occupancy) due to lack of EFN production.
-This means, while our vital rate models predict benefits for small ants, the assumptions we put in place for ant tending in our IPM means that these benefits are extremely rare and therefore have next to no impact on the fitness of the cacti.
-This is not to say that growth and survival are not important vital rates, in fact they are likely to be the two most important in the model. 
-In the literature, growth and survival for established plants are the vital rates that fitness estimates are most sensitive to \cite{Eacker2017}.
-This indicates that the boosts recieved from ant tending of small plants would not be likely to make a large impact on fitness estimations even if this were allowed to occur in this model.
+%Despite the apparent benefits for small \textit{C. opuntiae} tended cacti, these benefits are almost never experienced due to chronic vacancy of small cacti (Figure \ref{fig:Ant_Transition}).
+%We estimated that cacti between the approximate sizes of -$5 - 5 log(cm^3)$($0.01 - 148 cm^3$) would have important advantages in growth and survival if tended by \textit{C. opuntiae} ants.
+%At these sizes plants are very rarely tended ($< 10\%$ occupancy) due to lack of EFN production.
+%This means, while our vital rate models predict benefits for small ants, the assumptions we put in place for ant tending in our IPM means that these benefits are extremely rare and therefore have next to no impact on the fitness of the cacti.
+%This is not to say that growth and survival are not important vital rates, in fact they are likely to be the two most important in the model. 
+%In the literature, growth and survival for established plants are the vital rates that fitness estimates are most sensitive to \cite{Eacker2017}.
+%This indicates that the boosts recieved from ant tending of small plants would not be likely to make a large impact on fitness estimations even if this were allowed to occur in this model.
+
+% Portfolio Effect
+Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits. 
+Yet, we find no evidence of portfolio effect within our system. 
+When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
+We did not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure how the effects of each partner on cholla demography fluctuate through time, and whether those fluctuations are correlated.
+\tom{We found that each ant partner had different effects on different vital rates across time, indicating they exhibit different reactions to varying environments, leading us to believe that portfolio effect could be possible in this system.
+We compared the fitness of cholla under simulations where the variation in benefits across time reflected what we found in nature to siulations where each ant was forced to react the same way to temporal variation (excluding the possibility of portfolio effect).
+We found the benefits were not stronger across time when the ants were allowed to vary naturally than when they were forced to react synchronously, indicating that portfolio effect does not explain the benefits of diversity in our system.}{This is restating methods and results.
+Make this more interpretive.}
+\tom{Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found }{??}
+\tom{One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature.}{I like this point, and this is a good example of a discussion paragraph zooming out with a final sentence that makes a larger point.}
 
@@
 Aka do I need to reshape this paragraph more?}
+\tom{}{I would cut this and the next paragraph (``what remains now'') and replace them with a summary of how the results are robust to different assumptions about whether or not turnover patterns are driven by competition, and what we know and don't know about the mechanisms of turnover.
+You have some of that here but it can come through clearer.}
 The processes which drive the actual turnover frequency and directions remain a mystery in this system.
@@
 
-% Portfolio Effect
-Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits.
-Yet, we find no evidence of portfolio effect within our system.
-When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
-In our system, we do not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure the effects of each partner on the cholla across time, allowing us to statistically quantify how the benefits recieved from each partner vary across time.
-We found that each ant partner had different effects on different vital rates across time,  indicating they exhibit different reactions to varying environments, leading us to believe that portfolio effect could be possible in this system.
-We compared the fitness of cholla under simulations where the variation in benefits across time reflected what we found in nature to siulations where each ant was forced to react the same way to temporal variation (excluding the possibility of portfolio effect).
-We found the benefits were not stronger across time when the ants were allowed to vary naturally than when they were forced to react synchronously, indicating that portfolio effect does not explain the benefits of diversity in our system.
-Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found 
-One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
-This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in real systems.
-
 % Value of Long-term data
@@
 Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-The long term dataset we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
-
-First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that despite the overwhelming  frequency of \textit{L.
+\tom{The long term data set we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is a great message to emphasize.}
+By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of specific demographic effects. 
+For example, our previous study suggested that \textit{C.
+opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set \citep{Ohm2014}.
+Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages, indicates that this species is clearly a mutualist and nearly as strong a mutualist as \textit{L.
+apiculatum}.
+
+%First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that despite the overwhelming  frequency of \textit{L.
 apiculatum} they do not impact the fitness of the cholla more than any other ant observed.
-Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
-Second, our long-term data set allowed us to test the portfolio effect.
-This data is required, because we must show that the partners react asynchronously to temporal environmental stochasticity for portfolio effect to even be possible.
-This is a great example of how long-term data may offer more power to test underlying mechanisms in similar systems.
-Finallly, previous studies by \cite{Miller2007} hypothesized that \textit{C. opuntiae} may be a parasite based on the effects on the cholla cacti.
-Using more years of consecutive data in this system, we were able to show that \textit{C.
+%Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
+%Second, our long-term data set allowed us to test the portfolio effect.
+%This data is required, because we must show that the partners react asynchronously to temporal environmental stochasticity for portfolio effect to even be possible.
+%This is a great example of how long-term data may offer more power to test underlying mechanisms in similar systems.
+%Finallly, previous studies by \cite{Miller2007} hypothesized that \textit{C. opuntiae} may be a parasite based on the effects on the cholla cacti.
+%Using more years of consecutive data in this system, we were able to show that \textit{C.
 opuntiae}, while the worst individual partner in terms of the overall cholla fitness (Figure \ref{fig:LambdaMeans}), increases the growth and survival rates of cacti prior to reproduction.
-The length of our data set has allowed us to dive into individual vital rates as well as overall fitness with more clarity.
+%The length of our data set has allowed us to dive into individual vital rates as well as overall fitness with more clarity.
 
@@
 Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti.
-Despite this, we do not directly account for herbivory in our system. 
-Rather, we assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
+We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
 In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
 This is particularly true in light of our consideration of portfolio effect. 
-In the same way that different species of ants likely react to temporal environmental variation differently, the same can be said for herbivores \citep{Thibaut2012}.
+\tom{In the same way that different species of ants likely react to temporal environmental variation differently, the same can be said for herbivores \citep{Thibaut2012}.
 Herbivores have variations in their population reflective of the environment changing temporally, meaning there are "good" and "bad" years for herbivores just like for ant defenders \citep{Batstone2018}, so the strength of their negative effect varies.
@@
 We found no evidence of this mechanism (Figure \ref{fig:Portfolio}), but a more herbivore explicit model in the future may reveal more about driving mechanisms. 
+}{This confusing to me.
+Herbivores ARE the portfolio effect.
+We are not modeling ants becomes more or less abundant, we are modeling their demographic effects fluctuating, which could only be due to herbivores fluctuating.}
 
@@
 However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find the presence of ants can lead to direct increases in survival \citep{Beattie1985}, reproductive efforts \citep{Ford2015}, growth \citep{Frederickson2005}, etc.
-In the future, it would be interesting to verify the causation of our results through experimental manipulations.
+\tom{In the future, it would be interesting to verify the causation of our results through experimental manipulations.}{Ali, we have done these!
+I think this paragraph needs to be re-writted to acknowledge that, because it gives us a lot more confidence in interpreting the observational trends.}
 One further expansion could include nectar analysis of these cacti in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
 
-Another limitation of our study is that we are simulating populations through an IPM to estimate fitness, meaning we have to extrapolate beyond observed interaction.
+\tom{Another limitation of our study is that we are simulating populations through an IPM to estimate fitness, meaning we have to extrapolate beyond observed interaction.
 In our study, small plants are rarely seen with partners, yet we report on the expected effects of each ant on the growth and survival of plants from the size of new recruits to the largest plants. 
@@
 opuntiae}, yet we offer an estimation based on the effects of this partner in a diverse system.
-The many years of data and 10467 unique plant-year observations included in our analysis allow us to have relatively high confidence in our outputs despite extrapolations.
+The many years of data and 10467 unique plant-year observations included in our analysis allow us to have relatively high confidence in our outputs despite extrapolations.}{I am not convinced this paragraph is necessary.
+The topic of creating ``counterfactual'' scenarios could be incorporated into the paragraph I am suggesting above, intepreting the different scenarios of turnover.}
+
+\tom{}{I strongly suggest a conclusion paragraph that brings the reader back to the big picture, what we learn from this study in the broadest sense, and why it matters.
+Is this how mutualisms work generally?
+How commonly and under what conditions should we expect diversity to not matter, as it appears not to matter here?}
 
 \section*{Acknowledgments}
-This should be drafted.
+\tom{This should be drafted.}{}
 %%%%%%%%%%%%%%%%%%%%%
@@
 \section*{Data and Code Availability}
-This should be drafted.
+\tom{This should be drafted.}{}
 

### 2024-09-28 Manuscript/ant_cactus_AmNat.tex
Manuscript Updates

@@
 \usepackage{amsmath}
-\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
+%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
 \usepackage{fullpage}
-%\usepackage[authoryear,sectionbib,sort]{natbib}
 \usepackage[authoryear]{natbib}
@@
 \newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}  
+\usepackage{hyperref}
 
@@
 We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in the wild.
+We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in nature.
 The hierarchical models revealed that different ant partners had different impacts on the cholla vital rates. 
 Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
-The IPM results revealed that scenario which included \textit{L. apiculatum} only resulted in the highest possible fitness for the tree cholla. 
+The IPM results revealed that despite these differences in vital rates, the presence of any ant partner is beneficial, the identity and number of partners do not matter.
 This suggests that there are no benefits of diversity in this system.
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$). 
-\tom{}{You should cite the data package here.
-The published package only goes up to 2018, so we should talk about updating that package with you as lead author, which should be easy to do as long as the data are QAQC'd.}
-%These data were used to fit vital rate models (survival, growth, reproduction) as functions of plant size and ant occupancy state.
+In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
 
@@
 Vital rate models included spatial and temporal random effects associated with plot and year variation, respectively (only year variation is used in the IPM), and included plant size (the natural logarithm of volume, $log(cm^3)$; $x,x'$), ant partner state ($a,a'$), or both as fixed-effect predictor variables.
-%In addition to vital rate models describing plant demographic performance, we also fit a sub-model to predict transitions between ant states conditional on plant size and previous ant state. 
 As in the IPM, our statistical modeling assumed that demographic effects of ant occupancy are limited to survival, growth, and flowerbud viability.
-
-%Vital rate models for survival, growth, probability of flowering,  take the form of the function:
-%$$ f(\mu) = \beta_0 + \beta_1 x + ... + u + w $$
-%\noindent where $u \sim N(0,\sigma_{yr }^{2})$ is the year random effect with year specific deviates of $\sigma_{yr}$ (which parameterizes the $\tau$ vector), and $w \sim N(0,\sigma_{plot}^{2})$ is the plot random effect with plot specific variance of $\sigma_{plot}$.
-%Some models deviate from this general model, which will be specified below.
 
@@
 For each model, we obtained three chains of 10,000 iterations, discarding the first 1,500 iterations. 
-%We did not thin the chains, thus all samples were retained following burn-in. 
-\tom{We visually assessed parameter convergence between and within chains (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).}{These figures need to be labeled so that it is clear that they are in the Appendix.
-Am Nat's template should allow you to do this automatically.}\ali{}{They are referencing the figures, do you mean another way?
-I am not quite sure what this means.}
-%All estimated parameters are available in Table \ref{tab:Params}.
+We visually assessed parameter convergence between and within chains (Appendix \ref{appendix:D}: Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Appendix \ref{appendix:D}: Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
 
@@
 This means that, in scenarios that remove species from the partner community, remaining species gain access to plants that the removed species would have tended. 
-In Appendix XX, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
-
-%\tom{We used this model to estimate probabilities which are counterfactual (such as the probability of being \textit{L. apiculatum} tended or vacant with no other options). 
-%To do this we assumed that the level of vacancy was constant across all states and that whatever ants were present in our simulation would tend the gaps which would otherwise be tended by another ant species.
-%This assumes that competitive exclusion is the mechanism which determines the ants we see tending cacti, other potential approaches are outlined in Apendix XX.}{I would move this material to the model analysis section.}
+In Appendix \ref{appendix}, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
 
@@
 In this scenario, good years and bad years can differ between ant states, according to the empirical parameter estimates. 
-\tom{We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.}{Worth doing!} 
-\tom{Second, we re-fit the ant-dependent vital rates (survival, growth, flowerbud viability) without ant-specific year random effects, thus assuming that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
-}{I think this is not what you are doing -- you are averaging them.}
+We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.
+Second, we averaged the ant-dependent vital rates (survival, growth, flowerbud viability) across ant-specific year random effects, thus ensuring that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
 We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
@@
 \section*{Results}
-\tom{}{All throughout the results section, your assertions are not sufficiently supported with evidence. This will be an important thing to improve, and it should not be hard to do.}
 \subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
-\tom{}{For growth, I suggest using one of my figure options, v1 or v2.
-Also, why do all of the vital rate and turnover figures have different x axis ranges?
-They should all show the entire size distribution.}
 Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
-\tom{Plants tended by \textit{C.
+Plants tended by \textit{C.
 opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
-apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).}{You cannot actually see this in the figure.}
+apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).
 For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean. 
-Similarly, ant visitation enhanced cactus survival, and ant partner identity had \tom{a significant impact}{What is the basis for this being ``significant''?} on survival for smaller plants (Figure \ref{fig:Surv}).
+Similarly, for plants which were large enough to have ant visitors (larger than $0.8$ $log(cm^3)$), visitation enhanced cactus survival (Figure \ref{fig:Surv}).
+Plants that were smaller than $-2$ $log(cm^3)$ (indicated by the dashed lines for ant estimates, Figure \ref{fig:Surv}) were estimated to experience survival boosts from vacancy.
+Due to the assumptions in our models based on real ant presence data, tended plants of the size where vacancy would offer a boost are never seen and therefore do not influence the fitness estimates of cacti.
+\ali{}{I'm not 100\% certain on how to discuss this.
+I know that this still shows that crem offer a survival advantage, but I see that it could look like vacancy offers survival advantages.
+}
+The smallest plants showed evidence that vacancy resulted in the highest survival rates, however these plants are never tended, meaning the models of ant-tended cacti survival before the solid lines are extrapolation.
 Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
-\textit{C. opuntiae}-occupied plants had a survival advantage over other ant states, particularly at smaller sizes, consistent with the positive effects on growth. 
+\textit{C. opuntiae}-occupied plants had a survival advantage over other ant-tended plants, particularly at smaller sizes, consistent with the positive effects on growth. 
 At larger sizes, plants in any state of ant occupancy had a survival advantage over vacant plants.
@@
 \begin{figure}
-	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour.png}
+	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour_v2.png}
 	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
@@
 
-We found evidence that ant visitation leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
+We found evidence that ant visitation by \textit{L. apiculatum} ants leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
 We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:viab}).
 Ant partners influence the mean viability rate of flowers, with \textit{L.
-apiculatum}-tended plants experiencing the highest mean viability rate (86\%, Figure \ref{fig:Viab}b), followed by vacant and Other tended plants (at 60-61\%, Figure \ref{fig:Viab}b,c), \tom{and \textit{C.
-opuntiae} tended plants had the lowest floral viability rate (57\%, Figure \ref{fig:Viab}d).}{This contradicts the first sentence of the paragraph.
-This figure legend needs to be updated.}
-\tom{Plants tended by \textit{L.
-apiculatum} had floral viability advantages, while plants in states of \textit{C.
-opuntiae}, Other ants, and vacancy had very similar floral viability rates.
-}{This sentence is restating the previous one.}
+apiculatum}-tended plants experiencing the highest mean viability rate (86\%), followed by vacant and Other tended plants (at 60-61\%), and \textit{C.
+opuntiae} tended plants had the lowest floral viability rate (57\%).
 
@@
 We analyzed all other vital rates mentioned in the methods as well.
-The results of each of these as well as the posterior predictive checks are all included in the Apendix A (Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}).
-
-We analyzed the \tom{correlation coefficients}{You are not actually providing any correlation coefficients.} of all models which included ant state as a predictor and found that the annually varying effects of each ant on survival were the least correlated, and that the effects on growth were the most correlated.
-\tom{Low correlation indicates synchronicity}{This is not correct.}, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
-Across growth, survival, and viability models, the annually varying random effects of Other ants were the most asynchronous (the least correlated with the effects of other partners), while vacant random effects were the most synchronous.
-This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed \tom{low synchronicity}{You have not actualy quantified what you mean by low synchronicity.}, particularly in the survival model.
+The results of each of these as well as the posterior predictive checks are all included in the Apendix \ref{appendix:A} : Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
+
+We analyzed the correlation coefficients of all models which included ant state as a predictor and found that the annually varying effects of each ant on growth were the most correlated (correlation coefficients averaged 0.63), and that the effects on survival were the least correlated (correlation coefficients averaged 0.36).
+High correlation indicates synchronicity, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
+For the growth model, annually varying randomn effects of Other ants were the least synchronous (correlation coefficients ranging from 0.28 to 0.39) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.39 to 0.81).
+For the viability model, annually varying random effects of \textit{L.
+apiculatum} ants were the least synchronous (correlation coefficients ranging from 0.06 to 0.44) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.44 to 0.62).
+For the survival model, annually varying random effects of \textit{C.
+opuntiae} ants were the least synchronous (correlation coefficients ranging from 0 to 0.15) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.13 to 2.6).
+This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed low synchronicity (low correlation of annual effects, indicating that ant identity impacts annual random effects), particularly in the survival model.
 
@@
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-\tom{We found that 40\% of plants experienced an ant state transition on average}{I am not sure what exactly this is quantifying -- 40\% of individuals in the long-term data had at least one ant transition, is that right?
-I would have expected higher.}, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
+We found that 55\% of individual plants surveyed in the long-term data experienced at least one ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
 Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
@@
 The lowest mean fitness was $\lambda_{NS,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-Across all 1+ partner scenarios, \tom{we are 100\% confident}{I know what you are saying here but it would be better to rephrase this in the language of the contrasts described in the methods.} that any scenario of ant visitation elevates fitness.
-Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \tom{\ref{fig:LambdaMeans}).}{All multi-panel figures should be labeled with letters, and they should be referenced specifically.
-Here you are actually referencing the third panel, not the whole figure.
-I suggest re-ordering to make this panel first, since you describe it first.
-I also thought we decided to use the uncertainty version of this figure.}
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_Full.png}
-	\caption{Panel a) shows the mean values of the estimated $\lambda_{NS}$ (filled in circles) and $\lambda_{S}$ (empty circles with an X) for each simulated combination of ant partners.
-The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+Across all 1+ partner scenarios, we found that the posterior $\lambda_{NS,Occupied}$ distributions were greater than $\lambda_{NS,Vacant}$ distributions 100\% of the time.
+This indicates that ant visitation elevates fitness no matter the identity or number of partners.
+Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
+From the means alonet appears that there is a reduction in fitness when all ants are present compared to 1 or 2 partners.
+The contrasts of $\lambda_{NS}$ posterior distributions of 0 - 2 partner diversity scenarios to full diversity show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
+This is highlighted in Figure \ref{fig:LambdaMeans} panel b.
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_lines.png}
+	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+The lines show the posterior distribution spread of estimated $\lambda$ values.
+The letters in the legend correspond to what ant partners are present (V = Vacant, C = \textit{C.
 opuntiae}, L = \textit{L.
 apiculatum}, O = other).
-Panel b) shows the mean values of the estimated $\lambda_{NS}$ for }
+}
 	\label{fig:LambdaMeans}
@@
 
-\tom{We found that there was little benefit of any specific partner over another.
-The identity and number of partners does not appear to significantly affect the fitness of the cacti. 
-This indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.}{You are not referencing any figures or providing any results in the text, so there is nothing backing up your assertions here.
-The methods does a nice job of describing the contrasts that are the focus of the analysis but you do not ever provide those results here.
-You never address the apparent reduction in fitness at the full partner scenario.}
+We found that the difference in posterior distributions of any ant scenario ($\lambda_{NS,A} - \lambda_{NS,B}$, where A and B are arbitrary ant diversity scenarios that are not vacancy) were between 31\% and 83\% different.
+The posterior distributions of $\lambda$ for each diversity scenario are visualized in Figure \ref{fig:LambdaMeans} b.
+This high level of overlap between posterior distributions indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.
 
@@
 
-We found no evidence of portfolio effect, meaning the presence of \tom{multiple partners did not buffer against the potentially negative effects of annual fluctuations}{The portfoilio effect is not about buffering negative effects of fluctuations.}.
+We found no evidence of portfolio effect.
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
@@
 The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
-opuniae} tended plants \tom{experiencing advantages in growth}{I don't think this is true if you look at my v1 figure.} and survival when small, and \textit{L.
+opuniae} tended plants \tom{experiencing slight advantages in growth}{I don't think this is true if you look at my v1 figure.} and survival when small, and \textit{L.
 apiculatum}-tended plants experiencing floral viability advantages.
@@
 % Connect to broader patterns in the lit
-\tom{Previous studies have reported complementarity \citep{Palmer2010,Rezende2007,Hooper2005,Cardinale2007,Afkhami2021} while we found that there are no benefits of diversity in our system.
-}{This senetence cites mutualism and non-mutualism studies.
-It is important that your discussion be focused on effects of partner diversity in mutualisms -- that is the frame of reference against which you are interpreting your results.}
+Previous studies have reported complementarity \citep{Palmer2010,Rezende2007,Afkhami2021} while we found that there are no benefits of diversity in our system.
 Vital rate results indicated that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
@@
 Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
-Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in survival at small sizes, where mortality risk was high (Figure \ref{}) and the probability of ant tending was low (Figure \ref{}).
+Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{fig:Viab}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in survival at small sizes, where mortality risk was high (Figure \ref{fig:Surv}) and the probability of ant tending was low (Figure \ref{fig:Ant_Transition}).
 At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of ant identity, and that result dominates integrated measures of fitness because of the high sensitvity of established plant survival.
@@
 We did not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure how the effects of each partner on cholla demography fluctuate through time, and whether those fluctuations are correlated.
-\tom{We found that each ant partner had different effects on different vital rates across time, indicating they exhibit different reactions to varying environments, leading us to believe that portfolio effect could be possible in this system.
-We compared the fitness of cholla under simulations where the variation in benefits across time reflected what we found in nature to siulations where each ant was forced to react the same way to temporal variation (excluding the possibility of portfolio effect).
-We found the benefits were not stronger across time when the ants were allowed to vary naturally than when they were forced to react synchronously, indicating that portfolio effect does not explain the benefits of diversity in our system.}{This is restating methods and results.
-Make this more interpretive.}
-\tom{Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found }{??}
-\tom{One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
-This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature.}{I like this point, and this is a good example of a discussion paragraph zooming out with a final sentence that makes a larger point.}
-
-%Turnover
-\ali{}{I could always reframe this to be We see little difference numerically based on ant partner identity, but ants have interesting turnover dynamics or something.
-Aka do I need to reshape this paragraph more?}
-\tom{}{I would cut this and the next paragraph (``what remains now'') and replace them with a summary of how the results are robust to different assumptions about whether or not turnover patterns are driven by competition, and what we know and don't know about the mechanisms of turnover.
-You have some of that here but it can come through clearer.}
-The processes which drive the actual turnover frequency and directions remain a mystery in this system.
+The vital rate results led us to believe that portfolio effect may be at play due to the differences in ant responses to annual variation.
+However, there is no stabilization of fitness across temporal heterogeneity and therefore no benefits of diversity due to the portfolio effect.
+Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found that asynchrony was linked to weak evidence of portfolio effect.
+One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature.
+
+% These results are robust to different assumptions about turnover patterns & what we don't know about turnover
+Partner turnover is likely a significant driver behind the fitness we see within Cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
 In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
-In some systems high freqency of turnover is necessary for resiliency and leads to higher fitness benefits \cite{Trojelsgaard2015}, while in other systems loyalty is the most beneficial \cite{Batstone2018}.
-We found that some ants had higher frequency of turnover than others, specifically the other category appeared to have high frequency of turnover, rarely returning to the same plant, while \textit{C.
-opuntiae} and \textit{L.
-apiculatum} often returned.
 The direction of partner turnover is also important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
-In our study we found that there are distinct patterns to the direction of partner turnover. 
-Specifically, most plants are tended by \textit{L.
-apiculatum} ants at some point, because vacant plants, other tended plants, and \textit{L.
-apiculatum} tended plants are all most likely to be tended by \textit{L.
-apiculatum} ants.
-This indicates that while \textit{L. apiculatum} ants are loyal to their own plants and return multiple years in a row to the same ones, they are also strong colonizers.
-
-% rephrase this to be more relevant
-What now remains is to determine the underlying processes that lead to the emergence of these characterized patterns. 
-Throughout the literature, it is reported that non-symbiotic ants and non-mutualist ants are highly attracted to monosaccharide heavy EFN \citep{Heil2004,Kautz2009} and that plants often produce monosaccharide EFN to avoid exploitation \citep{Heil2010}.
-It has been previously reported that when reproducing, the nectar of stem segments has a greater disaccharide vs monosaccharide ratio, while flowerbuds have a lower ratio \citep{Miller2008}. 
-This could potentially have a significant impact on the partners who tend a plant in a given season, though determining the impacts of nectar on directional turnover requires further study. 
-We conducted alternative simulations where the driving mechanism for ant occupation was not competitive exclusion, these results are included in Apendix C (Figures \ref{app:FreqLambdaMeans} - \ref{app:EqualLambdaMeans}).
+In our system we found that there are distinct patterns to the directions and frequency of partner turnover, but we still don't understand the mechanisms driving these turnovers.
+In the model presented in this paper we assumed that competitive exclusion was the driving mechanism behind the patterns we see.
+This assumption would indicate that ants are competing for each individual cacti, likely based on EFN quality and composition \citep{Heil2004,Heil2010}, and that the more common an ant is the better a competitor it is.
+In our simulations this means that when one ant is removed the ants which are left occupy the newly vacated cacti so that the proportion of vacant cacti never changes.
+We included this assumption because we believe it to be the most biologically reasonable based on the personal observations of multiple researchers.
+However, it is possible this assumption is incorrect, so we created two alternative simulations to determine how robust our results were to various assumptions about partner turnover drivers.
+The first alternative we tested is a model that assumes all ants have the same competitive ability rather than the differences in competitive ability we assume cause observed variety in ant frequency.
+In this hypothetical simulation all ants have equal probability of occupying any given cactus, but the level of vacancy across the population remains consistent with observed data.
+This simulation was designed to test if the overwhelming frequency of \textit{L. apiculatum} may be overriding any potential advantages of other partners.
+The results of this simulation (Appendix \ref{appendix:C}: Figure \ref{app:EqualLambdaMeans}) were effectively the same as the competitive exclusion model presented in the body of this paper (Figure: \ref{fig:LambdaMeans}).
+The second alternative we tested was a model that assumed rather than competition driving ant occupations, it was the population size of each ant species alone.
+In this hypothetical simulation when one ant species is removed from the model the cacti are left vacant.
+We believe this is an unlikely scenario as there appear to be vast populations of the ant species in the area. \ali{}{This is just personal observation so it might be best to remove this?}
+This simulation was designed to test if the fitness was more sensitive to changes in vacancy across the population than the frequency of any particular ant species.
+The results of this simulation revealed that when vacancy increased sampling effect is at play (Appendix \ref{appendix:C}: Figure \ref{app:FreqLambdaMeans}).
+This result highlights that the primary driver behind the results we presented are ant presence and that it is only in cases oh unrealistically high vacancy that there are benefits to partner diversity.
+Further information about both of these models can be found in Appendix C.
+Together, these results indicate that our assertions that the fitness of the cacti are more sensitive to changes in vacancy than partner identity are robust across multiple assumptions.
 
@@
 Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-\tom{The long term data set we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.}{This is a great message to emphasize.}
+The long term data set we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
 By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of specific demographic effects. 
@@
 
-%First, we were able to look at many years of data to reveal the complex dynamics of partner turnover and identity, revealing that despite the overwhelming  frequency of \textit{L.
-apiculatum} they do not impact the fitness of the cholla more than any other ant observed.
-%Fewer years of data would not have been able to show the distinct patterns of \textit{L. apiculatum} taking over cacti that were previously tended by any other partners with as much clarity.
-%Second, our long-term data set allowed us to test the portfolio effect.
-%This data is required, because we must show that the partners react asynchronously to temporal environmental stochasticity for portfolio effect to even be possible.
-%This is a great example of how long-term data may offer more power to test underlying mechanisms in similar systems.
-%Finallly, previous studies by \cite{Miller2007} hypothesized that \textit{C. opuntiae} may be a parasite based on the effects on the cholla cacti.
-%Using more years of consecutive data in this system, we were able to show that \textit{C.
-opuntiae}, while the worst individual partner in terms of the overall cholla fitness (Figure \ref{fig:LambdaMeans}), increases the growth and survival rates of cacti prior to reproduction.
-%The length of our data set has allowed us to dive into individual vital rates as well as overall fitness with more clarity.
-
 % Herbivory
-Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti.
+Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti \citep{Miller2009}.
 We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
+Analysis of our observational herbivore data shows that vacant plants experience elevated levels of herbivory compared to tended ones (Appendix \ref{appendix:B}: Figure \ref{app:herb}).
 In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
-This is particularly true in light of our consideration of portfolio effect. 
-\tom{In the same way that different species of ants likely react to temporal environmental variation differently, the same can be said for herbivores \citep{Thibaut2012}.
-Herbivores have variations in their population reflective of the environment changing temporally, meaning there are "good" and "bad" years for herbivores just like for ant defenders \citep{Batstone2018}, so the strength of their negative effect varies.
-The fluctuations in herbivore populations in conjunction with ant-specific defenses agains herbivores provide a solid foundation to study portfolio effect in this system. 
-We found no evidence of this mechanism (Figure \ref{fig:Portfolio}), but a more herbivore explicit model in the future may reveal more about driving mechanisms. 
-}{This confusing to me.
-Herbivores ARE the portfolio effect.
-We are not modeling ants becomes more or less abundant, we are modeling their demographic effects fluctuating, which could only be due to herbivores fluctuating.}
 
@@
 As with any study, there are limitations to consider when interpreting the results reported here.
-These results are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
-We are able to report evidence that plants experience higher growth when associating with a particular partner, but we are not able to determine if the presence of the ant is the cause of the higher growth observed.
-It is possible, for example, that plants which attract \textit{L.
-apiculatum} through osition have a different pattern of resource allocation to growth and survival and this is what we are observing rather than a direct affect of \textit{L.
-apiculatum} on the plant.
-This is a typical limitation of observational studies, but does not reduce the power or importance of our results, just the interpretation.
-However, we believe that the ants likely have direct impacts on the plant demography based on other studies which find the presence of ants can lead to direct increases in survival \citep{Beattie1985}, reproductive efforts \citep{Ford2015}, growth \citep{Frederickson2005}, etc.
-\tom{In the future, it would be interesting to verify the causation of our results through experimental manipulations.}{Ali, we have done these!
-I think this paragraph needs to be re-writted to acknowledge that, because it gives us a lot more confidence in interpreting the observational trends.}
-One further expansion could include nectar analysis of these cacti in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
-
-\tom{Another limitation of our study is that we are simulating populations through an IPM to estimate fitness, meaning we have to extrapolate beyond observed interaction.
-In our study, small plants are rarely seen with partners, yet we report on the expected effects of each ant on the growth and survival of plants from the size of new recruits to the largest plants. 
-While we are confident in our model, there remains the small possibility of interactions which have never been observed occurring (such as a seedling tended by \textit{L. apiculatum}).
-This means that our fitness estimations are based on extrapolations of our observed data, and while we have done what we can to be confident in our model, it is possible that our assumptions that observed patterns would continue beyond observed interactions could be incorrect.
-Another example of this includes our single partner simulations.
-We have no data on what the fitness of cholla would look like if the only ant partner possible was \textit{C.
-opuntiae}, yet we offer an estimation based on the effects of this partner in a diverse system.
-The many years of data and 10467 unique plant-year observations included in our analysis allow us to have relatively high confidence in our outputs despite extrapolations.}{I am not convinced this paragraph is necessary.
-The topic of creating ``counterfactual'' scenarios could be incorporated into the paragraph I am suggesting above, intepreting the different scenarios of turnover.}
-
-\tom{}{I strongly suggest a conclusion paragraph that brings the reader back to the big picture, what we learn from this study in the broadest sense, and why it matters.
-Is this how mutualisms work generally?
-How commonly and under what conditions should we expect diversity to not matter, as it appears not to matter here?}
+These results highlighted in the paper are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
+However, we have previously conducted experimental manupulations which revealed that ant presence has direct impacts on plant demography in our system \citep{Miller2007}.
+This combination of observational data backed up by experimental results gives us greater confidence in our causational interpretations than if we had only observational data.
+Specifically, we are confident in claiming that ant presence causes increased performance in cactus vital rates and the fitness signals we are detecting are directly related to ant partner presence.
+One further expansion could include nectar analysis in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+
+% tie back paragraph.
+This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangable to a focal mutualist.
+The individual estimation of each vital rate and explicit inclusion of ant partners in this IPM allowed us to determine that partner diversity is not beneficial in this system and may highlight circumstances under which we would expect similar results in other systems.
+The individual vital rate estimation approach allowed us to determine that while different ant partners did impact each vital rate differently, the largest differences were estimated to occur at sizes when ants would not be present in real life or in vital rates that do not have a significant impact on the overall fitnes.
+Considering partner imapcts on vital rates across the entire life cycle of the plant allows more power in determining the importance of these differences.
+Based on our results we would expect to see that partner diversity does not matter when partner impacts occur during low sensitivity life stages.
 
 \section*{Acknowledgments}
-\tom{This should be drafted.}{}
+We are grateful to the Sevilleta community (LTER and US Fish and Wildlife Service) for providing a stimulating research environment and logistical support. 
+Jeremiah Dye identified ants and Massa Takahashi identified herbivores. 
+We acknowledge the many students and technicians who have helped collect annual census data for this project, including M.
+Donald, J.
+Fowler, T.
+Jordan-Millet, J.
+Moutouama, C.
+Oxley, K.
+Schraeder, B.
+Scherick, A.
+Sears, M.
+Tucker, and J.
+Xiong.
+Financial support for this work came from the Sevilleta LTER (NSF DEB-0217774)
+
 %%%%%%%%%%%%%%%%%%%%%
@@
 \section*{Data and Code Availability}
-\tom{This should be drafted.}{}
+The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1
 
@@
 
-\section*{Appendix A: Additional Methods and Parameters}
-In addition to the models described in the body of the paper, we fit several other simpler models using data from previous studies.
+\appendix
+\label{appendix}
+\section*{Appendix A: Additional Methods and Parameters} \label{appendix:A}
+In addition to the models described in the body of the paper, we fit several other models using data from previous studies.
 These models are described below.
@@
 \paragraph{Seeds Per Fruit}
-With data\cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+With data from \cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
 We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
@@
 
-\section*{Appendix B: Observed Herbivory Data}
-One type of data we took every year is herbivory data, which involved looking at the plant for evidence of recent herbivory (fresh bites) and noting the type and quantity of herbivores observed.
-
-\section*{Appendix C: Alternative Ant Transition Simulations}
+\section*{Appendix B: Observed Herbivory Data} \label{appendix:B}
+Herbivory is an important driver in this population and shapes the range and demography of cholla.
+Herbivore presence has been shown to negatively impact growth and fecundity of cholla populations \citep{Miller2009}.
+Ant visitors are believed to offer defensive benefits to the plants they tend in this system, leading to the hypothesis that ant presence would be correlated with reduced herbivory.
+Herbivory data was collected during censuses any time herbivores were identified on a plant. 
+This involved noting the type and quantity of herbivores observed. 
+This data has been taken consistently since 2017, so the analysis below considers 6 years of data.
+We considered only plants which were reproducing, as they were likely to produce the highest quantity of EFN \citep{Miller2014}.
+The proportion of reproducing plants that experienced herbivory was calculated for each ant state separately.
+Analysis showed that ant presence is correlated with lower herbivore visitation.
+40\% of vacant cacti experienced herbivory.
+Plants tended by Other ants experienced similar, though lower, levels of herbivory on reproducing plants, with herbivores detected on 37.5\% of plants.
+Herbivores were detected on 25\% of plants tended by \textit{C. opuntiae} ants and on 11\% of plants tended by \textit{L. apiculatum} ants.
+These results indicate that ant presence is correlated with lower levels of herbivory and that partner identity has an impact on the level of herbivory.
+They also indicate that the partner correlated with the lowest levels of herbivory is \textit{L. apiculatum} ants.
+These findings are consistent with literature findings which show that \textit{L.
+apiculatum} ants are the most aggressive (therefore the most effective against herbivores), but differ from previous findings that \textit{C.
+opuntiae} may not offer defensive benefits \citep{Miller2007}.
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/herb_all.png}
+	\caption{Shows the proportion of reproducing plants which are visited by herbivores. Each bar represents the subset of the cacti population in a different ant state.  }
+	\label{app:herb}
+\end{figure}
+
+\section*{Appendix C: Alternative Ant Transition Simulations}\label{appendix:C}
 In addition to the competitive exclusion model defined and analyzed in the body of the paper, we simulated results from several other potential models. 
@@
 \begin{figure}
-\includegraphics[width=0.61\linewidth]{Figures/Lambdas_freq.png}
-\includegraphics[width=0.39\linewidth]{Figures/freq_num_partners.png}
-\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
-The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Freq_lines.png}
+	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+The lines show the posterior distribution spread of estimated $\lambda$ values.
+The letters in the legend correspond to what ant partners are present (V = Vacant, C = \textit{C.
 opuntiae}, L = \textit{L.
 apiculatum}, O = other).
-Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+}
 \label{app:FreqLambdaMeans}
@@
 Despite very different proportions, we found very similar outcomes to the competitive exclusion model analyzed in the paper. 
-All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:EqualLambdaMeans}a).
-Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:EqualLambdaMeans}b).
-
-\begin{figure}
-\includegraphics[width=0.61\linewidth]{Figures/Lambdas_equal.png}
-\includegraphics[width=0.39\linewidth]{Figures/equal_num_partners.png}
-\caption{Panel a) shows the mean values of the estimated $\lambda_{S}$ (filled in circles) for each simulated combination of ant partners.
-The letters above the points correspond to what ant partners are present (V = Vacant, C = \textit{C.
+All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:EqualLambdaMeans}b).
+Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:EqualLambdaMeans}a).
+
+\begin{figure}
+\includegraphics[width=0.91\linewidth]{Figures/Lambdas_equal_lines.png}
+	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+The lines show the posterior distribution spread of estimated $\lambda$ values.
+The letters in the legend correspond to what ant partners are present (V = Vacant, C = \textit{C.
 opuntiae}, L = \textit{L.
 apiculatum}, O = other).
-Panel b) shows the mean values of the estimated $\lambda_{S}$ for }
+}
 \label{app:EqualLambdaMeans}
@@
 
-\section*{Appendix D: Posterior Checks and Model Validation}
+\section*{Appendix D: Posterior Checks and Model Validation}\label{appendix:D}
 For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 
@@
 Each viable flower on a plant produces between 97 and 257 seeds.
-\tom{This number is affected by the ant partner present. 
-	\textit{C. opuntiae} tended plants produce a mean of 148 seeds per flower. 
-	\textit{L.
-apiculatum} tended plants produce a mean of 149 seeds per flower.
-}{These results are not consistent with Ohm and Miller, where Crem had lower seeds than Liom.
-I would check this.
-This section should also reference that paper because these are not new results.}
-Vacant plants produce a mean of 160 seeds per flower. 
-We are 73\% and 78\% confident that vacant plants produce more seeds per flower on average than plants tended by \textit{C. opuntiae} and \textit{L. apiculatum} ants respectively.
-
-\tom{%% Precensus Survival\
+This number is affected by the ant partner present, as shown in previous work \citep{Ohm2014}. 
+\textit{C. opuntiae} tended plants produce a mean of 115 seeds per flower. 
+\textit{L. apiculatum} tended plants produce a mean of 143 seeds per flower.
+Vacant plants produce a mean of 148 seeds per flower. 
+Comparison between posterior distributions revealed that \textit{C.
+opuntiae} tended plants produced fewer seeds per flower than \textit{L.
+apiculatum} tended plants and vacant plants 80\% and 87\% of the time.
+Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
+We are confident that \textit{C. opuntiae} tended plants produce the fewest seeds per flowers.
+
+%% Precensus Survival\
 	\paragraph{Precensus Survival Model}
@@
 	%% Recruit size distribution
-	New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.}{Move to an appendix. These results are not relevant for the questions at hand.}
+	New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.
 

### 2024-10-01 Manuscript/ant_cactus_AmNat.tex
Updates

@@
 \noindent The first term in Eq.
-\tom{\ref{eqn:IPM2}}{We should label equations.
+\tom{\ref{eqn:IPM2}}{We should label s.
 I am not sure why the equation label is not working here and I did not try to figure it out.
@@
 We found evidence that ant visitation by \textit{L. apiculatum} ants leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
-We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:viab}).
+We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:Viab}).
 Ant partners influence the mean viability rate of flowers, with \textit{L.
@@
 \begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/viab.png}
+	\includegraphics[width=0.95\linewidth]{Figures/Viab_v2.png}
 	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
@@
 \begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/year_ant_timeseries.png}
+	\includegraphics[width=0.95\linewidth]{Figures/ant_RFX.png}
 	\caption{This figure shows the mean affect of each ant partner on a) the estimated next size, b) the estimated survival, and c) the floral viability of cacti across every year of our study.
@@
 From the means alonet appears that there is a reduction in fitness when all ants are present compared to 1 or 2 partners.
-The contrasts of $\lambda_{NS}$ posterior distributions of 0 - 2 partner diversity scenarios to full diversity show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
+The contrasts of $\lambda_{NS}$ posterior distributions of 0 - 2 partner diversity scenarios to $\lambda_{NS}$ posterior distribution of 3 partners show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
 This is highlighted in Figure \ref{fig:LambdaMeans} panel b.
@@
 The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are 98\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scenarios.
+We are are 100\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scenarios.
 When subtracting these two resulting vectors from each other (($\lambda_{NS,All} - \lambda_{NS,Vacant}$) - ($\lambda_{S,All} - \lambda_{S,Vacant}$)), we found that there is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.

### 2024-10-01 01_cholla_ant_IPM_setup.R

@@ -38,7 +38,7 @@ library(ggpubr)
 library(reshape)
 library(reshape2)
 library(stats)
-knitr::oknitr::oknitr::opts_chunk$set(echo = TRUE)
+# knitr::oknitr::oknitr::opts_chunk$set(echo = TRUE)
 # check the workign directory and make sure it is set to the right location
 #getwd()
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")

### 2024-10-01 03_cholla_ant_IPM_params_functions.R

@@ -10,7 +10,7 @@
 ################################################################################
 ## Source the IPM vital rates code 
 #setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
-#source("02_cholla_ant_IPM_vital_rates.R")
+# source("02_cholla_ant_IPM_vital_rates.R")
 ## Set conditions for the IPM 
 cholla_min<- min((cactus$logsize_t), na.rm = TRUE)  ## minsize 
 cholla_max<- max((cactus$logsize_t), na.rm = TRUE)  ## maxsize 

### 2024-10-01 05_cholla_ant_IPM_figures_tables.R

@@ -106,61 +106,46 @@ y_vac_subset_grow <- subset(y_subset, ant_t == "vacant")
 y_other_subset_grow <- subset(y_subset, ant_t == "other")
 ## Predicted sizes for each partner condition
 omega <- exp(mean(params$grow_sig0) + size_dummy*mean(params$grow_sig1))
+omega_other <- exp(mean(params$grow_sig0) + size_other*mean(params$grow_sig1))
+omega_crem <- exp(mean(params$grow_sig0) + size_crem*mean(params$grow_sig1))
+omega_liom <- exp(mean(params$grow_sig0) + size_liom*mean(params$grow_sig1))
+omega_vac <- exp(mean(params$grow_sig0) + size_vac*mean(params$grow_sig1))
 alpha <- exp(mean(params$grow_alp0) + size_dummy*mean(params$grow_alp1))
-# omega_crem <- exp(mean(params$grow_sig01) + size_dummy * mean(params$grow_sig11))
-# omega_liom <- exp(mean(params$grow_sig02) + size_dummy * mean(params$grow_sig12))
-# omega_other <- exp(mean(params$grow_sig03) + size_dummy * mean(params$grow_sig13))
-# omega_vac <- exp(mean(params$grow_sig04) + size_dummy * mean(params$grow_sig14))
-# alpha_crem <- mean(params$grow_alp01) + size_dummy * mean(params$grow_alp11)
-# alpha_liom <- mean(params$grow_alp02) + size_dummy * mean(params$grow_alp12)
-# alpha_other <- mean(params$grow_alp03) + size_dummy * mean(params$grow_alp13)
-# alpha_vac <- mean(params$grow_alp04) + size_dummy * mean(params$grow_alp14)
+alpha_other <- exp(mean(params$grow_alp0) + size_other*mean(params$grow_alp1))
+alpha_crem <- exp(mean(params$grow_alp0) + size_crem*mean(params$grow_alp1))
+alpha_liom <- exp(mean(params$grow_alp0) + size_liom*mean(params$grow_alp1))
+alpha_vac <- exp(mean(params$grow_alp0) + size_vac*mean(params$grow_alp1))
 # Other
+size_other <- seq(min(y_other_subset_grow$logsize_t1),max(y_other_subset_grow$logsize_t1), by = 0.1)
 y_other_mean_grow <- mean(params$grow_beta03) + (size_dummy) * mean(params$grow_beta13 ) + (size_dummy)^2 * mean(params$grow_beta23)
[…]

### 2024-10-02 Manuscript/ant_cactus_AmNat.tex
results edits

@@
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We measured reproduction by counting flowerbuds, and we distinguished between flowerbuds that were viable and aborted. 
+We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable and aborted. 
 We recorded the ant species present (or vacancy if no ants present).
@@
 In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
+In addition to missing the year 2020, there are gaps in the time series where were switched plots or plants (and thus broke up transition years for growth and survival) or where we did not distinguish between viable and aborted flower buds (Table \ref{}).
 
@@
 For stochastic analyses, we estimated the approximating matrix corresponding to each $t$ to $t+1$ transition year. 
-To estimate population mean fitness in a stochastic environment where ants responded to annual variation uniquely \tom{($\lambda_{NS}$)}{Let's discuss this symbology.} we simulated population dynamics for 500 years by randomly sampling among the 16 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
+To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the 15 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
 Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
 We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$ \citep{Mark2009}.
-We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{NS}$ and other derived quantities.
+We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
 
 \subsubsection*{Partner diversity simulation experiments}
-Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness in a stochastic environment ($\lambda_{NS}$).
+Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness. 
 From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created treatments corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
@@
 For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants across the size distribution.
+We refer to stochastic fitness associated with partner number or identity using superscripts, e.g.
+$\lambda^{0}_{S}$ for vacant plants (zero partners), $\lambda^{1+}_{S}$ for any state of ant tending, $\lambda^{C}_{S}$ for tending by only \textit{C.
+opuntiae}, $\lambda^{CO}_{S}$ for tending by \textit{C.
+opuntiae} and Other ants, etc.
 
@@
 Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
-We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis by exploring two scenarios of environmental variation.
-The 'non-synchronous' (NS) version included ants effects that varied uniquely across time and the 'synchronous' (S) version included  ant effects that were forced to be the same across species. 
-First, we evaluated the model using empirical estimates for the $\pmb{\tau}_{a}$ vectors that describe ant-specific year deviates. 
-In this scenario, good years and bad years can differ between ant states, according to the empirical parameter estimates. 
-We also quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states.
-Second, we averaged the ant-dependent vital rates (survival, growth, flowerbud viability) across ant-specific year random effects, thus ensuring that plants in all ant states fluctuated synchronously in response to temporal environmental variation.
-We evaluated a second, ``synchronized'' version of the model that effectively turns off any portfolio effect, holding all else equal. 
+We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis.
+The baseline, 'non-synchronous' model described above included ants effects that could vary uniquely across time. 
+In this scenario, good years and bad years can differ between ant states, according to the parameter estimates for the random effects ($\pmb{\tau}_{a}$). 
+We quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states for each vital rate.
+The 'synchronous' (S) version included  ant effects that were forced to be the same across species. 
+To synchronize ant states, we averaged the ant-specific year random effects, thus ensuring that plants in all ant states fluctuated synchronously in response to temporal environmental variation. 
+This second, ``synchronized'' version of the model effectively turns off any portfolio effect, holding all else equal. 
 Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above. 
+We indicate stochastic fitness under the synchonrized scenario with $\lambda^{C(s)}_{S}$, for example.
 
@@
 The range of models we created generated many outputs; we focus our inference on the following specific contrasts. 
-First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{NS}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
 If cactus fitness increases with partner richness, this would be interpreted as evidence for benefits of partner diversity. 
@@
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with an opportunity cost of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{NS}$ of the full (four-state) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity (as a measure of how much benefit is recieved from all partners being present).
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy (as a measure of how much benefit is receieved from all partners being present) for synchronized and non-synchronized responses to temporal stochasticity.
 If the portfolio effect confers a benefit of diversity, the fitness advantage of having all vs. no partners should be greater when temporal fluctuations are not synchronized across ant states.
@@
 We base our statistical inferences on the posterior probability distributions of the contrasts described above. 
-For example, the contrast of \textit{C. opuntiae} with vacancy yields a posterior distribution of the difference in $\lambda_{NS}$ ($\Delta\lambda_{C-V}$). 
-We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda_{C-V}>0)$. 
+For example, the contrast of \textit{C. opuntiae} ($\lambda^{C}_{S}$) with vacancy ($\lambda^{V}_{S}$) yields a posterior distribution of the difference $\Delta\lambda^{C-V}_{S}$. 
+We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda^{C-V}_{S}>0)$. 
 We apply similar logic to other contrasts described above.
@@
 \section*{Results}
-\subsection*{What are the demographic effects of association with alternative parnters and how do these effects fluctuate across years?}
+\subsection*{What are the demographic effects of association with alternative partners and how do these effects fluctuate across years?}
 Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
-Plants tended by \textit{C.
-opuntiae} had a growth advantage, particularly for small plants, while plants in states of \textit{L.
-apiculatum}, Other ants, and vacancy had indistinguishable growth trajectories (Figure \ref{fig:Grow}).
+Ant-tended plants had a growth advantage over vacant plants, especially at smaller sizes (Figure \ref{fig:Grow}).
+Over the size range for which plants were likely ant-tended (larger than the minimum observed size of $0.8$ $log(cm^3)$; solid lines in Figure \ref{fig:Grow}) there were modest differences between partner species, with the greatest growth advantage associated with \textit{C.
+opuntiae} followed by \textit{L.
+apiculatum} and then Other ants.
+At large sizes the growth trajectories of ant-tended and vacant plants were nearly indistinguishable. 
 For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean.
-Similarly, for plants which were large enough to have ant visitors (larger than $0.8$ $log(cm^3)$), visitation enhanced cactus survival (Figure \ref{fig:Surv}).
-Plants that were smaller than $-2$ $log(cm^3)$ (indicated by the dashed lines for ant estimates, Figure \ref{fig:Surv}) were estimated to experience survival boosts from vacancy.
-Due to the assumptions in our models based on real ant presence data, tended plants of the size where vacancy would offer a boost are never seen and therefore do not influence the fitness estimates of cacti.
-\ali{}{I'm not 100\% certain on how to discuss this.
-I know that this still shows that crem offer a survival advantage, but I see that it could look like vacancy offers survival advantages.
-}
-The smallest plants showed evidence that vacancy resulted in the highest survival rates, however these plants are never tended, meaning the models of ant-tended cacti survival before the solid lines are extrapolation.
+
+Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). 
 Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
@@
 At larger sizes, plants in any state of ant occupancy had a survival advantage over vacant plants. 
+Plants that were smaller than $-2$ $log(cm^3)$ were predicted to experience a survival advantage from vacancy, but apparent benefits of occupancy were based on extrapolated survival estimates of ant-tended plants.
+Plants in this size range were never observed to be ant-tended and, because the IPM preserves the size-dependence of vacancy, benefits of vacancy at small sizes are never realized in the IPM.
 
@@
 
-We found evidence that ant visitation by \textit{L. apiculatum} ants leads to increased floral viability rates and that ant identity can influence the strength of viability benefits.
+We found evidence that ant tending was associated with increased floral viability rates and that ant identity can influence the strength of viability benefits.
 We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:Viab}).
-Ant partners influence the mean viability rate of flowers, with \textit{L.
-apiculatum}-tended plants experiencing the highest mean viability rate (86\%), followed by vacant and Other tended plants (at 60-61\%), and \textit{C.
-opuntiae} tended plants had the lowest floral viability rate (57\%).
+\textit{L.
+apiculatum}-tended plants had the highest mean viability rate (86.1\% [95\% credible interval: 77.6--92.4\%]), and there were similar viability rates for vacant (60.0\% [44.3--75.0\%]), Other-tended plants (at 60.6\% [43.7--75.5\%]), and \textit{C.
+opuntiae} tended plants (57.1\% [40.6--72\%]).
+Furthermore, \textit{C.
+opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
+apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure \ref{app:AppA_Seeds_Per_Fruit}).
 
@@
 
-We analyzed all other vital rates mentioned in the methods as well.
-The results of each of these as well as the posterior predictive checks are all included in the Apendix \ref{appendix:A} : Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
-
-We analyzed the correlation coefficients of all models which included ant state as a predictor and found that the annually varying effects of each ant on growth were the most correlated (correlation coefficients averaged 0.63), and that the effects on survival were the least correlated (correlation coefficients averaged 0.36).
-High correlation indicates synchronicity, which is necessary for portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
-For the growth model, annually varying randomn effects of Other ants were the least synchronous (correlation coefficients ranging from 0.28 to 0.39) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.39 to 0.81).
-For the viability model, annually varying random effects of \textit{L.
-apiculatum} ants were the least synchronous (correlation coefficients ranging from 0.06 to 0.44) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.44 to 0.62).
-For the survival model, annually varying random effects of \textit{C.
-opuntiae} ants were the least synchronous (correlation coefficients ranging from 0 to 0.15) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.13 to 2.6).
-This variety of synchronicity across ant states and vital rates indicates there is potential for portfolio effect as many of the ants effects revealed low synchronicity (low correlation of annual effects, indicating that ant identity impacts annual random effects), particularly in the survival model.
+The results of all other vital rates as well as the posterior predictive checks are all included in the Apendix \ref{appendix:A} : Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
+
+We analyzed the correlation coefficients of all models that included ant state as a predictor and found that, across vital rates, the annually varying effects on growth were the most correlated between ant states (mean pairwise correlation: 0.63) and effects on survival were the least correlated (0.36) (Figure \ref{fig:Annual_Ant}).
+%Weakly or uncorrelated fluctuations are necessary for the portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
+For growth, annually varying random effects of Other ants were the least synchronous with other ants states (correlation coefficients ranging from 0.28 to 0.39) while the effects of vacancy were the most synchronous with other states (correlation coefficients ranging from 0.39 to 0.81).
+For survival, effects of \textit{C.
+opuntiae} were the least synchronous with other ant states (correlation coefficients ranging from 0 to 0.15) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.13 to 2.6).
+For floral viability, effects of \textit{L.
+apiculatum} were the least synchronous with other ant states (correlation coefficients ranging from 0.06 to 0.44) while effects of vacancy were the most synchronous (correlation coefficients ranging from 0.44 to 0.62).
+
+The variation in synchrony across ant states and vital rates indicates there is potential for portfolio effect, as many of the ant effects revealed low synchronicity, particularly in survival.
 
@@
 We found that 55\% of individual plants surveyed in the long-term data experienced at least one ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
-Vacancy is the most likely ant state of small plants ($\leq 10 log(cm^3)$).
-Even when small plants are ant-tended at the start of the transition year, they are most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
-The probability of becoming ant-tended increases with size, though it is not equally likely to be tended by all partners. 
-For large plants that are initially vacant or tended by \textit{L.
+Vacancy was the most likely ant state of small plants ($\leq 10 log(cm^3)$).
+Even when small plants were ant-tended at the start of the transition year, they were most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
+The probability of becoming ant-tended increases with size, though it was not equally likely to be tended by all partners.  
+For large plants that were initially vacant or tended by \textit{L.
 apiculatum} or Other ants, \textit{L.
-apiculatum} is the most likely next partner, suggesting that this partner species is able to colonize plants that were previously vacant or occupied by Other ants, and effectively retain plants that it previously occupied.
+apiculatum} was the most likely next partner, suggesting that this species is able to colonize plants that were previously vacant or occupied by Other ants, and effectively retain plants that it previously occupied.
 \textit{C.
@@
 opuntiae} and Other  were often transitions to other ants.
-This suggests a competitive hierarchy whereby \textit{L. apiculatum} may abandon low-value plants with litte nectar production but is almost never displaced from high-value plants.
+This suggests a competitive hierarchy whereby \textit{L. apiculatum} may abandon low-value plants with little nectar production but is almost never displaced from high-value plants.
 
@@
 \subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
-By integrating vital rate results and ant transition dynamics into the multi-state IPM we can evaluate the fitness implications of different scenarios of partner diversity and identity. 
+By integrating vital rate results, temporal fluctuations, and ant transition dynamics into the stochastic, multi-state IPM we can evaluate the fitness implications of different scenarios of partner diversity and identity.
 First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
-The lowest mean fitness was $\lambda_{NS,Vacant}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}b).
-Across all 1+ partner scenarios, we found that the posterior $\lambda_{NS,Occupied}$ distributions were greater than $\lambda_{NS,Vacant}$ distributions 100\% of the time.
+The lowest mean stochastic fitness was $\lambda^{0}_{S}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}a).
+Across all 1+ partner scenarios, we found that the $\lambda^{1+}_{S}$ posterior distributions were greater than $\lambda^{0}_{S}$ with nearly 100\% certainty.
 This indicates that ant visitation elevates fitness no matter the identity or number of partners.
-Furthermore, we find no benefits of patner diversity here, with the fitness of the cholla with 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
-From the means alonet appears that there is a reduction in fitness when all ants are present compared to 1 or 2 partners.
-The contrasts of $\lambda_{NS}$ posterior distributions of 0 - 2 partner diversity scenarios to $\lambda_{NS}$ posterior distribution of 3 partners show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
-This is highlighted in Figure \ref{fig:LambdaMeans} panel b.
+Furthermore, we found no benefits of partner diversity, with the fitness of cholla 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
+The 1- and 2-partner scenarios were statistically consistent but there was a modest reduction in fitness in the 3-partner scenario compared to 2 partners ($Pr(\Delta\lambda^{2-3}_{S}=0.95)$).
+%The contrasts of $\lambda_{NS}$ posterior distributions of 1 or 2 partner diversity scenarios to $\lambda_{NS}$ posterior distribution of 3 partners show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
+%This is highlighted in Figure \ref{fig:LambdaMeans} panel b.
 
@@
 \end{figure}
+
+Among the one-partner scenarios, there was no strong evidence for any single best partner species. 
+While \textit{L.
+apiculatum}-tended plants had the highest mean fitness, it was not significantly higher than \textit{C.
+opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) tended plants.
+Furthermore, the fitness of \textit{L. apiculatum}-tended plants was consistent with 1- and 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
+However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$).
 

### 2024-10-02 Manuscript/ant_cactus_AmNat.tex
Figures Updates

@@
  The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
-\textit{Cylindropuntia imbracata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
+\textit{Cylindropuntia imbricata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
 We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
@@
 
-This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindriopuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
+This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
 Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \cite{Miller2009}, and that ant defense reduces herbivore damage \cite{Miller2007}. 
@@
 apiculatum}-tended plants, and vice versa, provoke aggressive responses by residents (A.
-Cambpell, \textit{personal observation}).
+Campbell, \textit{personal observation}).
 %Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
@@
 While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
-We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explictly.
+We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explicitly.
 
@@
 
-Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants (cite Tom's data).
+Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants \citep{Miller2006}.
 This data only included  \textit{L. apiculatum} and \textit{C. opuntiae} ants, so we had to make an assumption about the seeds per fruit on plants tended by Other ants.
 We chose to use the vacancy estimates for plants tended by other ants, a decision which does not have a significant impact on the final results. 
-Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix XX.
+Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A \ref{appendix:A}.
 
@@
 Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above. 
-We indicate stochastic fitness under the synchonrized scenario with $\lambda^{C(s)}_{S}$, for example.
+We indicate stochastic fitness under the synchronized scenario with $\lambda^{C(s)}_{S}$, for example.
 
@@
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with an opportunity cost of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy (as a measure of how much benefit is receieved from all partners being present) for synchronized and non-synchronized responses to temporal stochasticity.
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy (as a measure of how much benefit is received from all partners being present) for synchronized and non-synchronized responses to temporal stochasticity.
 If the portfolio effect confers a benefit of diversity, the fitness advantage of having all vs. no partners should be greater when temporal fluctuations are not synchronized across ant states.
@@
 	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
-The solid colored lines (seen in all panels) are the next mean predicted size of cholla.
+The colored lines (seen in all panels) are the next mean predicted size of cholla.
+Dashed lines indicate extrapolation while solid lines indicate estimations within the range of observed data.
 The points (seen in panels a-d) are the observed data which informs these estimates.
-The black countour lines (seen in a-d) appear at 5\% increments showing where 5\%, 10\%, etc.
+The black contour lines (seen in a-d) appear at 5\% increments showing where 5\%, 10\%, etc.
 of the data is expected to fall.
-They grey dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).
-}
+They gray dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).}
 	\label{fig:Grow}
@@
 	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
-The solid lines indicate the estimated viability distribution.
-The colored histograms represent the observed viability rates of plants with that partner.
-}
+The clouds of points show the observed data, with the color representing the ant state.
+The solid lines indicate 95\% of the estimated viability posterior distribution.}
 	\label{fig:Viab}
@@
 	\includegraphics[width=0.95\linewidth]{Figures/ant_RFX.png}
-	\caption{This figure shows the mean affect of each ant partner on a) the estimated next size, b) the estimated survival, and c) the floral viability of cacti across every year of our study.
+	\caption{The colored lines in each panel show the mean effect of each ant partner on a vital rate across years (blank periods indicate a lack of transition data).
+The blue squares in the upper corner of each panel show the correlation matrix of how similar ant-specific annual responses are for each vital rate.
+Panel a) shows the effects on the estimated next size, b) the effects on estimated survival, and c) the effects on floral viability of cacti across every year of our study.
 These values are estimated from the fitted random effects of ant and year in our models.
@@
 	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+The lines show the posterior distribution spread of the estimated $\lambda$ values.
 Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
@@
 opuntiae}, L = \textit{L.
-apiculatum}, O = other).
-}
+apiculatum}, O = other).}
 	\label{fig:LambdaMeans}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
-	\caption{Shows the distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
+	\caption{Shows the estimated posterior distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
 	\label{app:AppA_Seeds_Per_Fruit}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/rec_size.png}
-	\caption{Shows the distribution of recruit sizes.
-The mean  size of a recruit is marked by the dark line in the green box, the green box shows the interquartile range of recruit sizes, and the dashed lines show the minimum and maximum of estimated recruit sizes.
-The points beyond these dashed lines are considered outliers.
-}
+	\caption{Shows the estimated posterior distribution of recruit sizes as the $log(volume)$ of cholla. }
 	\label{app:AppA_Recruit_Dist}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/germination.png}
-	\caption{Shows the probability of a plant germinating during its first year in the seedbank (left) or during the second year in the seedbank (right).
-The mean germination rates are shown by the dark line near the bottom of the green box.
-The green boxes show the interquartile ranges of the germination rates for year 1 and 2 plants.
-The dashed line shows the maximum germination rates observed.
-The points above that show the outliers of our data.
-}
+	\caption{Shows the estimated posterior distributions for germination probability of seeds in the first year seedbank (blue) and seeds in the second year seedbank (green).}
 	\label{app:AppA_Germ}
@@
 	\caption{Shows the distribution of the probability of plants which germinated between censuses surviving to the next census to be counted.
-The green line shows the predicted pre-census survival and the grey dashed line shows the actual observed pre-census survival.
-}
+The green line shows the estimated posterior distribution of pre-census survival.}
 	\label{app:AppA_Pre_Census_Surv}
@@
 In addition to the competitive exclusion model defined and analyzed in the body of the paper, we simulated results from several other potential models. 
-We chose to include competitive eclusion as our primary results in the paper because we believe it to be the most biologically realistic.
+We chose to include competitive exclusion as our primary results in the paper because we believe it to be the most biologically realistic.
 However, in building and testing of alternative models we found that the method of ants occupying plants significantly impacts the fitness of the population. 
@@
 opuntiae}, L = \textit{L.
-apiculatum}, O = other).
-}
+apiculatum}, O = other).}
 \label{app:FreqLambdaMeans}
@@
 opuntiae}, L = \textit{L.
-apiculatum}, O = other).
-}
+apiculatum}, O = other).}
 \label{app:EqualLambdaMeans}
@@
 \subsection*{Statistical Models -- Results}
-Below are the results reorted of all statistical models not described in the main body of the text.
+Below are the results reported of all statistical models not described in the main body of the text.
 
@@
 %% Precensus Survival\
-	\paragraph{Precensus Survival Model}
+	\paragraph{Pre-census Survival Model}
 	Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.

### 2024-10-02 05_cholla_ant_IPM_figures_tables.R

@@ -100,10 +100,14 @@ dev.off()
 
 ## Format the original data
 y_subset <- growth_data[,c("logsize_t1","ant_t", "logsize_t")]
-y_crem_subset_grow <- subset(y_subset, ant_t == "crem")
-y_liom_subset_grow <- subset(y_subset, ant_t == "liom")
-y_vac_subset_grow <- subset(y_subset, ant_t == "vacant")
-y_other_subset_grow <- subset(y_subset, ant_t == "other")
+y_crem_subset_surv <- subset(survival_data, ant_t == "crem")
+y_liom_subset_surv <- subset(survival_data, ant_t == "liom")
+y_vac_subset_surv <- subset(survival_data, ant_t == "vacant")
+y_other_subset_surv <- subset(survival_data, ant_t == "other")
+size_crem = seq(min((y_crem_subset_surv$logsize_t), na.rm = TRUE), max ((y_crem_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
+size_other = seq(min((y_other_subset_surv$logsize_t), na.rm = TRUE), max ((y_other_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
+size_liom = seq(min((y_liom_subset_surv$logsize_t), na.rm = TRUE), max ((y_liom_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
+size_vac = seq(min((y_vac_subset_surv$logsize_t), na.rm = TRUE), max ((y_vac_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
 ## Predicted sizes for each partner condition
 omega <- exp(mean(params$grow_sig0) + size_dummy*mean(params$grow_sig1))
 omega_other <- exp(mean(params$grow_sig0) + size_other*mean(params$grow_sig1))
@@ -116,19 +120,15 @@ alpha_crem <- exp(mean(params$grow_alp0) + size_crem*mean(params$grow_alp1))
 alpha_liom <- exp(mean(params$grow_alp0) + size_liom*mean(params$grow_alp1))
 alpha_vac <- exp(mean(params$grow_alp0) + size_vac*mean(params$grow_alp1))
 # Other
-size_other <- seq(min(y_other_subset_grow$logsize_t1),max(y_other_subset_grow$logsize_t1), by = 0.1)
[…]

### 2024-10-04 Manuscript/ant_cactus_AmNat.tex
results edits

@@
 We apply similar logic to other contrasts described above. 
+We interpret contrast probabilities $\geq0.95$ as statistically significant differences.
 
@@
 apiculatum} were the least synchronous with other ant states (correlation coefficients ranging from 0.06 to 0.44) while effects of vacancy were the most synchronous (correlation coefficients ranging from 0.44 to 0.62).
-
 The variation in synchrony across ant states and vital rates indicates there is potential for portfolio effect, as many of the ant effects revealed low synchronicity, particularly in survival.
@@
 \end{figure}
-\ali{}{Honestly not sure if I should include an image for this one or just report some values? We should discuss. }
 
@@
 
-Among the one-partner scenarios, there was no strong evidence for any single best partner species. 
+Partner identity and composition were not strongly consequential for plant fitness. 
+Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans}a). 
 While \textit{L.
@@
 opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) tended plants.
-Furthermore, the fitness of \textit{L. apiculatum}-tended plants was consistent with 1- and 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
+Furthermore, the fitness of \textit{L. apiculatum}-tended plants was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
 However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$). 
-
-We found that the difference in posterior distributions of any ant scenario ($\lambda_{NS,A} - \lambda_{NS,B}$, where A and B are arbitrary ant diversity scenarios that are not vacancy) were between 31\% and 83\% different.
-The posterior distributions of $\lambda$ for each diversity scenario are visualized in Figure \ref{fig:LambdaMeans} b.
-This high level of overlap between posterior distributions indicates that while each of these partners are beneficial, there does not appear to be a significant benefit of partner diversity within this system.
-
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
 Using the simulations where all ants had equal frequencies across sizes (further explained and analyzed in Appendix C), we found the same fitness patterns as in the competitive exclusion model described above.
-Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
-Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.
-
-We found no evidence of portfolio effect.
-The effect of all ant partners can be measured as $\lambda_{All} - \lambda_{Vacant}$ (Figure \ref{fig:Portfolio}).
-We are are 100\% confident that when all ants are present the cholla experience higher fitness than when no ants are present according to both the synchronized (S) and non-synchronized (NS) model scenarios.
-When subtracting these two resulting vectors from each other (($\lambda_{NS,All} - \lambda_{NS,Vacant}$) - ($\lambda_{S,All} - \lambda_{S,Vacant}$)), we found that there is no real difference between the two scenarios, meaning we have no evidence of portfolio effect.
+%Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
+%Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.
+
+We found no evidence that the portfolio effect generated positive effects of partner diversity, as effects of partner richness and composition were highly consistent between the baseline model and the model version that synchronized all ant states (Figure \ref{fig:LambdaMeans}).
+The effect of all ant partners can be measured as $\lambda^{3}_{S} - \lambda^{0}_{S}$.
+We have high confidence that this contrast is positive, and equally so for the synchronous and non-synchronous scenarios \tom{(Figure \ref{fig:Portfolio}).}{Now that the lambda figure shows synchronized and non-synchronized, I am not sure you need the histogram figure for portfolio effect.
+Maybe appendix.}
 
@@
 Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
-Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{fig:Viab}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in survival at small sizes, where mortality risk was high (Figure \ref{fig:Surv}) and the probability of ant tending was low (Figure \ref{fig:Ant_Transition}).
-At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of ant identity, and that result dominates integrated measures of fitness because of the high sensitvity of established plant survival.
+Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{fig:Viab}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high (Figure \ref{fig:Surv}) and the probability of ant tending was low (Figure \ref{fig:Ant_Transition}).
+At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitvity of $\lambda$ established plant survival.
 The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness.
@@
 However, we have previously conducted experimental manupulations which revealed that ant presence has direct impacts on plant demography in our system \citep{Miller2007}.
-This combination of observational data backed up by experimental results gives us greater confidence in our causational interpretations than if we had only observational data.
+This combination of observational data backed up by experimental results gives us greater confidence in our causal interpretations than if we had only observational data.
 Specifically, we are confident in claiming that ant presence causes increased performance in cactus vital rates and the fitness signals we are detecting are directly related to ant partner presence.
-One further expansion could include nectar analysis in conjuntion with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+One further expansion could include nectar analysis in conjunction with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
 
 % tie back paragraph.
-This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangable to a focal mutualist.
+This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangable in their net effects on fitness of a shared mutualist.
 The individual estimation of each vital rate and explicit inclusion of ant partners in this IPM allowed us to determine that partner diversity is not beneficial in this system and may highlight circumstances under which we would expect similar results in other systems.
@@
 Donald, J.
-Fowler, T.
+Fowler, M.
+Saucedo, K.
+Dickens, T.
 Jordan-Millet, J.

### 2024-10-12 Manuscript/ant_cactus_AmNat.tex
Manuscript Updates -- Discussion

@@
  The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
-\textit{Cylindropuntia imbricata} (tree cholla) produce extrafloral nectar and various ant partners provide defense from herbivores and seed predators. 
+ To study multi-species demographic effects, we considered \textit{Cylindropuntia imbricata} (tree cholla), which produce extrafloral nectar and various ant partners ( \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) provide defense from herbivores and seed predators.
 We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-We constructed an Integral Projection Model in which we simulate different combinations of ant partners that don’t occur in nature.
-The hierarchical models revealed that different ant partners had different impacts on the cholla vital rates. 
+We constructed an Integral Projection Model (IPM) in which we simulate different combinations of ant partners that don’t occur in nature.
+Different ant partners had different impacts on the cholla vital rates. 
 Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
 The IPM results revealed that despite these differences in vital rates, the presence of any ant partner is beneficial, the identity and number of partners do not matter.
-This suggests that there are no benefits of diversity in this system.
+This suggests that there are no benefits of diversity in this system and therefore than none of the highlighted mechanisms explain the benefits of multi-species mutualisms.
 This study highlights that partner diversity does not always increase the benefits a focal mutualist receives.
@@
 \section*{Introduction}
+Cooperations between species, aka mutualisms, occur on a pairwise basis or in a more complex network and often play integral and foundational roles in ecological communities.
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \cite{Yeung2006,Barrett2015,Ushio2020}.
-First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \cite{Frederickson2013}.
-This can lead to an apparent benefit of diversity driven by a sampling effect \cite{Batstone2018} -- but there is no benefit of diversity \emph{per se}, only better and worse partners. 
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Yeung2006,Barrett2015,Ushio2020}.
+First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
+This can lead to an apparent benefit of diversity driven by a sampling effect \citep{Batstone2018} -- but there is no benefit of diversity \emph{per se}, only better and worse partners. 
 If partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
-Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \cite{Stachowicz2005,Bronstein2006,Stanton2003}. 
-This can lead to a positive effect of partner diversity through complementarity of alternative functions \cite{Batstone2018}. 
-Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \cite{Afkhami2014}. 
+Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \citep{Stachowicz2005,Bronstein2006,Stanton2003}. 
+This can lead to a positive effect of partner diversity through complementarity of alternative functions \citep{Batstone2018}. 
+Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}. 
 Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \cite{Batstone2018,Lazaro2022,Horvitz1990}.
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
 
 Partner diversity can have different effects depending on whether partners are present simultaneously or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
-Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \cite{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
+Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \citep{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
 Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
@@
 Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
-Finally, herbivore identity and pressure can vary inter-annually \cite{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
+Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
 Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
@@
 This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
-Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \cite{Miller2009}, and that ant defense reduces herbivore damage \cite{Miller2007}. 
+Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \citep{Miller2009}, and that ant defense reduces herbivore damage \citep{Miller2007}. 
 Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
@@
 	\begin{enumerate}	
-		\item{What are the demographic effects of association with alternative partners and how do these effects fluctuate across years?}
+		\item{What are the vital rate effects of association with alternative partners and how do these effects fluctuate across years?}
 		\item{What are the frequency and direction of partner turnover across the plant life cycle?}	
@@
 Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a')$) is dependent on ant state $a$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a)$) is dependent on ant state $a$. 
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
@@
 \end{linenomath*}
-Here, the location parameter for the $i$th observation $\hat{G_i}$ is defined as a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship (note that the location parameter of the skew-normal is not the mean, but the mean can be derived as $\hat{G} + \frac{\sigma\alpha}{\sqrt{1+\alpha^2}} \sqrt{\frac{2}{\pi}}$).
+Here, the location parameter for the $i$th observation $\hat{G_i}$ is defined as a second-order polynomial with ant-size interactions because  preliminary analysis found this was an improvement over a linear relationship.
+The location parameter of the skew-normal is not the mean, but the mean can be derived as $\hat{G} + \frac{\sigma\alpha}{\sqrt{1+\alpha^2}} \sqrt{\frac{2}{\pi}}$. 
 The year- and ant-specific random effect $u$ (which parameterizes the $\pmb{\tau}^G_{a}$ vectors) and plot-specific random effect $w$ are normally distributed with variances $\sigma^2_{year}$ and $\sigma^2_{plot}$, respectively.
@@
 We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
-This model used a logit link function and included random effects for plot and year following the same structure as the growth and survival models, with ant-specific year random effects.
+This model used a logit link function and included random effects for plot and year following the same structure as the survival model, with ant-specific year random effects.
 
@@
 \paragraph{Parameter estimation}
-We fit models using Stan run through version 4.0.2 of R \cite{Rcite,Rstancite}. 
+We fit models using Stan run through version 4.0.2 of R \citep{Rcite,Rstancite}. 
 We used vague priors for all parameters. 
 For each model, we obtained three chains of 10,000 iterations, discarding the first 1,500 iterations. 
-We visually assessed parameter convergence between and within chains (Appendix \ref{appendix:D}: Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Appendix \ref{appendix:D}: Figures \ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post} a).
+We visually assessed parameter convergence between and within chains (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}a).
 
@@
 In our model there is additional complexity in the form of transitions between $A$ ant states and two additional discrete states (year one and year two seed banks), leading to a matrix size of $A(b+2) \times A(b+2)$.
-We used $b = 500$ bins, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \cite{Williams2012}.
+We used $b = 500$ bins, which we found to be sufficient for numerically stable outputs, and extended the integration limits beyond the minimum ($L$) and maximum ($U$) observed sizes to avoid unintentional eviction using the ``floor-and-ceiling'' method \citep{Williams2012}.
 
@@
 To estimate population mean fitness in a stochastic environment ($\lambda_{S}$) we simulated population dynamics for 500 years by randomly sampling among the 15 annual transition matrices, discarding the first 100 years of the simulation to minimize the influence of initial conditions.
-Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \cite{metcalf2015statistical}.
-We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as $log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$ \citep{Mark2009}.
+Sampling observed transition matrices (rather than independently sampling regression coefficients) produces demographic time series that realistically capture inter-annual variation by preserving correlations between vital rates \citep{metcalf2015statistical}.
+We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as 
+$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \citep{Mark2009}.
 We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
@@
 This means that, in scenarios that remove species from the partner community, remaining species gain access to plants that the removed species would have tended. 
-In Appendix \ref{appendix}, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
+In Appendix C, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
 
@@
 \section*{Results}
-\subsection*{What are the demographic effects of association with alternative partners and how do these effects fluctuate across years?}
+\subsection*{What are the effects of association with alternative partners on vital rates and how do these effects fluctuate across years?}
 Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
@@
 	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour_v2.png}
-	\caption{This figure shows the next predicted size of cholla based on previous size with each individual ant partner.
+	\caption{The next predicted size of cholla based on previous size with each individual ant partner.
 The colored lines (seen in all panels) are the next mean predicted size of cholla.
@@
 of the data is expected to fall.
-They gray dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).}
+Panel e shows all mean estimates from panels a-d on one graph.
+The gray dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).}
 	\label{fig:Grow}
@@
 \includegraphics[width=0.95\linewidth]{Figures/survival_plot.png}
-\caption{This figure shows the estimated survival rates based on the size of the cactus with each individual ant partner.
+\caption{The estimated survival rates based on the size of the cactus with each individual ant partner.
 The solid colored lines (shown on all panels) indicate the mean estimated survival rates.
@@
 The colored dots are the real data binned by size to show how our estimates align with real survival observations.
-A larger circle means we had more data on survival of plants of this size with this partner.}
+A larger circle means we had more data on survival of plants of this size with this partner.
+The mean estimates from panels a-d are combined on one graph in panel e.}
 \label{fig:Surv}
@@
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
-apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure \ref{app:AppA_Seeds_Per_Fruit}).
+apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure S\ref{app:AppA_Seeds_Per_Fruit}).
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/Viab_v2.png}
-	\caption{This figure shows the estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
+	\caption{The estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
 The clouds of points show the observed data, with the color representing the ant state.
@@
 
-The results of all other vital rates as well as the posterior predictive checks are all included in the Apendix \ref{appendix:A} : Figures \ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
+The results of all other vital rates as well as the posterior predictive checks are all included in the Appendix A : Figures S\ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/transition.png}
-	\caption{This figure shows the probability of being tended by each ant partner or vacant based on the size of the plant.
+	\caption{The probability of being tended by each ant partner or vacant based on the size of the plant.
 Each panel shows these probabilities for a different previous ant state.
@@
 	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_lines.png}
-	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+	\caption{Panel a) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
 The lines show the posterior distribution spread of the estimated $\lambda$ values.
-Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+Panels b-c) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
 The lines show the posterior distribution spread of estimated $\lambda$ values.
@@
 The effect of all ant partners can be measured as $\lambda^{3}_{S} - \lambda^{0}_{S}$.
-We have high confidence that this contrast is positive, and equally so for the synchronous and non-synchronous scenarios \tom{(Figure \ref{fig:Portfolio}).}{Now that the lambda figure shows synchronized and non-synchronized, I am not sure you need the histogram figure for portfolio effect.
-Maybe appendix.}
-
-\begin{figure}
-	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
-	\caption{This figure shows the distribution of $\lambda_{NS,All}-\lambda_{NS,Vacant}$ in pink and $\lambda_{S,All}-\lambda_{S,Vacant}$ in green.
-The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
-	\label{fig:Portfolio}
-\end{figure}
+We have high confidence that this contrast is positive, and equally so for the synchronous and non-synchronous scenarios (Figure S\ref{fig:Portfolio}).
 
@@
 Here we show that while alternative partners may be ecologically different, their fitness effects on a shared mutualist can be effectively equivalent and interchangeable.
-The results of our heirarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
-opuniae} tended plants \tom{experiencing slight advantages in growth}{I don't think this is true if you look at my v1 figure.} and survival when small, and \textit{L.
+The results of our hierarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
+opuntiae} tended plants experiencing slight advantages in growth and survival when small, and \textit{L.
 apiculatum}-tended plants experiencing floral viability advantages.
@@
 % Connect to broader patterns in the lit
-Previous studies have reported complementarity \citep{Palmer2010,Rezende2007,Afkhami2021} while we found that there are no benefits of diversity in our system.
-Vital rate results indicated that different ant partners affected different vital rates uniquely (Figures \ref{fig:Grow},\ref{fig:Surv},\ref{fig:Viab}).
-On their own, these results suggested a likely benefit of diversity through complementarity, such that each ant partner offered a unique form of herbivore protection across ontogeny and or across different life history processes.
-However, our multi-state IPM results, which synthesize all life history processes and accouns for partner identity and turnover, revealed an overall positive effect of ant visitation but no apparent benefit of diversity, raising questions about how to resolve the tension between these two sets of results.
+Our conclusion that partner diversity does not benefit this system is novel in a filed where most studies highlight the benefits of diversity.
+Studies on multi-species mutualisms have shown that complementarity is a common mechanism explaining the benefits of diversity \citep{Afkhami2021,Rezende2007,Agrawal1998,Barrett2015,Stachowicz2005,Thibaut2012,Tornos2023}.
+In similar systems (plant-animal multi-species mutualisms) previous studies have reported complementarity \citep{Palmer2010,Frederickson2005,rogers2014} and portfolio effect \citep{Alarcon2008,Lazaro2022}.
+Only one other ant-plant study, to our knowledge, has highlighted a lack of benefit of diversity \citep{Bruna2014}, but they hypothesized these results occurred due to the extremely negative impacts of one partner and did not consider hypothetical scenarios without this partner to determine if there would be a benefit otherwise.
+Our finding is the only which has discovered and published a lack of diversity benefits when all partners are true mutualists.
+Vital rate results indicated that different ant partners affected different vital rates uniquely.
+On their own, these results suggested a likely benefit of diversity through complementarity, such that each ant partner offered a unique form of herbivore protection across ontogeny  and or across different life history processes (as seen in other studies \citep{Dejean2008,Barton2010,Boege2005}).
+However, our multi-state IPM results, which synthesize all life history processes and accounts for partner identity and turnover, revealed an overall positive effect of ant visitation but no apparent benefit of diversity, raising questions about how to resolve the tension between these two sets of results.
 
+% why we see these effects in OUR population
 We speculate that our results are driven by the vital rate sensitivity structure of this population. 
 Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
-Differences between alternative partners were most pronounced either in reproductive rates (Figure \ref{fig:Viab}), which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high (Figure \ref{fig:Surv}) and the probability of ant tending was low (Figure \ref{fig:Ant_Transition}).
-At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitvity of $\lambda$ established plant survival.
+Differences between alternative partners were most pronounced either in reproductive rates, which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high and the probability of ant tending was low.
+At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitivity of $\lambda$ established plant survival.
 The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness.
 
-% why we don't see the vital rate differences in real life
-%Despite the apparent benefits for small \textit{C. opuntiae} tended cacti, these benefits are almost never experienced due to chronic vacancy of small cacti (Figure \ref{fig:Ant_Transition}).
-%We estimated that cacti between the approximate sizes of -$5 - 5 log(cm^3)$($0.01 - 148 cm^3$) would have important advantages in growth and survival if tended by \textit{C. opuntiae} ants.
-%At these sizes plants are very rarely tended ($< 10\%$ occupancy) due to lack of EFN production.
-%This means, while our vital rate models predict benefits for small ants, the assumptions we put in place for ant tending in our IPM means that these benefits are extremely rare and therefore have next to no impact on the fitness of the cacti.
-%This is not to say that growth and survival are not important vital rates, in fact they are likely to be the two most important in the model. 
-%In the literature, growth and survival for established plants are the vital rates that fitness estimates are most sensitive to \cite{Eacker2017}.
-%This indicates that the boosts recieved from ant tending of small plants would not be likely to make a large impact on fitness estimations even if this were allowed to occur in this model.
+% opportunity costs of multipl epartners
+There was modest evidence of opportunity costs imposed by the full diversity scenario.
+Cholla under the full diversity scenario experienced lower $\lambda_S$ than cholla with either one or two possible partners.
+The population fitness would be optimized by including any combination of twp partners rather than all three.
+There is an opportunity cost associated with only the full diversity scenario, indicating there is some mechanism that reduces fitness benefits when all partners are present.
+This could be due an over-saturation of benefits \citep{Stanton2003} or some type of conflict between partners which results in reduced benefits \citep{Afkhami2014}.
+No matter the mechanism behind this reduction of benefits, the optimal diversity scenario for these cholla is any scenario with only two partners present.
 
 % Portfolio Effect
-Our work explcitly incorporated temporal environmental stochasticity, which raises the opportunity for portfoilio effect as a mechanism of diversity benefits. 
+Our work explicitly incorporated temporal environmental stochasticity, which raises the opportunity for portfolio effect as a mechanism of diversity benefits. 
 Yet, we find no evidence of portfolio effect within our system. 
@@
 Partner turnover is likely a significant driver behind the fitness we see within Cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
-In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \cite{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
-The direction of partner turnover is also important when the identity of partners impacts the quality of benefits recieved \cite{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
+In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \citep{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
+The direction of partner turnover is also important when the identity of partners impacts the quality of benefits received \citep{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
 In our system we found that there are distinct patterns to the directions and frequency of partner turnover, but we still don't understand the mechanisms driving these turnovers.
@@
 This simulation was designed to test if the overwhelming frequency of \textit{L. apiculatum} may be overriding any potential advantages of other partners.
-The results of this simulation (Appendix \ref{appendix:C}: Figure \ref{app:EqualLambdaMeans}) were effectively the same as the competitive exclusion model presented in the body of this paper (Figure: \ref{fig:LambdaMeans}).
+The results of this simulation (Figure S\ref{app:EqualLambdaMeans}) were effectively the same as the competitive exclusion model presented in the body of this paper.
 The second alternative we tested was a model that assumed rather than competition driving ant occupations, it was the population size of each ant species alone.
@@
 This simulation was designed to test if the fitness was more sensitive to changes in vacancy across the population than the frequency of any particular ant species.
-The results of this simulation revealed that when vacancy increased sampling effect is at play (Appendix \ref{appendix:C}: Figure \ref{app:FreqLambdaMeans}).
+The results of this simulation revealed that when vacancy increased sampling effect is at play (Figure S\ref{app:FreqLambdaMeans}).
 This result highlights that the primary driver behind the results we presented are ant presence and that it is only in cases oh unrealistically high vacancy that there are benefits to partner diversity.
@@
 This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
-Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \cite{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
-Separate studies have analyzed how inter-annual variability impacts focal mutualists \cite{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
+Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \citep{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
+Separate studies have analyzed how inter-annual variability impacts focal mutualists \citep{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
 The long term data set we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
@@
 We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
-Analysis of our observational herbivore data shows that vacant plants experience elevated levels of herbivory compared to tended ones (Appendix \ref{appendix:B}: Figure \ref{app:herb}).
+Analysis of our observational herbivore data shows that vacant plants experience elevated levels of herbivory compared to tended ones (Figure S\ref{app:herb}).
 In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
@@
 These results highlighted in the paper are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
-However, we have previously conducted experimental manupulations which revealed that ant presence has direct impacts on plant demography in our system \citep{Miller2007}.
+However, we have previously conducted experimental manipulations which revealed that ant presence has direct impacts on plant demography in our system \citep{Miller2007}.
 This combination of observational data backed up by experimental results gives us greater confidence in our causal interpretations than if we had only observational data.
 Specifically, we are confident in claiming that ant presence causes increased performance in cactus vital rates and the fitness signals we are detecting are directly related to ant partner presence.
-One further expansion could include nectar analysis in conjunction with ant interactions to look not only at the ant parnter impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+One further expansion could include nectar analysis in conjunction with ant interactions to look not only at the ant partner impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
 
 % tie back paragraph.
-This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangable in their net effects on fitness of a shared mutualist.
+This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangeable in their net effects on fitness of a shared mutualist.
 The individual estimation of each vital rate and explicit inclusion of ant partners in this IPM allowed us to determine that partner diversity is not beneficial in this system and may highlight circumstances under which we would expect similar results in other systems.
-The individual vital rate estimation approach allowed us to determine that while different ant partners did impact each vital rate differently, the largest differences were estimated to occur at sizes when ants would not be present in real life or in vital rates that do not have a significant impact on the overall fitnes.
-Considering partner imapcts on vital rates across the entire life cycle of the plant allows more power in determining the importance of these differences.
+The individual vital rate estimation approach allowed us to determine that while different ant partners did impact each vital rate differently, the largest differences were estimated to occur at sizes when ants would not be present in real life or in vital rates that do not have a significant impact on the overall fitness.
+Considering partner impacts on vital rates across the entire life cycle of the plant allows more power in determining the importance of these differences.
 Based on our results we would expect to see that partner diversity does not matter when partner impacts occur during low sensitivity life stages.
@@
 
-% I am creating a table here to include all parameter estimations and descriptions
-  \begin{table}[]
-  \begin{tabular}{l|l|l}
-    \textbf{Parameter} & \textbf{Median ($95\%$ CI)} & \textbf{Prior Distribution} \\
-    \hline
-    %% Growth Parameters
-    growth xi intercept vacant $\beta_{01}^g$ & $-5.210899 (-5.686865, -5.491787)$ & sDE\\
-    growth xi intercept other $\beta_{02}^g$ & $-5.8288 (-5.956217, 1.766021) $&asdf \\
-    growth xi intercept \textit{C. opuntiae} $\beta_{03}^g$ & $-4.529523 (-6.0770390, 0.1222112)$ & asdf\\
-    growth xi intercept \textit{L. apiculatum} $\beta_{04}^g$ & $-5.106802 (-5.4499944, 0.5453901)$ & asdf\\
-    growth xi size dependent vacant $\beta_{11}^g$ & asdf&asdf \\
-    growth xi size dependent other $\beta_{12}^g$ & asdf&asdf \\
-    growth xi size dependent \textit{C. opuntiae} $\beta_{13}^g$ & asdf&asdf \\
-    growth xi size dependent \textit{L. apiculatum} $\beta_{14}^g$ &sadf &asdf \\
-    growth omega intercept $\omega_0^g$ & & \\
-    growth omega size dependent $\omega_1^g$ & & \\
-    growth alpha intercept $\alpha_0^g$ & & \\
-    growth alpha size dependent $\alpha_1^g$ & & \\
-    \hline
-    %% Germination Parameters
-    1-year germination intercept $\alpha^{\gamma_1}$ & & \\
-    2-year germination intercept $\alpha^{\gamma_2}$ & & \\
-    \hline
-    %% Survival Parameters
-    survival intercept vacant $\beta_{01}^s$ & & \\
-    survival intercept other $\beta_{02}^s$ & & \\
-    survival intercept \textit{C.opuntiae} $\beta_{03}^s$ & & \\
-    survival intercept \textit{L. apiculatum} $\beta_{04}^s$ & & \\
-    survival size dependent vacant $\beta_{11}^s$ & & \\
-    survival size dependent other $\beta_{12}^s$ & & \\
-    survival size dependent \textit{C. opuntiae} $\beta_{13}^s$ & & \\
-    survival size dependent \textit{L. apiculatum} $\beta_{14}^s$ & & \\
-    \hline
-    %% Probability of Flowering Parameters
-    flowering intercept $\beta_0^f$ & & \\
-    flowering size dependent $\beta_1^f$ & & \\
-    \hline
-    %% Floral Viability Parameters
-    viability intercept vacant $\beta_01^v$ & & \\
-    viability intercept other $\beta02^v$ & & \\
-    viability intercept \textit{C. opuntiae} $\beta_03^v$ & & \\
-    viability intercept \textit{L. apiculatum} $\beta_04^v$ & &
-
-  \end{tabular}
-  \caption{This table includes the median estimates, the 95$\%$ confidence intervals, and the prior distribution for each parameter in each model.}
-  \label{tab:Params}
-  \end{table}
-
-\section*{Figure legends}
-
 \appendix
@@
 	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
-	\caption{Shows the estimated posterior distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
+	\caption{The estimated posterior distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
 	\label{app:AppA_Seeds_Per_Fruit}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/rec_size.png}
-	\caption{Shows the estimated posterior distribution of recruit sizes as the $log(volume)$ of cholla. }
+	\caption{The estimated posterior distribution of recruit sizes as the $log(volume)$ of cholla. }
 	\label{app:AppA_Recruit_Dist}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/germination.png}
-	\caption{Shows the estimated posterior distributions for germination probability of seeds in the first year seedbank (blue) and seeds in the second year seedbank (green).}
+	\caption{The estimated posterior distributions for germination probability of seeds in the first year seedbank (blue) and seeds in the second year seedbank (green).}
 	\label{app:AppA_Germ}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/seed_surv.png}
-	\caption{Shows the distribution of the probability of plants which germinated between censuses surviving to the next census to be counted.
+	\caption{The distribution of the probability of plants which germinated between censuses surviving to the next census to be counted.
 The green line shows the estimated posterior distribution of pre-census survival.}
 	\label{app:AppA_Pre_Census_Surv}
+\end{figure}
+
+\paragraph{Portfolio Effect}
+\begin{figure}
+	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
+	\caption{The distribution of $\lambda_{NS,All}-\lambda_{NS,Vacant}$ in pink and $\lambda_{S,All}-\lambda_{S,Vacant}$ in green.
+The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
+	\label{fig:Portfolio}
 \end{figure}
@@
 	\includegraphics[width=0.91\linewidth]{Figures/herb_all.png}
-	\caption{Shows the proportion of reproducing plants which are visited by herbivores. Each bar represents the subset of the cacti population in a different ant state.  }
+	\caption{The proportion of reproducing plants which are visited by herbivores. Each bar represents the subset of the cacti population in a different ant state.  }
 	\label{app:herb}
@@
 When only  \textit{C. opuntiae}, Other ants, or both ants are present, there is very little difference in the fitness of the cacti from when no partners are present. 
-Only when \textit{L. apiculatum} ants are present do we see an increase in the fitness of the focal mutualist (Figure \ref{fig:FreqLambdaMeans}a).
+Only when \textit{L. apiculatum} ants are present do we see an increase in the fitness of the focal mutualist (Figure \ref{app:FreqLambdaMeans}a).
 In this simulation, the more partners that are present the higher the fitness of the focal mutualist is, confirming that partner diversity would be beneficial through sampling effect if this transition model were correct.
-(Figure \ref{fig:FreqLambdaMeans}b).
+(Figure \ref{app:FreqLambdaMeans}b).
 
@@
 \includegraphics[width=0.91\linewidth]{Figures/Lambdas_Freq_lines.png}
-	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+	\caption{Panel a) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
 The lines show the posterior distribution spread of the estimated $\lambda$ values.
-Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+Panels b-c) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
 The lines show the posterior distribution spread of estimated $\lambda$ values.
@@
 \includegraphics[width=0.91\linewidth]{Figures/Lambdas_equal_lines.png}
-	\caption{Panel a) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
+	\caption{Panel a) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
 The lines show the posterior distribution spread of the estimated $\lambda$ values.
-Panels b-c) shows the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
+Panels b-c) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
 The lines show the posterior distribution spread of estimated $\lambda$ values.
@@
 \section*{Appendix D: Posterior Checks and Model Validation}\label{appendix:D}
-For each model fitted, we conducted two tests to determing if the fit was acceptable to use in our IPM. 
+For each model fitted, we conducted two tests to determine if the fit was acceptable to use in our IPM. 
 First, we checked the convergence of each parameter.
@@
 We show these posterior checks below, split by ant partner where relevant.
+Each model was run with 10,000 total iterations per chain, 5,000 warmups, and thinning set to 2, leaving us with a total of 2,500 iterations post warmup.
 %% Growth Figure
@@
 \includegraphics[width = 0.45\linewidth]{Figures/seed_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/seed_ant_post.png}
+\includegraphics[width=0.45\linewidth]{Figures/seed_post.png}
 \caption{The a) posterior convergence of the parameters estimated by the seeds per fruit model and the b) posterior distributions of seeds per fruit estimates (pink lines) for each ant species (1 = \textit{C.

### 2024-10-12 05_cholla_ant_IPM_figures_tables.R

@@ -3,6 +3,7 @@
 ################################################################################
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
 source( "/Users/alicampbell/Documents/GitHub/ant_cactus_demography/03_cholla_ant_IPM_params_functions.R")
+bayesplot::color_scheme_set(scheme = "pink")
 
 ## Color Codes
 ## Retro bright
@@ -23,7 +24,7 @@ cols <- c(vcol, ccol, lcol, ocol, lccol, locol, cocol, acol)
 str(cactus)
 ##### Size variable used in most visualizations
 size_dummy <- seq(min(cactus$logsize_t, na.rm = T), max(cactus$logsize_t, na.rm = TRUE), by = 0.1)
-
+scenario = c("none","cremvac","liomvac","othervac","liomcremvac","liomvacother","othercremvac","all")
 ################################################################################
 ################################################################################
 ####
@@ -42,7 +43,7 @@ bayesplot::mcmc_trace(fit_grow_skew,pars=c("d_0",
                                            "a_0",
                                            "a_size",
                                            "beta0[1]","beta0[2]","beta0[3]","beta0[4]",
-                                           "beta1[1]","beta1[2]","beta1[3]","beta1[4]"))
+                                           "beta1[1]","beta1[2]","beta1[3]","beta1[4]"))+labs(x = "iterations", y = "estimated value", title = "a)")
 dev.off()
 # # Check the different quantile fits of the model to make sure not only the mean but also other quantiles fit well with the real data
 # # real data moments
@@ -58,7 +59,7 @@ dev.off()
 # obs_sd<-Q.sd(q.fit[3,],q.fit[5,])
 # obs_skew<-Q.skewness(q.fit[2,],q.fit[4,],q.fit[6,])
 # obs_kurt<-Q.kurtosis(q.fit[1,],q.fit[3,],q.fit[5,],q.fit[7,])
[…]

### 2024-11-10 Manuscript/ant_cactus_AmNat.tex
Letter
Manuscript Updates

@@
 % 200 words max
- The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
+We expect  having multiple mutualistic partners to be beneficial for a species.
+These benefits be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
  To study multi-species demographic effects, we considered \textit{Cylindropuntia imbricata} (tree cholla), which produce extrafloral nectar and various ant partners ( \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) provide defense from herbivores and seed predators.
 We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-We constructed an Integral Projection Model (IPM) in which we simulate different combinations of ant partners that don’t occur in nature.
+We constructed an Integral Projection Model (IPM) in which we simulate different combinations of ant partners (both those that do and don't occur in nature).
 Different ant partners had different impacts on the cholla vital rates. 
@@
 They are widespread species interactions but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022,Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010}.
-Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
-
-Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}.
+Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by species' identities \citep{Noe1994,Leigh2010}.
+Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
 Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
 
@@
 Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
+When different partners have different quality or types of benefits, the order in which interactions takes place is extremely important.
 
@@
 For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of a diverse guild of defensive mutualists are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
+The length and order of interactions are just as important as the identities of the interacting partners.
+
+This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
+There are two common ant partners and several additional rarer species, all of which collect EFN in exchange for defense against herbivores.
+These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
+Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
+No other study has integrated analyses of the demographic effects of partners on entire life cycle with inter-annual stochastic population levels to understand multiple possible mechanistic dynamics of a system.
+Our novel approach allows for a deeper understanding of the mechanisms at play and the ability to pinpoint our results to specific life cycle stages.
 
@@
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
-
-This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
-Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \citep{Miller2009}, and that ant defense reduces herbivore damage \citep{Miller2007}. 
-Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
-These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
-Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{L.
-apiculatum} providing strong anti-herbivore defense and \textit{C.
-opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, all of our previous studies in this system have focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects of a diverse partner guild \citep[e.g.,][]{Palmer2010}.
-To our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
 
@@
 We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable and aborted. 
-We recorded the ant species present (or vacancy if no ants present).
+We recorded the ant species presence and absence (or vacancy if no ants present).
 Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
-In addition to missing the year 2020, there are gaps in the time series where were switched plots or plants (and thus broke up transition years for growth and survival) or where we did not distinguish between viable and aborted flower buds (Table \ref{}).
+In total, the data set includes a total of 9,787 observations of 1,141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
+In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth and survival) or where we did not distinguish between viable and aborted flower buds (Table \ref{}).
 
@@
 The ``Other'' category was made up of \textit{Forelius pruinosus} (3.5\% of observations), unidentified species belonging to the genera \textit{Camponotus} (0.9\%), \textit{Aphaenogaster} (0.4\%), \textit{Myrmecocystus} (0.08\%), \textit{Tetramorium} (0.02\%), \textit{Brachymyrmex} (0.02\%), and additional ants not identified to genus or species (2.8\%).
+We initially tried to split this category up further, defining other ant states, but found that error estimates were too high given low frequencies.
 
@@
 \begin{linenomath*}
+%	\begin{equation}
 	$$
-	B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
+	\overbrace{	B^1_{t+1}}^\text{First Year Seedbank}
+ =  \delta \sum_{a=1}^{A} \int_L^U 
+ \overbrace{ \kappa(a)}^\text{\# seeds per fruit}
+ \overbrace{P(x;\pmb{\tau}^P) }^\text{prob. of flowering}
+ \overbrace{F(x;\pmb{\tau}^F)}^\text{\# flowerbuds}
+ \overbrace{ V(a;\pmb{\tau}^V_{a})}^\text{flower viability}
+ \overbrace{ n(x,a)_{t} }^\text{\# size x plants} dx \\
+	$$
+
+	$$
+		\overbrace{	B^1_{t+1}}^\text{First Year Seedbank (yr t+1)}
+	=  \overbrace{ \kappa(a) P(x;\pmb{\tau}^P)F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a})  n(x,a)_{t}}^\text{\# of viable seeds produced by plants of size x (yr t)}
+	$$
+	$$
+	\overbrace{B^2_{t+1} }^\text{Second Year Seedbank (yr t+1)}
+	=
+	\overbrace{(1 - \gamma_1)B^1_{t}}^\text{Seeds which did not germinate in the first year in the seedbank}
 	$$
@@
 	\label{eqn:IPM1}
+	%\end{equation}
 \end{linenomath*}
@@
 \begin{linenomath*} 
+	%	\begin{equation} 		\label{eqn:IPM2}
 			$$
-	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + \\
+				\overbrace{n(x',a')_{t+1}}^\text{Plants of size x' and ant state a'}
+				= 
+				\overbrace{ (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a') }^\text{new plants coming from the seedbanks}
+				+
 			$$
+
 			$$
-	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
+				\overbrace{\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx}^\text{Already existing plants growing}
 			$$
-	\label{eqn:IPM2}
+	%	\end{equation}
 \end{linenomath*}
@@
 \noindent The first term in Eq.
-\tom{\ref{eqn:IPM2}}{We should label s.
-I am not sure why the equation label is not working here and I did not try to figure it out.
-It is probably something with the linenomath formatting.} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
+\ref{eqn:IPM2} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
 This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ ($\sum\rho_{0}(a')=1$). 
@@
 Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
-We assume growth variance and skewness were not dependent on ant occupancy state.
+After several rounds of model selection, we chose to model  growth variance and skewness as not dependent on ant occupancy state.
 
@@
 \begin{figure}
-	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour_v2.png}
+	\includegraphics[width = 0.75\linewidth]{Figures/grow_contour_v2.png}
 	\caption{The next predicted size of cholla based on previous size with each individual ant partner.
 The colored lines (seen in all panels) are the next mean predicted size of cholla.
-Dashed lines indicate extrapolation while solid lines indicate estimations within the range of observed data.
 The points (seen in panels a-d) are the observed data which informs these estimates.
@@
 Panel e shows all mean estimates from panels a-d on one graph.
+Dashed colored lines indicate extrapolation while solid lines indicate estimations within the range of observed data.
+The vacant (yellow) line is solid across the entire range because we have observed vacant plants at all sizes.
+The other lines must be extrapolated to cover the entire size range of the plants as we do not observe ants on small plants often.
 The gray dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).}
@@
 \begin{figure}[H]
-\includegraphics[width=0.95\linewidth]{Figures/survival_plot.png}
+\includegraphics[width=0.75\linewidth]{Figures/survival_plot.png}
 \caption{The estimated survival rates based on the size of the cactus with each individual ant partner.
@@
 
+We found that the annually varying effects of ants on vital rates differed.
 We analyzed the correlation coefficients of all models that included ant state as a predictor and found that, across vital rates, the annually varying effects on growth were the most correlated between ant states (mean pairwise correlation: 0.63) and effects on survival were the least correlated (0.36) (Figure \ref{fig:Annual_Ant}).
@@
 Here we show that while alternative partners may be ecologically different, their fitness effects on a shared mutualist can be effectively equivalent and interchangeable.
-The results of our hierarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
+The results of our hierarchical models revealed that different ant partners had subtly different effects on vital rates, with \textit{C.
 opuntiae} tended plants experiencing slight advantages in growth and survival when small, and \textit{L.
 apiculatum}-tended plants experiencing floral viability advantages.
-The results of our synchronous and non-synchronous stochastic IPM revealed that all diversity scenarios which included any partners resulted in the highest possible fitness for tree cholla, suggesting that while ant visitation is beneficial, partner identity and diversity are inconsequential in this system.
-These results highlight that partners can differ meaningfully while still resulting in the same fitness benefits for a focal mutualist.
+The results of our synchronous and non-synchronous stochastic IPM revealed that all scenarios which included any partners resulted in the highest possible fitness for tree cholla, suggesting that while ant visitation is beneficial, partner identity and diversity are inconsequential in this system.
+These results highlight that partners can differ while still resulting in the same fitness benefits for a focal mutualist.
 
-% Connect to broader patterns in the lit
-Our conclusion that partner diversity does not benefit this system is novel in a filed where most studies highlight the benefits of diversity.
-Studies on multi-species mutualisms have shown that complementarity is a common mechanism explaining the benefits of diversity \citep{Afkhami2021,Rezende2007,Agrawal1998,Barrett2015,Stachowicz2005,Thibaut2012,Tornos2023}.
-In similar systems (plant-animal multi-species mutualisms) previous studies have reported complementarity \citep{Palmer2010,Frederickson2005,rogers2014} and portfolio effect \citep{Alarcon2008,Lazaro2022}.
-Only one other ant-plant study, to our knowledge, has highlighted a lack of benefit of diversity \citep{Bruna2014}, but they hypothesized these results occurred due to the extremely negative impacts of one partner and did not consider hypothetical scenarios without this partner to determine if there would be a benefit otherwise.
-Our finding is the only which has discovered and published a lack of diversity benefits when all partners are true mutualists.
-Vital rate results indicated that different ant partners affected different vital rates uniquely.
-On their own, these results suggested a likely benefit of diversity through complementarity, such that each ant partner offered a unique form of herbivore protection across ontogeny  and or across different life history processes (as seen in other studies \citep{Dejean2008,Barton2010,Boege2005}).
-However, our multi-state IPM results, which synthesize all life history processes and accounts for partner identity and turnover, revealed an overall positive effect of ant visitation but no apparent benefit of diversity, raising questions about how to resolve the tension between these two sets of results.
-
-% why we see these effects in OUR population
-We speculate that our results are driven by the vital rate sensitivity structure of this population. 
-Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
-Differences between alternative partners were most pronounced either in reproductive rates, which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high and the probability of ant tending was low.
-At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitivity of $\lambda$ established plant survival.
-The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness.
+% conection to literature
+The field of multi-species mutualisms is rich with many studies highlighting unique results.
+Very few studies have quantified the effects of partner diversity on focal mutualist fitness.
+% Benefits of diversity
+The vast majority of multi-species mutualism studies have found that partner diversity is beneficial through complementarity \citep{Yang2024,Hernandez2020,Larimer2014,Afkhami2016,Fehling2022,Gustafson2005,McKeon2012,Palmer2010,Afkhami2021,Stachowicz2005} or portfolio effect \citep{Stevens2024,Thibaut2012,rogers2014,Lazaro2022}.
+% Costs of Diversity
+Three studies have reported costs of diversity.
+Keller et al. \citeyear{Keller2018} found that having multiple partners resulted in equivalent fitness to no partners and worse fitness than having a single partner.
+Barrett et al. \citeyear{Barrett2015} found that having multiple partners is worse than having a single partner, but better than having none.
+Bruna et al. \citeyear{Bruna2014} found that having multiple partners resulted in worse fitness than the best partner but had no vacancy scenario to compare.
+Our study revealed modest evidence of opportunity costs imposed by the full diversity scenario.
+The combination of a slight cost in a full diversity scenario compared to partial diversity scenarios is not unique (as addressed above), but the finding that the full diversity scenario is still better than vacancy does appear to be a unique finding.
+% Neutrality
+One study has previously reported a neutral effect of diversity \citep{Ford2015}, finding that vacant plants have the same fitness as plants guarded by any combination of partners.
+% Individual vital rates
+Most of the studies which quantitatively compare the fitness of individual partnered hosts with multi-partnered hosts used individual vital rates to estimate a fitness of the hosts.
+These measures ranged from growth rates, survival, biomass, to photosynthetic rates.
+Very few studies within the literature have used integrative fitness estimate approaches which consider the impacts of partners across the entire life cycle \citep{Palmer2010,Stevens2024,Ford2015,Thibaut2012,rogers2014,Lazaro2022}.
+% Sequential Partners
+Across these papers, only three considered sequential partnering, where diversity of interactions accumulate across time \citep{Palmer2010,Ford2015,Bruna2014}.
+There are not enough studies calculating and comparing the fitness of hosts with individual partners to host fitness with multiple partners in sequential systems.
+% Position in lit
+Not only are our results entirely unique in terms of benefits and neutrality, but our study adds to the literature studying sequential multi-species mutualisms using integrative methods considering the entire life cycle.
+No other study has reported that there are clear benefits from interacting with any partner, combination with little importance tied to partner identity is unique.
+The finding that the optimal scenario includes multiple partners, but there is a slight cost to the full diversity scenario is also unique. \ali{}{this may not be the best way to describe this}
+Our study offers contrasting results to other studies in sequentially partnered systems, indicating that we need more studies to quantitatively determine the fitness effects of partner diversity in sequentially partnered systems.
 
 % opportunity costs of multipl epartners
-There was modest evidence of opportunity costs imposed by the full diversity scenario.
-Cholla under the full diversity scenario experienced lower $\lambda_S$ than cholla with either one or two possible partners.
-The population fitness would be optimized by including any combination of twp partners rather than all three.
-There is an opportunity cost associated with only the full diversity scenario, indicating there is some mechanism that reduces fitness benefits when all partners are present.
-This could be due an over-saturation of benefits \citep{Stanton2003} or some type of conflict between partners which results in reduced benefits \citep{Afkhami2014}.
-No matter the mechanism behind this reduction of benefits, the optimal diversity scenario for these cholla is any scenario with only two partners present.
-
-% Portfolio Effect
-Our work explicitly incorporated temporal environmental stochasticity, which raises the opportunity for portfolio effect as a mechanism of diversity benefits. 
-Yet, we find no evidence of portfolio effect within our system. 
-When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
-We did not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure how the effects of each partner on cholla demography fluctuate through time, and whether those fluctuations are correlated.
-The vital rate results led us to believe that portfolio effect may be at play due to the differences in ant responses to annual variation.
-However, there is no stabilization of fitness across temporal heterogeneity and therefore no benefits of diversity due to the portfolio effect.
-Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found that asynchrony was linked to weak evidence of portfolio effect.
-One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
-This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature.
 
 % These results are robust to different assumptions about turnover patterns & what we don't know about turnover
+Due to the sequential nature of our system, partner turnover is an important aspect of the dynamics we observed.
 Partner turnover is likely a significant driver behind the fitness we see within Cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
@@
 
+% why we see these effects in OUR population
+We speculate that the reason our results are mote sensitive to changes in vacancy than any individual partner is due to the vital rate sensitivity structure of this population. 
+Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
+Differences between alternative partners were most pronounced either in reproductive rates, which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high and the probability of ant tending was low.
+At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitivity of $\lambda$ established plant survival.
+The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness, as seen in other studies \citep{Ford2015}.
+
+% Portfolio Effect
+Our work explicitly incorporated temporal environmental stochasticity, which raises the opportunity for portfolio effect as a mechanism of diversity benefits. 
+Yet, we find no evidence of portfolio effect within our system. 
+Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti \citep{Miller2009} (Figure \ref{app:herb}).
+We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
+As herbivore populations fluctuate across time, the vital rate level benefits offered by different ant partners also fluctuate across time (Figure \ref{fig:Annual_Ant}).
+When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
+We did not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure how the effects of each partner on cholla demography fluctuate through time, and whether those fluctuations are correlated.
+The vital rate results led us to believe that portfolio effect may be at play due to the differences in ant responses to annual variation.
+However, there is no stabilization of fitness across temporal heterogeneity and therefore no benefits of diversity due to the portfolio effect in our system.
+Other portfolio effect studies \citep{Lazaro2022,Tornos2023,Thibaut2012,rogers2014} found that asynchrony was linked to weak evidence of portfolio effect.
+One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
+This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature. \ali{}{not sure if I need a better transition here}
+
 % Value of Long-term data
@@
 apiculatum}.
-
-% Herbivory
-Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti \citep{Miller2009}.
-We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
-Analysis of our observational herbivore data shows that vacant plants experience elevated levels of herbivory compared to tended ones (Figure S\ref{app:herb}).
-In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
 
@@
 Despite very different proportions, we found very similar outcomes to the competitive exclusion model analyzed in the paper. 
-All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{fig:EqualLambdaMeans}b).
-Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{fig:EqualLambdaMeans}a).
+All ants are beneficial, but having more than one is not necessarily any better than having an individual species as a partner (Figure \ref{app:EqualLambdaMeans}b).
+Partner presence is beneficial, but neither identity nor number of partners appears to be important (Figure \ref{app:EqualLambdaMeans}a).
 

### 2024-11-10 05_cholla_ant_IPM_figures_tables.R

@@ -630,7 +630,8 @@ n_draws = 1000
 flow_sim <- matrix(NA, n_draws,stan_data_flow_trunc$N)
 for(i in 1:n_draws){
   for(n in 1:stan_data_flow_trunc$N){
-    flow_sim[i,n] <- sample(x=1:n_draws,size=1,replace=T,prob=dnbinom(1:n_draws, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i]) / (1 - dnbinom(0, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i])))
+    flow_sim[i,n] <- sample(x=1:n_draws,size=1,replace=T,
+                            prob=dnbinom(1:n_draws, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i]) / (1 - dnbinom(0, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i])))
   }
 }
 ## Plot the posterior distributions

### 2024-11-11 Manuscript/ant_cactus_AmNat.tex
figure updates

@@
 We chose to use the vacancy estimates for plants tended by other ants, a decision which does not have a significant impact on the final results. 
-Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A \ref{appendix:A}.
+Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A.
 

### 2024-11-11 02_cholla_ant_IPM_vital_rates.R

@@ -8,7 +8,7 @@
 ## Read the data in
 setwd("/Users/alicampbell/Documents/GitHub/ant_cactus_demography")
 source("01_cholla_ant_IPM_setup.R")
-cactus <- read.csv("Data Analysis/Data/cholla_demography_20042023_cleaned.csv", header = TRUE,stringsAsFactors=T)
+cactus <- read.csv("Data/cholla_demography_20042023_cleaned.csv", header = TRUE,stringsAsFactors=T)
 ################################################################################
 ## Growth Model Selection with WAIC
 ################################################################################
@@ -335,8 +335,8 @@ seed$ant_state <- NA
 seed$ant_state[seed$ant.access=="n"]<-"Vacant"
 seed$ant_state[seed$ant.access=="y" & seed$species=="c"]<-"Crem"
 seed$ant_state[seed$ant.access=="y" & seed$species=="l"]<-"Liom"
-seed_data <- seed %>% select(plant,ant_state,seed_count) %>% drop_na()
-
+seed_data <- seed[,c("plant","ant_state","seed_count")]
+seed_data <- drop_na(seed_data)
 ## Create Stan Data
 stan_data_seed <- list(N = nrow(seed_data),                            ## number of observations
                        N_plants = length(unique(seed_data$plant)),
@@ -453,7 +453,9 @@ stan_data_fruit <- list(N = nrow(fruit_data),                                   
 # these tag IDs are the only ones that start with H
 seed_add<-cactus %>% filter(substr(TagID,1,1)=="H") %>% drop_na()
 #2. the main census, where new plants get called recruits or not by observers
-recruits<-cactus %>% filter(Recruit==1) %>% select(logsize_t1) %>% drop_na()
+recruits <- cactus[,c("Recruit","logsize_t1")]
+recruits <- subset(recruits,recruits$Recruit == 1)
+recruits <- drop_na(recruits)
 seedling.dat <- c(recruits$logsize_t1,seed_add$logsize_t)
 
 ## Create Stan Data

### 2024-11-11 05_cholla_ant_IPM_figures_tables.R

@@ -101,14 +101,14 @@ dev.off()
 
 ## Format the original data
 y_subset <- growth_data[,c("logsize_t1","ant_t", "logsize_t")]
-y_crem_subset_surv <- subset(survival_data, ant_t == "crem")
-y_liom_subset_surv <- subset(survival_data, ant_t == "liom")
-y_vac_subset_surv <- subset(survival_data, ant_t == "vacant")
-y_other_subset_surv <- subset(survival_data, ant_t == "other")
-size_crem = seq(min((y_crem_subset_surv$logsize_t), na.rm = TRUE), max ((y_crem_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
-size_other = seq(min((y_other_subset_surv$logsize_t), na.rm = TRUE), max ((y_other_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
-size_liom = seq(min((y_liom_subset_surv$logsize_t), na.rm = TRUE), max ((y_liom_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
-size_vac = seq(min((y_vac_subset_surv$logsize_t), na.rm = TRUE), max ((y_vac_subset_surv$logsize_t), na.rm = TRUE), by = 0.1)
+y_crem_subset_grow <- subset(growth_data, ant_t == "crem")
+y_liom_subset_grow <- subset(growth_data, ant_t == "liom")
+y_vac_subset_grow <- subset(growth_data, ant_t == "vacant")
+y_other_subset_grow <- subset(growth_data, ant_t == "other")
+size_crem = seq(min((y_crem_subset_grow$logsize_t), na.rm = TRUE), max ((y_crem_subset_grow$logsize_t), na.rm = TRUE), by = 0.1)
+size_other = seq(min((y_other_subset_grow$logsize_t), na.rm = TRUE), max ((y_other_subset_grow$logsize_t), na.rm = TRUE), by = 0.1)
+size_liom = seq(min((y_liom_subset_grow$logsize_t), na.rm = TRUE), max ((y_liom_subset_grow$logsize_t), na.rm = TRUE), by = 0.1)
+size_vac = seq(min((y_vac_subset_grow$logsize_t), na.rm = TRUE), max ((y_vac_subset_grow$logsize_t), na.rm = TRUE), by = 0.1)
 ## Predicted sizes for each partner condition
[…]

### 2024-12-05 Manuscript/ant_cactus_AmNat.tex
tom's edits - mostly discussion and abstract

--- 2024-10-12 Manuscript/ant_cactus_AmNat.tex
@@
 % 200 words max
- The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms: portfolio effect, complementarity, and sampling effect.
- To study multi-species demographic effects, we considered \textit{Cylindropuntia imbricata} (tree cholla), which produce extrafloral nectar and various ant partners ( \textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) provide defense from herbivores and seed predators.
-We used plant demographic censuses to parameterize a series of Bayesian hierarchical generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. 
-We constructed an Integral Projection Model (IPM) in which we simulate different combinations of ant partners that don’t occur in nature.
-Different ant partners had different impacts on the cholla vital rates. 
-Specifically, \textit{C. opuntiae} tended plants had advantages in both growth and survival when small, and large \textit{L. apiculatum} tended plants had floral viability advantages. 
-The IPM results revealed that despite these differences in vital rates, the presence of any ant partner is beneficial, the identity and number of partners do not matter.
-This suggests that there are no benefits of diversity in this system and therefore than none of the highlighted mechanisms explain the benefits of multi-species mutualisms.
-This study highlights that partner diversity does not always increase the benefits a focal mutualist receives.
+Mutualisms commonly involve multiple partner species but the ecological consequences of partner diversity remain poorly understood.
+Partner diversity may benefit a focal mutualist through mechanisms that mirror positive effects of species diversity on ecosystem function, such as complementarity or portfolio effect. 
+We studied the fitness consequences of partner diversity in the multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
+We used long-term data (2004--2023) and Bayesian hierarchical statistical models to determine the demographic impacts of different ant partners, and the dynamics of turnover from one exclusive partner association to another.
+We then constructed a stochastic, multi-state Integral Projection Model (IPM) with which we simulated tree cholla populations with different richness and composition of partners.
+We found that, while ant partners had different impacts on plant vital rates (\textit{C.
+opuntiae}-tended plants had advantages in growth and survival when small, and \textit{L.
+apiculatum}-tended plants had floral viability advantages), plant fitness was effectively insensitive to partner identity and number, because effects on the highest-sensitivity vital rates were consistent across partner species.
+Furthermore, there was no evidence that benefits of partner diversity manifest in the context of fluctuating environments (i.e., the portfolio effect). 
+This study highlights that demographic benefits of mutualism can be surprisingly robust to diversity and turnover of partners.
 
@@
 \section*{Introduction}
-Cooperations between species, aka mutualisms, occur on a pairwise basis or in a more complex network and often play integral and foundational roles in ecological communities.
 Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
@@
 This can lead to a positive effect of partner diversity through complementarity of alternative functions \citep{Batstone2018}. 
-Interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}. 
+Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}.
 Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
@@
 
-Partner diversity can have different effects depending on whether partners are present simultaneously or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
+Partner diversity can have different effects on a focal mutualist depending on whether partners are present simultaneously or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
 Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \citep{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
@@
 For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of a diverse guild of defensive mutualists are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
+Thus, the duration and order of interactions may be just as important as the identities of the interacting partners.
 
@@
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
-Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, consequently causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
 Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}
-However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the necessary long-term data are rarely available.
+Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}. 
+However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
 
 This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
-Previous studies have shown that herbivory by specialized insect herbivores negatively affects cactus fitness \citep{Miller2009}, and that ant defense reduces herbivore damage \citep{Miller2007}. 
-Tree cholla are tended by two common ant species (\textit{Liometopum apiculatum} and \textit{Crematogaster opuntiae}) and several additional rarer species, all of which collect EFN during foraging visits but their colonies are ground-nesting and not housed by the plants.
+Tree cholla are tended by two common ant partners and several additional rarer species, all of which collect EFN in exchange for defense against herbivores.
 These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
@@
 opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, all of our previous studies in this system have focused on single life stages (adult plants) or vital rates (seed production) and did not integrate the demographic effects of ant defense across the life cycle, which may be essential for understanding net fitness effects of a diverse partner guild \citep[e.g.,][]{Palmer2010}.
-To our knowledge no previous study has incorporated inter-annual stochasticity into models of ant-plant dynamics, which limits our understanding of diversity benefits that may arise through the portfolio effect.
+However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production), did not integrate the demographic effects of ant defense across the life cycle, and did not incorporate temporal stochasticity, all of which may be essential for understanding net fitness effects of a diverse partner guild in variable environments.
 
-Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal turnover in partner identity, and how the demographic effects of alternative partners varies across plant size structure and nearly 20 years of inter-annual fluctuations.
+Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover, and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
 We used this observational data set of plant demography and ant-plant associations, contextualized by previous ant exclusion experiments, to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
@@
 
-This study was conducted in the Los Pi$\tilde{n}$os mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research site (SEV-LTER) in central New Mexico, USA.
-This is an area characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B. gracilis}), yuccas, cacti, and junipers. 
-Tree cholla cacti are common in high Chihuahuan desert habitats of the southwestern USA \citep{Benson1982}. 
+This study was conducted at the Sevilleta Long-term Ecological Research site (SEV-LTER), located within the Sevilleta National Wildlife Refuge in central New Mexico, USA.
+Out study area in the Los Pi$\tilde{n}$os mountains \tom{()}{Give lat-long} is characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
+gracilis}), yuccas, cacti, and junipers.
+Tree cholla cacti are common in such high Chihuahuan desert habitats \citep{Benson1982}. 
 These arborescent plants produce cylindrical segments with large spines. 
-In the growing season (May to August in New Mexico), the plants initiate new vegetative segments and flowerbuds at the ends of existing segments. 
+In the growing season (May to August in New Mexico), they initiate new vegetative segments and flowerbuds at the ends of existing segments. 
 While most plants produce new segments every season, only those which are reproductively mature produce flowerbuds. 
@@
 
-Tree cholla EFN is harvested by various ground-nesting ant species. 
+Tree cholla EFN is collected by several ground-nesting ant species. 
 At SEV-LTER, cholla are visited primarily by two species, \textit{Crematogaster opuntiae} and \textit{Liometopum apiculatum}, as well as other rarer species, including \textit{Forelius pruinosus} and unidentified species in the genera \textit{Aphaenogaster} and \textit{Camponotus}.
@@
 Workers of different species rarely co-occur on individual plants, likely due to interspecific competition. 
-For example, staged introductions of \textit{C.
-opuntiae} to \textit{L.
-apiculatum}-tended plants, and vice versa, provoke aggressive responses by residents (A.
-Campbell, \textit{personal observation}).
+For example, staged introductions of \textit{C. opuntiae} to \textit{L. apiculatum}-tended plants, and vice versa, provoke aggressive responses by residents (\textit{personal observation}).
 %Each cholla is visited by a single ant species for the duration of a season, and the species of the visitors can change from one season to the next. 
@@
 \subsection*{Data Collection}
-This study is based on long-term demographic data spanning 2004 to 2023 at SEV-LTER. 
+This study is based on long-term data spanning 2004 to 2023 that documented plant demographic rates in relation to their ant partner status. 
 From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
@@
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable and aborted. 
-We recorded the ant species present (or vacancy if no ants present).
+We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted. 
+We recorded the ant species presence and identity (or vacancy if no ants present).
 Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1141 unique individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
-In addition to missing the year 2020, there are gaps in the time series where were switched plots or plants (and thus broke up transition years for growth and survival) or where we did not distinguish between viable and aborted flower buds (Table \ref{}).
+In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. 
+In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds \tom{(Table \ref{})}{Missing table reference.
+There is no table in the supplement, but there is an empty heading for ``Table''}.
 
-We used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
+To complete the tree cholla life cycle, we used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
 Ohm et al.
@@
 apiculatum} and \textit{C.
-opuntiae} (\textless8\% of observations) we combined all other ants into an ``other'' category, such that our multi-state IPM included four ant states: vacant, \textit{L.
+opuntiae} (\textless8\% of observations) we combined all other ants into an ``other'' category, such that our multi-state IPM included four possible ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
@@
 The ``Other'' category was made up of \textit{Forelius pruinosus} (3.5\% of observations), unidentified species belonging to the genera \textit{Camponotus} (0.9\%), \textit{Aphaenogaster} (0.4\%), \textit{Myrmecocystus} (0.08\%), \textit{Tetramorium} (0.02\%), \textit{Brachymyrmex} (0.02\%), and additional ants not identified to genus or species (2.8\%).
+Given our objective of quantifying demographic effects across the plant size distribution, there was not sufficient size representation of these low-frequency partner species in the long-term data to treat these as separate partner states.
 
-Ant state is included as a predictor variable in IPM sub-models where there are biologically realistic pathways through which ants could impact the outcome of that process. 
-For example, ant partners defend cacti from herbivores, and prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion.
+Ant state is included as a predictor variable in IPM sub-models where there are biologically realistic pathways through which ants could have an impact. 
+For example, prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion, while also potentially reducing pollinator visitation \citep{Miller2009,Ohm2014}.
 Therefore, ant state was included in sub-models for survival, growth, flowerbud viability, and seed number per flowerbud. 
@@
 Seed bank dynamics are given by:
-
 \begin{linenomath*}
-	$$
+	\begin{gather}
 B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
-	$$
-	$$
-	B^2_{t+1} =  (1 - \gamma_1)B^1_{t}\\
-	$$
+B^2_{t+1} =  (1 - \gamma_1)B^1_{t}
 	\label{eqn:IPM1}
+	\end{gather}
 \end{linenomath*}
-
 \noindent %In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
@@
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
-For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability and plants in different ant states can fluctuate independently (good years for \textit{L.
-apiculatum} -occupied plants may not be good years for \textit{C.
+For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability, and plants in different ant states can fluctuate independently (good years for \textit{L.
+apiculatum} -occupied plants may be bad years for \textit{C.
 opuntiae}-occupied plants, for example).
-Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible initial ant states ($A=4$) giving total seed production. 
+Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
 Seeds are multiplied by the probability of escaping post-dispersal seed predation ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
@@
 \begin{linenomath*} 
-	$$
-	n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + \\
-	$$
-	$$
-	\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx \\
-	$$
+	\begin{gather}
 		\label{eqn:IPM2}
+n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + 
+\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx 
+	\end{gather}
 \end{linenomath*}
-
 \noindent The first term in Eq.
-\tom{\ref{eqn:IPM2}}{We should label s.
-I am not sure why the equation label is not working here and I did not try to figure it out.
-It is probably something with the linenomath formatting.} estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
-This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ ($\sum\rho_{0}(a')=1$). 
+(\ref{eqn:IPM2}) estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
+This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ upon its first appearance in the census ($\sum\rho_{0}(a')=1$). 
 The second term represents all possible transitions from size $x$ and ant $a$ to size $x'$ and ant $a'$, conditioned on survival. 
@@
 Finally, ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
-This function includes inter-annual variability through year-specific intercepts which are consistent across initial ant states ($\pmb{\tau}^\rho$).
+\tom{This function includes inter-annual variability through year-specific intercepts which are consistent across initial ant states ($\pmb{\tau}^\rho$).}{Confirming that this model includes random year effects?}
 
@@
 \begin{linenomath*}
-	$$y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) $$
-	$$\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} $$
-	$$log(\sigma_i)  = \beta^3 + \beta^4 x_i $$
-	$$\alpha_i = \beta^5 + \beta^6 x_i$$
+	\begin{gather}
+	y_i^G \sim Skewed Normal(\hat{G_i},\sigma_i,\alpha_i) \\
+	\hat{G_i} = \beta^0_{a[i]} + \beta^1_{a[i]} x_i + \beta^2_{a[i]} x_i^2 + u_{year[i],a[i]} + w_{plot[i]} \\
+	log(\sigma_i)  = \beta^3 + \beta^4 x_i \\
+	\alpha_i = \beta^5 + \beta^6 x_i
 	\label{eqn:growth}
+	\end{gather}
 \end{linenomath*}
@@
 Parameters $\sigma_i$ and $\alpha_i$  control residual variance and skewness, respectively, and were defined as linear functions of initial size $x_i$ ($\sigma_i$ is strictly positive and was modeled with a log link function).
-We assume growth variance and skewness were not dependent on ant occupancy state.
+Due to data constraints, we modeled growth variance and skewness as size-dependent but not dependent on ant occupancy state.
 
@@
 
-The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the probability that flowerbuds produced in year $t$ remain viable (not aborted), with fixed effects of ant partner $a$ in year $t$.
-We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
+The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the probability that flowerbuds produced in year $t$ remain viable, with fixed effects of ant partner $a$ in year $t$.
+We fit this model to floral viability data using a \tom{binomial}{For the inevitable revision, I would like to try beta-binomial.} distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
 This model used a logit link function and included random effects for plot and year following the same structure as the survival model, with ant-specific year random effects.
 
-Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants \citep{Miller2006}.
-This data only included  \textit{L. apiculatum} and \textit{C. opuntiae} ants, so we had to make an assumption about the seeds per fruit on plants tended by Other ants.
-We chose to use the vacancy estimates for plants tended by other ants, a decision which does not have a significant impact on the final results. 
-Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A \ref{appendix:A}.
+Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants from flowering plants \citep{Ohm2014}.
+This experiment only included  vacant or \textit{L.
+apiculatum} and \textit{C.
+opuntiae} -tended plants, so we assumed that plants tended by Other ants had the same number of seeds per fruit as vacant plants (our results showed very little sensitivity to this assumption).
+Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A.
 
@@
 
-\paragraph{Parameter estimation}
-We fit models using Stan run through version 4.0.2 of R \citep{Rcite,Rstancite}. 
+\paragraph{Model-fitting}
+We fit all models using Stan run through version 4.0.2 of R \citep{Rcite,Rstancite}. 
 We used vague priors for all parameters. 
 For each model, we obtained three chains of 10,000 iterations, discarding the first 1,500 iterations. 
-We visually assessed parameter convergence between and within chains (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}a).
+\tom{We visually assessed parameter convergence between and within chains (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}a).}{Not sure these are all necessary in the supplement - also I think these are not labeled correctly here (the a and b).}
 
@@
 We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as 
-$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \citep{Mark2009}.
-We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and other derived quantities.
+$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \tom{\citep{Mark2009}}{Mark is Mark Rees' first name. This bibtex entry needs to be updated.}.
+We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and the derived quantities described next.
 
@@
 
-In all scenarios that included any ant partners, we preserved the observed pattern of size-dependent vacancy/occupancy (estimated through the ant transition sub-model) and manipulated partner identity conditional on occupancy.
+In all scenarios that included any ant partners, we preserved the observed pattern of size-dependent vacancy/occupancy (estimated through the ant transition sub-model) and manipulated partner identity conditional on occupancy by any ant.
 This means, for example, that the \textit{C. opuntiae}-only scenario included two possible states, vacancy and occupied by \textit{C. opuntiae}. 
@@
 The partner diversity treatment scenarios required additional assumptions about the mechanisms that give rise to observed occupancy patterns. 
-Based on evidence that EFN-bearing cacti are nearly always ant-occupied \citep{Miller2014}, we assume that ant partners competitively exclude one another from EFN-bearing cacti and that competition is zero-sum.
+Based on evidence that EFN-bearing cacti are nearly always ant-occupied \citep{Miller2014}, we assume that ant partners competitively exclude one another from EFN-bearing cacti and that competition for plant partners is zero-sum.
 This means that, in scenarios that remove species from the partner community, remaining species gain access to plants that the removed species would have tended. 
@@
 Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation, yielding an advantage of a diverse ``portfolio'' of partners when the environment fluctuates.
-Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
+%Our statistical estimation of ant-specific year random effects in the vital rates allows for this possibility. 
 We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis.
-The baseline, 'non-synchronous' model described above included ants effects that could vary uniquely across time. 
-In this scenario, good years and bad years can differ between ant states, according to the parameter estimates for the random effects ($\pmb{\tau}_{a}$). 
+The baseline, 'non-synchronous' model described above included ants effects that could vary uniquely across time, according to the parameter estimates for the random effects ($\pmb{\tau}_{a}$). 
 We quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states for each vital rate.
-The 'synchronous' (S) version included  ant effects that were forced to be the same across species. 
+The 'synchronous' version included  ant effects that were forced to be the same across ant states. 
 To synchronize ant states, we averaged the ant-specific year random effects, thus ensuring that plants in all ant states fluctuated synchronously in response to temporal environmental variation. 
-This second, ``synchronized'' version of the model effectively turns off any portfolio effect, holding all else equal. 
+This second, synchronous version of the model effectively turns off any portfolio effect, holding all else equal. 
 Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above. 
-We indicate stochastic fitness under the synchronized scenario with $\lambda^{C(s)}_{S}$, for example.
+%We indicate stochastic fitness under the synchronized scenario with $\lambda^{C(s)}_{S}$, for example.
 
 \subsubsection*{Statistical inference on fitness consequences of partner identity and diversity}
-The range of models we created generated many outputs; we focus our inference on the following specific contrasts. 
-First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels (zero, one, two, three), averaging over composition scenarios within each level.
+The range of models we created could generate many outputs; we focus our inference on the following specific contrasts. 
+First, to determine whether ant occupancy and partner diversity are beneficial, we calculated a posterior distribution of $\lambda_{S}$ for each of four partner richness levels ($\lambda^{0}_{S}$, $\lambda^{1}_{S}$, $\lambda^{2}_{S}$, $\lambda^{3}_{S}$), averaging over composition scenarios within each level.
 If cactus fitness increases with partner richness, this would be interpreted as evidence for benefits of partner diversity. 
-Second, to determine whether each partner, in isolation, confers a fitness advantage and to rank alternative partners, we contrasted the fitness of each single partner scenario (\textit{C.
-opuntiae} only, \textit{L.
-apiculatum} only, Other only) against vacancy (zero partners).
+Second, to determine whether each partner, in isolation, confers a fitness advantage and to rank alternative partners, we contrasted the fitness of each single partner scenario ($\lambda^{C}_{S}$, $\lambda^{L}_{S}$, $\lambda^{O}_{S}$) against vacancy ($\lambda^{0}_{S}$).
 Third, to determine whether any benefits of diversity are due to the sampling effect or complementarity, we contrasted the fitness of multi-partner scenarios against the single best partner scenario. 
-If the best multi-partner scenario exceeds the fitness associated with the best single partner, this would be interpreted as evidence of complementarity, a true benefit of diversity \emph{per se}. 
+If the best multi-partner scenario exceeds the fitness associated with the best single partner, this would be interpreted as evidence of complementarity. 
 Alternatively, the sampling effect hypothesis predicts that no multi-partner scenario yields higher plant fitness then the best single partner. 
 It is also possible that multi-partner scenarios yield lower fitness than the single best partner, which would be consistent with an opportunity cost of diversity. 
-Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy (as a measure of how much benefit is received from all partners being present) for synchronized and non-synchronized responses to temporal stochasticity.
+Fourth, to quantify any contribution of the portfolio effect, we contrasted $\lambda_{S}$ of the full (four-state) scenario to vacancy for synchronized and non-synchronized responses to temporal stochasticity.
 If the portfolio effect confers a benefit of diversity, the fitness advantage of having all vs. no partners should be greater when temporal fluctuations are not synchronized across ant states.
@@
 We base our statistical inferences on the posterior probability distributions of the contrasts described above. 
-For example, the contrast of \textit{C. opuntiae} ($\lambda^{C}_{S}$) with vacancy ($\lambda^{V}_{S}$) yields a posterior distribution of the difference $\Delta\lambda^{C-V}_{S}$. 
-We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda^{C-V}_{S}>0)$. 
+For example, the contrast of \textit{C. opuntiae} ($\lambda^{C}_{S}$) with vacancy ($\lambda^{0}_{S}$) yields a posterior distribution of the difference $\Delta\lambda^{C-0}_{S}$. 
+We can quantify from this distribution our certainty in the mutualistic effect of \textit{C. opuntiae}, given the data, as $Pr(\Delta\lambda^{C-0}_{S}>0)$. 
 We apply similar logic to other contrasts described above. 
@@
 \subsection*{What are the effects of association with alternative partners on vital rates and how do these effects fluctuate across years?}
-Over the 20-year data set, we found that ant partners influenced demographic performance of cactus hosts, and different ant partners had contrasting demographic effects across host vital rates. 
+Over the 20-year data set, we found that ant partners influenced cactus demographic performance, and different ant partners had contrasting effects across host vital rates. 
 Ant-tended plants had a growth advantage over vacant plants, especially at smaller sizes (Figure \ref{fig:Grow}).
-Over the size range for which plants were likely ant-tended (larger than the minimum observed size of $0.8$ $log(cm^3)$; solid lines in Figure \ref{fig:Grow}) there were modest differences between partner species, with the greatest growth advantage associated with \textit{C.
+For the smallest sizes that were likely to be ant-tended (minimum observed size of ant-tended plants was $0.8$ $log(cm^3)$; solid lines in Figure \ref{fig:Grow}) there were modest differences between partner species, with the greatest growth advantage associated with \textit{C.
 opuntiae} followed by \textit{L.
 apiculatum} and then Other ants.
-At large sizes the growth trajectories of ant-tended and vacant plants were nearly indistinguishable. 
-For all ant states, growth was left-skewed, with transitions to sizes below the mean were more common than sizes above the mean.
+At the largest sizes, growth trajectories of ant-tended and vacant plants were nearly indistinguishable. 
+For all ant states, growth was left-skewed, with transitions to sizes below the mean were more frequent than sizes above the mean.
 
-Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). 
+\tom{Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). }{Stay consistent with use of dashed vs solid lines.}
 Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
-\textit{C. opuntiae}-occupied plants had a survival advantage over other ant-tended plants, particularly at smaller sizes, consistent with the positive effects on growth. 
+\textit{C. opuntiae}-occupied plants had a survival advantage over other ant-tended plants, particularly at smaller sizes, consistent with positive effects on growth. 
 At larger sizes, plants in any state of ant occupancy had a survival advantage over vacant plants. 
-Plants that were smaller than $-2$ $log(cm^3)$ were predicted to experience a survival advantage from vacancy, but apparent benefits of occupancy were based on extrapolated survival estimates of ant-tended plants.
+Plants smaller than $-2$ $log(cm^3)$ were predicted to experience negative effects of ant visitation, but this was based entirely on extrapolated survival estimates of ant-tended plants. 
 Plants in this size range were never observed to be ant-tended and, because the IPM preserves the size-dependence of vacancy, benefits of vacancy at small sizes are never realized in the IPM.
@@
 \begin{figure}
-	\includegraphics[width = 0.95\linewidth]{Figures/grow_contour_v2.png}
-	\caption{The next predicted size of cholla based on previous size with each individual ant partner.
-The colored lines (seen in all panels) are the next mean predicted size of cholla.
-Dashed lines indicate extrapolation while solid lines indicate estimations within the range of observed data.
-The points (seen in panels a-d) are the observed data which informs these estimates.
-The black contour lines (seen in a-d) appear at 5\% increments showing where 5\%, 10\%, etc.
-of the data is expected to fall.
-Panel e shows all mean estimates from panels a-d on one graph.
-The gray dashed line (in panel e only) shows the line where the next predicted size is the same as the previous (aka there is no growth on this line and below this line is shrinkage).}
+	\includegraphics[width = 0.75\linewidth]{Figures/grow_contour_v2.png}
+	\caption{The next predicted size of tree cholla based on previous size in relation to ant partner (a, \textit{C.
+opuntiae}; b, \textit{L.
+apiculatum}; c, Other; d, Vacant, e, all states combined).
+Points show observed size transitions, colored lines are the mean next size, and contours show probability density of the skewed normal growth model.
+	Dashed lines indicate extrapolation while solid lines indicate the range of observed data.}
 	\label{fig:Grow}
@@
 \begin{figure}[H]
-\includegraphics[width=0.95\linewidth]{Figures/survival_plot.png}
-\caption{The estimated survival rates based on the size of the cactus with each individual ant partner.
-The solid colored lines (shown on all panels) indicate the mean estimated survival rates.
-The dashed lines (shown in panel e) indicate extrapolations beyond existing data (where we estimated survival for plants tended by ants where we had never seen a tended cactus of that size).
-The grey area around the solid lines (shown in panels a-d) show the 90\% confidence interval for the estimates.
-The colored dots are the real data binned by size to show how our estimates align with real survival observations.
-A larger circle means we had more data on survival of plants of this size with this partner.
-The mean estimates from panels a-d are combined on one graph in panel e.}
+\includegraphics[width=0.75\linewidth]{Figures/survival_plot.png}
+\caption{Probability of survival in relation to size and ant partner.
+Layout as in Fig.
+\ref{fig:Grow}, except here points show data as binned means, where point size is proportional to number of observations.
+Grey areas show the 90\% confidence interval for the mean.}
 \label{fig:Surv}
@@
 
-We found evidence that ant tending was associated with increased floral viability rates and that ant identity can influence the strength of viability benefits.
-We observed mean viability rates of flowers between 40\% and 92\% (Figure \ref{fig:Viab}).
+Ant visitation was associated with increased floral viability and ant identity determined the strength of viability benefits.
+\tom{Mean viability rates spanned 40--92\% (Figure \ref{fig:Viab}).}{I am not sure where this range comes from. The group means look like they span about 55--88.}
 \textit{L.
-apiculatum}-tended plants had the highest mean viability rate (86.1\% [95\% credible interval: 77.6--92.4\%]), and there were similar viability rates for vacant (60.0\% [44.3--75.0\%]), Other-tended plants (at 60.6\% [43.7--75.5\%]), and \textit{C.
-opuntiae} tended plants (57.1\% [40.6--72\%]).
+apiculatum}-tended plants had the highest mean viability rate (86.1\% [95\% credible interval: 77.6--92.4\%]), while there were similar viability rates for vacant (60.0\% [44.3--75.0\%]), Other-tended (60.6\% [43.7--75.5\%]), and \textit{C.
+opuntiae}-tended plants (57.1\% [40.6--72\%]).
 Furthermore, \textit{C.
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
-apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure S\ref{app:AppA_Seeds_Per_Fruit}).
+apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure S\ref{app:AppA_Seeds_Per_Fruit}), likely due to pollinator interference \citep{Ohm2014}.
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/Viab_v2.png}
-	\caption{The estimated distributions of floral viability rates compared to observed distributions of floral viability rates of cholla based on ant partner identity.
-The clouds of points show the observed data, with the color representing the ant state.
-The solid lines indicate 95\% of the estimated viability posterior distribution.}
+	\caption{Flowerbud viability in relation to ant partner status.
+Points show observed data (fraction of initiated buds that flower).
+Points and bars show the means and 95\% credible intervals for each group.}
 	\label{fig:Viab}
@@
 
-The results of all other vital rates as well as the posterior predictive checks are all included in the Appendix A : Figures S\ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
+%The results of all other vital rates as well as the posterior predictive checks are all included in the Appendix A : Figures S\ref{app:AppA_Seeds_Per_Fruit} -- \ref{app:AppA_Pre_Census_Surv}.
 
-We analyzed the correlation coefficients of all models that included ant state as a predictor and found that, across vital rates, the annually varying effects on growth were the most correlated between ant states (mean pairwise correlation: 0.63) and effects on survival were the least correlated (0.36) (Figure \ref{fig:Annual_Ant}).
-%Weakly or uncorrelated fluctuations are necessary for the portfolio effect to occur because a central thesis of this effect is that partners must react differently to temporal environmental stochasticity.
-For growth, annually varying random effects of Other ants were the least synchronous with other ants states (correlation coefficients ranging from 0.28 to 0.39) while the effects of vacancy were the most synchronous with other states (correlation coefficients ranging from 0.39 to 0.81).
-For survival, effects of \textit{C.
-opuntiae} were the least synchronous with other ant states (correlation coefficients ranging from 0 to 0.15) while the effects of vacancy were the most synchronous (correlation coefficients ranging from 0.13 to 2.6).
-For floral viability, effects of \textit{L.
-apiculatum} were the least synchronous with other ant states (correlation coefficients ranging from 0.06 to 0.44) while effects of vacancy were the most synchronous (correlation coefficients ranging from 0.44 to 0.62).
-The variation in synchrony across ant states and vital rates indicates there is potential for portfolio effect, as many of the ant effects revealed low synchronicity, particularly in survival.
+Inter-annual fluctuations in demographic rates (estimated as statistical random effects) were generally positively correlated between ant states, limiting the potential for benefits of diversity through the portfolio effect (Figure \ref{fig:Annual_Ant}).
+However, the degree of correlation depended on the vital rate and pair of ant partners. 
+Across vital rates, random effects for cactus growth were the most strongly correlated between ant states (mean pairwise correlation: 0.63) and random effects for survival were the least correlated (mean pairwise correlation: 0.36).
+Certain partner pair / vital rate combinations were effectively independent in their yearly fluctuations (such as survival of \textit{C.
+opuntiae} and Other-tended plants) while others were almost completely synchronous (such as growth of Vacant and \textit{L.
+apiculatum}-tended plants).
 
@@
 	\includegraphics[width=0.95\linewidth]{Figures/ant_RFX.png}
-	\caption{The colored lines in each panel show the mean effect of each ant partner on a vital rate across years (blank periods indicate a lack of transition data).
-The blue squares in the upper corner of each panel show the correlation matrix of how similar ant-specific annual responses are for each vital rate.
-Panel a) shows the effects on the estimated next size, b) the effects on estimated survival, and c) the effects on floral viability of cacti across every year of our study.
-These values are estimated from the fitted random effects of ant and year in our models.
-Each point represents the mean of the random effect of the identified model, ant, and year (e.g.
-the lowest dot in panel b) represents the mean effect of vacancy on survival rates in year 2011).}
+	\caption{Year-specific random effects from statistical models for growth (a), survival (b), and floral viability (c) in relation to ant partner status.
+Inset panels show correlation coefficients (Pearson's \textit{r}) for each pair of time series.
+All random effects have mean zero such that positive values indicate better-than-average years and negative values indicate worse-than-average years.
+Values are missing where there were no data available for that year.}
 	\label{fig:Annual_Ant}
@@
 \subsection*{What are the frequency and direction of partner turnover across the plant life cycle?}
-We found that 55\% of individual plants surveyed in the long-term data experienced at least one ant state transition on average, with very distinct size-dependence and directional patterns (Figure \ref{fig:Ant_Transition}).
+A majority (55\%) of plants surveyed in the long-term data experienced at least one ant state transition, with distinct size-dependent and directional patterns (Figure \ref{fig:Ant_Transition}). 
 Vacancy was the most likely ant state of small plants ($\leq 10 log(cm^3)$).
 Even when small plants were ant-tended at the start of the transition year, they were most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
-The probability of becoming ant-tended increases with size, though it was not equally likely to be tended by all partners.  
+The probability of becoming ant-tended increased with size, though it was not equally likely to be tended by any partner.  
 For large plants that were initially vacant or tended by \textit{L.
@@
 opuntiae} were also able to retain plants they previously occupied, but not as well as \textit{L.
-apiculatum}: for plants that begin the transition year with \textit{C.
+apiculatum}: for plants that began the transition year with \textit{C.
 opuntiae}, the probability that those plants remain occupied by \textit{C.
 opuntiae} at the end of the transition year is only slightly greater than the probability of take-over by \textit{L.
-apiculatum}, while take-over in the other direction is extremely rare.
+apiculatum}, while take-over in the other direction was extremely rare.
 It is also notable that transitions away from the initial state of \textit{L.
@@
 	\includegraphics[width=0.95\linewidth]{Figures/transition.png}
-	\caption{The probability of being tended by each ant partner or vacant based on the size of the plant.
-Each panel shows these probabilities for a different previous ant state.
-The solid lines represent the mean probability of being tended by a specific partner.
-The colored points are the real data binned by size to show how our estimates align with real visitation observations.
-A larger circle means we had more data on visitation of plants of this size with this previous partner.}
+	\caption{The probability of ant state based on size and previous ant state.
+Each panel shows size-dependent probabilities of the next ant state conditional on the previous ant state.
+Solid lines represent predictions of the multinomial statistical model and points show data binned over size intervals, where point size is proportional to the number of observations.}
 	\label{fig:Ant_Transition}
@@
 \subsection*{What is the net effect of partner diversity on plant fitness, and what mechanism(s) explain(s) this effect?}
-By integrating vital rate results, temporal fluctuations, and ant transition dynamics into the stochastic, multi-state IPM we can evaluate the fitness implications of different scenarios of partner diversity and identity.
+By integrating vital rates, temporal fluctuations, and ant transition dynamics into the stochastic, multi-state IPM we can evaluate the fitness implications of different scenarios of partner identity and diversity (Figure \ref{fig:LambdaMeans}).
 First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
-The lowest mean stochastic fitness was $\lambda^{0}_{S}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}a).
+The lowest mean stochastic fitness was $\lambda^{0}_{S}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}A).
 Across all 1+ partner scenarios, we found that the $\lambda^{1+}_{S}$ posterior distributions were greater than $\lambda^{0}_{S}$ with nearly 100\% certainty.
-This indicates that ant visitation elevates fitness no matter the identity or number of partners.
-Furthermore, we found no benefits of partner diversity, with the fitness of cholla 1, 2, and 3 partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
-The 1- and 2-partner scenarios were statistically consistent but there was a modest reduction in fitness in the 3-partner scenario compared to 2 partners ($Pr(\Delta\lambda^{2-3}_{S}=0.95)$).
+This indicates that ant visitation elevates fitness no matter the number of partners.
+Furthermore, we found no benefit of partner diversity, with the fitness associated with one, two, or three partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
+The one- and two-partner scenarios were statistically consistent but there was a modest reduction in fitness in the three-partner scenario compared to two partners ($Pr(\Delta\lambda^{2-3}_{S}=0.95)$), consistent with a weak cost of diversity at the highest level.
+Patterns of $\lambda_{S}$ in relation to partner richness were generally consistent between scenarios of non-synchronous and synchronous inter-annual fluctuations. 
 %The contrasts of $\lambda_{NS}$ posterior distributions of 1 or 2 partner diversity scenarios to $\lambda_{NS}$ posterior distribution of 3 partners show that the overlap ranges from 58\% - 100\%, indicating high similarity between thr full partner diversity scenario and others.
@@
 	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_lines.png}
-	\caption{Panel a) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ for different numbers of partners (0-3) for the synchronous IPM (black circle) and the non-synchronous IPM (red diamond).
-The lines show the posterior distribution spread of the estimated $\lambda$ values.
-Panels b-c) the mean estimated $\lambda_{NS}$ and $\lambda_{S}$ respectively, for each simulated combination of ant partner as the filled in circles.
-The lines show the posterior distribution spread of estimated $\lambda$ values.
-The letters in the legend correspond to what ant partners are present (V = Vacant, C = \textit{C.
+	\caption{Tree cholla stochastic growth rate ($\lambda_{S}$) corresponding to number (A) or number and identity (B-C) of ant partners for synchronous and non-synchronous inter-annual fluctuations.
+Points and intervals show mean and 95\% credible intervals from the posterior distributions of $\lambda_{S}$.
+In B-C, letters in the legend correspond to ant partners as follows (V = Vacant, C = \textit{C.
 opuntiae}, L = \textit{L.
@@
 
-Partner identity and composition were not strongly consequential for plant fitness. 
-Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans}a). 
-While \textit{L.
-apiculatum}-tended plants had the highest mean fitness, it was not significantly higher than \textit{C.
-opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) tended plants.
-Furthermore, the fitness of \textit{L. apiculatum}-tended plants was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
+Partner identity and composition were not strongly consequential for plant fitness (Figure \ref{fig:LambdaMeans}b-c). 
+Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans}b). 
+While a hypothetical \textit{L.
+apiculatum}-only population had the highest mean fitness, it was not significantly higher than \textit{C.
+opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) only populations.
+Furthermore, the fitness of \textit{L. apiculatum}-only was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
 However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$). 
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
-Using the simulations where all ants had equal frequencies across sizes (further explained and analyzed in Appendix C), we found the same fitness patterns as in the competitive exclusion model described above.
+Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above.
 %Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
@@
 Mutualisms commonly involve multiple partners but the ecological consequences of partner diversity remain poorly understood. 
-Here we show that while alternative partners may be ecologically different, their fitness effects on a shared mutualist can be effectively equivalent and interchangeable.
-The results of our hierarchical models revealed that different ant partners had different effects on vital rates, with \textit{C.
-opuntiae} tended plants experiencing slight advantages in growth and survival when small, and \textit{L.
+Here we show that while alternative partners may be ecologically different, their fitness effects on a shared mutualist can be effectively interchangeable and redundant.
+The results of our hierarchical models revealed that different ant partners exhibit different effects on vital rates, with \textit{C.
+opuntiae}-tended plants experiencing advantages in growth and survival when small, and \textit{L.
 apiculatum}-tended plants experiencing floral viability advantages.
-The results of our synchronous and non-synchronous stochastic IPM revealed that all diversity scenarios which included any partners resulted in the highest possible fitness for tree cholla, suggesting that while ant visitation is beneficial, partner identity and diversity are inconsequential in this system.
-These results highlight that partners can differ meaningfully while still resulting in the same fitness benefits for a focal mutualist.
-
-% Connect to broader patterns in the lit
-Our conclusion that partner diversity does not benefit this system is novel in a filed where most studies highlight the benefits of diversity.
-Studies on multi-species mutualisms have shown that complementarity is a common mechanism explaining the benefits of diversity \citep{Afkhami2021,Rezende2007,Agrawal1998,Barrett2015,Stachowicz2005,Thibaut2012,Tornos2023}.
-In similar systems (plant-animal multi-species mutualisms) previous studies have reported complementarity \citep{Palmer2010,Frederickson2005,rogers2014} and portfolio effect \citep{Alarcon2008,Lazaro2022}.
-Only one other ant-plant study, to our knowledge, has highlighted a lack of benefit of diversity \citep{Bruna2014}, but they hypothesized these results occurred due to the extremely negative impacts of one partner and did not consider hypothetical scenarios without this partner to determine if there would be a benefit otherwise.
-Our finding is the only which has discovered and published a lack of diversity benefits when all partners are true mutualists.
-Vital rate results indicated that different ant partners affected different vital rates uniquely.
-On their own, these results suggested a likely benefit of diversity through complementarity, such that each ant partner offered a unique form of herbivore protection across ontogeny  and or across different life history processes (as seen in other studies \citep{Dejean2008,Barton2010,Boege2005}).
-However, our multi-state IPM results, which synthesize all life history processes and accounts for partner identity and turnover, revealed an overall positive effect of ant visitation but no apparent benefit of diversity, raising questions about how to resolve the tension between these two sets of results.
+These results, alone, would suggest some potential benefits of partner diversity through complementarity.  
+Yet, our stochastic, multi-state IPM revealed that all scenarios which included any partners resulted in similarly high cactus fitness, with a statistically significant but quantitatively weak cost of diversity at the highest level.
+Furthermore, evidence from nearly two decades of demographic data failed to support the ``portfolio effect'' hypothesis, whereby benefits of diversity manifest in the context of environmental fluctuations.
+Overall, our results indicate that while ant visitation is strongly consequential for plant fitness, partner identity and diversity are much less so.
 
 % why we see these effects in OUR population
-We speculate that our results are driven by the vital rate sensitivity structure of this population. 
+We attribute weak effects of partner identity and diversity to the vital rate sensitivity structure of this population. 
 Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
-Differences between alternative partners were most pronounced either in reproductive rates, which contribute relatively weakly to fitness (because a long reproductive lifespan overrides individual reproductive bouts), or in growth and survival at small sizes, where mortality risk was high and the probability of ant tending was low.
-At larger sizes with low mortality risk, ant-tended plants had a modest survival advantage over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness because of the high sensitivity of $\lambda$ established plant survival.
-The very real demographic differences between partner species in other vital rates and at other sizes do not register nearly as strongly in the currency of fitness.
+Differences between alternative partners were most pronounced either in reproductive rates (floral viability and seed number), which contribute relatively weakly to fitness because a long reproductive lifespan overrides individual reproductive bouts, or in growth and survival at small sizes, where mortality risk is relatively high.
+At larger sizes with low mortality risk, ant-tended plants had consistent growth and survival advantages over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness; differences between partner species in other vital rates and at other sizes do not register nearly as strongly.
+The weak reduction in fitness at the highest partner diversity level is harder to explain, but suggests some form of missed opportunity cost in a more crowded partner environment.
 
-% opportunity costs of multipl epartners
-There was modest evidence of opportunity costs imposed by the full diversity scenario.
-Cholla under the full diversity scenario experienced lower $\lambda_S$ than cholla with either one or two possible partners.
-The population fitness would be optimized by including any combination of twp partners rather than all three.
-There is an opportunity cost associated with only the full diversity scenario, indicating there is some mechanism that reduces fitness benefits when all partners are present.
-This could be due an over-saturation of benefits \citep{Stanton2003} or some type of conflict between partners which results in reduced benefits \citep{Afkhami2014}.
-No matter the mechanism behind this reduction of benefits, the optimal diversity scenario for these cholla is any scenario with only two partners present.
+Our results broaden the literature on multi-species mutualisms, a majority of which has demonstrated positive effects of partner diversity through complementarity \citep{Yang2024,Hernandez2020,Larimer2014,Afkhami2016,Fehling2022,Gustafson2005,McKeon2012,Palmer2010,Afkhami2021,Stachowicz2005} or portfolio effect \citep{Stevens2024,Thibaut2012,rogers2014,Lazaro2022}.
+For example, working in another sequentially-partnered ant protective mutualism of long-lived plants (\textit{Acacia drepanolobium}), Palmer et al.
+\citeyear{Palmer2010} showed that synergies between alternative partners elevate plant fitness above what could be expected from the single best partner.
+However, the \textit{Acacia} partner guild includes a ``castrating'' parasite that is uniquely able to boost plant growth by suppressing reproduction. 
+We speculate that partners that differ in degree but not type, such as those in our system, may be less likely to synergize. 
+Fewer studies have shown costs of mutualist diversity. 
+Bruna et al.
+\citeyear{Bruna2014} showed that ant diversity can depress fitness of the sequentially-partnered ant-plant \textit{Maieta guianensis} due to reduced interactions with the better defender.
+Costs of diversity have also been documented in systems in which multiple partners interact simultaneously with a shared mutualist, likely due to interference via the host resource budget \citep{Keller2018} or negative higher-order interactions \citep{Barrett2015}.
+One previous study (in a different ant-cactus system) showed neutral effects of partner diversity, but this study also showed that ant visitation was itself neutral, with no detectable effects on plant fitness \citep{Ford2015}.
+To our knowledge, ours is the first study documenting a strong, positive fitness effect of mutualism, regardless of mutualist identity or guild composition. 
+As this field grows, particularly with studies that integrate multiple vital rates across the life cycle, we will be better positioned to understand whether positive, neutral, or negative effects of partner diversity predominate, and the conditions under which each might be more likely.
 
 % Portfolio Effect
-Our work explicitly incorporated temporal environmental stochasticity, which raises the opportunity for portfolio effect as a mechanism of diversity benefits. 
-Yet, we find no evidence of portfolio effect within our system. 
-When partners exhibit different reactions to varying environments, interacting with multiple partners can lead to more consistent benefits across time \citep{Batstone2018}.
-We did not directly measure the reactions of ant partners to temporal environmental stochasticity, rather we measure how the effects of each partner on cholla demography fluctuate through time, and whether those fluctuations are correlated.
-The vital rate results led us to believe that portfolio effect may be at play due to the differences in ant responses to annual variation.
-However, there is no stabilization of fitness across temporal heterogeneity and therefore no benefits of diversity due to the portfolio effect.
-Other portfolio effect studies \citep{Lazaro2022,Tornos2023} found that asynchrony was linked to weak evidence of portfolio effect.
-One study \cite{dallas2022temporal} found that while portfolio effect was easy to show in theoretical models, it is often very weak or nonexistent in empirical data across many systems. 
-This indicates that it may be very difficult to detect, disguised by different mechanisms, or uncommon in nature.
-
-% These results are robust to different assumptions about turnover patterns & what we don't know about turnover
-Partner turnover is likely a significant driver behind the fitness we see within Cholla populations, however the processes which drive the actual turnover frequency and directions remain a mystery in this system.
-In the literature, it is clear that the frequency of partner turnover can have big effects on the fitness of the focal mutualist \citep{Fiala1994, Horvitz1986, Oliveira1999, Sachs2004}.
-The direction of partner turnover is also important when the identity of partners impacts the quality of benefits received \citep{Fonseca2003, Alonso1998, Dejean2008, Noe1994}.
-In our system we found that there are distinct patterns to the directions and frequency of partner turnover, but we still don't understand the mechanisms driving these turnovers.
-In the model presented in this paper we assumed that competitive exclusion was the driving mechanism behind the patterns we see.
-This assumption would indicate that ants are competing for each individual cacti, likely based on EFN quality and composition \citep{Heil2004,Heil2010}, and that the more common an ant is the better a competitor it is.
-In our simulations this means that when one ant is removed the ants which are left occupy the newly vacated cacti so that the proportion of vacant cacti never changes.
-We included this assumption because we believe it to be the most biologically reasonable based on the personal observations of multiple researchers.
-However, it is possible this assumption is incorrect, so we created two alternative simulations to determine how robust our results were to various assumptions about partner turnover drivers.
-The first alternative we tested is a model that assumes all ants have the same competitive ability rather than the differences in competitive ability we assume cause observed variety in ant frequency.
-In this hypothetical simulation all ants have equal probability of occupying any given cactus, but the level of vacancy across the population remains consistent with observed data.
-This simulation was designed to test if the overwhelming frequency of \textit{L. apiculatum} may be overriding any potential advantages of other partners.
-The results of this simulation (Figure S\ref{app:EqualLambdaMeans}) were effectively the same as the competitive exclusion model presented in the body of this paper.
-The second alternative we tested was a model that assumed rather than competition driving ant occupations, it was the population size of each ant species alone.
-In this hypothetical simulation when one ant species is removed from the model the cacti are left vacant.
-We believe this is an unlikely scenario as there appear to be vast populations of the ant species in the area. \ali{}{This is just personal observation so it might be best to remove this?}
-This simulation was designed to test if the fitness was more sensitive to changes in vacancy across the population than the frequency of any particular ant species.
-The results of this simulation revealed that when vacancy increased sampling effect is at play (Figure S\ref{app:FreqLambdaMeans}).
-This result highlights that the primary driver behind the results we presented are ant presence and that it is only in cases oh unrealistically high vacancy that there are benefits to partner diversity.
-Further information about both of these models can be found in Appendix C.
-Together, these results indicate that our assertions that the fitness of the cacti are more sensitive to changes in vacancy than partner identity are robust across multiple assumptions.
+When partners exhibit different reactions to varying environments, interacting with a diverse portfolio of partners can lead to more consistent benefits across time \citep{Batstone2018}.
+Our work explicitly incorporated temporal environmental stochasticity, yet we find no evidence for the portfolio effect as a mechanism of diversity benefits. 
+In ant-plant defensive mutualisms, any portfolio effect would emerge from fluctuations in the community of herbivores, and ant-specific responses to each herbivore species. 
+Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population -- notably, weather -- override any fluctuations in herbivore composition and ant-herbivore interactions.
+Dallas et al.
+\citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is often \tom{very weak or nonexistent in empirical data across many systems.
+}{Does this mean it is actually weak/nonexistent, or just difficult to detect?
+(I don't know the paper.)}
+This indicates that it may be difficult to detect, disguised by different mechanisms, or uncommon in nature.
 
 % Value of Long-term data
-This paper shows the importance of long-term datasets in investigating species interactions and calls for further use of long-term data. 
+This study highlights the value of long-term data in investigating species interactions. 
 Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \citep{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
 Separate studies have analyzed how inter-annual variability impacts focal mutualists \citep{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
-The long term data set we used gave us the unique ability to consider the combined effects of partner identity, partner turnover, and temporal stochasticity.
-By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of specific demographic effects. 
-For example, our previous study suggested that \textit{C.
+The long term data set we used gave us the ability to consider, for the first time, the combined effects of partner identity, partner turnover, and temporal stochasticity. 
+By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of ant visitation. 
+For example, our previous study that focused only on reproductive vital rates suggested that \textit{C.
 opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set \citep{Ohm2014}.
-Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages, indicates that this species is clearly a mutualist and nearly as strong a mutualist as \textit{L.
-apiculatum}.
-
-% Herbivory
-Herbivory is an important driver of the fitness of cholla in this system, as herbivores directly impact the growth, survival, and reproductive efforts of the cacti \citep{Miller2009}.
-We have not explicitly incorporated herbivory but rather assume that herbivory can be captured in the effects they have on growth, survival, and reproductive effort observations.
-Analysis of our observational herbivore data shows that vacant plants experience elevated levels of herbivory compared to tended ones (Figure S\ref{app:herb}).
-In the future, further studies on the direct impacts of herbivory would bolster the results reported here and expand our knowledge within this system.
+Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.
 
 % Limitations
-As with any study, there are limitations to consider when interpreting the results reported here.
-These results highlighted in the paper are based on observational data regarding ant effects on plant demography rather than experimental data, meaning we are able to determine correlations but not causation.
-However, we have previously conducted experimental manipulations which revealed that ant presence has direct impacts on plant demography in our system \citep{Miller2007}.
-This combination of observational data backed up by experimental results gives us greater confidence in our causal interpretations than if we had only observational data.
-Specifically, we are confident in claiming that ant presence causes increased performance in cactus vital rates and the fitness signals we are detecting are directly related to ant partner presence.
-One further expansion could include nectar analysis in conjunction with ant interactions to look not only at the ant partner impacts on the plant demography, but also at the ability of plants to attract specific partners through nectar composition shifts.
+As with any study, there are limitations to consider when interpreting our results.
+First, we used observational data to infer ant effects on plant demography. 
+However, we have previously conducted experimental manipulations which revealed that ant presence has causal effects on plant demographic rates through anti-herbivore defense \citep{Miller2007,Ohm2014}.
+The combination of long-term observational data backed up by experimental results gives us greater confidence in our causal interpretations. 
+Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
+Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
+Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
+As more long-term data accumulate, we may gain more information with which to further unpack this group of species.
 
 % tie back paragraph.
-This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangeable in their net effects on fitness of a shared mutualist.
-The individual estimation of each vital rate and explicit inclusion of ant partners in this IPM allowed us to determine that partner diversity is not beneficial in this system and may highlight circumstances under which we would expect similar results in other systems.
-The individual vital rate estimation approach allowed us to determine that while different ant partners did impact each vital rate differently, the largest differences were estimated to occur at sizes when ants would not be present in real life or in vital rates that do not have a significant impact on the overall fitness.
-Considering partner impacts on vital rates across the entire life cycle of the plant allows more power in determining the importance of these differences.
-Based on our results we would expect to see that partner diversity does not matter when partner impacts occur during low sensitivity life stages.
+\paragraph{Conclusion}
+This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangeable in their net effects on fitness of a shared mutualist based on where their effects occur in the life cycle.
+Contrary to much of the literature on multi-species mutualisms, we find that while ant partners are beneficial, partner diversity has no measurable benefit to their shared mutualist and may even be somewhat costly.
+As studies accumulate across different types of multi-partner mutualisms, particularly those spanning simultaneous/sequential partnerships and leveraging demographic data from the complete life cycle, our field will be better poised to understand whether, how, and under what conditions mutualist diversity is neutral, costly, or beneficial.
 
 \section*{Acknowledgments}
-We are grateful to the Sevilleta community (LTER and US Fish and Wildlife Service) for providing a stimulating research environment and logistical support. 
-Jeremiah Dye identified ants and Massa Takahashi identified herbivores. 
+Financial support for this work came from the Sevilleta Long-Term Ecological Research Program (NSF DEB awards 1655499 and 1748133).
 We acknowledge the many students and technicians who have helped collect annual census data for this project, including M.
@@
 Xiong.
-Financial support for this work came from the Sevilleta LTER (NSF DEB-0217774)
+J. Moutouama, A. Bradley, B. Scherick, A. Gough, J. Martin, and V. Rudolf provided helpful feedback on the manuscript.
 
@@
 \section*{Data and Code Availability}
-The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1
+The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} on the Environmental Data Initiative website, package id: \tom{knb-lter-sev.323.1}{UPdate this so that it's a normal author-year citation.
+And update the data package!}
 

### 2024-12-19 Manuscript/ant_cactus_AmNat.tex
Updates

@@
 This study was conducted at the Sevilleta Long-term Ecological Research site (SEV-LTER), located within the Sevilleta National Wildlife Refuge in central New Mexico, USA.
-Out study area in the Los Pi$\tilde{n}$os mountains \tom{()}{Give lat-long} is characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
+Out study area in the Los Pi$\tilde{n}$os mountains ($34^\circ20'5.3''$N, $106^\circ37'53.2''$W) is characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
 gracilis}), yuccas, cacti, and junipers.
@@
 In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. 
-In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds \tom{(Table \ref{})}{Missing table reference.
-There is no table in the supplement, but there is an empty heading for ``Table''}.
+In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds.
 
@@
 n(x',a')_{t+1} = (\gamma_1 B^1_{t} + \gamma_2 B^2_{t} ) \eta(x') \omega \rho_{0}(a')  + 
-\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a';\pmb{\tau^{\epsilon}}) n(x,a)_t dx 
+\sum_{a=1}^{A} \int_L^U S(x,a;\pmb{\tau^S_{a}}) G(x',x,a;\pmb{\tau^G_{a}}) \rho(x,a,a') n(x,a)_t dx 
 	\end{gather}
@@
 As above, these functions include inter-annual variability through year-specific deviates that can vary by ant state ($\pmb{\tau}_{a}$). 
-Finally, ant transition function $\rho(a',a,x;\pmb{\tau}^{\rho})$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$. 
-\tom{This function includes inter-annual variability through year-specific intercepts which are consistent across initial ant states ($\pmb{\tau}^\rho$).}{Confirming that this model includes random year effects?}
+Finally, ant transition function $\rho(a',a,x)$ gives the probability that an individual transitions from ant state $a$ to $a'$ in the next census, conditional on initial size $x$.
 
@@
 The flowerbud viability sub-model ($V(a;\pmb{\tau}^{V}_{a})$) estimates the probability that flowerbuds produced in year $t$ remain viable, with fixed effects of ant partner $a$ in year $t$.
-We fit this model to floral viability data using a \tom{binomial}{For the inevitable revision, I would like to try beta-binomial.} distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
+We fit this model to floral viability data using a binomial distribution where trials and successes are given by the total number of flowerbuds and the number that are viable, respectively.
 This model used a logit link function and included random effects for plot and year following the same structure as the survival model, with ant-specific year random effects.
@@
 For each model, we obtained three chains of 10,000 iterations, discarding the first 1,500 iterations. 
-\tom{We visually assessed parameter convergence between and within chains (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}b) and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data (Figures S\ref{fig:Surv_post} -- \ref{fig:Pre_Surv_post}a).}{Not sure these are all necessary in the supplement - also I think these are not labeled correctly here (the a and b).}
+We visually assessed parameter convergence between and within chains and assessed overall model fit with posterior predictive checks to examine how well the fitted model can generate simulated data similar to the real data.
 
@@
 We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as 
-$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \tom{\citep{Mark2009}}{Mark is Mark Rees' first name. This bibtex entry needs to be updated.}.
+$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \citep{Rees2009}
 We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and the derived quantities described next.
@@
 
-\tom{Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). }{Stay consistent with use of dashed vs solid lines.}
+Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). 
 Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\includegraphics[width = 0.75\linewidth]{Figures/grow_contour_v2.png}
@@
 Ant visitation was associated with increased floral viability and ant identity determined the strength of viability benefits.
-\tom{Mean viability rates spanned 40--92\% (Figure \ref{fig:Viab}).}{I am not sure where this range comes from. The group means look like they span about 55--88.}
+Mean viability rates spanned 55--81\% (Figure \ref{fig:Viab}).
 \textit{L.
@@
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
-apiculatum}-tended plants (142.4[100.7--200.2] seeds) (Figure S\ref{app:AppA_Seeds_Per_Fruit}), likely due to pollinator interference \citep{Ohm2014}.
+apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.
 
@@
 \begin{figure}[H]
-	\includegraphics[width=0.95\linewidth]{Figures/ant_RFX.png}
+	\includegraphics[width=0.75\linewidth]{Figures/ant_RFX.png}
 	\caption{Year-specific random effects from statistical models for growth (a), survival (b), and floral viability (c) in relation to ant partner status.
@@
 Vacancy was the most likely ant state of small plants ($\leq 10 log(cm^3)$).
-Even when small plants were ant-tended at the start of the transition year, they were most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition}b-d). 
+Even when small plants were ant-tended at the start of the transition year, they were most likely to transition back to vacancy (Figure \ref{fig:Ant_Transition} b-d). 
 The probability of becoming ant-tended increased with size, though it was not equally likely to be tended by any partner.  
@@
 It is also notable that transitions away from the initial state of \textit{L.
-apiculatum} were almost always transitions to vacancy (Figure \ref{fig:Ant_Transition}d), while transitions away from the initial states of \textit{C.
+apiculatum} were almost always transitions to vacancy (Figure \ref{fig:Ant_Transition} d), while transitions away from the initial states of \textit{C.
 opuntiae} and Other  were often transitions to other ants.
@@
 First, there was strong evidence that ant visitation had mutualistic fitness effects on plant partners. 
-The lowest mean stochastic fitness was $\lambda^{0}_{S}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans}A).
+The lowest mean stochastic fitness was $\lambda^{0}_{S}$, the fitness of the cholla with no partners (Figure \ref{fig:LambdaMeans} a).
 Across all 1+ partner scenarios, we found that the $\lambda^{1+}_{S}$ posterior distributions were greater than $\lambda^{0}_{S}$ with nearly 100\% certainty.
@@
 	\includegraphics[width=0.91\linewidth]{Figures/Lambdas_Comp_lines.png}
-	\caption{Tree cholla stochastic growth rate ($\lambda_{S}$) corresponding to number (A) or number and identity (B-C) of ant partners for synchronous and non-synchronous inter-annual fluctuations.
+	\caption{Tree cholla stochastic growth rate ($\lambda_{S}$) corresponding to number (a) or number and identity (b-c) of ant partners for synchronous and non-synchronous inter-annual fluctuations.
 Points and intervals show mean and 95\% credible intervals from the posterior distributions of $\lambda_{S}$.
-In B-C, letters in the legend correspond to ant partners as follows (V = Vacant, C = \textit{C.
+In b-c letters in the legend correspond to ant partners as follows (V = Vacant, C = \textit{C.
 opuntiae}, L = \textit{L.
@@
 
-Partner identity and composition were not strongly consequential for plant fitness (Figure \ref{fig:LambdaMeans}b-c). 
-Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans}b). 
+Partner identity and composition were not strongly consequential for plant fitness (Figure \ref{fig:LambdaMeans} b-c). 
+Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans} b). 
 While a hypothetical \textit{L.
@@
 Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population -- notably, weather -- override any fluctuations in herbivore composition and ant-herbivore interactions.
-Dallas et al.
-\citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is often \tom{very weak or nonexistent in empirical data across many systems.
-}{Does this mean it is actually weak/nonexistent, or just difficult to detect?
-(I don't know the paper.)}
+Dallas et al. \citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is often very difficult to detect in empirical data across many systems. 
 This indicates that it may be difficult to detect, disguised by different mechanisms, or uncommon in nature.
@@
 \section*{Data and Code Availability}
-The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} on the Environmental Data Initiative website, package id: \tom{knb-lter-sev.323.1}{UPdate this so that it's a normal author-year citation.
-And update the data package!}
+The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1
 
@@
 
-\section*{Tables}
-\renewcommand{\thetable}{\arabic{table}}
-\setcounter{table}{0}
-
-\renewcommand{\thetable}{\arabic{table}}
-\setcounter{table}{0}
-
 \appendix
@@
 $$y^{\kappa} \sim  Negative Binomial(\hat{\kappa},\hat{\phi})$$
-%$$ \hat{\kappa } = \beta_{0}^{\kappa} \times a'$$
 $$\hat{\phi} = \beta_{0}^{\phi}$$
@@
 apiculatum} tended plants (142.4 per fruit), and finally, \textit{C.
-opuntiae} tended plants (115.0 per fruit) (Figure \ref{app:AppA_Seeds_Per_Fruit}).
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
-	\caption{The estimated posterior distribution of seeds per fruit on plants tended by \textit{C. opuntiae} (purple), \textit{L. apiculatum} (Teal), and Vacant plants (Yellow). }
-	\label{app:AppA_Seeds_Per_Fruit}
-\end{figure}
+opuntiae} tended plants (115.0 per fruit).
 
 \paragraph{Recruit Size Distribution}
-%The recruit size model ($n(x',a')$) estimates the size distribution of all recruits } from a given year $t+1$, with no fixed or random effects. 
 We fit this model to recruit size data $y^{\eta}$ using a Normal distribution with the identity link function: 
 $$y^{\eta} ~\sim N(\hat{\eta},\hat{\sigma})$$
-%$$\hat{\eta} = \beta_{0}^{\eta}$$
 where $\hat{\sigma}$ is estimated with a non-informative prior.
 
-We found that the mean size of recruits is $log(-2.097) m^3$ with an interquartile range from $log(-2.173) m^3$ to $log(-1.712) m^3$ (Figure \ref{app:AppA_Recruit_Dist}).
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/rec_size.png}
-	\caption{The estimated posterior distribution of recruit sizes as the $log(volume)$ of cholla. }
-	\label{app:AppA_Recruit_Dist}
-\end{figure}
+We found that the mean size of recruits is $log(-2.097) m^3$ with an interquartile range from $log(-2.173) m^3$ to $log(-1.712) m^3$.
 
@@
 $$y^{\gamma_2} \sim Binomial(\hat{\gamma_2})$$
-%$$logit(\hat{\gamma_1}) = \beta_{0}^{\gamma_1}$$
-%$$logit(\hat{\gamma_2}) = \beta_{0}^{\gamma_2}$$
 
 We found that the mean germination rates for seeds in the seedbank for one year  is 0\%, with an interquartile range of 0\% and 1\%.
-We found that the mean germination rates for seeds in the seedbank for a second year is 0\%, with an interquartile range of 0\% to 0.4\% (Figure \ref{app:AppA_Germ}).
+We found that the mean germination rates for seeds in the seedbank for a second year is 0\%, with an interquartile range of 0\% to 0.4\%.
 Seeds are more likely to germinate in their first year in the seedbank, but most seeds will never germinate.
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/germination.png}
-	\caption{The estimated posterior distributions for germination probability of seeds in the first year seedbank (blue) and seeds in the second year seedbank (green).}
-	\label{app:AppA_Germ}
-\end{figure}
 
@@
 $$y^{\delta} ~ Bern(\hat{\delta})$$
-%$$logit(\hat{\delta}) = \beta_{0}^{\delta} + m$$
 where $m \sim N(0, \sigma_{transect}^2)$ is the random effect of transect where the recruited individual was analyzed for survival.
@@
 We found that plants have a 16.2\% probability of surviving from germination to the next census.
-Our model estimated this very well, expecting a 16.3\% probability (Figure \ref{app:AppA_Pre_Census_Surv}).
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/seed_surv.png}
-	\caption{The distribution of the probability of plants which germinated between censuses surviving to the next census to be counted.
-The green line shows the estimated posterior distribution of pre-census survival.}
-	\label{app:AppA_Pre_Census_Surv}
-\end{figure}
+Our model estimated this very well, expecting a 16.3\% probability.
 
 \paragraph{Portfolio Effect}
-\begin{figure}
+We compared the effects of partner diversity (estimated as $\lambda_S^4 - \lambda_S^0$) for the synchronized and non-synchronized models.
+We determined that non-synchronous effects of ant partners are greater than synchronous effects of ant partners only 50\% of the time.
+This means there is no evidence of portfolio effect in this system.
+
+\begin{figure}[H]
 	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
-	\caption{The distribution of $\lambda_{NS,All}-\lambda_{NS,Vacant}$ in pink and $\lambda_{S,All}-\lambda_{S,Vacant}$ in green.
+	\caption{The distributions of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in pink and green respectively.
 The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
@@
 Ant visitors are believed to offer defensive benefits to the plants they tend in this system, leading to the hypothesis that ant presence would be correlated with reduced herbivory.
+Therefore, herbivore protection is the likely driver behind the increased survival, growth, and reproductive rates of tended plants. 
+Fluctuations in herbivore communities across time would also lead to potentially important fluctuations in ant visitor responses and therefore cholla demographic responses.
+
+Here we conduct  an analyses of herbivory levels on plants with different ant visitors and contrast them to herbivory levels on plants with no visitors.
+Evidence that herbivory is decreased for tended plants further strengthens the results presented in this paper that ant presence is beneficial on a demographic level for cholla.
+
 Herbivory data was collected during censuses any time herbivores were identified on a plant. 
@@
 
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/herb_all.png}
+\begin{figure}[H]
+	\includegraphics[width=0.7\linewidth]{Figures/herb_all.png}
 	\caption{The proportion of reproducing plants which are visited by herbivores. Each bar represents the subset of the cacti population in a different ant state.  }
@@
 For each model fitted, we conducted two tests to determine if the fit was acceptable to use in our IPM. 
-First, we checked the convergence of each parameter.
-Below we show the convergence of all $\beta$ terms listed in the Statistical Modeling subsection of Methods.
+First, we checked the convergence of each parameter and determined they were acceptable.
 Second, we checked the posterior fit, comparing the estimated values of each model to the $y$ values of the actual data.
-We show these posterior checks below, split by ant partner where relevant.
 Each model was run with 10,000 total iterations per chain, 5,000 warmups, and thinning set to 2, leaving us with a total of 2,500 iterations post warmup.
-%% Growth Figure
-%% Survival Figure
-\begin{figure}
-\includegraphics[width = 0.45\linewidth]{Figures/surv_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/surv_post.png}
-\caption{The a) posterior convergence of the parameters estimated by the survival model and the b) posterior distribution of survival estimates (pink lines) for each ant species (1 = \textit{C.
-opuntiae}, 2 = \textit{L.
-apiculatum}, 3 = other, 4 = vacant) compared to the mean survival distribution (black line) of the real data.}
-\label{fig:Surv_post}
-\end{figure}
-%% Reproduction Figure
-\begin{figure}
-\includegraphics[width = 0.45\linewidth]{Figures/repro_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/repro_post.png}
-\caption{The a) posterior convergence of the parameters estimated by the reproduction model and the b) posterior distribution of reproductive status estimates (pink lines) for each ant species (1 = \textit{C.
-opuntiae}, 2 = \textit{L.
-apiculatum}, 3 = other, 4 = vacant) compared to the mean reproductive status distribution (black line) of the real data.}
-\label{fig:Repro_post}
-\end{figure}
-%% Flowering Figure
-\begin{figure}
-\includegraphics[width = 0.45\linewidth]{Figures/flow_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/flow_post.png}
-\caption{The a) posterior convergence of the parameters estimated by the number of flowers model and the b) posterior distribution of the number of flowers estimated (pink lines) compared to the mean distribution of observed flowers (black line).}
-\label{fig:Flow_post}
-\end{figure}
-%% Viability Figure
-\begin{figure}
-\includegraphics[width = 0.45\linewidth]{Figures/viab_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/viab_post.png}
-\caption{The a) posterior convergence of the parameters estimated by the viability model and the b) posterior distributions of floral viability estimates (pink lines) for each ant species (1 = \textit{C.
-opuntiae}, 2 = \textit{L.
-apiculatum}, 3 = other, 4 = vacant) compared to the mean floral viability distribution (black line) of the real data.}
-\label{fig:Viab_post}
-\end{figure}
-%% Seeds Per Fruit
-\begin{figure}
-\includegraphics[width = 0.45\linewidth]{Figures/seed_conv.png}
-\includegraphics[width=0.45\linewidth]{Figures/seed_post.png}
-\caption{The a) posterior convergence of the parameters estimated by the seeds per fruit model and the b) posterior distributions of seeds per fruit estimates (pink lines) for each ant species (1 = \textit{C.
-opuntiae}, 2 = \textit{L.
-apiculatum}, 3 = vacant) compared to the mean seeds per fruit distribution (black line) of the real data.}
-\label{fig:seed_post}
-\end{figure}
-%% Fruit survival 
-% \begin{figure}
-%% 	\includegraphics[width = 0.45\linewidth]{Figures/                .png}
-%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
-% 	\caption{The a) posterior convergence of the parameters estimated by the fruit survival model and the b) posterior distributions of fruit survival estimates (pink lines) compared to the mean fruit survival distribution (black line) of the real data.}
-% 	\label{fig:fruit_surv_post}
-% \end{figure}
-% %% Ant Transitions
-% \begin{figure}
-%% 	\includegraphics[width = 0.45\linewidth]{Figures/             .png}
-%% 	\includegraphics[width=0.45\linewidth]{Figures/               .png}
-% 	\caption{The a) posterior convergence of the parameters estimated by the ant transitions model and the b) posterior distributions of next ant partners estimates (pink lines) for each previous ant species (1 = \textit{C.
-opuntiae}, 2 = \textit{L.
-apiculatum}, 3 = other, 4 = vacant) compared to the mean next ant partner distribution (black line) of the real data.}
-% 	\label{fig:Transitions_post}
-% %% Germination
-\begin{figure}
-	\includegraphics[width = 0.45\linewidth]{Figures/germ1_conv.png}
-	\includegraphics[width=0.45\linewidth]{Figures/germ2_conv.png}\\
-	\includegraphics[width=0.45\linewidth]{Figures/germ1_post.png} 	\includegraphics[width=0.45\linewidth]{Figures/germ2_post.png}
-	\caption{The a-b) posterior convergence of the parameters estimated by the germination from year one seedbank and germination from year two seedbank models respectively.
-The c-d) posterior distributions of floral viability estimates (pink lines) compared to the mean germination distribution (black line) of the real data for first year germinants and second year germinants respectively.}
-	\label{fig:Germ_post}
-\end{figure}
-
-%% Pre-Census Survival
-\begin{figure}
-	\includegraphics[width = 0.45\linewidth]{Figures/seed_surv_conv.png}
-	\includegraphics[width=0.45\linewidth]{Figures/seed_surv_post.png}
-	\caption{The a) posterior convergence of the parameters estimated by the pre-census survival model and the b) posterior distribution of the pre-census survival estimated (pink lines) compared to the mean distribution of observed pre-census survival (black line).}
-	\label{fig:Pre_Surv_post}
-\end{figure}
-%%
 

### 2024-12-19 03_cholla_ant_IPM_params_functions.R

@@ -32,7 +32,7 @@ years <- unique(cactus$Year_t)
 ## -------- read in MCMC output ---------------------- ##
 ## Choose your pathway to pull from 
 #Ali
-# mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
+ mcmc_dir <- "/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/"
 mcmc_dir <- "/Users/alicampbell/Library/CloudStorage/GoogleDrive-amc49@rice.edu/Shared drives/Miller Lab/Sevilleta/Cholla/Model Outputs/"
 #Tom
 # mcmc_dir <- "G:/Shared drives/Miller Lab/Sevilleta/Cholla/Model Outputs/"

### 2024-12-19 05_cholla_ant_IPM_figures_tables.R

@@ -589,19 +589,23 @@ png("Manuscript/Figures/survival_plot.png")
 par(mar=c(3,3,3,1),oma=c(2,2,0,0))
 layout(matrix(c(1,2,3,4,5,5),ncol = 3, byrow = TRUE), heights = c(1.5,1.5), widths = c(3.9,3.9,3.9))
 # Crem
-plot(x = size_crem, y = invlogit(quantile(surv.params$beta0[,1],.5) + size_crem * quantile(surv.params$beta1[,1],.5)),lty = 2, type = "l", col = cremcol, lwd = 4, ylim = c(0,1), xlim = c(-5,15),cex.main = 2, main = "a)                        ")
+plot(x = size_crem, y = invlogit(quantile(surv.params$beta0[,1],.5) + size_crem * quantile(surv.params$beta1[,1],.5)),lty = 1, type = "l", col = cremcol, lwd = 4, ylim = c(0,1), xlim = c(-5,15),cex.main = 2, main = "a)                        ")
+lines(x = size_dummy, y = invlogit(quantile(surv.params$beta0[,1],.5) + size_dummy * quantile(surv.params$beta1[,1],.5)), col = cremcol, lwd = 4, lty = 2)
 points(surv_plot_crem$mean_size,surv_plot_crem$surv,pch=16,cex=surv_plot_crem$N_mod,col= alpha(cremcol, 0.4))
 polygon(c(size_dummy,rev(size_dummy)),c(invlogit(percentiles("crem",0.95)), rev(invlogit(percentiles("crem",0.05)))), col = rgb(red = 0.2, blue = 0.2, green = 0.2,alpha = 0.1), border = NA)
 # Liom
-plot(x = size_liom, y = invlogit(quantile(surv.params$beta0[,2],.5) + size_liom * quantile(surv.params$beta1[,2],.5)),lty = 2, type = "l", col = liomcol, lwd = 4, ylim = c(0,1), xlim = c(-5,15),cex.main = 2, main = "b)                        ")
+plot(x = size_liom, y = invlogit(quantile(surv.params$beta0[,2],.5) + size_liom * quantile(surv.params$beta1[,2],.5)),lty = 1, type = "l", col = liomcol, lwd = 4, ylim = c(0,1), xlim = c(-5,15),cex.main = 2, main = "b)                        ")
[…]

### 2024-12-19 Data_Package_Script.R

@@ -1,47 +1,10 @@
-################################################################################ 
-################################################################################
-## The purpose of this script is to load all required packages, load the data 
-## and fix any errors (such as excel taking numbers as dates or mistyping ant 
-## species names etc.), write a "clean" data file which has all errors 
-## fixed and can be called directly and run through all other r scripts, and 
-## finally load in any other data files.
-################################################################################
-################################################################################
-## Load All Necessary Packages Here
-# stan models
-library(rstan)
-options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = TRUE)
-# distributions
-library(sn)
-library(sgt)
-library(qgam)
-library(brms)
-library(loo)
-library(simcausal)
-# ipm analysis
-library(popbio)
-# visual model checks
-library(moments)
-library("posterior")
-library(gridExtra)
-library(grid)
-library(MASS)
-library(patchwork)
-library(tidyverse)
-# data management
-library(dplyr)
-library(magrittr)
-# graphics
-library(RColorBrewer)
-library(corrplot)
-library(ggpubr)
-library(reshape)
-library(reshape2)
-library(stats)
-# knitr::oknitr::oknitr::opts_chunk$set(echo = TRUE)
-# check the workign directory and make sure it is set to the right location
-#getwd()
+#############################################################
+## The purpose of this script is to accompany a data package.
+## This script will go through the process of cleaning errors 
+## in the data entry. d
+#############################################################
+
+# Set the working directory 
[…]

### 2024-12-19 05_cholla_ant_IPM_figures_tables.R
Update 05_cholla_ant_IPM_figures_tables.R

@@ -1426,7 +1426,7 @@ herbivory %>%
   group_by(Year_t1) %>%
   summarise(h = sum(herb_YN))
 # use 2014 and on 
-herbivory <- subset(herbivory,herbivory$Year_t1 >= 2014)
+herbivory <- subset(herbivory,herbivory$Year_t1 == 2014)
 
 
 ## Subset by ant

### 2024-12-23 Manuscript/ant_cactus_AmNat.tex
Manuscript

@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Yeung2006,Barrett2015,Ushio2020}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020}.
 First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.

### 2024-12-23 Manuscript/ant_cactus_AmNat.tex
Manuscript

@@
 We attribute weak effects of partner identity and diversity to the vital rate sensitivity structure of this population. 
-Like other long-lived, iteroparous species \citep{franco2004comparative}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
+Like other long-lived, iteroparous species \citep{Franco2004}, tree cholla fitness is most sensitive to the growth and survival of established individuals \citep{Miller2009,elderd2016quantifying}, which are virtually guaranteed many years of reproductive opportunities once they reach a size that is protected from mortality.
 Differences between alternative partners were most pronounced either in reproductive rates (floral viability and seed number), which contribute relatively weakly to fitness because a long reproductive lifespan overrides individual reproductive bouts, or in growth and survival at small sizes, where mortality risk is relatively high.

### 2024-12-24 Manuscript/ant_cactus_AmNat.tex
Tom's final push

@@
 Partner diversity may benefit a focal mutualist through mechanisms that mirror positive effects of species diversity on ecosystem function, such as complementarity or portfolio effect. 
-We studied the fitness consequences of partner diversity in the multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
-We used long-term data (2004--2023) and Bayesian hierarchical statistical models to determine the demographic impacts of different ant partners, and the dynamics of turnover from one exclusive partner association to another.
+We studied the fitness consequences of partner diversity and turnover in the multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
+We used long-term data and Bayesian hierarchical statistical models to determine the demographic impacts of different ant partners, and the dynamics of turnover from one exclusive partner association to another.
 We then constructed a stochastic, multi-state Integral Projection Model (IPM) with which we simulated tree cholla populations with different richness and composition of partners.
@@
 Furthermore, there was no evidence that benefits of partner diversity manifest in the context of fluctuating environments (i.e., the portfolio effect). 
-This study highlights that demographic benefits of mutualism can be surprisingly robust to diversity and turnover of partners.
+In contrast to much of the previous literature, this study highlights that demographic benefits of mutualism can be surprisingly robust to diversity and turnover of partners.
 
@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \tom{\citep{Barrett2015,Ushio2020}.}{Cite the Winfree chapter?}
 First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
 This can lead to an apparent benefit of diversity driven by a sampling effect \citep{Batstone2018} -- but there is no benefit of diversity \emph{per se}, only better and worse partners. 
-If partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
 Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \citep{Stachowicz2005,Bronstein2006,Stanton2003}. 
@@
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}. 
+Previous studies have investigated how ant partner diversity affects plant fitness \tom{\citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}.
+}{The Afkhami paper cited here is not an ant-plant study.}
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
@@
 This study was conducted at the Sevilleta Long-term Ecological Research site (SEV-LTER), located within the Sevilleta National Wildlife Refuge in central New Mexico, USA.
-Out study area in the Los Pi$\tilde{n}$os mountains ($34^\circ20'5.3''$N, $106^\circ37'53.2''$W) is characterized by steep, rocky slopes, and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
+Our study area in the Los Pi$\tilde{n}$os mountains ($34^\circ20'5.3''$N, $106^\circ37'53.2''$W) is characterized by steep, rocky slopes and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
 gracilis}), yuccas, cacti, and junipers.
@@
 Flowerbuds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
-Smaller, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation at the time of our census) \citep{Miller2014}.
+Smaller, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation) \citep{Miller2014}.
 
@@
 This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
-Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown). 
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD tree cholla per plot: \tom{}{Give the mean and SD of cactus density per plot (plots 1-8) here.}) has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
 For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
@@
 In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. 
-In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds.
+In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds.
 
@@
 apiculatum} and \textit{C.
-opuntiae} (\textless8\% of observations) we combined all other ants into an ``other'' category, such that our multi-state IPM included four possible ant states: vacant, \textit{L.
+opuntiae} (\textless8\% of observations) we combined all other ants into an ``Other'' category, such that our multi-state IPM included four possible ant states: vacant, \textit{L.
 apiculatum}, \textit{C.
@@
 We modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates.
-The models also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
+The model also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
 Seed bank dynamics are given by:
@@
 \noindent The first term in Eq.
-(\ref{eqn:IPM2}) estimates the number of individuals recruiting from a one or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
+(\ref{eqn:IPM2}) estimates the number of individuals recruiting from a one- or two-year seed bank to a plant of size $x'$ and ant state $a'$ based on the recruit size distribution $\eta(x')$ and the probability of over-winter seedling survival ($\omega$) from germination (late summer) to the census (May).
 This term is multiplied by $\rho_{0}(a')$, which gives the probability that a new recruit has ant state $a'$ upon its first appearance in the census ($\sum\rho_{0}(a')=1$). 
@@
 \paragraph{Reproduction}
-The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects size $x$ in year $t$ and random effects of plot and year.
+The flowering sub-model ($P(x;\pmb{\tau}^{P})$) estimates the probability of reproducing in year $t$, with fixed effects of size $x$ in year $t$ and random effects of plot and year.
 We fit this model to the reproductive status data (vegetative or flowering) using a Bernoulli distribution and a logit link function, similar to the survival model above but with no ant effects.  
@@
 We constructed two versions of the stochastic, multi-state IPMs that allowed us to test this hypothesis.
-The baseline, 'non-synchronous' model described above included ants effects that could vary uniquely across time, according to the parameter estimates for the random effects ($\pmb{\tau}_{a}$). 
+The baseline, ``non-synchronous'' model described above included ants effects that could vary uniquely across time, according to the parameter estimates for the random effects ($\pmb{\tau}_{a}$). 
 We quantified from the fitted random effects how tightly inter-annual variation was correlated between ant states for each vital rate.
-The 'synchronous' version included  ant effects that were forced to be the same across ant states. 
+The ``synchronous'' version included temporal fluctuations that were forced to be the same across ant states. 
 To synchronize ant states, we averaged the ant-specific year random effects, thus ensuring that plants in all ant states fluctuated synchronously in response to temporal environmental variation. 
@@
 	\caption{The probability of ant state based on size and previous ant state.
-Each panel shows size-dependent probabilities of the next ant state conditional on the previous ant state.
+Each panel shows size-dependent probabilities of the next ant state conditional on the previous ant state and size.
 Solid lines represent predictions of the multinomial statistical model and points show data binned over size intervals, where point size is proportional to the number of observations.}
@@
 
-We found no evidence that the portfolio effect generated positive effects of partner diversity, as effects of partner richness and composition were highly consistent between the baseline model and the model version that synchronized all ant states (Figure \ref{fig:LambdaMeans}).
+We found no evidence that the portfolio effect generated positive effects of partner diversity, as effects of partner richness and composition were highly consistent between the baseline, non-synchronous model and the model version that synchronized all ant states (Figure \ref{fig:LambdaMeans}).
 The effect of all ant partners can be measured as $\lambda^{3}_{S} - \lambda^{0}_{S}$.
@@
 Furthermore, evidence from nearly two decades of demographic data failed to support the ``portfolio effect'' hypothesis, whereby benefits of diversity manifest in the context of environmental fluctuations.
-Overall, our results indicate that while ant visitation is strongly consequential for plant fitness, partner identity and diversity are much less so.
+Overall, our results indicate that while ant visitation is strongly consequential, partner composition is much less so, suggesting that the fitness benefits of mutualism from the plant's perspective may be surprisingly resilient to declines in partner diversity.
 
@@
 Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population -- notably, weather -- override any fluctuations in herbivore composition and ant-herbivore interactions.
-Dallas et al. \citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is often very difficult to detect in empirical data across many systems. 
-This indicates that it may be difficult to detect, disguised by different mechanisms, or uncommon in nature.
+Dallas et al. \citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is infrequently detected in empirical data across many systems. 
+This indicates that the portfolio effect may be difficult to detect, disguised by different mechanisms, or uncommon in nature.
 
@@
 Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
-Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
+Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation \tom{(Figure \ref{app:herb})}{Note that this is referencing Figure 2, which is not correct.}, and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
 Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
@@
 \section*{Data and Code Availability}
-The data that support the findings of this study are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1
+Most of the data that support the findings of this study (through 2019) are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1.
+An updated data package, including recent years, is presently under review and we expect will be available in early 2025. 
+For now, all of our data and code are available in our public github repo: \url{https://github.com/acampbell110/ant_cactus_demography}.
 
@@
 This data has been taken consistently since 2017, so the analysis below considers 6 years of data.
-We considered only plants which were reproducing, as they were likely to produce the highest quantity of EFN \citep{Miller2014}.
+We considered only plants which were reproducing, since prior work shows that flowering significantly elevates herbivore pressure \citep{Miller2006}.
 The proportion of reproducing plants that experienced herbivory was calculated for each ant state separately.
@@
 
-\section*{Appendix D: Posterior Checks and Model Validation}\label{appendix:D}
-For each model fitted, we conducted two tests to determine if the fit was acceptable to use in our IPM. 
-First, we checked the convergence of each parameter and determined they were acceptable.
-Second, we checked the posterior fit, comparing the estimated values of each model to the $y$ values of the actual data.
-Each model was run with 10,000 total iterations per chain, 5,000 warmups, and thinning set to 2, leaving us with a total of 2,500 iterations post warmup.
-
-\subsection*{Statistical Models -- Results}
+\section*{}\label{appendix:D}
 Below are the results reported of all statistical models not described in the main body of the text.
@@
 \paragraph{Reproduction Model}
-The probability of a plant reproducing in a given year is highly size dependent. 
+\tom{The probability of a plant reproducing in a given year is highly size dependent. 
 The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
+}{What about the flowerbud model?
+I think there is value is showing the data and model fit (here in the appendix) for both Pf(flowering) and \# flowerbuds vs size.}
 
@@
 \paragraph{Seeds Per Flower Model}
-Each viable flower on a plant produces between 97 and 257 seeds.
+\tom{Each viable flower on a plant produces between 97 and 257 seeds.}{You had a figure showing mean seeds per fruit by ant group. I think you should keep that figure here.}
 This number is affected by the ant partner present, as shown in previous work \citep{Ohm2014}. 
@@
 Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
-We are confident that \textit{C. opuntiae} tended plants produce the fewest seeds per flowers.
+\tom{We are confident that \textit{C. opuntiae} tended plants produce the fewest seeds per flowers.}{Confident how? What is this based on?}
 
@@
 	\paragraph{Germination Model}
-	Seeds have a significantly higher probability of germinating in year one than in year two.
+	\tom{Seeds have a significantly higher probability of germinating in year one than in year two.}{Not sure why you say significantly higher.
+The posteriors you describe in the next sentence are completely overlapping.}
 	Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.

### 2025-01-01 Manuscript/ant_cactus_AmNat.tex
Final Updates

@@
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \tom{\citep{Barrett2015,Ushio2020}.}{Cite the Winfree chapter?}
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020,Winfree2020}.
 First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
@@
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \tom{\citep{Palmer2010,Afkhami2014,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}.
-}{The Afkhami paper cited here is not an ant-plant study.}
+Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}. 
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
@@
 This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
-Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD tree cholla per plot: \tom{}{Give the mean and SD of cactus density per plot (plots 1-8) here.}) has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
 For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
@@
 Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
-Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation \tom{(Figure \ref{app:herb})}{Note that this is referencing Figure 2, which is not correct.}, and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
+Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
 Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
@@
 
+\renewcommand{\thefigure}{A\arabic{figure}}
+
 \begin{figure}[H]
@@
 
+\renewcommand{\thefigure}{A\arabic{figure}}
+
 \begin{figure}
@@
 
-\section*{}\label{appendix:D}
+\section*{Appendix D}\label{appendix:D}
 Below are the results reported of all statistical models not described in the main body of the text.
@@
 \paragraph{Reproduction Model}
-\tom{The probability of a plant reproducing in a given year is highly size dependent. 
-The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes.
-}{What about the flowerbud model?
-I think there is value is showing the data and model fit (here in the appendix) for both Pf(flowering) and \# flowerbuds vs size.}
+
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/repro_panel.png}
+	\caption{The probability of a cholla plant reproducing based on its size.
+The green line shows the average probability, the gray surrounding area shows the 90\% confidence interval, and the black points show the actual data.
+}
+	\label{app:repro}
+\end{figure}
+
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/flow.png}
+	\caption{The number of flowers produced by a cholla plant based on its size.
+The green line shows the average number of flowers produced, the gray surrounding area shows the 90\% confidence interval, and the green points show the binned data.
+}
+	\label{app:flow}
+\end{figure}
+
+The probability of a plant reproducing in a given year is highly size dependent. 
+The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes (Figure \ref{app:repro}).
+
+The number of flowers produced by a reproducing plant is also highly size dependent. 
+The mean number of flowers produced are 0 until the plant reaches a medium size, after which the mean number of flowers produced increases to about 45 at the largest sizes (Figure \ref{app:flow}).
 
@@
 \paragraph{Seeds Per Flower Model}
-\tom{Each viable flower on a plant produces between 97 and 257 seeds.}{You had a figure showing mean seeds per fruit by ant group. I think you should keep that figure here.}
+
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
+	\caption{The mean number of seeds produced per fruit by the ant partner present.
+The teal,purple, and yellow lines show the mean seeds produced by each flower on plants tended by \textit{L.
+apiculatum} ants, \textit{C.
+opuntiae} ants, and no ants respectively.
+}
+	\label{app:seeds}
+\end{figure}
+
+Each viable flower on a plant produces between 97 and 257 seeds (Figure \ref{app:seeds}).
 This number is affected by the ant partner present, as shown in previous work \citep{Ohm2014}. 
@@
 Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
-\tom{We are confident that \textit{C. opuntiae} tended plants produce the fewest seeds per flowers.}{Confident how? What is this based on?}
 
@@
 	\paragraph{Germination Model}
-	\tom{Seeds have a significantly higher probability of germinating in year one than in year two.}{Not sure why you say significantly higher.
-The posteriors you describe in the next sentence are completely overlapping.}
+	Seeds have a slightly higher probability of germinating in year one than in year two.
 	Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.

### 2025-01-01 01_cholla_ant_IPM_setup.R

@@ -331,4 +331,13 @@ for(i in 1:nrow(cactus)){
     summarise(ants <- sum(antsum,na.rm=T)) -> sum_data
 nrow(subset(sum_data, sum_data$`ants <- sum(antsum, na.rm = T)`>1))/nrow(sum_data)
 
-
+# Count the number of plants per plot per year
+plant_counts <- cactus %>%
+  group_by(Plot, Year_t1) %>%
+  summarise(PlantCount = n(), .groups = "drop")
+summary_stats <- plant_counts %>%
+  group_by(Plot) %>%
+  summarise(
+    MeanPlantCount = mean(PlantCount),
+    SDPlantCount = sd(PlantCount)
+  )

### 2025-01-01 05_cholla_ant_IPM_figures_tables.R

@@ -635,7 +635,7 @@ flow_sim <- matrix(NA, n_draws,stan_data_flow_trunc$N)
 for(i in 1:n_draws){
   for(n in 1:stan_data_flow_trunc$N){
     flow_sim[i,n] <- sample(x=1:n_draws,size=1,replace=T,
-                            prob=dnbinom(1:n_draws, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i]) / (1 - dnbinom(0, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_surv$vol[n]), size=flow.params$phi[i])))
+                            prob=dnbinom(1:n_draws, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_flow_trunc$vol[n]), size=flow.params$phi[i]) / (1 - dnbinom(0, mu = exp(flow.params$beta0[i] + flow.params$beta1[i]*stan_data_flow_trunc$vol[n]), size=flow.params$phi[i])))
   }
 }
 ## Plot the posterior distributions

### 2025-06-18 Manuscript/Ant_Cactus_RESPONSE.tex
Manuscript Updates

% ======================================================= %
% Document: TEMPLATE FOR RESPONSES TO REVIEWERS
% Author: Andrea Ballatore
% Date: Jan 7, 2013
% Source: https://raw.githubusercontent.com/ucd-spatial/Datasets/master/tex_response_to_reviewers_template/responses_to_reviewers.tex
% Modified by Eduard Szöcs, 10.03.2015
% ======================================================= %
\documentclass[12pt]{article}

% packages
\usepackage{xr-hyper}
\usepackage{hyperref}
\externaldocument[ms-]{ant_cactus_SUBMISSION2}

\usepackage{graphicx}
\usepackage{url}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{color}
\definecolor{mygray}{gray}{0.6}
\usepackage[utf8]{inputenc}
\usepackage[onehalfspacing]{setspace}
\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	colon,	% (default) to separate multiple citations with colons;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they
]{natbib}
\usepackage[%disable
	]{todonotes}

\usepackage{anysize}
\marginsize{2.5cm}{2.5cm}{1.5cm}{2.5cm}

% macros
% add a counter
\newcounter{cN}
\setcounter{cN}{0}

\newcommand{\comment}[1]{
	\vspace{2em}
	\refstepcounter{cN} % incrment counter
	\noindent \hangindent=0em \textbf{\textcolor{Maroon}{\uline{Comment \thecN}:~}}\emph{``#1''}
	}

\newcommand{\response}[1]{
	\\[0.25em]
	\hangindent=2.3em \textbf{\textcolor{NavyBlue}{\uline{Response}:~}}#1
	}

\newcommand{\revise}[1]{{\color{Mahogany}{#1}}}

\usepackage[normalem]{ulem}
\definecolor{darkred}{rgb}{1,.6,.6}
\DeclareRobustCommand\problemline{\bgroup\markoverwith{\textcolor{darkred}{\rule[-0.9ex]{4pt}{3pt}}}\ULon}
\DeclareRobustCommand{\problem}[1]{\problemline{#1}} % soul
\setcounter{secnumdepth}{-1}

\begin{document}
% ======================================================= %
\title{Manuscript 62524 --- Response to reviewers}

\maketitle
% ======================================================= %
\noindent To the editorial board,

Thank you for the opportunity to submit a revision of our manuscript for your consideration. Our major changes include the following:
\begin{enumerate}
	\item We have added a new analysis that explores under what conditions portfolio effect could be shown. 
	We found *************
	\item We updated the introduction to be more accessible for readers and to better set up the rest of the manuscript.
\end{enumerate}

We describe these and other changes in greater detail below, where we reproduce comments from the associate editor and reviewers and provide our point-by-point responses. 
All of our changes are denoted in the manuscript with \revise{Mahogany font}.
We think the review process has greatly strengthened our manuscript such that it is now suitable for publication.
We hope you agree.

\vspace{2em}
\hfill On behalf of myself and T. Miller,

\hfill Ali Campbell
\newpage

% ======================================================= %

\section{Response to Dr. Megan E. Frederickson}
\vspace{-2em}

\comment{Both reviewers mentioned that the introduction does not quite set up the methods and results of the paper, and I agree, and I want to quickly expand on their comments. 
The introduction is mainly concerned with setting up the question of whether partner diversity in a mutualism can benefit a focal species (via sampling, complementarity, or portfolio effects). 
I agree that this is an interesting question, but I thought that the introduction undersells the results, because the demographic data and IPM offer more than an answer to this one question. 
I think the authors could ‘sell’ their demographic dataset more (just a single intro sentence says “the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available” at Line 101). 
I also thought they could explain in the introduction why even the quantification of the effects of single ant partners across the tree cholla lifecycle, and how ant species turn over on plants, are unique and valuable contributions to the field.}
\response{We have edited the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data3} and ~\ref{ms-res:long-term-data2}). 
We have also expanded on why the quantification of single ant partners is unique and how this is a valuable contribution to the field (line ~\ref{ms-res:L111})
}

\comment{Like the reviewers, especially Reviewer \#1, I could also have used a little more detail on some of the data collection methods. Reviewer \#1 notes that little information is provided on how seed production was estimated, and I agree. For my part, I could also have used a little more detail on how floral viability was measured. The methods say “we distinguished between flowerbuds that were viable versus aborted” but I wasn’t clear how.}
\response{We distinguished between viable and aborted flowerbuds through visual assessment. We have gone through the methods data collection section and added detail to our data collection descriptions which we hope will help readers better understand.}

\comment{I also encourage the authors to pay close attention to responding to several important points raised by Reviewer \#2, especially with regards to 1) what evidence for the portfolio effect would look like in a mutualism in which partners participate sequentially, and 2) providing the reader with a little more ‘hand holding’ when it comes to the rather complex model. Like the reviewer, I cannot claim that I would spot an error in the model if there were one, and it may be helpful to build up to the full Integral Projection Model in more stages, as Reviewer \#2 very helpfully suggests. }
\response{We agree with the assessments of Reviewer 2 and Dr. Frederickson. We have edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism.}

\comment{Line 184: Clarify whether the searches of these plots were exhaustive. Have you captured all plants within each plot in the dataset?}
\response{We have clarified this (line ~\ref{ms-res:exhaustive}).}

\comment{Lines 329 and 364: I realize that this is commonplace in some parts of the literature, but I would prefer not to call different simulation scenarios “experiments”}
\response{We have changed the language throughout to "simulations" rather than "experiments".}

\comment{Fig. 1: Is the gray dashed line in panel e) the 1:1 line that would indicate no growth? It is not labelled in the legend or the caption.}
\response{The grey dashed line does indicate no growth. We have updated the caption to indicate this.}

\comment{Figure 4: I wondered if taking the ant legend and inset correlation coefficients out of the plots and making them separate panels in a multi-panel figure, would increase the readability of this figure. It is very hard to see what is going on with the random effects across years because of the overlapping colours and the large y-axis scale, which creates space for the inset panels and legend, but squishes the coloured dots and lines together.}
\response{**** Honestly not sure, I would like to discuss with you Tom.}

\comment{There is very little mention of some of the supplemental results in the main text (e.g., the herbivory analysis). Please make sure all supplemental results are clearly described in the Results section of the main text.}
\response{**** Again, I would like to discuss where you think would be most appropriate}

\section{Response to Reviewer 1}
\vspace{-2em}

\comment{The study "Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism" uses a long-term dataset to look at the demography of a plant with mutualistic ants that provide defense to the plant. 
The estimated parameters and used a stochastic, multi-state IPM to look at the effects of ant mutualists on plant fitness. 
Most significantly, the authors found all ant species increased fitness (finite rate of growth), irrespective of their identity or turnover over the life of the plant. 
Overall, I found the study to be high-quality and novel, and the main finding is somewhat perplexing and would be of interest of the journal's readership. 
Admittedly, I found the framing of the introduction to not quite set up the study strongly and I was worried about the precision of the modeling, but the data and their implications are compelling and interesting. 
The level of precision of the modeling that concerned me was philosophical rather than practical. 
Specifically, I spent a few days thinking about this study with respect to Levins' (1966) strategies of model building in population biology. 
I originally thought that this population model was one that focused on *precision* and either generality or realism, but upon reflection I see this as one that sacrifices precision for generality and realism because the results and interpretation are largely focused on the qualitative outcome of the model, which are the useful and meaningful results. 
I think that it was really just the first paragraph of the introduction that was a sticking point for me given that mutualism "deterioration" and context dependency didn't seem to be important for the study.}
\response{We appreciate the positive response to our results and the concept of our manuscript. We have refined the introduction in response to your  }

\comment{Methodologically, the long-term data are incredible. 
What the authors were able to measure—especially for such a long period of time—greatly increase the strength of the study. 
Because of these methods for instance, the authors were able to observe many plant-switching events, and I was really surprised by just how often tending/occupancy switches on plants. 
How the data were collected on seeds was not clear to me, however. 
On L. 191 the paragraph describes using other study's data on seed production, but this wasn't related to this study's data collection, which is pretty important. 
I hope that data on seed production was measured on the same plants from which other demographic variables were measured. 
If not, was seed production just estimated based on flower viability?  
If that's the case, I think it would be important to state that; especially because an idea behind the multi-mutualist framework is that sometimes mutualist guilds can be antagonistic (e.g., defense guild decreasing pollination). 
The statistical modeling and IPM were explained clearly.}
\response{We have expanded on the source of our seed data to clear up any confusions as to their origins and the methods of obtainment (line~\ref{ms-res:data_origins}). }

\comment{I think the results of this study are compelling. 
Much thought (including older thought like seed-dispersal effectiveness (Schupp 1993)) on mutualism presumes interspecific variation within a guild. 
This is even true from what I know of the ant-plant-tending literature, where we observe differences in quality of protection among species. 
Ostensibly, this could be due to the typical short-term studies. 
Given the longevity and quality of data showing no differences among ant defenders, I think this could be an important contribution.}
\response{We are pleased you see the value and novelty of our results.}

\comment{L. 22: I feel like this paragraph doesn't add much and that if you nix everything after the first sentence and open with what's in paragraph 2, that would be more suitable for this study. 
It also frees some room to discuss generalized/diffuse mutualisms further.}
\response{Thank you for the helpful suggestion. 
We have streamlined the introduction paragraph by removing extraneous details after the first sentence and opened the section with the content previously in paragraph 2. 
This revision improves the focus of the introduction on the core aspects of our study. 
Additionally, we have expanded the discussion on generalized/diffuse mutualisms as you recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}

\comment{L. 29: I think that Chamberlain et al. didn't find that the interaction strength of mutualism wasn't more context dependent that competition. 
But overall, I interpreted their findings to be that other interactions are context dependent: not just mutualisms. 
Sometimes mutualism showed more CD than the others, but it's constantly discussed in mutualism, but not in the another two antagonistic interactions.}
\response{I removed this sentence based on their recommendations so I am not sure what I should say here... *********}

\comment{L. 53: Not certain I understand within a guild how benefits can vary by type.}
\response{We have revised the manuscript to explicitly explain that different partner species within the same guild can provide distinct benefits, not just differences in the magnitude of similar benefits (line~\ref{ms-res:L53}).}

\comment{L. 111-2: First time writing scientific names, please write out the full genus.}
\response{We have corrected this. (line~\ref{ms-res:L111})}

\comment{L 420-422: If the small-plant data are extrapolated, then I would not include that in the results.}
\response{Not sure how to respond to this given the need to include expected small plant reactions in the model despite slim probabilities of occurrence. ********}

\comment{L. 429: Is there a figure equivalent to figs. 1 and 2 that show seed production/set?}
\response{Yes, we have this figure in Appendix D (Figure ~\ref{ms-app:seeds}).}

\comment{Fig. 5, caption: ". . . based on [plant] size".}
\response{We have revised the caption of Figure 5 to clarify what is meant by "based on plant size" and to explicitly describe the relationship being visualized (Figure~\ref{ms-fig:Ant_Transition}).}

\section{Response to Reviewer 2}
\vspace{-2em}

\comment{I feel that the major research questions are poorly framed in the abstract and introduction. 
The problem is that the reader really needs to be primed to deal with the portfolio effect in sequential mutualisms with asynchronous environmental effects on mutualists. 
I was deep into the methods, which is quite dense, before I felt like I could understand the core research question.}
\response{We have expanded the Introduction and Abstract to more explicitly detail portfolio effects and sequential partner interactions.
These revisions better prime the reader to understand the core research question and the rationale behind our modeling approach, which integrates partner turnover and environmental stochasticity. We believe this strengthens the manuscript’s conceptual clarity and relevance.}

\comment{I would like to know what evidence for the portfolio effect would look like, so that I can carry that through the methods and see that this is what you are measuring, and into the results, so I can compare it to what you actually found.}
\response{We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).}

\comment{Further, following up on this, I'd like to see better discussion of what went wrong. 
Under what conditions, different from the ones in this study system, would portfolio effects have emerged? 
For example, I bet portfolios would be more important if asynchronous climate fluctuations were sufficient to convert a large fraction of trees from ant-bearing to vacant. 
Or, additionally, if the escape size of trees, where mortality drops off, were larger, such that both reproduction during bad years and survival advantages of ants were more important.  
Or, if an environmental condition that results in lower seed production promoted the spread of an ant that prevents seed loss (complementarity in sequence).
Can I see a positive control, so that I can (1) put the negative result into context, and not overgeneralize the result and (2) feel confident that the model could represent portfolio effects, so that their absence is meaningful.}
\response{*** Simulations \& updated discussions}

\comment{Abstract: What is the big research question? 
Which "demographic impacts of different ant partners" are of most interest w/r this question. 
What is meant by the "dynamics of turnover from one exclusive partner association to another?" Is there a simpler way of saying this?}
\response{We have revised the abstract to foreground the overarching question—how partner turnover influences host demography and fitness—and to clarify the specific demographic outcomes of interest. 
We also simplified the language around partner turnover.
We believe this revised version more clearly communicates the goals, approach, and key findings of the study.}

\comment{Overall comment: You have a complicated model. 
What training does the reader need to best make sense of it. 
I would argue: (1) portfolio effects as a mechanism of diversity-function relationships in mutualism as compared to (2) sampling effects (more species = more chance of having best mutualist) and associated (3) "missed opportunity costs", or suboptimal outcome from diversity compared to exclusive association with best mutualist. 
Make sure that the introduction of portfolio effects anticipates your hypothesis that they should be manifest only in the asynchronous ant model. 
This stuff is spread throughout more general material that I would cut or at least substantially reduce, so that the reader is ready for the methods.}
\response{**** Not sure if I have done this effectively.... But I edited mostly the paragraphs introducing the background and the mechanisms.
We have revised the Introduction to present the mechanisms more clearly, namely we tied the definition of the mechanism more closely to the mathematical definitions we use in the methods to show evidence of each mechanism (lines ~\ref{ms-res:mechanims_math} - ~\ref{ms-res:portfolio_asynchronous}), emphasizing that portfolio effects specifically depend on asynchronous partner responses to environmental variation (line~\ref{ms-res:portfolio_asynchronous}).
This framing provides clearer conceptual grounding for the models and we hope it better prepares readers for the methods that follow.}

\comment{A bit jarring out the gate. The first sentence states: everyone in mutualism benefits (full stop). The second sentence is : except when they don't.}
\response{We have revised the first sentence (line ~ \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits. We believe the language of net benefit allows for the further elaboration than the magnitude of benefits vary based on several conditions (aka. except when they don't).}

\comment{27: Compared to which interactions ("such as…")? I think this intro material obstructs you from rolling out the background we really need.}
\response{We have revised the first paragraph to reduce unnecessary and/or uninformative information to speed the reader into what is needed to understand the rest of the paper (line ~\ref{ms-res:L27}).}

\comment{39: "remain open questions within relevance in applied settings." 
What does this mean? 
What applied settings? 
How relevant?}
\response{We believe the questions we pose have relevance in regards to studying mutualisms under environmental variability, where different partners may confer benefits unevenly across space or time.
We have revised the introduction to better reflect these applied settings and further contextualize this study (line ~\ref{ms-res:L39}). }

\comment{45. Wordy sentences. Do we need both "hierarchy of fitness effects" and "a consistent ranking…".}
\response{We have edited to reflect this suggestion on line ~\ref{ms-res:L45}.}

\comment{63: Your paragraph doesn't deliver on its opening. 
Right out of this opening sentence I want to know: What is the effect of partner diversity when partners are present sequentially? 
How is this different from when partners are present simultaneously?}
\response{We have revised this paragraph to better answer your questions and deliver on its opening. 
We explain that in sequential associations, partner conflict is significantly reduced and associations are more stable -- leading to overall increased benefits for the focal mutualist (line ~\ref{ms-res:L63}).}

\comment{Some kind of brief lay-summary would go well here. 
Some kind of model overview that avoids too much jargon, ties into the introduction, and gives the reader who will not slog through all the simulation model details a way to engage in the model findings. 
I know the complex bits have to be there and that mature, quantitative readers exist, but this reviewer cleaves more to a lower common denominator}
\response{**** Not really sure where to put this or if I should instead modify existing text???}

\comment{Before you get into details, make sure I understand:
(1)     Plants die, or else survive, grow, flower, fruit, and add seeds to the recruitment bank, where new plants come from.
(2)     Growth, survival, fruiting and flowering are a product of plant size, whereas
(3)     Both size and ant species identity influence growth and survival, with ant species exclusively determining flower viability and number of seeds per fruit
(4)     Environments fluctuate stochastically in ways that affect plant vital rates, both directly by influencing flowering, growth, and survival, and indirectly by influencing colonization of vacant plants and/or the displacement of a resident ant species by a different ant species
(5)     These environmental fluctuations can either affect all ant species in the same way (the synchronous model:  what's good for one ant species is good for all ant species) or in different ways (the asynchronous model: what's good for one ant species may be bad for another).
(6)     An expectation from the portfolio effect is that in an asynchronous model, there ought to be a benefit to having a diversity of ant species. When one species does poorly, the other can "step-up" and fill in.
(7)     This model's outputs are used to see if this portfolio effect emerges. }
\response{Your understanding of our model and analysis is accurate.}

% ======================================================= %
\end{document}
% ======================================================= %

### 2025-06-18 Manuscript/Previous Version/ant_cactus.tex
Manuscript Updates

--- 2025-01-01 Manuscript/ant_cactus_AmNat.tex
@@
 \usepackage{fullpage}
-\usepackage[authoryear]{natbib}
+\usepackage[authoryear,sectionbib,sort]{natbib}
 \linespread{1.7}
@@
 
-\title{Thank you, next: demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism}
+\title{Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism}
 % title suggestion, trying to move away from focus on IPM
 
-\author{Alexandra Campbell$^{1,\dagger}$ \\ 
-	Tom E.X. Miller$^{1,\ast}$}
+%\author{Alexandra Campbell$^{1,\dagger}$ \\ 
+%	Tom E.X. Miller$^{1,\ast}$}
 
@@
 
-	\noindent{} 1. Program in Ecology and Evolutionary Biology, Department of BioSciences, Rice University, Houston, Texas 77005;
+%	\noindent{} 1. Program in Ecology and Evolutionary Biology, Department of BioSciences, Rice University, Houston, Texas 77005;
 
-	\noindent{} $\dagger$ e-mail: amc49@rice.edu\\
-	\noindent{} $\ast$ e-mail: tom.miller@rice.edu
+%	\noindent{} $\dagger$ e-mail: amc49@rice.edu\\
+%	\noindent{} $\ast$ e-mail: tom.miller@rice.edu
 
@@
 \section*{Data and Code Availability}
-Most of the data that support the findings of this study (through 2019) are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1.
-An updated data package, including recent years, is presently under review and we expect will be available in early 2025. 
-For now, all of our data and code are available in our public github repo: \url{https://github.com/acampbell110/ant_cactus_demography}.
+This section will be filled out once double blind reviews are complete. 
+%Most of the data that support the findings of this study (through 2019) are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1.
+%An updated data package, including recent years, is presently under review and we expect will be available in early 2025. 
+%For now, all of our data and code are available in our public github repo: \url{https://github.com/acampbell110/ant_cactus_demography}.
 

### 2025-06-18 Manuscript/References_Cacti.tex
Manuscript Updates

@@
 }
+
+@article{Loreau1999,
+	title = {Biodiversity and Ecosystem Functioning: Effects of Species Diversity on Productivity and Stability in Plant Communities},
+	author = {Loreau, Michel},
+	journal = {Proceedings of the Royal Society B: Biological Sciences},
+	volume = {264},
+	number = {1381},
+	pages = {127–130},
+	year = {1999},
+	doi = {10.1098/rspb.1999.0601}
+}
+@article{Tilman1994,
+	title = {Biodiversity and Stability in Grasslands},
+	author = {Tilman, David and Wedin, David and Knops, Johannes},
+	journal = {Nature},
+	volume = {367},
+	pages = {363–365},
+	year = {1994},
+	doi = {10.1038/367363a0}
+}
+@article{Hautier2014,
+	title = {Eutrophication weakens stabilizing effects of diversity in natural grasslands},
+	author = {Hautier, Yann and Isbell, Forest and Borer, Elizabeth T. and Seabloom, Eric W. and Harpole, William S. and Lind, Eric M. and MacDougall, Andrew S. and Stevens, Carly J. and Tilman, David},
+	journal = {Nature},
+	volume = {508},
+	pages = {521–525},
+	year = {2014},
+	doi = {10.1038/nature13014}
+}

### 2025-06-18 Manuscript/ant_cactus_SUBMISSION2.tex
Manuscript Updates

--- 2025-06-18 Manuscript/Previous Version/ant_cactus.tex
@@
 
-	\section*{Abstract}
+	\section*{Abstract} \label{res:abstract}
 % 200 words max
-Mutualisms commonly involve multiple partner species but the ecological consequences of partner diversity remain poorly understood.
-Partner diversity may benefit a focal mutualist through mechanisms that mirror positive effects of species diversity on ecosystem function, such as complementarity or portfolio effect. 
-We studied the fitness consequences of partner diversity and turnover in the multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
-We used long-term data and Bayesian hierarchical statistical models to determine the demographic impacts of different ant partners, and the dynamics of turnover from one exclusive partner association to another.
-We then constructed a stochastic, multi-state Integral Projection Model (IPM) with which we simulated tree cholla populations with different richness and composition of partners.
+The ecological consequences of sequential multi-species mutualisms, which involve multiple partners across time, with regards to partner diversity and turnover remain poorly understood.
+In such sequential mutualisms, partner identity can change over time due to environmental variability or competitive displacement. 
+This raises a key question: how does partner turnover influence host demography and long-term fitness?
+We studied the fitness consequences of partner diversity and turnover in the sequential multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
+We used along-term dataset on plant vital rates and partner identity and Bayesian hierarchical statistical models to quantify the demographic consequences of association with different ant partners and the frequency and direction of turnover between mutually exclusive partner associations across the plant life cycle.
+We then constructed a stochastic, multi-state Integral Projection Model (IPM) to simulate how different patterns of partner richness, composition, and sequential turnover affect population fitness under fluctuating environmental conditions.
 We found that, while ant partners had different impacts on plant vital rates (\textit{C.
@@
 \newpage{}
-\section*{Introduction}
-Mutualisms are species interactions where all participants receive net benefits, leading to higher individual fitness and increased population growth rates. 
-They are widespread species interactions but can deteriorate into commensalism or parasitism under conditions that elevate costs or dampen benefits \citep{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022,Bronstein1994,Chamberlain2014,Frederickson2013,Axelrod1981,Leigh2010}.
-Mutualisms are considered more context dependent than other species interactions \citep{Chamberlain2014,Frederickson2013}, meaning the magnitude and sign of interaction strength are often determined by environmental conditions and species' identities \citep{Noe1994,Leigh2010}.
+\section*{Introduction} \label{res:introduction}
+%% introduce multispecies mutualisms and sequential mutualisms with definitions
+Mutualisms, species interactions where all participants receive net benefits, are key components of ecological communities and can enhance the fitness and population growth rates of the partners involved.
+\linelabel{res:27}
+While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies networks in which a focal species interacts with multiple partners over space or time \citep{Bascompte2009, Palmer2010}.
+In such systems, partner diversity can arise either from simultaneous interactions or sequential ones, where individual hosts experience turnover in partner identity over time \citep{Djieto-Lordon2005, Ness2006}.
+\linelabel{res:L22}
+A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
+Understanding how this kind of sequential partner diversity affects host fitness is a pressing question in mutualism ecology—especially under environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
+\linelabel{res:L39}
 
-Mutualism is defined at the level of a species pair (+/+) but these interactions are embedded within multi-species communities, and growing evidence suggests that pairwise interactions are poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}.
-A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
-Whether and how partner diversity modifies the demographic effects of mutualistic interactions remain open questions within relevance in applied settings \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
+%% Look at mechanisms which show how diversity impacts demography
+In this study, we evaluate several mechanisms through which partner diversity might influence mutualistic function, drawing from broader biodiversity–ecosystem function theory \citep{Barrett2015,Ushio2020,Winfree2020}.
+First, sampling effect occurs in systems with a distinct hierarchy of fitness effects (a consistent ranking of best to worst mutualists), when diversity increases the likelihood of including a high-quality partner \citep{Frederickson2013,Batstone2018}.
+\linelabel{res:mechanims_math} \linelabel{res:L45}
+There is no benefit of diversity here \emph{per se}, rather, missed opportunity costs arise in systems with sequential partner associations, where less beneficial partners may preclude colonization by better mutualists \citep{Miller2007}.
+Second, complementarity occurs in systems where partners differ functionally, when diversity increases the net benefits received by focal mutualist beyond what a single partner species could offer \citep{Batstone2018}.
+Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}.
+Even within a single mutualist guild, the benefits conferred by alternative partner species can vary by method , efficiency, or timing of benefit conferral \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
+\linelabel{res:L53}
+Third, portfolio effects stabilize host fitness across variable environments when different mutualists respond asynchronously to environmental fluctuations \citep{Batstone2018, Loreau1999, Lazaro2022,Ollerton2006,Alarcon2008}.
+\linelabel{res:portfolio_asynchronous}
+In this context, the portfolio effect implies that inter-annual variation in partner identity—especially if driven by environmental conditions—could stabilize overall plant fitness across years.
+For example, if a high quality defensive partner is more likely to occur during dry years and a moderately defensive partner dominates in wet years, the shifting presence of these species could buffer the focal plant from herbivore damage under variable conditions.
+\linelabel{res:portfolio_theory}
+Critically, portfolio effects require asynchronous environmental responses among partners, such that declines in one partner's benefits are offset by others, stabilizing the focal mutualist's fitness over time \citep{Loreau1999, Tilman1994, Hautier2014}.
+This concept is well developed in biodiversity–ecosystem functioning literature, but has rarely been applied to mutualisms in which partner diversity is realized through turnover rather than coexistence.
+While portfolio effects have been well studied in plant and animal communities, their role in mutualisms—particularly where partner diversity is sequential (turnover) rather than simultaneous—is less explored \citep{Ushio2020}.
 
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020,Winfree2020}.
-First, when there is a hierarchy of fitness effects (a consistent ranking of best to worst mutualists), a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
-This can lead to an apparent benefit of diversity driven by a sampling effect \citep{Batstone2018} -- but there is no benefit of diversity \emph{per se}, only better and worse partners. 
-However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
-Second, even within a single mutualist guild, the benefits conferred by alternative partner species can vary in type and not just degree \citep{Stachowicz2005,Bronstein2006,Stanton2003}. 
-This can lead to a positive effect of partner diversity through complementarity of alternative functions \citep{Batstone2018}. 
-Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}.
-Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
-
-Partner diversity can have different effects on a focal mutualist depending on whether partners are present simultaneously or sequentially (partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014,Barrett2015,Ushio2020,Dattilo2014}.
-Sequential associations are likely when alternative partners engage in interference competition for access to a shared mutualist \citep{Kiers2003,Batstone2018,Tgaard2015,Wulff2008}. 
+%% Partner turnover
+Partner diversity can affect a focal mutualist differently depending on whether partners are present simultaneously or sequentially (i.e., via partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014, Barrett2015, Ushio2020, Dattilo2014}.
+In simultaneous associations partner interactions frequently lead to competition and interference \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}, reducing the net benefits received by the focal mutualist.
+\linelabel{res:L63}
+Sequential associations have been shown to reduce these competitive partner-partner interactions, thereby boosting the focal mutualists benefits \citep{Wolfin2023,Stachowicz2005_2} and leading to more stable interactions \cite{Johnson2019}.
 Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
-The frequency of partner turnover can impact the level of benefits received by the focal mutualist, particularly if the benefits continue to accumulate with successive turnover (e.g., when sequential partners provide complementary functions) or if they saturate over time \citep{Sachs2004,Fiala1994}.
-Directionality of turnover can also influence effects of partner diversity if partner identity changes consistently across ontogeny of a focal mutualist \citep{Fonseca2003,Noe1994,Dejean2008}.
+The timing, frequency, and directionality of partner turnover can influence the net fitness benefits to the host, particularly if turnover allows accumulation of complementary benefits or buffers environmental variability via portfolio effects \citep{Sachs2004,Fiala1994, Ushio2020}.
+If partner identity changes systematically across ontogeny, such as through succession of partners aligned with host life stages, this may further affect fitness outcomes \citep{Fonseca2003,Noe1994,Dejean2008}.
 For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of a diverse guild of defensive mutualists are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
@@
 
+%% Defensive ant-plant mutualisms
 Defensive ant-plant mutualisms -- where plants provide food and/or housing to ants that in turn defend them from enemies -- are widespread interactions that offer valuable model systems for the ecology and evolution of mutualism \citep{Bronstein1998, Bronstein2006}.
@@
 However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
+Specifically, sequential associations require much longer-term data in order to analyze due to an accumulation of partners across time rather than the ability to capture multiple at once.
+\linelabel{res:long-term-data}
 
+%% What this study set out to do
 This study examined the consequences of partner diversity in a food-for-protection mutualism between the tree cholla cactus (\textit{Cylindropuntia imbricata}), a long-lived EFN-bearing plant, and multiple species of ant partners.
@@
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{L.
-apiculatum} providing strong anti-herbivore defense and \textit{C.
-opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
-However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production), did not integrate the demographic effects of ant defense across the life cycle, and did not incorporate temporal stochasticity, all of which may be essential for understanding net fitness effects of a diverse partner guild in variable environments.
+Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+\linelabel{res:L111}
+Even the basic quantification of how different ant species affect host demography across the plant life cycle is rarely possible, particularly in long-lived perennials. 
+Most studies evaluate short-term fitness proxies like seed set or herbivory at a single life stage \citep{Palmer2010, Bruna2014, Trager2010}, while few trace demographic effects across multiple stages.
+Similarly, while turnover of mutualist partners is known to occur \citep{Ness2006, Djieto-Lordon2005}, long-term data documenting patterns and consequences of partner switching are uncommon, especially in field systems.
+\linelabel{res:long-term-data3}
+Thus, this study provides unique empirical insight into how mutualist quality and temporal partner dynamics shape host population trajectories.
 
+%% Models and questions
 Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover, and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
-We used this observational data set of plant demography and ant-plant associations, contextualized by previous ant exclusion experiments, to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
+The long-term nature of this observational dataset gives us the power to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
+\linelabel{res:long-term-data2}
 Specifically, we asked:
@@
 Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
-For all plants, in May or early June of each year we recorded plant survival since the last survey and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
+These censuses were exhaustive, tracking every single plant and searching the entire plot for any new ones. \linelabel{res:exhaustive}
+For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh left it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted. 
+We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (this differentiation was a visual assessment). 
 We recorded the ant species presence and identity (or vacancy if no ants present).
@@
 To complete the tree cholla life cycle, we used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
+These studies were conducted at the same sites using the same plants as our study. \linelabel{res:data_origins}
 Ohm et al.
@@
 
-\subsubsection*{Partner diversity simulation experiments}
-Using the fully parameterized multi-state IPM, we conducted simulation experiments to quantify how diversity and identity of ant partners influenced plant fitness. 
+\subsubsection*{Partner diversity simulations}
+Using the fully parameterized multi-state IPM, we conducted simulations to quantify how diversity and identity of ant partners influenced plant fitness. 
 From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created treatments corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
@@
 apiculatum}, and Other); and three partners (observed scenario of all ant states).
-These simulation experiments were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed.
+These simulations were made possible by extrapolating ant-specific demographic performance across the size distribution, even for combinations of size and ant occupancy that were rarely observed. 
 For example, the no-partner scenario modeled a hypothetically ant-free cactus population, even though no such population exists to our knowledge, by applying the statistical knowledge gleaned from vacant plants across the size distribution.
@@
 Small, non-reproductive plants are typically vacant because they do not produce extrafloral nectar, and once plants begin producing nectar they are nearly always ant-tended \citep{Miller2014}. 
-Our simulation experiments preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature.
+Our simulations preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature.
 
@@
 Points show observed size transitions, colored lines are the mean next size, and contours show probability density of the skewed normal growth model.
-	Dashed lines indicate extrapolation while solid lines indicate the range of observed data.}
+	Dashed lines indicate extrapolation while solid lines indicate the range of observed data.
+	They gray dashed line indicates no growth (areas above this line show growth while areas under it show shrinkage).}
 	\label{fig:Grow}
@@
 	\includegraphics[width=0.95\linewidth]{Figures/transition.png}
-	\caption{The probability of ant state based on size and previous ant state.
+	\caption{The probability of ant state based on plant size and previous ant state.
 Each panel shows size-dependent probabilities of the next ant state conditional on the previous ant state and size.

### 2025-06-18 01_cholla_ant_IPM_setup.R

@@ -338,6 +338,6 @@ plant_counts <- cactus %>%
 summary_stats <- plant_counts %>%
   group_by(Plot) %>%
   summarise(
-    MeanPlantCount = mean(PlantCount),
-    SDPlantCount = sd(PlantCount)
+    mean = mean(PlantCount),
+    sd = sd(PlantCount)
   )

### 2025-06-25 Manuscript/ant_cactus_SUBMISSION2.tex

@@
 We studied the fitness consequences of partner diversity and turnover in the sequential multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
-We used along-term dataset on plant vital rates and partner identity and Bayesian hierarchical statistical models to quantify the demographic consequences of association with different ant partners and the frequency and direction of turnover between mutually exclusive partner associations across the plant life cycle.
+We used a long-term dataset on plant vital rates and partner identity and Bayesian hierarchical statistical models to quantify the demographic consequences of association with different ant partners and the frequency and direction of turnover between mutually exclusive partner associations across the plant life cycle.
 We then constructed a stochastic, multi-state Integral Projection Model (IPM) to simulate how different patterns of partner richness, composition, and sequential turnover affect population fitness under fluctuating environmental conditions.
@@
 \linelabel{res:mechanims_math} \linelabel{res:L45}
-There is no benefit of diversity here \emph{per se}, rather, missed opportunity costs arise in systems with sequential partner associations, where less beneficial partners may preclude colonization by better mutualists \citep{Miller2007}.
+There is no benefit of diversity here \emph{per se}, rather, missed opportunity costs arise in systems with sequential partner associations where less beneficial partners may preclude colonization by better mutualists \citep{Miller2007}.
 Second, complementarity occurs in systems where partners differ functionally, when diversity increases the net benefits received by focal mutualist beyond what a single partner species could offer \citep{Batstone2018}.
-Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than the expected from the sum of complementary functions \citep{Afkhami2014}.
-Even within a single mutualist guild, the benefits conferred by alternative partner species can vary by method , efficiency, or timing of benefit conferral \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
+Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than expected from the sum of complementary functions \citep{Afkhami2014}.
+Even within a single mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing of benefit conferral \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
 \linelabel{res:L53}
@@
 Partner diversity can affect a focal mutualist differently depending on whether partners are present simultaneously or sequentially (i.e., via partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014, Barrett2015, Ushio2020, Dattilo2014}.
-In simultaneous associations partner interactions frequently lead to competition and interference \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}, reducing the net benefits received by the focal mutualist.
+In simultaneous associations, partner interactions frequently lead to competition and interference \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}, reducing the net benefits received by the focal mutualist.
 \linelabel{res:L63}
@@
 Defensive ant-plant mutualisms are commonly multi-species, where a guild of ant partner species share, and often compete for, a plant mutualist \citep{Bronstein1998, Beattie1985, Trager2010, Agrawal1998}.
-Ant partners can vary in their ability to deter herbivores \citep{Bruna2014}, and visitation by low quality ant partners can prevent visitation by higher quality partners, causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
+Ant partners can vary in their ability to deter herbivores \citep{Bruna2014} and visitation by low quality ant partners can prevent visitation by higher quality partners, causing a reduction in fitness through missed opportunity costs \citep{Fraser2001, Frederickson2005}.
 Susceptibility to herbivory can also vary significantly throughout the life stages of the plant \citep{Boege2005}, suggesting that the order and timing of successive partners is important to the fitness impacts of the combined partner guild \citep{Barton2010, Boege2005, Fonseca2003}.
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}. 
-However, little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
-Specifically, sequential associations require much longer-term data in order to analyze due to an accumulation of partners across time rather than the ability to capture multiple at once.
+Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
+Specifically, sequential associations require much longer-term data in order to analyze, due to an accumulation of partners across time rather than the ability to capture multiple at once.
 \linelabel{res:long-term-data}
@@
 Tree cholla are tended by two common ant partners and several additional rarer species, all of which collect EFN in exchange for defense against herbivores.
-These ant species locally co-occur but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
+These ant species locally co-occur, but individual plants are typically tended by only one species that patrols the plant around-the-clock and maintains control of the plant's nectar resources, usually for an entire growing season \citep{Ohm2014, Donald2022}.
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
@@
 %% Models and questions
-Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover, and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
-The long-term nature of this observational dataset gives us the power to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
+Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
+The long-term nature of this observational dataset gives us the power to investigate whether, and through which mechanism(s), partner diversity affects the fitness benefits of ant visitation.
 \linelabel{res:long-term-data2}
@@
 To answer these questions we used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
-We then used a stochastic, multi-state integral projection model (IPM) that combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
+We then used a stochastic, multi-state integral projection model (IPM) which combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
 

### 2025-07-01 Manuscript/Ant_Cactus_RESPONSE.tex
response edits

@@
 
+\comment{This paper uses 20 years of data on the demography of an ant-defended cactus to model the effects of ant partner diversity and turnover on plant vital rates.
+I really like what this paper is trying to do, and I see this paper as an important contribution to the field.
+There are very few long-term datasets on the demography of mutualisms, especially for ant-plant interactions.
+So, the demographic data alone are a valuable contribution to the literature.
+However, this paper also leverages this dataset to develop a model of how ant partner diversity influences plant vital rates and therefore fitness.
+\\
+\\	
+This is already a very good paper, but with a few revisions, I think it could be an exceptional one.}
+\response{We thank Dr.
+Frederickson for their positive comments and critical feedback on our data and manuscript.
+We strived, in his revision, to take the manuscript from ``very good'' to ``exceptional''.}
+
 \comment{Both reviewers mentioned that the introduction does not quite set up the methods and results of the paper, and I agree, and I want to quickly expand on their comments. 
@@
 I also thought they could explain in the introduction why even the quantification of the effects of single ant partners across the tree cholla lifecycle, and how ant species turn over on plants, are unique and valuable contributions to the field.}
-\response{We have edited the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data3} and ~\ref{ms-res:long-term-data2}). 
-We have also expanded on why the quantification of single ant partners is unique and how this is a valuable contribution to the field (line ~\ref{ms-res:L111})
+\response{These are great suggestions -- thank you.
+We have revised the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data3} and ~\ref{ms-res:long-term-data2}).
+We have also expanded on why the quantification of the complete demographic effects of even single ant partners is a unique and valuable contribution to the field (line ~\ref{ms-res:L111})
 }
@@
 The methods say “we distinguished between flowerbuds that were viable versus aborted” but I wasn’t clear how.}
-\response{We distinguished between viable and aborted flowerbuds through visual assessment.
-We have gone through the methods data collection section and added detail to our data collection descriptions which we hope will help readers better understand.}
+\response{Thank you for these suggestions. 
+	We agree that methods for estimating seed production were a little under-developed in our previous submission because we relied too heavily on citations of our prior work. 
+	We now describe these methods more thoroughly (lines \ref{}). 
+	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{}). 
+	We have gone through the data collection methods section and added greater detail throughout, which we hope will make our work more understandable and reproducible.}
 
@@
 }
-\response{We agree with the assessments of Reviewer 2 and Dr.
-Frederickson.
-We have edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism.}
+\response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
+	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines XX and Figure XX).
+	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{}). 
+	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{}). }
 
 \comment{Line 184: Clarify whether the searches of these plots were exhaustive. Have you captured all plants within each plot in the dataset?}
-\response{We have clarified this (line ~\ref{ms-res:exhaustive}).}
+\response{We follow a consistent search protocol, so while it's possible we miss some recruits we think this source of observation error is small and roughly consistent across years.
+We have clarified this (line ~\ref{ms-res:exhaustive}).}
 
@@
 \comment{Fig. 1: Is the gray dashed line in panel e) the 1:1 line that would indicate no growth? It is not labelled in the legend or the caption.}
-\response{The grey dashed line does indicate no growth. We have updated the caption to indicate this.}
+\response{Good catch. Yes, the gray dashed line indicates no change in size. We have updated the caption to indicate this.}
 
@@
 It is very hard to see what is going on with the random effects across years because of the overlapping colours and the large y-axis scale, which creates space for the inset panels and legend, but squishes the coloured dots and lines together.}
-\response{**** Honestly not sure, I would like to discuss with you Tom.}
+\response{We agree that this figure was hard to interpret and tried to do too much in too little space. We have re-made this figure following these very good suggestions.}
 
@@
 Please make sure all supplemental results are clearly described in the Results section of the main text.}
-\response{**** Again, I would like to discuss where you think would be most appropriate}
+\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{}. \ref{}).}
 
@@
 I think that it was really just the first paragraph of the introduction that was a sticking point for me given that mutualism "deterioration" and context dependency didn't seem to be important for the study.}
-\response{We appreciate the positive response to our results and the concept of our manuscript. We have refined the introduction in response to your  }
+\response{We appreciate the reviewer's positive comments about our manuscript and their thoughtful reflection of how our work relates to broader philosophical considerations of ecological modeling. 
+	We agree with the Reviewer's conclusion that, in the context of Levins' triangle, our work leans toward generality and realism at the expense of precision. 
+That is, we have built a general model of demographic transitions (births, deaths, growth, and partner turnover) that is realistic with respect to our study system, but we query the model for qualitative inference (e.g., asking whether a multi-partner scenario yields higher or lower fitness than the single best partner) and not quantitative precision.
+We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}).}
 
@@
 The statistical modeling and IPM were explained clearly.}
-\response{We have expanded on the source of our seed data to clear up any confusions as to their origins and the methods of obtainment (line~\ref{ms-res:data_origins}). }
+\response{We appreciate the positive comments about our long-term data set. 
+	As described above, we now provide more detailed information about methods for estimating seed production.
+	Because these data required time-consuming fruit dissections, we were not able to collect seed production data as part of the long-term study. 
+	Instead, these data come from plants in different ant states during a single year. 
+	We have expanded on the source of our seed data to make all of this clearer (line~\ref{ms-res:data_origins}). 
+	The seed data do support the idea of defense decreasing pollination, and this is accounted for in our modeling framework. 
+	Because we lack long-term data on seed production, we must assume that this cost of defense is temporally static; we now state this assumption explicitly (line \ref{}).}
 
@@
 Given the longevity and quality of data showing no differences among ant defenders, I think this could be an important contribution.}
-\response{We are pleased you see the value and novelty of our results.}
+\response{We are pleased that the reviewer recognizes value and novelty in our results.}
 
@@
 This revision improves the focus of the introduction on the core aspects of our study. 
-Additionally, we have expanded the discussion on generalized/diffuse mutualisms as you recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}
+Additionally, we have expanded the discussion on generalized/diffuse mutualisms, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}
 
@@
 Sometimes mutualism showed more CD than the others, but it's constantly discussed in mutualism, but not in the another two antagonistic interactions.}
-\response{I removed this sentence based on their recommendations so I am not sure what I should say here... *********}
+\response{I removed this sentence based on their recommendations so I am not sure what I should say here...
+********* Tom: I will check the manuscript text, but probably simplest to change the way this paper gets cited so that we are not asserting that mutualism is more context-dependent.}
 
 \comment{L. 53: Not certain I understand within a guild how benefits can vary by type.}
-\response{We have revised the manuscript to explicitly explain that different partner species within the same guild can provide distinct benefits, not just differences in the magnitude of similar benefits (line~\ref{ms-res:L53}).}
+\response{We intended that partners within a guild can have qualitatively different effects on a shared mutualist (e.g., increasing growth vs.
+seed set) rather than simply different degrees of the same effect (increasing growth a little or a lot).
+We have revised this sentence so that our meaning is clearer (line~\ref{ms-res:L53}).}
 
 \comment{L. 111-2: First time writing scientific names, please write out the full genus.}
-\response{We have corrected this. (line~\ref{ms-res:L111})}
+\response{Thank you, we have corrected this. (line~\ref{ms-res:L111})}
 
 \comment{L 420-422: If the small-plant data are extrapolated, then I would not include that in the results.}
-\response{Not sure how to respond to this given the need to include expected small plant reactions in the model despite slim probabilities of occurrence. ********}
+\response{The simplest response we can offer here is that the small plant data are \emph{effectively} not extrapolated, so the expected performance of tiny ant-tended plants has little bearing on our results.
+	The more complete, technical response is that some amount of extrapolation is an inevitable feature of the Integral Projection Model framework. 
+	Because the state space is defined by size and ant status, we must tell the model what to expect in all combinations of those states. 
+	However, because state transitions are guided by empirical data, the model ``knows'' which state combinations are likely and unlikely. 
+	Tiny, ant-tended plants do exist in the model as a possible state combination, because they must -- but these are at such low frequency (as informed by the data) that the model is effectively insensitive to what we assume about their demographic performance.
+	We have tried to better communicate this nuance in our presentation of the model (line \ref{}).}
 
 \comment{L. 429: Is there a figure equivalent to figs. 1 and 2 that show seed production/set?}
-\response{Yes, we have this figure in Appendix D (Figure ~\ref{ms-app:seeds}).}
+\response{Yes, we have this figure in Appendix D (Figure ~\ref{ms-app:seeds}). 
+	We could have done a better job in our first submission of linking the main text and the supporting material. 
+	Hopefully these connections are clearer now (lines \ref{}).}
 
@@
 
-\comment{I feel that the major research questions are poorly framed in the abstract and introduction. 
+\comment{The authors' used a time-series dataset on defensive ant-plant mutualism to parameterize a simulation model.
+They used this model to determine the impact of ant species identity and diversity on the performance of the plant host under variable environments.
+They found (1) that ants provided benefits to their hosts; (2) ant species differed in the benefits they provide, but (3) there is no evidence that diversity or even ant identity mattered, even when environmental fluctuations had ant-species specific effects.
+\\
+\\	
+This is an excellent study relevant to anyone who studies partner diversity on the x-axis and host response on the y.
+The authors do not conclude that portfolio effects do not exist or that mutualist identity generally doesn't matter.
+But they do present compelling evidence that, in a system where it seems like mutualist identity and diversity ought to matter, they may not.
+They suggest that portfolio effects, even when eminently plausible, should not be assumed.}
+\response{We are pleased that this reviewer appreciated the novelty and significance of our findings, and we are grateful for their constructive feedback below.}
+
+\comment{But it's not perfect. Revisions should focus on presentation, starting with the introduction.
+\\
+\\
+I feel that the major research questions are poorly framed in the abstract and introduction. 
 The problem is that the reader really needs to be primed to deal with the portfolio effect in sequential mutualisms with asynchronous environmental effects on mutualists. 
 I was deep into the methods, which is quite dense, before I felt like I could understand the core research question.}
-\response{We have expanded the Introduction and Abstract to more explicitly detail portfolio effects and sequential partner interactions.
+\response{This is really helpful feedback. 
+	We have revised the Introduction and Abstract to more explicitly detail portfolio effects and sequential partner interactions (lines \ref{}, \ref{}).
 	These revisions better prime the reader to understand the core research question and the rationale behind our modeling approach, which integrates partner turnover and environmental stochasticity. 
@@
 \comment{I would like to know what evidence for the portfolio effect would look like, so that I can carry that through the methods and see that this is what you are measuring, and into the results, so I can compare it to what you actually found.}
-\response{We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
-Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).}
+\response{Great suggestion. 
+	We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
+	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory})[TOM: let's work on this, it's a little clunky.].
+	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{}). }
 
@@
 Can I see a positive control, so that I can (1) put the negative result into context, and not overgeneralize the result and (2) feel confident that the model could represent portfolio effects, so that their absence is meaningful.}
-\response{*** Simulations \& updated discussions}
+\response{Thank you for this intriguing suggestion, which got us thinking about our results in a way we hadn't before. 
+In response to this suggestion, we have added several new analyses that provide a ``positive control'', demonstrating the conditions under which our model reveals a portfolio effect.
+[more details]}
 
@@
 What is meant by the "dynamics of turnover from one exclusive partner association to another?" Is there a simpler way of saying this?}
-\response{We have revised the abstract to foreground the overarching question—how partner turnover influences host demography and fitness—and to clarify the specific demographic outcomes of interest. 
+\response{Thank you for pushing us to make the Abstract more direct and effective. 
+	Abstracts tend to come together late in the manuscript preparation process, but this is a healthy reminder that they require as much effort and attention as all the rest. 
+	We have re-written the abstract to clarify the overarching question—how partner turnover influences host demography and fitness[TOM: is this actually the question???
+let's work on this]—and to clarify the specific demographic outcomes of interest.
 We also simplified the language around partner turnover.
@@
 This stuff is spread throughout more general material that I would cut or at least substantially reduce, so that the reader is ready for the methods.}
-\response{**** Not sure if I have done this effectively.... But I edited mostly the paragraphs introducing the background and the mechanisms.
+\response{Thank you for pushing us to improve the communication of our modeling work. 
 We have revised the Introduction to present the mechanisms more clearly, namely we tied the definition of the mechanism more closely to the mathematical definitions we use in the methods to show evidence of each mechanism (lines ~\ref{ms-res:mechanims_math} - ~\ref{ms-res:portfolio_asynchronous}), emphasizing that portfolio effects specifically depend on asynchronous partner responses to environmental variation (line~\ref{ms-res:portfolio_asynchronous}).
@@
 We believe the language of net benefit allows for the further elaboration than the magnitude of benefits vary based on several conditions (aka.
-except when they don't).}
+except when they don't)[TOM: I think we can make this revision and response smoother].}
 
@@
 How relevant?}
-\response{We believe the questions we pose have relevance in regards to studying mutualisms under environmental variability, where different partners may confer benefits unevenly across space or time.
-We have revised the introduction to better reflect these applied settings and further contextualize this study (line ~\ref{ms-res:L39}). }
+\response{We have edited this sentence to be more specific about ``applied settings'' we referenced previously. 
+For example, prior work has examine how diversity of pollinator assemblages affects productivity of crop plants (see line \ref{}). }
 
 \comment{45. Wordy sentences. Do we need both "hierarchy of fitness effects" and "a consistent ranking…".}
-\response{We have edited to reflect this suggestion on line ~\ref{ms-res:L45}.}
+\response{We have edited to reflect this suggestion (line ~\ref{ms-res:L45}).}
 
@@
 \response{We have revised this paragraph to better answer your questions and deliver on its opening. 
-We explain that in sequential associations, partner conflict is significantly reduced and associations are more stable -- leading to overall increased benefits for the focal mutualist (line ~\ref{ms-res:L63}).}
+We explain that in sequential associations, partner conflict is significantly reduced and associations are more stable [TOM: not sure I agree] -- leading to overall increased benefits for the focal mutualist (line ~\ref{ms-res:L63}).}
 
@@
 I know the complex bits have to be there and that mature, quantitative readers exist, but this reviewer cleaves more to a lower common denominator}
-\response{**** Not really sure where to put this or if I should instead modify existing text???}
+\response{We appreciate this comment and we have taken inspiration from the reviewer's next comment to add the lay summary suggested here.}
 
@@
 (7)     This model's outputs are used to see if this portfolio effect emerges. }
-\response{Your understanding of our model and analysis is accurate.}
+\response{Your understanding of our model and analysis is accurate. 
+	In fact, this is such a good overview that it inspired a new section we added (lines \ref{}) to introduce the model, so that readers can first understand the ``story'' of the model before seeing any equations.}
 

### 2025-07-11 Manuscript/ant_cactus_SUBMISSION2.tex
abstract edits

@@
 \newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}  
+\newcommand{\revise}[1]{{\color{olive}{#1}}}
 \usepackage{hyperref}
@@
 % 200 words max
-The ecological consequences of sequential multi-species mutualisms, which involve multiple partners across time, with regards to partner diversity and turnover remain poorly understood.
-In such sequential mutualisms, partner identity can change over time due to environmental variability or competitive displacement. 
-This raises a key question: how does partner turnover influence host demography and long-term fitness?
-We studied the fitness consequences of partner diversity and turnover in the sequential multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
-We used a long-term dataset on plant vital rates and partner identity and Bayesian hierarchical statistical models to quantify the demographic consequences of association with different ant partners and the frequency and direction of turnover between mutually exclusive partner associations across the plant life cycle.
-We then constructed a stochastic, multi-state Integral Projection Model (IPM) to simulate how different patterns of partner richness, composition, and sequential turnover affect population fitness under fluctuating environmental conditions.
+Mutualisms commonly involve multiple partner species but the ecological consequences of partner diversity remain poorly understood.
+Partner diversity may benefit a focal mutualist through mechanisms that mirror positive effects of species diversity on ecosystem function, such as sampling effect, complementarity or portfolio effect.
+\revise{In sequential mutualisms, partner identity can change over time due to environmental variability or competitive displacement. 
+Here we asked: how do partner diversity and turnover influence host demography and long-term fitness, and which mechanisms explain these effects?}
+We studied the sequential multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
+We used a 20-year dataset on plant vital rates and partner identity to quantify the demographic consequences of association with different ant partners and the frequency and direction of turnover between mutually exclusive partner associations across the plant life cycle.
+We then constructed a stochastic, multi-state Integral Projection Model (IPM) to investigate how different patterns of partner richness, composition, and sequential turnover affect population fitness under fluctuating environmental conditions.
 We found that, while ant partners had different impacts on plant vital rates (\textit{C.
 opuntiae}-tended plants had advantages in growth and survival when small, and \textit{L.
-apiculatum}-tended plants had floral viability advantages), plant fitness was effectively insensitive to partner identity and number, because effects on the highest-sensitivity vital rates were consistent across partner species.
+apiculatum}-tended plants had floral viability advantages), plant fitness was similarly elevated by ant visitation for any level of partner identity and richness, because effects on the highest-sensitivity vital rates were consistent across partner species.
 Furthermore, there was no evidence that benefits of partner diversity manifest in the context of fluctuating environments (i.e., the portfolio effect). 

### 2025-07-14 Manuscript/ant_cactus_SUBMISSION2.tex
Intro edits

@@
 \newcommand{\ali}[2]{{\color{blue}{#1}}\footnote{\textit{\color{blue}{#2}}}}  
-\newcommand{\revise}[1]{{\color{olive}{#1}}}
+\newcommand{\revise}[1]{{\color{blue}{#1}}}
 \usepackage{hyperref}
@@
 Furthermore, there was no evidence that benefits of partner diversity manifest in the context of fluctuating environments (i.e., the portfolio effect). 
-In contrast to much of the previous literature, this study highlights that demographic benefits of mutualism can be surprisingly robust to diversity and turnover of partners.
+In contrast to much of the previous literature, this study highlights that demographic benefits of mutualism can be surprisingly robust to diversity and turnover of partner species.
 
@@
 %% introduce multispecies mutualisms and sequential mutualisms with definitions
-Mutualisms, species interactions where all participants receive net benefits, are key components of ecological communities and can enhance the fitness and population growth rates of the partners involved.
+\revise{Mutualisms, species interactions where all participants receive net benefits, are key components of ecological communities and can enhance the fitness and population growth rates of the partners involved.
 \linelabel{res:27}
-While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies networks in which a focal species interacts with multiple partners over space or time \citep{Bascompte2009, Palmer2010}.
-In such systems, partner diversity can arise either from simultaneous interactions or sequential ones, where individual hosts experience turnover in partner identity over time \citep{Djieto-Lordon2005, Ness2006}.
+While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal species interacts with multiple partner species.
+Growing evidence suggests that pairwise interactions may be poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}. 
+A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
+Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}.
+Partner diversity can arise either from simultaneous interactions or sequential ones, where individual hosts experience turnover in partner identity over time \citep{Djieto-Lordon2005, Ness2006}.
 \linelabel{res:L22}
-A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
-Understanding how this kind of sequential partner diversity affects host fitness is a pressing question in mutualism ecology—especially under environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
-\linelabel{res:L39}
+Understanding how partner diversity and turnover affect the overall fitness benefits of mutualism is an open question -— especially in the context of environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
+\linelabel{res:L39}}
 
 %% Look at mechanisms which show how diversity impacts demography
-In this study, we evaluate several mechanisms through which partner diversity might influence mutualistic function, drawing from broader biodiversity–ecosystem function theory \citep{Barrett2015,Ushio2020,Winfree2020}.
-First, sampling effect occurs in systems with a distinct hierarchy of fitness effects (a consistent ranking of best to worst mutualists), when diversity increases the likelihood of including a high-quality partner \citep{Frederickson2013,Batstone2018}.
-\linelabel{res:mechanims_math} \linelabel{res:L45}
-There is no benefit of diversity here \emph{per se}, rather, missed opportunity costs arise in systems with sequential partner associations where less beneficial partners may preclude colonization by better mutualists \citep{Miller2007}.
-Second, complementarity occurs in systems where partners differ functionally, when diversity increases the net benefits received by focal mutualist beyond what a single partner species could offer \citep{Batstone2018}.
+There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020,Winfree2020}.
+First, when there is a consistent ranking of best to worst mutualists, a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
+This can lead to an apparent benefit of diversity driven by a sampling effect \citep{Batstone2018}.  
+However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
+\revise{Second, even within a mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
+\linelabel{res:L53}\tom{}{There are two different Stachowicz references for the same paper -- be sure to clean this up.}
+This can lead to a positive effect of partner diversity through complementarity, whereby diversity increases the net benefits received by focal mutualist beyond what a single partner species could offer \citep{Batstone2018}.
 Positive diversity effects through complementarity need not be additive: interference or synergies between partners can make their combined effect different than expected from the sum of complementary functions \citep{Afkhami2014}.
-Even within a single mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing of benefit conferral \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
-\linelabel{res:L53}
-Third, portfolio effects stabilize host fitness across variable environments when different mutualists respond asynchronously to environmental fluctuations \citep{Batstone2018, Loreau1999, Lazaro2022,Ollerton2006,Alarcon2008}.
+Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
+Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
+In the context of temporally fluctuating environments, the portfolio effect implies that inter-annual variation in partner identity could stabilize overall plant fitness across years.
 \linelabel{res:portfolio_asynchronous}
-In this context, the portfolio effect implies that inter-annual variation in partner identity—especially if driven by environmental conditions—could stabilize overall plant fitness across years.
-For example, if a high quality defensive partner is more likely to occur during dry years and a moderately defensive partner dominates in wet years, the shifting presence of these species could buffer the focal plant from herbivore damage under variable conditions.
-\linelabel{res:portfolio_theory}
-Critically, portfolio effects require asynchronous environmental responses among partners, such that declines in one partner's benefits are offset by others, stabilizing the focal mutualist's fitness over time \citep{Loreau1999, Tilman1994, Hautier2014}.
-This concept is well developed in biodiversity–ecosystem functioning literature, but has rarely been applied to mutualisms in which partner diversity is realized through turnover rather than coexistence.
-While portfolio effects have been well studied in plant and animal communities, their role in mutualisms—particularly where partner diversity is sequential (turnover) rather than simultaneous—is less explored \citep{Ushio2020}.
+Critically, portfolio effects require asynchronous fluctuations among partners, such that declines in one partner's benefits are offset by others, stabilizing the focal mutualist's fitness over time \citep{Loreau1999, Tilman1994, Hautier2014}.
+%This concept is well developed in biodiversity–ecosystem functioning literature, but has rarely been applied to mutualisms in which partner diversity is realized through turnover rather than coexistence.
+While portfolio effects have been well studied in plant and animal communities, their role in mutualisms —- particularly where partner diversity is sequential rather than simultaneous —- is less explored \citep{Ushio2020}.}
 
 %% Partner turnover
-Partner diversity can affect a focal mutualist differently depending on whether partners are present simultaneously or sequentially (i.e., via partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014, Barrett2015, Ushio2020, Dattilo2014}.
-In simultaneous associations, partner interactions frequently lead to competition and interference \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}, reducing the net benefits received by the focal mutualist.
+%Partner diversity can affect a focal mutualist differently depending on whether partners are present simultaneously or sequentially (i.e., via partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014, Barrett2015, Ushio2020, Dattilo2014}.
+%In simultaneous associations, interference between partners may reduce the net benefits received by a focal mutualist \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}. \linelabel{res:L63}
+%Exclusive, sequential associations can reduce antagonistic partner-partner interactions \citep{Wolfin2023,Stachowicz2005_2} and enhance mutualism stability \cite{Johnson2019}.
+
+%%What if we skip the contrast of simultaneous vs sequential, since it is really the sequential ideas that set up our study?
+\revise{For exclusive, sequential associations, the timing, frequency, and directionality of partner turnover can influence the net fitness benefits to a focal mutualist, particularly if turnover allows accumulation of complementary benefits or buffers environmental variability via portfolio effects \citep{Sachs2004,Fiala1994, Ushio2020}.
 \linelabel{res:L63}
-Sequential associations have been shown to reduce these competitive partner-partner interactions, thereby boosting the focal mutualists benefits \citep{Wolfin2023,Stachowicz2005_2} and leading to more stable interactions \cite{Johnson2019}.
-Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}. 
-The timing, frequency, and directionality of partner turnover can influence the net fitness benefits to the host, particularly if turnover allows accumulation of complementary benefits or buffers environmental variability via portfolio effects \citep{Sachs2004,Fiala1994, Ushio2020}.
-If partner identity changes systematically across ontogeny, such as through succession of partners aligned with host life stages, this may further affect fitness outcomes \citep{Fonseca2003,Noe1994,Dejean2008}.
+\tom{Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}.}{any more recent examples we could cite here?} 
+If partner identity changes systematically across ontogeny, such as through succession of partners aligned with host life stages, this may further affect fitness outcomes associated with partner diversity \citep{Fonseca2003,Noe1994,Dejean2008}.
 For example, plant susceptibility to enemies can change across life stages \citep{Boege2005,Barton2010}, so the benefits of a diverse guild of defensive mutualists are greatest when more defensive partner species align with more vulnerable life stages \citep{Djieto-Lordon2005,Dejean2008}.
-Thus, the duration and order of interactions may be just as important as the identities of the interacting partners.
+Thus, the duration and order of interactions may be just as important as the identities of the interacting partners.}
 
@@
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available.
-Specifically, sequential associations require much longer-term data in order to analyze, due to an accumulation of partners across time rather than the ability to capture multiple at once.
-\linelabel{res:long-term-data}
+\revise{Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because long-term demographic studies on ant-plant interactions are \tom{exceedingly rare}{Any other examples we can cite here?
+How long was the Palmer Acacia study?}.
+
+Long-term demographic data are essential to understanding the fitness consequences of multi-species mutualisms for several reasons. \linelabel{res:long-term-data}
+First, the fitness consequences of sequential partnering play out over long time scales due to the accumulation of partners across time rather than the effects of multiple partners at once. 
+While turnover of mutualist partners is known to occur \citep{Ness2006, Djieto-Lordon2005,Fonseca2003}, long-term data documenting patterns and consequences of partner switching are uncommon, especially in field settings \citep{Palmer2010}.
+Second, certain effects of partner diversity are fluctuation-dependent (i.e., portfolio effect) and would only be detectable in the context of temporally stochastic environments in which alternative partners may fluctuate asynchronously.
+Quantifying contributions from the portfolio effect therefore requires long-term data on partner (a)synchrony across inter-annual fluctuations. 
+Finally, alternative partners may have different demographic effects (e.g., one that enhances survival and another that enhances reproduction) that contribute differently to overall fitness of a focal mutualist.
+\tom{Most studies evaluate short-term fitness proxies like seed set or herbivory at a single life stage \citep{Palmer2010, Bruna2014, Trager2010}, while few trace demographic effects across multiple stages.}{I like this sentence but it is unclear whether you are using the citations as examples of studies that only consider single life stages, or the opposite.
+This would be good to clarify, since you don't want to mis-cite people.}
+Integrating long-term data from across the life cycle within a stochastic demographic modeling framework therefore provides a powerful opportunity to diagnose the fitness consequences of partner diversity and turnover.
+}
 
@@
 \linelabel{res:L111}
-Even the basic quantification of how different ant species affect host demography across the plant life cycle is rarely possible, particularly in long-lived perennials. 
-Most studies evaluate short-term fitness proxies like seed set or herbivory at a single life stage \citep{Palmer2010, Bruna2014, Trager2010}, while few trace demographic effects across multiple stages.
-Similarly, while turnover of mutualist partners is known to occur \citep{Ness2006, Djieto-Lordon2005}, long-term data documenting patterns and consequences of partner switching are uncommon, especially in field systems.
-\linelabel{res:long-term-data3}
-Thus, this study provides unique empirical insight into how mutualist quality and temporal partner dynamics shape host population trajectories.
+However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production), did not integrate the demographic effects of ant defense across the life cycle, and did not incorporate temporal stochasticity -- gaps we aimed to fill in the present study.
 
 %% Models and questions
-Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
-The long-term nature of this observational dataset gives us the power to investigate whether, and through which mechanism(s), partner diversity affects the fitness benefits of ant visitation.
+Here we used a unique long-term data set that allows us to explore mutualistic associations with multiple partner species, longitudinal dynamics of partner turnover, and demographic effects of alternative partners that vary across plant size structure and across nearly 20 years of inter-annual fluctuations.
+The long-term nature of this observational dataset gives us the power to investigate whether and through which mechanism(s) partner diversity affects the fitness benefits of ant visitation.
 \linelabel{res:long-term-data2}
@@
 To answer these questions we used a hierarchical Bayesian statistical approach to estimate demographic vital rates for hosts in different states of ant occupancy and to quantify state-dependent partner turnover from the long-term data.
-We then used a stochastic, multi-state integral projection model (IPM) which combines diverse effects on vital rates and pathways of partner turnover to quantify effects of partner diversity on plant fitness.
+\revise{We then built a stochastic, multi-state integral projection model (IPM) which combines diverse effects on vital rates and pathways of partner turnover. 
+Finally, we used the model to conduct simulation `experiments' that quantify plant fitness under different scenarios of partner richness, composition, and synchrony. }
 

### 2025-07-18 Manuscript/ant_cactus_SUBMISSION2.tex
methods to end edits

@@
 Long-term demographic data are essential to understanding the fitness consequences of multi-species mutualisms for several reasons. \linelabel{res:long-term-data}
-First, the fitness consequences of sequential partnering play out over long time scales due to the accumulation of partners across time rather than the effects of multiple partners at once. 
+First, fitness effects of sequential partnering play out over long time scales due to the accumulation of partners across time rather than the effects of multiple partners at once. 
 While turnover of mutualist partners is known to occur \citep{Ness2006, Djieto-Lordon2005,Fonseca2003}, long-term data documenting patterns and consequences of partner switching are uncommon, especially in field settings \citep{Palmer2010}.
-Second, certain effects of partner diversity are fluctuation-dependent (i.e., portfolio effect) and would only be detectable in the context of temporally stochastic environments in which alternative partners may fluctuate asynchronously.
+\tom{Second, certain effects of partner diversity are fluctuation-dependent (i.e., portfolio effect) and would only be detectable in the context of temporally stochastic environments in which alternative partners may fluctuate asynchronously.
 Quantifying contributions from the portfolio effect therefore requires long-term data on partner (a)synchrony across inter-annual fluctuations.
-Finally, alternative partners may have different demographic effects (e.g., one that enhances survival and another that enhances reproduction) that contribute differently to overall fitness of a focal mutualist.
+}{This might be a place to more thoroughly address R2's comment about what evidence for portfolio effect would look like.}
+Finally, alternative partners may have different demographic effects (e.g., one that enhances survival and another that enhances reproduction), and these effects may contribute differently to overall fitness of a focal mutualist.
 \tom{Most studies evaluate short-term fitness proxies like seed set or herbivory at a single life stage \citep{Palmer2010, Bruna2014, Trager2010}, while few trace demographic effects across multiple stages.}{I like this sentence but it is unclear whether you are using the citations as examples of studies that only consider single life stages, or the opposite.
 This would be good to clarify, since you don't want to mis-cite people.}
-Integrating long-term data from across the life cycle within a stochastic demographic modeling framework therefore provides a powerful opportunity to diagnose the fitness consequences of partner diversity and turnover.
+Integrating long-term data from across the life cycle within a demographic modeling framework therefore provides a powerful opportunity to diagnose the fitness consequences of partner diversity and turnover in the context of environmental stochasticity.
 }
@@
 \revise{We then built a stochastic, multi-state integral projection model (IPM) which combines diverse effects on vital rates and pathways of partner turnover. 
-Finally, we used the model to conduct simulation `experiments' that quantify plant fitness under different scenarios of partner richness, composition, and synchrony. }
+Finally, we used the model to conduct simulation `experiments' that quantify plant fitness under different scenarios of partner richness, composition, and synchrony.}
 
@@
 These arborescent plants produce cylindrical segments with large spines. 
-In the growing season (May to August in New Mexico), they initiate new vegetative segments and flowerbuds at the ends of existing segments. 
+Early each growing season (May to August in New Mexico), they initiate new vegetative segments and flowerbuds at the ends of existing segments. 
 While most plants produce new segments every season, only those which are reproductively mature produce flowerbuds. 
@@
 Flowerbuds produce more and higher-quality EFN than vegetative segments, making reproductive cholla valuable mutualist partners \citep{Miller2014}. 
-Smaller, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation) \citep{Miller2014}.
+Small, non-flowering cholla produce little to no EFN and are commonly vacant (no ant visitation) \citep{Miller2014}.
 
@@
 This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
-Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) has since been censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
-These censuses were exhaustive, tracking every single plant and searching the entire plot for any new ones. \linelabel{res:exhaustive}
-For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh left it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
+For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
-We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (this differentiation was a visual assessment). 
+We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (aborted flowerbuds are distinguishable by dry and withered true leaves and petals).
 We recorded the ant species presence and identity (or vacancy if no ants present).
 Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
-Plots were searched for new recruits each year, and these were added to the census.
+\linelabel{res:exhaustive} Each year, the field team searched exhaustively for new recruits by walking slowly in a standard formation across the plot area, and these were added to the census (cactus seedlings are small ($<2 cm$ in height) and cryptic so it is likely that some are missed even with a standardized search protocol).
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. 
-In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds.
+\tom{In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
+}{Update with new data package citation, and also update data availability statement.}
+In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds (and therefore are missing some vital rate information for those years).
 
-To complete the tree cholla life cycle, we used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
-These studies were conducted at the same sites using the same plants as our study. \linelabel{res:data_origins}
-Ohm et al.
+\linelabel{res:data_origins1}To complete the tree cholla life cycle, we used additional, smaller data sets from previously published studies to estimate seed and seed bank parameters. 
+First, Ohm et al.
 \citeyear{Ohm2014} provide data on the number of seeds per fruit for plants tended by \textit{L.
@@
 opuntiae}, or no ants (experimental exclusion), accounting for their effects on pollinator visitation.
-Elderd and Miller \citeyear{elderd2016quantifying} provide data on seed entry to the seed bank and seedling germination and survival rates.
+Second, Elderd and Miller \citeyear{elderd2016quantifying} provide data on seed entry to the seed bank and seedling germination and survival rates based on surveys of fruit predation and seed addition experiments.
+Because these seed and seed bank data were not collected as part of our long-term study, these demographic parameters are based on data from a single year and are assumed to be temporally constant.
 
 \subsection*{Multi-state Integral Projection Model}
+\subsubsection*{Model overview}
 The demographic data were used to parameterize a multi-state Integral Projection Model (IPM).
@@
 While IPMs are a natural choice for populations with continuous size structure, they can also be modified to accommodate a combination of continuous and discrete state variables, as we do here. 
-We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explicitly.
+We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species in isolation and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explicitly.
 
@@
 
-We modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates.
+\revise{\linelabel{res:model_story}The biological ``story'' of the model is as follows: 
+The tree cholla population is comprised of plants characterized by their size, which can take any continuous value between the lower and upper limits, and ant visitation status, which can take one of the four discrete levels described above.
+An individual's size and ant status jointly predict its demographic fate such as survival, growth, and reproductive contributions to the seed bank. 
+Plants may be lost to mortality and new individuals may be gained through recruitment from the seed bank. 
+Size and ant status are updated annually.
+The environment fluctuates from year to year such that demographic rates exhibit stochastic deviations from their long-term average. 
+Plants in different ant states can fluctuate independently; for example, a year that is better than average for \textit{L.
+apiculatum}-tended plants may be worse than average for \textit{C.
+opuntiae}, and vice versa, which could occur if there are year-to-year fluctuations in the natural enemies against which each defender species is most effective.
+Our long-term demographic data define the ``rules'' of tree cholla demography and ant partner transition. 
+In the analyses that follow, we tinker with those rules to create novel scenarios of partner diversity and inter-annual synchrony.}
+
+\subsubsection*{Mathematical formulation}
+Mathematically, we modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates.
 The model also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
@@
 Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
-The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a)$) is dependent on ant state $a$. 
+The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$. 
 The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
-For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability, and plants in different ant states can fluctuate independently (good years for \textit{L.
-apiculatum} -occupied plants may be bad years for \textit{C.
-opuntiae}-occupied plants, for example).
+For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability, and plants in different ant states can fluctuate independently.
+% (good years for \textit{L. apiculatum} -occupied plants may be bad years for \textit{C. opuntiae}-occupied plants, for example). 
 Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
@@
 \paragraph{Growth}
-We fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account for left-skewed size transitions (at some initial sizes, transitions below the expected future size were more common than transitions above it).
+Following best practices for modeling size transitions \citep{miller2025my}, we fit the growth sub-model ($G(x',x,a;\pmb{\tau^G_{a}})$) to data on size in year $t+1$ ($y^G$) using the skewed normal distribution to account for left-skewed size transitions (at some initial sizes, transitions below the expected future size were more common than transitions above it).
 The skew-normal has three parameters corresponding to location ($\hat{G}$), shape ($\sigma$), and scale ($\alpha$):
@@
 Estimates for the number of seeds per fruit were obtained from a field experiment which excluded ants from flowering plants \citep{Ohm2014}.
-This experiment only included  vacant or \textit{L.
-apiculatum} and \textit{C.
-opuntiae} -tended plants, so we assumed that plants tended by Other ants had the same number of seeds per fruit as vacant plants (our results showed very little sensitivity to this assumption).
-Additional reproductive parameters for the number of seeds per fruit, probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A.
+This experiment only included \textit{L. apiculatum} and \textit{C. opuntiae} -tended plants.
+We therefore assumed that plants tended by Other ants had the same number of seeds per fruit as vacant plants (our results showed very little sensitivity to this assumption). 
+Additional reproductive parameters for probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A. 
+\revise{\linelabel{res:data_origins2}All of these parameters are assumed to be temporally constant.}
 
@@
 
-\subsubsection*{Partner diversity simulations}
+\subsubsection*{Partner diversity scenarios}
 Using the fully parameterized multi-state IPM, we conducted simulations to quantify how diversity and identity of ant partners influenced plant fitness. 
-From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created treatments corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
+From the full version of the model (described above) corresponding to the observed assemblage of partners and observed frequencies of partner transition, we created \revise{alternative model versions} corresponding to all eight ``counter-factual'' scenarios of diversity and composition: no ant partners (complete vacancy); one ant partner (\textit{C.
 opuntiae} only, \textit{L.
@@
 This means, for example, that the \textit{C. opuntiae}-only scenario included two possible states, vacancy and occupied by \textit{C. opuntiae}. 
-While our statistical models allow us to extrapolate the demographic performance of ant-tended plants to small sizes that are typically vacant, the natural history of this system tells us that this is not biologically sensible.
+Our statistical models allow us to extrapolate the demographic performance of ant-tended plants to small sizes that are typically vacant, but the natural history of this system tells us that this is not biologically sensible.
 Small, non-reproductive plants are typically vacant because they do not produce extrafloral nectar, and once plants begin producing nectar they are nearly always ant-tended \citep{Miller2014}. 
 Our simulations preserved this basic biology, avoiding tiny ant-occupied plants that do not and could not occur in nature. 
+\revise{\linelabel{res:extrap}While technically such plants must exist in the model (because all ant states share the same $L$ and $U$ size limits), they are so infrequent, as guided by our data, as to have virtually no bearing on the results.}
 
-The partner diversity treatment scenarios required additional assumptions about the mechanisms that give rise to observed occupancy patterns. 
+The partner diversity scenarios required additional assumptions about the mechanisms that give rise to observed occupancy patterns. 
 Based on evidence that EFN-bearing cacti are nearly always ant-occupied \citep{Miller2014}, we assume that ant partners competitively exclude one another from EFN-bearing cacti and that competition for plant partners is zero-sum.
@@
 
-\subsubsection*{Temporal stochasticity experiments}
+\subsubsection*{Partner synchrony scenarios}
 Under the portfolio effect hypothesis, partner diversity may confer a fitness advantage when the benefits of alternative partners are not perfectly synchronized across temporal environmental variation, yielding an advantage of a diverse ``portfolio'' of partners when the environment fluctuates.
@@
 Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above. 
+\revise{\linelabel{res:portfolio_methods}In Appendix D, we explore additional scenarios that vary the correlation of ant-specific fluctuations, to determine if contributions from the portfolio effect are constrained by high correlations seen in the long-term data (Results).}
 %We indicate stochastic fitness under the synchronized scenario with $\lambda^{C(s)}_{S}$, for example.
@@
 
-Similarly, for plants which were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). 
-Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants the most vulnerable to mortality. 
+Similarly, for plants that were large enough to have ant visitors, visitation enhanced cactus survival (Figure \ref{fig:Surv}). 
+Mean survival rates ranged from 7.7\% to 99.9\%, with the smallest plants being the most vulnerable to mortality. 
 \textit{C. opuntiae}-occupied plants had a survival advantage over other ant-tended plants, particularly at smaller sizes, consistent with positive effects on growth. 
@@
 opuntiae}-tended plants (57.1\% [40.6--72\%]).
-Furthermore, \textit{C.
+\tom{Furthermore, \textit{C.
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
-apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.
+apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.}{I think a reviewer asked to see this result, so be sure to reference the figure in the appendix and flag this as a revision in the response letter.
+Also, the figure legend should be updated to indicate that we are looking at posterior probability distributions, which some readers will not understand if you don't explicitly state.}
 
@@
 
-Inter-annual fluctuations in demographic rates (estimated as statistical random effects) were generally positively correlated between ant states, limiting the potential for benefits of diversity through the portfolio effect (Figure \ref{fig:Annual_Ant}).
+\tom{Inter-annual fluctuations in demographic rates (estimated as statistical random effects) were generally positively correlated between ant states, limiting the potential for benefits of diversity through the portfolio effect (Figure \ref{fig:Annual_Ant}).}{Be sure to update this figure and flag it as a revision in the response letter.}
 However, the degree of correlation depended on the vital rate and pair of ant partners. 
 Across vital rates, random effects for cactus growth were the most strongly correlated between ant states (mean pairwise correlation: 0.63) and random effects for survival were the least correlated (mean pairwise correlation: 0.36).
-Certain partner pair / vital rate combinations were effectively independent in their yearly fluctuations (such as survival of \textit{C.
-opuntiae} and Other-tended plants) while others were almost completely synchronous (such as growth of Vacant and \textit{L.
-apiculatum}-tended plants).
+Certain partner pair / vital rate combinations were effectively independent in their yearly fluctuations (\tom{such as survival of \textit{C.
+opuntiae} and Other-tended plants}{Give the correlation coefficient.}) while others were almost completely synchronous (\tom{such as growth of Vacant and \textit{L.
+apiculatum}-tended plants}{Give the correlation coefficient.}).
 
@@
 We have high confidence that this contrast is positive, and equally so for the synchronous and non-synchronous scenarios (Figure S\ref{fig:Portfolio}). 
+\revise{\linelabel{res:portfolio_results}In Appendix D, we show that contributions from the portfolio effect were constrained by positively correlated fluctuations between ant states (Figure \ref{fig:Annual_Ant}).
+In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model (Figures), as expected under the portfolio effect hypothesis.}
 
@@
 Yet, our stochastic, multi-state IPM revealed that all scenarios which included any partners resulted in similarly high cactus fitness, with a statistically significant but quantitatively weak cost of diversity at the highest level.
-Furthermore, evidence from nearly two decades of demographic data failed to support the ``portfolio effect'' hypothesis, whereby benefits of diversity manifest in the context of environmental fluctuations.
-Overall, our results indicate that while ant visitation is strongly consequential, partner composition is much less so, suggesting that the fitness benefits of mutualism from the plant's perspective may be surprisingly resilient to declines in partner diversity.
+Furthermore, evidence from nearly two decades of demographic data failed to support the ``portfolio effect'' hypothesis, whereby benefits of diversity manifest in the context of environmental fluctuations, \revise{because fluctuations in the demographic effects of alternative partners are too synchronized}.
+Overall, our results indicate that while ant visitation is strongly consequential for plant demography, partner composition is much less so, suggesting that the fitness benefits of mutualism from the plant's perspective may be surprisingly resilient to declines in partner diversity.
 
@@
 To our knowledge, ours is the first study documenting a strong, positive fitness effect of mutualism, regardless of mutualist identity or guild composition. 
-As this field grows, particularly with studies that integrate multiple vital rates across the life cycle, we will be better positioned to understand whether positive, neutral, or negative effects of partner diversity predominate, and the conditions under which each might be more likely.
+As this field grows, particularly with long-term studies that integrate multiple vital rates across the life cycle, we will be better positioned to understand whether positive, neutral, or negative effects of partner diversity predominate, and the conditions under which each might be more likely.
 
 % Portfolio Effect
-When partners exhibit different reactions to varying environments, interacting with a diverse portfolio of partners can lead to more consistent benefits across time \citep{Batstone2018}.
+\tom{When}{I think you will want to expand this paragraph to discuss the new results.} partners exhibit different reactions to varying environments, interacting with a diverse portfolio of partners can lead to more consistent benefits across time \citep{Batstone2018}.
 Our work explicitly incorporated temporal environmental stochasticity, yet we find no evidence for the portfolio effect as a mechanism of diversity benefits. 
@@
 For example, our previous study that focused only on reproductive vital rates suggested that \textit{C.
-opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set \citep{Ohm2014}.
+opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed \tom{set}{Reference the appendix figure here.} \citep{Ohm2014}.
 Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.
@@
 Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
-Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
+\tom{Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
+}{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.}
 Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
@@
 \paragraph{Conclusion}
-This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangeable in their net effects on fitness of a shared mutualist based on where their effects occur in the life cycle.
+This study highlights that while partners within a mutualistic guild can be ecologically different, they may still be interchangeable in their net effects on fitness of a shared mutualist based on where their effects occur in the life cycle \revise{-- highlighting the importance of a demographic perspective in understanding the fitness consequences of mutualisms}.
 Contrary to much of the literature on multi-species mutualisms, we find that while ant partners are beneficial, partner diversity has no measurable benefit to their shared mutualist and may even be somewhat costly.
@@
 J. Moutouama, A. Bradley, B. Scherick, A. Gough, J. Martin, and V. Rudolf provided helpful feedback on the manuscript. 
+\tom{}{Never forget to thank reviewers in the acknowledgments of resubmissions!}
 
@@
 \section*{Data and Code Availability}
-This section will be filled out once double blind reviews are complete. 
+\tom{This section will be filled out once double blind reviews are complete. }{Be sure to update this with the new data package.}
 %Most of the data that support the findings of this study (through 2019) are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1.

### 2025-07-18 Manuscript/Ant_Cactus_RESPONSE.tex
Update Ant_Cactus_RESPONSE.tex

@@
 
-\newcommand{\revise}[1]{{\color{Mahogany}{#1}}}
-
+\newcommand{\revise}[1]{{\color{blue}{#1}}}
+\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}} 
 \usepackage[normalem]{ulem}
@@
 \begin{enumerate}
-	\item We have added a new analysis that explores under what conditions portfolio effect could be shown. 
-	We found *************
-	\item We updated the introduction to be more accessible for readers and to better set up the rest of the manuscript.
+	\item In response to Reviewer 2, we have added a new analysis that explores under what conditions portfolio effect could generate benefits of partner diversity.
+We found that contributions from the portfolio effect are constrained by strong positive correlations in the demographic effects of alternative partners.
+Scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated generated results consistent with the portfolio effect hypothesis.
+	\item In response to both reviewers and the handling editor, we have heavily revised the introduction to be more accessible for readers and to better set up the rest of the manuscript.
+	\item We have re-drawn Figure XX to make it more easily interpretable, as suggested by the handling editor.
 \end{enumerate}
@@
 We describe these and other changes in greater detail below, where we reproduce comments from the associate editor and reviewers and provide our point-by-point responses. 
-All of our changes are denoted in the manuscript with \revise{Mahogany font}.
+All of our changes are denoted in the manuscript with \revise{blue font}.
 We think the review process has greatly strengthened our manuscript such that it is now suitable for publication.
@@
 	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{}). 
-	We have gone through the data collection methods section and added greater detail throughout, which we hope will make our work more understandable and reproducible.}
+	We have gone through the data collection methods section and added greater detail throughout (lines \ref{}), which we hope will make our work more understandable and reproducible.}
 
@@
 \comment{Lines 329 and 364: I realize that this is commonplace in some parts of the literature, but I would prefer not to call different simulation scenarios “experiments”}
-\response{We have changed the language throughout to "simulations" rather than "experiments".}
+\response{We have changed the language throughout to "simulations" or "scenarios" rather than "experiments".}
 
@@
 	Instead, these data come from plants in different ant states during a single year. 
-	We have expanded on the source of our seed data to make all of this clearer (line~\ref{ms-res:data_origins}). 
+	We have expanded on the source of our seed and seed bank data to make all of this clearer (line~\ref{ms-res:data_origins1}). 
 	The seed data do support the idea of defense decreasing pollination, and this is accounted for in our modeling framework. 
-	Because we lack long-term data on seed production, we must assume that this cost of defense is temporally static; we now state this assumption explicitly (line \ref{}).}
+	Because we lack long-term data on seed production, we must assume that this cost of defense is temporally static; we now state this assumption explicitly (line \ref{ms-res:data_origins1}, \ref{ms-res:data_origins2}).}
 
@@
 This revision improves the focus of the introduction on the core aspects of our study. 
-Additionally, we have expanded the discussion on generalized/diffuse mutualisms, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}
+Additionally, \tom{we have expanded the discussion on generalized/diffuse mutualisms}, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}{I don't really see any discussion of generalized/diffuse mutualisms.}
 
@@
 Sometimes mutualism showed more CD than the others, but it's constantly discussed in mutualism, but not in the another two antagonistic interactions.}
-\response{I removed this sentence based on their recommendations so I am not sure what I should say here...
-********* Tom: I will check the manuscript text, but probably simplest to change the way this paper gets cited so that we are not asserting that mutualism is more context-dependent.}
+\response{This is a great point. Because we have heavily revised the introduction this concern is no longer relevant.}
 
@@
 \response{The simplest response we can offer here is that the small plant data are \emph{effectively} not extrapolated, so the expected performance of tiny ant-tended plants has little bearing on our results.
-	The more complete, technical response is that some amount of extrapolation is an inevitable feature of the Integral Projection Model framework. 
+	The more complete, technical response is that some amount of extrapolation is an inevitable feature of the Integral Projection Model framework, and that is why we show the extrapolated vital rate functions in the figure.
 	Because the state space is defined by size and ant status, we must tell the model what to expect in all combinations of those states. 
 	However, because state transitions are guided by empirical data, the model ``knows'' which state combinations are likely and unlikely. 
-	Tiny, ant-tended plants do exist in the model as a possible state combination, because they must -- but these are at such low frequency (as informed by the data) that the model is effectively insensitive to what we assume about their demographic performance.
-	We have tried to better communicate this nuance in our presentation of the model (line \ref{}).}
+	Tiny, ant-tended plants do exist in the model as a possible state combination, because they must -- but these occur at such low frequency (as informed by the data) that the model is effectively insensitive to what we assume about their demographic performance.
+	We have tried to better communicate this nuance in our presentation of the model (line \ref{ms-res:extrap}).}
 
@@
 	We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
-	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory})[TOM: let's work on this, it's a little clunky.].
-	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{}). }
+	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).
+	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{}).
+	\tom{}{I think there is more work to do to address this comment.}}
 
@@
 \response{Thank you for this intriguing suggestion, which got us thinking about our results in a way we hadn't before. 
-In response to this suggestion, we have added several new analyses that provide a ``positive control'', demonstrating the conditions under which our model reveals a portfolio effect.
-[more details]}
+In response to this suggestion, we have added new analyses that provide a ``positive control'', demonstrating the conditions under which our model reveals a portfolio effect.
+Specifically, we conducted new simulations in which we re-arranged ant-specific fluctuations to make them less positively correlated than we observe in the long-term data (line~\ref{ms-res:portfolio_methods} and new Appendix D).
+As predicted by theory, we detected significant benefits of diversity via the portfolio effect when partner fluctuations were uncorrelated or negatively correlated (line~\ref{ms-res:portfolio_results}, new Figures).
+These additions enrich our understand of the model results and provide the ``positive control'' that the reviewer is looking for.}
 
@@
 	Abstracts tend to come together late in the manuscript preparation process, but this is a healthy reminder that they require as much effort and attention as all the rest. 
-	We have re-written the abstract to clarify the overarching question—how partner turnover influences host demography and fitness[TOM: is this actually the question???
-let's work on this]—and to clarify the specific demographic outcomes of interest.
+	We have re-written the abstract to clarify the overarching question—how partner turnover influences host demography and fitness—and to clarify the specific demographic outcomes of interest. 
 We also simplified the language around partner turnover.
-We believe this revised version more clearly communicates the goals, approach, and key findings of the study.}
+We believe this revised version more clearly communicates the goals, approach, and key findings of the study.
+\tom{}{See the new abstract because I edited heavily. You may want to adjust this response accordingly.}
+}
 
@@
 \comment{A bit jarring out the gate. The first sentence states: everyone in mutualism benefits (full stop). The second sentence is : except when they don't.}
-\response{We have revised the first sentence (line ~ \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.
-We believe the language of net benefit allows for the further elaboration than the magnitude of benefits vary based on several conditions (aka.
-except when they don't)[TOM: I think we can make this revision and response smoother].}
+\response{We have revised the first sentence (line ~ \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.}
 
@@
 How relevant?}
-\response{We have edited this sentence to be more specific about ``applied settings'' we referenced previously. 
-For example, prior work has examine how diversity of pollinator assemblages affects productivity of crop plants (see line \ref{}). }
+\response{In this sentence we were referencing a study of how diversity of pollinator assemblages affects productivity of crop plants (Rogers et al.
+2014).
+However, rather than go into the details of that study, we have simplified the text and no longer mention ``applied settings'' (though we continue to cite Rogers et al.)  (line ~\ref{ms-res:L39}).}
 
@@
 How is this different from when partners are present simultaneously?}
-\response{We have revised this paragraph to better answer your questions and deliver on its opening. 
-We explain that in sequential associations, partner conflict is significantly reduced and associations are more stable [TOM: not sure I agree] -- leading to overall increased benefits for the focal mutualist (line ~\ref{ms-res:L63}).}
+\response{We have heavily revised this paragraph to better answer your questions and deliver on its opening (line ~\ref{ms-res:L63}). 
+Because our study examined a sequentially-partnered mutualism, we have sharpened the Introduction to focus on this scenario, rather than explore the contrast between simultaneous and sequential partnering.}
 
@@
 I know the complex bits have to be there and that mature, quantitative readers exist, but this reviewer cleaves more to a lower common denominator}
-\response{We appreciate this comment and we have taken inspiration from the reviewer's next comment to add the lay summary suggested here.}
+\response{We appreciate this comment and we have taken inspiration from the reviewer's next comment to add the lay summary suggested here (line \ref{ms-res:model_story}).}
 
@@
 (7)     This model's outputs are used to see if this portfolio effect emerges. }
-\response{Your understanding of our model and analysis is accurate. 
-	In fact, this is such a good overview that it inspired a new section we added (lines \ref{}) to introduce the model, so that readers can first understand the ``story'' of the model before seeing any equations.}
+\response{The reviewer's understanding of our model and analysis is accurate. 
+	In fact, this is such a good overview that it inspired a new section we added (lines \ref{ms-res:model_story}) to introduce the model, so that readers can first understand the ``story'' of the model before seeing any equations.}
 

### 2025-08-25 Manuscript/Ant_Cactus_RESPONSE.tex
Updates
Manuscript Response

@@
 	We agree that methods for estimating seed production were a little under-developed in our previous submission because we relied too heavily on citations of our prior work. 
-	We now describe these methods more thoroughly (lines \ref{}). 
-	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{}). 
-	We have gone through the data collection methods section and added greater detail throughout (lines \ref{}), which we hope will make our work more understandable and reproducible.}
+	We now describe these methods more thoroughly (lines \ref{ms-res:seed_prod}). 
+	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{ms-res:viable}). 
+	We have gone through the data collection methods section and added greater detail throughout (lines \ref{ms-res:repro1}, \ref{ms-res:repro2},  \ref{ms-res:viable}, \ref{ms-res:exhaustive}), which we hope will make our work more understandable and reproducible.}
 
@@
 \response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
-	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines XX and Figure XX).
-	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{}). 
-	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{}). }
+	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines \ref{ms-res:portfolio_evidence} and Figure \ref{ms-app:portfolio}).
+	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{ms-res:portfolio_asynchronous}).
+	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{ms-res:portfolio_methods}). }
 
@@
 Please make sure all supplemental results are clearly described in the Results section of the main text.}
-\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{}. \ref{}).}
+\response{Thank you for pointing this out.
+We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{res:portfolio_methods}, \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}
 

### 2025-08-25 Manuscript/ant_cactus_SUBMISSION2.tex
Updates
Manuscript Response

@@
 \linelabel{res:L22}
-Understanding how partner diversity and turnover affect the overall fitness benefits of mutualism is an open question -— especially in the context of environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
+Understanding how partner diversity and turnover affect the overall fitness benefits of mutualism is an open question, especially in the context of environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}.
 \linelabel{res:L39}}
@@
 Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
+\linelabel{res:repro1}
 For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
-Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
+Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. \linelabel{res:repro2}
 We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (aborted flowerbuds are distinguishable by dry and withered true leaves and petals).
+\linelabel{res:viable}
 We recorded the ant species presence and identity (or vacancy if no ants present).
@@
 Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
-Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.
+Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.  \linelabel{res:seed_prod}
 
@@
 \caption{Probability of survival in relation to size and ant partner.
+These are the posterior probability distributions.
 Layout as in Fig.
@@
 opuntiae}-tended plants (57.1\% [40.6--72\%]).
-\tom{Furthermore, \textit{C.
+Furthermore, \textit{C.
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
-apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.}{I think a reviewer asked to see this result, so be sure to reference the figure in the appendix and flag this as a revision in the response letter.
-Also, the figure legend should be updated to indicate that we are looking at posterior probability distributions, which some readers will not understand if you don't explicitly state.}
+apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.
+\linelabel{res:viability}
+\ali{}{I am not sure where in the appendix you want this to be referenced?}
 
@@
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
-Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above.
+Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above. \linelabel{res:appendixc}
 %Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
@@
 For example, our previous study that focused only on reproductive vital rates suggested that \textit{C.
-opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed \tom{set}{Reference the appendix figure here.} \citep{Ohm2014}.
+opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set (Figure \ref{app:seeds}) \citep{Ohm2014}.
+\linelabel{res:seeds}
 Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.
@@
 \tom{Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
-}{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.}
+}{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.} \linelabel{res:herbivory}
 Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
@@
 
+\section*{Appendix E}\label{appendix:E}
+To further evaluate the potential for a portfolio effect in this system, we conducted an additional analysis in which we artificially increased the degree of difference among ant partners. 
+Specifically, we manipulated the correlation structure of ant-specific year random effects in the stochastic IPM. 
+In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
+In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.
+
+For each of these scenarios, we quantified the effect of partner diversity as the difference between projected population growth with and without partner diversity ($\lambda_S^4 - \lambda_S^0$), and compared results under synchronous versus non-synchronous environmental variation.
+Both scenarios resulted in greater benefits of partner diversity under Non-synchronous conditions, though the evidence was much stronger when the ants were anti-correlated.
+
+This shows that portfolio effect is possible in this system if the ant partners diverged more greatly in their responses to annual variation. \linelabel{res:portfolio_evidence}
+
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/cor0_Portfolio_Effect.png}
+	\caption{The mean estimations of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in green and pink respectively. }
+	\label{app:portfolio}
+\end{figure}
+
 \end{document}

### 2025-08-25 Manuscript/ant_cactus_SUBMISSION2.tex
Updates

@@
 \revise{In sequential mutualisms, partner identity can change over time due to environmental variability or competitive displacement. 
-Here we asked: how do partner diversity and turnover influence host demography and long-term fitness, and which mechanisms explain these effects?}
+Here we asked how partner diversity and turnover influence host demography and long-term fitness, and which mechanisms explain these effects.}
 We studied the sequential multi-species food-for-protection mutualism involving an EFN-bearing plant (tree cholla cactus \textit{Cylindropuntia imbricata}) and multiple species of nectar-feeding ants (\textit{Crematogaster opuntiae}, \textit{Liometopum apiculatum}, and other less frequent ants) that provide defense from herbivores.
@@
 %This concept is well developed in biodiversity–ecosystem functioning literature, but has rarely been applied to mutualisms in which partner diversity is realized through turnover rather than coexistence.
-While portfolio effects have been well studied in plant and animal communities, their role in mutualisms —- particularly where partner diversity is sequential rather than simultaneous —- is less explored \citep{Ushio2020}.}
+While portfolio effects have been well studied in plant and animal communities, their role in mutualisms  (particularly where partner diversity is sequential rather than simultaneous) is less explored \citep{Ushio2020}.}
 

### 2025-08-25 Manuscript/Ant_Cactus_RESPONSE.tex
Submissions

@@
 	\item In response to both reviewers and the handling editor, we have heavily revised the introduction to be more accessible for readers and to better set up the rest of the manuscript.
-	\item We have re-drawn Figure XX to make it more easily interpretable, as suggested by the handling editor.
+	\item We have re-drawn Figure \ref{ms-fig:Annual_Ant} to make it more easily interpretable, as suggested by the handling editor.
 \end{enumerate}
@@
 \response{We follow a consistent search protocol, so while it's possible we miss some recruits we think this source of observation error is small and roughly consistent across years.
-We have clarified this (line ~\ref{ms-res:exhaustive}).}
+We have clarified this (line \ref{ms-res:exhaustive}).}
 
@@
 That is, we have built a general model of demographic transitions (births, deaths, growth, and partner turnover) that is realistic with respect to our study system, but we query the model for qualitative inference (e.g., asking whether a multi-partner scenario yields higher or lower fitness than the single best partner) and not quantitative precision.
-We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}).}
+We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}).
+******** Tom have you already done this?
+Or is this something you would like me to do?}
 
@@
 This revision improves the focus of the introduction on the core aspects of our study. 
-Additionally, \tom{we have expanded the discussion on generalized/diffuse mutualisms}, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).}{I don't really see any discussion of generalized/diffuse mutualisms.}
+Additionally, we have expanded the discussion on generalized/diffuse mutualisms}, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).
 
@@
 \comment{L. 429: Is there a figure equivalent to figs. 1 and 2 that show seed production/set?}
-\response{Yes, we have this figure in Appendix D (Figure ~\ref{ms-app:seeds}). 
+\response{Yes, we have this figure in Appendix D (Figure \ref{ms-app:seeds}). 
 	We could have done a better job in our first submission of linking the main text and the supporting material. 
-	Hopefully these connections are clearer now (lines \ref{}).}
+	Hopefully these connections are clearer now (lines \ref{ms-res:seeds}).}
 
@@
 \response{This is really helpful feedback. 
-	We have revised the Introduction and Abstract to more explicitly detail portfolio effects and sequential partner interactions (lines \ref{}, \ref{}).
+	We have revised the Introduction and Abstract to more explicitly detail portfolio effects and sequential partner interactions (lines \ref{ms-res:portfolio_criteria}, \ref{ms-res:portfolio},\ref{ms-res:sequential}, \ref{ms-res:abstract}).
 	These revisions better prime the reader to understand the core research question and the rationale behind our modeling approach, which integrates partner turnover and environmental stochasticity. 
@@
 	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).
-	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{}).
-	\tom{}{I think there is more work to do to address this comment.}}
+	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{ms-res:portfolio_criteria}).}}
 
@@
 We also simplified the language around partner turnover.
-We believe this revised version more clearly communicates the goals, approach, and key findings of the study.
-\tom{}{See the new abstract because I edited heavily. You may want to adjust this response accordingly.}
-}
+We believe this revised version more clearly communicates the goals, approach, and key findings of the study.}
 
@@
 \comment{A bit jarring out the gate. The first sentence states: everyone in mutualism benefits (full stop). The second sentence is : except when they don't.}
-\response{We have revised the first sentence (line ~ \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.}
+\response{We have revised the first sentence (line \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.}
 
 \comment{27: Compared to which interactions ("such as…")? I think this intro material obstructs you from rolling out the background we really need.}
-\response{We have revised the first paragraph to reduce unnecessary and/or uninformative information to speed the reader into what is needed to understand the rest of the paper (line ~\ref{ms-res:L27}).}
+\response{We have revised the first paragraph to reduce unnecessary and/or uninformative information to speed the reader into what is needed to understand the rest of the paper (line \ref{ms-res:L27}).}
 
@@
 2014).
-However, rather than go into the details of that study, we have simplified the text and no longer mention ``applied settings'' (though we continue to cite Rogers et al.)  (line ~\ref{ms-res:L39}).}
+However, rather than go into the details of that study, we have simplified the text and no longer mention ``applied settings'' (though we continue to cite Rogers et al.)  (line \ref{ms-res:L39}).}
 
 \comment{45. Wordy sentences. Do we need both "hierarchy of fitness effects" and "a consistent ranking…".}
-\response{We have edited to reflect this suggestion (line ~\ref{ms-res:L45}).}
+\response{We have edited to reflect this suggestion (line \ref{ms-res:L45}).}
 
@@
 How is this different from when partners are present simultaneously?}
-\response{We have heavily revised this paragraph to better answer your questions and deliver on its opening (line ~\ref{ms-res:L63}). 
+\response{We have heavily revised this paragraph to better answer your questions and deliver on its opening (line \ref{ms-res:L63}). 
 Because our study examined a sequentially-partnered mutualism, we have sharpened the Introduction to focus on this scenario, rather than explore the contrast between simultaneous and sequential partnering.}

### 2025-08-25 Manuscript/ant_cactus_SUBMISSION2.tex
Submissions

@@
 While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal species interacts with multiple partner species.
+These broader, network-structured interactions are characteristic of generalized or diffuse mutualisms, in which mutualists interact with multiple partners rather than engaging in strict one‐to‐one pairings \citep{Aizen2014, Prior2015}.
 Growing evidence suggests that pairwise interactions may be poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}. 
@@
 \revise{Second, even within a mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
-\linelabel{res:L53}\tom{}{There are two different Stachowicz references for the same paper -- be sure to clean this up.}
+\linelabel{res:L53}
 This can lead to a positive effect of partner diversity through complementarity, whereby diversity increases the net benefits received by focal mutualist beyond what a single partner species could offer \citep{Batstone2018}.
@@
 \linelabel{res:portfolio_asynchronous}
-Critically, portfolio effects require asynchronous fluctuations among partners, such that declines in one partner's benefits are offset by others, stabilizing the focal mutualist's fitness over time \citep{Loreau1999, Tilman1994, Hautier2014}.
-%This concept is well developed in biodiversity–ecosystem functioning literature, but has rarely been applied to mutualisms in which partner diversity is realized through turnover rather than coexistence.
-While portfolio effects have been well studied in plant and animal communities, their role in mutualisms  (particularly where partner diversity is sequential rather than simultaneous) is less explored \citep{Ushio2020}.}
-
-%% Partner turnover
-%Partner diversity can affect a focal mutualist differently depending on whether partners are present simultaneously or sequentially (i.e., via partner turnover) \citep{Djieto-Lordon2005, Ness2006, Bruna2014, Barrett2015, Ushio2020, Dattilo2014}.
-%In simultaneous associations, interference between partners may reduce the net benefits received by a focal mutualist \citep{Kiers2003,Batstone2018, Tgaard2015, Wulff2008}. \linelabel{res:L63}
-%Exclusive, sequential associations can reduce antagonistic partner-partner interactions \citep{Wolfin2023,Stachowicz2005_2} and enhance mutualism stability \cite{Johnson2019}.
+For such a portfolio effect to operate, two conditions must be met: (i) partner species must differ in the timing or magnitude of their responses to environmental variation (i.e., their benefits fluctuate asynchronously), and (ii) the combined effect of these fluctuations must result in increased net benefit to the host across years relative to the net benefit expected if all partners responded in parallel \citep{Loreau1999, Tilman1994, Hautier2014}.
+\linelabel{res:portfolio_criteria}
+While portfolio effects have been well studied in plant and animal communities, their role in mutualisms—particularly where partner diversity is expressed sequentially rather than simultaneously—remains less explored \citep{Ushio2020}.}
 
@@
 \revise{For exclusive, sequential associations, the timing, frequency, and directionality of partner turnover can influence the net fitness benefits to a focal mutualist, particularly if turnover allows accumulation of complementary benefits or buffers environmental variability via portfolio effects \citep{Sachs2004,Fiala1994, Ushio2020}.
-\linelabel{res:L63}
-\tom{Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986}.}{any more recent examples we could cite here?} 
+\linelabel{res:L63} \linelabel{res:sequential}
+Turnover can happen at different timescales, from minutes to years \citep{Oliveira1999,Horvitz1986,VentreLespiaucq2021,Nakazawa2020}.
 If partner identity changes systematically across ontogeny, such as through succession of partners aligned with host life stages, this may further affect fitness outcomes associated with partner diversity \citep{Fonseca2003,Noe1994,Dejean2008}.
@@
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-\revise{Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because long-term demographic studies on ant-plant interactions are \tom{exceedingly rare}{Any other examples we can cite here?
-How long was the Palmer Acacia study?}.
+\revise{Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because long-term demographic studies on ant-plant interactions are exceedingly rare \cite{Donald2022,Moraes2009}.
+}
 
-Long-term demographic data are essential to understanding the fitness consequences of multi-species mutualisms for several reasons. \linelabel{res:long-term-data}
+\revise{Long-term demographic data are essential to understanding the fitness consequences of multi-species mutualisms for several reasons. \linelabel{res:long-term-data}
 First, fitness effects of sequential partnering play out over long time scales due to the accumulation of partners across time rather than the effects of multiple partners at once. 
 While turnover of mutualist partners is known to occur \citep{Ness2006, Djieto-Lordon2005,Fonseca2003}, long-term data documenting patterns and consequences of partner switching are uncommon, especially in field settings \citep{Palmer2010}.
-\tom{Second, certain effects of partner diversity are fluctuation-dependent (i.e., portfolio effect) and would only be detectable in the context of temporally stochastic environments in which alternative partners may fluctuate asynchronously.
+Second, certain effects of partner diversity are fluctuation-dependent (i.e., portfolio effect) and would only be detectable in the context of temporally stochastic environments in which alternative partners may fluctuate asynchronously.
 Quantifying contributions from the portfolio effect therefore requires long-term data on partner (a)synchrony across inter-annual fluctuations. 
-}{This might be a place to more thoroughly address R2's comment about what evidence for portfolio effect would look like.}
+In practical terms, evidence for a portfolio effect would include: 
+(i) partner species showing asynchronous variation in their benefits to the focal host across years, and 
+(ii) the combined effect of these fluctuations must increase the net benefits of partner diversity to the host relative to the variance expected if all partners responded in parallel.
+\linelabel{res:portfolio}
+Detecting these patterns typically involves either correlation analyses of inter-annual partner benefits, or null-model simulations comparing observed host fitness variance to expectations under random or synchronous partner turnover.
 Finally, alternative partners may have different demographic effects (e.g., one that enhances survival and another that enhances reproduction), and these effects may contribute differently to overall fitness of a focal mutualist.
-\tom{Most studies evaluate short-term fitness proxies like seed set or herbivory at a single life stage \citep{Palmer2010, Bruna2014, Trager2010}, while few trace demographic effects across multiple stages.}{I like this sentence but it is unclear whether you are using the citations as examples of studies that only consider single life stages, or the opposite.
-This would be good to clarify, since you don't want to mis-cite people.}
+Most studies evaluate short-term fitness proxies, such as seed set or herbivory, often focusing on a single life stage \citep{Palmer2010, Trager2010}, while fewer trace demographic effects across multiple stages \citep{Bruna2014}.
 Integrating long-term data from across the life cycle within a demographic modeling framework therefore provides a powerful opportunity to diagnose the fitness consequences of partner diversity and turnover in the context of environmental stochasticity.
@@
 Switches between partner species, or between vacancy and ant occupancy, commonly occur from one growing season to the next \citep{Miller2007}. 
-Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
+\revise{Prior experiments suggested a hierarchy of mutualist quality, with \textit{Liometopum apiculatum} providing strong anti-herbivore defense and \textit{Crematogaster opuntiae} having net negative effects because herbivore deterrence is outweighed by deterrence of pollinators \citep{Miller2007,Ohm2014}.
 \linelabel{res:L111}
 However, previous studies in this system focused on single life stages (adult plants) or vital rates (seed production), did not integrate the demographic effects of ant defense across the life cycle, and did not incorporate temporal stochasticity -- gaps we aimed to fill in the present study.
+}
 
@@
 \subsection*{Data Collection}
-This study is based on long-term data spanning 2004 to 2023 that documented plant demographic rates in relation to their ant partner status. 
+\revise{This study is based on long-term data spanning 2004 to 2023 that documented plant demographic rates in relation to their ant partner status. 
 From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
@@
 These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
-\tom{In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}.
-}{Update with new data package citation, and also update data availability statement.}
+In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. 
 In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds (and therefore are missing some vital rate information for those years).
@@
 Second, Elderd and Miller \citeyear{elderd2016quantifying} provide data on seed entry to the seed bank and seedling germination and survival rates based on surveys of fruit predation and seed addition experiments.
-Because these seed and seed bank data were not collected as part of our long-term study, these demographic parameters are based on data from a single year and are assumed to be temporally constant.
+Because these seed and seed bank data were not collected as part of our long-term study, these demographic parameters are based on data from a single year and are assumed to be temporally constant. }
 
@@
 Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
-Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.  \linelabel{res:seed_prod}
+\revise{Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.  \linelabel{res:seed_prod}}
 
@@
 opuntiae}-tended plants (57.1\% [40.6--72\%]).
-Furthermore, \textit{C.
+\revise{Furthermore, \textit{C.
 opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
 apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.
-\linelabel{res:viability}
+\linelabel{res:viability}}
 \ali{}{I am not sure where in the appendix you want this to be referenced?}
@@
 
-\tom{Inter-annual fluctuations in demographic rates (estimated as statistical random effects) were generally positively correlated between ant states, limiting the potential for benefits of diversity through the portfolio effect (Figure \ref{fig:Annual_Ant}).}{Be sure to update this figure and flag it as a revision in the response letter.}
+Inter-annual fluctuations in demographic rates (estimated as statistical random effects) were generally positively correlated between ant states, limiting the potential for benefits of diversity through the portfolio effect (Figure \ref{fig:Annual_Ant}).
 However, the degree of correlation depended on the vital rate and pair of ant partners. 
 Across vital rates, random effects for cactus growth were the most strongly correlated between ant states (mean pairwise correlation: 0.63) and random effects for survival were the least correlated (mean pairwise correlation: 0.36).
-Certain partner pair / vital rate combinations were effectively independent in their yearly fluctuations (\tom{such as survival of \textit{C.
-opuntiae} and Other-tended plants}{Give the correlation coefficient.}) while others were almost completely synchronous (\tom{such as growth of Vacant and \textit{L.
-apiculatum}-tended plants}{Give the correlation coefficient.}).
+Certain partner pair / vital rate combinations were effectively independent in their yearly fluctuations (such as survival of \textit{C.
+opuntiae} and Other-tended plants; mean pairwise correlation: -0.0016) while others were almost completely synchronous (such as growth of Other and \textit{L.
+apiculatum}-tended plants; mean pairwise correlation: 0.093).
 
@@
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
-Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above. \linelabel{res:appendixc}
+\revise{Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above. \linelabel{res:appendixc}}
 %Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
@@
 % Portfolio Effect
-\tom{When}{I think you will want to expand this paragraph to discuss the new results.} partners exhibit different reactions to varying environments, interacting with a diverse portfolio of partners can lead to more consistent benefits across time \citep{Batstone2018}.
+When partners exhibit different reactions to varying environments, interacting with a diverse portfolio of partners can lead to more consistent benefits across time \citep{Batstone2018}.
 Our work explicitly incorporated temporal environmental stochasticity, yet we find no evidence for the portfolio effect as a mechanism of diversity benefits. 
 In ant-plant defensive mutualisms, any portfolio effect would emerge from fluctuations in the community of herbivores, and ant-specific responses to each herbivore species. 
-Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population -- notably, weather -- override any fluctuations in herbivore composition and ant-herbivore interactions.
+Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population ( notably, weather) override any fluctuations in herbivore composition and ant-herbivore interactions.
+To further explore the potential for portfolio effects, we conducted simulations in which ant responses to herbivory were artificially manipulated to be independent (correlation = 0) or anti-synchronous (correlation = -1).
+We found evidence of the portfolio effect (the net benefit of partner diversity was greater when ants responded to inter-annual fluctuations non-synchronously than when they responded synchronously) in both scenarios.
+These results were particularly strong when the ants were anti-synchronous.
+These results indicate that while the portfolio effect can emerge under conditions of asynchronous partner responses, natural drivers in our study system result in synchronous partner effects, masking the potential benefits of partner diversity.
 Dallas et al. \citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is infrequently detected in empirical data across many systems. 
-This indicates that the portfolio effect may be difficult to detect, disguised by different mechanisms, or uncommon in nature.
+This indicates that the portfolio effect may be difficult to detect (particularly in similar systems where ant differences are only somewhat non-synchronous), disguised by different mechanisms, or uncommon in nature.
 
@@
 By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of ant visitation. 
-For example, our previous study that focused only on reproductive vital rates suggested that \textit{C.
+\revise{For example, our previous study that focused only on reproductive vital rates suggested that \textit{C.
 opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set (Figure \ref{app:seeds}) \citep{Ohm2014}.
-\linelabel{res:seeds}
+\linelabel{res:seeds}}
 Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.
@@
 % Limitations
-As with any study, there are limitations to consider when interpreting our results.
+\revise{As with any study, there are limitations to consider when interpreting our results.  
 First, we used observational data to infer ant effects on plant demography. 
@@
 The combination of long-term observational data backed up by experimental results gives us greater confidence in our causal interpretations.  
-Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
-\tom{Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction.
-}{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.} \linelabel{res:herbivory}
-Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences.
-As more long-term data accumulate, we may gain more information with which to further unpack this group of species.
+Second, we have not explicitly incorporated ant-herbivore interactions, even though these are the primary pathway through which ants influence plant demography. 
+Surveys of herbivore damage from our long-term data (Figure \ref{app:herb}) support the protective benefits of ant visitation. 
+Specifically, herbivore pressure was highest on plants in which no ant was present (vacancy), intermediate under Other and \textit{C.
+opuntiae} tending, and lowest under \textit{L.
+apiculatum} tending.
+This gradient aligns with the expected defensive hierarchy of ant species, providing indirect evidence that differences among ant partners influence herbivory rates and, by extension, plant demographic outcomes \citep{Miller2007,Ohm2014}.
+Future work explicitly incorporating ant-herbivore-plant interactions in the demographic models could more directly quantify how variation in ant species composition mediates herbivore pressure and plant fitness.
+\linelabel{res:herbivory}
+Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an Other category that might mask important differences.
+As more long-term data accumulate, we may gain additional information to further unpack this group of species.}
 
@@
 J. Moutouama, A. Bradley, B. Scherick, A. Gough, J. Martin, and V. Rudolf provided helpful feedback on the manuscript. 
-\tom{}{Never forget to thank reviewers in the acknowledgments of resubmissions!}
+We also thank the reviewers for taking the time to provide thoughtful comments and suggestions on earlier versions of this manuscript.
 
@@
 
-\section*{Appendix E}\label{appendix:E}
+\revise{\section*{Appendix E}\label{appendix:E}
 To further evaluate the potential for a portfolio effect in this system, we conducted an additional analysis in which we artificially increased the degree of difference among ant partners. 
@@
 This shows that portfolio effect is possible in this system if the ant partners diverged more greatly in their responses to annual variation. \linelabel{res:portfolio_evidence}
+}
 

### 2025-08-28 Manuscript/ant_cactus_SUBMISSION2.tex
intro

@@
 \linelabel{res:27}
-While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal species interacts with multiple partner species.
-These broader, network-structured interactions are characteristic of generalized or diffuse mutualisms, in which mutualists interact with multiple partners rather than engaging in strict one‐to‐one pairings \citep{Aizen2014, Prior2015}.
+While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal mutualist interacts with multiple partner species.
+%These broader, network-structured interactions are characteristic of generalized or diffuse mutualisms, in which mutualists interact with multiple partners rather than engaging in strict one‐to‐one pairings \citep{Aizen2014, Prior2015}.
+%tom: I would avoid discussion of 'networks'
 Growing evidence suggests that pairwise interactions may be poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}.
+
 A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species).
@@
 Third, partner species can have species-specific responses to environmental variation, either spatially \citep{Ollerton2006} or temporally \citep{Alarcon2008}. 
-Multiple partners can therefore act as a 'portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
+Multiple partners can therefore act as a `portfolio' that stabilizes fitness benefits across spatial or temporal heterogeneity, leading to positive effects of partner diversity through the portfolio effect \citep{Batstone2018,Lazaro2022,Horvitz1990}.
 In the context of temporally fluctuating environments, the portfolio effect implies that inter-annual variation in partner identity could stabilize overall plant fitness across years.
@@
 Finally, herbivore identity and pressure can vary inter-annually \citep{Wetzel2023}, much like mutualist identity and presence, meaning the threat plants face can vary just as much as the protection they receive due to temporal stochasticity.
-\revise{Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover and temporal stochasticity, particularly because long-term demographic studies on ant-plant interactions are exceedingly rare \cite{Donald2022,Moraes2009}.
+\revise{Previous studies have investigated how ant partner diversity affects plant fitness \citep{Palmer2010,Fiala1994,Gaume1998,Dattilo2014,Ludka2015}, however little is known about the combined effects of partner identity, directional partner turnover, and temporal stochasticity, particularly because long-term demographic studies on ant-plant interactions are exceedingly rare \cite{Donald2022,Moraes2009}.
 }
@@
 Quantifying contributions from the portfolio effect therefore requires long-term data on partner (a)synchrony across inter-annual fluctuations. 
-In practical terms, evidence for a portfolio effect would include: 
-(i) partner species showing asynchronous variation in their benefits to the focal host across years, and 
-(ii) the combined effect of these fluctuations must increase the net benefits of partner diversity to the host relative to the variance expected if all partners responded in parallel.
+%In practical terms, evidence for a portfolio effect would include: 
+%(i) partner species showing asynchronous variation in their benefits to the focal host across years, and 
+%(ii) the combined effect of these fluctuations must increase the net benefits of partner diversity to the host relative to the variance expected if all partners responded in parallel.
 \linelabel{res:portfolio}
-Detecting these patterns typically involves either correlation analyses of inter-annual partner benefits, or null-model simulations comparing observed host fitness variance to expectations under random or synchronous partner turnover.
+%Detecting these patterns typically involves either correlation analyses of inter-annual partner benefits, or null-model simulations comparing observed host fitness variance to expectations under random or synchronous partner turnover.
 Finally, alternative partners may have different demographic effects (e.g., one that enhances survival and another that enhances reproduction), and these effects may contribute differently to overall fitness of a focal mutualist.
@@
 \revise{We then built a stochastic, multi-state integral projection model (IPM) which combines diverse effects on vital rates and pathways of partner turnover. 
-Finally, we used the model to conduct simulation `experiments' that quantify plant fitness under different scenarios of partner richness, composition, and synchrony.}
+Finally, we used the model to quantify plant fitness under different scenarios of partner richness, composition, and synchrony.}
 

### 2025-08-28 Manuscript/Ant_Cactus_RESPONSE.tex
tom's edits on resubmission

@@
 \response{These are great suggestions -- thank you.
-We have revised the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data3} and ~\ref{ms-res:long-term-data2}).
+We have revised the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data2}).
 We have also expanded on why the quantification of the complete demographic effects of even single ant partners is a unique and valuable contribution to the field (line ~\ref{ms-res:L111})
@@
 	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines \ref{ms-res:portfolio_evidence} and Figure \ref{ms-app:portfolio}).
-	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{ms-res:portfolio_asynchronous}).
+	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute evidence for this mechanism (lines \ref{ms-res:portfolio_asynchronous}).
 	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{ms-res:portfolio_methods}). }
@@
 \comment{Lines 329 and 364: I realize that this is commonplace in some parts of the literature, but I would prefer not to call different simulation scenarios “experiments”}
-\response{We have changed the language throughout to "simulations" or "scenarios" rather than "experiments".}
+\response{We have changed the language throughout to ``simulations'' or ``scenarios'' rather than ``experiments''.}
 
@@
 \response{Thank you for pointing this out.
-We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{res:portfolio_methods}, \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}
+We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}
 
@@
 We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}). 
-******** Tom have you already done this?
-Or is this something you would like me to do?}
+%******** Tom have you already done this? Or is this something you would like me to do?
+}
 
@@
 \response{Thank you for the helpful suggestion. 
-We have streamlined the introduction paragraph by removing extraneous details after the first sentence and opened the section with the content previously in paragraph 2. 
-This revision improves the focus of the introduction on the core aspects of our study. 
-Additionally, we have expanded the discussion on generalized/diffuse mutualisms}, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).
+We have streamlined the introduction paragraph by removing extraneous details after the first sentence. We now open with the content previously in paragraph 2.} 
+%Additionally, we have expanded the discussion on generalized/diffuse mutualisms, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).
 
@@
 \comment{Fig. 5, caption: ". . . based on [plant] size".}
-\response{We have revised the caption of Figure 5 to clarify what is meant by "based on plant size" and to explicitly describe the relationship being visualized (Figure~\ref{ms-fig:Ant_Transition}).}
+\response{We have revised the caption of Figure 5 to clarify what is meant by ``based on plant size'' and to explicitly describe the relationship being visualized (Figure~\ref{ms-fig:Ant_Transition}).}
 
@@
 	We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
-	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).
-	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{ms-res:portfolio_criteria}).}}
+	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_asynchronous}).
+	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{ms-res:portfolio_criteria}).}
 
@@
 \comment{27: Compared to which interactions ("such as…")? I think this intro material obstructs you from rolling out the background we really need.}
-\response{We have revised the first paragraph to reduce unnecessary and/or uninformative information to speed the reader into what is needed to understand the rest of the paper (line \ref{ms-res:L27}).}
+\response{We have heavily revised the first paragraph as described above.}
 

### 2025-08-28 Manuscript/ant_cactus_SUBMISSION2.tex
tom's edits on resubmission

@@
 This study was conducted at the Sevilleta Long-term Ecological Research site (SEV-LTER), located within the Sevilleta National Wildlife Refuge in central New Mexico, USA.
-Our study area in the Los Pi$\tilde{n}$os mountains ($34^\circ20'5.3''$N, $106^\circ37'53.2''$W) is characterized by steep, rocky slopes and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
+Our study area in the Los Pi\~nos mountains ($34^\circ20'5.3''$N, $106^\circ37'53.2''$W) is characterized by steep, rocky slopes and perennial vegetation including grasses (\textit{Bouteloua eriopoda} and \textit{B.
 gracilis}), yuccas, cacti, and junipers.
@@
 This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
-Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
+Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown).
 \linelabel{res:repro1}
-For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
+For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm).
 Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. \linelabel{res:repro2}
@@
 Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
+We also quantified the abundances and identities of any cactus herbivores observed, and recorded the occurrence of herbivore damage (described further in Appendix B). 
 \linelabel{res:exhaustive} Each year, the field team searched exhaustively for new recruits by walking slowly in a standard formation across the plot area, and these were added to the census (cactus seedlings are small ($<2 cm$ in height) and cryptic so it is likely that some are missed even with a standardized search protocol).
@@
 \revise{\linelabel{res:model_story}The biological ``story'' of the model is as follows: 
-The tree cholla population is comprised of plants characterized by their size, which can take any continuous value between the lower and upper limits, and ant visitation status, which can take one of the four discrete levels described above.
+The tree cholla population is comprised of plants characterized by their size, which can take any continuous value between lower and upper limits, and ant visitation status, which can take one of the four discrete levels described above.
 An individual's size and ant status jointly predict its demographic fate such as survival, growth, and reproductive contributions to the seed bank. 
@@
 We tallied the total population size at each time step as  $N_{t} = B^1_{t} + B^2_{t} + \sum_{a=1}^{A}\int n(x,a)_{t}dx$ and calculated the stochastic growth rate as 
-$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \citep{Rees2009}
+$$log(\lambda_S) = E[log(\frac{N_{t+1}}{N_{t}})]$$ \citep{Rees2009}.
 We propagated uncertainty from the vital rate models using 100 draws from the joint posterior distribution of model parameters, resulting in a posterior distribution of $\lambda_{S}$ and the derived quantities described next.
@@
 At the largest sizes, growth trajectories of ant-tended and vacant plants were nearly indistinguishable. 
-For all ant states, growth was left-skewed, with transitions to sizes below the mean were more frequent than sizes above the mean.
+For all ant states, growth was left-skewed, with transitions to sizes below the mean being more frequent than sizes above the mean.
 
@@
 	Dashed lines indicate extrapolation while solid lines indicate the range of observed data.
-	They gray dashed line indicates no growth (areas above this line show growth while areas under it show shrinkage).}
+	They gray dashed line indicates no change in size (areas above this line show growth while areas under it show shrinkage).}
 	\label{fig:Grow}
@@
 \caption{Probability of survival in relation to size and ant partner.
-These are the posterior probability distributions.
 Layout as in Fig.
@@
 \revise{Furthermore, \textit{C.
-opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
+opuntiae}-tended plants had lower mean seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L.
 apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.
 \linelabel{res:viability}}
-\ali{}{I am not sure where in the appendix you want this to be referenced?}
+\ali{}{I am not sure where in the appendix you want this to be referenced?
+--TOM: I don't know what you are asking.
+There is a figure of the posterior distribution for the means.
+YOu should reference the figure.}
 
@@
 	\caption{Flowerbud viability in relation to ant partner status.
-Points show observed data (fraction of initiated buds that flower).
+Points show observed data (fraction of initiated buds that were viable).
 Points and bars show the means and 95\% credible intervals for each group.}
@@
 opuntiae} and Other-tended plants; mean pairwise correlation: -0.0016) while others were almost completely synchronous (such as growth of Other and \textit{L.
-apiculatum}-tended plants; mean pairwise correlation: 0.093).
+apiculatum}-tended plants; mean pairwise correlation: \tom{0.093}{SHould this be 0.93?}).
 
@@
 opuntiae}, the probability that those plants remain occupied by \textit{C.
-opuntiae} at the end of the transition year is only slightly greater than the probability of take-over by \textit{L.
+opuntiae} at the end of the transition year was only slightly greater than the probability of take-over by \textit{L.
 apiculatum}, while take-over in the other direction was extremely rare.
@@
 Furthermore, we found no benefit of partner diversity, with the fitness associated with one, two, or three partners roughly equivalent (Figure \ref{fig:LambdaMeans} a).
-The one- and two-partner scenarios were statistically consistent but there was a modest reduction in fitness in the three-partner scenario compared to two partners ($Pr(\Delta\lambda^{2-3}_{S}=0.95)$), consistent with a weak cost of diversity at the highest level.
+The one- and two-partner scenarios were statistically consistent but there was a modest reduction in fitness in the three-partner scenario compared to two partners ($Pr(\Delta\lambda^{2-3}_{S} > 0)=0.95$), consistent with a weak cost of diversity at the highest level.
 Patterns of $\lambda_{S}$ in relation to partner richness were generally consistent between scenarios of non-synchronous and synchronous inter-annual fluctuations. 
@@
 apiculatum}-only population had the highest mean fitness, it was not significantly higher than \textit{C.
-opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) only populations.
-Furthermore, the fitness of \textit{L. apiculatum}-only was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
-However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$). 
+opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}>0)=0.7$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}>0)=0.69$) only populations.
+Furthermore, the fitness of \textit{L. apiculatum}-only was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S}>0)$ ranged from 0.37 to 0.83).
+However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S}>0)>0.95$).
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
-\revise{Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above. \linelabel{res:appendixc}}
+\revise{\tom{Using the simulations where all ants had equal frequencies across sizes}{Can you make this more helpful to readers who have not (and may not) see the appendix?
+It is not clear how ``equal frequencies across sizes'' is relevant here or how it should be interpreted.} (further explained in Appendix C), we found the same fitness patterns as described above.
+\linelabel{res:appendixc}}
 %Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
@@
 The effect of all ant partners can be measured as $\lambda^{3}_{S} - \lambda^{0}_{S}$.
-We have high confidence that this contrast is positive, and equally so for the synchronous and non-synchronous scenarios (Figure S\ref{fig:Portfolio}). 
+We have high confidence that this contrast is positive and equally so for the synchronous and non-synchronous scenarios (Figure S\ref{fig:Portfolio}), indicating that the full suite of partners is similarly beneficial with and without the portfolio effect operating.
 \revise{\linelabel{res:portfolio_results}In Appendix D, we show that contributions from the portfolio effect were constrained by positively correlated fluctuations between ant states (Figure \ref{fig:Annual_Ant}).
-In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model (Figures), as expected under the portfolio effect hypothesis.}
+In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model \tom{(Figures)}{Update this.}, as expected under the portfolio effect hypothesis.}
 
@@
 Furthermore, evidence from nearly two decades of demographic data failed to support the ``portfolio effect'' hypothesis, whereby benefits of diversity manifest in the context of environmental fluctuations, \revise{because fluctuations in the demographic effects of alternative partners are too synchronized}.
-Overall, our results indicate that while ant visitation is strongly consequential for plant demography, partner composition is much less so, suggesting that the fitness benefits of mutualism from the plant's perspective may be surprisingly resilient to declines in partner diversity.
+Overall, our results indicate that while ant visitation is strongly consequential for plant demography, partner composition is much less so, suggesting that the fitness benefits of mutualism from the plant's perspective may be surprisingly resilient to changes in partner diversity.
 
@@
 Differences between alternative partners were most pronounced either in reproductive rates (floral viability and seed number), which contribute relatively weakly to fitness because a long reproductive lifespan overrides individual reproductive bouts, or in growth and survival at small sizes, where mortality risk is relatively high.
-At larger sizes with low mortality risk, ant-tended plants had consistent growth and survival advantages over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness; differences between partner species in other vital rates and at other sizes do not register nearly as strongly.
+At intermediate sizes, ant-tended plants had consistent growth and survival advantages over vacant plants regardless of partner identity, and that result dominates our integrative measures of fitness; differences between partner species in other vital rates and at other sizes do not register nearly as strongly.
 The weak reduction in fitness at the highest partner diversity level is harder to explain, but suggests some form of missed opportunity cost in a more crowded partner environment.
@@
 In ant-plant defensive mutualisms, any portfolio effect would emerge from fluctuations in the community of herbivores, and ant-specific responses to each herbivore species. 
-Instead, we found that inter-annual fluctuations in vital rates were largely synchronized across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population ( notably, weather) override any fluctuations in herbivore composition and ant-herbivore interactions.
-To further explore the potential for portfolio effects, we conducted simulations in which ant responses to herbivory were artificially manipulated to be independent (correlation = 0) or anti-synchronous (correlation = -1).
-We found evidence of the portfolio effect (the net benefit of partner diversity was greater when ants responded to inter-annual fluctuations non-synchronously than when they responded synchronously) in both scenarios.
-These results were particularly strong when the ants were anti-synchronous.
-These results indicate that while the portfolio effect can emerge under conditions of asynchronous partner responses, natural drivers in our study system result in synchronous partner effects, masking the potential benefits of partner diversity.
+Instead, we found that inter-annual fluctuations in vital rates were generally positively correlated across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population (notably, weather) override any fluctuations in herbivore composition and ant-herbivore interactions.
+\revise{To further explore the potential for portfolio effects, we conducted simulations in which inter-annual fluctuations in ant effects were manipulated to be independent or negatively correlated; we found evidence of the portfolio effect in both scenarios, but especially when ant effects were negatively correlated (Appendix E).
+These results indicate that while the portfolio effect can emerge under hypothetical conditions of anti-synchronous partner responses, shared drivers of inter-annual variability in our study system dampen this potential benefit of partner diversity.
+}
 Dallas et al. \citeyear{dallas2022temporal} found that while the portfolio effect was easy to show in theoretical models, it is infrequently detected in empirical data across many systems. 
-This indicates that the portfolio effect may be difficult to detect (particularly in similar systems where ant differences are only somewhat non-synchronous), disguised by different mechanisms, or uncommon in nature.
+This indicates that the portfolio effect may be difficult to detect, disguised by other mechanisms, or uncommon in nature.
 
@@
 J. Moutouama, A. Bradley, B. Scherick, A. Gough, J. Martin, and V. Rudolf provided helpful feedback on the manuscript. 
-We also thank the reviewers for taking the time to provide thoughtful comments and suggestions on earlier versions of this manuscript.
+We also thank two reviewers for taking the time to provide thoughtful comments and suggestions on earlier versions of this manuscript.
 
@@
 
-\paragraph{Portfolio Effect}
+\tom{\paragraph{Portfolio Effect}
 We compared the effects of partner diversity (estimated as $\lambda_S^4 - \lambda_S^0$) for the synchronized and non-synchronized models.
 We determined that non-synchronous effects of ant partners are greater than synchronous effects of ant partners only 50\% of the time.
-This means there is no evidence of portfolio effect in this system.
+This means there is no evidence of portfolio effect in this system.}{I do not udnerstand why this is here. It does not fit in an appendix about Additional Methods and Parameters.}
 
@@
 Here we conduct  an analyses of herbivory levels on plants with different ant visitors and contrast them to herbivory levels on plants with no visitors.
-Evidence that herbivory is decreased for tended plants further strengthens the results presented in this paper that ant presence is beneficial on a demographic level for cholla.
+Evidence that herbivory is decreased for tended plants further strengthens the results presented in the main text that ant presence is beneficial.
 
-Herbivory data was collected during censuses any time herbivores were identified on a plant. 
+Herbivory data was collected during annual demographic censuses. 
 This involved noting the type and quantity of herbivores observed. 
@@
 The proportion of reproducing plants that experienced herbivory was calculated for each ant state separately.
-Analysis showed that ant presence is correlated with lower herbivore visitation.
-40\% of vacant cacti experienced herbivory.
+\tom{Analysis showed that ant presence is correlated with lower herbivore visitation.
+40\% of vacant cacti experienced herbivory.}{It is unclear what quantity is shown in the figure. It is important to clarify this.}
 Plants tended by Other ants experienced similar, though lower, levels of herbivory on reproducing plants, with herbivores detected on 37.5\% of plants.
@@
 They also indicate that the partner correlated with the lowest levels of herbivory is \textit{L. apiculatum} ants.
-These findings are consistent with literature findings which show that \textit{L.
-apiculatum} ants are the most aggressive (therefore the most effective against herbivores), but differ from previous findings that \textit{C.
-opuntiae} may not offer defensive benefits \citep{Miller2007}.
+These findings are consistent with literature findings which show that \textit{L. apiculatum} ants are the most aggressive (therefore the most effective against herbivores) \citep{Miller2007}.
 
@@
 \revise{\section*{Appendix E}\label{appendix:E}
-To further evaluate the potential for a portfolio effect in this system, we conducted an additional analysis in which we artificially increased the degree of difference among ant partners. 
+To further evaluate the potential for a portfolio effect, we conducted an additional analysis in which we increased the degree of difference among ant partners. 
 Specifically, we manipulated the correlation structure of ant-specific year random effects in the stochastic IPM. 
-In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
-In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.
+\tom{In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
+In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.}{Need to say more about how you did this.}
 
-For each of these scenarios, we quantified the effect of partner diversity as the difference between projected population growth with and without partner diversity ($\lambda_S^4 - \lambda_S^0$), and compared results under synchronous versus non-synchronous environmental variation.
-Both scenarios resulted in greater benefits of partner diversity under Non-synchronous conditions, though the evidence was much stronger when the ants were anti-correlated.
-
+For each of these scenarios, we quantified the effect of partner diversity as the difference between projected population growth with no partners and with the full suite of partner diversity ($\lambda_S^4 - \lambda_S^0$), and compared results under synchronous versus non-synchronous environmental variation.
+Both scenarios resulted in greater benefits of partner diversity under non-synchronous conditions, though the evidence was much stronger when the ants were anti-correlated, as expected under the portfolio effect hypothesis.
 This shows that portfolio effect is possible in this system if the ant partners diverged more greatly in their responses to annual variation. \linelabel{res:portfolio_evidence}

### 2025-08-28 Manuscript/ant_cactus_SUBMISSION2.tex
fixing appendix lettering

@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thefigure}{B\arabic{figure}}
 
@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thefigure}{C\arabic{figure}}
 
@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
-
 \begin{figure}
@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thefigure}{D\arabic{figure}}
 
@@
 \end{figure}
-
-\renewcommand{\thefigure}{A\arabic{figure}}
 
@@
 %% Seeds Produced
-\paragraph{Seeds Per Flower Model}
-
-\renewcommand{\thefigure}{A\arabic{figure}}
+\paragraph{Seeds Per Fruit Model}
 
@@
 opuntiae} tended plants produced fewer seeds per flower than \textit{L.
-apiculatum} tended plants and vacant plants 80\% and 87\% of the time.
-Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
+apiculatum} tended plants and vacant plants with 80\% and 87\% confidence, respectively.
+%Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
 
@@
 	Seeds have a slightly higher probability of germinating in year one than in year two.
-	Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
-	Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.
+	\tom{Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
+	Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.}{I do not understand what these ranges are.}
 
@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thefigure}{E\arabic{figure}}
 

### 2025-08-29 Manuscript/Ant_Cactus_RESPONSE.tex
Updates

@@
 That is, we have built a general model of demographic transitions (births, deaths, growth, and partner turnover) that is realistic with respect to our study system, but we query the model for qualitative inference (e.g., asking whether a multi-partner scenario yields higher or lower fitness than the single best partner) and not quantitative precision.
-We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}). 
-%******** Tom have you already done this? Or is this something you would like me to do?
+We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{ms-res:gen1}, \ref{ms-res:model_story}). 
 }
@@
 \comment{A bit jarring out the gate. The first sentence states: everyone in mutualism benefits (full stop). The second sentence is : except when they don't.}
-\response{We have revised the first sentence (line \ref{ms-res:R2_1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.}
+\response{We have revised the first sentence (line \ref{ms-res:L1}) to reflect that in mutualisms everyone receives a net benefit rather than just saying everyone benefits.}
 

### 2025-08-29 Manuscript/ant_cactus_SUBMISSION2.tex
Updates

@@
 % 200 words max
-Mutualisms commonly involve multiple partner species but the ecological consequences of partner diversity remain poorly understood.
+Mutualisms commonly involve multiple partner species but the ecological consequences of partner diversity remain poorly understood. \linelabel{res:L1}
 Partner diversity may benefit a focal mutualist through mechanisms that mirror positive effects of species diversity on ecosystem function, such as sampling effect, complementarity or portfolio effect.
@@
 %% Look at mechanisms which show how diversity impacts demography
-There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020,Winfree2020}.
+\revise{There are multiple mechanisms by which partner diversity can influence the net benefits accrued by a focal mutualist, mirroring the mechanisms by which, at a larger scale of organization, biodiversity can influence ecosystem function \citep{Barrett2015,Ushio2020,Winfree2020}.
+\linelabel{res:mechanims_math}
 First, when there is a consistent ranking of best to worst mutualists, a more diverse sample of the partner community may be more likely to include the best partner \citep{Frederickson2013}.
+\linelabel{res:L45}
 This can lead to an apparent benefit of diversity driven by a sampling effect \citep{Batstone2018}.  
 However, if partner associations are mutually exclusive (a focal mutualist may engage with only one partner at a time), then partner diversity may impose opportunity costs, leading to negative effects of a diverse mutualist assemblage relative to exclusive association with the single best partner \citep{Miller2007}.
-\revise{Second, even within a mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
+Second, even within a mutualist guild, the benefits conferred by alternative partner species can vary by method, efficiency, or timing \citep{Stachowicz2005,Bronstein2006,Stanton2003}.
 \linelabel{res:L53}
@@
 We constructed a stochastic, multi-state IPM that stitches together population structure associated with plant size and ant state, allowing us to determine the individual fitness effects of each ant species in isolation and the composite effects of multiple partners, with ant transition dynamics and inter-annual variability modeled explicitly.
+\linelabel{res:gen1}
 
@@
 
-Ant state is included as a predictor variable in IPM sub-models where there are biologically realistic pathways through which ants could have an impact. 
+Ant state is included as a predictor variable in IPM sub-models where there are biologically realistic pathways through which ants could have an impact. \linelabel{res:gen2}
 For example, prior experimental work indicates that ant tending can reduce vegetative tissue loss and floral abortion, while also potentially reducing pollinator visitation \citep{Miller2009,Ohm2014}.
@@
 opuntiae} and Other-tended plants; mean pairwise correlation: -0.0016) while others were almost completely synchronous (such as growth of Other and \textit{L.
-apiculatum}-tended plants; mean pairwise correlation: \tom{0.093}{SHould this be 0.93?}).
+apiculatum}-tended plants; mean pairwise correlation: 0.93.
 
 \begin{figure}[H]
-	\includegraphics[width=0.75\linewidth]{Figures/ant_RFX.png}
-	\caption{Year-specific random effects from statistical models for growth (a), survival (b), and floral viability (c) in relation to ant partner status.
-Inset panels show correlation coefficients (Pearson's \textit{r}) for each pair of time series.
+	\includegraphics[width=0.75\linewidth]{Figures/corr_RFX.png}
+	\caption{
+		The correlation coefficients (Pearson's \textit{r}) of ant-specific year-random effects from statistical models for growth (a), survival (b), and floral viability (c).
+		Dark red indicates negative correlation, dark blue indicates positive correlation, white indicates no correlation.
+		The year-specific random effects for each ant state  from the statistical models are shown in a time series for growth (d), survival (e), and floral viability (f).
 		All random effects have mean zero such that positive values indicate better-than-average years and negative values indicate worse-than-average years. 
@@
 The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}.
-\revise{\tom{Using the simulations where all ants had equal frequencies across sizes}{Can you make this more helpful to readers who have not (and may not) see the appendix?
-It is not clear how ``equal frequencies across sizes'' is relevant here or how it should be interpreted.} (further explained in Appendix C), we found the same fitness patterns as described above.
+
+\revise{Using simulations in which all ant species were artificually forced to be equally frequent (further explained in Appendix C), we found the same fitness patterns as described above.
 \linelabel{res:appendixc}}
@@
 \revise{\linelabel{res:portfolio_results}In Appendix D, we show that contributions from the portfolio effect were constrained by positively correlated fluctuations between ant states (Figure \ref{fig:Annual_Ant}).
-In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model \tom{(Figures)}{Update this.}, as expected under the portfolio effect hypothesis.}
+In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model (Figure \ref{app:portfolio}), as expected under the portfolio effect hypothesis.}
 
@@
 \section*{Appendix A: Additional Methods and Parameters} \label{appendix:A}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
 In addition to the models described in the body of the paper, we fit several other models using data from previous studies.
@@
 \paragraph{Seeds Per Fruit}
-With data from \cite{Miller2006}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+With data from \cite{Ohm2014}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
 We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
@@
 The data used for this model did not include data on ants in the ``other" category, so we used the data from vacant plants to parameterize seeds per flower for plants with ``other" ants in the IPM.
+
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
+	\caption{The posterior distributions of seeds produced per fruit on reproducing plants.
+		These estimates are ant specific with the yellow representing vacant plant seed production, purple representing \textit{C.
+opuntiae} tended plant seed production, and teal representing \textit{L.
+apiculatum} tended plant seed production.
+		 }
+	\label{app:seeds}
+\end{figure}
 
@@
 
-\tom{\paragraph{Portfolio Effect}
-We compared the effects of partner diversity (estimated as $\lambda_S^4 - \lambda_S^0$) for the synchronized and non-synchronized models.
-We determined that non-synchronous effects of ant partners are greater than synchronous effects of ant partners only 50\% of the time.
-This means there is no evidence of portfolio effect in this system.}{I do not udnerstand why this is here. It does not fit in an appendix about Additional Methods and Parameters.}
+%% Reproduction Model
+\paragraph{Probability of Flowering}
+We fit a model for the probability of flowering ($F$) in year $t+1$ based on the size ($x$) in year $t+1$ with random effects of year $\tau_a$ using a Bernoulli distribution with a logit link function:
+$$y^{F}~ Bern(\hat{F})$$
+$$\hat{F_i} = \beta^0 x_i + \tau_{a[i]}$$
 
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
-	\caption{The distributions of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in pink and green respectively.
-The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
-	\label{fig:Portfolio}
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/repro_panel.png}
+	\caption{The probability of a cholla plant reproducing based on its size.
+The green line shows the average probability, the gray surrounding area shows the 90\% confidence interval, and the black points show the actual data.
+}
+	\label{app:repro}
 \end{figure}
 
+\begin{figure}
+	\includegraphics[width=0.91\linewidth]{Figures/flow.png}
+	\caption{The number of flowers produced by a cholla plant based on its size.
+The green line shows the average number of flowers produced, the gray surrounding area shows the 90\% confidence interval, and the green points show the binned data.
+}
+	\label{app:flow}
+\end{figure}
+
+The probability of a plant reproducing in a given year is highly size dependent. 
+The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes (Figure \ref{app:repro}).
+
+The number of flowers produced by a reproducing plant is also highly size dependent. 
+The mean number of flowers produced are 0 until the plant reaches a medium size, after which the mean number of flowers produced increases to about 45 at the largest sizes (Figure \ref{app:flow}).
+
 \section*{Appendix B: Observed Herbivory Data} \label{appendix:B}
+\renewcommand{\thefigure}{B\arabic{figure}}
+
 Herbivory is an important driver in this population and shapes the range and demography of cholla.
@@
 This involved noting the type and quantity of herbivores observed. 
-This data has been taken consistently since 2017, so the analysis below considers 6 years of data.
+Presence of herbivores is considered to be herbivore pressure as it is assumed they will damage the plant if they are present.
+This data has been taken consistently since 2014, so the analysis below considers 9 years of data.
 We considered only plants which were reproducing, since prior work shows that flowering significantly elevates herbivore pressure \citep{Miller2006}.
 The proportion of reproducing plants that experienced herbivory was calculated for each ant state separately.
-\tom{Analysis showed that ant presence is correlated with lower herbivore visitation.
-40\% of vacant cacti experienced herbivory.}{It is unclear what quantity is shown in the figure. It is important to clarify this.}
-Plants tended by Other ants experienced similar, though lower, levels of herbivory on reproducing plants, with herbivores detected on 37.5\% of plants.
+Analysis showed that ant presence is correlated with lower herbivore visitation.
+We observed herbivores on 40\% of vacant cacti.
+Plants tended by Other ants experienced similar, though lower, frequency of observed herbivores on reproducing plants (herbivores detected on 37.5\% of plants).
 Herbivores were detected on 25\% of plants tended by \textit{C. opuntiae} ants and on 11\% of plants tended by \textit{L. apiculatum} ants.
@@
 
-\renewcommand{\thefigure}{B\arabic{figure}}
-
 \begin{figure}[H]
 	\includegraphics[width=0.7\linewidth]{Figures/herb_all.png}
-	\caption{The proportion of reproducing plants which are visited by herbivores. Each bar represents the subset of the cacti population in a different ant state.  }
+	\caption{The proportion of reproducing plants which are visited by herbivores.
+	Each bar represents the observed proportion of cacti in each ant state with observed herbivores.}
 	\label{app:herb}
@@
 \section*{Appendix C: Alternative Ant Transition Simulations}\label{appendix:C}
+\renewcommand{\thefigure}{C\arabic{figure}}
+
 In addition to the competitive exclusion model defined and analyzed in the body of the paper, we simulated results from several other potential models. 
@@
 (Figure \ref{app:FreqLambdaMeans}b).
-
-\renewcommand{\thefigure}{C\arabic{figure}}
 
@@
 
-\section*{Appendix D}\label{appendix:D}
-Below are the results reported of all statistical models not described in the main body of the text.
-
-%% Reproduction Model
-\paragraph{Reproduction Model}
-
+\section*{The Portfolio Effect}\label{appendix:D}
 \renewcommand{\thefigure}{D\arabic{figure}}
 
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/repro_panel.png}
-	\caption{The probability of a cholla plant reproducing based on its size.
-The green line shows the average probability, the gray surrounding area shows the 90\% confidence interval, and the black points show the actual data.
-}
-	\label{app:repro}
+This appendix includes our evaluation of evidence for the portfolio effect under realistic conditions (those observed in the field) as well as simulated conditions.
+
+\paragraph{Portfolio Effect Evidence}
+	We compared the effects of partner diversity (estimated as $\lambda_S^4 - \lambda_S^0$) for the synchronized and non-synchronized models.
+	We determined that non-synchronous effects of ant partners are greater than synchronous effects of ant partners only 50\% of the time.
+	This means there is no evidence of portfolio effect in this system.
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{Figures/portfolio_effect.png}
+	\caption{The distributions of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in pink and green respectively.
+The vertical dashed line shows where the effect of partners on the fitness of the cholla is 0 (to the left the partners have a negative effect, to the right the partners are beneficial).}
+	\label{fig:Portfolio}
 \end{figure}
 
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/flow.png}
-	\caption{The number of flowers produced by a cholla plant based on its size.
-The green line shows the average number of flowers produced, the gray surrounding area shows the 90\% confidence interval, and the green points show the binned data.
-}
-	\label{app:flow}
-\end{figure}
-
-The probability of a plant reproducing in a given year is highly size dependent. 
-The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes (Figure \ref{app:repro}).
-
-The number of flowers produced by a reproducing plant is also highly size dependent. 
-The mean number of flowers produced are 0 until the plant reaches a medium size, after which the mean number of flowers produced increases to about 45 at the largest sizes (Figure \ref{app:flow}).
-
-%% Seeds Produced
-\paragraph{Seeds Per Fruit Model}
-
-\begin{figure}
-	\includegraphics[width=0.91\linewidth]{Figures/Seeds_Per_Fruit.png}
-	\caption{The mean number of seeds produced per fruit by the ant partner present.
-The teal,purple, and yellow lines show the mean seeds produced by each flower on plants tended by \textit{L.
-apiculatum} ants, \textit{C.
-opuntiae} ants, and no ants respectively.
-}
-	\label{app:seeds}
-\end{figure}
-
-Each viable flower on a plant produces between 97 and 257 seeds (Figure \ref{app:seeds}).
-This number is affected by the ant partner present, as shown in previous work \citep{Ohm2014}. 
-\textit{C. opuntiae} tended plants produce a mean of 115 seeds per flower. 
-\textit{L. apiculatum} tended plants produce a mean of 143 seeds per flower.
-Vacant plants produce a mean of 148 seeds per flower. 
-Comparison between posterior distributions revealed that \textit{C.
-opuntiae} tended plants produced fewer seeds per flower than \textit{L.
-apiculatum} tended plants and vacant plants with 80\% and 87\% confidence, respectively.
-%Vacant plants produced more seeds per flower than \textit{L. apiculatum} tended plants only 57\% of the time.
-
-%% Precensus Survival\
-	\paragraph{Pre-census Survival Model}
-	Pre-census seed survival rates fall between 0\% and 95\% with the mean pre-census seed survival at 18\%.
-
-	%% Germination
-	\paragraph{Germination Model}
-	Seeds have a slightly higher probability of germinating in year one than in year two.
-	\tom{Seeds in year one experience germination rates between 50\% and 100\% with a mean of 62\% germination.
-	Seeds in year two experience germination rates between 50\% and 98\% with a mean of 58\% germination.}{I do not understand what these ranges are.}
-
-	%% Recruit size distribution
-	New recruits are expected to be between the sizes of 0.11 $cm^3$ and 0.38 $cm^3$ with a mean size of 0.20 $cm^3$.
-
-\revise{\section*{Appendix E}\label{appendix:E}
 To further evaluate the potential for a portfolio effect, we conducted an additional analysis in which we increased the degree of difference among ant partners. 
 Specifically, we manipulated the correlation structure of ant-specific year random effects in the stochastic IPM. 
-\tom{In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
-In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.}{Need to say more about how you did this.}
+In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
+In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.
+To reach a zero correlation and a maximally negative one, we reordered the vector of mean ant-specific year random effects in each year so that high values for one ant corresponded to low values for another, and vice versa, while preserving the original distribution of effects.
+We continued reordering each vector until the correlations reached the desired values. 
+The resulting manipulated vectors allowed us to evaluate the fitness of the cacti under different partner diversity scenarios when the ants experienced different asynchronous responses to inter-annual fluctuations, allowing us to determine if portfolio effect was possible under different ant conditions.
 
@@
 This shows that portfolio effect is possible in this system if the ant partners diverged more greatly in their responses to annual variation. \linelabel{res:portfolio_evidence}
-}
-
-\renewcommand{\thefigure}{E\arabic{figure}}
 
@@
 	\includegraphics[width=0.91\linewidth]{Figures/cor0_Portfolio_Effect.png}
-	\caption{The mean estimations of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in green and pink respectively. }
+	\caption{
+	A comparison of synchronous and non-synchronous cactus responses to partner diversity across inter-annual fluctuations. 
+	If the non-synchronous effects of partner diversity are greater than the synchronous ones under fluctuation environments, evidence of the portfolio effect is found.
+	Panel a) shows the mean effects of partner diversity ($\lambda_{S}^4-\lambda_{S}^0$) under realistic levels of non-synchronicity (what was estimated from the field data), while panels b) and c) shows the mean effects of partner diversity under no synchronicity (where we artificially forced a correlation of 0 for the ant-specific year random effects) and negative synchronicity (where we artificially forced a correlation close to -1 for the ant-specific year random effects).
+	The mean estimations of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in green and pink respectively. 
+	}
 	\label{app:portfolio}

### 2025-08-29 05_cholla_ant_IPM_figures_tables.R

@@ -879,7 +879,7 @@ dev.off()
 ################################################################################
 png("Manuscript/Figures/Seeds_Per_Fruit.png")
 plot(density(exp(seed.params$beta0[,3])),lwd=3,col=vaccol,
-     xlab="Mean seeds per fruit",main=" ",xlim=c(50,250))
+     xlab="Seeds Per Fruit",main=" ",xlim=c(50,250))
 lines(density(exp(seed.params$beta0[,2])),lwd=3,col=liomcol)
 lines(density(exp(seed.params$beta0[,1])),lwd=3,col=cremcol)
 legend("topright",legend = c("Vacant",expression(italic("C. opuntiae")),expression(italic("L. apiculatum"))), fill = c(vaccol, cremcol, liomcol))
@@ -1516,7 +1516,7 @@ vac_prop <- sum(vac_herb$herb_YN, na.rm = T)/nrow(vac_herb)
 png("Manuscript/Figures/herb_all.png")
 barplot(height = c(crem_prop,liom_prop,other_prop,vac_prop), names.arg = c(expression(italic("C. opuntiae")),expression(italic("L. apiculatum")),"Other","Vacant"), col = c(cremcol,liomcol,othercol,vaccol))
 mtext("Ant Partner",side=1,line=-1.5,outer=TRUE,cex=1.5)
-mtext("Herbivore Prevalence",side=2,line=-1.5,outer=TRUE,cex=1.5,las=0)
+mtext("Herbivore Presence",side=2,line=-1.5,outer=TRUE,cex=1.5,las=0)
 dev.off()
 
 

### 2025-08-29 06_Portfolio_Sim.R

@@ -969,44 +969,67 @@ dev.off()
 
 
 ######### FULL VISUALS ##############
+####### Real Correlation
+lams_stoch <- read.csv("Model Outputs/lambda_comp.csv")
+lams_stoch <- lams_stoch[,-c(1)]
+## Competitive Exclusion Null Stochastic
+lams_stoch_sync <- read.csv("Model Outputs/lambda_sync_comp.csv")
+lams_stoch_sync <- lams_stoch_sync[,-c(1)]
+
+####### 0 Correlation
 ## Competitive Exclusion
 lams_comp_stoch <- read.csv("Model Outputs/cor0_lambda_comp.csv")
 lams_comp_stoch <- lams_comp_stoch[,-c(1)]
 ## Competitive Exclusion Null Stochastic
 lams_comp_stoch_sync <- read.csv("Model Outputs/cor0_lambda_sync_comp.csv")
 lams_comp_stoch_sync <- lams_comp_stoch_sync[,-c(1)]
-
+####### Negative Correlation
 ## Competitive Exclusion
-lams_comp_stoch_1 <- read.csv("Model Outputs/lambda_comp.csv")
+lams_comp_stoch_1 <- read.csv("Model Outputs/cor_neg_lambda_comp.csv")
 lams_comp_stoch_1 <- lams_comp_stoch_1[,-c(1)]
 lams_comp_stoch_1 <- colMeans(lams_comp_stoch_1)
 lams_comp_stoch_1 <- as.data.frame(t(lams_comp_stoch_1))
 ## Competitive Exclusion Null Stochastic
-lams_comp_stoch_sync_1 <- read.csv("Model Outputs/lambda_sync_comp.csv")
+lams_comp_stoch_sync_1 <- read.csv("Model Outputs/cor_neg_lambda_sync_comp.csv")
 lams_comp_stoch_sync_1 <- lams_comp_stoch_sync_1[,-c(1)]
 lams_comp_stoch_sync_1 <- colMeans(lams_comp_stoch_sync_1)
 lams_comp_stoch_sync_1 <- as.data.frame(t(lams_comp_stoch_sync_1))
-
+x1 <- c(1,2)
 a <- lams_comp_stoch_sync[1,8] - lams_comp_stoch_sync[1,1]
 b <- lams_comp_stoch[1,8] - lams_comp_stoch[1,1]
 c <- lams_comp_stoch_sync_1[1,8] - lams_comp_stoch_sync_1[1,1]
[…]

### 2025-08-29 05_cholla_ant_IPM_figures_tables.R
Update 05_cholla_ant_IPM_figures_tables.R

@@ -879,7 +879,7 @@ dev.off()
 ################################################################################
 png("Manuscript/Figures/Seeds_Per_Fruit.png")
 plot(density(exp(seed.params$beta0[,3])),lwd=3,col=vaccol,
-     xlab="Seeds Per Fruit",main=" ",xlim=c(50,250))
+     xlab="Seeds Per ruit",main=" ",xlim=c(50,250))
 lines(density(exp(seed.params$beta0[,2])),lwd=3,col=liomcol)
 lines(density(exp(seed.params$beta0[,1])),lwd=3,col=cremcol)
 legend("topright",legend = c("Vacant",expression(italic("C. opuntiae")),expression(italic("L. apiculatum"))), fill = c(vaccol, cremcol, liomcol))
@@ -1492,7 +1492,7 @@ herbivory %>%
   group_by(Year_t1) %>%
   summarise(h = sum(herb_YN))
 # use 2014 and on 
-herbivory <- subset(herbivory,herbivory$Year_t1 == 2014)
+herbivory <- subset(herbivory,herbivory$Year_t1 >= 2014)
 
 
 ## Subset by ant
@@ -1808,11 +1808,6 @@ dev.off()
 ################################################################################
 ## Visualize Equal Likelihood Model
 ################################################################################
-
-png("Manuscript/Figures/Lambdas_Equal_lines.png", height = 340, width = 880)
-par(mar=c(5,5,1,1))
-layout(matrix(c(1,2,3),
-              ncol = 3, nrow = 1), heights = c(1,1,1))
 # number of partners vs fitness
 no_part <- colMeans(lams_equal_stoch)[1]
 no_part_low <- min(lams_equal_stoch[,1])
@@ -1838,50 +1833,68 @@ two_part_high_sync <- max(lams_equal_stoch_null[,c(5:7)])
 three_part_sync <- colMeans(lams_equal_stoch_null)[8]
 three_part_low_sync <- min(lams_equal_stoch_null[,8])
 three_part_high_sync <- max(lams_equal_stoch_null[,8])
-plot(c(0.9,1.9,2.9,3.9),c(no_part, one_part, two_part, three_part), pch = 16, cex = 3, 
[…]

### 2025-12-23 Manuscript/ant_cactus_SUBMISSION2.tex
Updates

@@
 Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species.
-We also quantified the abundances and identities of any cactus herbivores observed, and recorded the occurrence of herbivore damage (described further in Appendix B). 
+We also quantified the abundances and identities of any cactus herbivores observed, and recorded the occurrence of herbivore damage (described further in Appendix \ref{appendix:B}). 
 \linelabel{res:exhaustive} Each year, the field team searched exhaustively for new recruits by walking slowly in a standard formation across the plot area, and these were added to the census (cactus seedlings are small ($<2 cm$ in height) and cryptic so it is likely that some are missed even with a standardized search protocol).
@@
 We therefore assumed that plants tended by Other ants had the same number of seeds per fruit as vacant plants (our results showed very little sensitivity to this assumption). 
-Additional reproductive parameters for probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix A. 
+Additional reproductive parameters for probability of entry to the seed bank, germination rates, and recruit size were estimated following methods described in Appendix \ref{appendix:A}. 
 \revise{\linelabel{res:data_origins2}All of these parameters are assumed to be temporally constant.}
@@
 This means that, in scenarios that remove species from the partner community, remaining species gain access to plants that the removed species would have tended. 
-In Appendix C, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
+In Appendix \ref{appendix:C}, we present results under an alternative assumption, that ant visitation is limited by factors other than availability of cactus EFN (e.g., nesting sites or off-plant dietary resources), such that when a species is removed from the partner community, the plants it would have occupied remain vacant.
 
@@
 Both scenarios of temporal stochasticity, non-synchronized and synchronized, were run for all eight ant partner scenarios described above. 
-\revise{\linelabel{res:portfolio_methods}In Appendix D, we explore additional scenarios that vary the correlation of ant-specific fluctuations, to determine if contributions from the portfolio effect are constrained by high correlations seen in the long-term data (Results).}
+\revise{\linelabel{res:portfolio_methods}In Appendix \ref{appendix:D}, we explore additional scenarios that vary the correlation of ant-specific fluctuations, to determine if contributions from the portfolio effect are constrained by high correlations seen in the long-term data (Results).}
 %We indicate stochastic fitness under the synchronized scenario with $\lambda^{C(s)}_{S}$, for example.
@@
 \linelabel{res:viability}}
-\ali{}{I am not sure where in the appendix you want this to be referenced?
---TOM: I don't know what you are asking.
-There is a figure of the posterior distribution for the means.
-YOu should reference the figure.}
 
@@
 
-\revise{Using simulations in which all ant species were artificually forced to be equally frequent (further explained in Appendix C), we found the same fitness patterns as described above.
+\revise{Using simulations in which all ant species were artificually forced to be equally frequent (further explained in Appendix \ref{appendix:C}), we found the same fitness patterns as described above.
 \linelabel{res:appendixc}}
-%Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
-%Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.
-
 We found no evidence that the portfolio effect generated positive effects of partner diversity, as effects of partner richness and composition were highly consistent between the baseline, non-synchronous model and the model version that synchronized all ant states (Figure \ref{fig:LambdaMeans}).
@@
 We have high confidence that this contrast is positive and equally so for the synchronous and non-synchronous scenarios (Figure S\ref{fig:Portfolio}), indicating that the full suite of partners is similarly beneficial with and without the portfolio effect operating.
-\revise{\linelabel{res:portfolio_results}In Appendix D, we show that contributions from the portfolio effect were constrained by positively correlated fluctuations between ant states (Figure \ref{fig:Annual_Ant}).
+\revise{\linelabel{res:portfolio_results}In Appendix \ref{appendix:D}, we show that contributions from the portfolio effect were constrained by positively correlated fluctuations between ant states (Figure \ref{fig:Annual_Ant}).
 In scenarios in which partner fluctuations were uncorrelated or, especially, negatively correlated, the non-synchronous model exhibited a stronger benefit of diversity than the synchronous model (Figure \ref{app:portfolio}), as expected under the portfolio effect hypothesis.}
@@
 Instead, we found that inter-annual fluctuations in vital rates were generally positively correlated across plants in different ant states, suggesting other drivers of inter-annual variation that are shared across the population (notably, weather) override any fluctuations in herbivore composition and ant-herbivore interactions.
-\revise{To further explore the potential for portfolio effects, we conducted simulations in which inter-annual fluctuations in ant effects were manipulated to be independent or negatively correlated; we found evidence of the portfolio effect in both scenarios, but especially when ant effects were negatively correlated (Appendix E).
+\revise{To further explore the potential for portfolio effects, we conducted simulations in which inter-annual fluctuations in ant effects were manipulated to be independent or negatively correlated; we found evidence of the portfolio effect in both scenarios, but especially when ant effects were negatively correlated (Appendix \ref{appendix:D}).
 These results indicate that while the portfolio effect can emerge under hypothetical conditions of anti-synchronous partner responses, shared drivers of inter-annual variability in our study system dampen this potential benefit of partner diversity.
@@
 \section*{Data and Code Availability}
-\tom{This section will be filled out once double blind reviews are complete. }{Be sure to update this with the new data package.}
+This section will be filled out once double blind reviews are complete. 
 %Most of the data that support the findings of this study (through 2019) are openly available as a \href{https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sev.323.1}{data package} \citep{DataCholla} on the Environmental Data Initiative website, package id: knb-lter-sev.323.1.
@@
 \paragraph{Seeds Per Fruit}
-With data from \cite{Ohm2014}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
+With data from \cite{Miller2007}, we fit a model for the number of seeds produced by every fruit on a cholla ($\kappa(a')$) in year $t+1$ based on the ant partner $a'$ in year $t+1$.
 We fit this model to seed data $y^{\kappa}$ using a Negative Binomial distribution and the log link function: 
@@
 \paragraph{Germination}
-With germination data \cite{Miller2007}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
+With germination data \cite{Miller2009}, we fit two models for the probability of germinating from the first year seedbank ($\gamma_1$) or the second year seedbank ($\gamma_2$) in year $t+1$, with no fixed or random effects.
 These models were fit to germination data $y^{\gamma_1}, y^{\gamma_2}$  using the binomial distribution with logit link functions:
@@
 \paragraph{Probability of Flowering}
-We fit a model for the probability of flowering ($F$) in year $t+1$ based on the size ($x$) in year $t+1$ with random effects of year $\tau_a$ using a Bernoulli distribution with a logit link function:
-$$y^{F}~ Bern(\hat{F})$$
-$$\hat{F_i} = \beta^0 x_i + \tau_{a[i]}$$
+We fit a model for the probability of flowering ($P$) in year $t+1$ based on the size ($x$) in year $t+1$ with random effects of year $\tau_a$ using a Bernoulli distribution with a logit link function:
+$$y^{P}~ Bern(\hat{P})$$
+$$\hat{P_i} = \beta^0 x_i + \tau_{a[i]}$$
 
@@
 The mean probability of reproducing remains at about 0\% until the plant reaches a medium size, after which the mean probability of reproducing increases steadily before reaching about 100\% at large sizes (Figure \ref{app:repro}).
+
+\paragraph{Number of Flowers Produced}
+We fit a model for the flowers produced ($F$) in year $t+1$ based on the size ($x$) in year $t+1$ with random effects of year $\tau_a$ using a zero truncated negative binomial distribution: 
+$$y^{F}~ Bern(\hat{F})$$
+$$\hat{F_i} = \beta^0  x_i + \tau_{a[i]}$$
 

### 2025-12-23 02_cholla_ant_IPM_vital_rates.R

@@ -174,26 +174,26 @@ stan_data_surv <- list(N = nrow(survival_data),                                 
                        plot = as.integer(as.factor(survival_data$Plot)),           ## predictor plots
                        year = as.integer(as.factor(survival_data$Year_t))          ## predictor years
 ) 
-# ## Run the survival model with a bernoulli distribution ---- fixed effects: previous size and ant state; random effects: plot and year
-# surv_model <- stan_model("STAN Models/surv_code.stan")
-# fit_surv<-sampling(surv_model, data = stan_data_surv,chains=3,
-#                    control = list(adapt_delta=0.99,stepsize=0.1),
-#                    iter=10000,cores=3,thin=2,
-#                    pars = c("beta0","beta1","u","w"),   #location coefficients)
-#                    save_warmup=F)
-# # ##Save the RDS file which saves all parameters, draws, and other information
-# saveRDS(fit_surv, "G:/Shared drives/Miller Lab/Sevilleta/Cholla/Model Outputs/fit_surv.rds")
-# saveRDS(fit_surv, "/Users/Labuser/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_surv.rds")
-# saveRDS(fit_surv,"/Users/alicampbell/Dropbox/Ali and Tom -- cactus-ant mutualism project/Model Outputs/fit_surv.rds")
-# surv_model_null <- stan_model("Data Analysis/STAN Models/surv_code_null.stan")
-# fit_surv_null<-sampling(surv_model_null, data = stan_data_surv,chains=3,
-#                    control = list(adapt_delta=0.99,stepsize=0.1),
-#                    iter=10000,cores=3,thin=2,
-#                    pars = c("beta0","beta1","u","w"),   #location coefficients)
-#                    save_warmup=F)
+## Run the survival model with a bernoulli distribution ---- fixed effects: previous size and ant state; random effects: plot and year
+surv_model <- stan_model("STAN Models/surv_code.stan")
+fit_surv<-sampling(surv_model, data = stan_data_surv,chains=3,
+                   control = list(adapt_delta=0.99,stepsize=0.1),
+                   iter=1000,cores=3,thin=2,
+                   pars = c("beta0","beta1","u","w"),   #location coefficients)
+                   save_warmup=F)
[…]

### 2025-12-23 05_cholla_ant_IPM_figures_tables.R

@@ -462,12 +462,11 @@ colMeans(b)
 colMeans(c)
 
 
-
-
-
-
-
 png("Manuscript/Figures/corr_RFX.png", width=1800, height=1200, res=200)
+
+# define breaks and color ramp for correlation legend
+breaks <- seq(-1, 1, length.out = 101)  
+my_colors <- colorRampPalette(c("blue", "white", "red"))(length(breaks)-1)
 
 layout(matrix(c(1,2,3,4,5,6,7,8), nrow=2, byrow=TRUE), widths=c(1,1,1,1))
 par(mar=c(1,3,4,1.5), oma=c(1,5,5,0))
@@ -488,13 +487,18 @@ corrplot(c, method="shade", col.lim=c(-1,1), type="upper", diag=FALSE,
          addCoef.col=NULL, cl.pos="n")
 mtext("c)", side=3, adj=0, line=1, cex=1.8)
 
-par(mar=c(2,2,2,8))  # margins for legend
+# color ramp legend
+par(mar=c(2,2,2,8))  
 x <- seq(0, 1, length = 2)
 y <- breaks
 z <- matrix(breaks[-length(breaks)], nrow = 1)
+
 image(x, y, z, col=my_colors, xaxt="n", yaxt="n", xlab="", ylab="")
 axis(4, at=seq(-1,1,by=0.5), labels=seq(-1,1,by=0.5), cex.axis=2, las=1)
-mtext("   Growth          Survival           Viability                  ", side=3, line=1, outer=TRUE, cex=2)
+mtext("Correlation \nCoefficient", side=3, line=1.2, cex=1.5)
+
+mtext("   Growth          Survival           Viability                  ", 
+      side=3, line=1, outer=TRUE, cex=2)
 
 ############### Timeseries
 par(mar=c(3,3,1,.5))
@@ -534,6 +538,8 @@ legend(x=0, y=.85, legend=c(expression(italic("C. opuntiae")),
        fill=c(cremcol, liomcol, othercol, vaccol), cex=1.4)
 
 dev.off()
+
+
 
 
 png("Manuscript/Figures/year_ant_timeseries.png",width = 480, height = 280)
@@ -1745,6 +1751,69 @@ legend("bottomright", legend = scenario_abv, fill = cols)
 mtext("Number of Partners",side = 1, cex = 2.2, line = 3.5, adj = -50)
 dev.off()
 
+
[…]
